### 2018-05-24 InequalitiesPaper/Inequality3.tex
Adding all research and expository material so far

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,geometry,graphicx,enumerate}

\def\Name{Edward Kim}  % Your name

%for code%
\usepackage{listings}
\lstset{language=python}

\title{Proof of Inequality 2}
\author{\Name}

\textheight=9in
\textwidth=6.5in
\topmargin=-.75in
\oddsidemargin=0.25in
\evensidemargin=0.25in
\DeclareMathSizes{11}{19}{13}{9}
\setlength\parindent{0pt}

\begin{document}
\maketitle

First, let us state the problem and develop some motivation for it: \newline

Let $n \in \mathbb{N}$. Then the following holds provided that $a_i \geq 1$ for $1 \leq i \leq n$.

\begin{equation}
 \sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \frac{n}{1 + (a_1...a_n)^{\frac{1}{n}}}
\end{equation}

\textbf{Proposition: } The inequality holds for all $n \in  \mathbb{N}$. \newline

\textbf{Proof: } We shall use the property that the inequality holds for all $n = 2^k$ for some $k \in \mathbb{N}$.

Let $m$ be a power of 2 such that $m > n$. Observe the following modification to the original sum as follows.

\begin{gather}
  \sum_{i=1}^{n} \frac{1}{1 + a_i} + \sum^{m-n} \frac{1}{1 + (a_1...a_n)^{\frac{1}{n}}} 
\end{gather}

Since $(a_1...a_n)^{\frac{1}{n}} \geq 1$ and there are $m$ terms in the sum, we can apply the inequality:

\begin{gather}
\sum_{i=1}^{n} \frac{1}{1 + a_i} + \frac{m-n}{1 + (a_1...a_n)^{\frac{1}{n}}} \geq \frac{m}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} \\ 
\sum_{i=1}^{n} \frac{1}{1 + a_i} + \frac{m-n}{1 + (a_1...a_n)^{\frac{1}{n}}} \geq \frac{n}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} + \frac{m-n}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} \\
\sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \frac{n}{1 + (a_1...a_n)^{\frac{1}{n}}}
\end{gather}

Thus, the inequality holds for any $n \in \mathbb{N}$.

\end{document}

### 2018-05-24 InequalitiesPaper/Proof1.tex
Adding all research and expository material so far

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,geometry,graphicx,enumerate}

\def\Name{Edward Kim}  % Your name

%for code%
\usepackage{listings}
\lstset{language=python}

\title{Proof of Inequality 1}
\author{\Name}

\textheight=9in
\textwidth=6.5in
\topmargin=-.75in
\oddsidemargin=0.25in
\evensidemargin=0.25in
\DeclareMathSizes{11}{19}{13}{9}

\begin{document}
\maketitle

\textbf{Proof:}
Let $a \geq 1$ and $b \geq 1$. We shall directly prove the following inequality.

\begin{equation}
\frac{1}{1+a} + \frac{1}{1+b} \geq \frac{2}{1 + \sqrt{ab}} 
\end{equation}

We first get the expanded version after multiplying both sides by all the terms. We notice that this preserves the inequality since all terms are positive.

\begin{equation}
(1+b)(1 + \sqrt{ab}) + (1+a)(1 + \sqrt{ab}) \geq 2(1+a)(1+b)
\end{equation}

After we expand each term and simplify, we get the following terms:

\begin{gather}
 2\sqrt{ab} + a + b + b\sqrt{ab} + a\sqrt{ab} \geq 2a + 2b + 2ab \\
 2\sqrt{ab} - a - b + b\sqrt{ab} + a\sqrt{ab} - 2(\sqrt{ab})^2 \geq 0
\end{gather}

Factor out $\sqrt{ab} - 1$:

\begin{gather}
 (\sqrt{ab} - 1)(a+b) + 2(\sqrt{ab} - (\sqrt{ab})^2) \geq 0 \\ 
 (\sqrt{ab} - 1)(a+b) + 2(\sqrt{ab})(1 - \sqrt{ab}) \geq 0 \\
 (\sqrt{ab} - 1)(a+b) - 2(\sqrt{ab})(\sqrt{ab} - 1) \geq 0 
\end{gather}

Now divide both sides by the common factor:

\begin{gather}
 (a + b) - 2\sqrt{ab} \geq 0 \\
 (a + b) \geq 2\sqrt{ab} \\
 \frac{a + b}{2} \geq \sqrt{ab}
\end{gather}

However, this is exactly the Arithmetic-Geometric Mean Inequality which holds when $a \geq 1$ and $b \geq 1$.

Thus, the inequality is proven.

\end{document}

### 2020-04-26 P590/final.tex
Half of final report on Schur-Weyl duality

\documentclass[12pt]{article}%
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{physics}
\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.2cm, right=2.2cm]%
{geometry}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{ytableau}

\newtheorem{thm}{Theorem}
\newtheorem{lemma}{Lemma}
%\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}

\begin{document}

\title{On the Schur-Weyl Duality and its Role in Quantum Information}
\author{Edward Kim}
\date{\today}
\maketitle
\begin{abstract}
  The Schur-Weyl Duality describes the decomposition of representations of $GL(d,\mathbb{C})$ and the symmetric group $S_n$ into subspaces of irreducible tensor representations indiced through partitions determined by Young diagrams. In particular, we are interested in the decomposition of a $n$-tensored $d$-dimensional Hilbert space $\mathcal{H}^{\otimes n}$ and the subsequent operators acting on it as described by the two groups above. These tools allow us to derive slick proofs for Schumacher's compression rate and the sub-additivity properties of von-Neumann entropy.
\end{abstract}

\tableofcontents
\newpage

\section{The Schur-Weyl Duality}

Formal definitions of the Schur-Weyl Duality revolve the commutation property of the actions of $GL(d,\mathbb{C})$ and $S_n$ for some $n,d \geq 1$. First, we define the following basic notions to state the theorem: \newline

Let $\mathcal{H}$ denote a $d$-dimensional complex Hilbert space with orthonormal basis $\{i\}_{1 \leq i \leq d}$ and let $H^{\otimes n}$ denote its corresponding $n$-tensored form.
This allows us to define the following representations:
\begin{gather*}
  R: GL(d,\mathbb{C}) \rightarrow GL(\mathcal{H}^{\otimes n}): \\
  R(g) \cdot \ket{\psi_1} \otimes \ket{\psi_2} \otimes ... \otimes \ket{\psi_n} = g \cdot \ket{\psi_1} \otimes g \cdot \ket{\psi_2} \otimes ... \otimes g \cdot \ket{\psi_n}, \quad \ket{\psi_i} \in \mathcal{H}
\end{gather*}
and
\begin{gather*}
  U: S_n \rightarrow GL(\mathcal{H}^{\otimes n}): \\
  R(\sigma) \cdot \ket{s_1} \otimes \ket{s_2} \otimes ... \otimes \ket{s_n} = \ket{s_{\sigma(1)}} \otimes \ket{s_{\sigma(2)}} \otimes ... \otimes  \ket{s_{\sigma(n)}}, \quad s_i \in \{i\}_{1 \leq i \leq d}
\end{gather*}
The Schur-Weyl Duality is based upon the following commutation property: For a subset $A \subseteq GL(\mathcal{H}^{\otimes n})$, define
$$ Comm(A) = \{s \in End(\mathcal{H}^{\otimes n})\mid s \cdot x = x \cdot s, \; \forall x \in A \} $$
One can check that this is an associative unital $\mathbb{C}$-algebra. The commutant theorem shown by Schur states that:
\begin{thm}
  Let $\mathcal{A} = \text{span }{R(GL(d,\mathbb{C}))}$ and $\mathcal{B} = \text{span }U(S_n)$, then $Comm(\mathcal{A}) = \mathcal{B}$ and $Comm(\mathcal{B}) = \mathcal{A}$
\end{thm}

We omit the proof here for the sake of brevity, but the reader can find a proof in \cite{GW} using the Double Commutant Theorem. A direct corollary of this is the decompostion of $\mathcal{H}^{\otimes n}$ into tensor products of irreducible representations of $GL(d,\mathbb{C}), S_n$. Here, we borrow our notation from \cite{botero}.
\begin{gather}
  \mathcal{H}^{\otimes n} = \bigoplus_{\lambda \vdash_d } U_\lambda \otimes V_\lambda
\end{gather}
where the $U_\lambda, V_\lambda$ refer to irreducible representations of $GL(d, \mathbb{C}), S_n$ respectively and $\lambda \vdash_d n$ refers to the partitions of $n$ having at most $d$ rows. The following block decompositions also hold:

\begin{gather}\label{matdecomp}
    R(g) = \bigoplus_{\lambda \vdash_d n} \pi_{U_{\lambda}}(g) \otimes 1_{V_{\lambda}} \\
    U(\sigma) = \bigoplus_{\lambda \vdash_d n} 1_{U_{\lambda}} \otimes \pi_{V_{\lambda}}(\sigma)
\end{gather}

\noindent where the maps

\begin{gather*}
  \pi_{U_{\lambda}}: GL(d, \mathbb{C}) \rightarrow GL(U_\lambda) \\ \pi_{V_{\lambda}}: S_n \rightarrow GL(V_\lambda)
\end{gather*} refer to the matrices corresponding to the linear maps describing the action on the corresponding irreducible representations.

The core idea behind the utilization strategy of the Schur-Weyl duality involves inferring on the dimensions of the irreducible representations to deduce bounds on the spectra of density operators $\rho \in \mathcal{D}(\mathcal{H})$.

\subsection{Young Tableaux}
We now describe the formal definition of $\lambda \vdash_d n$, namely the partitions of $n$ having at most $d$ rows. A \textit{partition} $\lambda = (\lambda_1,...,\lambda_d)$ is described as a sequence of non-increasing non-negative integers such that:
$$ \lambda_1 \geq \lambda_2 \geq ... \geq \lambda_d, \quad \sum_{i=1}^d \lambda = n $$
These partitions will define the methods in which calculate the dimensions of the irreducible representations present in $U_\lambda \otimes V_\lambda$.
The \textit{Young Tableaux} provides a relatively simple combinatorial framework to facilitate these calculations. Suppose we are given a partition as defined above, then one can construct a corresponding \textit{Young frame} as follows: each $\lambda_i$ represents the number of cells present in the $i^{th}$ row of the Young frame. Furthermore, our rows index from top to bottom. For example, Young diagram shown in Figure \ref{fig:yd} represents a $(3,3,2,1)$ partition.

\begin{figure}[ht]
    \centering
    \ydiagram{3, 3, 2, 1}
    \caption{Young diagram of a (3,3,2,1) partition}
    \label{fig:yd}
\end{figure}

This gives a bijective correspondence between Young frames and partitions in $\lambda \vdash_d n$. Once we have determined the proper Young diagram for $\lambda$, the dimensions of the irreducible representations are calculated through filling these diagrams under specific rules respecting the symmetry imposed by the two groups.

A more detailed exposition on the calculation of these Young Tableaux is given in \cite{FH}.

\subsection{Bounds on Dimension of $U_\lambda$ and $Y_\lambda$}
Based on the Young Tableaux filling rules explained in the previous section, we can explicitly calculate $\dim U_\lambda$ and $\dim V_\lambda$ for a given a partition $\lambda$. For the irreducible $Gl(d, \mathbb{C})$-represenation $U_\lambda$:

$$ \dim U_\lambda = \frac{\prod_{1 \leq i<j \leq d} (\lambda_i - \lambda_j - i + j)}{\prod_{i=1}^{d-1} i!} $$

A simple analysis shows that $(\lambda_j - \lambda_i +i - j) \leq (n+1)$ and there are $d(d-1)/2$ distinct pairs. This immediately gives us the bound:
\begin{equation}\label{ubound}
  \dim U_\lambda \leq (n + 1)^{d(d-1)/2}
\end{equation}

The calculation for the dimension of the irreducible $S_n$-representation $V_\lambda$ requires an auxiliary computation called the \textit{Hook-length} of a box. Formally speaking:
\begin{gather*}
  \dim V_\lambda = \frac{n!}{\prod_{i} h(i)}
\end{gather*}
where the product walks over all boxes in the Young diagram of $\lambda$. The function value $h(i)$ refers to the hook length of box $i$ which is defined to be the number of cells vertically below and horizontally to the right of box $i$. We can bound this value by observing that we can simply disregard the vertical cells below each box in a row. Thus, given all boxes in row $i$, the hook length can be bounded from below:
$$ \lambda_i! \leq \prod_{r \in row_\lambda(i)} h(r) $$ where $row_\lambda(i)$ is the set of all cells comprising row $i$ of the Young diagram representing partition $\lambda$. By applying this definition for each row, we get the following upper bound:

\begin{equation}\label{vbound}
  \dim V_\lambda \leq \frac{n!}{\prod_i \lambda_i!}
\end{equation}

Similarly, we can add the total number of rows to the analysis to get the upper bound:
$$\prod_{r \in row_\lambda(i)} h(r) \leq (\lambda_i + d)!$$ which subsequently gives the lower bound:
$$ \frac{n!}{\prod_i (\lambda +d)!} \leq \dim V_{\lambda} $$
The bounds given for $\dim V_\lambda$ are particularly important for our connection to quantum information. To see this, note that we can reexpress our bounds as follows:
$${n \choose \lambda_1+d,\lambda_2+d,...,\lambda_d+d} \leq \dim V_{\lambda} \leq {n \choose \lambda_1,\lambda_2,...,\lambda_d} $$ where the two bounds are expressed as multinomial coefficients. From classical Shannon entropy, we know that these multinomial coeffients are asymptotic to $2^{nH(\bar{\lambda})}$ where $H$ is the usual Shannon entropy function. Explicitly applying this to our coeffients above shows that:
$$ {n \choose \lambda_1,\lambda_2,...,\lambda_d} \sim 2^{nH(\bar{\lambda})}$$
where $\bar{\lambda} = (\frac{\lambda_1}{n}, \frac{\lambda_2}{n},...,\frac{\lambda_d}{n})$ is the \textit{reduced form} of the partition $\lambda$. Hence, the bounds above directly show that indeed:

\begin{equation}
  \dim V_{\lambda} \sim 2^{nH(\bar{\lambda})}
\end{equation}

\subsection{Schur Polynomials}
One final tool we need involves taking the trace of the $\pi_{U_{\lambda}}(g)$ for some $g \in GL(d, \mathbb{C})$. These are calculated through the \textit{Schur polynomials}:
\begin{equation}\label{schurdef}
    Tr(\pi_{U_\lambda}(g)) = s_{\lambda}(r_1,...,r_d)
\end{equation}
where $spec(g) = (r_1,...,r_d)$. $s_\lambda$ is defined by the following sum over all semi-standard Young Tableau $\Gamma(\lambda)$ over Young frame $\lambda$:
\begin{equation}
  s_{\lambda}(x_1,...,x_d) = \sum_{t \in \Gamma(\lambda)} x_1^{w_1(t)}x_2^{w_2(t)}...x_d^{w_d(t)}
\end{equation}
where $w_i: \Gamma(\lambda) \rightarrow \{1,...,n\}$ denotes the function outputting the weight of integer $i$ assigned in tableau $t$. \newline

Using this definition we can bound the Schur polynomials as follows:
\begin{equation} \label{schur}
  \prod_{i=1}^d r_i^{\lambda_i} \leq s_{\lambda}(r_1,....,r_d) \leq \dim{U_\lambda} \prod_{i=1}^d r_i^{\lambda_i}
\end{equation}

\section{Quantum Compression}

\subsection{Keyl-Werner Theorem}
Perhaps the foundations of our bridge between quantum information and representation theory roots itself in the celebrated Keyl-Werner Theorem. As mentioned in their original paper \cite{KW}, Keyl and Werner were originally motivation by the problem of seeking a set of observables which could be used to output good approximations of the spectra of an arbitrary density operator $\rho \in \mathcal{D}(\mathcal{H})$.  Roughly speaking, the theorem shows that measuring on the subspaces of the form $U_\lambda \otimes V_\lambda$ where the Young frames of $\lambda$ are ``close" to $spec(\rho)$ gives good estimates of $spec(\rho)$. The formal proof of the theorem was refined by Hayashi and Matsumoto and corrected slightly by Christandl and Mitchison \cite{CM}. We follow the latters' formulation.
\begin{thm}
  Given a density operator $\rho$ with spectrum $r  = spec(\rho)$ and $\epsilon_1 > 0$, define projection operator $P_X = \sum_{\lambda \vdash_d n, \bar{\lambda} \in B_{\epsilon_1}(r)} P_{U_{\lambda} \otimes V_{\lambda}}$. Then for $\epsilon_2 > 0$, there exists sufficiently large $k$ such that:
  $$  Tr(P_X \rho^{\otimes k}) > 1 - \epsilon_2$$
\end{thm}

\noindent where we recall that $\bar{\lambda}$ is reduced partition as first defined in the previous section. To show this, we need a lemma.  Assume that $spec(\rho)$ is ordered in non-increasing order henceforth.

\begin{lemma}
  Let $\rho \in \mathcal{D}(\mathcal{H})$ and $P_{\lambda}$ be the projection on $U_{\lambda} \otimes Y_{\lambda}$ described by Young frame $\lambda$. Then:
  $$Tr(P_{\lambda}\rho^{\otimes n}) \leq (n+1)^{d(d-1)/2} \cdot 2^{-n D(\bar{\lambda}||spec(\rho))} $$
  where $D(\bar{\lambda}||spec(\rho))$ refers to the Kullback-Liebler distance between distributions $\bar{\lambda},spec(\rho)$
\end{lemma}

\begin{proof}
 By Property \ref{matdecomp},
 $$ \rho^{\otimes n} = \bigoplus_{\lambda \vdash_d n} \pi_{U_\lambda}(\rho) \otimes 1_{V_\lambda}$$
 We can calculate the probability of projecting on $P_{\lambda}$:
 \begin{gather*}
   Tr(P_\lambda\rho^{\otimes n}) = Tr(\pi_{U_\lambda})Tr(1_{V_\lambda}) = s_{\lambda}(r_1,...,r_d)\dim_{V_\lambda}
 \end{gather*}
  Now the bounds derived for the Schur polynomials at (\ref{schur}) and the dimension bounds mentioned at (\ref{ubound}) and (\ref{vbound}) give us that:
  \begin{gather*}
    Tr(P_\lambda\rho^{\otimes n}) \leq \dim{V_\lambda}\dim{U_\lambda} \prod_{i} r_i^{\lambda_i} \\
    \leq (n+1)^{d(d-1)/2}\frac{n!}{\prod_i \lambda_i!} \prod_{i} r_i^{\lambda_i} \\
    \leq (n+1)^{d(d-1)/2}\cdot 2^{-n D(\bar{\lambda}||spec(\rho))}
  \end{gather*}
\end{proof}

\subsection{Towards Compression}
We now further develop the connection between quantum information and representation theory through an illustrative example concerning the Quantum Compression rate.

\section{Subadditivity Properties of von-Neumann entropy}

\bibliography{biblio}
\bibliographystyle{ieeetr}

\end{document}

### 2020-04-27 P590/final.tex
Update on Final

@@
 
+\setlength{\parskip}{1ex}
+
 \newtheorem{thm}{Theorem}
 \newtheorem{lemma}{Lemma}
+\newtheorem{coro}{Corollary}
 %\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
@@
 \section{The Schur-Weyl Duality}
-
 Formal definitions of the Schur-Weyl Duality revolve the commutation property of the actions of $GL(d,\mathbb{C})$ and $S_n$ for some $n,d \geq 1$.
-First, we define the following basic notions to state the theorem: \newline
+First, we define the following basic notions to state the theorem:
 
@@
 A direct corollary of this is the decompostion of $\mathcal{H}^{\otimes n}$ into tensor products of irreducible representations of $GL(d,\mathbb{C}), S_n$.
-Here, we borrow our notation from \cite{botero}.
+Here, we borrow our notation and formulation from \cite{botero}.
 \begin{gather}
@@
 
-This gives a bijective correspondence between Young frames and partitions in $\lambda \vdash_d n$.
-Once we have determined the proper Young diagram for $\lambda$, the dimensions of the irreducible representations are calculated through filling these diagrams under specific rules respecting the symmetry imposed by the two groups.
+This gives a bijective correspondence between Young frames containing $n$ cells with at most $d$ rows and partitions in $\lambda \vdash_d n$.
+Hence, we can simply refer to the Young frame $\lambda$ rather than the partition $\lambda$.
+ Once we have determined the proper Young diagram for $\lambda$, the dimensions of the irreducible representations are calculated through filling these frames under specific rules respecting the symmetry imposed by the two groups.
+These rules are enforced via the \textit{standard Young tableau (SYT)} for $S_n$-representations and the \textit{semi-standard Young tableau (SSYT)} for $GL(d, \mathbb{C})$-representations.
 
-A more detailed exposition on the calculation of these Young Tableaux is given in \cite{FH}.
+\subsubsection{Standard Young Tableau}
+Given Young frame $\lambda$ indicing an irreducible $S_n$-representation, a filling of $\lambda$ via the standard Young tableau follows the following conditions:
+\begin{enumerate}
+  \item Cells are filled with any integer from $\{1,...,n\}$ where $|\lambda|$ is simply the number of cells in the frame.
+Each integer can only occur once in the tableau i.e can only be of weight one.
+  \item The rows must be strictly increasing from left to right and along columns from top to bottom.
+\end{enumerate}
+
+\noindent Figure \ref{syt} is borrowed from \cite{botero} and gives fillings of a $(3,2)$ frame via this standard Young tableau.
+
+\begin{figure}[ht]
+  \centering
+  \label{syt}
+  \begin{ytableau}
+  1 & 2 & 3 \\
+  4 & 5
+  \end{ytableau} \quad
+  \begin{ytableau}
+  1 & 3 & 4  \\
+  2 & 5
+  \end{ytableau} \quad
+  \begin{ytableau}
+  1 & 3 & 5  \\
+  2 & 4
+  \end{ytableau} \quad
+  \begin{ytableau}
+  1 & 2 & 5  \\
+  3 & 4
+  \end{ytableau} \quad
+  \begin{ytableau}
+  1 & 3 & 4  \\
+  2 & 5
+  \end{ytableau} \quad
+  \begin{ytableau}
+  1 & 2 & 4  \\
+  3 & 5
+  \end{ytableau}
+  \caption{Fillings of a $(3,2)$ partition via SYT}
+\end{figure}
+
+\subsubsection{Semi-Standard Young Tableau}
+Given Young frame $\lambda$ indicing an irreducible $GL(d,\mathbb{C})$-representation, a filling of $\lambda$ via the semi-standard Young tableau follows the following conditions:
+\begin{enumerate}
+  \item Cells are filled with any integer from $\{1.,,,.,d\}$. These integers can be of arbitrary weight i.e occur more than once or not at all.
+  \item Rows are non-decreasing along rows from left to right and strictly increasing along columns from top to bottom.
+\end{enumerate}
+
+\noindent The following uses the same frame as Figure \ref{syt} except we look at with a filling with integers $\{1,2,3\}$
+\begin{figure}[ht]
+  \centering
+  \begin{ytableau}
+    1 & 2 & 3 \\
+    2 & 3
+  \end{ytableau} \quad
+  \begin{ytableau}
+    1 & 1 & 3 \\
+    2 & 3
+  \end{ytableau} \quad
+  \begin{ytableau}
+    1 & 1 & 2 \\
+    2 & 3
+  \end{ytableau} \quad
+  \begin{ytableau}
+    2 & 2 & 2 \\
+    3 & 3
+  \end{ytableau} \quad
+  \begin{ytableau}
+    1 & 1 & 1 \\
+    3 & 3
+  \end{ytableau} \quad
+  \begin{ytableau}
+    1 & 1 & 1 \\
+    2 & 2
+  \end{ytableau}
+  \caption{Some fillings of $(3,2)$ parition with $\{1,2,3\}$ under SSYT}
+\end{figure}
+
+\noindent A more detailed exposition on the calculation of these Young Tableaux is given in \cite{FH}.
 
 \subsection{Bounds on Dimension of $U_\lambda$ and $Y_\lambda$}
-Based on the Young Tableaux filling rules explained in the previous section, we can explicitly calculate $\dim U_\lambda$ and $\dim V_\lambda$ for a given a partition $\lambda$.
+Based on the Young Tableaux filling rules explained in the previous section, we can explicitly calculate $\dim U_\lambda$ and $\dim V_\lambda$ for a given a Young frame $\lambda$.
 For the irreducible $Gl(d, \mathbb{C})$-represenation $U_\lambda$:
@@
 \end{gather*}
-where the product walks over all boxes in the Young diagram of $\lambda$.
+where the product walks over all boxes in the Young frame $\lambda$.
 The function value $h(i)$ refers to the hook length of box $i$ which is defined to be the number of cells vertically below and horizontally to the right of box $i$.
@@
 $$ \lambda_i!
-\leq \prod_{r \in row_\lambda(i)} h(r) $$ where $row_\lambda(i)$ is the set of all cells comprising row $i$ of the Young diagram representing partition $\lambda$.
+\leq \prod_{r \in row_\lambda(i)} h(r) $$ where $row_\lambda(i)$ is the set of all cells comprising row $i$ of the Young frame $\lambda$.
 By applying this definition for each row, we get the following upper bound:
@@
 
-Similarly, we can add the total number of rows to the analysis to get the upper bound:
+\noindent Similarly, we can add the total number of rows to the analysis to get the upper bound:
 $$\prod_{r \in row_\lambda(i)} h(r) \leq (\lambda_i + d)!$$ which subsequently gives the lower bound:
@@
 $$ {n \choose \lambda_1,\lambda_2,...,\lambda_d} \sim 2^{nH(\bar{\lambda})}$$
-where $\bar{\lambda} = (\frac{\lambda_1}{n}, \frac{\lambda_2}{n},...,\frac{\lambda_d}{n})$ is the \textit{reduced form} of the partition $\lambda$. Hence, the bounds above directly show that indeed:
+where $\bar{\lambda} = (\frac{\lambda_1}{n}, \frac{\lambda_2}{n},...,\frac{\lambda_d}{n})$ is the \textit{reduced form} of the Young frame $\lambda$. Hence, the bounds above directly show that indeed:
 
-\begin{equation}
+\begin{equation} \label{asymp}
   \dim V_{\lambda} \sim 2^{nH(\bar{\lambda})}
@@
 \subsection{Schur Polynomials}
-One final tool we need involves taking the trace of the $\pi_{U_{\lambda}}(g)$ for some $g \in GL(d, \mathbb{C})$. These are calculated through the \textit{Schur polynomials}:
+The final tool we need involves taking the trace of the $\pi_{U_{\lambda}}(g)$ for some $g \in GL(d, \mathbb{C})$. These are calculated through the \textit{Schur polynomials}:
+
 \begin{equation}\label{schurdef}
@@
 \end{equation}
-where $spec(g) = (r_1,...,r_d)$. $s_\lambda$ is defined by the following sum over all semi-standard Young Tableau $\Gamma(\lambda)$ over Young frame $\lambda$:
+
+\noindent where $spec(g) = (r_1,...,r_d)$. The polynomials $s_\lambda$ are defined by the following sum over all semi-standard Young Tableau $\Gamma(\lambda)$ over Young frame $\lambda$:
+
 \begin{equation}
@@
 \end{equation}
-where $w_i: \Gamma(\lambda) \rightarrow \{1,...,n\}$ denotes the function outputting the weight of integer $i$ assigned in tableau $t$. \newline
 
-Using this definition we can bound the Schur polynomials as follows:
+\noindent where $w_i: \Gamma(\lambda) \rightarrow \{1,...,n\}$ denotes the function outputting the weight of integer $i$ assigned in tableau $t$.
+In other words, $w_i(t)$ counts the number of occurances of integer $i$ in tableau $t$.
+Using this definition we can bound the Schur polynomials.
+Without loss of generality, assume that we order $spec(\rho) = (r_1,...,r_d)$ in non-increasing order $r_1 \geq r_2 \geq ...
+\geq r_d$.
+
 \begin{equation} \label{schur}
@@
 \end{equation}
+
+The left-hand side simply follows from the definition of the Schur polynomials and the filling rule imposed by the semi-standard Young tableau.
+As $s_\lambda$ runs over all such tableau $t$, it must be at least as large as one such filling obeying the the semi-standard Young tableau rule.
+On the other hand, recalling that the number of semi-standard Young tableau is equal to $\dim U_{\lambda}$ and noticing that the product gives the maximum value over any such filling of $\lambda$ under the semi-standard rule gives us the right-hand side.
 
@@
 As mentioned in their original paper \cite{KW}, Keyl and Werner were originally motivation by the problem of seeking a set of observables which could be used to output good approximations of the spectra of an arbitrary density operator $\rho \in \mathcal{D}(\mathcal{H})$.
-Roughly speaking, the theorem shows that measuring on the subspaces of the form $U_\lambda \otimes V_\lambda$ where the Young frames of $\lambda$ are ``close" to $spec(\rho)$ gives good estimates of $spec(\rho)$.
-The formal proof of the theorem was refined by Hayashi and Matsumoto and corrected slightly by Christandl and Mitchison \cite{CM}.
+Roughly speaking, the theorem shows that measuring on the subspaces of the form $U_\lambda \otimes V_\lambda$ where the Young frames of $\lambda$ are ``close" to $spec(\rho)$ give good estimates of $spec(\rho)$.
+The formal proof of the theorem was refined by Hayashi and Matsumoto \cite{HM} and corrected slightly by Christandl and Mitchison \cite{CM}.
 We follow the latters' formulation.
-\begin{thm}
-  Given a density operator $\rho$ with spectrum $r  = spec(\rho)$ and $\epsilon_1 > 0$, define projection operator $P_X = \sum_{\lambda \vdash_d n, \bar{\lambda} \in B_{\epsilon_1}(r)} P_{U_{\lambda} \otimes V_{\lambda}}$.
+
+\begin{thm} \label{keyl}
+  Given a density operator $\rho$ with spectrum $r  = spec(\rho)$ and $\epsilon_1 > 0$, define projection operator $P_X = \sum_{\lambda \vdash_d n, \bar{\lambda} \in B_{\epsilon_1}(r)} P_{\lambda}$ where $P_{\lambda}$ is the projection on $U_{\lambda} \otimes V_{\lambda}$.
 Then for $\epsilon_2 > 0$, there exists sufficiently large $k$ such that:
@@
   Let $\rho \in \mathcal{D}(\mathcal{H})$ and $P_{\lambda}$ be the projection on $U_{\lambda} \otimes Y_{\lambda}$ described by Young frame $\lambda$. Then:
-  $$Tr(P_{\lambda}\rho^{\otimes n}) \leq (n+1)^{d(d-1)/2} \cdot 2^{-n D(\bar{\lambda}||spec(\rho))} $$
+  $$Tr(P_{\lambda}\rho^{\otimes n}) \leq (n+1)^{d(d-1)/2} \cdot exp \left\{-n D(\bar{\lambda}||spec(\rho)) \right\} $$
   where $D(\bar{\lambda}||spec(\rho))$ refers to the Kullback-Liebler distance between distributions $\bar{\lambda},spec(\rho)$
@@
     \leq (n+1)^{d(d-1)/2}\frac{n!}{\prod_i \lambda_i!} \prod_{i} r_i^{\lambda_i} \\
-    \leq (n+1)^{d(d-1)/2}\cdot 2^{-n D(\bar{\lambda}||spec(\rho))}
+    \leq (n+1)^{d(d-1)/2}\cdot exp\left\{{-n D(\bar{\lambda}||spec(\rho))}\right\}
   \end{gather*}
@@
 
+\noindent From the above lemma, we arrive at the following corollary.
+
+\begin{coro}
+  Let $P_X = \sum_{\lambda \vdash_d n: \bar{\lambda} \in \mathcal{S}} P_\lambda$. Then:
+  $$ Tr(P_X\rho^{\otimes n}) \leq (n+1)^{d(d+1)/2} \cdot exp\left\{-n \min_{\lambda \vdash_d n, \bar{\lambda} \in \mathcal{S}}{D(\bar{\lambda}||spec(\rho))} \right\}$$
+where $\mathcal{S} \subset [0,1]^d$.
+\end{coro}
+
+\begin{proof}
+Observe the following string of inequalities:
+  \begin{gather*}
+      Tr(P_X\rho^{\otimes n}) = \sum_{\lambda \vdash_d n: \bar{\lambda} \in \mathcal{S}} Tr(P_\lambda\rho^{\otimes n}) \leq \sum_{\lambda \vdash_d n: \bar{\lambda} \in \mathcal{S}} (n+1)^{d(d-1)/2} \cdot exp \left\{ -n D(\bar{\lambda}||spec(\rho)) \right\} \\
+      \leq
+            \sum_{\lambda \vdash_d n} (n+1)^{d(d-1)/2} \cdot exp \left\{ -n \min_{\lambda \vdash_d n, \bar{\lambda} \in \mathcal{S}} D(\bar{\lambda}||spec(\rho)) \right\} \\
+      \leq (n+1)^d  (n+1)^{d(d-1)/2} \cdot exp \left\{ -n \min_{\lambda \vdash_d n, \bar{\lambda} \in \mathcal{S}} D(\bar{\lambda}||spec(\rho)) \right\} \\
+      = (n+1)^{d(d+1)/2} exp\left\{-n \min_{\lambda \vdash_d n, \bar{\lambda} \in \mathcal{S}}{D(\bar{\lambda}||spec(\rho))} \right\}
+  \end{gather*}
+\end{proof}
+
+\noindent We see that by taking our subset $\mathcal{S}$ to be the complement of the ball $\mathcal{B}_{\epsilon_1}(spec(\rho))$, projecting onto the subspaces indiced by Young frames which lie outside of of this ball tend to zero by the corollary above.
+This gives us Theorem \ref{keyl} as required.
+
 \subsection{Towards Compression}
 We now further develop the connection between quantum information and representation theory through an illustrative example concerning the Quantum Compression rate.
+Recall the setup of the Quantum Compresson problem where we have a quantum source describe by ensemble $\{p_X(x), \ket{\psi_i}\bra{\psi_i}\}$.
+This gives us the density operator $\rho = \sum_x p_X(x)\ket{\psi_x}\bra{\psi_x}$.
+The resulting state from $n$-outputs from the quantum source can thus be described as $\rho^{\otimes n}$.
+We wish to determine a compression rate $R$ such that we can project onto a subspace of $2^{nR}$.
+We wish for $R < \log{d}$; that is, ideally the subspace, say $W$, we wish to project onto is of smaller dimension than $d$.
+Furthermore, this subspace must capture enough of the quantum information encoded into $\rho$ in the sense that:
+\begin{equation*}
+  \limsup_{n \rightarrow \infty} || \rho^{\otimes n} - \Pi_{W}\rho_{\otimes n}\Pi_W ||_1 = 0
+\end{equation*}
+
+We can view this setup as the quantum analogue of the classical Shannon source compression result.
+This connection turns out to be deeper as Schumacher's compression theorem extends the concept of typical sequences to \textit{typical subspaces} \cite{sch}.
+In fact, the Keyl-Werner result tells that these subspaces can be taken to be the subspaces given by \textit{typical Young frames}.
+We formalize this below.
+Assume once again that $spec(\rho)$ is ordered in non-increasing order.
+
+Define the projection operator $$ \Pi(T_{\epsilon}) = \bigoplus_{\lambda: \; |\bar{\lambda} - spec(\rho)| < \epsilon} P_{\lambda} $$
+
+where $P_{\lambda}$ is the projection operator for the subspace $U_{\lambda} \otimes V_{\lambda}$. By the Keyl-Werner Theorem (Theorem \ref{keyl}), we know that
+
+$$ \lim_{n \rightarrow \infty} ||\rho^{\otimes n} - \Pi(T_\epsilon) \rho^{\otimes n} \Pi(T_\epsilon) ||_1 = 0$$
+
+Furthermore, we use (\ref{ubound}), (\ref{vbound}) to calculate the dimension of the support of $\Pi(T_\epsilon)$:
+
+\begin{gather} \label{epsilbound}
+  Tr(\Pi(T_\epsilon)) \leq \sum_{\lambda: \; |\bar{\lambda} - spec(\rho)| < \epsilon} \dim_{U_\lambda} \dim_{Y_\lambda} \leq |\mathcal{I}_{\epsilon,\rho}| (n+1)^{d(d-1)/2} \frac{n!}{\prod_{i=1}^d
+   \lambda_i!}
+\end{gather}
+where $\mathcal{I}_{\epsilon,\rho}$ is the set of all Young frames which are $\epsilon$-close to $spec(\rho)$.
+Here we can take $(\lambda_1,...,\lambda_d)$ to be an $\epsilon$-typical Young frame such that $\prod_{i=1}^d \lambda_i!$ is minimized.
+Now this set is bounded by the total number of Young frames of $n$ cells having at most $d$ rows:
+$$ |\{\lambda: \lambda \vdash_d n \}| \leq (n+1)^d$$
+
+\noindent Hence by (\ref{asymp}), we see that the polynomial factors vanish asympotically to give way to exponential factor:
+$$ Tr(\Pi(T_\epsilon)) \sim 2^{nH({\bar{\lambda}})}  $$
+\noindent which yields that:
+$$ Tr(\Pi(T_\epsilon)) \sim 2^{n(S(\rho) + \epsilon)}  $$
+
+The derivation aligns with Schumachers result as we see that the dimension of our Young frame subspace is asymptotic to the von-Neumann entropy $S(\rho)$.
+As $n \rightarrow \infty$, we observe that we cannot compress better than the rate given by $S(\rho)$.
+The reader can find a more rigorous approach in \cite{harrow}.
 
@@
 
+\subsection{Clebsch-Gordon Decomposition}
+An elegant proof of the subadditivity property of von-Neumann entropy arises from the combination of the \textit{Clebsch-Gordon decomposition} and the Keyl-Werner Theorem.
+We simply state the decomposition whose proof can be found in \cite{FH}.
+
+\begin{thm}
+  Let $V_{\gamma}, V_{\omega}$ be two irreducible $S_n$-representations indiced by partitions $\gamma,\omega \vdash_d n$. Then there exists an $S_n$-isomorphism:
+  $$ V_{\gamma} \otimes V_{\omega} \cong \bigoplus_{\lambda \vdash n} g_{\lambda\gamma\omega} V_{\lambda}$$ where $g_{\lambda\gamma\omega}$ are the Kronecker coefficients indicating the multiplicity of $V_\lambda$.
+\end{thm}
+\noindent Note that there are no restrictions on the number of rows for frame $\lambda$.
+The Kronecker coefficients have another interpretation through the Schur-Weyl duality applied to the analysis of tripartite entanglement \cite{CSW}.
+
+\noindent Using this decompostion along with the Keyl-Werner Theorem, Christandl and Mitchison proved the following \cite{CM}:
+
+\begin{thm} \label{converge}
+  Let $\rho_{AB} \in \mathcal{D}(\mathcal{H}_A \otimes \mathcal{H}_B)$ for $d$-dimensional Hilbert spaces $\mathcal{H}_A, \mathcal{H}_B$.
+Then there exists a sequence of partitions $(\lambda_i, \gamma_i, \omega_i)$ with $|\lambda_i| = |\gamma_i| = |\omega_i|$ such that $g_{\lambda_i\gamma_i\omega_i} \neq 0$, and
+  \begin{gather*}
+    \lim_{i \rightarrow \infty} \bar{\lambda_i} = spec(\rho_{AB}) \\
+    \lim_{i \rightarrow \infty} \bar{\gamma_i} = spec(\rho_A) \\
+    \lim_{i \rightarrow \infty} \bar{\omega_i} =  spec(\rho_B)
+  \end{gather*}
+\end{thm}
+\subsection{Subadditivity and SSA}
+
+\noindent Using the above theorem, we can give a simple proof of subadditivity:
+\begin{thm}
+  von-Neumann entropy is subadditive: $S(\rho_{AB}) \leq S(\rho_A) + S(\rho_B)$
+\end{thm}
+\begin{proof}
+  Theorem \ref{converge} gives us a sequence of partitions $(\lambda_i,\gamma_i, \omega_i)$ converging to the spectra of $\rho_{AB}, \rho_A, \rho_B$ respectively.
+As $g_{\lambda_i\gamma_i\omega_i} \neq 0$, we have that:
+  $$ \dim{V_{\lambda_i}} \leq \dim{V_{\gamma_i}}\dim{V_{\omega_i}} $$ as $i \rightarrow \infty$.
+
+  \noindent By (\ref{vbound}), (\ref{asymp}) given above, we have that
+  $$2^{nS(\rho_{AB})} \leq 2^{nS(\rho_A)S(\rho_B)}$$
+  where $n = |\lambda_i| = |\gamma_i| = |\omega_i|$. The result follows after taking the logarithm of both sides.
+\end{proof}
+
+In fact, a similar line of thought using recoupling coeffcients is used to prove the strong subadditivity property of von-Neumann entropy as well \cite{CSW}.
+These recoupling coefficients arise from a direct generalization of the Clebsch-Gordon decomposition introduced above, but the starting ingredients are more-or-less the same.
+
+\newpage
 \bibliography{biblio}

### 2020-04-29 P590/final.tex
Final version of p590

@@
 \setlength{\parskip}{1ex}
+\setlength{\intextsep}{5ex}
 
@@
 \begin{abstract}
-  The Schur-Weyl Duality describes the decomposition of representations of $GL(d,\mathbb{C})$ and the symmetric group $S_n$ into subspaces of irreducible tensor representations indiced through partitions determined by Young diagrams.
-In particular, we are interested in the decomposition of a $n$-tensored $d$-dimensional Hilbert space $\mathcal{H}^{\otimes n}$ and the subsequent operators acting on it as described by the two groups above.
-These tools allow us to derive slick proofs for Schumacher's compression rate and the sub-additivity properties of von-Neumann entropy.
+  The Schur-Weyl Duality describes the decomposition of representations of $GL(d,\mathbb{C})$ and the symmetric group $S_n$ into a direct sum of irreducible tensor representations indiced by partitions described by Young frames.
+In particular, we are interested in the decomposition of the $n$-tensored $d$-dimensional Hilbert space $\mathcal{H}^{\otimes n}$ and the subsequent operators acting on it as described by the two groups above.
+Calculation of the dimensions of these irreducible representations reduces to a combinatorial problem, allowing us to derive useful asympotic bounds.
+These bounds provide elegant proofs for Schumacher's compression rate and the subadditivity properties of von-Neumann entropy.
 \end{abstract}
@@
 \section{The Schur-Weyl Duality}
-Formal definitions of the Schur-Weyl Duality revolve the commutation property of the actions of $GL(d,\mathbb{C})$ and $S_n$ for some $n,d \geq 1$.
+Formal definitions of the Schur-Weyl Duality revolve around the commutation property of the actions of $GL(d,\mathbb{C})$ and $S_n$ for $n,d \geq 1$.
 First, we define the following basic notions to state the theorem:
@@
 
-We omit the proof here for the sake of brevity, but the reader can find a proof in \cite{GW} using the Double Commutant Theorem.
-A direct corollary of this is the decompostion of $\mathcal{H}^{\otimes n}$ into tensor products of irreducible representations of $GL(d,\mathbb{C}), S_n$.
+\noindent We omit the proof here for the sake of brevity, but the reader can find a proof using the Double Commutant Theorem \cite{GW}.
+A direct corollary is the decompostion of $\mathcal{H}^{\otimes n}$ into tensor products of irreducible representations of $GL(d,\mathbb{C}), S_n$.
 Here, we borrow our notation and formulation from \cite{botero}.
@@
 \noindent where the maps
-
 \begin{gather*}
   \pi_{U_{\lambda}}: GL(d, \mathbb{C}) \rightarrow GL(U_\lambda) \\ \pi_{V_{\lambda}}: S_n \rightarrow GL(V_\lambda)
-\end{gather*} refer to the matrices corresponding to the linear maps describing the action on the corresponding irreducible representations.
+\end{gather*}
 
-The core idea behind the utilization strategy of the Schur-Weyl duality involves inferring on the dimensions of the irreducible representations to deduce bounds on the spectra of density operators $\rho \in \mathcal{D}(\mathcal{H})$.
+\noindent refer to the matrices corresponding to the linear maps describing the action on the corresponding irreducible representations.
 
 \subsection{Young Tableaux}
-We now describe the formal definition of $\lambda \vdash_d n$, namely the partitions of $n$ having at most $d$ rows.
+We now describe the formal definition of the partitions $\lambda$ such that $\lambda \vdash_d n$, namely the partitions of $n$ having at most $d$ rows.
 A \textit{partition} $\lambda = (\lambda_1,...,\lambda_d)$ is described as a sequence of non-increasing non-negative integers such that:
-$$ \lambda_1 \geq \lambda_2 \geq ... \geq \lambda_d, \quad \sum_{i=1}^d \lambda = n $$
-These partitions will define the methods in which calculate the dimensions of the irreducible representations present in $U_\lambda \otimes V_\lambda$.
-The \textit{Young Tableaux} provides a relatively simple combinatorial framework to facilitate these calculations.
+$$ \lambda_1 \geq \lambda_2 \geq ... \geq \lambda_d, \quad \sum_{i=1}^d \lambda_i = n $$
+These partitions will provide the visual groundwork for calculating the dimensions of the irreducible representations present in $U_\lambda \otimes V_\lambda$.
+The \textit{Young Tableaux} provide a relatively simple combinatorial framework to facilitate these calculations.
 Suppose we are given a partition as defined above, then one can construct a corresponding \textit{Young frame} as follows: each $\lambda_i$ represents the number of cells present in the $i^{th}$ row of the Young frame.
-Furthermore, our rows index from top to bottom.
-For example, Young diagram shown in Figure \ref{fig:yd} represents a $(3,3,2,1)$ partition.
+Furthermore, our rows index from top to bottom, and our cells index from left to right.
+For example, the Young diagram shown in Figure \ref{fig:yd} represents a $(3,3,2,1)$ partition.
 
@@
 
-This gives a bijective correspondence between Young frames containing $n$ cells with at most $d$ rows and partitions in $\lambda \vdash_d n$.
+\noindent This gives a bijective correspondence between Young frames containing $n$ cells with at most $d$ rows and partitions $\lambda$ such that $\lambda \vdash_d n$.
 Hence, we can simply refer to the Young frame $\lambda$ rather than the partition $\lambda$.
@@
 \subsubsection{Standard Young Tableau}
-Given Young frame $\lambda$ indicing an irreducible $S_n$-representation, a filling of $\lambda$ via the standard Young tableau follows the following conditions:
+Given Young frame $\lambda$ indicing an irreducible $S_n$-representation, a filling of $\lambda$ via the \textit{standard Young tableau (SYT)} obeys the following conditions:
 \begin{enumerate}
-  \item Cells are filled with any integer from $\{1,...,n\}$ where $|\lambda|$ is simply the number of cells in the frame.
-Each integer can only occur once in the tableau i.e can only be of weight one.
-  \item The rows must be strictly increasing from left to right and along columns from top to bottom.
+  \item Cells are filled with any integer from $\{1,...,n\}$ where $n$ will be the total number of cells in the frame. Each integer can only occur once in the tableau i.e can only be of weight one.
+  \item The rows must be \textit{strictly increasing} from left to right and along columns from top to bottom.
 \end{enumerate}
 
-\noindent Figure \ref{syt} is borrowed from \cite{botero} and gives fillings of a $(3,2)$ frame via this standard Young tableau.
+\noindent Figure \ref{fig:syt} is borrowed from \cite{botero} and gives fillings of a $(3,2)$ frame via this standard Young tableau.
 
@@
   \centering
-  \label{syt}
   \begin{ytableau}
@@
   \caption{Fillings of a $(3,2)$ partition via SYT}
+  \label{fig:syt}
 \end{figure}
-
+\newpage
 \subsubsection{Semi-Standard Young Tableau}
-Given Young frame $\lambda$ indicing an irreducible $GL(d,\mathbb{C})$-representation, a filling of $\lambda$ via the semi-standard Young tableau follows the following conditions:
+Given Young frame $\lambda$ indicing an irreducible $GL(d,\mathbb{C})$-representation, a filling of $\lambda$ via the \textit{semi-standard Young tableau (SSYT)} follows the following conditions:
 \begin{enumerate}
-  \item Cells are filled with any integer from $\{1.,,,.,d\}$. These integers can be of arbitrary weight i.e occur more than once or not at all.
-  \item Rows are non-decreasing along rows from left to right and strictly increasing along columns from top to bottom.
+  \item Cells are filled with any integer from $\{1,,...,d\}$. These integers can be of arbitrary weight i.e occur more than once or not at all.
+  \item Rows are \textit{non-decreasing} along rows from left to right and \textit{strictly increasing} along columns from top to bottom.
 \end{enumerate}
 
-\noindent The following uses the same frame as Figure \ref{syt} except we look at with a filling with integers $\{1,2,3\}$
+\noindent Figure \ref{fig:ssyt} uses the same frame as Figure \ref{fig:syt} except we adhere to the semi-standard Young tableau with integers $\{1,2,3\}$
 \begin{figure}[ht]
@@
   \caption{Some fillings of $(3,2)$ parition with $\{1,2,3\}$ under SSYT}
+  \label{fig:ssyt}
 \end{figure}
 
-\noindent A more detailed exposition on the calculation of these Young Tableaux is given in \cite{FH}.
+\noindent A more detailed exposition on the calculation of these Young Tableaux is given in \cite{GW}, \cite{FH}.
 
 \subsection{Bounds on Dimension of $U_\lambda$ and $Y_\lambda$}
-Based on the Young Tableaux filling rules explained in the previous section, we can explicitly calculate $\dim U_\lambda$ and $\dim V_\lambda$ for a given a Young frame $\lambda$.
-For the irreducible $Gl(d, \mathbb{C})$-represenation $U_\lambda$:
+Based on the Young Tableau filling rules explained in the previous section, we can explicitly calculate $\dim U_\lambda$ and $\dim V_\lambda$ for a given Young frame $\lambda$ by counting the number of valid fillings possible.
+For the irreducible $Gl(d, \mathbb{C})$-represenation $U_\lambda$, we calculate the number of SSYT possible for Young frame $\lambda$:
 
@@
 
-A simple analysis shows that $(\lambda_j - \lambda_i +i - j) \leq (n+1)$ and there are $d(d-1)/2$ distinct pairs. This immediately gives us the bound:
+\noindent A simple analysis shows that $(\lambda_j - \lambda_i +i - j) \leq (n+1)$.
+Furthermore, there are $d(d-1)/2$ distinct pairs considered in the numerator's product.
+This immediately gives us the bound:
 \begin{equation}\label{ubound}
@@
 
-The calculation for the dimension of the irreducible $S_n$-representation $V_\lambda$ requires an auxiliary computation called the \textit{Hook-length} of a box. Formally speaking:
+\noindent The calculation for the dimension of the irreducible $S_n$-representation $V_\lambda$ requires an auxiliary computation called the \textit{Hook-length} of a box. Formally speaking:
 \begin{gather*}
@@
 where the product walks over all boxes in the Young frame $\lambda$.
-The function value $h(i)$ refers to the hook length of box $i$ which is defined to be the number of cells vertically below and horizontally to the right of box $i$.
+The function value $h(i)$ refers to the hook length of box $i$ which is defined to be the number of cells \textit{below} and to the \textit{right} of box $i$.
 We can bound this value by observing that we can simply disregard the vertical cells below each box in a row.
@@
 $$ {n \choose \lambda_1,\lambda_2,...,\lambda_d} \sim 2^{nH(\bar{\lambda})}$$
-where $\bar{\lambda} = (\frac{\lambda_1}{n}, \frac{\lambda_2}{n},...,\frac{\lambda_d}{n})$ is the \textit{reduced form} of the Young frame $\lambda$. Hence, the bounds above directly show that indeed:
+where $\bar{\lambda} = (\frac{\lambda_1}{n}, \frac{\lambda_2}{n},...,\frac{\lambda_d}{n})$ is the \textit{reduced form} of the Young frame $\lambda$. Hence, the bounds above directly yield that:
 
@@
 \subsection{Schur Polynomials}
-The final tool we need involves taking the trace of the $\pi_{U_{\lambda}}(g)$ for some $g \in GL(d, \mathbb{C})$. These are calculated through the \textit{Schur polynomials}:
+The final tool we need involves taking the trace of the $\pi_{U_{\lambda}}(g)$ for some $g \in GL(d, \mathbb{C})$. These are determined through the \textit{Schur polynomials}:
 
@@
 \subsection{Keyl-Werner Theorem}
-Perhaps the foundations of our bridge between quantum information and representation theory roots itself in the celebrated Keyl-Werner Theorem.
-As mentioned in their original paper \cite{KW}, Keyl and Werner were originally motivation by the problem of seeking a set of observables which could be used to output good approximations of the spectra of an arbitrary density operator $\rho \in \mathcal{D}(\mathcal{H})$.
+Perhaps the foundation of our bridge between quantum information and representation theory roots itself in the celebrated Keyl-Werner Theorem.
+As mentioned in their original paper \cite{KW}, Keyl and Werner were originally motivation by the problem of seeking a set of observables which could be used to output good approximations of the spectra of an arbitrary density operator $\rho \in \mathcal{D}(\mathcal{H})$ for $d$-dimensional Hilbert space $\mathcal{H}$.
 Roughly speaking, the theorem shows that measuring on the subspaces of the form $U_\lambda \otimes V_\lambda$ where the Young frames of $\lambda$ are ``close" to $spec(\rho)$ give good estimates of $spec(\rho)$.
@@
       \leq (n+1)^d  (n+1)^{d(d-1)/2} \cdot exp \left\{ -n \min_{\lambda \vdash_d n, \bar{\lambda} \in \mathcal{S}} D(\bar{\lambda}||spec(\rho)) \right\} \\
-      = (n+1)^{d(d+1)/2} exp\left\{-n \min_{\lambda \vdash_d n, \bar{\lambda} \in \mathcal{S}}{D(\bar{\lambda}||spec(\rho))} \right\}
+      = (n+1)^{d(d+1)/2} \cdot exp\left\{-n \min_{\lambda \vdash_d n, \bar{\lambda} \in \mathcal{S}}{D(\bar{\lambda}||spec(\rho))} \right\}
   \end{gather*}
 \end{proof}
-
-\noindent We see that by taking our subset $\mathcal{S}$ to be the complement of the ball $\mathcal{B}_{\epsilon_1}(spec(\rho))$, projecting onto the subspaces indiced by Young frames which lie outside of of this ball tend to zero by the corollary above.
+\noindent The last inequality comes from the bound
+\begin{equation}\label{framebound}
+|\{\lambda: \lambda \vdash_d n \}| \leq (n+1)^d
+\end{equation}
+which just follows from the binomial expansion.
+\noindent We see that by taking our subset $\mathcal{S}$ to be the complement of the ball $\mathcal{B}_{\epsilon_1}(spec(\rho))$, the probability of projecting onto the subspaces indiced by Young frames which lie outside of this ball tend to zero by the corollary above.
 This gives us Theorem \ref{keyl} as required.
@@
 We wish to determine a compression rate $R$ such that we can project onto a subspace of $2^{nR}$.
-We wish for $R < \log{d}$; that is, ideally the subspace, say $W$, we wish to project onto is of smaller dimension than $d$.
-Furthermore, this subspace must capture enough of the quantum information encoded into $\rho$ in the sense that:
+Furthermore, we wish for $R < \log{d}$; that is, ideally the subspace $W$ we wish to project onto is of smaller dimension than $d$.
+Finally, this subspace must capture enough of the quantum information encoded into $\rho$ in the sense that:
 \begin{equation*}
@@
 
-We can view this setup as the quantum analogue of the classical Shannon source compression result.
-This connection turns out to be deeper as Schumacher's compression theorem extends the concept of typical sequences to \textit{typical subspaces} \cite{sch}.
+We can view this setup as the quantum analogue of the classical Shannon source compression.
+This connection turns out to be deeper as Schumacher's compression theorem extends the concept of typical sequences in classical Shannon theory to \textit{typical subspaces} \cite{sch}.
 In fact, the Keyl-Werner result tells that these subspaces can be taken to be the subspaces given by \textit{typical Young frames}.
@@
 
-Define the projection operator $$ \Pi(T_{\epsilon}) = \bigoplus_{\lambda: \; |\bar{\lambda} - spec(\rho)| < \epsilon} P_{\lambda} $$
+\noindent Define the projection operator $$ \Pi(T_{\epsilon}) = \bigoplus_{\lambda: \; |\bar{\lambda} - spec(\rho)| < \epsilon} P_{\lambda} $$
 
-where $P_{\lambda}$ is the projection operator for the subspace $U_{\lambda} \otimes V_{\lambda}$. By the Keyl-Werner Theorem (Theorem \ref{keyl}), we know that
+\noindent where $P_{\lambda}$ is the projection operator for the subspace $U_{\lambda} \otimes V_{\lambda}$. By the Keyl-Werner Theorem (Theorem \ref{keyl}), we know that
 
@@
 
-Furthermore, we use (\ref{ubound}), (\ref{vbound}) to calculate the dimension of the support of $\Pi(T_\epsilon)$:
+\noindent Furthermore, we use (\ref{ubound}), (\ref{vbound}) to calculate the dimension of the support of $\Pi(T_\epsilon)$:
 
@@
 Here we can take $(\lambda_1,...,\lambda_d)$ to be an $\epsilon$-typical Young frame such that $\prod_{i=1}^d \lambda_i!$ is minimized.
-Now this set is bounded by the total number of Young frames of $n$ cells having at most $d$ rows:
-$$ |\{\lambda: \lambda \vdash_d n \}| \leq (n+1)^d$$
+Now this set is bounded by the total number of Young frames of $n$ cells having at most $d$ rows given by (\ref{framebound}).
 
-\noindent Hence by (\ref{asymp}), we see that the polynomial factors vanish asympotically to give way to exponential factor:
+\noindent Hence by (\ref{asymp}), we see that the polynomial factors vanish asympotically to give way to the exponential factor:
 $$ Tr(\Pi(T_\epsilon)) \sim 2^{nH({\bar{\lambda}})}  $$
@@
 
-The derivation aligns with Schumachers result as we see that the dimension of our Young frame subspace is asymptotic to the von-Neumann entropy $S(\rho)$.
+\noindent The derivation aligns with Schumacher's result by noting that the dimension of our Young frame subspace is asymptotic to the von-Neumann entropy $S(\rho)$.
 As $n \rightarrow \infty$, we observe that we cannot compress better than the rate given by $S(\rho)$.
@@
 
-\newpage
+\section{Conclusion}
+We introduced the Schur-Weyl Duality as a powerful tool allowing us to not only decompose $\mathcal{H}^{\otimes n}$ into tensor products of irreducible representations but also allowing us to utilize the Young Tableaux to easily calculate the dimensions of these irreducible representations.
+These methods yield deeper insight of the spectra of density operators as seen through the Keyl-Werner result.
+Instead of thinking about tensor products of eigenspaces of $\rho$, we can give good approximations through subspaces described by the more intuitive framework given by Young frames.
+These techniques give us slick proofs of the subadditivty property of von-Neumann entropy and a different interpretation of the Quantum Source Compression rate first derived by Schumacher.
+However, this is only scratching the bare surface of the deep intertwining of representation theory and quantum information theory.
+For a more detailed tract on the field, the reader is encouraged to read Aram Harrow's thesis \cite{harrow}.
+Matthias Christandl's thesis offers more insight into the role of Kronecker coefficients naturally arising from dimension calculations involving multipartite entangled systems \cite{chris}.
+
 \bibliography{biblio}

### 2020-06-29 KAA-Arch/samplepaper.tex
Added current Kaa draft.

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[EPiC]{easychair}
\usepackage{graphicx}
\usepackage[defaultlines=4,all]{nowidow}
\usepackage{booktabs}
\usepackage{amssymb,amsmath}
\usepackage{hyperref, subcaption}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Kaa: A Python Implementation of Reachable Set Computation Using Bernstein Polynomials}
%
\titlerunning{Kaa: Python Reachability Tool}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Edward Kim
%\orcidID{0000-1111-2222-3333} 
\and
Parasara Sridhar Duggirala
%\orcidID{1111-2222-3333-4444}}
}
%
\authorrunning{Kim and Duggirala}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science\\
University of North Carolina at Chapel Hill\\
\email{\{ehkim,psd\}@cs.unc.edu}
%\url{http://www.springer.com/gp/computer-science/lncs} \and
%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%\email{\{abc,lncs\}@uni-heidelberg.de}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Reachable set computation is one of the commonly used technique for verification of safety properties of dynamical system. 
%
One of the simplest algorithms for computing reachable set for discrete nonlinear systems uses parallelotope bundles and Bernstein polynomials.
%For discrete nonlinear systems, using Bernstein polynomials and parallelotope bundles for computing reachable set have been investigated in the literature.
%One of the reachable set computation techniques that has received considerable attention uses Bernstein polynomials and parallelotope bundles.
%
In this paper, we describe Kaa, a terse Python implementation of reachable set computation algorithm that leverages widely used symbolic package sympy.
%
Additionally, we simplify the user interface with the tool and provide easy to use plotting utilities.
%
%We also provide an easy to use interfact for plotting the reachable sets.
%
We believe that our tool, Kaa, has pedagogical use given the simplicity of the implementation and user-friendliness.
%
%In addition to the simplicity of the implementation, our tool, named Kaa, also achieves better performance on standard benchmarks.
\end{abstract}

\keywords{Nonlinear Dynamical Systems; \and Reachable Set Computation; \and Bernstein Polynomials.}

%
%
%
\section{Introduction}

Reachable set computation is one of the many important tools available for verification of dynamical and hybrid systems.
%
One of the simplest and easy-to-understand reachable set computation algorithm uses Bernstein polynomials and parallelotopes~\cite{dang2009image}.
%
Several cumulative improvements~\cite{dang2012reachability,sassi2012reachability,dang2014parameter,dreossi2016parallelotope,dreossi2017reachability} have been made to improve the accuracy and efficiency of nonlinear systems reachability.
%
The tool SAPO~\cite{dreossi2017sapo} that implements these algorithms in \cplusplus is also available publically.
%
In this paper, we present a Python implementation of algorithm presented in~\cite{dreossi2017reachability}.

%
Our primary motivation for reimplementing the algorithms in Python is two-fold.
%
First, using some of the standard libraries for symbolic manipulation in Python, the algorithm for reachability can be implemented very concisely.
%
In fact, Kaa is roughly only $650$ lines of code.
%
Additionally, the main algorithm for computing the reachable set is completely provided in just one file with $150$ lines of code.
%
In contrast, SAPO is more than $2000$ lines of \cplusplus code with memory and pointer management performed completely by the developer.
%
We believe that such a simple implementation has pedagogical value and can be used for teaching one of the main algorithms in reachability of nonlinear systems.
%
% Second, because of its implementation in C++, SAPO requires recompiling the code for every reachable set computation.
% %
Second, reimplementing the tool in Python makes the user interaction significantly easy. 
%
In SAPO, for computing the reachable set, one has to recompile the code along with the model file to obtain the binary and execute the binary to compute reachable set.
%
Additionally, SAPO generates a MATLAB script for visualizing time or phase plots of the reachable set.
%
In contrast, we provide a Jupyter notebook interface for Kaa that is very easy to use and leverage the matplotlib library for visualizing the reachable set.
%
%Therefore, the user has to 
%we believe that for reimplementing it in Python improves the quality of user interaction with the tool and removes the barrier to use the tools.
%

Bernstein polynomials are also considered an active area of research in the domain of global optimization~\cite{nataray2002algorithm,garloff2003bernstein,nataraj2007new}.
%
There have been several algorithms proposed that perform heuristics for improving the performance of optimization using Bernstein polynomials and also improve the accuracy using sequential refinement~\cite{smith2009fast,munoz2013formalization}.
%
Having an accessible and flexible tool would make it more conducive to implement these heuristics and determine specific heuristics helpful in the domain of reachability.
%
We believe Kaa can be used as a easy first step for introducing the reachability due to its simplicity of implementation and its usability.

\section{Preliminaries}

The state of a system, denoted as $x$, lies in a domain $D \subseteq \mathbb{R}^n$. A discrete-time polynomial nonlinear system is denoted as

\begin{equation}
    x^{+} = f(x)
\label{eq:system}
\end{equation}

where $f(x) : \mathbb{R}^n \rightarrow \mathbb{R}^n$ is polynomial in $x$. 
%
The trajectory of the system that evolves according to Equation~\ref{eq:system}, denoted as $\xi(x_0)$, is the sequence $x_0, x_1, \ldots$ where $x_{i+1} = f(x_i)$. 
%
The $k^{th}$ element in this sequence is denoted as $\xi(x_0,k)$.
%
Given an initial set $\Theta$, the reachable set at time $k$, denoted as $\Theta_{k} = \{~ \xi(x_0,k)~|~ x_0 \in \Theta\}$.

A parallelotope $P$ is a set of states in $\mathbb{R}^{n}$ denoted as $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ and $c \in \mathbb{R}^{2n}$, $\Lambda_{i+n} = -\Lambda_{i}$ and $i \in \{1, \ldots, n\}$ such that

$$
x \in P \mbox{ if and only if } \Lambda x \leq c.
$$
$\Lambda$ is called the \emph{direction matrix} $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$. $c$ is called the \emph{offset}. For a parallelotope $P$, we denote the affine transformation that maps $[0,1]^{n}$ to $P$ as $T_{p}$. A parallelotope bundle $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$ where $Q = \cap_{i=1}^{m}P_i$.

Given two multi-indices $i$ and $d$ of size $n$, where $i \leq d$, Bernstein polynomial of degree $d$ and index $i$ is 
$$
\mathcal{B}_{i,d} = \beta_{i_1,d_1}(x_1) \beta_{i_2,d_2}(x_2)\ldots \beta_{i_n,d_n}(x_n).
$$
%
where $\beta_{i_m,d_m}(x_m) = \binom{d_m}{i_m}x_{m}^{i_m}(1-x_m)^{d_m - i_m}$. 
%
Any polynomial function can be expressed in Bernstein basis. 
%
The primary advantage of Bernstein representation of a function is that the an upper bound on supremum and lower bound on infimum of a function in $[0,1]^{n}$ can be computed purely by looking at the coefficients of the function in the Bernstein basis.

Given a function $h(x) = \Sigma_{j=1}^{m} \alpha_j \mathcal{B}_{j}$, then
$$
min \{\alpha_1, \ldots, \alpha_m\} ~~\leq~~ inf_{x \in [0,1]^n} h(x) ~~~~~\leq~~~~ sup_{x \in [0,1]^{n}} h(x) ~~\leq~~ max \{\alpha_1, \ldots, \alpha_m\}.
$$

Any parallelotope $P$ can be considered as an affine transformation $T_{p}$ from $[0,1]^{n}$ to $P$. 
%
Therefore, upper bounds on the suprenum of a function $h$ over $P$ is equivalent to upper bound of $h \circ T_{p}$ over $[0,1]^{n}$.
%
Similar argument follows for lower bound on infimum.
%
In the remainder of the document, we assume that by using functional composition and Bernstein representation, we can easily compute upper bound on supremum and lower bound on infimum of polynomial functions over parallelotopes.
%
We denote the procedures for calculating such upper and lower bounds as $\mathsf{BernsteinUpper}$ and $\mathsf{BernsteinLower}$ respectively.

\section{Reachability of Nonlinear Systems Using Parallelotope Bundles and Bernstein Polynomials}

Given a set represented as a parallelotope bundle $Q = \{P_1, P_2, \ldots, P_m\}$, and a discrete dynamical system $x^{+} = f(x)$, we now present the method for computing an overapproximation of $f(Q)$ as parallelotope bundle $Q' = \{P_1', P_2', \ldots, P_m'\}$.
%
We assume that the direction matrix of $P_i'$ is same as $P_i$ and the computation is required only to compute the offsets. 
%
We denote the $j^{th}$ offset of $P_i$ and $P_i'$ as $c_{j,i}$ and $c'_{j,i}$ respectively. %
We also denote the $j^{th}$ direction in $P_i$ (same as $P_i'$) as $\Lambda_{j,i}$.
%
Therefore, given $\Lambda$, transformation $f$, and offsets $c_{j,i}$ for $P_i$, we have to compute the values of $c'_{j,i}$.

%
%Let us denote the $j^{th}$ offset in $P_i'$ and $c'_{j,i}$ and the direction be $\Lambda_{j,i}$.
%
If $j \leq n$, any offset $c'_{j,i}$ such that $\forall_{x\in Q} \Lambda_{j,i} \cdot f(x) \leq c'_{j,i}$ is valid.
%
%For any valid $c'_{j,i}$, it should be the case that, if $j \leq n$ $ \forall_{x \in Q} \Lambda_{j,i} \cdot f(x) \leq c'_{j,i}$.
%
If $j > n$, then any offset $c'_{j,i}$ such that $\forall_{x \in Q} \Lambda_{j-n,i} \cdot f(x) \geq c'_{j,i}$ is valid.
%
Therefore, one can compute the $j^{th}$ offset of $P_i'$, i.e., $c'_{j,i}$ for $j \leq n$ by computing an upper bound of the function $\Lambda_{j,i} \cdot f(x)$ over each parallelotope $x\in P_i$.
%The algorithm for computing the $j^{th}$ offset ($j \leq n$) for $P'_i$ is the following: iterature over all the parallelotopes in the bundle and compute the upper bounds of $\Lambda_{j,i} \cdot f(x)$ over each parallotope. 
%
Similarly, the $j+n^{th}$ offset can be computed from a lower bound of the function $\Lambda_{j,i} \cdot f(x)$ over each parallelotope $x\in P_i$.
%
This is given in Equations~\ref{eq:upper} and~\ref{eq:lower}.
% Sridhar here; I am done with changes.
% Sent you an email with specific directions for submission.
% Submit the paper and let's see what they say.

\begin{eqnarray}
c_{j,i} &=& min_{l = 1}^{m} \Big\{ \mathsf{BernsteinUpper}\big(\Lambda_{j,i} \cdot f(x) \mbox{ over } P_l \big) \Big\} \mbox{ if } j \leq n. \label{eq:upper} \\
c_{j+n,i} &=& max_{l = 1}^{m} \Big\{ \mathsf{BernsteinLower}\big(\Lambda_{j,i} \cdot f(x) \mbox{ over } P_l \big) \Big\} \mbox{ otherwise. } \label{eq:lower}
\end{eqnarray}

\subsection{Python Implementation of Reachable Set Computation}

One of the primary reasons for preferring Python as an implementation language for this algorithm is the presence of powerful, well-tested symbolic and matrix-computation libraries.
%
Python's \emph{numpy} libraries are popular matrix-computation libraries which allow higher-level manipulation of matrices. This gives us an avenue of overcoming the verbosity and the possibility of memory leaks inherent in implementing identical features in C++. We believe that overcoming these obstacles results in a more readable, more compact approach to parallelotope reachability suitable for graduate students and curious practitioners. Furthermore, these libraries are known for their extensive amount of useful functionalities, allowing us to avoid reimplementing many fundamental operations.

%
As we have seen before, there are two main sub-routines for computing the reachable set using Bernstein polynomials.
%
First, is performing functional composition for computing upper bound of a function on a parallelotope.
%
Second, computing the Bernstein representation of a polynomial.
%
The library of \emph{sympy} has powerful symbolic manipulation tools which allow us to comfortably perform many sensitive symbolic subsititutions into polynomials. 
%
We use \emph{sympy}'s internal representation of polynomials to (a) perform functional composition and (b) compute the Bernstein representations of the resulting polynomial.
%
Additionally, \emph{matplotlib} library has easy plotting facilities that we integrate into our tool for visualizing the reachable set. In particular, \emph{matplotlib} facilitates the ability to visualize several reachable sets simultaneously. This will aid others in performing more complex bundle-transformation experiments in future.

%
Finally, Python has a rich set of multiprocessing libraries, namely \emph{multiprocessing} or \emph{mp} for short. In the future, we plan to exploit the parallelizable nature of bundle-transformtion computations using \emph{mp}'s ability to provide powerful multiprocessing features without much verbosity.

% c_{j+n,i} &=& max_{l = 1}^{m} \big{ \mathsf{BernsteinLower} \Lambda_{j,i}\cdot f(x) \mbox{ over } P_l \big} \mbox{ otherwsie. } \label{eq:upper}

% \subsection{A Subsection Sample}
% Please note that the first paragraph of a section or subsection is
% not indented. The first paragraph that follows a table, figure,
% equation etc. does not need an indent, either.

% Subsequent paragraphs, however, are indented.

% \subsubsection{Sample Heading (Third Level)} Only two levels of
% headings should be numbered. Lower level headings remain unnumbered;
% they are formatted as run-in headings.

\section{Evaluations}
We evaluate our tool on the benchmarks that are shipped along with SAPO. With each benchmark, the reachable sets of SAPO are juxtaposed with that of Kaa's to demonstrate the accuracy and quality of our reachable set. A comparison of time taken by Kaa and SAPO for each benchmark is provided in Table~\ref{tab:one}.

\subsection{SIR Epidemic Model}
\noindent The SIR Epidemic model is a 3-dimensional dynamical system governed by the following dynamics:
\begin{align} \label{eqn:sir}
    \begin{split}
        s_{k+1} &= s_k - (\beta s_k i_k)\Delta \\
        i_{k+1} &= i_k + (\beta s_k i_k - \gamma i_k)\Delta \\
        r_{k+1} &= r_k + (\gamma i_k)\Delta
    \end{split}
\end{align} 
where $s,i,r$ represent the fractions of a population of individuals designated as \textit{susceptible}, \textit{infected}, and \textit{recovered} respectively. There are two parameters, namely $\beta$ and $\gamma$, which influence the evolution of the system. $\beta$ is labeled as the contraction rate and $1/\gamma$ is mean infective period. Finally, $\Delta$ is simply the discretization step. For the benchmarks, we set $\beta = 0.34$, $\gamma=0.05$, and $\Delta=0.5$. The reachable sets are displayed in Figure \ref{fig1}. The table of numerical value comparisions between Sapo and Kaa is shown in Table \ref{val1}.

\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|c|c|}
    \hline
    Time Steps & Kaa (OffU) & Sapo (OffU) & Kaa (OffL) & Kaa (OffL)  \\
    \hline
     50 & 0.470716 & 0.470716 & -0.435191 & -0.43519 \\
    \hline
     51 & 0.475839 & 0.475839 & -0.439599 & -0.439599 \\ 
    \hline
     52 & 0.480906 & 0.480906 & -0.443945 & -0.443945\\
    \hline
     53 & 0.485915 &  0.485915 & -0.448227 &  -0.448227 \\ 
    \hline
     54 & 0.490862 & 0.490862 & -0.452443 & -0.452443 \\ 
    \hline
     55 & 0.495747 & 0.495747 & -0.456591 & -0.456591 \\
    \hline
     56 & 0.500566 & 0.500566 & -0.460669 & -0.460669 \\ 
    \hline
     57 & 0.505317 & 0.505317 & -0.464675 & -0.464675\\ 
    \hline
     58 & 0.509999 & 0.509999  & -0.4686075 & -0.468608 \\ 
    \hline
     59 & 0.514610 & 0.514610 & -0.472465 & -0.472465\\
    \hline
     60 & 0.519147 & 0.519147 & -0.476246 & -0.476246 \\
    \hline
\end{tabular}
\caption{Comparision of offu, offl values along variable $i$ between steps 50-60}
\label{val1}
\end{table}
\noindent Here, OffU is the vector of upper offsets of the parallelotope and OffL is the vector for the lower offsets. We define the upper facets of parallelotope $P$ as the $c_i$ for $i \leq n$ where $n$ is the dimension of the system and $P = \langle \Lambda, c \rangle$ is the parallelotope as defined above. Similarly, the lower facets are defined as the $c_{i+n}$ for $i \leq n$. In this case, we are looking at values in $c_2$ and $c_{2+3} = c_5$.

\subsection{Rossler Model}
\noindent The Rossler model is another 3-dimensional system governed under the dynamics:
\begin{align} \label{rossler}
    \begin{split}
        x_{k+1} &= x_k + -(y-z)\Delta \\
        y_{k+1} &= y_k + (x_k + ay_k)\Delta \\
        z_{k+1} &= z_k + (b + z_k(x_k - c))\Delta
    \end{split}
\end{align}
where $a,b,c$ are parameters which we set to $a = 0,1$, $b =0.1$, and $c = 14$. We set our discretization step to be $\Delta= 0.025$.
The reachable sets are displayed in Figure \ref{fig2} and the table showing the comparisions between the numerical values are shown in Table \ref{val2}.

% Sridhar here: I think just giving two of these tables should be fine. No need to give tables for all of them.
% Also, there is one figure missing for KAA. Do you plan to add that?
% One thing to do is to do a screen capture of the image and upload the screen capture. Supremely hacky!
% Cool.
% Will do. Thanks.
% Yes. One figure fails to upload to Overleaf for some reason. I have to find an alternative way to upload to Overlead somehow.
% Good point. I'll look into it once I finish this table here.

\begin{table}[hbt!]
\centering
\begin{tabular}{|c|c|c|c|c|}
    \hline
    Time Steps & Kaa (OffU) & Sapo (OffU) & Kaa (OffL) & Kaa (OffL)  \\
    \hline
    50 & 1.95908 & 1.96043  & -1.9209 & -1.9193 \\
    \hline
    51 & 1.83552 & 1.83688  & -1.7963 & -1.7947 \\
    \hline
    52 & 1.71044 & 1.71181  & -1.67016 & -1.6685 \\
    \hline
    53 & 1.58392 & 1.58531  & -1.54255 & -1.5409 \\
    \hline
    54 & 1.45604 & 1.45744  & -1.41355 & -1.4119 \\
    \hline
    55 & 1.32687 & 1.32829  & -1.28324 & -1.2815\\
    \hline
    56 & 1.19649 & 1.19792  & -1.15168 & -1.1500\\
    \hline
    57 & 1.06499 & 1.06642  & -1.01896 & -1.0172 \\
    \hline
    58 & 0.932432 & 0.933877 & -0.885157 & -0.88339 \\
    \hline
    59 & 0.798905 & 0.800358  & -0.75035 &  -0.74857 \\
    \hline
    60 & 0.664489 & 0.665949  & -0.614619 & -0.61282\\
    \hline
\end{tabular}
\caption{Comparision of offU, offL values along variable $y$ between steps 50-60}
\label{val2}
\end{table}

\begin{figure}[h]

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_S.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/SIR/KaaSIR_S.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_I.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/SIR/KaaSIR_I.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_R.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/SIR/KaaSIR_R.png}
    \end{subfigure}

    \caption{Figure depicting the reachable set computation of the SIR model.} 
    \label{fig1}
\end{figure}

\begin{figure}[h]

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerX.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/KaaRosslerX.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerY.jpg}
    \end{subfigure}
    \begin{subfigure}{0.65\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/KaaRosslerY.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerZ.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/KaaRosslerZ.png}
    \end{subfigure}

    \caption{Figure depicting the reachable set computation of the Rossler model.} 
    \label{fig2}
\end{figure}

\subsection{Quadcopter Model}
\noindent The Quadcopter model is a 17-dimensional dynamical system with the following variables:
A set of inertial positions: $(p_n,p_e,h)$, linear velocities $(u,v,w)$, the quaternions expressing the Euler angles $(q_0,q_1,q_2,q_3)$, the angular velocities $(p,q,r)$, and finally the parameters $(h_I, u_I, v_I, \psi_I)$. The complete set of dynamics and relevant parameters is found in~\cite{dreossi2017reachability} and reachable sets are shown in Figure \ref{fig3}.

\begin{figure}[h]

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_H.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Quad/KaaQuad_H.jpg}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_W.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Quad/KaaQuad_W.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_HI.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Quad/KaaQuad_HI.png}
    \end{subfigure}

    \caption{Figure depicting the reachable set computation of the Quadcopter model.} 
    \label{fig3}
\end{figure}

% Sridhar: I added the figures, it wasn't that much effort.
% I see. Thank you.

\subsection{Lotka-Volterra Model}
\noindent We also test on the competetitive Lotka-Volterra models for predator-prey biological systems. Our instance is a 5-dimensional sysmtem governed as below:
\begin{align}
    \begin{split}
        x_{k+1} &= x_k + (x_k(1 - (x_k + \alpha y_k + \beta \ell_k))) \Delta \\
        y_{k+1} &= y_k + (y_k(1 - (y_k + \alpha z_k + \beta x_k))) \Delta \\
        z_{k+1} &= z_k + (z_k(1 - (z_k + \alpha h_k + \beta y_k))) \Delta \\
        h_{k+1} &= h_k + (h_k(1 - (h_k + \alpha \ell_k + \beta z_k))) \Delta \\
        \ell_{k+1} &= \ell_k + (\ell_k(1 - (\ell_k + \alpha x_k + \beta h_k))) \Delta
    \end{split}
\end{align}
where $\alpha,\beta$ are parameters set to $\alpha=0.85$ and $\beta=0.5$. We set $\Delta = 0.01$.
The relevant reachable sets are shown in Figure \ref{fig4}.

\begin{figure}[h]

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X1.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/LV/KaaLV_X1.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X2.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/LV/KaaLV_X2.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X3.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/LV/KaaLV_X3.png}
    \end{subfigure}

    \caption{Figure depicting the reachable set computation of the Lotka-Volterra model.} 
    \label{fig4}
\end{figure}

\subsection{Phosphoraley Model}
The Phosphoraley model describes a certain cellular regulatory system. It is captured by seven variables governed by the following dynamics:
\begin{align}
    \begin{split}
        x^1_{k+1} &= x^1_k + ( -\alpha x^1_k + \beta x^3_k x^4_k)\Delta \\
        x^2_{k+1} &= x^2_k + (  \alpha x^1_k - x^2_k)\Delta \\
        x^3_{k+1} &= x^3_k + ( x^2_k \beta x^3_k x^4_k)\Delta \\
        x^4_{k+1} &= x^4_k + ( \beta x^5_k x^6_k - \beta x^3_k x^4_k)\Delta \\
        x^5_{k+1} &= x^5_k + ( -\beta x^5_k x^6_k + \beta x^3_k x^4_k)\Delta \\
        x^6_{k+1} &= x^6_k + ( \alpha x^7_k - \beta x^5_k x^6_k)\Delta \\
        x^7_{k+1} &= x^7_k + ( -\alpha x^7_k + \beta x^5_k x^6_k)\Delta
    \end{split}
\end{align}
where $\alpha,\beta$ are two parameters assigned as $\alpha=0.5$ and $\beta=5$. We set $\Delta =0.01$ here. The relevant figures can be found in Figure \ref{fig5}.

\begin{figure}[h]

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X1.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Phos/KaaPhos_X1.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X2.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Phos/KaaPhos_X2.png}
    \end{subfigure}

    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X3.jpg}
    \end{subfigure}
    \begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{SapoFigures/Phos/KaaPhos_X3.png}
    \end{subfigure}

    \caption{Figure depicting the reachable set computation of the Phosphoraley model.} 
    \label{fig5}
\end{figure}

\begin{table}[h]
\centering
\caption{Reachable Set Computation Time of Benchmarks}\label{tab:one}
\begin{tabular}{|l|l|l|}
\hline
Model & Kaa & sapo (C++)\\
\hline
SIR &  11.41 sec & 0.16 sec\\
\hline
Rossler &  41.92 sec & 1.17 sec\\
\hline
Quadcopter & 78.21 sec & 11.98 sec \\
\hline
Lotka-Volterra & 18 min 95.05 sec  & 57.48 sec\\
 \hline
 Phosphoraley & 103.81 sec & 24.86 sec \\
\hline
\end{tabular}
\end{table}

\noindent It is evident from Table~\ref{tab:one} that while implementation in Python is very intuitive and concise, the current implementation incurs severe performance penalties. 
%
Nevertheless, we pursued this activity because of the pedagogical value in making this an accessible tool for implementing and understanding reachability. To this end, the repository contains a Jupyter notebook with the name \href{https://github.com/Tarheel-Formal-Methods/kaa/blob/master/kaa-intro.ipynb}{kaa-intro} designed to interactively introduce graduate students and practitioners to the usage of Kaa. The notebook even contains a subset of the examples shown above, giving the auidence a chance to run the code for themselves. An immediate next step is to deploy extensive profiling to find performance bottlenecks and subsequently improve on them.

\section{Conclusions}

We present Kaa, a Python implementation of reachable set computation of nonlinear systems that is focussed towards accessibility and pedagocial use.
%
The usage of inbuilt \emph{sympy} libraries makes the implementation very short and simple (only 650 LOC).
%
While we do incur performance drawbacks from selecting Python for implementing this algorithm, we believe that it aids in fast prototyping and enables students to easily build on top of the library. In particular, Python's readability and extendibility will allow curious students to experiment with more complex bundle-transformations. 
%
The code can be accessed easily at \url{https://github.com/Tarheel-Formal-Methods/kaa}

% \begin{figure}
% \includegraphics[width=\textwidth]{Figures/fig1.eps}
% \caption{A figure caption is always placed below the illustration.
% Please note that short captions are centered, while long ones are
% justified by the macro package automatically.} \label{fig1}
% \end{figure}

% For citations of references, we prefer the use of square brackets
% and consecutive numbers. Citations using labels or the author/year
% convention are also acceptable. The following bibliography provides
% a sample reference list with entries for journal
% articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
% book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
% and a homepage~\cite{ref_url1}. Multiple citations are grouped
% \cite{ref_article1,ref_lncs1,ref_book1},
% \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{archbib}
%
% \begin{thebibliography}{8}
% \bibitem{ref_article1}
% Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

% \bibitem{ref_lncs1}
% Author, F., Author, S.: Title of a proceedings paper. In: Editor,
% F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
% Springer, Heidelberg (2016). \doi{10.10007/1234567890}

% \bibitem{ref_book1}
% Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
% Location (1999)

% \bibitem{ref_proc1}
% Author, A.-B.: Contribution title. In: 9th International Proceedings
% on Proceedings, pp. 1--2. Publisher, Location (2010)

% \bibitem{ref_url1}
% LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
% Oct 2017
% \end{thebibliography}
\end{document}

### 2020-07-06 KAA-Arch/samplepaper.tex
New draft of Kaa

@@
 \begin{abstract}
-Reachable set computation is one of the commonly used technique for verification of safety properties of dynamical system. 
+Reachable set computation is one of the widely used techniques for verification of safety properties of dynamical system. 
 %
-One of the simplest algorithms for computing reachable set for discrete nonlinear systems uses parallelotope bundles and Bernstein polynomials.
+One of the simplest algorithms for computing reachable sets for discrete nonlinear systems uses parallelotope bundles and Bernstein polynomials.
 %For discrete nonlinear systems, using Bernstein polynomials and parallelotope bundles for computing reachable set have been investigated in the literature.
@@
 %
-In this paper, we describe Kaa, a terse Python implementation of reachable set computation algorithm that leverages widely used symbolic package sympy.
+In this paper, we describe Kaa, a terse Python implementation of reachable set computation which leverages the widely used symbolic package sympy.
 %
@@
 %
-We believe that our tool, Kaa, has pedagogical use given the simplicity of the implementation and user-friendliness.
+We believe that our tool has pedagogical use given the simplicity of the implementation and its user-friendliness.
 %
@@
 %
-One of the simplest and easy-to-understand reachable set computation algorithm uses Bernstein polynomials and parallelotopes~\cite{dang2009image}.
+One of the simpler and easier-to-understand reachable set computation algorithms utilizes Bernstein polynomials and parallelotopes~\cite{dang2009image}.
 %
@@
 %
-The tool SAPO~\cite{dreossi2017sapo} that implements these algorithms in \cplusplus is also available publically.
+The tool SAPO~\cite{dreossi2017sapo} which implements these algorithms in C\texttt{++} is also available publically.
 %
@@
 %
-First, using some of the standard libraries for symbolic manipulation in Python, the algorithm for reachability can be implemented very concisely.
+First, using some of the standard libraries for symbolic manipulation in Sympy, the algorithm for reachability can be implemented very concisely.
 %
@@
 %
-Additionally, the main algorithm for computing the reachable set is completely provided in just one file with $150$ lines of code.
+Additionally, the main algorithm for computing the reachable set is provided in just one file with $150$ lines of code.
 %
-In contrast, SAPO is more than $2000$ lines of \cplusplus code with memory and pointer management performed completely by the developer.
+In contrast, SAPO is more than $2000$ lines of C\texttt{++} code with memory and pointer management performed completely by the developer.
 %
-We believe that such a simple implementation has pedagogical value and can be used for teaching one of the main algorithms in reachability of nonlinear systems.
+We believe that such a simple implementation has pedagogical value and can be used for teaching one of the main algorithms for reachability of nonlinear systems.
 %
@@
 % %
-Second, reimplementing the tool in Python makes the user interaction significantly easy. 
+Second, reimplementing the tool in Python makes user interaction significantly easy. 
 %
@@
 %
-Additionally, SAPO generates a MATLAB script for visualizing time or phase plots of the reachable set.
+Additionally, SAPO generates a separate MATLAB script for visualizing time or phase plots of the reachable set.
+This script must be fed separately into either MATLAB or Octave to illustrate the reachable set.
 %
-In contrast, we provide a Jupyter notebook interface for Kaa that is very easy to use and leverage the matplotlib library for visualizing the reachable set.
+In contrast, we provide a Jupyter notebook interface for Kaa to create an intuitive platform facilitating experimentation.
+Juypter notebooks are simple to create and known for their straightforward user interface.
+By leveraging the matplotlib library for visualizing the reachable set, we were able to design an engaging interactive tutorial and experimentation platform for visualzing reachable sets of non-linear systems.
 %
@@
 %
-There have been several algorithms proposed that perform heuristics for improving the performance of optimization using Bernstein polynomials and also improve the accuracy using sequential refinement~\cite{smith2009fast,munoz2013formalization}.
+There have been several algorithms proposed for performing heuristics which improve the performance of optimization using Bernstein polynomials.
+Furthermore, these algorithms also improve the accuracy using sequential refinement techniques~\cite{smith2009fast,munoz2013formalization}.
 %
@@
 %
-We believe Kaa can be used as a easy first step for introducing the reachability due to its simplicity of implementation and its usability.
+In light of these features and advantages, we believe Kaa can be used as a easy first step for introducing reachability due to its simplicity of implementation and its usability.
 
@@
 %
-The trajectory of the system that evolves according to Equation~\ref{eq:system}, denoted as $\xi(x_0)$, is the sequence $x_0, x_1, \ldots$ where $x_{i+1} = f(x_i)$. 
+The trajectory of the system that evolves according to Equation~\eqref{eq:system}, denoted as $\xi(x_0)$, is the sequence $x_0, x_1, \ldots$ where $x_{i+1} = f(x_i)$. 
 %
-The $k^{th}$ element in this sequence is denoted as $\xi(x_0,k)$.
+The $k^{th}$ element in this sequence is denoted as $x_k = \xi(x_0,k)$.
 %
@@
 
-Given two multi-indices $i$ and $d$ of size $n$, where $i \leq d$, Bernstein polynomial of degree $d$ and index $i$ is 
+Given two multi-indices $i$ and $d$ of size $n$, where $i \leq d$, the Bernstein polynomial of degree $d$ and index $i$ is 
 $$
@@
 %
-Any polynomial function can be expressed in Bernstein basis. 
+Any polynomial function can be expressed in the Bernstein basis. 
 %
-The primary advantage of Bernstein representation of a function is that the an upper bound on supremum and lower bound on infimum of a function in $[0,1]^{n}$ can be computed purely by looking at the coefficients of the function in the Bernstein basis.
+The primary advantage of the Bernstein representation of a polynomial $p(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $p(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis.
 
-Given a function $h(x) = \Sigma_{j=1}^{m} \alpha_j \mathcal{B}_{j}$, then
+In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
+The upper and lower bounds of $p(x_1,\ldots,x_n)$ over $[0,1]^n$ are exactly bounded by the Bernstein coefficients:
 $$
-min \{\alpha_1, \ldots, \alpha_m\} ~~\leq~~ inf_{x \in [0,1]^n} h(x) ~~~~~\leq~~~~ sup_{x \in [0,1]^{n}} h(x) ~~\leq~~ max \{\alpha_1, \ldots, \alpha_m\}.
+min \{b_1, \ldots, b_m\} ~~\leq~~ inf_{x \in [0,1]^n} h(x) ~~~~~\leq~~~~ sup_{x \in [0,1]^{n}} h(x) ~~\leq~~ max \{b, \ldots, b_m\}.
 $$
 
-Any parallelotope $P$ can be considered as an affine transformation $T_{p}$ from $[0,1]^{n}$ to $P$. 
+Any parallelotope $P$ can be considered as a suitable affine transformation $T_{p}$ from $[0,1]^{n}$ to $P$. 
 %
@@
 %
-Similar argument follows for lower bound on infimum.
+A similar argument follows for the lower bound on infimum.
 %
-In the remainder of the document, we assume that by using functional composition and Bernstein representation, we can easily compute upper bound on supremum and lower bound on infimum of polynomial functions over parallelotopes.
+For the remainder of the document, we assume that by using functional composition and the Bernstein representation, we can easily compute the upper bound on supremum and lower bound on infimum of polynomial functions over parallelotopes.
 %
-We denote the procedures for calculating such upper and lower bounds as $\mathsf{BernsteinUpper}$ and $\mathsf{BernsteinLower}$ respectively.
+
+Furthermore, we denote the procedures for calculating such upper and lower bounds for a polynomial $p$ over some parallelotope $S$ as $\mathsf{BernsteinUpper}(p,S)$ and $\mathsf{BernsteinLower}(p,S)$ respectively.
 
@@
 
-Given a set represented as a parallelotope bundle $Q = \{P_1, P_2, \ldots, P_m\}$, and a discrete dynamical system $x^{+} = f(x)$, we now present the method for computing an overapproximation of $f(Q)$ as parallelotope bundle $Q' = \{P_1', P_2', \ldots, P_m'\}$.
+Given a set represented as a parallelotope bundle $Q = \{P_1, P_2, \ldots, P_m\}$ and a discrete dynamical system $x^{+} = f(x)$, we now present the method for computing an overapproximation of $f(Q)$ as a parallelotope bundle $Q' = \{P_1', P_2', \ldots, P_m'\}$.
 %
-We assume that the direction matrix of $P_i'$ is same as $P_i$ and the computation is required only to compute the offsets. 
+We exert that the direction matrix of $P_i'$ is same as $P_i$ and the computation is required only to compute the offsets. 
 %
-We denote the $j^{th}$ offset of $P_i$ and $P_i'$ as $c_{j,i}$ and $c'_{j,i}$ respectively. %
-We also denote the $j^{th}$ direction in $P_i$ (same as $P_i'$) as $\Lambda_{j,i}$.
+Denote the $j^{th}$ offset of $P_i$ and $P_i'$ as $c_{j,i}$ and $c'_{j,i}$ respectively and the $j^{th}$ direction in $P_i$ (same as $P_i'$) as $\Lambda_{j,i}$.
 %
@@
 %
-Therefore, one can compute the $j^{th}$ offset of $P_i'$, i.e., $c'_{j,i}$ for $j \leq n$ by computing an upper bound of the function $\Lambda_{j,i} \cdot f(x)$ over each parallelotope $x\in P_i$.
+Therefore, one can compute the $j^{th}$ offset of $P_i'$, i.e., $c'_{j,i}$ for $j \leq n$ by computing an upper bound of the function $\Lambda_{j,i} \cdot f(x)$ over the $x\in P_i$.
 %The algorithm for computing the $j^{th}$ offset ($j \leq n$) for $P'_i$ is the following: iterature over all the parallelotopes in the bundle and compute the upper bounds of $\Lambda_{j,i} \cdot f(x)$ over each parallotope.
@@
 \begin{eqnarray}
-c_{j,i} &=& min_{l = 1}^{m} \Big\{ \mathsf{BernsteinUpper}\big(\Lambda_{j,i} \cdot f(x) \mbox{ over } P_l \big) \Big\} \mbox{ if } j \leq n. \label{eq:upper} \\
-c_{j+n,i} &=& max_{l = 1}^{m} \Big\{ \mathsf{BernsteinLower}\big(\Lambda_{j,i} \cdot f(x) \mbox{ over } P_l \big) \Big\} \mbox{ otherwise. } \label{eq:lower}
+c_{j,i} &=& min_{l = 1}^{m} \Big\{ \mathsf{BernsteinUpper}\big(\Lambda_{j,i} \cdot f(x), P_l \big) \Big\} \mbox{ if } j \leq n. \label{eq:upper} \\
+c_{j+n,i} &=& max_{l = 1}^{m} \Big\{ \mathsf{BernsteinLower}\big(\Lambda_{j,i} \cdot f(x), P_l \big) \Big\} \mbox{ otherwise. } \label{eq:lower}
 \end{eqnarray}
@@
 We believe that overcoming these obstacles results in a more readable, more compact approach to parallelotope reachability suitable for graduate students and curious practitioners.
-Furthermore, these libraries are known for their extensive amount of useful functionalities, allowing us to avoid reimplementing many fundamental operations.
+In addition, these libraries are known for their extensive set of useful functionalities which allow us to avoid reimplementing many fundamental operations.
 
@@
 %
-First, is performing functional composition for computing upper bound of a function on a parallelotope.
+First, performing functional composition for computing uthe pper bound of a polynomial over a parallelotope.
 %
@@
 %
-Additionally, \emph{matplotlib} library has easy plotting facilities that we integrate into our tool for visualizing the reachable set.
+Additionally, the \emph{matplotlib} library has easy plotting facilities that we integrate into our tool for visualizing the reachable set.
 In particular, \emph{matplotlib} facilitates the ability to visualize several reachable sets simultaneously.
@@
 \centering
-\begin{tabular}{|c|c|c|c|c|}
+\begin{tabular}{|c c c c c|}
     \hline
-    Time Steps & Kaa (OffU) & Sapo (OffU) & Kaa (OffL) & Kaa (OffL)  \\
+    {\bf Time Steps} & {\bf Kaa (offu) }  & { \bf Sapo (offu) } & {\bf Kaa (offl)} & {\bf Kaa (offl) }  \\
     \hline
      50 & 0.470716 & 0.470716 & -0.435191 & -0.43519 \\
-    \hline
+
      51 & 0.475839 & 0.475839 & -0.439599 & -0.439599 \\
-    \hline
+
      52 & 0.480906 & 0.480906 & -0.443945 & -0.443945\\
-    \hline
+
      53 & 0.485915 &  0.485915 & -0.448227 &  -0.448227 \\
-    \hline
+
      54 & 0.490862 & 0.490862 & -0.452443 & -0.452443 \\
-    \hline
+
      55 & 0.495747 & 0.495747 & -0.456591 & -0.456591 \\
-    \hline
+
      56 & 0.500566 & 0.500566 & -0.460669 & -0.460669 \\
-    \hline
+
      57 & 0.505317 & 0.505317 & -0.464675 & -0.464675\\
-    \hline
+
      58 & 0.509999 & 0.509999  & -0.4686075 & -0.468608 \\
-    \hline
+
      59 & 0.514610 & 0.514610 & -0.472465 & -0.472465\\
-    \hline
+
      60 & 0.519147 & 0.519147 & -0.476246 & -0.476246 \\
@@
 \end{tabular}
-\caption{Comparision of offu, offl values along variable $i$ between steps 50-60}
+\caption{Comparision of offu, offl values along variable $i$ for the SIR model. We select the steps 50-60}
 \label{val1}
 \end{table}
-\noindent Here, OffU is the vector of upper offsets of the parallelotope and OffL is the vector for the lower offsets.
-We define the upper facets of parallelotope $P$ as the $c_i$ for $i \leq n$ where $n$ is the dimension of the system and $P = \langle \Lambda, c \rangle$ is the parallelotope as defined above.
+\noindent Here, offu is the vector of upper offsets of the parallelotope and offl is the vector for the lower offsets.
+We define the upper facets of parallelotope $P$ as the $c_i$ for $i \leq n$ where $n$ is the dimension of the system and $P = \langle \Lambda, c \rangle$ is the parallelotope.
 Similarly, the lower facets are defined as the $c_{i+n}$ for $i \leq n$.
-In this case, we are looking at values in $c_2$ and $c_{2+3} = c_5$.
+In the case above, we are looking at values in $c_2$ and $c_{2+3} = c_5$.
 
@@
 \centering
-\begin{tabular}{|c|c|c|c|c|}
+\begin{tabular}{|c c c c c|}
     \hline
-    Time Steps & Kaa (OffU) & Sapo (OffU) & Kaa (OffL) & Kaa (OffL)  \\
+    {\bf Time Steps } & {\bf Kaa (offu) } & {\bf Sapo (offu) } & {\bf Kaa (offl) } & {\bf Kaa (offl) }  \\
     \hline
     50 & 1.95908 & 1.96043  & -1.9209 & -1.9193 \\
-    \hline
+
     51 & 1.83552 & 1.83688  & -1.7963 & -1.7947 \\
-    \hline
+
     52 & 1.71044 & 1.71181  & -1.67016 & -1.6685 \\
-    \hline
+
     53 & 1.58392 & 1.58531  & -1.54255 & -1.5409 \\
-    \hline
+
     54 & 1.45604 & 1.45744  & -1.41355 & -1.4119 \\
-    \hline
+
     55 & 1.32687 & 1.32829  & -1.28324 & -1.2815\\
-    \hline
+
     56 & 1.19649 & 1.19792  & -1.15168 & -1.1500\\
-    \hline
+
     57 & 1.06499 & 1.06642  & -1.01896 & -1.0172 \\
-    \hline
+
     58 & 0.932432 & 0.933877 & -0.885157 & -0.88339 \\
-    \hline
+
     59 & 0.798905 & 0.800358  & -0.75035 &  -0.74857 \\
-    \hline
+
     60 & 0.664489 & 0.665949  & -0.614619 & -0.61282\\
@@
 \end{tabular}
-\caption{Comparision of offU, offL values along variable $y$ between steps 50-60}
+\caption{Comparision of offu, offl values along variable $y$ for the Rossler model. We select the steps 50-60}
 \label{val2}
 \end{table}
-
-\begin{figure}[h]
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_S.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/SIR/KaaSIR_S.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_I.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/SIR/KaaSIR_I.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_R.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/SIR/KaaSIR_R.png}
-    \end{subfigure}
-
-    \caption{Figure depicting the reachable set computation of the SIR model.} 
-    \label{fig1}
-\end{figure}
-
-\begin{figure}[h]
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerX.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/KaaRosslerX.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerY.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.65\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/KaaRosslerY.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerZ.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/KaaRosslerZ.png}
-    \end{subfigure}
-
-    \caption{Figure depicting the reachable set computation of the Rossler model.} 
-    \label{fig2}
-\end{figure}
 
@@
 The complete set of dynamics and relevant parameters is found in~\cite{dreossi2017reachability} and reachable sets are shown in Figure \ref{fig3}.
-
-\begin{figure}[h]
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_H.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Quad/KaaQuad_H.jpg}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_W.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Quad/KaaQuad_W.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_HI.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Quad/KaaQuad_HI.png}
-    \end{subfigure}
-
-    \caption{Figure depicting the reachable set computation of the Quadcopter model.} 
-    \label{fig3}
-\end{figure}
 
@@
 
-\begin{figure}[h]
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X1.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/LV/KaaLV_X1.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X2.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/LV/KaaLV_X2.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X3.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/LV/KaaLV_X3.png}
-    \end{subfigure}
-
-    \caption{Figure depicting the reachable set computation of the Lotka-Volterra model.} 
-    \label{fig4}
-\end{figure}
-
 \subsection{Phosphoraley Model}
@@
 
-\begin{figure}[h]
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X1.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Phos/KaaPhos_X1.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X2.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Phos/KaaPhos_X2.png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X3.jpg}
-    \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Phos/KaaPhos_X3.png}
-    \end{subfigure}
-
-    \caption{Figure depicting the reachable set computation of the Phosphoraley model.} 
-    \label{fig5}
-\end{figure}
-
-\begin{table}[h]
-\centering
-\caption{Reachable Set Computation Time of Benchmarks}\label{tab:one}
-\begin{tabular}{|l|l|l|}
-\hline
-Model & Kaa & sapo (C++)\\
-\hline
-SIR &  11.41 sec & 0.16 sec\\
-\hline
-Rossler &  41.92 sec & 1.17 sec\\
-\hline
-Quadcopter & 78.21 sec & 11.98 sec \\
-\hline
-Lotka-Volterra & 18 min 95.05 sec  & 57.48 sec\\
- \hline
- Phosphoraley & 103.81 sec & 24.86 sec \\
-\hline
-\end{tabular}
-\end{table}
+\subsection{Times}
 
@@
 
+\begin{table}[h]
+\centering
+\caption{Reachable Set Computation Time of Benchmarks}\label{tab:one}
+\begin{tabular}{|l | l l|}
+\hline
+{\bf Model} &{\bf Kaa } & {\bf SAPO } (C\texttt{++})\\
+\hline
+SIR &  11.41 sec & 0.16 sec\\
+
+Rossler &  41.92 sec & 1.17 sec\\
+
+Quadcopter & 78.21 sec & 11.98 sec \\
+
+Lotka-Volterra & 18 min 95.05 sec  & 57.48 sec\\
+
+ Phosphoraley & 103.81 sec & 24.86 sec \\
+\hline
+\end{tabular}
+\end{table}
+
 \section{Conclusions}
@@
 While we do incur performance drawbacks from selecting Python for implementing this algorithm, we believe that it aids in fast prototyping and enables students to easily build on top of the library.
-In particular, Python's readability and extendibility will allow curious students to experiment with more complex bundle-transformations.
+In particular, Python's readability and extendibility will allow curious students to experiment with more involved bundle-transformations.
 %
-The code can be accessed easily at \url{https://github.com/Tarheel-Formal-Methods/kaa}
+The code can be accessed through \url{https://github.com/Tarheel-Formal-Methods/kaa}
+
+\begin{figure}[h]
+    \hspace{-15ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_S.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.2\textwidth]{SapoFigures/SIR/KaaSIR_S.png}
+    \end{subfigure}
+
+    \hspace{-15ex}   
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_I.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.2\textwidth]{SapoFigures/SIR/KaaSIR_I.png}
+    \end{subfigure}
+
+    \hspace{-15ex}  
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{SapoFigures/SIR/SapoSIR_R.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.2\textwidth]{SapoFigures/SIR/KaaSIR_R.png}
+    \end{subfigure}
+
+    \caption{Figure depicting the reachable set computation of the SIR model.} 
+    \label{fig1}
+\end{figure}
+
+\begin{figure}[h]
+    \hspace{-15ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerX.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.1\textwidth]{SapoFigures/Rosslert/KaaRosslerX.png}
+    \end{subfigure}
+
+     \hspace{-15ex}  
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerY.jpg}
+    \end{subfigure}
+    \hspace{-3ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.4\textwidth,height=0.9\textwidth]{SapoFigures/Rosslert/KaaRosslerY.png}
+    \end{subfigure}
+
+    \hspace{-15ex}  
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerZ.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.25\textwidth]{SapoFigures/Rosslert/KaaRosslerZ.png}
+    \end{subfigure}
+
+    \caption{Figure depicting the reachable set computation of the Rossler model.} 
+    \label{fig3}
+\end{figure}
+
+\begin{figure}[h]
+    \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.2\textwidth]{SapoFigures/Quad/SapoQuad_H.jpg}
+    \end{subfigure}
+    \hspace{5ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.13\textwidth]{SapoFigures/Quad/KaaQuad_H.jpg}
+    \end{subfigure}
+
+     \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.2\textwidth]{SapoFigures/Quad/SapoQuad_W.jpg}
+    \end{subfigure}
+    \hspace{5ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.13\textwidth]{SapoFigures/Quad/KaaQuad_W.png}
+    \end{subfigure}
+
+    \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.15\textwidth]{SapoFigures/Quad/SapoQuad_HI.jpg}
+    \end{subfigure}
+    \hspace{5ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.13\textwidth]{SapoFigures/Quad/KaaQuad_HI.png}
+    \end{subfigure}
+
+    \caption{Figure depicting the reachable set computation of the Quadcopter model.} 
+    \label{fig3}
+\end{figure}
+
+\begin{figure}[h]
+    \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.2\textwidth]{SapoFigures/LV/SapoLV_X1.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.14\textwidth]{SapoFigures/LV/KaaLV_X1.png}
+    \end{subfigure}
+
+     \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.2\textwidth]{SapoFigures/LV/SapoLV_X2.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.14\textwidth]{SapoFigures/LV/KaaLV_X2.png}
+    \end{subfigure}
+
+    \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.15\textwidth]{SapoFigures/LV/SapoLV_X3.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.14\textwidth]{SapoFigures/LV/KaaLV_X3.png}
+    \end{subfigure}
+
+    \caption{Figure depicting the reachable set computation of the Lotka-Volterra model.} 
+    \label{fig3}
+\end{figure}
+
+\begin{figure}[h]
+    \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.15\textwidth]{SapoFigures/Phos/SapoKaa_X1.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.12\textwidth]{SapoFigures/Phos/KaaPhos_X1.png}
+    \end{subfigure}
+
+     \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.15\textwidth]{SapoFigures/Phos/SapoPhos_X2.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.12\textwidth]{SapoFigures/Phos/KaaPhos_X2.png}
+    \end{subfigure}
+
+    \hspace{-12ex}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.15\textwidth]{SapoFigures/Phos/SapoPhos_X3.jpg}
+    \end{subfigure}
+    \begin{subfigure}{0.6\textwidth}
+    \centering
+    \includegraphics[width=1.12\textwidth]{SapoFigures/Phos/KaaPhos_X3.png}
+    \end{subfigure}
+
+    \caption{Figure depicting the reachable set computation of the Phosporley model.} 
+    \label{fig3}
+\end{figure}
 

### 2020-07-25 KAA-Arch/samplepaper.tex
Final version of Kaa paper

@@
 %
-We believe that our tool has pedagogical use given the simplicity of the implementation and its user-friendliness.
+We believe that our tool has pedagogical value given the simplicity of the implementation and its user-friendliness.
 %
@@
 
-\keywords{Nonlinear Dynamical Systems; \and Reachable Set Computation; \and Bernstein Polynomials.}
+% \begin{keywords}
+% Nonlinear Dynamical Systems; Reachable Set Computation; Bernstein Polynomials.
+% \end{keywords}
+%\keywords{Nonlinear Dynamical Systems; Reachable Set Computation; Bernstein Polynomials.}
 
@@
 
-Reachable set computation is one of the many important tools available for verification of dynamical and hybrid systems.
+Reachable set computation is one of the 
+%many 
+important tools %available 
+for verification of safety properties of dynamical and hybrid systems.
 %
-One of the simpler and easier-to-understand reachable set computation algorithms utilizes Bernstein polynomials and parallelotopes~\cite{dang2009image}.
+A simpler and easier-to-understand reachable set computation algorithm that utilizes Bernstein polynomials and parallelotopes has been presented in~\cite{dang2009image}.
 %
-Several cumulative improvements~\cite{dang2012reachability,sassi2012reachability,dang2014parameter,dreossi2016parallelotope,dreossi2017reachability} have been made to improve the accuracy and efficiency of nonlinear systems reachability.
+Several cumulative improvements~\cite{dang2012reachability,sassi2012reachability,dang2014parameter,dreossi2016parallelotope,dreossi2017reachability} have been proposed to improve its accuracy and efficiency.
 %
-The tool SAPO~\cite{dreossi2017sapo} which implements these algorithms in C\texttt{++} is also available publically.
+The tool SAPO~\cite{dreossi2017sapo} which implements these algorithms in C\texttt{++} is available publicly.
 %
@@
 %
-Our primary motivation for reimplementing the algorithms in Python is two-fold.
+Our motivation for reimplementing the algorithms in Python is two-fold.
 %
-First, using some of the standard libraries for symbolic manipulation in Sympy, the algorithm for reachability can be implemented very concisely.
+First, using standard libraries for symbolic manipulation in Sympy, the algorithm for reachability can be implemented very concisely.
 %
-In fact, Kaa is roughly only $650$ lines of code.
+In fact, the main algorithm for computing the reachable set has be implemented in only one file with $150$ lines of code.
 %
-Additionally, the main algorithm for computing the reachable set is provided in just one file with $150$ lines of code.
+The total code base for Kaa is roughly only $650$ lines of code.
+%
+%Additionally, the main algorithm for computing the reachable set is provided in just one file with $150$ lines of code.
 %
@@
 %
-We believe that such a simple implementation has pedagogical value and can be used for teaching one of the main algorithms for reachability of nonlinear systems.
+We believe that such a simple implementation can be used for teaching one of the main algorithms for reachability of nonlinear systems.
 %
@@
 %
-In SAPO, for computing the reachable set, one has to recompile the code along with the model file to obtain the binary and execute the binary to compute reachable set.
+In SAPO, one has to recompile the code along with the model file to generate the binary and the user has to execute the binary for computing the reachable set.
 %
-Additionally, SAPO generates a separate MATLAB script for visualizing time or phase plots of the reachable set.
-This script must be fed separately into either MATLAB or Octave to illustrate the reachable set.
+Additionally, visualizing the reachable set using SAPO is a two step process.
+In the first step, SAPO generates a MATLAB script that contains projections of reachable set for visualizing time or phase plots of the reachable set.
+In the second step, the user feeds the generated script into either MATLAB or Octave to illustrate the reachable set.
 %
-In contrast, we provide a Jupyter notebook interface for Kaa to create an intuitive platform facilitating experimentation.
-Juypter notebooks are simple to create and known for their straightforward user interface.
-By leveraging the matplotlib library for visualizing the reachable set, we were able to design an engaging interactive tutorial and experimentation platform for visualzing reachable sets of non-linear systems.
+In contrast, Kaa has very intuitive interface for computing and visualizing reachable set.
+%
+The models and the code for computing reachable set using Kaa are provided in Python and do not read any compilation. 
+%
+We also leverage the matplotlib library for visualizing the reachable set. 
+%
+Integrating these two, we provide an easy to use Jupyter notebook interface for computing and visualizing reachable set with Kaa for some of the standard benchmark models.
+%
+%Juypter notebooks are simple to create and known for their straightforward user interface. %
+%we were able to design an engaging interactive tutorial and experimentation platform for visualzing reachable sets of non-linear systems.
 %
@@
 
-Bernstein polynomials are also considered an active area of research in the domain of global optimization~\cite{nataray2002algorithm,garloff2003bernstein,nataraj2007new}.
+Bernstein polynomials are also an active area of research in the domain of global optimization~\cite{nataray2002algorithm,garloff2003bernstein,nataraj2007new}.
 %
-There have been several algorithms proposed for performing heuristics which improve the performance of optimization using Bernstein polynomials.
-Furthermore, these algorithms also improve the accuracy using sequential refinement techniques~\cite{smith2009fast,munoz2013formalization}.
+Several heuristics have been proposed for improving the performance of optimization using Bernstein polynomials~\cite{kodiak,smith2009fast,munoz2013formalization}. 
 %
-Having an accessible and flexible tool would make it more conducive to implement these heuristics and determine specific heuristics helpful in the domain of reachability.
+%Furthermore, these algorithms also improve the accuracy using sequential refinement techniques~\cite{}.
 %
-In light of these features and advantages, we believe Kaa can be used as a easy first step for introducing reachability due to its simplicity of implementation and its usability.
+Having an accessible and flexible tool would make it more conducive to implement these heuristics and determine if they are helpful in the domain of reachability.
+%
+In light of these features and advantages, we believe Kaa can be used as a easy first step for introducing reachability in a pedagogical setting.
 
@@
 \end{equation}
-
 where $f(x) : \mathbb{R}^n \rightarrow \mathbb{R}^n$ is polynomial in $x$. 
@@
 $$
-$\Lambda$ is called the \emph{direction matrix} $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$.
-$c$ is called the \emph{offset}.
-For a parallelotope $P$, we denote the affine transformation that maps $[0,1]^{n}$ to $P$ as $T_{p}$.
+$\Lambda$ is called the \emph{direction matrix} $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$. $c$ is called the \emph{offset}. 
+%
+Alternatively, a parallelotope can also be represented in vertex-generator representation as $\langle v, g_1, \ldots, g_n\rangle$.
+Here $v \in \mathbb{R}^n$ is called vertex and $g_1, \ldots, g_n$, $g_i \in \mathbb{R}^n$, are called generators.
+The parallelotope is defined as
+$$
+P \triangleq \{ x ~|~ \exists \alpha_1, \ldots, \alpha_n, x = v + \alpha_1 g_1 + \ldots + \alpha_n g_n, 0 \leq \alpha_i \leq 1\}
+$$
+This representation is very similar to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}. 
+%
+Notice that for a parallelotope $P$, the vertex-generator representation also defines the affine transformation that maps $[0,1]^{n}$ to $P$. 
+%
+We denote this affine transformation as $T_{p}$. 
+%
 A parallelotope bundle $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$ where $Q = \cap_{i=1}^{m}P_i$.
@@
 %
-The primary advantage of the Bernstein representation of a polynomial $p(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $p(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis.
+The primary advantage of the Bernstein representation of a polynomial $h(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $h(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis.
 
-In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
-The upper and lower bounds of $p(x_1,\ldots,x_n)$ over $[0,1]^n$ are exactly bounded by the Bernstein coefficients:
+In other words, given a polynomial $h(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $h(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $h(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
+%
+The upper and lower bounds of $h(x_1,\ldots,x_n)$ over $[0,1]^n$ are bounded by the Bernstein coefficients:
 $$
@@
 
-Any parallelotope $P$ can be considered as a suitable affine transformation $T_{p}$ from $[0,1]^{n}$ to $P$. 
+As mentioned earlier, a parallelotope $P$ can also be represented as an affine transformation $T_{p}$ from $[0,1]^{n}$ to $P$. 
 %
@@
 %
-For the remainder of the document, we assume that by using functional composition and the Bernstein representation, we can easily compute the upper bound on supremum and lower bound on infimum of polynomial functions over parallelotopes.
+
+For the remainder of the document, we assume that by using functional composition and the Bernstein representation, we can compute the upper bound on supremum and lower bound on infimum of polynomial functions over parallelotopes.
 %
-
-Furthermore, we denote the procedures for calculating such upper and lower bounds for a polynomial $p$ over some parallelotope $S$ as $\mathsf{BernsteinUpper}(p,S)$ and $\mathsf{BernsteinLower}(p,S)$ respectively.
+We denote the procedures for calculating such upper and lower bounds for a polynomial $h$ over some parallelotope $P$ as $\mathsf{BernsteinUpper}(h,P)$ and $\mathsf{BernsteinLower}(h,P)$ respectively.
 
@@
 
-Given a set represented as a parallelotope bundle $Q = \{P_1, P_2, \ldots, P_m\}$ and a discrete dynamical system $x^{+} = f(x)$, we now present the method for computing an overapproximation of $f(Q)$ as a parallelotope bundle $Q' = \{P_1', P_2', \ldots, P_m'\}$.
+Given a set represented as a parallelotope bundle $Q = \{P_1, P_2, \ldots, P_m\}$ and a discrete dynamical system $x^{+} = f(x)$, we now present the method for computing an overapproximation of the image $f(Q)$ as a new parallelotope bundle $Q' = \{P_1', P_2', \ldots, P_m'\}$.
 %
-We exert that the direction matrix of $P_i'$ is same as $P_i$ and the computation is required only to compute the offsets. 
+We ensure that direction matrix of $P_i'$ is same as $P_i$ and the computation is required only to compute the offsets. 
 %
-Denote the $j^{th}$ offset of $P_i$ and $P_i'$ as $c_{j,i}$ and $c'_{j,i}$ respectively and the $j^{th}$ direction in $P_i$ (same as $P_i'$) as $\Lambda_{j,i}$.
+Let us denote the $j^{th}$ offset of $P_i$ and $P_i'$ as $c_{j,i}$ and $c'_{j,i}$ respectively and the $j^{th}$ direction in $P_i$ (same as $P_i'$) as $\Lambda_{j,i}$.
 %
@@
 %
-First, performing functional composition for computing uthe pper bound of a polynomial over a parallelotope.
+First, performing functional composition for computing the upper bound of a polynomial over a parallelotope.
 %
@@
 %
-The library of \emph{sympy} has powerful symbolic manipulation tools which allow us to comfortably perform many sensitive symbolic subsititutions into polynomials. 
+The library of \emph{sympy} has powerful symbolic manipulation tools which allow us to comfortably perform many sensitive symbolic function composition of polynomials. 
 %
@@
 Additionally, the \emph{matplotlib} library has easy plotting facilities that we integrate into our tool for visualizing the reachable set.
-In particular, \emph{matplotlib} facilitates the ability to visualize several reachable sets simultaneously.
+In particular, \emph{matplotlib} facilitates the ability to plot several reachable sets simultaneously.
 This will aid others in performing more complex bundle-transformation experiments in future.
@@
 \centering
-\begin{tabular}{|c c c c c|}
+\begin{tabular}{|c| c c c c|}
     \hline
@@
 
-\noindent It is evident from Table~\ref{tab:one} that while implementation in Python is very intuitive and concise, the current implementation incurs severe performance penalties. 
+\noindent It is evident from Table~\ref{tab:one} that while the current implementation in Python is very intuitive and concise, it incurs severe performance penalties. 
 %
@@
 \caption{Reachable Set Computation Time of Benchmarks}\label{tab:one}
-\begin{tabular}{|l | l l|}
+\begin{tabular}{|l | c c|}
 \hline
@@
 
-We present Kaa, a Python implementation of reachable set computation of nonlinear systems that is focussed towards accessibility and pedagocial use.
+We present Kaa, a Python implementation of reachable set computation of nonlinear systems which is focused towards accessibility and pedagogical use.
 %
-The usage of inbuilt \emph{sympy} libraries makes the implementation very short and simple (only 650 LOC).
+The usage of inbuilt \emph{sympy} libraries makes the implementation short and simple (only 650 LOC).
 %
@@
 The code can be accessed through \url{https://github.com/Tarheel-Formal-Methods/kaa}
+\newline
+\newline
+{\bf Acknowledgements: } This material is based upon work supported by the Air Force Office of Scientific Research under award number FA9550-19-1-0288 and National Science Foundation (NSF) under grant numbers CNS 1739936, 1935724.
+Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the United States Air Force or National Science Foundation.
+
+\bibliographystyle{splncs04}
+\bibliography{archbib}
+\newpage
 
 \begin{figure}[h]
-    \hspace{-15ex}
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-15ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
@@
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
@@
 
-    \hspace{-15ex}   
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-15ex}   
+    \begin{subfigure}{0.47\textwidth}
     \centering
@@
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.2\textwidth]{SapoFigures/SIR/KaaSIR_I.png}
+    \includegraphics[width=1.2\textwidth,height=0.8\textwidth]{SapoFigures/SIR/KaaSIR_I.png}
     \end{subfigure}
 
-    \hspace{-15ex}  
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-15ex}  
+    \begin{subfigure}{0.47\textwidth}
     \centering
@@
 \begin{figure}[h]
-    \hspace{-15ex}
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-15ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerX.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRossler_X.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.1\textwidth]{SapoFigures/Rosslert/KaaRosslerX.png}
+    \includegraphics[width=1.3\textwidth]{SapoFigures/Rosslert/KaaRosslerX.png}
     \end{subfigure}
 
-     \hspace{-15ex}  
-    \begin{subfigure}{0.6\textwidth}
+%     \hspace{-15ex}  
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerY.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRossler_Y.jpg}
     \end{subfigure}
-    \hspace{-3ex}
-    \begin{subfigure}{0.6\textwidth}
+    % \hspace{-3ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.4\textwidth,height=0.9\textwidth]{SapoFigures/Rosslert/KaaRosslerY.png}
+    \includegraphics[width=1.3\textwidth,height=0.8\textwidth]{SapoFigures/Rosslert/KaaRosslerY.png}
     \end{subfigure}
 
-    \hspace{-15ex}  
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-15ex}  
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRosslerZ.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Rosslert/SapoRossler_Z.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.25\textwidth]{SapoFigures/Rosslert/KaaRosslerZ.png}
+    \includegraphics[width=1.3\textwidth]{SapoFigures/Rosslert/KaaRosslerZ.png}
     \end{subfigure}
@@
     \caption{Figure depicting the reachable set computation of the Rossler model.} 
-    \label{fig3}
+    \label{fig2}
 \end{figure}
@@
 \begin{figure}[h]
-    \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-12ex}
+    \begin{subfigure}{0.45\textwidth}
     \centering
-    \includegraphics[width=1.2\textwidth]{SapoFigures/Quad/SapoQuad_H.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_H.jpg}
     \end{subfigure}
-    \hspace{5ex}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.13\textwidth]{SapoFigures/Quad/KaaQuad_H.jpg}
+    \includegraphics[width=1.1\textwidth,height=0.8\textwidth]{SapoFigures/Quad/KaaQuad_H.jpg}
     \end{subfigure}
 
-     \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%     \hspace{-12ex}
+    \begin{subfigure}{0.45\textwidth}
     \centering
-    \includegraphics[width=1.2\textwidth]{SapoFigures/Quad/SapoQuad_W.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_W.jpg}
     \end{subfigure}
-    \hspace{5ex}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.13\textwidth]{SapoFigures/Quad/KaaQuad_W.png}
+    \includegraphics[width=1.1\textwidth,height=0.8\textwidth]{SapoFigures/Quad/KaaQuad_W.png}
     \end{subfigure}
 
-    \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-12ex}
+    \begin{subfigure}{0.45\textwidth}
     \centering
-    \includegraphics[width=1.15\textwidth]{SapoFigures/Quad/SapoQuad_HI.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Quad/SapoQuad_HI.jpg}
     \end{subfigure}
-    \hspace{5ex}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.13\textwidth]{SapoFigures/Quad/KaaQuad_HI.png}
+    \includegraphics[width=1.1\textwidth,height=0.85\textwidth]{SapoFigures/Quad/KaaQuad_HI.png}
     \end{subfigure}
@@
 \begin{figure}[h]
-    \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-12ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.2\textwidth]{SapoFigures/LV/SapoLV_X1.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X1.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.14\textwidth]{SapoFigures/LV/KaaLV_X1.png}
+    \includegraphics[width=1.1\textwidth,height=0.85\textwidth]{SapoFigures/LV/KaaLV_X1.png}
     \end{subfigure}
 
-     \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%     \hspace{-12ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.2\textwidth]{SapoFigures/LV/SapoLV_X2.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X2.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.14\textwidth]{SapoFigures/LV/KaaLV_X2.png}
+    \includegraphics[width=1.1\textwidth,height=0.82\textwidth]{SapoFigures/LV/KaaLV_X2.png}
     \end{subfigure}
 
-    \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-12ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.15\textwidth]{SapoFigures/LV/SapoLV_X3.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/LV/SapoLV_X3.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.14\textwidth]{SapoFigures/LV/KaaLV_X3.png}
+    \includegraphics[width=1.1\textwidth,height=0.82\textwidth]{SapoFigures/LV/KaaLV_X3.png}
     \end{subfigure}
@@
     \caption{Figure depicting the reachable set computation of the Lotka-Volterra model.} 
-    \label{fig3}
+    \label{fig4}
 \end{figure}
@@
 \begin{figure}[h]
-    \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%    \hspace{-12ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.15\textwidth]{SapoFigures/Phos/SapoKaa_X1.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoKaa_X1.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.12\textwidth]{SapoFigures/Phos/KaaPhos_X1.png}
+    \includegraphics[width=1.1\textwidth,height=0.82\textwidth]{SapoFigures/Phos/KaaPhos_X1.png}
     \end{subfigure}
 
-     \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%   \hspace{-12ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.15\textwidth]{SapoFigures/Phos/SapoPhos_X2.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X2.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.12\textwidth]{SapoFigures/Phos/KaaPhos_X2.png}
+    \includegraphics[width=1.1\textwidth,height=0.82\textwidth]{SapoFigures/Phos/KaaPhos_X2.png}
     \end{subfigure}
 
-    \hspace{-12ex}
-    \begin{subfigure}{0.6\textwidth}
+%   \hspace{-12ex}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.15\textwidth]{SapoFigures/Phos/SapoPhos_X3.jpg}
+    \includegraphics[width=\textwidth]{SapoFigures/Phos/SapoPhos_X3.jpg}
     \end{subfigure}
-    \begin{subfigure}{0.6\textwidth}
+    \begin{subfigure}{0.47\textwidth}
     \centering
-    \includegraphics[width=1.12\textwidth]{SapoFigures/Phos/KaaPhos_X3.png}
+    \includegraphics[width=1.1\textwidth,height=0.82\textwidth]{SapoFigures/Phos/KaaPhos_X3.png}
     \end{subfigure}
 
-    \caption{Figure depicting the reachable set computation of the Phosporley model.} 
-    \label{fig3}
+    \caption{Figure depicting the reachable set computation of the Phosporaley model.} 
+    \label{fig5}
 \end{figure}
@@
 %
-\bibliographystyle{splncs04}
-\bibliography{archbib}
+
 %

### 2020-11-08 CS590/project/paper.tex
First commit of Circuit Complexity paper for Spring2020

\documentclass{amsart}
\usepackage{complexity}

\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-2in}
\calclayout

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{xca}[theorem]{Exercise}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\numberwithin{equation}{section}

\begin{document}

\title{On the Fourier Analysis of Boolean Functions and the Linai-Mansour-Nisan Theorem}

\author{Edward Kim}
\email{ehkim@cs.unc.edu}

\begin{abstract}
  We briefly summarize the basic tools of Fourier Analysis applied to the finite abelian group $\mathbb{Z}_2^n$. By viewing boolean functions $f:\{0,1\}^n \rightarrow \{0,1\}$ as square-integrable maps in the Hilbert space $L^2(\mathbb{Z}_2^n)$ under the uniform probability measure, we derive the Linai-Mansour-Nisan theorem concerning the tail of the Fourier spectrum of boolean functions in $\AC^0$. Finally, we highlight some applications of the LMN theorem in the context of sensitivity and influence of functions in $\AC^0$.
\end{abstract}

\maketitle

\section{Fourier Analysis of Boolean Functions}

\subsection{Characters}

We introduce some stepping stones defining the tools used in the Fourier Analysis of boolean functions. Many of these tools can be defined using techniques from Representation Theory, but for the sake of briefity, we will not delve into those concepts here in detail. Instead, we will provide remarks and snippets on relevant topics as needed along with references to the standard literature at the end of the section.  \newline

We first define some general concepts for any finite abelian group $G$. A group homomorphism from $\chi: G \rightarrow \mathbb{C}^*$ is called a \emph{character} of $G$ where $\mathbb{C}^* = \mathbb{C}/\{0\}$ is the multiplicative group of the complex numbers. We call the homomorphism $\chi_0:G \rightarrow \mathbb{C}^*, \; \chi_0 = 1$ the \emph{trivial character} of $G$. By definition, for $a,b \in G$,
\begin{equation} \label{homo}
\chi(a + b) = \chi(a)\chi(b)
\end{equation}
for all characters $\chi$. \newline

Let us now consider the simplest finite abelian group, $\mathbb{Z}_n$ for some $n \in \mathbb{N}$. It follows from basic properties of cyclic groups that the only characters of $\mathbb{Z}_n$ are:
%
\begin{equation}
  \chi_j(x) = e^{ 2\pi i jx/n} \quad j \in [n], \; x \in \mathbb{Z}_n
\end{equation}
%
We can decompose any finite abelian group $G$ into a direct product of finite cyclic groups $G \cong \mathbb{Z}_{n_1} \times \mathbb{Z}_{n_2} \cdots \times\mathbb{Z}_{n_k}$. If we decompose $x = \sum_{i \in [k]} x_i$ for $x \in G$, then the characters of $G$ are defined as
%
\begin{equation}
  \chi_a(x) = \prod_{i \in [k]} e^{2 \pi i a_k x_k/ n_k}
\end{equation}
%
We see that every element $a \in G$ has an unique associated character $\chi_a$. Furthermore
%
\begin{gather} \label{chargroup}
  \chi_{a+b}(x)  = \chi_{a}(x)\chi_{b}(x) \\
  \chi_{a^{-1}}(x) = \frac{1}{\chi_a(x)}
\end{gather}
%
This shows that the set whose elements are the characters of $G$ endowed with pointwise multiplication is an abelian group. Denote this set as $\hat{G}$. From the map $a \mapsto \chi_a$, we have a group isomorphism between $G \cong \hat{G}$. The group $\hat{G}$ is known as the \emph{character group} of $G$. \newline

To define the Fourier transform, we must consider the Hilbert space $L_2(G)$ where $G$ endowed with the uniform probability measurel i.e the discrete measure mapping each subset $H \subseteq G$ to $\frac{|H|}{|G|}$. This yields the standard inner product for maps $f,g \in L_2(G)$:
%
\begin{equation}
  \langle f,g \rangle = \frac{1}{|G|}\sum_{x\in G} f(x)\overline{g(x)}= \mathbb{E}_x f(x)\overline{g(x)}
\end{equation}
Note that this induces the norm $||f||_2 = \sqrt{\langle f, f \rangle} = \mathbb{E}_x |f(x)|^2$ on $L_2(G)$. It is simple to check that the characters of $G$ indeed lie in $L_2(G)$. However, we can prove something stronger: the characters of $G$, $\chi_a$ form an \emph{orthonormal basis} of $L_2(G)$.
%

\begin{theorem}
  The characters $\chi_a \in \hat{G}$ form an orthonormal basis for $L_2(G)$
\end{theorem}

\begin{proof}
  To show this, we must employ a lemma:

\end{proof}

We restrict our attention to the finite abelian group $\mathbb{Z}_2^n$ as a natural group to define boolean functions $f: \{0,1\}^n \rightarrow \{0,1\}$. By the observations above, we conclude that our characters for $\mathbb{Z}_2^n$ are defined as:

\begin{equation}
  \chi_a(x) = (-1)^{\sum_{i \in [n]} x_ia_i} = (-1)^{\sum_{i \in [n], \; a_i = 1} x_i}
\end{equation}
where $x_i$ is the $i^{th}$ bit of $x \in \{0,1\}^n$.

%\begin{remark}
%The above properties on characters can be shown through %techniques in Representation Theory. For instance, we %know representations of finite groups are group %homomorphisms into the group of linear isomorphisms of a %suitable complex vector space $\rho: G \rightarrow %GL(\mathbb{C}^n)$. The \emph{characters} of $\rho$ are %defined as $\chi(x) = Tr(\rho(x))$ where $Tr$ is the %trace operator. It turns out that the representations of %finite abelian groups are one-dimensional i.e isomorphic %to $\mathbb{C}$. Thus, $\rho(x)$ will just be multiplying %$\mathbb{C}$ by some scalar $\lambda_x \in \mathbb{C}^*$. %In this case, we can show that indeed for any $a,b \in G$ %and $\chi$ any character of $G$:
%$$ \chi(a+b) = Tr(\rho(a+b)) = Tr(\rho(a)\circ\rho(b)) = %\lambda_a \lambda_b = Tr(\rho(a))Tr(\rho(b)) = %\chi(a)\chi(b)$$ Note that this relation does not hold %for arbitrary finite groups. For more information on the %Representation Theory of Finite Groups, see \ref{serre}.
%\end{remark}

%\section{Linai-Mansour-Nisan Theorem}

%\section{Immediate Applications}

%    Bibliographies can be prepared with BibTeX using amsplain,
%    amsalpha, or (for "historical" overviews) natbib style.
\bibliographystyle{amsplain}
%    Insert the bibliography data here.

\end{document}

### 2020-11-09 CS590/project/paper.tex
Added FT notes to CS590

@@
 \numberwithin{equation}{section}
+
+\theoremstyle{remark}
+\newtheorem{notation}[theorem]{Notation}
 
@@
 
-Let us now consider the simplest finite abelian group, $\mathbb{Z}_n$ for some $n \in \mathbb{N}$. It follows from basic properties of cyclic groups that the only characters of $\mathbb{Z}_n$ are:
+Consider the set whose elements are the characters of $G$ and whose binary operation is pointwise multiplication of complex-valued functions on $G$.
+This set is also an \emph{abelian group} by the following theorem:
+%
+\begin{theorem}
+  If $G$ is a finite abelian group, then the characters of $G$ form an abelian group under pointwise multiplication.
+\end{theorem}
+%
+\begin{proof}
+  We first show that if $\chi, \phi$ are characters of $G$, then $\chi\phi$ is also a character of $G$. But this immediately follows from \ref{homo}:
+  $$ \chi\phi(a+b) = \chi(a)\chi(b)\phi(a)\phi(b) = [\chi(a)\phi(a)][\chi(b)\phi(b)] = \chi\phi(a)\chi\phi(b)$$
+  Hence, $\chi\phi$ is also a character of $G$. Furthermore, $\chi^{-1} = \frac{1}{\chi} = \overline{\chi}$ is also a character of $G$ since
+  $$ \chi^{-1}(a+b) = \frac{1}{\chi(a+b)} = \frac{1}{\chi(a)\chi(b)} = \frac{1}{\chi(a)}\frac{1}{\chi(b)} = \chi^{-1}(a)\chi^{-1}(b)$$
+  showing that the set has inverses.
+\end{proof}
+%
+\begin{remark}
+  We abuse notation a bit and just note that $\chi^{-1}$ is not the inverse of the $\chi$ as a \emph{map}. Rather, $\chi^{-1}$ will refer to the \emph{group inverse} of character $\chi$ from hereon.
+\end{remark}
+
+%
+Denote this set as $\hat{G}$.
+The group $\hat{G}$ is known as the \emph{character group} of $G$.
+It also will turn out that there is a bijective map between $G$ and $\hat{G}$ such that the map is also a \emph{group isomorphism}.
+We will prove this later on.
+\newline
+
+Let us now consider the simplest finite abelian group, $\mathbb{Z}_n = \mathbb{Z}/n\mathbb{Z}$ for some $n \in \mathbb{N}$.
+It follows from basic properties of cyclic groups that the only characters of $\mathbb{Z}_n$ are:
 %
@@
 \end{equation}
-%
+where $x_i$ refers to the $i^{th}$ index of $x$. \newline
+
 We see that every element $a \in G$ has an unique associated character $\chi_a$. Furthermore
@@
 \begin{gather} \label{chargroup}
-  \chi_{a+b}(x)  = \chi_{a}(x)\chi_{b}(x) \\
-  \chi_{a^{-1}}(x) = \frac{1}{\chi_a(x)}
-\end{gather}
-%
-This shows that the set whose elements are the characters of $G$ endowed with pointwise multiplication is an abelian group.
-Denote this set as $\hat{G}$.
-From the map $a \mapsto \chi_a$, we have a group isomorphism between $G \cong \hat{G}$.
-The group $\hat{G}$ is known as the \emph{character group} of $G$.
-\newline
+  \chi_{a+b}(x)  = \chi_{a}(x)\chi_{b}(x)
+\end{gather} \newline
 
-To define the Fourier transform, we must consider the Hilbert space $L_2(G)$ where $G$ endowed with the uniform probability measurel i.e the discrete measure mapping each subset $H \subseteq G$ to $\frac{|H|}{|G|}$.
+To define the Fourier transform, we must consider the Hilbert space $L_2(G)$ where $G$ endowed with the uniform probability measure i.e the discrete measure mapping each subset $H \subseteq G$ to $\frac{|H|}{|G|}$.
 This yields the standard inner product for maps $f,g \in L_2(G)$:
@@
 \end{equation}
-Note that this induces the norm $||f||_2 = \sqrt{\langle f, f \rangle} = \mathbb{E}_x |f(x)|^2$ on $L_2(G)$.
+Note that this induces the norm $||f||_2^2 = \langle f, f \rangle = \mathbb{E}_x |f(x)|^2$ on $L_2(G)$.
 It is simple to check that the characters of $G$ indeed lie in $L_2(G)$.
@@
 %
-
-\begin{theorem}
-  The characters $\chi_a \in \hat{G}$ form an orthonormal basis for $L_2(G)$
-\end{theorem}
-
+First, we prove a lemma:
+\begin{lemma} \label{totalsum}
+  Let $\chi$ be a non-trivial character of $G$. Then $\sum_{x \in G} \chi(x) = 0$
+\end{lemma}
+%
 \begin{proof}
-  To show this, we must employ a lemma:
-
+  Given any non-trivial character $\chi$, the following holds for all $y \in G$:
+    %
+    \[ \chi(y)\sum_{x \in G} \chi(x) = \sum_{x \in G} \chi(y+x) = \sum_{x \in G} \chi(x) \]
+    %
+    If $\sum_{x \in G} \chi(x) \neq 0$, then this would force $\chi(y) = 1$ for all $y \in G$. However, this would contradict the non-triviality of the $\chi$.
 \end{proof}
 
-We restrict our attention to the finite abelian group $\mathbb{Z}_2^n$ as a natural group to define boolean functions $f: \{0,1\}^n \rightarrow \{0,1\}$.
+\begin{theorem} \label{orthonormal}
+  The characters $\chi \in \hat{G}$ form an orthonormal basis for $L_2(G)$
+\end{theorem}
+%
+\begin{proof}
+We first note that $\langle \chi, \chi \rangle = \mathbb{E}_x|\chi(x)|^2 = 1$ since the image of character $\chi$ lies on the complex unit circle by the discussions above.
+Now given two non-trivial characters $\chi, \phi$
+$$ \langle \chi, \phi \rangle = \mathbb{E}_x \chi(x)\overline{\phi(x)} = \frac{1}{|G|} \sum_{x \in G} \chi\phi^{-1}(x)  = 0$$
+The last equality follows from Lemma \ref{totalsum} since $\chi\phi^{-1}$ must also be a non-trivial character if $\chi, \phi$ are non-trivial characters.
+The case where one of the character is trivial also follows from the lemma.
+Finally, observe that $L_2(G)$ is $|G|$-dimensional as a vector space by considering the dual of $G$.
+We know that there are at least $|G|$-characters, showing that the characters span $L_2(G)$.
+However, from the lemma, the characters must also be pairwise orthgonal to each other.
+This yields that the characters of $G$ form an \emph{orthonormal basis} of $L_2(G)$ and that there must be exactly $|G|$ characters in $\hat{G}$.
+Finally, by using \ref{chargroup}, it follows that the map $a \mapsto \chi_a$ must be a group isomorphism, so $G \cong \hat{G}$.
+\end{proof}
+
+To finish, we restrict our attention to the finite abelian group $\mathbb{Z}_2^n$ as a natural group to define boolean functions $f: \{0,1\}^n \rightarrow \{0,1\}$.
 By the observations above, we conclude that our characters for $\mathbb{Z}_2^n$ are defined as:
-
+%
 \begin{equation}
@@
 where $x_i$ is the $i^{th}$ bit of $x \in \{0,1\}^n$.
+%
+\begin{notation}
+  Instead of considering $x \in \{0,1\}^n$ as a bit-string, we would like to use the following notation: Let $S \subseteq [n]$ be such that $S_x = \{i \mid x_i = 1\}$.
+As there is a bijective correspondence between all such subsets $S_x$ and $x \in \{0,1\}^n$, we will start to identify bit strings with their subset counterparts for notational convenience.
+\end{notation}
+
+\subsection{The Fourier Transform}
+
+We have shown above that the members of the character group $\hat{G}$ form an orthonormal basis of $L_2(G)$.
+This bestows us the power of decomposing any $f \in L_2(G)$ into its orthogonal counterparts.
+The restriction to the case where $G = \mathbb{Z}_2^n$ is what motivates us as any $n$-ary boolean function $f: \{0,1\}^n \rightarrow \{0,1\}$ must lie in $L_2(\mathbb{Z}_2^n)$.
+As before, we begin by considering the general case for any finite abelian group $G$  \newline
+
+Let $\hat{f}: \hat{G} \rightarrow \mathbb{C}$ be the complex-valued function defined as:
+%
+\begin{equation}
+  \hat{f}(\chi_a) =  \langle f, \chi_a \rangle, \; a \in G
+\end{equation}
+%
+We can iterate this for all characters to get a direct sum decomposition for any $f \in L_2(G)$
+%
+\begin{equation}
+  f = \sum_{a \in G} \hat{f}(\chi_a)\chi_a
+\end{equation}
+
+It can be show that the map $f \mapsto \hat{f}$ is linear by the left-linearity of the inner product. This linear map is the \emph{Fourier Transform on $G$}.
+%
+\begin{notation}
+  We will sometimes denote $\hat{f}(a) = \hat{f}(\chi_a)$.
+\end{notation}
+
+A simple application of the Fourier Transform yields an useful identity which will be instrumental in our formulation of the Fourier tail of a boolean function.
+
+\begin{theorem}{(Parseval's Identity)}
+  Let $f \in L_2(G)$. Then
+  $$ ||f||_2^2 = \sum_{a \in G} |\hat{f}(a)|^2$$
+\end{theorem}
+%
+\begin{proof}
+  $$ ||f||^2_2 = \langle f, f \rangle = \langle \sum_{a \in G} \hat{f}(a) \chi_a,  \sum_{b \in G} \hat{f}(b) \chi_b \rangle = \sum_{a,b \in G} \hat{f}(a)\overline{\hat{f}(b)}\langle \chi_a, \chi_b \rangle = \sum_{a \in G} |\hat{f}(a)|^2 $$
+  The last equality is just the orthonormality of the characters (\ref{orthonormal}).
+\end{proof}
+
+%by the remark in the previous section, we will identify characters of %$\mathbb{Z}_2^n$ as $\chi_A$ for $A \subseteq [n]$.
 

### 2020-11-10 CS590/paper.tex
Fourier section finished for cs590.

--- 2020-11-09 CS590/project/paper.tex
@@
   \chi_{a+b}(x)  = \chi_{a}(x)\chi_{b}(x)
-\end{gather} \newline
+\end{gather}
 
@@
 \end{equation}
+%
 Note that this induces the norm $||f||_2^2 = \langle f, f \rangle = \mathbb{E}_x |f(x)|^2$ on $L_2(G)$.
@@
 \begin{proof}
-  Given any non-trivial character $\chi$, the following holds for all $y \in G$:
+  Given any non-trivial character $\chi$, the following holds for all $y \in G$: \newline
     %
-    \[ \chi(y)\sum_{x \in G} \chi(x) = \sum_{x \in G} \chi(y+x) = \sum_{x \in G} \chi(x) \]
+    \[ \chi(y)\sum_{x \in G} \chi(x) = \sum_{x \in G} \chi(y+x) = \sum_{x \in G} \chi(x) \]\newline
     %
@@
   Instead of considering $x \in \{0,1\}^n$ as a bit-string, we would like to use the following notation: Let $S \subseteq [n]$ be such that $S_x = \{i \mid x_i = 1\}$.
-As there is a bijective correspondence between all such subsets $S_x$ and $x \in \{0,1\}^n$, we will start to identify bit strings with their subset counterparts for notational convenience.
+As there is a bijective correspondence between all such subsets $S_x$ and $x \in \{0,1\}^n$, we will sometimes identify bit strings with their subset counterparts for notational convenience.
 \end{notation}
@@
 %
-We can iterate this for all characters to get a direct sum decomposition for any $f \in L_2(G)$
+We can iterate this for all characters to get a direct sum decomposition for any $f \in L_2(G)$ which we call the \emph{Fourier Inversion Formula}\newline.
 %
@@
 
-It can be show that the map $f \mapsto \hat{f}$ is linear by the left-linearity of the inner product. This linear map is the \emph{Fourier Transform on $G$}.
+It can be show that the map $f \mapsto \hat{f}$ is linear by the left-linearity of the inner product.
+This linear map is the \emph{Fourier Transform on $G$} and (\ref{orthonormal}) shows that the decomposition is unique.
+Thus the Fourier Transform is a linear isomorphism as $\hat{}:L_2(G) \rightarrow L_2(\hat{G})$.
+The complex value $\hat{f}(\chi_a)$ is called the \emph{Fourier coefficient} associated to $\chi_a$.
 %
 \begin{notation}
-  We will sometimes denote $\hat{f}(a) = \hat{f}(\chi_a)$.
+  We will sometimes denote $\hat{f}(a)$ for $\hat{f}(\chi_a)$.
 \end{notation}
+%
+From the above definition, we directly calculate that
+\begin{equation}
+  \hat{f}(0) = \langle f, \chi_0 \rangle = \mathbb{E}_x f(x)
+\end{equation}
 
@@
 \end{proof}
+%
+By combining Parseval's Identity with the observation that $||\hat{f}||_2^2 = \langle \hat{f}, \hat{f} \rangle = \frac{1}{|G|}\sum_{a \in G} |\hat{f}(a)|^2$, we arrive at the \emph{Plancheral Formula}
 
-%by the remark in the previous section, we will identify characters of %$\mathbb{Z}_2^n$ as $\chi_A$ for $A \subseteq [n]$.
+\begin{theorem}{(Plancheral Formula)}
+  $$||\hat{f}||_2 = \sqrt{G}||f||_2$$
+\end{theorem}
+%
+\begin{example}{\emph{(Fourier coefficients for Parity)}}
+  Let $G = \mathbb{Z}_2^n$ and $f = \parity_n$, the parity function on $n$ bits.
+  We wish to calculate the Fourier coefficients of $f$. We begin by calculating $\hat{f}(0^n)$ and $\hat{f}(1^n)$
+  \begin{align}
+    & \hat{f}(0^n) = \frac{1}{2^n}\sum_{x \in G} f(x) = \frac{2^{n-1}}{2^n} = \frac{1}{2} \\
+    & \hat{f}(1^n) = \frac{1}{2^n}\sum_{x \in G} (-1)^{|x|}f(x) = \frac{-2^{n-1}}{2^n} = -\frac{1}{2}
+  \end{align}
+  Now for $a \neq 0^n,1^n$, there must exist two indices $i,j$ such that $a_i = 0$ and $a_j = 1$.
+Define $e_i, e_j$ to be the $n$-bit strings such that all indices are set to zero except for the $i^{th},j^{th}$ places respectively.
+Note that:
+  $$ \hat{f}(a) = \frac{1}{|G|} \sum_{x \in G} f(x)\overline{\chi_a(x)} = \frac{1}{|G|}\sum_{x \in G} f(x + e_0 + e_1) \chi_a(x + e_0 + e_1)$$
+  since translating by $e_0 + e_1$ rearranges the lefthand sum. This yields the equality:
+  $$ \hat{f}(a) = \frac{1}{2}\mathbb{E}_x[f(x)\chi_a(x) + f(x + e_0 + e_1) \chi_a(x + e_0 + e_1)] $$
+  As flipping bits at indices $i,j$ does not change the parity for any $x \in \{0,1\}^n$, $f(x) = f(x + e_0 + e_1)$.
+If $x_j = 1$, then adding $e_j$ will flip the bit to $x_j = 0$.
+Recalling the definition of the character $\chi_a(x) = (-1)^{\sum_{i \in [n], a_i = 1}}$ shows that a factor of $-1$ disappears from the product since $a_j = 1$.
+Similarly, if initially $x_j = 0$, then the product gains a factor of $-1$.
+This proves that $\chi_a(x + e_0 + e_1) = - \chi_a(x)$ for all $x \in \{0,1\}^n$ which in turn proves that $\hat{f}(a) = 0$ for all such $a \neq 0^n,1^n$.
+By Parseval's identity:
+  $$||\parity_n||_2 = \frac{1}{\sqrt{2}} $$
+\end{example}
+%
+%\begin{example}{\emph{(Fourier coefficients for Majority)}}
+%  We proceed through a similar analysis for the parity case.
+%  \begin{align*}
+%    & \hat{f}(0) = \mathbb{E}_x f(x) = \frac{{n \choose \frac{n}{2}} + {n \choose \frac{n}{2} + 1} + ... + {n \choose n}}{2^n} = \frac{2^{n-1}}{2^n} = \frac{1}{2} \\
+%    & \hat{f}(1) =
+%  \end{align*}
+%\end{example}
 
-%\begin{remark}
-%The above properties on characters can be shown through %techniques in Representation Theory.
-For instance, we %know representations of finite groups are group %homomorphisms into the group of linear isomorphisms of a %suitable complex vector space $\rho: G \rightarrow %GL(\mathbb{C}^n)$.
-The \emph{characters} of $\rho$ are %defined as $\chi(x) = Tr(\rho(x))$ where $Tr$ is the %trace operator.
-It turns out that the representations of %finite abelian groups are one-dimensional i.e isomorphic %to $\mathbb{C}$.
-Thus, $\rho(x)$ will just be multiplying %$\mathbb{C}$ by some scalar $\lambda_x \in \mathbb{C}^*$.
-%In this case, we can show that indeed for any $a,b \in G$ %and $\chi$ any character of $G$:
-%$$ \chi(a+b) = Tr(\rho(a+b)) = Tr(\rho(a)\circ\rho(b)) = %\lambda_a \lambda_b = Tr(\rho(a))Tr(\rho(b)) = %\chi(a)\chi(b)$$ Note that this relation does not hold %for arbitrary finite groups.
-For more information on the %Representation Theory of Finite Groups, see \ref{serre}.
-%\end{remark}
+\begin{remark}{\emph{(Generalizations of Fourier Analysis)}}
+The above properties on characters can be shown through techniques in Representation Theory.
+For instance, we know representations of finite groups are group homomorphisms into the group of linear automorphisms of a suitable complex vector space $\rho: G \rightarrow GL(\mathbb{C}^n)$.
+The \emph{characters} of $\rho$ are defined as $\chi(x) = Tr(\rho(x))$ where $Tr$ is the trace operator.
+It turns out that the representations of finite abelian groups are one-dimensional i.e isomorphic to $\mathbb{C}$.
+Thus, $\rho(x)$ will just be multiplying $\mathbb{C}$ by some scalar $\lambda_x \in \mathbb{C}^*$.
+In this case, we can show that indeed for any $a,b \in G$ and $\chi$ any character of $G$:
+$$ \chi(a+b) = Tr(\rho(a+b)) = Tr(\rho(a)\circ\rho(b)) = \lambda_a \lambda_b = Tr(\rho(a))Tr(\rho(b)) = \chi(a)\chi(b)$$ Note that this relation does not hold for arbitrary finite groups.
+For more information on the Representation Theory of Finite Groups, see \ref{serre}.
+\end{remark}
 

### 2020-11-11 CS590/paper.tex
Finished LMN section

@@
 \newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
 \newtheorem{lemma}[theorem]{Lemma}
@@
 
-\title{On the Fourier Analysis of Boolean Functions and the Linai-Mansour-Nisan Theorem}
+\title{On the Fourier Analysis of Boolean Functions and the Linial-Mansour-Nisan Theorem}
 
@@
 Instead, we will provide remarks and snippets on relevant topics as needed along with references to the standard literature at the end of the section.
+Our presentation follows that of \cite{hatami}.
 \newline
 
-We first define some general concepts for any finite abelian group $G$.
+Let us first define some general concepts for any finite abelian group $G$.
 A group homomorphism from $\chi: G \rightarrow \mathbb{C}^*$ is called a \emph{character} of $G$ where $\mathbb{C}^* = \mathbb{C}/\{0\}$ is the multiplicative group of the complex numbers.
@@
   $$ \chi^{-1}(a+b) = \frac{1}{\chi(a+b)} = \frac{1}{\chi(a)\chi(b)} = \frac{1}{\chi(a)}\frac{1}{\chi(b)} = \chi^{-1}(a)\chi^{-1}(b)$$
-  showing that the set has inverses.
+  showing that the set contains inverses.
 \end{proof}
@@
 To define the Fourier transform, we must consider the Hilbert space $L_2(G)$ where $G$ endowed with the uniform probability measure i.e the discrete measure mapping each subset $H \subseteq G$ to $\frac{|H|}{|G|}$.
-This yields the standard inner product for maps $f,g \in L_2(G)$:
+This yields the standard inner product for maps $f,g \in L_2(G)$: \newline
 %
@@
 %
-Note that this induces the norm $||f||_2^2 = \langle f, f \rangle = \mathbb{E}_x |f(x)|^2$ on $L_2(G)$.
+This induces the norm $||f||_2^2 = \langle f, f \rangle = \mathbb{E}_x |f(x)|^2$ on $L_2(G)$.
 It is simple to check that the characters of $G$ indeed lie in $L_2(G)$.
@@
 As there is a bijective correspondence between all such subsets $S_x$ and $x \in \{0,1\}^n$, we will sometimes identify bit strings with their subset counterparts for notational convenience.
+This changes the notation for characters of $\mathbb{Z}_2^n$ as below
+  $$ \chi_A(x) = (-1)^{\sum_{i \in A} x_i}, \quad A \subseteq [n]$$
 \end{notation}
@@
 
-It can be show that the map $f \mapsto \hat{f}$ is linear by the left-linearity of the inner product.
+It can be shown that the map $f \mapsto \hat{f}$ is linear by the left-linearity of the inner product.
 This linear map is the \emph{Fourier Transform on $G$} and (\ref{orthonormal}) shows that the decomposition is unique.
-Thus the Fourier Transform is a linear isomorphism as $\hat{}:L_2(G) \rightarrow L_2(\hat{G})$.
+Hence the Fourier Transform can be viewed as a linear isomorphism $\hat{}:L_2(G) \rightarrow L_2(\hat{G})$.
 The complex value $\hat{f}(\chi_a)$ is called the \emph{Fourier coefficient} associated to $\chi_a$.
@@
 \begin{theorem}{(Plancheral Formula)}
-  $$||\hat{f}||_2 = \sqrt{G}||f||_2$$
+  $$||\hat{f}||_2 = \sqrt{|G|}\cdot||f||_2$$
 \end{theorem}
@@
 %
-%\begin{example}{\emph{(Fourier coefficients for Majority)}}
-%  We proceed through a similar analysis for the parity case.
-%  \begin{align*}
-%    & \hat{f}(0) = \mathbb{E}_x f(x) = \frac{{n \choose \frac{n}{2}} + {n \choose \frac{n}{2} + 1} + ... + {n \choose n}}{2^n} = \frac{2^{n-1}}{2^n} = \frac{1}{2} \\
-%    & \hat{f}(1) =
-%  \end{align*}
-%\end{example}
+\begin{example}{\emph{(Fourier coefficients for $\text{Maj}_3$)}}
+ We proceed through a similar analysis for the parity case. Let $f = \text{Maj}_3$.
+ \begin{align}
+   & \hat{f}(0^3) = \mathbb{E}_x[f(x)] = \frac{1}{2} \\
+   & \hat{f}(\{001,010,100\}) = - \frac{1}{4} \\
+   & \hat{f}(\{011,110,101\}) = 0 \\
+   & \hat{f}(1^3) = \frac{1}{8}\sum_{x \in \{0,1\}^3} (-1)^{|x|}f(x) = \frac{1}{4}
+ \end{align}
+ There is an exact formula for calculating the Fourier coefficients for Maj$_{n}$. It can be found in \cite{odonnell} page 108.
+\end{example}
+%
+We end with a defintion. Recall that we will identify $n$-bit strings with subsets $S \subseteq [n]$.
+%
+\begin{definition}
+  Let $f:\{0,1\}^n \rightarrow \{0,1\}$ be an $n$-ary boolean function. The \emph{Fourier degree} of $f$, denoted by def$_{\mathcal{F}}(f)$, is the largest $|S|$ such that $\hat{f}(S) \neq 0$.
+\end{definition}
+%
+The Fourier Transform here is over the group $\mathbb{Z}_2^n$.
+
+\begin{remark}{\emph{(Further Topics)}}
+  There are still many essential concepts in the Fourier Analysis of boolean functions which we have not covered in this chapter.
+For instance, one can endow a multiplication operator to $L_2(G)$ called the \emph{convolution operator}
+  $$ (f * g)(x) = \frac{1}{|G|}\sum_{\substack{h,j \in G \\ hj = x}} f(h)g(j) = \frac{1}{|G|}\sum_{y \in G} f(xy^{-1})g(y), \quad f,g \in L_2(G)$$
+   The convolution $f * g$ can be interpreted as the average of $f(h)\cdot g(j)$ over all elements $h,j \in G, \; h + j = x$.
+Many natural, interesting functions can be defined through the convolution of two boolean functions.
+For instance, let ${\bf 1}_{B_r}$ be the indictator function for a closed ball centered at $0$ of radius $r$ in respect to the Hamming metric on $\mathbb{Z}_2^n$.
+Then for any $n$-ary boolean function $f:\{0,1\}^n \rightarrow \{0,1\}$, $f * {\bf 1}_{B_r}(x)$ will be the expected value of $f$ over the closed ball of radius $r$ centered at $x$.
+If one wished to proceed further, he/she can show that the convolution operator satisfies many fundamental algebraic properties such as associativity and commutativity, turning $L_2(G)$ into a commutative $\mathbb{C}$-algebra.
+   For a more detailed introduction to these topics, see \cite{hatami},\cite{odonnell}. For a more general approach to harmonic analysis over groups, see \cite{rudin}, \cite{ceccherini}.
+\end{remark}
 
@@
 The above properties on characters can be shown through techniques in Representation Theory.
-For instance, we know representations of finite groups are group homomorphisms into the group of linear automorphisms of a suitable complex vector space $\rho: G \rightarrow GL(\mathbb{C}^n)$.
+For instance, we know representations of finite groups are group homomorphisms into the group of linear automorphisms of a suitable finite-dimensional complex vector space $\rho: G \rightarrow GL(\mathbb{C}^n)$.
 The \emph{characters} of $\rho$ are defined as $\chi(x) = Tr(\rho(x))$ where $Tr$ is the trace operator.
@@
 $$ \chi(a+b) = Tr(\rho(a+b)) = Tr(\rho(a)\circ\rho(b)) = \lambda_a \lambda_b = Tr(\rho(a))Tr(\rho(b)) = \chi(a)\chi(b)$$ Note that this relation does not hold for arbitrary finite groups.
-For more information on the Representation Theory of Finite Groups, see \ref{serre}.
+The genealizations of the Fourier Transform generally lie in \emph{Harmonic Analysis}.
+For more information on the Representation Theory of Finite Groups, see \cite{serre}, \cite{webb}.
 \end{remark}
 
-%\section{Linai-Mansour-Nisan Theorem}
+\section{Linial-Mansour-Nisan Theorem}
+We now present the LMN (Linial-Mansour-Nisan) Theorem concerning the Fourier tail of boolean functions in $\AC^0$.
+\begin{definition}
+  Let $k$ be a positive integer and $f:\mathbb{Z}_2^n \rightarrow \mathbb{C}$ a complex-valued function on $\mathbb{Z}_2^n$. We define: \newline
+  $$f^{\leq k} := \sum_{|S| \leq k} \hat{f}(S)\chi_{S} $$
+  The notation symbols $f^{=k},f^{\geq k}$ are defined in the same manner.
+\end{definition}
+%
+Let us state the relevant theorem first before proceeding further
+%
+\begin{theorem} (Linial, Mansour, Nisan \cite{lmn}) \label{lmn}
+    Let $f$ be a boolean fucntion computed by a circuit of depth $d$ and size $M$ and let $t$ be any non-negative integer. Then
+    \begin{equation}
+      \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2M2^{-t^{1/d}/20}
+    \end{equation}
+\end{theorem}
 
-%\section{Immediate Applications}
+The theorem reveals that the $t$-tails of the Fourier spectrum, i.e strings indexed by sets $|S| > t$, become exponentially small in $t$ for boolean functions in $\AC^0$.
+To prove the theorem, we need version of the H\aa stad's switching lemma shown in \cite{hastad}.
+%
+\begin{theorem}{(\emph{H\aa stad}, \cite{hastad})}
+  Let $f$ be given by a CNF-formula where each clause has size at most $t$, and choose a random restriction $\rho$ with parameter $p$ such that $Pr[\rho(x_i)] = p$ for all input variables $x_i$.
+With probability of at least $1 - (5pt)^s$, $f_{\rho}$ can be expressed as a DNF formula where each clause has size of at most $s$, and the clause all accept disjoint sets of inputs i.e no string $x \in \{0,1\}^n$ satisfies more than one clause.
+\end{theorem}
+From H\aa stad's switching lemma, we derive two integral corollaries:
+%
+\begin{corollary} \label{hastaddegree}
+Let $f$ be a boolean function computed by a CNF of bottom fan-in of at most $t$, and $\rho$ is a $p$-random restriction, then
+\begin{equation}
+  Pr[\text{deg}_{\mathcal{F}}(f_\rho) > s] < (5pt)^s
+\end{equation}
+\end{corollary}
+%
+\begin{proof}
+  By H\aa stad's switching lemma, we can convert the CNF computing $f$, into a DNF of bottom fan-in of at most $s$ with probability of at least $1 - (5pt)^s$, such that the clauses accept disjoint sets of inputs.
+Suppose that we had a subset $S \subseteq [n]$ such that $|S| > s$.
+Pick one such $\wedge$-clause $C$ and let $x_{C} \subseteq [n]$ denote the bit-indices of the input variables.
+By our assumptions, $|x_C| \leq s$.
+Now w.l.o.g set $S$ to be some subset of $[n]$ such that $x_C \subset S$ and $|S| = |x_C| + 1$.
+Since each $\wedge$-clause in the DNF accepts disjoint sets of inputs, they cannot contribute to the coefficient.
+Calculating the Fourier coefficients shows us that $\hat{f}(S) = 2,-2 \neq 0$, so $\text{deg}_{\mathcal{F}}(f_\rho) > s$.
+\end{proof}
+
+\begin{corollary} \label{fouriertail}
+  Let $f$ be a boolean function computred by a circuit of size $M$ and depth $d$. Then
+  $$ Pr[\text{deg}(f_{\rho}) > s] \leq M2^{-s} $$
+  where $\rho$ is a random restriction where $p = \frac{1}{10^ds^{d-1}}$
+\end{corollary}
+%
+\begin{proof}
+   First, we will repeatedly iterate the switching lemma to reduce the depth of $f$ through random restrictions such that the bottom fan-in at each step is at most $s$.
+This is accomplished by sampling the first round of restrictions $\rho_0$ to parameter $p_0 = \frac{1}{10}$ and by sampling the subsequent $d-1$ rounds $\rho_i$ according to parameter $p_i = \frac{1}{10s}$.
+We begin by claiming that, after the first restriction, the bottom gates' fan-ins are at most $s$.
+This can be decomposed into two cases:
+   %
+   \begin{enumerate}
+     \item Suppose a gate initially has fan-in of at least $2s$ and (w.l.o.g) suppose that it is a $\wedge$-gate.
+The probability of $\rho_0$ setting at least of the input variables to $0$ is at most $(\frac{1}{10} + \frac{9}{20})^{2s} = (0.55)^{2s} < 2^{-s}$.
+     \item Suppose a gate initially has fan-in of at most $2s$. The probability of $\rho$ setting at least $s$ of the input variables to $*$ is ${2s \choose s}(1/10)^s < 2^{-s}$
+   \end{enumerate}
+   %
+  Since we assume that the $\AC^0$ circuit computing $f$ is in alternating normal form, once we invoke the switching lemma for the bottom two layer gates, we can merge the two $\vee$-gates to reduce the depth by one.
+The switching tells us that with probability of at least $1 - (5\frac{1}{10s}s)^s < 2^{-s}$, we can reduce the depth in each step such that the gates of distance two from the input variables have bottom fan-in of at most $s$.
+By the end of our depth-reduction routine, we are left with a CNF with bottom fan-in of at most $s$.
+Invoking Corollary (\ref{hastaddegree}) tells us that our final restricted function $f_{\rho_0\rho_1...\rho_{d-1}}$ has Fourier degree of at most $s$ with probability $< 2^{-s}$.
+The final bound arises from the observation that we applied the beginning analysis for each of the bottom gates exactly once, the switching lemma exactly once for each gate except for the output gate, and we invoked the corollary for the output gate.
+Each step contributed probability of $< 2^{-s}$
+\end{proof}
+%
+Corollary \ref{fouriertail} tells us that the probability of the Fourier degree exceeding $s$ decays exponentially with $s$ given that we set our probability parameter to a specific quantity $p = \frac{1}{10^ds^{d-1}}$.
+This is reminiscent of the familiar Chernoff bounds encountered in
+probability theory.
+
+\begin{proof}{\emph{(Proof of Theorem \ref{lmn})}}
+We begin by first setting our probability parameter $p \leq \frac{1}{10^dk^{d-1}}$. The values $p, k$ will be fixed later to invoke Corollary \ref{fouriertail}.
+We sample a random restriction $\rho$ by sampling some $V \subseteq$ to be the indices which are not set to $*$.
+Each index has a $1-p$ probability of being set to either $0,1$.
+For each index in $V$, we uniformly sample some bit string in $\{0,1\}^{|V|}$ to fix the indices contained in $V$.
+If we recall that our characters are defined as $\chi_A(x) = (-1)^{\sum_{i \in A} x_i}$ and let $x_{V}$ be the restriction of string $x$ on the input indices to those found in $V$ \newline
+%
+\begin{align}
+  \chi_S(x) = (-1)^{\sum_{i \in S} x_i}
+  & = (-1)^{\sum_{i \in S \cap V} x_i + \sum_{i \in S / V} x_i}   \\
+  & = (-1)^{\sum_{i \in S \cap V} x_i}(-1)^{\sum_{i \in S / V} x_i} = \chi_{S \cap V}(x_V) \chi_{S / V}(x_{\overline{V}})
+\end{align}
+%
+Let $f_{x_V} = f(x_V,*)$ as a function $f_{x_V}: \{0,1\}^{|x_{\overline{V}}|} \rightarrow \{0,1\}$.
+We can rearrange the Fourier decomposition of any function $f$ as follows:
+%
+\begin{align}
+f(x)
+& = \sum_{S \subseteq [n]} \hat{f}(S)\chi_{S}(x) =  \sum_{V \sqcup \overline{V} \subseteq [n]} \hat{f}(S)\chi_{S \cap V}(x_V)\chi_{S / V}({x_{\overline{V}}}) \\
+& = \sum_{H \subseteq \overline{V}} \left( \sum_{J \subseteq V} \hat{f}(H \cup J) \chi_{J}(x_V) \right) \chi_H(x_{\overline{V}})
+\end{align}
+%
+The last equality just rearranges the sum via iterating through all subsets $S \subseteq [n]$ by considering susbets of each partition $V, \overline{V}$ separately.
+Since the Fourier decomposition is unique for $f_{x_V}$,
+\begin{equation}
+  \widehat{f_{x_V}}(H) = \sum_{J \subseteq V} \hat{f}(H \cup J)\chi_{J}(x_V), \quad H \subseteq \overline{V}
+\end{equation}
+By Parseval's identity on the function $x_T \mapsto \widehat{f_{x_T}}(H)$,
+\begin{equation}
+  \mathbb{E}_{x_V}|\widehat{f_{x_V}}(H)|^2 = \langle \hat{f}_{-}(H), \hat{f}_{-}(H) \rangle = \sum_{J \subseteq V} |\hat{f}(H \cup J)|^2
+\end{equation}
+%
+This along with another application of Parseval's identity yields the string of equalities: \newline
+\begin{align}
+  \mathbb{E}_{x_V}||f^{>k}_{x_V}||_2^2 = \mathbb{E}_{x_V}\sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} |\widehat{f_{x_V}}(H)|^2 = \sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} \sum_{J \subseteq V} |\hat{f}(H \cup J)|^2 = \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2
+\end{align}
+%
+Sampling over all such $p$-restrictions further shows that:
+\begin{equation} \label{coroinvoke}
+  \mathbb{E}_V\mathbb{E}_{x_V} ||f^{>k}_{x_V}||_2^2 \leq Pr[\text{deg}_{\mathcal{F}}(f_\rho) > k] \leq M2^{-k}
+\end{equation}
+The second-to-last inequality follows since $||f_\rho^{>k}||_2^2 \leq 1$.
+So we can just replace $||f_\rho^{>k}||_2^2$ with an indicator random variable flipping to one if deg$_{\mathcal{F}}(f_\rho) > k$ and zero otherwise.
+The last inequality is from Corollary (\ref{fouriertail}).
+
+\begin{lemma}
+  For any boolean function $f$ and $0 < p < 1$,
+  $$ \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2 \mathbb{E}_T \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2 $$
+  for $k = pt/2$
+\end{lemma}
+
+\begin{proof}
+  We know that:
+  \begin{align} \label{ineq1}
+    \mathbb{E}_T\sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2 = \sum_{S \subseteq [n]} Pr[|S \cap \overline{T}| > k]|\hat{f}(S)|^2 \geq \sum_{|S| > t} Pr[|S \cap \overline{T}| > k]|\hat{f}(S)|^2
+  \end{align}
+  By invoking a Chernoff bound on the probability that $|A \cap \overline{T}|$ will be $\leq pt/2$
+  $$  Pr[|A \cap \overline{T}| < (1- \frac{2|S| - t}{2|S|})p|S|] \leq  e^{-\frac{pt }{8}}$$ Thus, $Pr[|A \cap \overline{T}| > \frac{pt}{2}] \geq 1 - e^{-\frac{pt }{8}} \geq \frac{1}{2}$.
+This shows that:
+    $$ \sum_{|S| > t} \frac{1}{2}|\hat{f}(S)|^2 \leq \mathbb{E}_T \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2 $$ by (\ref{ineq1}).
+\end{proof}
+To finish the proof, set $p = \frac{1}{10t^{(d-1)/d}}$ and $k = t^{1/d}/20$.
+Direct calculation shows that $p \leq \frac{1}{10^dk^{d-1}}$.
+This allows to us to invoke Corollary $(\ref{fouriertail})$ through (\ref{coroinvoke}) to yield the desired outcome:
+\begin{equation}
+  \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2M2^{-t^{1/d}/20}
+\end{equation}
+\end{proof}
+
+\section{Basic Applications}
 
@@
 \bibliographystyle{amsplain}
-%    Insert the bibliography data here.
+\bibliography{biblio}
 

### 2020-11-11 CS590/paper.tex
Finished first draft of cS590 project

@@
 Then for any $n$-ary boolean function $f:\{0,1\}^n \rightarrow \{0,1\}$, $f * {\bf 1}_{B_r}(x)$ will be the expected value of $f$ over the closed ball of radius $r$ centered at $x$.
-If one wished to proceed further, he/she can show that the convolution operator satisfies many fundamental algebraic properties such as associativity and commutativity, turning $L_2(G)$ into a commutative $\mathbb{C}$-algebra.
+Additionally, the convolution operator satisfies many fundamental algebraic properties such as associativity and commutativity, turning $L_2(G)$ into a commutative $\mathbb{C}$-algebra.
    For a more detailed introduction to these topics, see \cite{hatami},\cite{odonnell}. For a more general approach to harmonic analysis over groups, see \cite{rudin}, \cite{ceccherini}.
@@
 The final bound arises from the observation that we applied the beginning analysis for each of the bottom gates exactly once, the switching lemma exactly once for each gate except for the output gate, and we invoked the corollary for the output gate.
-Each step contributed probability of $< 2^{-s}$
+Each step contributed probability of $< 2^{-s}$.
 \end{proof}
@@
 Corollary \ref{fouriertail} tells us that the probability of the Fourier degree exceeding $s$ decays exponentially with $s$ given that we set our probability parameter to a specific quantity $p = \frac{1}{10^ds^{d-1}}$.
-This is reminiscent of the familiar Chernoff bounds encountered in
-probability theory.
 
@@
 The last inequality is from Corollary (\ref{fouriertail}).
-
+%
 \begin{lemma}
@@
 \end{lemma}
-
+%
 \begin{proof}
@@
 
+\subsection{Approximation by Low-degree Polynomials}
+ Theorem (\ref{lmn}) shows that we can approximate functions in $f \in \AC^0$ by taking our polynomials to be $f^{\leq k}$ for some sufficiently large $k$.
+Here we make the distinction that our Fourier expansion of $f$ will be
+$$f = \sum_{S \subseteq [n]} \hat{f}(S) \prod_{i \in S} x_i $$ as a polynomial with complex coefficients.
+\begin{lemma}
+  Let $f \in \AC^0$ be a boolean function of polynomial size and depth $d$. Then there exists a complex polynomial of degree $\mathcal{O}((\log{n/\epsilon})^d)$ such that $||f - p ||_2 < \epsilon$
+\end{lemma}
+%
+\begin{proof}
+This follows from setting $t = \mathcal{O}((\log{n/\epsilon})^d)$ and invoking the LMN theorem to conclude that $||f - f^{\leq t}|| < \epsilon$.
+\end{proof}
+%
+This gives another scheme of approximating functions in $\AC^0$ through low-degree polynomials, complementing the results of \cite{razborov}, \cite{smolensky}.
+Here we make the extra distinction that our polynomials are also of low Fourier degree.
+
+\subsection{Sensitivity and Influence}
+LMN also shows that functions in $\AC^0$ have low average sensitivity i.e its output is not very sensitive to changes to the input.
+
+\begin{definition}
+  Let $f: \{0,1\}^n \rightarrow \{0,1\}$ be a boolean function.
+We define the \emph{sensitivity} of an input $x\in \{0,1\}^n$ in respect to $f$, $s_f(x)$ to be the number of indices $i$ such that $f(x) \neq f(x + e_i)$ where $e_i$ is the bit string with zeros everywhere except for the $i^{th}$ index.
+\end{definition}
+
+\begin{definition}
+  Let $f: \{0,1\}^n \rightarrow \{0,1\}$ be a boolean function. The \emph{influence} of $f$, $I_f$ is defined as the average sensitivity over all input bit strings
+  \begin{equation}
+    I_f = \mathbb{E}_x[s_f(x)]
+  \end{equation}
+\end{definition}
+
+The influence of $f$ can be expressed in terms of its Fourier coefficients:
+
+\begin{equation}
+  I_f = 4\sum_{S \subseteq [n]} |S||\hat{f}(s)|^2
+\end{equation}
+
+By combining this equivalence with Theorem (\ref{lmn}), we deduce the following upper bound on the influence of a function in $\AC^0$
+
+\begin{lemma}
+  Let $f \in \AC^0$ be of depth $d$. Then
+  \begin{equation}
+    I_f = \mathcal{O}((\log n)^d)
+  \end{equation}
+\end{lemma}
+
 %    Bibliographies can be prepared with BibTeX using amsplain,

### 2020-11-11 CS590/slides.tex
Finished first draft of cS590 project

\documentclass{beamer}
\usepackage{pgfpages}

\usetheme{Madrid}
\useinnertheme{rectangles}
\usefonttheme{serif}

\title{Linial-Mansour-Nisan Theorem}
\subtitle{Crash Course on Fourier Analysis on Boolean Functions}
\author{Edward Kim}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\label{contents}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{Fourier Analysis on Boolean Functions}
\subsection{sub a}

\begin{frame}
\frametitle{List}
\begin{itemize}
\pause
\item Point A
\pause
\item Point B
\begin{itemize}
\pause
\item part 1
\pause
\item part 2
\end{itemize}
\pause
\item Point C
\pause
\item Point D
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{More Lists}
\begin{enumerate}[(I)]
\item<1-> Point A
\item<2-> Point B
\begin{itemize}
\item<3-> part 1
\item<4-> part 2
\end{itemize}
\item<5-> Point C
\item<-2,4-5,7> Point D
\end{enumerate}
\end{frame}

\subsection{sub b}
\setbeamercovered{transparent}
\begin{frame}
\frametitle{Overlays}
\only<1>{First Line of Text}

\only<2>{Second Line of Text}

\only<3>{Third Line of Text}
\end{frame}

\begin{frame}
\frametitle{Overlay Compatible Commands}
\textbf<2>{Example Text}

\textit<3>{Example Text}

\textsl<4>{Example Text}

\textrm<5>{Example Text}

\textsf<6>{Example Text}

\textcolor<7>{orange}{Example Text}

\alert<8>{Example Text}

\structure<9>{Example Text}
\end{frame}

\begin{frame}
\label{columns}
\frametitle{Using Columns}
\begin{columns}
\column{0.5\textwidth}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
\column{0.5\textwidth}
\end{columns}
\end{frame}

\section{Section 2}

\begin{frame}
Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
\end{frame}

\begin{frame}
\frametitle{Descriptions}
\begin{description}
\item[API] Application Programming Interface
\item[LAN] Local Area Network
\item[ASCII] American Standard Code for Information Interchange
\end{description}
\end{frame}

\begin{frame}
\frametitle{Blocks}
\begin{block}{Block Title}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
\end{block}
\begin{alertblock}{Block Title}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
\end{alertblock}
\end{frame}

\begin{frame}
\frametitle{More Blocks}
\begin{definition}
A prime number is a number that...
\end{definition}
\begin{example}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
\end{example}
\end{frame}

\begin{frame}
\frametitle{Maths Blocks}
\begin{theorem}<1->[Pythagoras]
$ a^2 + b^2 = c^2$
\end{theorem}
\begin{corollary}<3->
$ x + y = y + x  $
\end{corollary}
\begin{proof}<2->
$\omega +\phi = \epsilon $
\end{proof}
\end{frame}

\begin{frame}[fragile]
\frametitle{Including Code}
\begin{semiverbatim}
\\begin\{frame\}
\\frametitle\{Outline\}
\\tableofcontents
\\end\{frame\}
\end{semiverbatim}
\end{frame}

\begin{frame}
\frametitle{buttons}
\hyperlink{contents}{\beamerbutton{contents page}}

\hyperlink{columns}{\beamergotobutton{columns page}}

\hyperlink{pictures}{\beamerskipbutton{pictures page}}

\hyperlink{pictures}{\beamerreturnbutton{pictures page}}
\end{frame}

\end{document}

### 2020-11-11 CS590/paper.tex
Actually first draft

@@
 
-\begin{lemma}
+\begin{lemma} \label{influence}
   Let $f \in \AC^0$ be of depth $d$. Then
@@
 
+Lemma (\ref{influence}) shows that functions in $\AC^0$  are not suitable for constructing universal hash functions (\cite{mansour}) and pseudorandom function generators (\cite{lmn}).
+We give a rough definition and sketch based on the treatment found in (\cite{lmn}.
+
+\begin{definition}
+  A function $f:\{0,1\}^m \times \{0,1\}^n \rightarrow \{0,1\}$ is called a \emph{pseudorandom function generator} if there exists no polynomial-time oracle Turing machine which can distinguish between the outputs from a true random oracle versus $f(s,_)$ for some random seed $s \in \{0,1\}^m$.
+\end{definition}
+
+\begin{lemma}
+  No pseudorandom function generators exist in $\AC^0$
+\end{lemma}
+By taking advantage of the low average sensitivity, we can simply perturb the input slightly and check if the output of $f$ changes. If it doesn't, there is a good chance that
+it lies in $\AC^0$.
+
+\begin{remark}{(\emph{Some concluding remarks})}
+  It is worth noting that the bounds shown by Linai-Mansour-Nisan have been improved by Tal (\cite{tal}).
+Specifically, he showed that for any $n$-ary boolean function $f$ with size $M$ and depth $d$,\
+  \begin{equation}
+    \sum_{|S| > t} |\hat{f}(S)|^2 = 2^{\Omega(k/\log^{d-1}{M})}
+  \end{equation}
+\end{remark}
 %    Bibliographies can be prepared with BibTeX using amsplain,

### 2020-11-12 CS590/paper.tex
Commiting finished presentation slides for circuit complexity class

@@
 \begin{corollary} \label{fouriertail}
-  Let $f$ be a boolean function computred by a circuit of size $M$ and depth $d$. Then
+  Let $f$ be a boolean function computed by a circuit of size $M$ and depth $d$. Then
   $$ Pr[\text{deg}(f_{\rho}) > s] \leq M2^{-s} $$
@@
 This is accomplished by sampling the first round of restrictions $\rho_0$ to parameter $p_0 = \frac{1}{10}$ and by sampling the subsequent $d-1$ rounds $\rho_i$ according to parameter $p_i = \frac{1}{10s}$.
-We begin by claiming that, after the first restriction, the bottom gates' fan-ins are at most $s$.
+We begin by claiming that, after the first restriction, the bottom gates' fan-ins are at most $s$ with probability of at least $1 - 2^{-s}$.
 This can be decomposed into two cases:
@@
 By Parseval's identity on the function $x_T \mapsto \widehat{f_{x_T}}(H)$,
-\begin{equation}
+\begin{align}
   \mathbb{E}_{x_V}|\widehat{f_{x_V}}(H)|^2 = \langle \hat{f}_{-}(H), \hat{f}_{-}(H) \rangle = \sum_{J \subseteq V} |\hat{f}(H \cup J)|^2
-\end{equation}
+\end{align}
 %
@@
 \begin{align}
-  \mathbb{E}_{x_V}||f^{>k}_{x_V}||_2^2 = \mathbb{E}_{x_V}\sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} |\widehat{f_{x_V}}(H)|^2 = \sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} \sum_{J \subseteq V} |\hat{f}(H \cup J)|^2 = \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2
+  \mathbb{E}_{x_V}||f^{>k}_{x_V}||_2^2 = \mathbb{E}_{x_V}\sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} |\widehat{f_{x_V}}(H)|^2
+  & = \sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} \sum_{J \subseteq V} |\hat{f}(H \cup J)|^2 \\
+  & = \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2
 \end{align}

### 2020-11-12 CS590/slides.tex
Commiting finished presentation slides for circuit complexity class

@@
 \usepackage{pgfpages}
+\usepackage{complexity}
 
@@
 \section{Fourier Analysis on Boolean Functions}
-\subsection{sub a}
+\subsection{Characters}
 
 \begin{frame}
-\frametitle{List}
-\begin{itemize}
-\pause
-\item Point A
-\pause
-\item Point B
-\begin{itemize}
-\pause
-\item part 1
-\pause
-\item part 2
-\end{itemize}
-\pause
-\item Point C
-\pause
-\item Point D
-\end{itemize}
+\frametitle{Characters}
+Let us first define some general concepts for any finite abelian group $G$.
+
+\begin{enumerate}
+\begin{definition}
+  A group homomorphism from $\chi: G \rightarrow \mathbb{C}^*$ is called a \emph{character} of $G$ where $\mathbb{C}^* = \mathbb{C}/\{0\}$
+\end{definition}
+
+\item We call the homomorphism $\chi_0:G \rightarrow \mathbb{C}^*, \; \chi_0 = 1$ the \emph{trivial character} of $G$.
+\item $\chi(a + b) = \chi(a)\chi(b)$
+\item  The characters of $G$ form an abelian group $\hat{G}$ under pointwise multiplication of complex-valued functions. The group $\hat{G}$ is known as the \emph{character group} of $G$.
+\end{enumerate}
 \end{frame}
@@
 \begin{frame}
-\frametitle{More Lists}
-\begin{enumerate}[(I)]
-\item<1-> Point A
-\item<2-> Point B
-\begin{itemize}
-\item<3-> part 1
-\item<4-> part 2
-\end{itemize}
-\item<5-> Point C
-\item<-2,4-5,7> Point D
+\frametitle{Characters}
+  \begin{enumerate}
+    \item It follows from basic properties of cyclic groups that the only characters of $\mathbb{Z}_n$ are ones of the form $$ \chi_j(x) = e^{ 2\pi i jx/n} \quad j \in [n], \; x \in \mathbb{Z}_n $$ Note that each character is associated to an element of the group $j \in \mathbb{Z}_n$
+    \begin{theorem}[Characters for finite abelian groups]
+    For any finite abelian group $G \cong \mathbb{Z}_{n_1} \times \mathbb{Z}_{n_2} \cdots \times\mathbb{Z}_{n_k}$: the characters for $G$ will be:
+    $$   \chi_a(x) = \prod_{i \in [k]} e^{2 \pi i a_k x_k/ n_k} $$
+    for $a \in \mathbb{Z}_{n_1} \times \mathbb{Z}_{n_2} \cdots \times\mathbb{Z}_{n_k}$
+    \end{theorem}
   \end{enumerate}
@@
 
-\subsection{sub b}
-\setbeamercovered{transparent}
 \begin{frame}
-\frametitle{Overlays}
-\only<1>{First Line of Text}
-
-\only<2>{Second Line of Text}
-
-\only<3>{Third Line of Text}
+\frametitle{Characters}
+  \begin{enumerate}
+    \item It turns out the characters live in the vector space $L_2(G)$ i.e maps from $\phi: G \rightarrow \mathbb{C}$ square-integrable in respect to the uniform probability measure: $\frac{1}{|G|}\sum_{x \in G} |\phi(x)|^2 < \infty$
+    \item Actually a Hilbert space endowed with the inner product:
+    $$ \langle f,g \rangle = \frac{1}{|G|}\sum_{x\in G} f(x)\overline{g(x)}= \mathbb{E}_x f(x)\overline{g(x)} $$
+    \begin{theorem}
+      The characters $\chi \in \hat{G}$ form an orthonormal basis for $L_2(G)$
+    \end{theorem}
+    \item
+    With any orthonormal basis, one can decompose any vector into its direct sum decomposition.
+  \end{enumerate}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Overlay Compatible Commands}
-\textbf<2>{Example Text}
+\frametitle{Characters}
+  \begin{enumerate}
+    \item We restrict our attention to the finite abelian group $\mathbb{Z}_2^n$ as a natural group to define boolean functions $f: \{0,1\}^n \rightarrow \{0,1\}$.
+    \item By the observations above, we conclude that our characters for $\mathbb{Z}_2^n = \mathbb{Z}_2 \times ... \times \mathbb{Z}_2$ are defined as
+    $$  \chi_a(x) = (-1)^{\sum_{i \in [n]} x_ia_i} = (-1)^{\sum_{i \in [n], \; a_i = 1} x_i}$$
+    where $x_i$ is the $i^{th}$ bit of $x \in \{0,1\}^n$.
+    \begin{block}{Some Notation}
+      Let $S \subseteq [n]$ be such that $S_x = \{i \mid x_i = 1\}$.
+As there is a bijective correspondence between all such subsets $S_x$ and $x \in \{0,1\}^n$, we will sometimes identify bit strings with their subset counterparts
+      $$ \chi_A(x) = (-1)^{\sum_{i \in A} x_i}, \quad A \subseteq [n]$$
+    \end{block}
+  \end{enumerate}
+\end{frame}
 
-\textit<3>{Example Text}
-
-\textsl<4>{Example Text}
-
-\textrm<5>{Example Text}
-
-\textsf<6>{Example Text}
-
-\textcolor<7>{orange}{Example Text}
-
-\alert<8>{Example Text}
-
-\structure<9>{Example Text}
+\subsection{The Fourier Transform}
+\begin{frame}
+\frametitle{The Fourier Transform}
+  \begin{enumerate}
+    \item Given any $f \in L_2(G)$, let $\hat{f}: \hat{G} \rightarrow \mathbb{C}$ be the complex-valued function such that
+    $$  \hat{f}(\chi_a) =  \langle f, \chi_a \rangle, \; a \in G $$
+    These are the projections onto the orthonormal basis of characters. We deem $\hat{f}$ the \emph{Fourier Transform} of $f$.
+    \item The direct sum decomposition of $f$ yields the following form known as the \emph{Fourier Inversion Formula}:
+    $$  f = \sum_{a \in G} \hat{f}(\chi_a)\chi_a $$
+    The complex value $\hat{f}(\chi_a)$ is called the \emph{Fourier coefficient} associated to $\chi_a$
+    \item From the above definition, we directly calculate that
+    $$  \hat{f}(0) = \langle f, \chi_0 \rangle = \mathbb{E}_x f(x) $$
+  \end{enumerate}
 \end{frame}
@@
 \begin{frame}
-\label{columns}
-\frametitle{Using Columns}
-\begin{columns}
-\column{0.5\textwidth}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-\column{0.5\textwidth}
-\end{columns}
-\end{frame}
-
-\section{Section 2}
-
-\begin{frame}
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\frametitle{The Fourier Transform}
+\begin{enumerate}
+  \item
+  Let $f = \text{Maj}_3$ where $G = \mathbb{Z}_2^n$.
+  \begin{align}
+    & \hat{f}(0^3) = \mathbb{E}_x[f(x)] = \frac{1}{2} \\
+    & \hat{f}(\{001,010,100\}) = - \frac{1}{4} \\
+    & \hat{f}(\{011,110,101\}) = 0 \\
+    & \hat{f}(1^3) = \frac{1}{8}\sum_{x \in \{0,1\}^3} (-1)^{|x|}f(x) = \frac{1}{4}
+  \end{align}
+  \begin{definition}
+    Let $f:\{0,1\}^n \rightarrow \{0,1\}$ be an $n$-ary boolean function. The \emph{Fourier degree} of $f$, denoted by def$_{\mathcal{F}}(f)$, is the largest $|S|$ such that $\hat{f}(S) \neq 0$.
+  \end{definition}
+  For the case of f = Maj$_3$, def$_{\mathcal{F}}(f) = 3$
+\end{enumerate}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Descriptions}
-\begin{description}
-\item[API] Application Programming Interface
-\item[LAN] Local Area Network
-\item[ASCII] American Standard Code for Information Interchange
-\end{description}
+\frametitle{The Fourier Transform}
+  \begin{theorem}[Parseval's Identity]
+    Let $f \in L_2(G)$. Then
+    $$ ||f||_2^2 = \sum_{a \in G} |\hat{f}(a)|^2$$
+  \end{theorem}
+  \begin{proof}
+    \begin{align*}
+      ||f||^2_2 = \langle f, f \rangle = \langle \sum_{a \in G} \hat{f}(a) \chi_a,  \sum_{b \in G} \hat{f}(b) \chi_b \rangle & = \sum_{a,b \in G} \hat{f}(a)\overline{\hat{f}(b)}\langle \chi_a, \chi_b \rangle \\
+      & = \sum_{a \in G} |\hat{f}(a)|^2
+    \end{align*}
+
+    The last equality is just the orthonormality of the characters.
+  \end{proof}
+\end{frame}
+
+\section{LMN Theorem}
+
+\begin{frame}
+  \frametitle{LMN Theorem}
+  \begin{theorem} (Linial, Mansour, Nisan) \label{lmn}
+      Let $f$ be a boolean fucntion computed by a circuit of depth $d$ and size $M$ and let $t$ be any non-negative integer. Then
+      \begin{equation}
+        \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2M2^{-t^{1/d}/20}
+      \end{equation}
+  \end{theorem}
+  The theorem reveals that the $t$-tails of the Fourier spectrum, i.e strings indexed by sets $|S| > t$, become exponentially small in $t$ for boolean functions in $\AC^0$.
 \end{frame}
@@
 \begin{frame}
-\frametitle{Blocks}
-\begin{block}{Block Title}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-\end{block}
-\begin{alertblock}{Block Title}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-\end{alertblock}
+  \frametitle{LMN Theorem}
+
+  \begin{theorem}[\emph{H\aa stad}]
+    Let $f$ be given by a CNF-formula where each clause has size at most $t$, and choose a random restriction $\rho$ with parameter $p$ such that $Pr[\rho(x_i)] = p$ for all input variables $x_i$.
+With probability of at least $1 - (5pt)^s$, $f_{\rho}$ can be expressed as a DNF formula where each clause has size of at most $s$, and the clause all accept disjoint sets of inputs i.e no string $x \in \{0,1\}^n$ satisfies more than one clause.
+  \end{theorem}
+
+  \begin{corollary} \label{hastaddegree}
+  Let $f$ be a boolean function computed by a CNF of bottom fan-in of at most $t$, and $\rho$ is a $p$-random restriction, then
+  \begin{equation}
+    Pr[\text{deg}_{\mathcal{F}}(f_\rho) > s] < (5pt)^s
+  \end{equation}
+  \end{corollary}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{More Blocks}
-\begin{definition}
-A prime number is a number that...
-\end{definition}
-\begin{example}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-\end{example}
+  \frametitle{LMN Theorem}
+  \begin{corollary}[Tail Degree Corollary] \label{fouriertail}
+    Let $f$ be a boolean function computed by a circuit of size $M$ and depth $d$. Then
+    $$ Pr[\text{deg}(f_{\rho}) > s] \leq M2^{-s} $$
+    where $\rho$ is a random restriction where $p = \frac{1}{10^ds^{d-1}}$
+  \end{corollary}
+  \pause
+  \begin{proof}[Proof Sketch]
+    Show that first random restriction of parameter $p_0 = \frac{1}{10}$, the bottom gates' fan-ins are at most $s$ with probability of at least $1 - 2^{-s}$.
+Then iterate H\aa stad's switching lemma with under $p_i = \frac{1}{10s}$ on each gate of distance two fromm the input variables to turn them into DNFs with disjoint inputs.
+Collapse a level and the lemma ensures that the new bottom fan-in is at most $s$.
+Stop when we are left with a CNF (depth-2) with bottom fan-in of at most $s$ and invoke the previous corollary.
+  \end{proof}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Maths Blocks}
-\begin{theorem}<1->[Pythagoras]
-$ a^2 + b^2 = c^2$
-\end{theorem}
-\begin{corollary}<3->
-$ x + y = y + x  $
-\end{corollary}
-\begin{proof}<2->
-$\omega +\phi = \epsilon $
-\end{proof}
-\end{frame}
-
-\begin{frame}[fragile]
-\frametitle{Including Code}
-\begin{semiverbatim}
-\\begin\{frame\}
-\\frametitle\{Outline\}
-\\tableofcontents
-\\end\{frame\}
-\end{semiverbatim}
+  \frametitle{Proof of LMN Theorem}
+  \begin{definition}
+    Let $k$ be a positive integer and $f:\mathbb{Z}_2^n \rightarrow \mathbb{C}$ a complex-valued function on $\mathbb{Z}_2^n$. We define: \newline
+    $$f^{\leq k} := \sum_{|S| \leq k} \hat{f}(S)\chi_{S} $$
+    The notation symbols $f^{=k},f^{\geq k}$ are defined in the same manner.
+  \end{definition}
+  \pause
+  \begin{enumerate}
+    \item We begin by first setting our probability parameter $p \leq \frac{1}{10^dk^{d-1}}$. The values $p, k$ will be fixed later to invoke the Tail Degree Corollary above.
+    \pause
+    \item
+    Recall how we sample a random restriction $\rho$ by sampling some $V \subseteq [n]$ to be the indices which are \emph{not} set to $*$.
+Each index has a $1-p$ probability of being set to either $0,1$.
+For each index in $V$, we uniformly sample some bit string in $\{0,1\}^{|V|}$ to fix the indices contained in $V$.
+    \pause
+    \item Let $x_{V}$ be the restriction of string $x$ on the input indices to those found in $V$
+  \end{enumerate}
 \end{frame}
@@
 \begin{frame}
-\frametitle{buttons}
-\hyperlink{contents}{\beamerbutton{contents page}}
+  \frametitle{Proof of LMN Theorem}
+  \begin{enumerate}
+    \item If we recall that our characters are defined as $\chi_A(x) = (-1)^{\sum_{i \in A} x_i}$
+    \pause
+    \begin{block}{How characters separate}
+      \begin{align*}
+        \chi_S(x) = (-1)^{\sum_{i \in S} x_i}
+        & = (-1)^{\sum_{i \in S \cap V} x_i + \sum_{i \in S / V} x_i}   \\
+        & = (-1)^{\sum_{i \in S \cap V} x_i}(-1)^{\sum_{i \in S / V} x_i} = \chi_{S \cap V}(x_V) \chi_{S / V}(x_{\overline{V}})
+      \end{align*}
+    \end{block}
+  \end{enumerate}
+\end{frame}
 
-\hyperlink{columns}{\beamergotobutton{columns page}}
+\begin{frame}
+  \frametitle{Proof of LMN Theorem}
+  \begin{enumerate}
+    \item If we set $x_V$ to some bit string, it makes sense to think of $f_{x_V} = f(x_V,*)$ as a function $f_{x_V}: \{0,1\}^{|x_{\overline{V}}|} \rightarrow \{0,1\}$.
+    \pause
+    \item By our Fourier Inversion Formula and the observation in the previous slide:
+    \begin{align*}
+    f(x)
+    & = \sum_{S \subseteq [n]} \hat{f}(S)\chi_{S}(x) =  \sum_{V \sqcup \overline{V} \subseteq [n]} \hat{f}(S)\chi_{S \cap V}(x_V)\chi_{S / V}({x_{\overline{V}}}) \\
+    & = \sum_{H \subseteq \overline{V}} \left( \sum_{J \subseteq V} \hat{f}(H \cup J) \chi_{J}(x_V) \right) \chi_H(x_{\overline{V}})
+  \end{align*}
+    \item
+    Since the Fourier decomposition is unique for $f_{x_V}$,
+    \begin{equation*}
+      \widehat{f_{x_V}}(H) = \sum_{J \subseteq V} \hat{f}(H \cup J)\chi_{J}(x_V), \quad H \subseteq \overline{V}
+    \end{equation*}
+  \end{enumerate}
+\end{frame}
 
-\hyperlink{pictures}{\beamerskipbutton{pictures page}}
+\begin{frame}
+  \frametitle{Proof of LMN Theorem}
+  \begin{enumerate}
+    \item By Parseval's identity on the function $x_T \mapsto \widehat{f_{x_T}}(H)$,
+    \begin{equation}
+      \mathbb{E}_{x_V}|\widehat{f_{x_V}}(H)|^2 = \langle \hat{f}_{-}(H), \hat{f}_{-}(H) \rangle = \sum_{J \subseteq V} |\hat{f}(H \cup J)|^2
+    \end{equation}
+    \pause
+    \item
+    This along with another application of Parseval's identity yields the string of equalities: \newline
+    \begin{align}
+      \mathbb{E}_{x_V}||f^{>k}_{x_V}||_2^2 = \mathbb{E}_{x_V}\sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} |\widehat{f_{x_V}}(H)|^2
+       & = \sum_{\substack {H \subseteq \overline{V} \\ |H| > k}} \sum_{J \subseteq V} |\hat{f}(H \cup J)|^2 \\
+       & = \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2
+    \end{align}
+  \end{enumerate}
+\end{frame}
 
-\hyperlink{pictures}{\beamerreturnbutton{pictures page}}
+\begin{frame}
+  \frametitle{Proof of LMN Theorem}
+  \begin{enumerate}
+    \item Sampling over all such $p$-restrictions further shows that the lefthand side is upper bounded by:
+    \begin{equation*} \label{coroinvoke}
+      \mathbb{E}_V\mathbb{E}_{x_V} ||f^{>k}_{x_V}||_2^2 = \mathbb{E}_{\rho} ||f^{>k}_{\rho}||_2^2 \leq Pr[\text{deg}_{\mathcal{F}}(f_\rho) > k] \leq M2^{-k}
+    \end{equation*}
+    \pause
+    \item A Chernoff bound argument shows that the righthand side is lower bounded by:
+    $$ \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2 \mathbb{E}_T \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2 $$
+    for $k = pt/2$
+    \pause
+    \item Setting our constants $p = \frac{1}{10t^{(d-1)/d}}$ and $k = t^{1/d}/20$ shows that $p \leq \frac{1}{10^dk^{d-1}}$.
+This allows us to invoke our Tail Degree Corollary which gives us the desired inequality
+    \begin{equation}
+      \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2M2^{-t^{1/d}/20}
+    \end{equation}
+  \end{enumerate}
+\end{frame}
+
+\section{Consequences}
+\subsection{Approximation by Low-degree Polynomials}
+
+\begin{frame}
+  \frametitle{Approximation by Low-degree Polynomials}
+   Theorem (\ref{lmn}) shows that we can approximate functions in $f \in \AC^0$ by taking our polynomials to be $f^{\leq k}$ for some sufficiently large $k$.
+Here we make the distinction that our Fourier expansion of $f$ will be
+  $$f = \sum_{S \subseteq [n]} \hat{f}(S) \prod_{i \in S} x_i $$ as a polynomial with complex coefficients.
+  \begin{lemma}
+    Let $f \in \AC^0$ be a boolean function of polynomial size and depth $d$. Then there exists a complex polynomial of degree $\mathcal{O}((\log{n/\epsilon})^d)$ such that $||f - p ||_2 < \epsilon$
+  \end{lemma}
+\end{frame}
+
+\subsection{Sensitivity and Influence}
+
+\begin{frame}
+  \frametitle{Sensitivity and Influence}
+  LMN also shows that functions in $\AC^0$ have low average sensitivity i.e its output is not very sensitive to changes to the input.
+  \begin{definition}
+    Let $f: \{0,1\}^n \rightarrow \{0,1\}$ be a boolean function.
+We define the \emph{sensitivity} of an input $x\in \{0,1\}^n$ in respect to $f$, $s_f(x)$ to be the number of indices $i$ such that $f(x) \neq f(x + e_i)$ where $e_i$ is the bit string with zeros everywhere except for the $i^{th}$ index.
+  \end{definition}
+
+  \begin{definition}
+    Let $f: \{0,1\}^n \rightarrow \{0,1\}$ be a boolean function. The \emph{influence} of $f$, $I_f$ is defined as the average sensitivity over all input bit strings
+    \begin{equation}
+      I_f = \mathbb{E}_x[s_f(x)]
+    \end{equation}
+  \end{definition}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Sensitivity and Influence}
+  The influence of $f$ can be expressed in terms of its Fourier coefficients:
+
+  \begin{equation}
+    I_f = 4\sum_{S \subseteq [n]} |S||\hat{f}(s)|^2
+  \end{equation}
+
+  By combining this equivalence with Theorem (\ref{lmn}), we deduce the following upper bound on the influence of a function in $\AC^0$
+
+  \begin{lemma} \label{influence}
+    Let $f \in \AC^0$ be of depth $d$. Then
+    \begin{equation}
+      I_f = \mathcal{O}((\log n)^d)
+    \end{equation}
+  \end{lemma}
+  The lemma shows that functions in $\AC^0$  are not suitable for constructing universal hash functions and pseudorandom function generators.
+\end{frame}
+
+\begin{frame}
+  \frametitle{Sensitivity and Influence}
+  \begin{definition}
+    A function $f:\{0,1\}^m \times \{0,1\}^n \rightarrow \{0,1\}$ is called a \emph{pseudorandom function generator} if there exists no polynomial-time oracle Turing machine which can distinguish between the outputs from a true random oracle versus $f(s,_)$ for some random seed $s \in \{0,1\}^m$.
+  \end{definition}
+
+  \begin{lemma}
+    No pseudorandom function generators exist in $\AC^0$
+  \end{lemma}
+  By taking advantage of the low average sensitivity, we can simply perturb the input slightly and check if the output of $f$ changes. If it doesn't, there is a good chance that
+  it lies in $\AC^0$.
 \end{frame}

### 2020-11-12 CS590/paper.tex
Uploading 2nd draft of CS590 paper

@@
 We introduce some stepping stones defining the tools used in the Fourier Analysis of boolean functions.
-Many of these tools can be defined using techniques from Representation Theory, but for the sake of briefity, we will not delve into those concepts here in detail.
-Instead, we will provide remarks and snippets on relevant topics as needed along with references to the standard literature at the end of the section.
+Many of these tools can be defined using techniques from Representation Theory of Finite Groups, but for the sake of briefity, we will not delve into those concepts here in detail.
+Instead, we will provide remarks and snippets on relevant topics along with references to some literature at the end of the section.
 Our presentation follows that of \cite{hatami}.
@@
 
-Consider the set whose elements are the characters of $G$ and whose binary operation is pointwise multiplication of complex-valued functions on $G$.
+Consider the set whose elements are the \emph{characters} of $G$ and whose binary operation is \emph{pointwise multiplication} of complex-valued functions on $G$.
 This set is also an \emph{abelian group} by the following theorem:
@@
 We can decompose any finite abelian group $G$ into a direct product of finite cyclic groups $G \cong \mathbb{Z}_{n_1} \times \mathbb{Z}_{n_2} \cdots \times\mathbb{Z}_{n_k}$.
-If we decompose $x = \sum_{i \in [k]} x_i$ for $x \in G$, then the characters of $G$ are defined as
+If we decompose $x = \sum_{i \in [k]} x_i$ for $x_i \in \mathbb{Z}_{n_i},\; x \in G$, then the characters of $G$ are defined as
 %
@@
 
-We see that every element $a \in G$ has an unique associated character $\chi_a$. Furthermore
+\noindent We see that every element $a \in G$ has an unique associated character $\chi_a$. Furthermore
 %
@@
   \chi_{a+b}(x)  = \chi_{a}(x)\chi_{b}(x)
-\end{gather}
+\end{gather} \newline
 
 To define the Fourier transform, we must consider the Hilbert space $L_2(G)$ where $G$ endowed with the uniform probability measure i.e the discrete measure mapping each subset $H \subseteq G$ to $\frac{|H|}{|G|}$.
-This yields the standard inner product for maps $f,g \in L_2(G)$: \newline
+These will be the complex-valued maps $f: G \rightarrow \mathbb{C}$ such that
+\begin{equation}
+  \frac{1}{|G|}\sum_{x \in G} |f(x)|^2 < \infty
+\end{equation}
+%
+\noindent This yields the standard inner product for maps $f,g \in L_2(G)$: \newline
 %
@@
 The last equality follows from Lemma \ref{totalsum} since $\chi\phi^{-1}$ must also be a non-trivial character if $\chi, \phi$ are non-trivial characters.
-The case where one of the character is trivial also follows from the lemma.
+The case where one of the characters is trivial also follows from the lemma.
 Finally, observe that $L_2(G)$ is $|G|$-dimensional as a vector space by considering the dual of $G$.
@@
 %
-We can iterate this for all characters to get a direct sum decomposition for any $f \in L_2(G)$ which we call the \emph{Fourier Inversion Formula}\newline.
+The function $\hat{f}$ is deemed as the \emph{Fourier Transform of $f$}.
+By iterating this for all characters, we get a direct sum decomposition for any $f \in L_2(G)$ which we call the \emph{Fourier Inversion Formula}\newline
 %
@@
 
-It can be shown that the map $f \mapsto \hat{f}$ is linear by the left-linearity of the inner product.
+\noindent It can be shown that the map $f \mapsto \hat{f}$ is linear by the left-linearity of the inner product.
 This linear map is the \emph{Fourier Transform on $G$} and (\ref{orthonormal}) shows that the decomposition is unique.
@@
 %
-From the above definition, we directly calculate that
+\noindent From the above definition, we directly calculate that
 \begin{equation}
@@
 
-A simple application of the Fourier Transform yields an useful identity which will be instrumental in our formulation of the Fourier tail of a boolean function.
+\noindent Additionally, a simple application of the Fourier Transform yields an useful identity which will be instrumental in our formulation of the Fourier tail of a boolean function.
 
@@
 \end{proof}
-%
-By combining Parseval's Identity with the observation that $||\hat{f}||_2^2 = \langle \hat{f}, \hat{f} \rangle = \frac{1}{|G|}\sum_{a \in G} |\hat{f}(a)|^2$, we arrive at the \emph{Plancheral Formula}
+
+\noindent By combining Parseval's Identity with the observation that $||\hat{f}||_2^2 = \langle \hat{f}, \hat{f} \rangle = \frac{1}{|G|}\sum_{a \in G} |\hat{f}(a)|^2$, we arrive at the \emph{Plancheral Formula}
 
@@
 \end{example}
-%
-We end with a defintion. Recall that we will identify $n$-bit strings with subsets $S \subseteq [n]$.
+\noindent We end with a definition. Recall that we will identify $n$-bit strings with subsets $S \subseteq [n]$.
 %
@@
 %
-The Fourier Transform here is over the group $\mathbb{Z}_2^n$.
+\noindent It's worth stressing that the Fourier Transform here is over the group $\mathbb{Z}_2^n$.
 
@@
 %
-Let us state the relevant theorem first before proceeding further
+\noindent Let us state the relevant theorem first before proceeding further
 %
@@
 
-The theorem reveals that the $t$-tails of the Fourier spectrum, i.e strings indexed by sets $|S| > t$, become exponentially small in $t$ for boolean functions in $\AC^0$.
+\noindent The theorem reveals that the $t$-tails of the Fourier spectrum, i.e strings indexed by sets $|S| > t$, become exponentially small in $t$ for boolean functions in $\AC^0$.
 To prove the theorem, we need version of the H\aa stad's switching lemma shown in \cite{hastad}.
@@
 \end{theorem}
-From H\aa stad's switching lemma, we derive two integral corollaries:
+\noindent From H\aa stad's switching lemma, we derive two integral corollaries:
 %
@@
 %
-Corollary \ref{fouriertail} tells us that the probability of the Fourier degree exceeding $s$ decays exponentially with $s$ given that we set our probability parameter to a specific quantity $p = \frac{1}{10^ds^{d-1}}$.
-
+\noindent Corollary \ref{fouriertail} tells us that the probability of the Fourier degree exceeding $s$ decays exponentially with $s$ given that we set our probability parameter to a specific quantity $p = \frac{1}{10^ds^{d-1}}$.
+%
 \begin{proof}{\emph{(Proof of Theorem \ref{lmn})}}
@@
 Let $f_{x_V} = f(x_V,*)$ as a function $f_{x_V}: \{0,1\}^{|x_{\overline{V}}|} \rightarrow \{0,1\}$.
-We can rearrange the Fourier decomposition of any function $f$ as follows:
+Rearrange the Fourier decomposition of any function $f$ as follows:
 %
@@
   For any boolean function $f$ and $0 < p < 1$,
-  $$ \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2 \mathbb{E}_T \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2 $$
+  $$ \sum_{|S| > t} |\hat{f}(S)|^2 \leq 2 \cdot \mathbb{E}_T \sum_{ \substack {S \subseteq [n] \\ |S \cap \overline{V}| > k}} |\hat{f}(S)|^2 $$
   for $k = pt/2$
@@
 \end{proof}
-To finish the proof, set $p = \frac{1}{10t^{(d-1)/d}}$ and $k = t^{1/d}/20$.
+\noindent To finish the proof, set $p = \frac{1}{10t^{(d-1)/d}}$ and $k = t^{1/d}/20$.
 Direct calculation shows that $p \leq \frac{1}{10^dk^{d-1}}$.
@@
 %
-This gives another scheme of approximating functions in $\AC^0$ through low-degree polynomials, complementing the results of \cite{razborov}, \cite{smolensky}.
+\noindent This gives another scheme of approximating functions in $\AC^0$ through low-degree polynomials, complementing the results of \cite{razborov}, \cite{smolensky}.
 Here we make the extra distinction that our polynomials are also of low Fourier degree.
@@
 
-The influence of $f$ can be expressed in terms of its Fourier coefficients:
+\noindent The influence of $f$ can be expressed in terms of its Fourier coefficients:
 
@@
 
-By combining this equivalence with Theorem (\ref{lmn}), we deduce the following upper bound on the influence of a function in $\AC^0$
+\noindent By combining this equivalence with Theorem (\ref{lmn}), we deduce the following upper bound on the influence of a function in $\AC^0$
 
@@
 
-Lemma (\ref{influence}) shows that functions in $\AC^0$  are not suitable for constructing universal hash functions (\cite{mansour}) and pseudorandom function generators (\cite{lmn}).
+\noindent Lemma (\ref{influence}) shows that functions in $\AC^0$  are not suitable for constructing universal hash functions (\cite{mansour}) and pseudorandom function generators (\cite{lmn}).
 We give a rough definition and sketch based on the treatment found in (\cite{lmn}.
@@
 \end{lemma}
-By taking advantage of the low average sensitivity, we can simply perturb the input slightly and check if the output of $f$ changes. If it doesn't, there is a good chance that
+\noindent By taking advantage of the low average sensitivity, we can simply perturb the input slightly and check if the output of $f$ changes. If it doesn't, there is a good chance that
 it lies in $\AC^0$.
@@
 \begin{remark}{(\emph{Some concluding remarks})}
-  It is worth noting that the bounds shown by Linai-Mansour-Nisan have been improved by Tal (\cite{tal}).
-Specifically, he showed that for any $n$-ary boolean function $f$ with size $M$ and depth $d$,\
+  he bounds shown by Linai-Mansour-Nisan have been improved by Tal (\cite{tal}). Specifically, he showed that for any $n$-ary boolean function $f$ with size $M$ and depth $d$,\
   \begin{equation}
@@
   \end{equation}
+  Furthermore, he showed a converse to the proof direction employed by Linial et.
+al \cite{lmn}.
+Recall that in our proof we started with a switching lemma which bound the probability of the randomly-restricted function's Fourier degree exceeding a certain value.
+We then used this switching lemma to bound the tail Fourier spectrum.
+Tal showed a form of the opposite direction in \cite{talshrinkage}.
+  \begin{theorem}{\emph{(Tal \cite{talshrinkage})}}
+    Let $f\{1,-1\}^n \rightarrow \{-1,1\}$ be a Boolean function, let t,$C > 0$ such that for all $k$, $\sum_{|S| > k} |\hat{f}(S)|^2 \leq C e^{k/t}$ and let $\rho$ be a $p$-random restriction, then for all $d$, $Pr[\text{deg}(f_{\rho})] \leq C \cdot (4pt)^d$.
+  \end{theorem}
+  His work uses results in Quantum Query Complexity to ultimately culminate to another proof of the shrinkage exponent of Demorgan Formulae under random restrictions:
+  \begin{theorem}{\emph{(Tal \cite{talshrinkage})}}
+    Let $f$ be a Boolean function. For $p > 0$,
+    \begin{equation}
+      \mathbb{E}_{\rho \sim \mathcal{R}_p}[\mathcal{L}(f_\rho)] = \mathcal{O} \left(p^2\mathcal{L}(f) + p\sqrt{\mathcal{L}(f)} \right)
+    \end{equation}
+  \end{theorem}
 \end{remark}

### 2020-11-19 PHYS790/final.tex
First commit of PHY790 term paper.

\documentclass[12pt]{amsart}
\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-2in}
\calclayout

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{xca}[theorem]{Exercise}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\numberwithin{equation}{section}

\theoremstyle{remark}
\newtheorem{notation}[theorem]{Notation}

\begin{document}

\title{Annotations on Quantum Pseudorandomness and $k$-Designs}
\author{Edward Kim}
\date{\today}
\maketitle

\begin{abstract}
  We investigate constructions of efficient approximate unitary $k$-designs or distributions over the unitary group which model the Haar distribution up to the $k^{th}$-moment. In particular, we review the paper by Harrow and Low as a directed reading conducted during the Spring 2020 semester.
\end{abstract}

\tableofcontents

\section{Introduction}

The power of randomness in computation is a cornerstone of contemporary computation theory as evidenced by decades of beautiful, celebrated results. Quantum Information is, by no means, a stranger to this trend. Numerous applications of random unitaries have been discovered in the areas of quantum tomography, cryptography, and, rather naturally, randomized algorithms. To gain access to uniformly-sampled elements from $\mathbb{U}(d)$, one must sample from the Haar distribution over $\mathbb{U}(d)$ where $\mathbb{U}(d)$ denotes the group of unitary $d \times d$ matrices. One method of sampling from the Haar distribution is to apply local random unitaries sampled from distributions over lower-dimensional unitary groups. For example, if we consider an $n$-qubit system, one method of sampling from the Haar distribution over $\mathbb{U}(2^n)$ would be through the following scheme:
\begin{enumerate}
  \item Sample from the Haar distribution over $\mathbb{U}(4)$ to produce a random gate $G$
  \item Uniformly pick two distinct qubits and locally apply $G$ to the pair.
\end{enumerate}
However, this becomes computationally infeasible as even reasonably approximating the Haar distribution under this manner requires an exponential number of local 2-qubit gates \cite{knill}. This hurdle motivates us to study the construction of efficient $k$-designs i.e distributions over $\mathbb{U}(4)$ such that sampling over these distributions and applying our scheme above will converge to good approximations of the $k^{th}$ moments of the Haar distributon over $\mathbb{U}(2^n)$. To this end, we begin with annotations of the seminal paper by Harrow and Low \cite{harrlow}. We first introduce concepts germane to the area of efficient $k$-designs and summarize the techniques employed in the analyses presented in the paper.
nally infeasible as the number of local random unitaries grows exponentially in the number of qubits.

\section{Approximate 2-designs}

\subsection{Definitions and Motivations}
In \cite{harrlow}, the authors tackle the problem of approximate $2$-designs, specifically \emph{unitary} $2$-designs. The central result of the paper claims that the scheme of applying sampled $2$-qubit unitaries from certain distributions with a $k$-gapped property converge to the second moment of the Haar distribution rapidly. The crux of the rapid convergence argument relies on mapping the evolution of an initial state under our random circuit to that of a Markov chain's.
Consequently, the analysis becomes amenable to classic techniques created to study mixing times of Markov chains. To begin, we first state the relevant definitions as employed by Harrow and Low.

\begin{definition}
  An ensemble of operators $\mathcal{E} = \{p_i, U_i\}$ over $\mathbb{U}(d)$ is called an \emph{(exact) unitary $k$-design} if $\mathcal{G}_W = \mathcal{G}_H$ where
  \begin{gather}
    \mathcal{G}_W(\rho) = \sum_{i} p_i U_i^{\otimes k}\rho (U_i^{\dagger})^{\otimes k} \\
    \mathcal{G}_H(\rho) = \int_{\mathbb{U}(d)} U^{\otimes k} \rho (U^\dagger)^{\otimes k} \; dU
  \end{gather}
\end{definition}
%
 We would naturally like to define some distance between $\mathcal{G}_W$ and $\mathcal{G}_H$. One formulation involves interpreting the two $\mathcal{G}_W, \mathcal{G}_H$ as quantum channels to utilize the diamond norm:
%
\begin{definition}
  The diamond norm of an operator $T$ is defined as
  $$||T||_{\diamond} = \sup_d ||T \otimes I_d ||_{\infty} $$
\end{definition}
%
\begin{definition}
  $\mathcal{G}_W$ is an \emph{$\epsilon$-approximate $k$-design} if
  \begin{equation}
    ||\mathcal{G}_W - \mathcal{G}_H ||_{\diamond}  \leq \epsilon
  \end{equation}
\end{definition}
%
As discussed in the introduction above, the unitary constructed by sampling from a universal gate set and applying them to random pairs of qubits will converge to a unitary sampled from Haar distribution. However, we would like to find \emph{efficient approximate} $k$-designs. Here, efficient refers to random circuits of length polynomial in the number of qubits. The authors actually prove some stronger, namely that \emph{$k$-copy gapped} distributions suffice.
%
\begin{definition}
  Let $\mu$ be a (possibly continuous) probability measure on $\mathbb{U}(d)$. If
  \begin{equation}
    \mathcal{G}_{\mu} = \int_{\mathbb{U}(d)} d\mu(U) U^{\otimes k} \otimes (U^*)^{\otimes k}
  \end{equation}
  then we say that $\mathcal{G}_\mu$ is $k$-copy gapped if $\mathcal{G}_\mu$ has only $k!$ unit eigenvalues.
\end{definition}
%
\noindent A more general definition of universality of gatesets can be defined as well:
%
\begin{definition}
  Let $\mu$ be a distribution over $\mathbb{U}(d)$. $\mu$ is deemed to be \emph{universal} for $\mathbb{U}(d)$ if for any open ball $B \subset \mathbb{U}(d)$, there exists a sufficiently large integer $\ell$ such that $\mu^{\star \ell}(B) > 0$
  where $\mu^{\star \ell}$ refers to the $\ell$-fold \emph{convolution} product of $\mu$:
\end{definition}
%
\begin{equation}
\mu^{\star \ell} = \int_{\mathbb{U}(d)} \delta_{U_1,\cdots, U_\ell} d_\mu(U_1)\cdots d_\mu(U_\ell)
\end{equation}
%
\begin{remark}
  This is the convolution product in the sense of measures. For example, if we let $\mu,\sigma$ be two Borel measures over $\mathbb{R}$, then the Borel measure $\mu \star \sigma$:
  $$ \mu \star \sigma(A) = \int_{\mathbb{R}}\int_{\mathbb{R}} \mathbb{I}_A(x+y) d\mu(x)d\sigma(y)$$
  where $\mathbb{I}$ is the characteristic function for measurable set $A \subset \mathbb{R}$.
\end{remark}
\noindent Intuitively, universality implies that sampling enough times over $\mu$ is enough to approximate any gate in $\mathbb{U}(d)$. This aligns with the definition introduced for discrete measures $\mu$ over finite gate sets. We will prove later that any universal gate set will be $k$-copy gapped on $\mathbb{U}(4)$ (Lemma \ref{universal}). For now, let us state the main theorem and its auxiliary lemma to guide our upcoming analysis:
%
\begin{theorem} \label{maintheorem}
  Let $\mu$ be a $2$-copy gapped distribution over $\mathbb{U}(4)$ and $W$ a random circuit of length $t$ drawn from sampling $2$-qubit unitaries from $\mu$ and applying them to random pairs of of a system of $n$-qubits. Then there exists a constant $C$ depending on $\mu$ such that for $\epsilon > 0$, $\mathcal{G}_W$ is an $\epsilon$-approximate $2$-design if $t \geq n(n + \log 1/\epsilon)$.
\end{theorem}

\begin{notation}
  Let $\rho$ be an initial state such that $\rho \in \mathbb{C}^{2n}$. We can expand $\rho$ in terms of the Pauli basis $\{\sigma_i\}_{0 \leq i \leq 3}$ as follows:
  \begin{equation}
    \rho = \frac{1}{2^n} \sum_{p_1,p_2} \gamma_0(p_1,p_2) \sigma_{p_1} \otimes \sigma_{p_2}, \quad \gamma_0(p_1,p_2) \in \in \mathbb{C}
   \end{equation}
  where $p_1,p_2$ run over all strings of Pauli indices of length $n$, $\{0,1,2,3\}^n$. Let $\gamma_W(p_1,p_2)$ be the coefficient of the term $\sigma_{p_1} \otimes \sigma_{p_2}$ for $\rho_W = W^{\otimes 2} \rho (W^\dagger)^{\otimes 2}$. Extending this notation allows us to express the expected coefficient after $t$-samples from our distribution $\mu$ as
  $\gamma_t(p_1,p_2) = \mathbb{E}_W [\gamma_W(p_1,p_2)]$, averaging over all $t$-length random circuits.
\end{notation}
%
\begin{lemma}
  Let $\rho$ be a an initial state such that $\gamma_0(p,p) \geq 0$ and $\sum_{p} \gamma_0(p,p) = 1$ with $\mu$ and $W$ as in Theorem \ref{maintheorem}. Then there exists constant $C$ such that for $\epsilon > 0$

  \begin{equation}
    \sum_{p_1,p_2 \neq 00} \left( \gamma_t(p_1,p_2) - \delta_{p_1p_2} \frac{1}{2^n(2^n + 1)} \right)^2 \leq \epsilon
  \end{equation}
  for $t \geq Cn\log 1/\epsilon$.

  \begin{equation}
    \sum_{p_1,p_2 \neq 00} \left| \gamma_t(p_1,p_2) - \delta_{p_1p_2} \frac{1}{2^n(2^n + 1)} \right| \leq \epsilon
  \end{equation}
  for $t \geq n(n + \log 1/\epsilon)$
\end{lemma}

\begin{corollary}
  Let $\mu$, $W$, $\gamma_W$ be as above. Then for \emph{any} initial state $\rho$, there exists constant $C$ such that for $\epsilon > 0$,
  \begin{align}
    & \sum_{p_1,p_2 \neq 00} \left( \gamma_t(p_1,p_2) - \delta_{p_1p_2}(\frac{\sum_{p\neq 0} \gamma_0(p,p)}{4^n - 1} \right)^2 \leq \epsilon \\
    & \sum_{p_1,p_2 \neq 00} \left| \gamma_t(p_1,p_2) - \delta_{p_1p_2}(\frac{\sum_{p\neq 0} \gamma_0(p,p)}{4^n - 1} \right| \leq \epsilon
  \end{align}
  both for $t \geq Cn(n + \log 1/ \epsilon)$
\end{corollary}

\subsection{Analysis of Moments}

\newpage
\bibliographystyle{amsplain}
\bibliography{biblio}

\end{document}

### 2020-11-21 PHYS790/final.tex
Finished definitions of T(p) and MC definitions.

@@
 \calclayout
+
+\usepackage{appendix}
 
@@
 The power of randomness in computation is a cornerstone of contemporary computation theory as evidenced by decades of beautiful, celebrated results.
-Quantum Information is, by no means, a stranger to this trend.
-Numerous applications of random unitaries have been discovered in the areas of quantum tomography, cryptography, and, rather naturally, randomized algorithms.
+Quantum Information has not become stranger to this trend.
+Numerous applications of random unitaries have been discovered in the areas of quantum tomography, quantum cryptography, and, rather naturally, quantum algorithms.
 To gain access to uniformly-sampled elements from $\mathbb{U}(d)$, one must sample from the Haar distribution over $\mathbb{U}(d)$ where $\mathbb{U}(d)$ denotes the group of unitary $d \times d$ matrices.
@@
 We first introduce concepts germane to the area of efficient $k$-designs and summarize the techniques employed in the analyses presented in the paper.
-nally infeasible as the number of local random unitaries grows exponentially in the number of qubits.
 
@@
 The crux of the rapid convergence argument relies on mapping the evolution of an initial state under our random circuit to that of a Markov chain's.
-Consequently, the analysis becomes amenable to classic techniques created to study mixing times of Markov chains. To begin, we first state the relevant definitions as employed by Harrow and Low.
+Consequently, the analysis becomes amenable to classic techniques created to study mixing times of Markov chains. To begin, we first state the relevant definitions as introduced by Harrow and Low.
+%
+\begin{remark}
+  As dissecting every proof in \cite{harrlow} may unnecessarily length this report, we occasionally omit proofs of lemmas.
+Some major theorems will have their proofs delegated to the appendix of this report.
+\end{remark}
 
@@
   An ensemble of operators $\mathcal{E} = \{p_i, U_i\}$ over $\mathbb{U}(d)$ is called an \emph{(exact) unitary $k$-design} if $\mathcal{G}_W = \mathcal{G}_H$ where
-  \begin{gather}
-    \mathcal{G}_W(\rho) = \sum_{i} p_i U_i^{\otimes k}\rho (U_i^{\dagger})^{\otimes k} \\
-    \mathcal{G}_H(\rho) = \int_{\mathbb{U}(d)} U^{\otimes k} \rho (U^\dagger)^{\otimes k} \; dU
-  \end{gather}
+  \begin{align}
+    & \mathcal{G}_W(\rho) = \sum_{i} p_i U_i^{\otimes k}\rho (U_i^{\dagger})^{\otimes k} \\
+    & \mathcal{G}_H(\rho) = \int_{\mathbb{U}(d)} U^{\otimes k} \rho (U^\dagger)^{\otimes k} \; dU
+  \end{align}
 \end{definition}
@@
  We would naturally like to define some distance between $\mathcal{G}_W$ and $\mathcal{G}_H$.
-One formulation involves interpreting the two $\mathcal{G}_W, \mathcal{G}_H$ as quantum channels to utilize the diamond norm:
+One formulation involves interpreting the two $\mathcal{G}_W, \mathcal{G}_H$ as quantum channels in order to utilize the diamond norm:
 %
@@
 \begin{remark}
-  This is the convolution product in the sense of measures. For example, if we let $\mu,\sigma$ be two Borel measures over $\mathbb{R}$, then the Borel measure $\mu \star \sigma$:
+  This is the convolution product in the context of \emph{measures}.
+For example, if we let $\mu,\sigma$ be two finite Borel measures over $\mathbb{R}$, then we have a new finite Borel measure on $\mathbb{R}$, $\mu \star \sigma$:
   $$ \mu \star \sigma(A) = \int_{\mathbb{R}}\int_{\mathbb{R}} \mathbb{I}_A(x+y) d\mu(x)d\sigma(y)$$
@@
 \end{theorem}
-
+%
 \begin{notation}
-  Let $\rho$ be an initial state such that $\rho \in \mathbb{C}^{2n}$. We can expand $\rho$ in terms of the Pauli basis $\{\sigma_i\}_{0 \leq i \leq 3}$ as follows:
+  Let $\rho$ be an initial state of an $n$-qubit system i.e $\rho \in (\mathbb{C}^{2})^{\otimes n}$. We can expand $\rho$ in terms of the Pauli basis $\{\sigma_i\}_{0 \leq i \leq 3}$ as follows:
   \begin{equation}
@@
   where $p_1,p_2$ run over all strings of Pauli indices of length $n$, $\{0,1,2,3\}^n$.
-Let $\gamma_W(p_1,p_2)$ be the coefficient of the term $\sigma_{p_1} \otimes \sigma_{p_2}$ for $\rho_W = W^{\otimes 2} \rho (W^\dagger)^{\otimes 2}$.
+Let $\gamma_W(p_1,p_2)$ be the coefficient of the term $\sigma_{p_1} \otimes \sigma_{p_2}$ for the output state after applying random circuit $W$, $\rho_W = W^{\otimes 2} \rho (W^\dagger)^{\otimes 2}$.
 Extending this notation allows us to express the expected coefficient after $t$-samples from our distribution $\mu$ as
-  $\gamma_t(p_1,p_2) = \mathbb{E}_W [\gamma_W(p_1,p_2)]$, averaging over all $t$-length random circuits.
+  $\gamma_t(p_1,p_2) = \mathbb{E}_W [\gamma_W(p_1,p_2)]$. Note that the expectation is taken over all $t$-length random circuits.
 \end{notation}
@@
 
+We turn back to the general case and begin to define concepts and definitions for group $\mathbb{U}(d)$ and for $d$-qudit states.
+Set a Hermitian basis of $d \times d$ to be $\sigma_0,\sigma_1,\cdots \sigma_{d^2 -1}$ such that $\sigma_0 = I$.
+These basis elements span the $d$-qudit states.
+Here we stress that we sample our unitaries over the \emph{Haar distribution} over $\mathbb{U}(d)$.
+We denote sampling from the Haar distribution with the notation $U \sim \mathbb{U}(d)$ when necessary.
+Furthermore, our Hermitian basis operators are such that $Tr(\sigma_p\sigma_q) = d \cdot \delta_{p,q}$ i,e orthogonal under the Hilbert-Schmidt inner product.
+\newline
+
+\noindent Let ${\bf p} = (p_1, \cdots, p_k) \in \{0,1,\cdots, d^2 -1\}^k$ denote a $k$-string of basis indices. We will be interested in the quantity:
+%
+\begin{equation}
+  T({\bf p}) := \mathbb{E}_{U \sim \mathbb{U}(d)} \left[ U^{\otimes k} \sigma_{p_1} \otimes \sigma_{p_2} \otimes \cdots \otimes \sigma_{p_k} (U^\dagger)^{\otimes k} \right]
+\end{equation}
+This will be the expected state after applying a sampled unitary $U$ from the uniform distribution over $\mathbb{U}(d)$. Naturally, we can reexpress this resulting state into the $\sigma_p$ basis:
+%
+\begin{equation}
+  T({\bf p}) = \sum_{{\bf q}} \widehat{G}(\textbf{p},\textbf{q}) \sigma_{q_1} \otimes \cdots \otimes \sigma_{q_k}
+\end{equation}
+where the same is taken over all $k$-strings in $\{0, \cdots, d^2-1\}^k$.
+This allows us to define the matrix $\widehat{G}$ where the rows and columns are indiced by $k$-strings with entries $\widehat{G}(\textbf{q},\textbf{p})$.
+This matrix turns out to be a \emph{symmetric projector}.
+
+\begin{lemma}
+  The matrix $\widehat{G}$ above is symmetric i.e. $\widehat{G}(\textbf{p},\textbf{q}) = \widehat{G}(\textbf{q},\textbf{p})$
+\end{lemma}
+%
+\begin{proof}
+  By orthogonality:
+  \begin{align}
+    \widehat{G}(\textbf{q},\textbf{p})
+    & = d^{-k} \cdot Tr(\sigma_{q_1} \otimes \cdots \otimes \sigma_{q_k} T(\textbf{p}) ) \\
+    & = d^{-k} \mathbb{E}_{U}\left[ Tr(\sigma_{q_1} \otimes \cdots \otimes \sigma_{q_k}) U^{\otimes k} (\sigma_{p_1} \otimes \sigma_{p_2} \otimes \cdots \otimes \sigma_{p_k}) (U^\dagger)^{\otimes k}\right]
+  \end{align}
+\end{proof}
+%
+Note that the expectation and the trace operator commute by linearity of latter. The result follows from the cyclic invariance of trace.
+%
+\begin{lemma} \label{permfix}
+$\widehat{G}$ fixes permutation operators $P_{\pi}, \; \pi \in S_k$.
+\end{lemma}
+%
+These two results are all that is required to show that $\widehat{G}$ is indeed a projector.
+%
+\begin{theorem} \label{g-proj}
+  $\widehat{G}$ is a projector i.e. $\widehat{G}^2 = \widehat{G}$
+\end{theorem}
+%
+\begin{proof}
+  Expanding the matrix product shows that
+  \begin{align}
+     \widehat{G}(\textbf{p},\textbf{q})
+     & = \sum_{\textbf{q'}} \widehat{G}(\textbf{p},\textbf{q'})\widehat{G}(\textbf{q'},\textbf{q}) = \\
+     & = \sum_{\textbf{q'}} \widehat{G}(\textbf{p},\textbf{q'})
+          \left( d^{-k} \cdot Tr(\sigma_{q_1'} \otimes \cdots \otimes \sigma_{q_k'}T(\textbf{q})) \right)
+  \end{align}
+  By Schur-Weyl Duality, $T(\text{q})$ must be a linear combination of permutation operators ($T(\textbf{p})$ commutes with all unitaries of the form $U^{\otimes k}$ by the invariance of the Haar integral).
+However, by Lemma (\ref{permfix}) above, $\widehat{G}$ fixes the permuatation operators.
+To understand this more explicitly, rewrite the Lemmma (\ref{permfix}) as follows:
+  \begin{equation}
+    \sum_{\textbf{q}} \widehat{G}(\textbf{p},\textbf{q})\cdot Tr(\sigma_{q_1} \otimes \cdots \otimes \sigma_{q_k} P_\pi) = Tr(\sigma_{p_1} \otimes \cdots \otimes \sigma_{p_k}P_\pi)
+  \end{equation}
+  This simply expressing the equality $\widehat{G}P_\pi = P_\pi$ for permutation operator $p_\pi, \; \pi \in S_{k}$. By linearity of trace, the equality above reduces to
+  \begin{align}
+    \sum_{\textbf{q'}} \widehat{G}(\textbf{p},\textbf{q'})
+         \left( d^{-k} \cdot Tr(\sigma_{q_1'} \otimes \cdots \otimes \sigma_{q_k'}T(\textbf{q})) \right)
+         & = Tr(\sigma_{p_1} \otimes \cdots \otimes \sigma_{p_k} T(\text{q})) \\
+         & = \widehat{G}(\textbf{p},\textbf{q})
+  \end{align}
+\end{proof}
+
+\begin{corollary}
+$\widehat{G}$ has eigenvalues 0 and 1
+\end{corollary}
+
+\subsubsection{Explicit calculation of $T(\textbf{p})$}
+The calculation is divided into the two pertinent cases $k=1,2$. \newline
+
+\begin{enumerate}
+  \item \emph{(For $k = 1$)} For this case, a random unitary sampled from the uniform distribution will randomize the initial state. Thus:
+  $$
+  T(p) = \begin{cases}
+    \sigma_0 & p = 0 \\
+    0 & p \neq 0
+  \end{cases}
+  $$
+  \item \emph{(For $k=2$)} From Lemma (\ref{permfix}) and Theorem (\ref{g-proj}), $\widehat{G}$ must fix the only two permutation operators $I, \mathcal{F}$ where $\mathcal{F}$ is the swap operator on two qubits.
+We note that we will normalize the swap operator
+  %
+  \begin{equation}
+    \mathcal{F} = \frac{1}{d^2 -1 }\sum_{1 \leq q \leq d^2 -1} \sigma_{q} \otimes \sigma_{q}
+  \end{equation}
+  %
+  We then deduce the following about $\widehat{G}$: \newline
+    \begin{enumerate}
+      \item $\widehat{G}(p_1,p_2;q_1,q_2) = 0, \quad p_1 \neq p_2$ or $q_1 \neq q_2$
+      \begin{proof}
+        To see this, observe that both $I, \mathcal{F}$ have vanishing $\sigma_{q_1} \otimes \sigma_{q_2}$ components for $q_1 \neq q_2$.
+Thus, multiplying $\widehat{G}$ with either one should also result in vanishing $\sigma_{q_1} \otimes \sigma_{q_2}$ components.
+This with the fact that all other operators are sent to zero by $\widehat{G}$ yields the result.
+      \end{proof}
+
+      \item $\widehat{G}(p;0) = \delta_{p0}$ (Here we denote $\widehat{G}(p;0) := \widehat{G}(p,p;0,0)$)
+      \begin{proof}
+        This just derives from $\widehat{G}$ acting on $I$
+      \end{proof}
+
+      \item $\widehat{G}(p,a) = \frac{1}{d^2-1}, \quad a,p \neq 0$
+      \begin{proof}
+        As above, this results from $\widehat{G}$ acting on $\mathcal{F}$ along with the condition that it must send all other operators to zero.
+      \end{proof}
+    \end{enumerate}
+    We conclude from the definition of $T(p,q)$ for $p,q \in \{0,\cdots,d^2-1\}$ that
+    %
+    \begin{equation}
+      T(p,q) = \begin{cases}
+        0 \quad & p \neq q \\
+        \sigma_0 \otimes \sigma_0 & p = q = 0 \\
+        \frac{1}{d^2 - 1}\sum_{\ell \neq 0}  \sigma_\ell \otimes \sigma_\ell & p = q \neq 0
+      \end{cases}
+    \end{equation}
+\end{enumerate}
+%
+Recall that $T(p,q)$ refers to the expected \emph{state} from applying unitary $U \sim \mathbb{U}(d)$ to the basis state $\sigma_p \otimes \sigma_q$.
+From the determination above, it is simple to see that $U$ uniformly samples from all diagonal basis states when acting on $\sigma_p \otimes \sigma_p, \; p \neq 0$.
+Strikingly enough, applying $U \sim \mathbb{U}(d)$ conserves the total sum of the coefficients of the diagonal basis states.
+%
+\begin{lemma}
+  The total sum of the coefficients $\sum_{p} \gamma_0(p,p)$ is conserved by $U \sim \mathbb{U}(d)$ for state $\rho = \frac{1}{d}\sum_{p_1,p_2} \gamma_0(p_1,p_2) \sigma_{p_1} \otimes \sigma_{p_2}$.
+\end{lemma}
+%
+Normalizing this sum allows us to view it as a \emph{probability distribution}.
+Furthermore, the evolution of our random circuit scheme is amenable to Markov chain analysis.
+As an illustrative example, suppose we have an $n$-qubit system and so far have appied $t$ steps of our scheme.
+As per usual, we pick our $2$-qubit unitary $U \sim \mathbb{U}(4)$.
+Now if we were to pick qubits $1,2$ for our local unitary transformation, the expected \emph{coefficients} for step $t+1$ will work out to be
+%
+\begin{align}
+    \gamma_{t+1}(p_1,\cdots,p_n,q_1,\cdots,q_n) = \begin{cases}
+       0 & (p_1 ,p_2) \neq (q_1, q_2) \\
+       \gamma_t(0,0,\cdots, p_n, 0,0,\cdots,q_n ) & (p_1,p_2) = (q_1, q_2) = (0,0) \\
+       \frac{1}{15}\sum_{\substack{r_1,r_2 \\ r_1r_2 \neq 00}} \gamma_t(r_1,r_2,\cdots p_n,r_1,r_2,\cdots,q_n) & (p_1,p_2) = (q_1,q_2) \neq (0,0)
+    \end{cases}
+\end{align}
+It is worth reiterating that our random $2$-qubit unitary is only acting on the first two qubits non-trivially whereas every other qubit is acted on by identity.
+Our desired Markov chain is then defined naturally as such \newline
+%
+\begin{enumerate}
+  \item The states of the Markov chain are $\Omega = \{0,1,2,3\}^n$
+  \item The transitions adhere to the guidelines below: Suppose we start with a state indexed by string $\alpha = \alpha_1\cdots\alpha_n$
+    \begin{enumerate}
+      \item Pick two distinct sites $i < j$ uniformly at random.
+      \item If $i = j = 0$, remain at the same state. Otherwise pick $\beta_i\beta_j = \{0,1,2,3\}^2/\{00\}$ uniformly at random. Transition to the string
+      $$\alpha_1\alpha_2,\cdots\alpha_{i-1}\beta_i\alpha_{i+1}\cdots \alpha_{j-1}\beta_{j}\alpha_{j+1}\cdots\alpha_n$$
+    \end{enumerate}
+\end{enumerate}
+
+\subsection{Convergence of the Random Circuit}
+
 \newpage

### 2020-11-22 PHYS790/final.tex
Finished half of MC section.

@@
 The power of randomness in computation is a cornerstone of contemporary computation theory as evidenced by decades of beautiful, celebrated results.
-Quantum Information has not become stranger to this trend.
+Quantum Information has, by no means, become stranger to this trend.
 Numerous applications of random unitaries have been discovered in the areas of quantum tomography, quantum cryptography, and, rather naturally, quantum algorithms.
@@
 %
-\begin{definition}
+\begin{definition} \label{gappeddef}
   Let $\mu$ be a (possibly continuous) probability measure on $\mathbb{U}(d)$. If
@@
 %
-\begin{definition}
+\begin{definition} \label{univdef}
   Let $\mu$ be a distribution over $\mathbb{U}(d)$.
@@
 \end{notation}
-%
-\begin{lemma}
+
+\begin{lemma} \label{mainlemma}
   Let $\rho$ be a an initial state such that $\gamma_0(p,p) \geq 0$ and $\sum_{p} \gamma_0(p,p) = 1$ with $\mu$ and $W$ as in Theorem \ref{maintheorem}.
@@
 
-\begin{corollary}
+\begin{corollary} \label{maincorollary}
   Let $\mu$, $W$, $\gamma_W$ be as above. Then for \emph{any} initial state $\rho$, there exists constant $C$ such that for $\epsilon > 0$,
@@
 %
-\begin{equation}
+\begin{equation} \label{Tdef}
   T({\bf p}) := \mathbb{E}_{U \sim \mathbb{U}(d)} \left[ U^{\otimes k} \sigma_{p_1} \otimes \sigma_{p_2} \otimes \cdots \otimes \sigma_{p_k} (U^\dagger)^{\otimes k} \right]
@@
 
-\subsubsection{Explicit calculation of $T(\textbf{p})$}
+\subsubsection{Explicit calculation of $T(\textbf{p})$} \label{eigenanalysis}
 The calculation is divided into the two pertinent cases $k=1,2$. \newline
@@
 Furthermore, the evolution of our random circuit scheme is amenable to Markov chain analysis.
-As an illustrative example, suppose we have an $n$-qubit system and so far have appied $t$ steps of our scheme.
+As an illustrative example, suppose we have an $n$-qubit system and so far have applied $t$ steps of our scheme.
 As per usual, we pick our $2$-qubit unitary $U \sim \mathbb{U}(4)$.
-Now if we were to pick qubits $1,2$ for our local unitary transformation, the expected \emph{coefficients} for step $t+1$ will work out to be
+Now if we were to pick qubits $1,2$, our random $2$-qubit unitary would only act on the first two qubits non-trivially whereas every other qubit would just transform via the identity map.
+The expected \emph{coefficients} for step $t+1$ will work out to be
 %
@@
 \end{align}
-It is worth reiterating that our random $2$-qubit unitary is only acting on the first two qubits non-trivially whereas every other qubit is acted on by identity.
-Our desired Markov chain is then defined naturally as such \newline
+ Oliveira et.al \cite{oliveira} had the insight to map this evolution to a Markov chain as follows: \newline
+%
+\begin{theorem} \label{haarevol}
+  Let $M$ be a Markov chain governing the evolution of states under $\widehat{G}$ where $\widehat{G}$ samples from the Haar distribution over $\mathbb{U}(2^n)$.
+Then $M$ exhibits the following transitions:
   %
@@
   \end{enumerate}
+\end{theorem}
 
 \subsection{Convergence of the Random Circuit}
+Before embarking on the Markov chain analysis, we must prove that our random circuit indeed converges to Haar distribution.
+This will be proven for the general case of a universal distribution (Definition (\ref{univdef})) on $\mathbb{U}(d)$.
+Specifcally, we will prove that any universal distribution on $\mathbb{U}(d)$ will be $k$-copy gapped (Definition (\ref{gappeddef})).
+In turn, this leads to convergence in sense of $\lim_{t \rightarrow \infty} (\widehat{G}_{\mu})^{t} = \widehat{G}_{\mathbb{U}(d)}$.
+To see this, observe that another formulation of a $k$-copy gapped distributon over $\mathbb{U}(d)$ is
+\begin{equation} \label{gappeddis}
+  ||G_{\mu} - G_{\mathbb{U}(d)}||_\infty < 1
+\end{equation}
+where $G_{\mu} = \mathbb{E}_{U \sim \mu} \left[ U^{\otimes k} \otimes (U^\dagger)^{\otimes k} \right]$.
+To justify this, first recall that $\widehat{G}$ and $G$ are unitarily-related by Definition \ref{Tdef}, so in proving \ref{gappeddis}, we can argue on the operators $\widehat{G}$ instead.
+In view of this, $\widehat{G}_{\mathbb{U}(d)}$ is a projector and thus has eigenvalues 0 and 1.
+If we can prove that any eigenvector of $\widehat{G}_{\mathbb{U}(d)}$ of eigenvalue one is also an eigenvector of $\widehat{G}_{\mu}$ of eigenvalue one i.e the $k$-symmetric subspace of $(\mathbb{C}^d)^{\otimes k}$ is also fixed by $\widehat{G}_{\mu}$.
+Thus, Equation $\ref{gappeddis}$ shows that indeed we can converge to $G_{\mathbb{U}(d)}$ by iterating our scheme (as denoted by $(\widehat{G}_\mu)^t$).
+Note that $\widehat{G}_{\mu}$ may not even be Hermitian, so we cannot directly argue on its spectrum (as the spectrum may not completely contain real values).
+%
+\begin{theorem} \label{convergence}
+  Let $\mu$ be a universal distribution over $\mathbb{U}(d)$.
+Then all eigenvectors of $G_{\mathbb{U}(d)}$ of eigenvalue one will also be eigenvectors of $G_{\mu}$ of eigenvalue one.
+Furthermore, $\mu$ will be $k$-copy gapped.
+\end{theorem}
+%
+\noindent For the sake of presentation, the proof of this will be delegated towards the appendix of this report.
+%
+\subsection{Markov Chain Analysis}
+As discussed in a previous section, the central idea of proving \emph{rapid} convergence involves mapping the evolution of the inital state under our random circuit to that of a Markov chain's.
+In fact, we can explicitly determine the operator $P$ capturing our dynamics.
+Let $\widehat{G}^{(ij)}$ denote the unitary transformation resulting from sampling some unitary $U \sim \mathbb{U}(d)$ and applying $U$ to qubits $i$ and $j$ while leaving all other qubits undisturbed.
+%
+\begin{equation} \label{schemechain}
+P = \frac{1}{n(n-1)}\sum_{i \neq j} \widehat{G}^{(ij)}
+\end{equation}
+as evidenced by the equality
+\begin{equation}
+  \gamma_{t+1}(\textbf{p}) = \frac{1}{n(n-1)}\sum_{i \neq j}\sum_{\textbf{q}} \widehat{G}^{(ij)}(\textbf{p},\textbf{q})\gamma_t(\textbf{q})
+\end{equation}
+where $\textbf{p},\textbf{q} \in \{0,1,2,3\}^2$ and $\gamma_t(\textbf{p})$ represents the expected coeffcients at timestep $t$.
+%
+A major caveat of interpreting $P$ as a \emph{transition matrix} for some suitable Markov chain on the diagonal states $\sigma_p \otimes \sigma_p$ is the possibility that our initial coefficients of diagonal states $\gamma_0(p,p)$ may not even sum to one or even be non-negative.
+It turns out that this doesn't bring too much trouble into the analysis as pointed out by Corollary \ref{maincorollary}.
+For now, let us tackle the case where the two conditions above are true and direct our focus on the evolution of the diagonal states.
+Of course, we need to also show that the expected coeffcients for the non-diagonal states $\gamma_0(p,q)$ vanish rapidly as well.
+These matters will be addressed in the upcoming section.
+
+\subsubsection{Preliminaries on Markov Chains}
+Let us introduce some defintions of properties of the Markov chains we are interested in.
+Although we assume familiarity with the definition of a Markov chain along with the its associated transition matrix $P$, we refer the reader to the reference \cite{cover} for greater detail.
+Recall that if $q^{(0)}$ is some initial  distribution over the $\Omega$ then $q^{(t)} = q^{(0)}P^t$.
+The \emph{stationary distribution} of a Markov chain $\pi$ is a probability distribution over $\Omega$ such that $\pi P = \pi$.
+%
+\begin{definition}
+  A Markov chain is said to be \emph{irreducible} if there exists a path $s \rightsquigarrow j$ for every pair of states $s,j \in \Omega$.
+\end{definition}
+%
+\begin{definition}
+  A Markov chain is said to be \emph{periodic} at state $i \in \Omega$ if there exists a positive integer $T > 1$ such for that some initial distribution $q^(0)$, if $q_i^{(t)} > 0$ then $t \in \{a + Tm \mid m \in \mathbb{N}\}$.
+A Markov chain is deemed as \emph{aperiodic} if none of its states are periodic.
+Equivaently, a Markov chain is aperiodic if all of its states contain self-loops.
+\end{definition}
+%
+\begin{definition}
+  A Markov chain is \emph{ergodic} if it is aperiodic and irreducible.
+\end{definition}
+\noindent As the below theorem reveals, ergodicity shows that any initial distribution over $\Omega$ converges to a unique stationary distribution.
+%
+\begin{theorem}
+  If a Markov chain is ergodic, then it converges to a unique stationary distributon $\pi$ i.e $P$ has a unique eigenvector with eigenvalue one.
+\end{theorem}
+%
+\noindent One can see this through a special case of the \emph{Perron-Frobenius Theorem for Regular Matrices}
+%
+\begin{theorem}
+  Suppose that $P$ is transition matrix with positive entries and also that $P_{ij}^{k} > 0$ for some sufficiently large $k$ and for all entries $i,j$ (Regularity). Then the following must hold:
+  \begin{enumerate}
+    \item The largest eigenvalue must be real and postive
+    \item The largest left eigenvector $\pi$ must be positive and unique
+    \item $\lim_{k \rightarrow \infty} qP^{(k)} = \pi$ for all initial distributions $q$.
+  \end{enumerate}
+\end{theorem}
+\noindent Since our matrices are stochastic i.e $\sum_{j} P_{ij} = 1$ for all $i$, our largest eigenvalue will be one.
+Furthermore, the left eigenvector of eigenvalue one will be the stationary distribution $\pi$.
+Since our Markov chain is both irreducible and aperiodic, it must be regular.
+%Mention counterexamples if time remains
+%
+\begin{definition}
+A Marko chain is \emph{reversible} if the following equality holds for all states $i,j \in \Omega$
+\begin{equation}
+  \pi(i)P_{ij} = \pi(j)P_{ji}
+\end{equation}
+\end{definition}
+\noindent If the stationary distributon is uniform, then it is simple to see that reversability is equivalent to $P$ being symmetric.
+Reversability is necessary to invoke certain mixing time lemmas which are introduced in the upcoming section.
+
+\subsubsection{Mixing Times}
+We introduce the definition of mixing times presented in \cite{harrlow}. For a deeper treatment of the properties below, see \cite{montenegro}.
+%
+\begin{definition}
+  Let $\pi$ be the stationary distribution of ergodic chain $P$. The 1-norm mixing time $\tau:[0,1] \rightarrow \mathbb{R}$ is real-valued function defined as
+  \begin{equation}
+    \tau(\epsilon) = \max_q \min_{t \geq 0} ||qP^t - \pi|| \leq \epsilon
+  \end{equation}
+  where $||p-q|| = \frac{1}{2}\sum_{i \in \Omega} |p_i - q_i|$
+\end{definition}
+%
+\noindent  There is also an associated 2-norm version as well:
+%
+\begin{definition}
+  Let $\pi$ be the stationary distribution of ergodic chain $P$. The 2-norm mixing time $\tau:[0,1] \rightarrow \mathbb{R}$ is real-valued function defined as
+  \begin{equation}
+    \tau(\epsilon) = \max_q \min_{t \geq 0} ||qP^t - \pi||_2 \leq \epsilon
+  \end{equation}
+\end{definition}
+%
+\noindent These definitions allow us to measure the length of time required for a random walk on an ergodic Markov chain to be within $\epsilon$ from the stationary distribution.
+Consquently, we can express a deep integral inequality bounding the mixing time of an ergodic, reversible chain using its \emph{spectral gap}.
+%
+\begin{theorem}[Refer to \cite{montenegro}, Corollary 1.15] \label{spectralgap}
+  \begin{equation}
+    \tau(\epsilon) \leq \frac{1}{\Delta}\log\frac{1}{\pi_*\epsilon}
+  \end{equation}
+  where $\pi_* = \min_{x \in \Omega} \pi(x)$ and $\Delta = \min{1 - \lambda_2, 1 + \lambda_{min}}$ Here $\lambda_2$ refers to the second largest eigenvalue of $P$ and $\lambda_{min}$ refers to the smallest.
+This $\Delta$ is known as the spectral gap.
+\end{theorem}
+\noindent For irreversible chains, the same holds for $PP^*$
+%
+\begin{theorem}[(Refer to \cite{montenegro}), Corollary 1.14]
+  \begin{equation}
+    \tau(\epsilon) \leq \frac{2}{\Delta_{PP^*}}\log\frac{1}{\pi_*\epsilon}
+  \end{equation}
+\end{theorem}
+\noindent The inequality can be converted for the 2-norm as well:
+%
+\begin{equation} \label{2spectralgap}
+  \tau_2(\epsilon) \leq \frac{1}{\Delta}\log(1/\epsilon)
+\end{equation}
+
+Finally, one more significant tool is required to lower bound the spectral gap of our matrix $P$ in Equation \ref{schemechain}.
+It is worth stressing that the Markov chain associated to $P_{\mu}$ which we will define as
+%
+\begin{equation}
+P_\mu - \frac{1}{n(n-1)}\sum_{i\neq j}\widehat{G_\mu}^{(ij)}
+\end{equation}
+For this chain, we sample from a $2$-copy gapped distribution $\mu $ over $\mathbb{U}(4)$ rather than over Haar distribution.
+It is unclear exactly how this chain evolves as we cannot explicitly compute $P_\mu$ as we did for $P_{\mathbb{U}(d)}$.
+However, with the \emph{Comparision Theorem}, we can at least lower bound the spectral gap associated with $P_\mu$, say $\Delta_\mu$, with $\Delta_{\mathbb{U}(d)}$.
+We first state a more general version of the theorem
+%
+\begin{theorem}[Refer to \cite{montenegro}, Theorem 2.3] \label{comptheorem}
+  Let $P$ and $\widehat{P}$ be two Markov chains with the same state space $\Omega$ and the same stationary distribution $\pi$.
+Then, for ever $x \neq y \in \Omega$ with $\widehat{P}_{xy} > 0$, construct a set of directed paths $\mathcal{P}_{xy}$ from $x$ to $y$ along edges in $P$.
+Let $\mathcal{P} = \bigsqcup_{x\neq y} \mathcal{P}_{xy}$.
+We define the flow function $f: \mathcal{P} \rightarrow [0,1]$ such that
+  \begin{equation}
+    \sum_{\gamma_{xy} \in \mathcal{P}_{xy}} f(\gamma_{xy}) = \widehat{P}_{xy}
+  \end{equation}
+  then
+  $$ \Delta \geq \widehat{\Delta}/A $$
+  where
+  $$ A = \max_{a \neq b, P_{ab} > 0}\frac{1}{\pi_aP_{ab}} \sum_{x \neq y}\sum_{
+                          \substack {
+                          \gamma_{xy} \in \mathcal{P}_{xy} \\
+                          (a,b) \in \gamma_{xy}}}  \pi(x)f(\gamma_{xy})|\gamma_{xy}|$$
+  where $|\gamma_{xy}|$ refers to the length of the path $\gamma_{xy}$.
+\end{theorem}
+\noindent Note that the bounds yielded will depend on the paths chosen for all $x \neq y \in \Omega$ such that $\widehat{P}_{xy} > 0$ and the values assigned to those paths by the constructed flow function $f$.
+\newline
+
+By Theorem \ref{convergence}, both $P_{\mu}$ and $P_{\mathbb{U}(4)}$ converge to same stationary distribution and both share the same state space.
+Using Theorem \ref{comptheorem}, we can bound the spectral gap of $P_\mu$, $\Delta_\mu$ with the with spectral gap of the $P_{\mathbb{U}(4)}$ chain, $\Delta_{\mathbb{U}(4)}$.
+However, we must first ensure that the Markov chain associated $P_\mu$ is ergodic.
+To achieve this, we simply delete the state $\sigma_0 \otimes \sigma_0$ from the state space.
+This will turn $P$ into an irreducible chain as $\sigma_0 \otimes \sigma_0$ the only isolated component in the chain.
+This chain has self-loops by Theorem \ref{haarevol} and thus, it must be ergodic.
+This puts us in the position to state the lower bound
+
+%Mention Hermitian conjugation at some point.
+
+\begin{theorem}
+  Let $\mu$ be any 2-copy gapped distribution of gates. If $\mu$ is invariant under Hermitian conjugation then let $\Delta_\mu$ be the spectral gap of resulting Markov chain $P_\mu$. Then
+  \begin{equation}
+    \Delta_\mu = \Omega(\Delta_{\mathbb{U}(4)})
+  \end{equation}
+  If $\mu$ is not invariant under Hermitian conjugation, then let $\Delta_{PP^*}$ be the spectral gap of the chain $PP^*$. Then
+  \begin{equation}
+    \Delta_{PP^*} = \Omega(\Delta_{\mathbb{U}(4)})
+  \end{equation}
+\end{theorem}
+\begin{proof}
+  We only show the symmetric case for now. \newline
+
+  First, let us start with an simple example on a 4-qubit system to demonstrate the construction of the relevant paths and flow function.
+Consider the valid transiton in $P_{\mathbb{U}(4)}$, $1200 \rightsquigarrow 1300$ accomplished by picking sites $1,2$ and transitioning to $1,3$.
+This shows that $P_{\mathbb{U}(4)}(1200,1300) > 0$.
+An immediate obstacle arises by observing that this may not a valid transition in the chain $P_\mu$.
+However, from Section \ref{eigenanalysis}, we know that $\hat{G}_{\mu}$ acts transitively on this subspace of two qubits $\mathbb{C}^4 \otimes \mathbb{C}^4$.
+Hence, no cycles can exist during the evolution i.e there exists no path $12 \rightsquigarrow \cdots \rightsquigarrow 12$ with length $< 14$.
+Consequently, only a constant number of transitions have to occur from $12 \rightsquigarrow 13$ and the length of the path cannot be more than $14$.
+Also, there exists two more possbile transition paths from $1200 \rightsquigarrow 1300$, namely the ones which pick sites $2,3$ and sites $2,4$.
+The Markov chain just has to transition from $20 \rightsquigarrow 30$ for both cases.
+This gives us \emph{three} paths from $1200 \rightsquigarrow 1300$ which motivates us to construct our flow function to assign $\frac{1}{3}$ to each path.
+The procedure for the general case follows similarly: \newline
+
+  We begin by noting that there two cases to consider.
+Let $h(x,y)$ denote the \emph{Hamming distance} between strings $x,y \in \{0,1,2,3\}^n$ defined to be $h(x,y) = |\{i \in [n] \mid x_i \neq y_i\}|$.
+There only two cases where $P_{\mathbb{U}(4)}(x,y) > 0$: \newline
+  %
+  \begin{enumerate}
+    \item (\emph{h(x,y) = 2}).
+Let $i,j$ be the unique indices such that $x_i \neq y_j$.
+By the discussion above, there exists a constant length path in $P_\mu$, $x \rightsquigarrow y$ acting on sites $i,j$.
+    \item (\emph{h(x,y) = 1}).
+Let $i$ be the index where $x_i \neq y_i$.
+There are two other subcases in this instance.
+If the inital state $x_i \neq 0$, then the chain choose any other of the $n-1$ sites to transiton to $y$.
+To see this, view the desired transition as
+    $$x_1 \cdots x_i \cdots x_n \rightsquigarrow y_1 \cdots y_i \cdots y_n$$
+    For any $x_j, j < i$ (w.l.o.g), we can transition by $x_jx_i \rightsquigarrow x_jy_i = y_jy_i$.
+Finally, there is the case when $x_i = 0$.
+This precludes us from choosing any site containing the identity state $(x_j = 0)$ as $00 \rightsquigarrow 00$ only.
+This shows us that we can use $n - z(x)$ sites where $z(x)$ denotes the number of zero sites in string $x$.
+  \end{enumerate}
+  We can now define the flow function $f: \mathcal{P} \rightarrow [0,1]$ through the observations above:
+  \begin{enumerate}
+    \item (\emph{h(x,y) = 2}) There is only one path $\sigma_{xy}$ in this case. So set $f(\sigma_{xy}) = P_{\mathbb{U}(4)}(x,y) = \Theta(1/n^2)$ by Theorem \ref{haarevol}.
+    \item (\emph{h(x,y) = 1}) For the case of non-zero initial value, there are $n-1$ constant-length paths.
+Equally distributing the weight across all paths shows that $f(\gamma_{xy}) = \frac{P_{\mathbb{U}(4)}(x,y)}{n-1} = \Theta(1/n^2)$ for $\gamma_{xy} \in \mathcal{P}_{xy}$ as $P_{\mathbb{U}(4)}(x,y) = \Theta(1/n)$.
+For the case of zero initial value: $f(\gamma_{xy}) = \frac{P_{\mathbb{U}(4)}(x,y)}{n-z(x)} = \Theta(1/n^2)$ since ${P_{\mathbb{U}}(4)}(x,y) = \Theta(\frac{n - z(x)}{n^2})$.
+  \end{enumerate}
+  All that remains is to calculate the number of edges $(a,b)$ such that $P_\mu(a,b) > 0$ and $(a,b) \in \gamma_{xy}$ for some path $\gamma_{xy} \in \mathcal{P}_{xy}$.
+First, we reduce the expression of the constant $A$ found in Theorem \ref{comptheorem}:
+  \begin{align} \label{simplified}
+    & A = \max_{a \neq b, \; P_\mu(a,b) > 0} A(a,b) \\
+    & A(a,b) = \frac{1}{P(a,b)}\sum_{\substack{\gamma_{xy} \in \sum_{x \neq y}\mathcal{P}_{xy} \\
+    (a,b) \in \gamma_{xy}}} f(\gamma_{xy})
+  \end{align}
+  This arises as $\pi$ is just the uniform distribution and the path lengths $|\gamma_{xy}|$ will always be constant.
+  \begin{enumerate}
+    \item (\emph{h(a,b) = 2}) Let $i,j$ be the two differing sites.
+Recall that our paths pick two sites and evolve accordingly.
+In other words, no path picks two sites and changes sites in the middle of the path (no path evolves like $1200 \rightsquigarrow 1010 \rightsquigarrow 1300$).
+Furthermore, any such $x,y$ must have $ h(x,y) = 1,2$ in order for it's path $\gamma_{xy}$ to use edge $(a,b)$.
+There are only a constant number of such pairs $x,y$ which fit the above criterion.
+Additionally, by the $h(x,y) =2$ case, there is a unique path which \emph{only uses} sites $i,j$ to transition from $x \rightsquigarrow y$.
+To see this, the case where $h(x,y) = 2$ is simple as there is only one path by construction above.
+If $h(x,y) = 1$ and (w.l.o.g) only differ at site $i$, then there is only one path from $x$ which picks the exact pair $i,j$ and transitions to $y$.
+Hence, there are is overall a constant number of paths which can use $(a,b)$.
+The sum in \ref{simplified} will reduce to $\frac{1}{\Theta(1/n^2)} \cdot \Theta(1/n^2) = \Theta(1)$ as $P_\mu(a,b) = \Theta(1/n^2)$.
+    \item (\emph{h(a,b) = 1}) Let $r$ be the number of allowed transitions between $a,b$ (not all pairs may yield direct transitions between $a,b$).
+We can reduce to the case above for every pair, showing that the sum cumulates to $\Theta(r/n^2)$.
+As $P_\mu(a,b) = \Theta(r/n^2)$, this once again shows that $A = \Theta(1)$.
+  \end{enumerate}
+\end{proof}
+
+A significant portion of \cite{harrlow} is dedicated to proving that $\Delta_{\mathbb{U}(4)} = \Theta(1/n)$.
+Invoking Lemma \ref{spectralgap} and \ref{2spectralgap} along with this tight bound on the spectral gap of the $\mathbb{U}(4)$ chain shows that
+%
+\begin{corollary}
+  The chain $P_\mu$ has 1-norm mixing time $\mathcal{O}(n(n+\log1/\epsilon))$ and 2-norm $\mathcal{O}(n\log\frac{n}{\epsilon})$
+\end{corollary}
+
+\subsection{Tight Analysis of $\mathbb{U}(4)$ gap}
+In the last section, we used the Comparision Theorem (Theorem \ref{comptheorem}) to bound the gap of the $P_\mu$ chain where $\mu$ is some 2-copy gapped distribution.
+The rest of this section will be devoted to proving tight bounds on the spectral gap of the $\mathbb{U}(4)$ chain, $\Delta_{\mathbb{U}(4)}$.
+We reiterate that the Markov chain transitions will be governed as described in Theorem \ref{haarevol}.
+\newline
+
+For this case, we will just sketch the overarching ideas of the section and leave some details to the appendix of this report and \cite{harrlow}.
+
+%\subsection{Tying Some Loose Ends}
 

### 2020-11-22 PHYS790/final.tex
Finished most of MC portion.

@@
 \usepackage{appendix}
+\usepackage{tikz}
+\usetikzlibrary{automata, positioning}
 
@@
 \begin{abstract}
-  We investigate constructions of efficient approximate unitary $k$-designs or distributions over the unitary group which model the Haar distribution up to the $k^{th}$-moment.
+  We investigate constructions of efficient approximate unitary $k$-designs which are defined as probability distributions over the unitary group which model the Haar distribution up to the $k^{th}$-moment.
 In particular, we review the paper by Harrow and Low as a directed reading conducted during the Spring 2020 semester.
@@
 
-\section{Approximate 2-designs}
+\section{Approximate 2-designs via Local Random Quantum Circuits}
 
@@
     %
-    \begin{equation}
+    \begin{equation} \label{concreteT}
       T(p,q) = \begin{cases}
@@
 Now if we were to pick qubits $1,2$, our random $2$-qubit unitary would only act on the first two qubits non-trivially whereas every other qubit would just transform via the identity map.
-The expected \emph{coefficients} for step $t+1$ will work out to be
+By the explicit calculation shown in Equation \ref{concreteT}, the expected \emph{coefficients} for step $t+1$ will work out to be
 %
-\begin{align}
+\begin{align} \label{evolexample}
     \gamma_{t+1}(p_1,\cdots,p_n,q_1,\cdots,q_n) = \begin{cases}
@@
 \end{align}
- Oliveira et.al \cite{oliveira} had the insight to map this evolution to a Markov chain as follows: \newline
+ Oliveira et.al \cite{oliveira} had the insight to map the evolution of the $\gamma_t(p,p)$ coeffcients to a Markov chain as follows:
 %
@@
 \end{theorem}
+In the context of the example above, if we start with the coeffcient $\gamma_t(\textbf{q},\textbf{q})$ for $q \in \{0,1,2,3\}^n$, the only two possibilities are $(\textbf{q},\textbf{q}) = (0,0)$ and $(\textbf{q},\textbf{q}) \neq (0,0)$ since it is impossible for the coefficient to vanish by the first case in Equation \ref{evolexample}.
+Thus, the transitions presented in Theorem \ref{haarevol} indeed lead to the correct expected coefficient value $\gamma_{t+1}(\textbf{p},\textbf{p})$.
+Note that this is not the complete value for $\gamma_{t+1}$ as we still have to take the expectation over all the remaining qubit pairs.
 
@@
 For now, let us tackle the case where the two conditions above are true and direct our focus on the evolution of the diagonal states.
-Of course, we need to also show that the expected coeffcients for the non-diagonal states $\gamma_0(p,q)$ vanish rapidly as well.
-These matters will be addressed in the upcoming section.
+The ultimate goal of this line of thought will be to first prove Lemma \ref{mainlemma} through a carefully-done investigation of the \emph{mixing time} of $P$.
+Of course, we need to also show that the expected coeffcients of the non-diagonal states $\gamma_0(p,q)$ vanish rapidly as well.
+These matters will be addressed in a future section.
 
@@
 %
-\begin{theorem}
+\begin{theorem} \label{perron}
   Suppose that $P$ is transition matrix with positive entries and also that $P_{ij}^{k} > 0$ for some sufficiently large $k$ and for all entries $i,j$ (Regularity). Then the following must hold:
@@
 Furthermore, the left eigenvector of eigenvalue one will be the stationary distribution $\pi$.
-Since our Markov chain is both irreducible and aperiodic, it must be regular.
+Since our Markov chain is both irreducible and aperiodic, it must be regular as well.
+%
+\begin{remark}
+  It is not difficult to find counterexamples to Theorem \ref{perron} if the Markov chain is not ergodic. For instance, consider the chain
+
+  \begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+      \node[state]  (one) {1};
+      \node[state, right=of one] (two) {2};
+
+      \draw[every loop]
+        (one) edge[loop above] node {1} (one)
+        (two) edge[loop above] node {1} (two);
+    \end{tikzpicture}
+    \caption{An aperiodic but reducible chain}
+  \end{figure}
+  The transition matrix will just be $P = I$.
+Although the largest eigenvalue is positive and real, its associated left eigenvector is not unique.
+Thus, it does not have an unique stationary distribution.
+Another simple chain is
+
+  \begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+      \node[state]  (one) {1};
+      \node[state, right=of one] (two) {2};
+
+      \draw[every loop]
+        (one) edge[bend right, auto=right] node {1} (two)
+        (two) edge[bend right, auto=right] node {1} (one);
+    \end{tikzpicture}
+    \caption{An irreducible but periodic chain}
+  \end{figure}
+
+  Once again, this chain does not have a unique left eigenvector of unit eigenvalue.
+Although the stationary distribution $\pi = [\frac{1}{2} \; \frac{1}{2}]$ is unique, no initial distribution converges to it besides $\pi$ itself.
+
+  \begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+      \node[state]  (one) {1};
+      \node[state, right=of one] (two) {2};
+
+      \draw[every loop]
+        (one) edge[loop above] node {1/2} (one)
+        (one) edge[bend right, auto=right] node {1/2} (two)
+        (two) edge[loop above] node {1} (two);
+    \end{tikzpicture}
+    \caption{An aperiodic, reducible chain which converges}
+  \end{figure}
+  The transition matrix will be $$ P = \left[ \begin{matrix} \frac{1}{2} & \frac{1}{2} \\
+  0 & 1 \end{matrix} \right]$$ which has eigenvalues $\lambda_1 = 1, \lambda_2 = \frac{1}{2}$.
+Every initial distribution also converges to the unique stationary distribution $\pi = [0 \; 1]$.
+Hence, this chain exhibits a counterexample to a converse of Theorem \ref{perron}.
+\end{remark}
+%
+%
 %Mention counterexamples if time remains
@@
 \begin{definition}
-A Marko chain is \emph{reversible} if the following equality holds for all states $i,j \in \Omega$
+A Markov chain is \emph{reversible} if the following equality holds for all states $i,j \in \Omega$
 \begin{equation}
@@
 
-Finally, one more significant tool is required to lower bound the spectral gap of our matrix $P$ in Equation \ref{schemechain}.
+Finally, one more integral tool is required to lower bound the spectral gap of our matrix $P$ in Equation \ref{schemechain}.
 It is worth stressing that the Markov chain associated to $P_{\mu}$ which we will define as
@@
 
-\begin{theorem}
+\begin{theorem} \label{gapbound}
   Let $\mu$ be any 2-copy gapped distribution of gates. If $\mu$ is invariant under Hermitian conjugation then let $\Delta_\mu$ be the spectral gap of resulting Markov chain $P_\mu$. Then
@@
 %
-\begin{corollary}
+\begin{corollary} \label{muchainmix}
   The chain $P_\mu$ has 1-norm mixing time $\mathcal{O}(n(n+\log1/\epsilon))$ and 2-norm $\mathcal{O}(n\log\frac{n}{\epsilon})$
 \end{corollary}
+In fact, by Theorem \ref{gapbound} and Lemma \ref{spectralgap}, the mixing time of the $P_{\mathbb{U}(4)}$ chain upper-bounds the mixing time of the $P_{\mu}$ chain.
+This mixing time bound is required to finally prove Lemma \ref{mainlemma} as intended.
 
@@
 For this case, we will just sketch the overarching ideas of the section and leave some details to the appendix of this report and \cite{harrlow}.
+Once again, we restrict our attention to the coefficients of diagonal states and delete the identity $\sigma_0 \otimes \sigma_0$ in order for the chain to become irreducible.
+The analysis starts with the consideration of an auxiliary chain called the \emph{zero chain}.
+The zero chain describes the number of zeros present in a string.
+The idea is that the number of zeroes slow the mixing process since picking two zeroes effectively changes nothing.
+If we can show that distribution of zeros mixes well-enough, it can be used to further infer about the mixing of the total chain.
+The transition probabilities are worked out in a lemma
+%
+\begin{lemma} \label{zerochain}
+  The zero chain has the following transition matrix $P$ on the state space $\Omega = \{1,2, \cdots,n\}$. This chain actually counts the number of non-zeros of state.
+  \begin{equation}
+    P(x,y) = \begin{cases}
+    1 - \frac{2x(3n-2x -1)}{5n(n-1)} & y = x \\
+    \frac{2x(x-1)}{5n(n-1)} & y = x - 1 \\
+    \frac{6x(n-x)}{5n(n-1)} & y = x + 1 \\
+    0 & \text{otherwise}
+    \end{cases}
+  \end{equation}
+  for $1 \leq x,y \leq n$
+\end{lemma}
+\begin{proof}
+  This can be derived through a simple counting argument.
+For example, for the number of zeros to decrease ($x$ increases), two sites with a non-zero value and a zero value must be chosen and the chain must transition to pair of non-zero values.
+$9$ out of the $15$ possible transition values consist of two non-zero values.
+There the probabilty of picking a zero and a non-zero value is $\frac{2x(n-x)}{n(n-1)}$, bringing the total value to $\frac{9}{15}\frac{2x(n-x)}{n(n-1)} = \frac{6x(n-x)}{5n(n-1)}$.
+\end{proof}
+% Mention stationary distribution
+%
+\noindent It is also simple to show that the stationary distribution $\pi_0$ of the zero chain will be
+%
+\begin{equation}
+    \pi_0(x) = \frac{3^x{n \choose x}}{4^n - 1}
+\end{equation}
+%
+The integral property about this zero chain is that it mixes rapidly
+%
+\begin{theorem} \label{zerochainmixing}
+  The zero chain mixes in $\Theta(n\log\frac{n}{\epsilon})$ time.
+\end{theorem}
+%
+\noindent which leads to the rapid mixing of the total chain
+\begin{theorem} \label{totalchainmixing}
+  The total chain mixes in time $\Theta(n \log\frac{n}{\epsilon})$
+\end{theorem}
+\begin{remark}
+  Diniz and Johnathan point out in \cite{diniz} a faulty assumption made by the authors of \cite{harrlow}. In the proof of Corollary \ref{totalchainmixing},
+\end{remark}
+The proofs of these two theorems will also be assigned a portion of the appendix.
+By the discussion of Corollary \ref{muchainmix}, this proves Lemma \ref{mainlemma}.
 
@@
 
-\newpage
+%Finally, we state the proof of Theorem \ref{maintheorem}. \newline
+%\begin{proof}[(Proof of Theorem \ref{maintheorem})]
+%\end{proof}
+
 \bibliographystyle{amsplain}

### 2020-11-22 PHYS790/final.tex
Added a bit on the comments paper.

@@
 Now if we were to pick qubits $1,2$, our random $2$-qubit unitary would only act on the first two qubits non-trivially whereas every other qubit would just transform via the identity map.
-By the explicit calculation shown in Equation \ref{concreteT}, the expected \emph{coefficients} for step $t+1$ will work out to be
+By the explicit calculation shown in Equation \ref{concreteT} and linearity, the expected \emph{coefficients} for step $t+1$ will work out to be
 %
@@
 \end{theorem}
+%
 In the context of the example above, if we start with the coeffcient $\gamma_t(\textbf{q},\textbf{q})$ for $q \in \{0,1,2,3\}^n$, the only two possibilities are $(\textbf{q},\textbf{q}) = (0,0)$ and $(\textbf{q},\textbf{q}) \neq (0,0)$ since it is impossible for the coefficient to vanish by the first case in Equation \ref{evolexample}.
@@
 \begin{remark}
-  Diniz and Johnathan point out in \cite{diniz} a faulty assumption made by the authors of \cite{harrlow}. In the proof of Corollary \ref{totalchainmixing},
+  Diniz and Johnathan point out in \cite{diniz} a faulty assumption made by the authors of \cite{harrlow}.
+In the proof of Corollary \ref{totalchainmixing}, Harrow and Low make the argument that once every site has been paired with another site with non-zero value, the total chain will be uniformly mixed.
+This would reduce the mixing time to a coupon collectors problem.
+However, the authors of \cite{diniz} show a simple counterexample to this statement.
+Let us focus our attention to a $3$-qubit system with initial state $(0,0,1)$.
+We want to see the resulting distribution after hitting all three sites with two transitions.
+  First, observe that we have the following transitions $(0,0,1) \rightsquigarrow (1,0,0)$
+  \begin{align*}
+    & 001 \rightsquigarrow 100 \rightsquigarrow 100 \\
+    & 001 \rightsquigarrow 010 \rightsquigarrow 100 \\
+    & 001 \rightsquigarrow 001 \rightsquigarrow 100
+  \end{align*}
+  whereas we have the transitions $(0,0,1) \rightsquigarrow (0,0,1)$
+  \begin{align*}
+    & 001 \rightsquigarrow 001 \rightsquigarrow 001 \\
+    & 001 \rightsquigarrow 001 \rightsquigarrow 001 \\
+  \end{align*}
+   by first hitting the sites $2,3$ and $1,3$. The second transition hits the sites in the reverse order.
 \end{remark}
+
 The proofs of these two theorems will also be assigned a portion of the appendix.

### 2020-11-23 PHYS790/final.tex
Added more figures for comparision theorem.

@@
 
-The power of randomness in computation is a cornerstone of contemporary computation theory as evidenced by decades of beautiful, celebrated results.
+The theory of randomness and its role in computation are cornerstones of contemporary computation theory as evidenced by decades of beautiful, celebrated results.
 Quantum Information has, by no means, become stranger to this trend.
@@
 The central result of the paper claims that the scheme of applying sampled $2$-qubit unitaries from certain distributions with a $k$-gapped property converge to the second moment of the Haar distribution rapidly.
-The crux of the rapid convergence argument relies on mapping the evolution of an initial state under our random circuit to that of a Markov chain's.
+The crux of the convergence argument relies on mapping the evolution of  diagonal states to that of a Markov chain's.
 Consequently, the analysis becomes amenable to classic techniques created to study mixing times of Markov chains. To begin, we first state the relevant definitions as introduced by Harrow and Low.
@@
 However, we would like to find \emph{efficient approximate} $k$-designs.
-Here, efficient refers to random circuits of length polynomial in the number of qubits.
+Here, efficient refers to random circuits of asymptotic length polynomial in the number of qubits and $\log 1/\epsilon$.
 The authors actually prove some stronger, namely that \emph{$k$-copy gapped} distributions suffice.
@@
   \end{equation}
-  then we say that $\mathcal{G}_\mu$ is $k$-copy gapped if $\mathcal{G}_\mu$ has only $k!$ unit eigenvalues.
+  then we say that $\mathcal{G}_\mu$ is \emph{$k$-copy gapped} if $\mathcal{G}_\mu$ has only $k!$ eigenvectors with unit eigenvalues.
 \end{definition}
 %
-\noindent A more general definition of universality of gatesets can be defined as well:
+\noindent A more general definition for the universality of gatesets can be defined as well:
 %
@@
   Let $\mu$ be a distribution over $\mathbb{U}(d)$.
-$\mu$ is deemed to be \emph{universal} for $\mathbb{U}(d)$ if for any open ball $B \subset \mathbb{U}(d)$, there exists a sufficiently large integer $\ell$ such that $\mu^{\star \ell}(B) > 0$
+Then $\mu$ is deemed to be \emph{universal} for $\mathbb{U}(d)$ if for any open ball $B \subset \mathbb{U}(d)$, there exists a sufficiently large integer $\ell$ such that $\mu^{\star \ell}(B) > 0$
   where $\mu^{\star \ell}$ refers to the $\ell$-fold \emph{convolution} product of $\mu$:
@@
   This is the convolution product in the context of \emph{measures}.
-For example, if we let $\mu,\sigma$ be two finite Borel measures over $\mathbb{R}$, then we have a new finite Borel measure on $\mathbb{R}$, $\mu \star \sigma$:
-  $$ \mu \star \sigma(A) = \int_{\mathbb{R}}\int_{\mathbb{R}} \mathbb{I}_A(x+y) d\mu(x)d\sigma(y)$$
-  where $\mathbb{I}$ is the characteristic function for measurable set $A \subset \mathbb{R}$.
+For example, if we let $\mu,\sigma$ be two finite Borel measures over $\mathbb{U}(d)$, then we have a new finite Borel measure on $\mathbb{U}(d)$, $\mu^{\star \ell}$:
+  $$ \mu^{\star \ell} (S) = \int_{\mathbb{U}(d)}\cdots\int_{\mathbb{U}(d)} \mathbb{I}_S(U_1\cdots U_{\ell}) d\mu(U_1) \cdots d\mu(U_\ell) $$
+  where $\mathbb{I}_S$ is the characteristic function for measurable set $S \subset \mathbb{U}(d)$.
 \end{remark}
+%
 \noindent Intuitively, universality implies that sampling enough times over $\mu$ is enough to approximate any gate in $\mathbb{U}(d)$.
 This aligns with the definition introduced for discrete measures $\mu$ over finite gate sets.
-We will prove later that any universal gate set will be $k$-copy gapped on $\mathbb{U}(4)$ (Lemma \ref{universal}).
+We will prove later that any universal gate set will be $k$-copy gapped on $\mathbb{U}(4)$ (Theorem  \ref{convergence}).
 For now, let us state the main theorem and its auxiliary lemma to guide our upcoming analysis:
@@
 \begin{notation}
-  Let $\rho$ be an initial state of an $n$-qubit system i.e $\rho \in (\mathbb{C}^{2})^{\otimes n}$. We can expand $\rho$ in terms of the Pauli basis $\{\sigma_i\}_{0 \leq i \leq 3}$ as follows:
+  Let $\rho$ be an initial state of an $2n$-qubit system. We can expand $\rho$ in terms of the Pauli basis $\{\sigma_i\}_{0 \leq i \leq 3}$ as follows:
+
   \begin{equation}
-    \rho = \frac{1}{2^n} \sum_{p_1,p_2} \gamma_0(p_1,p_2) \sigma_{p_1} \otimes \sigma_{p_2}, \quad \gamma_0(p_1,p_2) \in \in \mathbb{C}
+    \rho = \frac{1}{2^n} \sum_{p_1,p_2} \gamma_0(p_1,p_2) \sigma_{p_1} \otimes \sigma_{p_2}, \quad \gamma_0(p_1,p_2) \in \mathbb{C}
    \end{equation}
+   %
   where $p_1,p_2$ run over all strings of Pauli indices of length $n$, $\{0,1,2,3\}^n$.
@@
 \end{notation}
-
+%
 \begin{lemma} \label{mainlemma}
@@
   \end{equation}
-  for $t \geq n(n + \log 1/\epsilon)$
+  for $t \geq Cn(n + \log 1/\epsilon)$
 \end{lemma}
@@
 \begin{corollary} \label{maincorollary}
-  Let $\mu$, $W$, $\gamma_W$ be as above. Then for \emph{any} initial state $\rho$, there exists constant $C$ such that for $\epsilon > 0$,
+  Let $\mu$, $W$, $\gamma_W$ be as above. Then for \emph{\bf any} initial state $\rho$, there exists constant $C$ such that for $\epsilon > 0$,
   \begin{align}
-    & \sum_{p_1,p_2 \neq 00} \left( \gamma_t(p_1,p_2) - \delta_{p_1p_2}(\frac{\sum_{p\neq 0} \gamma_0(p,p)}{4^n - 1} \right)^2 \leq \epsilon \\
-    & \sum_{p_1,p_2 \neq 00} \left| \gamma_t(p_1,p_2) - \delta_{p_1p_2}(\frac{\sum_{p\neq 0} \gamma_0(p,p)}{4^n - 1} \right| \leq \epsilon
+    & \sum_{p_1,p_2 \neq 00} \left( \gamma_t(p_1,p_2) - \delta_{p_1p_2}\frac{\sum_{p\neq 0} \gamma_0(p,p)}{4^n - 1} \right)^2 \leq \epsilon \\
+    & \sum_{p_1,p_2 \neq 00} \left| \gamma_t(p_1,p_2) - \delta_{p_1p_2}\frac{\sum_{p\neq 0} \gamma_0(p,p)}{4^n - 1} \right| \leq \epsilon
   \end{align}
@@
 We denote sampling from the Haar distribution with the notation $U \sim \mathbb{U}(d)$ when necessary.
-Furthermore, our Hermitian basis operators are such that $Tr(\sigma_p\sigma_q) = d \cdot \delta_{p,q}$ i,e orthogonal under the Hilbert-Schmidt inner product.
+Furthermore, our Hermitian basis operators are such that $Tr(\sigma_p\sigma_q) = d \cdot \delta_{p,q}$ i.e orthogonal under the Hilbert-Schmidt inner product.
 \newline
@@
 \end{equation}
-This will be the expected state after applying a sampled unitary $U$ from the uniform distribution over $\mathbb{U}(d)$. Naturally, we can reexpress this resulting state into the $\sigma_p$ basis:
+This will be the expected state after applying a sampled unitary $U$ from the Haar distribution over $\mathbb{U}(d)$. Naturally, we can reexpress this resulting state into the $\sigma_p$ basis:
 %
@@
 where the same is taken over all $k$-strings in $\{0, \cdots, d^2-1\}^k$.
-This allows us to define the matrix $\widehat{G}$ where the rows and columns are indiced by $k$-strings with entries $\widehat{G}(\textbf{q},\textbf{p})$.
-This matrix turns out to be a \emph{symmetric projector}.
+This allows us to define the matrix $\widehat{G}$ where the rows and columns are indiced by $k$-strings with entries $\widehat{G}(\textbf{p},\textbf{q})$.
+By virtue of the invariance of the Haar distribution, this matrix turns out to be a \emph{symmetric projector}.
 
@@
 %
-Note that the expectation and the trace operator commute by linearity of latter. The result follows from the cyclic invariance of trace.
+\noindent Note that the expectation and the trace operator commute by linearity of latter. The result follows from the cyclic invariance of trace.
 %
 \begin{lemma} \label{permfix}
-$\widehat{G}$ fixes permutation operators $P_{\pi}, \; \pi \in S_k$.
+$\widehat{G}$ fixes the permutation operators $P_{\pi}, \; \pi \in S_k$.
 \end{lemma}
 %
-These two results are all that is required to show that $\widehat{G}$ is indeed a projector.
+\noindent These two results are all that is required to show that $\widehat{G}$ is indeed a projector.
 %
@@
 
-      \item $\widehat{G}(p;0) = \delta_{p0}$ (Here we denote $\widehat{G}(p;0) := \widehat{G}(p,p;0,0)$)
+      \item $\widehat{G}(\textbf{p};\textbf{0}) = \delta_{p0}$ (Here we denote $\widehat{G}(\textbf{p};\textbf{0}) := \widehat{G}(p,p;0,0)$. )
       \begin{proof}
@@
 
-      \item $\widehat{G}(p,a) = \frac{1}{d^2-1}, \quad a,p \neq 0$
+      \item $\widehat{G}(\textbf{p},\textbf{a}) = \frac{1}{d^2-1}, \quad \textbf{p},\textbf{a} \neq 0$
       \begin{proof}
@@
 %
-Recall that $T(p,q)$ refers to the expected \emph{state} from applying unitary $U \sim \mathbb{U}(d)$ to the basis state $\sigma_p \otimes \sigma_q$.
+Recall that $T(p,q)$ refers to the expected state from applying unitary $U \sim \mathbb{U}(d)$ to the basis state $\sigma_p \otimes \sigma_q$.
 From the determination above, it is simple to see that $U$ uniformly samples from all diagonal basis states when acting on $\sigma_p \otimes \sigma_p, \; p \neq 0$.
@@
 \subsection{Convergence of the Random Circuit}
-Before embarking on the Markov chain analysis, we must prove that our random circuit indeed converges to Haar distribution.
-This will be proven for the general case of a universal distribution (Definition (\ref{univdef})) on $\mathbb{U}(d)$.
-Specifcally, we will prove that any universal distribution on $\mathbb{U}(d)$ will be $k$-copy gapped (Definition (\ref{gappeddef})).
+Before embarking towards Markov chain techniques, we must prove that our random circuit construction indeed converges to Haar distribution.
+Specifcally, we will prove that any universal distribution (Definition \ref{univdef}) on $\mathbb{U}(d)$ will be $k$-copy gapped (Definition \ref{gappeddef}).
 In turn, this leads to convergence in sense of $\lim_{t \rightarrow \infty} (\widehat{G}_{\mu})^{t} = \widehat{G}_{\mathbb{U}(d)}$.
@@
 If we can prove that any eigenvector of $\widehat{G}_{\mathbb{U}(d)}$ of eigenvalue one is also an eigenvector of $\widehat{G}_{\mu}$ of eigenvalue one i.e the $k$-symmetric subspace of $(\mathbb{C}^d)^{\otimes k}$ is also fixed by $\widehat{G}_{\mu}$.
-Thus, Equation $\ref{gappeddis}$ shows that indeed we can converge to $G_{\mathbb{U}(d)}$ by iterating our scheme (as denoted by $(\widehat{G}_\mu)^t$).
-Note that $\widehat{G}_{\mu}$ may not even be Hermitian, so we cannot directly argue on its spectrum (as the spectrum may not completely contain real values).
+Thus, Equation $\ref{gappeddis}$ shows that indeed we can converge to $G_{\mathbb{U}(d)}$ as all eigenvectors shrink by iterating our construction.
+Note that $\widehat{G}_{\mu}$ may not even be Hermitian, so we cannot directly argue on its spectrum (as the spectrum may not completely contain real eigenvalues).
 %
@@
 \subsection{Markov Chain Analysis}
-As discussed in a previous section, the central idea of proving \emph{rapid} convergence involves mapping the evolution of the inital state under our random circuit to that of a Markov chain's.
+As discussed in a previous section, the central idea of proving \emph{rapid} convergence involves mapping the evolution of the diagonal states under our random circuit to that of a Markov chain's.
+Here, we refer to states of the form
+$$ \psi = \frac{1}{2^n}\sum_{p}\gamma(p,p) \sigma_{p} \otimes \sigma_{p}, \quad p \in \{0,1,2,3\}^n, \; \gamma(p,p) \in \mathbb{C} $$
 In fact, we can explicitly determine the operator $P$ capturing our dynamics.
-Let $\widehat{G}^{(ij)}$ denote the unitary transformation resulting from sampling some unitary $U \sim \mathbb{U}(d)$ and applying $U$ to qubits $i$ and $j$ while leaving all other qubits undisturbed.
+Let $\widehat{G}^{(ij)}$ denote the unitary transformation resulting from sampling some unitary $U \sim \mathbb{U}(4)$ and applying $U$ to qubits $i$ and $j$ while leaving all other qubits undisturbed.
 %
@@
 %
-A major caveat of interpreting $P$ as a \emph{transition matrix} for some suitable Markov chain on the diagonal states $\sigma_p \otimes \sigma_p$ is the possibility that our initial coefficients of diagonal states $\gamma_0(p,p)$ may not even sum to one or even be non-negative.
+A major caveat of interpreting $P$ as a \emph{transition matrix} for some suitable Markov chain on the diagonal states is the possibility that our initial coefficients of the diagonal states $\gamma_0(p,p)$ may not even sum to one or even be non-negative.
 It turns out that this doesn't bring too much trouble into the analysis as pointed out by Corollary \ref{maincorollary}.
@@
 \subsubsection{Preliminaries on Markov Chains}
-Let us introduce some defintions of properties of the Markov chains we are interested in.
+Let us introduce some definitions for properties of the Markov chains we are interested in.
 Although we assume familiarity with the definition of a Markov chain along with the its associated transition matrix $P$, we refer the reader to the reference \cite{cover} for greater detail.
@@
 A Markov chain is deemed as \emph{aperiodic} if none of its states are periodic.
-Equivaently, a Markov chain is aperiodic if all of its states contain self-loops.
 \end{definition}
+\noindent Hence, a Markov chain containing self-loops i.e transitions to itself with positive probability, is aperiodic.
 %
@@
 %
-\noindent One can see this through a special case of the \emph{Perron-Frobenius Theorem for Regular Matrices}
+\noindent One can view this as special case of the \emph{Perron-Frobenius Theorem for Regular Matrices}.
 %
@@
 Although the stationary distribution $\pi = [\frac{1}{2} \; \frac{1}{2}]$ is unique, no initial distribution converges to it besides $\pi$ itself.
+Finally, consider the chain
 
@@
   \end{figure}
-  The transition matrix will be $$ P = \left[ \begin{matrix} \frac{1}{2} & \frac{1}{2} \\
+  \noindent The transition matrix will be $$ P = \left[ \begin{matrix} \frac{1}{2} & \frac{1}{2} \\
   0 & 1 \end{matrix} \right]$$ which has eigenvalues $\lambda_1 = 1, \lambda_2 = \frac{1}{2}$.
@@
 For this chain, we sample from a $2$-copy gapped distribution $\mu $ over $\mathbb{U}(4)$ rather than over Haar distribution.
-It is unclear exactly how this chain evolves as we cannot explicitly compute $P_\mu$ as we did for $P_{\mathbb{U}(d)}$.
-However, with the \emph{Comparision Theorem}, we can at least lower bound the spectral gap associated with $P_\mu$, say $\Delta_\mu$, with $\Delta_{\mathbb{U}(d)}$.
+It is unclear exactly how this chain evolves as we cannot explicitly compute $P_\mu$ as we did for $P_{\mathbb{U}(4)}$.
+However, with the \emph{Comparision Theorem}, we can at least lower bound the spectral gap associated with $P_\mu$, say $\Delta_\mu$, with $\Delta_{\mathbb{U}(4)}$.
 We first state a more general version of the theorem
@@
 \noindent Note that the bounds yielded will depend on the paths chosen for all $x \neq y \in \Omega$ such that $\widehat{P}_{xy} > 0$ and the values assigned to those paths by the constructed flow function $f$.
-\newline
+Figure \ref{compfig} shows an example of creating paths as specified in the theorem.
+%
+\begin{figure}[h!] \label{compfig}
+  \centering
+  \begin{tikzpicture}[node distance=1.5]
+      \node[state] at (0,0) (start) {x};
+      \node[state, right=of start] at (6.5,0) (finish) {y};
+      \node[state, below=of start] at (-1,0) (one) {1};
+      \node[state, right=of one] (two) {2};
+      \node[draw=none, right=of two] (three) {$\cdots$};
+      \node[state, right=of three] (four) {4};
+      \node[state, right=of four] (five) {5};
+
+      \draw[every loop]
+            (start) edge[auto=left] node {$\widehat{P}(x,y) > 0$} (finish)
+            (start) edge[auto=right, line width=1mm] node {{\tiny $P(x,1) > 0$}} (one)
+            (one) edge[auto=right, line width=1mm] node {{\tiny $P(1,2) > 0$}} (two)
+            (two) edge[line width=1mm] node {} (three)
+            (three) edge[line width=1mm] node {} (four)
+            (four) edge[auto=right, line width=1mm] node {{\tiny $P(4,5) > 0$}} (five)
+            (five) edge[auto=right, line width=1mm] node {{\tiny $P(5,y) > 0$}} (finish);
+  \end{tikzpicture}
+  \caption{Creating paths in spirit of Theorem \ref{comptheorem}. Thick lines delineate the constructed path.}
+\end{figure}
 
@@
 This chain has self-loops by Theorem \ref{haarevol} and thus, it must be ergodic.
-This puts us in the position to state the lower bound
+This puts us in a good position to state the lower bound
 
-%Mention Hermitian conjugation at some point.
+%!!!Mention Hermitian conjugation at some point.
 
@@
   \end{align*}
-   by first hitting the sites $2,3$ and $1,3$. The second transition hits the sites in the reverse order.
+   by first hitting the sites $2,3$ and $1,3$.
+The second transition hits the sites in the reverse order.
+This reveals that the number of ways to get to states $(1,0,0),(0,0,1)$ will be in ratio $3:2$; the final state will not be uniformly distributed.
 \end{remark}

### 2020-11-23 773/cohom.tex
First commit of 773 project.

\documentclass[12pt]{amsart}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.2cm, right=2.2cm]%
{geometry}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usetikzlibrary{cd}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}{Proposition}[theorem]

\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{xca}[theorem]{Exercise}

\newtheorem{remark}[theorem]{Remark}
\numberwithin{equation}{section}

\newtheorem{notation}[theorem]{Notation}

\begin{document}

\title{Cohomology of Connected Compact Lie Groups}
\author{Edward Kim}
\date{\today}
\maketitle
\tableofcontents

\section{De Rham Cohomology}
We rapidly review some definitions and concepts from de Rham cohomology before introducing the main topics. Recall that for a $n$-dimensional smooth manifold $M$, we define $\Omega^p(M)=\Gamma(M,\bigwedge^p(T^*M))$ to be the \emph{$p^{th}$ differential forms} on $M$ i.e smooth sections of the $p$-alternating tensor bundle. The direct sum of these vector spaces:
  $$\Omega^*(M) = \bigoplus_{p = 0}^{n} \Omega^p(M)$$
along with the wedge product $\wedge: \Omega^n(M) \times \Omega^m(M) \rightarrow \Omega^{n+m}(M)$ turns $\Omega^*(M)$ into an associative, anticommutative graded $C^{\infty}(M)$-algebra. Furthermore, there exists an \emph{exterior derivative}: $d: \Omega^{n}(M) \rightarrow \Omega^{n+1}(M)$ defined as follows on basis forms:
$$ d(fdx^1 \wedge dx^2 \wedge \cdots \wedge dx^n) = df \wedge dx^1 \wedge dx^2 \wedge \cdots \wedge dx^n$$
We will frequently use the global formula for the exterior derivative on $\Omega^p(M)$:
\begin{align} \label{globalexter}
  d\omega(X_1,\cdots, X_{p+1})
  & = \sum_{i = 1}^{p+1} (-1)^i X_i\omega(X_1,\cdots,\widehat{X_i},\cdots,X_{p+1}) \\
  & + \sum_{i < j} (-1)^{i+j} \omega([X_i,X_j],X_1,\cdots, \widehat{X_i},\cdots,\widehat{X_j},\cdots,X_{p+1})
\end{align}
where $X_1,\cdots,X_{p+1}$ are smooth vector fields on $M$.
Now it turns out that $d \circ d = 0$, defining the following cochain complex:
%
\begin{definition}
  The $p^{th}$ \emph{de Rham cohomology group} is defined to be the quotient groups
  \begin{equation}
    H^p(M) = Z^p(M) / B^p(d: \Omega^{p-1}(M) \rightarrow \Omega^{p})
  \end{equation}
  where
  \begin{align*}
    & Z^p(M) = \text{Ker}\left(d: \Omega^p(M) \rightarrow \Omega^{p+1} \right) \\
    & B^p(M) = \text{Im}\left(d: \Omega^p(M) \rightarrow \Omega^{p+1} \right)
  \end{align*}
\end{definition}
For $\omega \in \Omega^p(M)$, recall that $\omega$ is called \emph{closed} if $d\omega = 0$ and is called \emph{exact} if there exists $\eta \in \Omega^{p-1}(M)$ such that $\omega = d\eta$. Thus, $Z^p(M)$ denotes the closed $p$-forms on $M$ while $B^p(M)$ denotes the exact $p$-forms on $M$. The following proposition shows that smooth maps induce homomorphisms between the cohomology groups via their pullbacks:
%
\begin{proposition}
  Let $F:M \rightarrow N$ be a smooth map. The pullback $F^*: \Omega^p(N) \rightarrow \Omega^p(M)$ takes forms from $Z^p(N)$ into $Z^p(M)$ and $B^p(N)$ into $B^p(M)$, inducing a homomorphism between cohomology groups $H^p(N) \rightarrow H^p(M)$.
\end{proposition}
%
\begin{proof}
  If $\omega \in \Omega^p(N)$ is closed, then $dF^*\omega = F^*d\omega = 0$ since the exterior derivative commutes with pullbacks . Similarly, if $\omega = d\eta$ for some $\eta \in \Omega^{p-1}(N)$, then $F^*\omega = F^*d\eta = dF^*\eta$. This shows the proposition with the induced cohomology map being defined as:
  $$ F^*[\omega] = [F^*\omega], \quad \omega \in \Omega^p(N) $$
\end{proof}
This further shows that diffeomorphisms between manifolds induce isomorphisms between their respective cohomology groups. For more information about background topics, see \cite{lee}.

\section{Left-invariant Forms}
We base a good bit of our presentation on the articles by Fok and Zhang (\cite{fok},\cite{zhang}). The discussions on invariant forms are based on the treatment in \cite{chevalley}. The original paper detailing these arguments is \cite{reeder}.

We begin by deeming $G$ to be a connected compact Lie group. First, we start with the definition of a left-invariant
%
\begin{definition}
  Let $\omega \in \Omega^p(G)$. Then $\omega$ is said to be \emph{left-invariant} if
  \begin{equation}
    L_g^*\omega = \omega, \quad g \in G
  \end{equation}
\end{definition}
%
\noindent We denote $\Omega_L^p(G)$ to be the subspace of left-invariant $p$-forms on $G$.
%
\begin{lemma}
  Let $\omega \in \Omega_L^p(G)$. Then $d\omega \in \Omega_L^{n+1}$
\end{lemma}
\begin{proof}
  For all $g \in G$:
  $$L^*_gd\omega = dL^*\omega = d\omega$$
\end{proof}
\noindent This shows that $\Omega_L^{*}(G)$ form a cochain complex under $d$. Define $H^p_L(G)$ be cohomology of $\Omega^p_L(G)$. Furthermore, note that since pullbacks distribute over the wedge product:
$$ L^*_g(\omega_1 \wedge \omega_2) = L^*_g\omega_1 \wedge L^*_g\omega_2  = \omega_1 \wedge \omega_2$$
This shows that when $\omega_1 \in \Omega_L^n(G)$ and $\omega_2 \in \Omega_L^m(G)$, then $\omega_1 \wedge \omega_2 \in \Omega_L^{n+m}(G)$. In fact, the wedge product becomes a well-defined map: $\wedge: H^n_L(G) \times H^{m}_L(G) \rightarrow H^{n+m}_L(G)$ given by $$([\omega_1] ,[\omega_2]) \mapsto [\omega_1 \wedge \omega_2]$$
This endows a ring structure on $H^{*}_L(G)$ and an identical argument shows that $H^*(G)$ is also a ring. Now as subspaces, there exists an inclusion map $i: \Omega_L^p(G) \hookrightarrow \Omega^p(G)$ which induces a cohomology map between complexes $i': H^*_L(G) \rightarrow H^*(G)$. In fact, $i'$ is a ring isomorphism:
%
\begin{theorem} \label{leftrep}
$i'$ is a ring isomorphism
\end{theorem}
%
\begin{proof}
  To show surjectivity, given a form $\omega \in \Omega^p(G)$, there is an invariant form $\omega' \in \Omega^p_L(G)$ defined as
  %
  \begin{equation}
    \omega' = \int_G L^*_g\omega \; dg
  \end{equation}
  Recall that a similar trick is used to create an invariant inner product from an existing inner product on some representation space. Then
  %Arugment zhang
  To show injectivity: let $\omega \in \Omega_L^p(G)$ such that $\omega = d\eta$ for some $\eta \in \Omega^{p-1}(G)$. Then
  $$\omega = \omega' = \int_G L^*_g\omega \; dg = \int_G L^*_g d\eta \; dg = d \int_G L^*_g \eta \; dg = d\eta' $$
  The first equality is by left-invariance of $\omega$. Then $\omega$ must be an exact in respect to cochain complex $\Omega^*_L(G)$, showing injectivity.
\end{proof}
%
Theorem \ref{leftrep} states that the classes of the de Rham complex of $G$, $H^p(G)$ can be represented through its left-invariant counterparts $H^p_L(G)$. The additional structure placed on the left-invariant forms on $G$ give rise to an isomorphism between alternating forms on the Lie algebra of $G$, $\mathfrak{g}$  and the left-invariant forms
\begin{lemma}
Define
      \begin{gather}
        \psi: \Omega_L^{*}(G) \rightarrow \bigwedge^* \mathfrak{g}^* \\
        \omega \mapsto \{\omega\} := \omega\vert_{\wedge^n T_eG}
      \end{gather}
or explicitly:
$\{\omega\}(X_1,\cdots X_n) = \omega_e((X_1)_e, \cdots, (X_n)_e)$. The map $\psi$ will be a ring isomorphism.
\end{lemma}
\begin{proof}
  Showing that $\psi$ is a ring homomorphism just involves reasoning about the wedge product. To show surjectivity, take some $\mu \in \bigwedge^n\mathfrak{g}^*$ and define the $p$-form
  $$ \omega_g((X_1)_g,\cdots,(X_n)_g) = \mu((L_{g^{-1}})_*(X_1)_g, \cdots, (L_{g^{-1}})_*(X_n)_g) $$
  where $(L_{g^{-1}})_*(X_i)$ is the pushforward vector field of $X_i$ in respect to the diffeomorphism $L_{g^{-1}}$. To check that $\omega \in \Omega_L^n(G)$:
  \begin{align*}
    (L^*_g\omega)_h((X_1)_h, \cdots ,(X_n)_h)
    & = \omega_{gh}((L_g)_*(X_1)_h,\cdots, (L_g)_*(X_n)_h) \\
    & = \mu((L_{h^{-1}g^{-1}}L_g)_*(X_1)_h, \cdots (L_{h^{-1}g^{-1}}L_g)_*(X_n)_h) \\
    & = \mu((L_{h^{-1}})_*(X_1)_h), \cdots, (L_{h^{-1}})_*(X_n)_h) \\
    & = \omega_h((X_1)_h, \cdots, (X_n)_h)
  \end{align*}
  so $L^*_g\omega = \omega$ as desired. As for injectivity, suppose we had $\alpha \in \Omega^n_L(G)$ such that $\{\alpha\} = 0$. By a similar line of thinking as above:
  \begin{align*}
    \alpha_h((X_1)_h,\cdots,(X_n)_h)
    & = (L^*_h\alpha)_e((L_{h^{-1}})_*(X_1)_h, \cdots, (L_{h^{-1}})_*(X_n)_h) \\
    &= \alpha_e((L_{h^{-1}})_*(X_1)_h, \cdots, (L_{h^{-1}})_*(X_n)_h) \\
    & = 0
  \end{align*}
\end{proof}

There is actually a stronger property based on this ring isomorphism $\psi$. We will now show that there exists a homomorphism $\delta: \bigwedge^n\mathfrak{g}^* \rightarrow \bigwedge^{n+1}\mathfrak{g}^*$ which commutes with the $\psi$, turning $\bigwedge^*\mathfrak{g}^*$ into a cochain complex.
%
\begin{lemma}
There exist a linear map $\delta$ which makes the below diagram commute:
\begin{figure}[h!]
  \centering
  %
  \begin{tikzcd}
    \Omega_L^n \arrow[r, "d"] \arrow[d, "\psi"] &
    \Omega_L^{n+1} \arrow[d, "\psi"] \\
    \bigwedge^n\mathfrak{g}^* \arrow[r, "\delta"] &
    \bigwedge^{n+1}\mathfrak{g}^*
  \end{tikzcd}
\end{figure}

with the explicit form
\begin{equation}
  \delta\omega(\xi_1,\cdots, \xi_{n+1}) = \sum_{i < j} \omega([\xi_i,\xi_j], \xi_1,\cdots,\widehat{\xi_i},\cdots,\widehat{\xi_j},\cdots, \xi_{n+1})
\end{equation}
\end{lemma}
%
\begin{proof}
  Let us first take the $\xi_1,\cdots,\xi_{n+1}$ which are tangent vectors living in $T_eG$ and extend them to left-invariant vector fields on $G$ via a pushforward, $(X_i)_g = (L_g)_*(\xi_i)$. Expand $\psi \circ d$ as follows:
  \begin{align*}
    (\psi d\omega)(\xi_1,\cdots, \xi_{n+1})
    & = d\omega(X_1,\cdots,X_{n+1}) \\
    & = \sum_{i < j} (-1)^{i+j} \omega([X_i,X_j],X_1,\cdots,X_i, \cdots, \widehat{X_i}, \cdots, \widehat{X_j}, \cdots,X_{n+1}) \\
    & = \sum_{i < j} (-1)^{i+j} \{\omega\}([(X_i)_e,(X_j)_e],(X_1)_e,\cdots,(X_i)_e, \cdots, \widehat{X_i}, \cdots, \widehat{X_j}, \cdots,(X_{n+1})_e) \\
    & = (\delta \psi \omega)(\xi_1,\cdots,\xi_{n+1})
  \end{align*}
  The first equality follows since the restriction is $(X_i)_e = \xi$ by construction. The second equality arises from the left-invariance of $\omega$ and the $X_i$. Recall the global formula in Equation \ref{globalexter} and the terms:
  \begin{align*}
     X_i\omega(X_1,\cdots,\widehat{X_i},\cdots,X_{n+1})
     & = \mathcal{L}_{X_i}(\omega(X_1,\cdots,\widehat{X_i},\cdots,X_{n+1}) \\
    & = (\mathcal{L}_{X_i})(\omega(X_1,\cdots,\widehat{X_i},\cdots,X_{n+1})
    + \sum_{j = 0, j \neq i } \omega(X_1,\cdots [X_i,X_j], \cdots X_{n+1}) \\
    & = \sum_{j = 0, j \neq i } \omega(X_1, \cdots, \widehat{X_i}, \cdots, [X_i,X_j], \cdots X_{n+1})
  \end{align*}
  since $\mathcal{L}_{X_i}\omega = 0$ and $\mathcal{L}_{X_i}X_i = [X_i,X_i] = 0$. Taking the total sum reveals that:
  \begin{equation}
    \sum_{i=0}^{n+1}X_i\omega(X_1,\cdots,\widehat{X_i},\cdots,X_{n+1}) = 0
  \end{equation}
  The third equality is once again just from the restrictions of the $X_i$ and the definition of $\psi$.
\end{proof}
\noindent The lemma shows that the $\bigwedge^*\mathfrak{g}^*$ under the $\delta$ map is a cochain complex with $\psi$ inducing a cochain complex isomorphism.
\begin{theorem}
  $H^*_L(G) \cong H^*(\bigwedge^*\mathfrak{g}^*,\delta) $
\end{theorem}
\noindent $H^*(\mathfrak{g}) = H^*(\bigwedge^*\mathfrak{g}^*,\delta)$ is called the \emph{Lie algebra cohomology} of $\mathfrak{g}$. We have shown that for connected compact groups, its Lie algabra cohomology isomorphic to its left-invariant cohomology groups.

\section{Bi-invariant Forms}
We can extend these results to the so-called \emph{Bi-invariant classes} of $G$
%
\begin{definition}
  Let $\omega \in \Omega^n(G)$. $\omega$ is bi-invariant if $\omega \in \Omega_B^n(G) = \Omega_L^n(G) \cap \Omega_R^n(G)$. That is:
  \begin{align*}
    & L^*_g\omega = \omega \\
    &  R^*_g\omega = \omega
  \end{align*}
  for all $g \in G$
\end{definition}
%
The use of the average to discover bi-invariant cohomology representatives in $[\omega]$ also apply:
\begin{equation}
  \omega' = \int_{G}\int_{G} L^*_{g}R^*_{g}\omega \; dg dg
\end{equation}
A similar argument to Theorem \ref{leftrep} yields that the inclusion of subspaces $\Omega_B^n(G) \hookrightarrow \Omega^n(G)$ is also a ring isomorphism
%
\begin{theorem}
  $H_B^{*}(G) \cong H^*(G)$ as rings
\end{theorem}

\bibliographystyle{amsplain}
\bibliography{biblio}

\end{document}

### 2021-01-26 Complexity/LMNTheorem/paper.tex
Reorganized papers into more appropriate categories

--- 2020-11-12 CS590/paper.tex
@@
 \subsection{Characters}
-
 We introduce some stepping stones defining the tools used in the Fourier Analysis of boolean functions.
@@
 %
-\noindent This yields the standard inner product for maps $f,g \in L_2(G)$: \newline
+\noindent which yields the standard inner product for maps $f,g \in L_2(G)$: \newline
 %
@@
 In this case, we can show that indeed for any $a,b \in G$ and $\chi$ any character of $G$:
-$$ \chi(a+b) = Tr(\rho(a+b)) = Tr(\rho(a)\circ\rho(b)) = \lambda_a \lambda_b = Tr(\rho(a))Tr(\rho(b)) = \chi(a)\chi(b)$$ Note that this relation does not hold for arbitrary finite groups.
-The genealizations of the Fourier Transform generally lie in \emph{Harmonic Analysis}.
+$$ \chi(a+b) = Tr(\rho(a+b)) = Tr(\rho(a)\circ\rho(b)) = \lambda_a \lambda_b = Tr(\rho(a))Tr(\rho(b)) = \chi(a)\chi(b)$$ Note that this relation does not hold for arbitrary finite groups since the trace is generally not multiplicative.
+The genealizations of Fourier Analysis generally lie in \emph{Harmonic Analysis}.
 For more information on the Representation Theory of Finite Groups, see \cite{serre}, \cite{webb}.
@@
 Invoking Corollary (\ref{hastaddegree}) tells us that our final restricted function $f_{\rho_0\rho_1...\rho_{d-1}}$ has Fourier degree of at most $s$ with probability $< 2^{-s}$.
-The final bound arises from the observation that we applied the beginning analysis for each of the bottom gates exactly once, the switching lemma exactly once for each gate except for the output gate, and we invoked the corollary for the output gate.
-Each step contributed probability of $< 2^{-s}$.
+The final bound arises from the observation that we applied the beginning analysis for each of the bottom gates exactly once, the switching lemma exactly once for each gate above the bottom gates except for the output gate, and we invoked the corollary for the output gate.
+Each step contributed probability $< 2^{-s}$.
 \end{proof}
@@
 \begin{remark}{(\emph{Some concluding remarks})}
-  he bounds shown by Linai-Mansour-Nisan have been improved by Tal (\cite{tal}). Specifically, he showed that for any $n$-ary boolean function $f$ with size $M$ and depth $d$,\
+  The bounds shown by Linai-Mansour-Nisan have been improved by Tal (\cite{tal}). Specifically, he showed that for any $n$-ary boolean function $f$ with size $M$ and depth $d$
   \begin{equation}
@@
   \begin{theorem}{\emph{(Tal \cite{talshrinkage})}}
-    Let $f\{1,-1\}^n \rightarrow \{-1,1\}$ be a Boolean function, let t,$C > 0$ such that for all $k$, $\sum_{|S| > k} |\hat{f}(S)|^2 \leq C e^{k/t}$ and let $\rho$ be a $p$-random restriction, then for all $d$, $Pr[\text{deg}(f_{\rho})] \leq C \cdot (4pt)^d$.
+    Let $f\{1,-1\}^n \rightarrow \{-1,1\}$ be a Boolean function, let $t,C > 0$ such that for all $k$, $\sum_{|S| > k} |\hat{f}(S)|^2 \leq C e^{k/t}$ and let $\rho$ be a $p$-random restriction, then for all $d$, $Pr[\text{deg}(f_{\rho})] \leq C \cdot (4pt)^d$.
   \end{theorem}
-  His work uses results in Quantum Query Complexity to ultimately culminate to another proof of the shrinkage exponent of Demorgan Formulae under random restrictions:
+  His work uses results in Quantum Query Complexity to ultimately culminate into another proof of the shrinkage exponent of Demorgan Formulae under random restrictions:
   \begin{theorem}{\emph{(Tal \cite{talshrinkage})}}

### 2021-01-26 Nonlinear/KAA-Arch/samplepaper.tex
Reorganized papers into more appropriate categories

--- 2020-07-25 KAA-Arch/samplepaper.tex
@@
 \begin{abstract}
-Reachable set computation is one of the widely used techniques for verification of safety properties of dynamical system. 
+Reachable set computation is one of the many widely-used techniques for the verification of safety properties of dynamical systems. 
 %
@@
 %
-In this paper, we describe Kaa, a terse Python implementation of reachable set computation which leverages the widely used symbolic package sympy.
+In this paper, we describe Kaa, a terse Python implementation of reachable set computation which leverages the widely used symbolic package \emph{sympy}.
 %
-Additionally, we simplify the user interface with the tool and provide easy to use plotting utilities.
+Additionally, we simplify the user interface and provide easy-to-use plotting utilities.
 %
@@
 In this paper, we present a Python implementation of algorithm presented in~\cite{dreossi2017reachability}.
-
 %
@@
 %
-First, using standard libraries for symbolic manipulation in Sympy, the algorithm for reachability can be implemented very concisely.
+First, using standard libraries for symbolic manipulation in Sympy, the algorithm for reachability can be implemented concisely.
 %
@@
 %
-In contrast, SAPO is more than $2000$ lines of C\texttt{++} code with memory and pointer management performed completely by the developer.
+In contrast, SAPO consists of more than $2000$ lines of C\texttt{++} code with memory and pointer management performed completely by the developer.
 %
@@
 % %
-Second, reimplementing the tool in Python makes user interaction significantly easy. 
+Second, reimplementing the tool in Python makes user interaction significantly simpler. 
 %
-In SAPO, one has to recompile the code along with the model file to generate the binary and the user has to execute the binary for computing the reachable set.
+In SAPO, one has to recompile the code along with the model file to generate the binary. Subsequently the user has to execute the binary to compute the reachable set.
 %
 Additionally, visualizing the reachable set using SAPO is a two step process.
-In the first step, SAPO generates a MATLAB script that contains projections of reachable set for visualizing time or phase plots of the reachable set.
-In the second step, the user feeds the generated script into either MATLAB or Octave to illustrate the reachable set.
+In the first step, SAPO generates a MATLAB script which plots projections of reachable set for visualizing time or phase plots of the reachable set.
+In the second step, the user feeds the generated script into either MATLAB or GNU Octave to illustrate the reachable set.
 %
-In contrast, Kaa has very intuitive interface for computing and visualizing reachable set.
+In contrast, Kaa offers plotting functionalities through the \emph{matplotlib} library for visualizing the reachable set.
 %
-The models and the code for computing reachable set using Kaa are provided in Python and do not read any compilation. 
+The models and the code for computing reachable set using Kaa are provided in Python and do not require any compilation steps.
 %
-We also leverage the matplotlib library for visualizing the reachable set. 
-%
-Integrating these two, we provide an easy to use Jupyter notebook interface for computing and visualizing reachable set with Kaa for some of the standard benchmark models.
+Integrating these two, we provide an easy-to-use Jupyter notebook interface for computing and visualizing reachable set with Kaa for some of the standard benchmark models.
 %
@@
 %
-In light of these features and advantages, we believe Kaa can be used as a easy first step for introducing reachability in a pedagogical setting.
+In light of these features and advantages, we believe Kaa could be used as a natural first step for introducing reachability in a pedagogical setting.
 
@@
 %
-A similar argument follows for the lower bound on infimum.
+A similar argument follows for the lower bound on the infimum.
 %
 
-For the remainder of the document, we assume that by using functional composition and the Bernstein representation, we can compute the upper bound on supremum and lower bound on infimum of polynomial functions over parallelotopes.
+For the remainder of the document, we assume that by using functional composition and the Bernstein representation, we can compute the upper bound on supremum and the lower bound on the infimum of polynomial functions over parallelotopes.
 %
@@
 
-One of the primary reasons for preferring Python as an implementation language for this algorithm is the presence of powerful, well-tested symbolic and matrix-computation libraries.
+One of the primary reasons for preferring Python as the implementation language for this algorithm is the presence of powerful, well-tested symbolic and matrix-computation libraries.
 %
@@
 %
-Additionally, the \emph{matplotlib} library has easy plotting facilities that we integrate into our tool for visualizing the reachable set.
+Additionally, the \emph{matplotlib} library has accessible plotting facilities that we integrate into our tool for visualizing the reachable set.
 In particular, \emph{matplotlib} facilitates the ability to plot several reachable sets simultaneously.
-This will aid others in performing more complex bundle-transformation experiments in future.
+This will aid others in performing more complex bundle-transformation experiments in the future.
 
 %
-Finally, Python has a rich set of multiprocessing libraries, namely \emph{multiprocessing} or \emph{mp} for short.
-In the future, we plan to exploit the parallelizable nature of bundle-transformtion computations using \emph{mp}'s ability to provide powerful multiprocessing features without much verbosity.
+Finally, Python contains a rich set of multiprocessing libraries, namely \emph{multiprocessing} or \emph{mp} for short.
+We plan to exploit the parallelizable nature of bundle-transformtion computations using \emph{mp}'s ability to provide powerful multiprocessing features without much verbosity.
 
@@
 We evaluate our tool on the benchmarks that are shipped along with SAPO.
-With each benchmark, the reachable sets of SAPO are juxtaposed with that of Kaa's to demonstrate the accuracy and quality of our reachable set.
+With each benchmark, the reachable sets of SAPO are juxtaposed with that of Kaa's to demonstrate the accuracy and quality of our reachable sets.
 A comparison of time taken by Kaa and SAPO for each benchmark is provided in Table~\ref{tab:one}.
@@
 \subsection{Lotka-Volterra Model}
-\noindent We also test on the competetitive Lotka-Volterra models for predator-prey biological systems. Our instance is a 5-dimensional sysmtem governed as below:
+\noindent We also test on the competetitive Lotka-Volterra models for predator-prey biological systems. Our instance is a 5-dimensional system governed as below:
 \begin{align}
@@
 \end{align}
-where $\alpha,\beta$ are two parameters assigned as $\alpha=0.5$ and $\beta=5$. We set $\Delta =0.01$ here. The relevant figures can be found in Figure \ref{fig5}.
+where $\alpha,\beta$ are two parameters assigned as $\alpha=0.5$ and $\beta=5$. We set $\Delta =0.01$ here. The figures can be found in Figure \ref{fig5}.
 
@@
 
-\noindent It is evident from Table~\ref{tab:one} that while the current implementation in Python is very intuitive and concise, it incurs severe performance penalties. 
+\noindent It is evident from Table~\ref{tab:one} that while the current implementation in Python is intuitive and concise, it incurs severe performance penalties. 
 %
-Nevertheless, we pursued this activity because of the pedagogical value in making this an accessible tool for implementing and understanding reachability.
+Nevertheless, we pursued this direction because of the pedagogical value in making this an accessible tool for implementing and understanding reachability.
 To this end, the repository contains a Jupyter notebook with the name \href{https://github.com/Tarheel-Formal-Methods/kaa/blob/master/kaa-intro.ipynb}{kaa-intro} designed to interactively introduce graduate students and practitioners to the usage of Kaa.
-The notebook even contains a subset of the examples shown above, giving the auidence a chance to run the code for themselves.
+The notebook even contains a subset of the examples shown above, giving the auidence a hands-on method to run the code for themselves.
 An immediate next step is to deploy extensive profiling to find performance bottlenecks and subsequently improve on them.
@@
 %
-The code can be accessed through \url{https://github.com/Tarheel-Formal-Methods/kaa}
+The code can be accessed through \url{https://github.com/Tarheel-Formal-Methods/kaa}.
 \newline

### 2021-01-26 misc/InequalitiesPaper/Chrystal.tex
Reorganized papers into more appropriate categories

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,geometry,graphicx,enumerate}

\def\Name{Edward Kim}  % Your name

%for code%
\usepackage{listings}
\lstset{language=python}

\title{Proof of Chrystal's Inequality}
\author{\Name}

\textheight=9in
\textwidth=6.5in
\topmargin=-.75in
\oddsidemargin=0.25in
\evensidemargin=0.25in
\DeclareMathSizes{11}{19}{13}{9}

\begin{document}
\maketitle

\textbf{Proof:}

We first note a couple things about the hypothesis. Firstly, \newline
\begin{equation}
  \prod_{i} a_i = \l^n \implies (\prod_{i} a_i)^{-n} = \l 
\end{equation}

We also know that the above expression is not undefined since $a > 0$. \newline

Now we shall invoke Theorm 40 (pg 39) to prove the claim.

Let our sets be the following: $(1, a_1), (1, a_2), ... , (1, a_n)$.
When we invoke the theorm, we get the following inequality:
\begin{equation}
 (1+a_1)^{-n}(1+a_2)^{-n}...(1+a_n)^{-n} > (\prod^{n}1^{-n} + \prod_{i}a_i^{-n}) \implies
 ((1+a_1)(1+a_2)...(1+a_n))^{-n} > (\prod^{n}1^{-n} + \prod_{i}a_i^{-n})
\end{equation}

However, by our observation above,

\begin{equation}
 ((1+a_1)(1+a_2)...(1+a_n))^{-n} > (1 + \l)
\end{equation}

After we exponentiate both sides by $n$, we get

\begin{equation}
  (1+a_1)(1+a_2)...(1+a_n)> (1 + \l)^{n}
\end{equation}

We know that the inequality is preserved since $a > 0$. 
Also if all $a_1 = a_2 = ... = a_n$, then $\frac{1}{a_1} = \frac{1}{a_2} = ... = \frac{1}{a_n}$. This would contradict our requirements for the 
validity of Theorm 40. So not all $a_i$'s can take one the same value.

\end{document}

### 2021-01-26 misc/InequalitiesPaper/Inequality2.tex
Reorganized papers into more appropriate categories

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,geometry,graphicx,enumerate}

\def\Name{Edward Kim}  % Your name

%for code%
\usepackage{listings}
\lstset{language=python}

\title{Proof of Inequality 2}
\author{\Name}

\textheight=9in
\textwidth=6.5in
\topmargin=-.75in
\oddsidemargin=0.25in
\evensidemargin=0.25in
\DeclareMathSizes{11}{19}{13}{9}
\setlength\parindent{0pt}

\begin{document}
 \maketitle

 (Holy crap. This induction took me hours. I was trying all these weird methods when the simplest answer was right in front of me.)
 \newline \newline
 \noindent

 Let us first prove the base case for when $n = 2$. This is proven in Inequality 1 so the base case is proven. Assume
 that the inequality holds for some $k$ so that $n = 2^k$. Now let us prove the inequality for 
 $2^{k+1}$. \newline

 \noindent

 We have the sum:

 \begin{equation}
  \sum_{i=1}^{2^{k+1}} \frac{1}{1+a_i}.
 \end{equation}

 Now choose any two partitions and relabel the $a_i$ such that the sum becomes:

  \begin{equation}
  \sum_{i=1}^{2^{k+1}} \frac{1}{1+a_i} = \sum_{i=1}^{2^{k}} \frac{1}{1+a_i} + \sum_{i=1}^{2^{k}} \frac{1}{1+\widetilde{a_i}}.
 \end{equation}

 Let us denote the sum as $\phi$. \newline

 By our induction hypothesis,

 \begin{equation}
  \phi \geq \frac{2^k}{1 + (a_1...a_{2^k})^{\frac{1}{2^k}}} + \frac{2^k}{1 + (\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}}}
 \end{equation}

Let the above sum be denoted as $R$. Factor out the $2^k$

\begin{equation}
 \phi \geq 2^k(\frac{1}{1 + (a_1...a_{2^k})^{\frac{1}{2^k}}} + \frac{1}{1 + (\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}}})
\end{equation}

However, $(a_1...a_{2^k})^{\frac{1}{2^k}} \geq 1$ and $(\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}} \geq 1$. So we can
use our induction hypothesis again!

\begin{equation}
 \phi \geq R \geq 2^k(\frac{2}{1 + \sqrt{((a_1...a_{2^{k}})^{\frac{1}{2^k}})((\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}})}})
\end{equation}

However, this can be simplified to

\begin{gather}
 \phi \geq 2^k(\frac{2}{1 + \sqrt{(a_1...a_{2^{k+1}})^{\frac{1}{2^{k}}}}}) \\
 \phi \geq \frac{2^{k+1}}{1 + (a_1...a_{2^{k+1}})^{\frac{1}{2^{k+1}}}}
\end{gather}

This proves the induction and, thus, we have proved the inequality.
\end{document}

### 2021-01-26 misc/InequalitiesPaper/Proof2.tex
Reorganized papers into more appropriate categories

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,geometry,graphicx,enumerate}

\def\Name{Edward Kim}  % Your name

%for code%
\usepackage{listings}
\lstset{language=python}

\title{Proof of Inequality 2}
\author{\Name}

\textheight=9in
\textwidth=6.5in
\topmargin=-.75in
\oddsidemargin=0.25in
\evensidemargin=0.25in
\DeclareMathSizes{11}{19}{13}{9}
\setlength\parindent{0pt}

\begin{document}
\maketitle

First, let us state the problem and develop some motivation for it: \newline

Let $n \in \mathbb{N}$. Then the following holds provided that $a_i \geq 1$ for $1 \leq i \leq n$.

\begin{equation}
 \sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \frac{n}{1 + (a_1...a_n)^{\frac{1}{n}}}
\end{equation}

First, let us prove a small lemma before we proceed. \newline

Let $\mathbb{P}$ denote the set of prime numbers. \newline

\textbf{Lemma:} If the inequality holds for the cases where $n \in \mathbb{P}$, then the inequality holds for all natural numbers $n \in \mathbb{N}$. \newline

\textbf{Proof:} Since $n \in \mathbb{N}$, by the Fundamental Theorm of Arithmetic, $n$ can decomposed into a product of primes unique up to order. Let $n = p_1^{l_i}...p_r^{l_r}$ where $l_i \geq 1$. We can expand the product so that each prime in the product will have powers of 1.
So $n = p_1...p_g$. We prove the inequality by the following steps. Let $p_i$ be any prime in the expanded product. We can easily partition the terms into $C$ partitions of $p_i$ terms where 
\begin{equation}
  C = \prod_{\substack{1 \leq j \leq g \\ j \neq i}} p_j
\end{equation}

However, since the inequality holds for any prime number, we get the following inequality:
\begin{equation}
 \sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \sum_{\mathcal{C}} \frac{p_i}{1 + ( a_{\sigma(1)}...a_{\sigma(p_i)} )^{\frac{1}{p_i}}}
\end{equation}

where $\mathcal{C}$ is the set of all $C$ paritions of the sum. $a_{\sigma(1)},...,a_{\sigma(p_i)}$ are the $p_i$ terms in each parition iterated over the sum. Now take another prime from $C$ which does not contain $p_i$. Let us call this new prime $p_r$.
Let $S$ be the new sum on the right side of the inequality of (3). We can think of $S$ as $C'$ paritions of $p_r$ terms where $C'$ is

\begin{equation}
 C' = \prod_{\substack{1 \leq j \leq g \\ j \neq i \\ j \neq r}} p_j
\end{equation}

Once again, since the inequality holds for all prime numbers, we get the following inequality:

\begin{gather}
\sum_{\mathcal{C}} \frac{p_i}{1 + ( a_{\sigma(1)}...a_{\sigma(p_i)} )^{\frac{1}{p_i}}} \geq p_i(\sum_{\mathcal{C'}} \frac{p_r}{1 + ( \prod_{\mathcal{C}}(a_{\sigma(1)},...,a_{\sigma(p_i)})^{\frac{1}{p_i}})^{\frac{1}{p_r}}}) 
\end{gather}

where $\mathcal{C}$ is the set of all $p_r$ partitions of $S$. We see that this valid since $(a_{\sigma(1)},...,a_{\sigma(p_i)})^{\frac{1}{p_i}} \geq 1$ for all such $p_i$ partitions. But the left side of the inequality is:

\begin{equation}
\sum_{\mathcal{C'}} \frac{p_ip_r}{1 + ( \prod_{\mathcal{C}}a_{\sigma(1)},...,a_{\sigma(p_i)})^{\frac{1}{p_rp_i}}}
\end{equation}

Also the product in the denominator contains a product of $p_ip_r$ terms which parition the set $\{a_1,...,a_n\}$. Thus, we proceed until the process terminates. This process must terminate since $n$ is the product of only a finite number of primes.
Our final result will be of the form:

\begin{equation}
\sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \frac{p_1...p_g}{1 + (a_1...a_{p_1...p_g})^{\frac{1}{p_1...p_g}}} = \frac{n}{1 + (a_1...a_{n})^{\frac{1}{n}}}
\end{equation}

Thus, this proves the lemma.

Now we shall show the inequality holds for all prime numbers. \newline

\textbf{Proposition: } The inequality holds for all $n \in  \mathbb{P}$. \newline

\textbf{Proof: } We shall use the property that the inequality holds for all $n = 2^k$ for some $k \in \mathbb{N}$.

Let $m$ be a power of 2 such that $m > n$. Observe the following modification to the original sum as follows.

\begin{gather}
  \sum_{i=1}^{n} \frac{1}{1 + a_i} + \sum^{m-n} \frac{1}{1 + (a_1...a_n)^{\frac{1}{n}}} 
\end{gather}

Since $(a_1...a_n)^{\frac{1}{m-n}} \geq 1$ and there are $m$ terms in the sum, we can apply the inequality:

\begin{gather}
\sum_{i=1}^{n} \frac{1}{1 + a_i} + \frac{m-n}{1 + (a_1...a_n)^{\frac{1}{n}}} \geq \frac{m}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} \\ 
\sum_{i=1}^{n} \frac{1}{1 + a_i} + \frac{m-n}{1 + (a_1...a_n)^{\frac{1}{n}}} \geq \frac{n}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} + \frac{m-n}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} \\
\sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \frac{n}{1 + (a_1...a_n)^{\frac{1}{n}}}
\end{gather}

\end{document}

### 2021-05-25 Nonlinear/formats2021/abstract.tex
Adding draft for formats2021 paper

Reachable set computation is an important technique for the verification of safety properties of dynamical systems.
%
% one method proposed for solving this problem, based on parallelotope bundle reachability for
%
In this paper, we investigate parallelotope bundle based reachable set computation for discrete nonlinear systems.
%
The algorithm relies on computing an upper bound on the supremum of a nonlinear function over a rectangular domain,
traditionally done using Bernstein polynomials.
%or nonlinear optimization tools.
 We strive to remove the manual step of parallelotope template selection to make the method fully automatic.
%
Further, we show that changing templates dynamically during computations can improve accuracy.
%
% One of the simplest algorithms for computing reachable sets for discrete nonlinear systems uses parallelotope bundles and Bernstein polynomials.
%
%Our main hypothesis in this paper is that generating templates in a dynamic manner would improve the accuracy of the reachable set.
%
We investigate two techniques for generating the template directions.
%
The first technique approximates the dynamics as a linear transformation and generates templates using this linear transformation.
%
The second technique uses Principle Component Analysis (PCA) of sample trajectories for generating templates.
%
We have implemented our approach in a Python based tool called Kaa and improve its performance by two main enhancements.
%
% have implemented two main enhancements to it.
%
The tool is modular and use two types of global optimization solvers, the first using Bernstein polynomials and the second using NASA's Kodiak nonlinear optimization library.
%
Second, we leverage the natural parallelism of the reachability algorithm and parallelize the Kaa implementation.
% thus leveraging the cloud computing resources.
%
We demonstrate the improved accuracy of our approach on several standard nonlinear benchmark systems.
%
% In this paper, we describe Kaa, a terse Python implementation of reachable set computation which leverages the widely used symbolic package \emph{sympy}.
% %
% Additionally, we simplify the user interface and provide easy-to-use plotting utilities.
% %
% %
% We believe that our tool has pedagogical value given the simplicity of the implementation and its user-friendliness.
%

### 2021-05-25 Nonlinear/formats2021/evaluation.tex
Adding draft for formats2021 paper

\vspace{-1em}
\section{Evaluation}
\label{sec:eval}
%\vspace{-0.5em}
 We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} \cite{kim2020kaa}. Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems. The optimization procedure for finding the direction offets is performed through the \emph{Kodiak} library. Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module. To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles. For systems of dimension fewer than or equal to three, we utilize scipy.spatial library's Convex hull routine.
 These routines are esstentially Python wrappers around their corresponding \emph{QHull} procedures. For higher-dimensional systems, we employ the volume of the tightest enveloping box around the intersection of the parallelotopes.
 The dimensions of this box are easily calculated through linear programming. The total volume estimate of the reachable set will be the sum of all the computed bundles' volume estimates.

 %For two-dimensional models, note that there are only five non-trivial templates. The trivial template here refers to the initial box which is defined exactly the axis-aligned directions. Refer to the non-axis aligned templates. At each every time, pca. lin app directions to
\vspace{-1em}
\subsubsection{Model Dynamics}
For benchmarking, we select six non-linear models with polynomial dynamics. Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability analysis with {\bf static} parallelotope bundles. In these cases, we directly compare the performance of our dynamic strategies with the static parallelotopes defined in Sapo. To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box. Here, {\bf diagonal directions} are defined to be the vectors created by adding and subtracting distinct pairs of all unit axis-aligned vectors from each other. By {\bf diagonal parallelotopes}, we refer to parallelotopes which are defined only by axis-aligned and diagonal directions. Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles which solely consists of diagonal parallelotopes. Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box. We refer to these parallelotopes as {\bf non-axis-aligned parallelotopes}.

%Maybe mention something about role of diagonal templates in program verfication here.
\begin{example}
In two dimensions, $\mathbb{R}^2$, we have the two unit axis-aligned directions, $[1,0]^T, [0,1]^T$. The diagonal directions will then

\[ [1,1]^T, \; [1,-1]^T\]
Consequently, the diagonal parallelotopes will be precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
% Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions $[1,0,0]^T, [0,1,0]^T,[0,0,1]^T$. Calculating the diagonal directions yields the following directions:
% \[ [1,1,0]^T, \; [1,-1,0]^T, \; [1,0,1]^T,\;  [1,0,-1]^T,\;  [0,1,1]^T, \; [0,1,-1]^T \]
% Thus, there are ${9 \choose 3} = 84$ diagonal parallelotopes.
%
% Note that for both examples, we do not consider the directions' negative counterparts since we define parallelotopes use both the positive and negative counterparts of a its chosen directions.
%
% For two dimensions, the non-axis-aligned parallelotopes would be all parallelotopes excluding the one defined by directions $[0,1]^T, [1,0]^T$. In particular, there are 5 non-axis-aligned diagonal parallelotopes.
\end{example}
Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional model COVID supermodel is explained in the subsequent subsection below.

\vspace{-1em}
\begin{table}[h!]
%\hspace{-5em}
  \centering
\begin{tabular}{|p{1.5cm}|c|p{1.5cm}|c|c|p{5cm}|}
\hline
Model & Dimension & Parameters & \# steps & $\Delta$ & \hspace{1.5cm}Initial Box \\
\hline
Vanderpol & 2 & - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
\hline
Jet Engine& 2 & - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
\hline
Neuron \cite{fitzhugh1961impulses}& 2 & - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
\hline
SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 & $s \in [0.79,0.8], i \in [0.19,0.2], r = 0$ \\
\hline
Coupled \newline Vanderpol & 4 & - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
\hline
COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & Stated Below\\
\hline
\end{tabular}
\caption{Benchmark models and relevant information}
\label{tab:modeldyns}
\end{table}

\vspace{-2em}
\noindent \textbf{COVID Supermodel:}
%
We further benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}. This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients. These patients have the ability to infect susceptible members with a fixed probability. The following dynamics account for this new group and its interactions with the traditional groups found in the SIR model. The continuous dynamics given in \cite{indiansuper2020supermodel} are discretized through the Euler method.

\begin{align}
  \begin{split}
   S_A & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
   S_I & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
   A & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
   I & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
   R_A & = R_A + (\gamma A)\cdot \Delta \\
   R_I & = R_I + (\gamma I)\cdot \Delta \\
   D & = D + (\eta I)\cdot \Delta
  \end{split}
\end{align}
where the variables denote the fraction of a population of individuals designated as \emph{Susceptible Transition to Asymptomatic $(S_A)$}, \emph{Susceptible Transition to Infected $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}. We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the average probablity of infection, $\gamma$ is the average removal rate, and $\eta$ is the average mortality rate. The parameters are set based on figures shown in \cite{ansumali2020modelling} for the USA. The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.

\subsubsection{Accuracy of Dynamic Strategies}
The results of testing our dynamic strategies against static ones are summarized in Table~\ref{tab:voltable}. For models previously defined in Sapo, we set the static parallelotopes to be exactly those found in Sapo.
If a model is not implemented in Sapo, we simply use the static parallelotopes defined in a model of equal dimension. To address the unavailability of a four-dimensional model implemented in Sapo, we sampled random subsets of five static non-axis-aligned parallelotopes and chose the flowpipe with smallest volume.
%
% As justifiction for this scheme, we wish to remark the time required in testing all possible subsets of diagonal parallelotopes.
%
A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is impossible.

% are $2{n \choose 2}$ diagonal directions for an $n$-dimensional system, giving a total of $n + 2 {n \choose 2}$ axis-aligned and diagonal directions. Thus, there are ${n + 2 {n \choose 2} \choose n} = {n^2 \choose n}$ possible diagonal parallelotopes for an $n$-dimensional system. This explosive growth in search space size compelled us to proceed by sampling the search space instead.

From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks. In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the PCA/LinApp ratio imparts differing effects on the reachable set.  Observe that using parallelotopes defined by Linear Approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5Lin_.png}
\caption{5 Lin}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_1PCA5Lin_.png}
\caption{1 PCA 5 Lin}
\end{subfigure}%

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5PCA_.png}
\caption{5 PCA}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_5PCA1Lin_.png}
\caption{5 PCA 1 Lin}
\end{subfigure}

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_Sapo_.png}
\caption{Sapo}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_Sapo_.png}
\caption{Sapo}
\end{subfigure}
\caption{Effect of varying ratio between the number of PCA and Linear Approximation parallelotopes. The Vanderpol (left) and the FitzHugh-Nagumo Neuron (right) phase plots are shown to illustrate differing effects of varying the PCA/LinApp ratio. The initial set for the Vanderpol model is set to $x \in [0,0.05], \, y \in [1.95,2]$}
\label{fig:PCALinAppRatio}
\end{figure}

\vspace{-1em}
\subsubsection{Performance under Increasing Initial Sets}
A key advantage of our dynamic strategies is the improved ability to control the wrapping error naturally arising from larger initial sets. Figure \ref{fig:InitVolReachComp} presents charts showcasing the effect of increasing initial sets on the total flowpipe volume. We vary the initial box dimensions to gradually increase the box's volume. We then plot the total flowpipe volume after running the benchmark. The same initial boxes are also fed as input into computation using only static parallelotopes borrowed from Sapo. The number of parallelotopes defined by PCA and Linear Approximation directions were chosen based on best performance as seen in Table \ref{tab:voltable}. We remark that our dynamic strategies perform better than static ones in controlling the total flowpipe volume as the initial set becomes larger. On the other hand, the performance of static parallelotopes tends to degrade rapidly as we increase the volume of the initial box.

\subsubsection{Performance against Random Static Templates}
We additionally benchmark our dynamic strategies against static random parallelotope bundles. We sample such parallelotopes in $n$ dimensions by first sampling a set of $n$ directions uniformly on the surface of the unit $(n-1)$-sphere, then defining our parallelotope using these sampled directions. We sample twenty of these parallelotopes for each trial and average the total flowpipe volumes. As shown in Figure \ref{fig:RanStaticStratComp}, our best-performing dynamic strategies consistently outperform static random strategies for all tested benchmarks.

\vspace{-1em}
\section{Conclusions}
In this paper we investigated two techniques for generating templates dynamically, first using linear approximation of the dynamics, and second using PCA.
%
We demonstrated that these techniques improve the accuracy of reachable set by an order of magnitude when compared to static or random template directions.
%
We also observed that both these techniques improve the accuracy of the reachable sets for different benchamrks.
%
In future, we intend to investigate Koopman linearization techniques for computing the template directions~\cite{bak2021reachability}.

\begin{table}[h!]
\hspace{1em}
\begin{subtable}[h]{0.45\textwidth}
     \centering
     \begin{tabular}{|c|c|}
     \hline
     Strategy & Total  Volume \\
     \hline\
     5 LinApp & 0.227911 \\
     \hline
     1 PCA, 4 LinApp& 0.225917 \\
     \hline
     2 PCA, 3 LinApp & 0.195573 \\
     \hline
     {\bf 3 PCA, 2 LinApp} & {\bf 0.188873} \\
     \hline
     4 PCA, 1 LinApp & 1.227753\\
     \hline
     5 PCA & 1.509897 \\
     \hline
     5 Static Diagonal(Sapo) & 2.863307  \\
     \hline
    \end{tabular}
    \caption{Vanderpol}
    \label{tab:vdpvol}
 \end{subtable}\hspace{1em}
 \begin{subtable}[h]{0.45\textwidth}
      \centering
      \begin{tabular}{|c|c|}
      \hline
      Strategy & Total  Volume \\
      \hline
      5 LinApp & 58199.62 \\
      \hline
      1 PCA, 4 LinApp & 31486.16 \\
      \hline
      {\bf 2 PCA, 3 LinApp} & {\bf 5204.09}\\
      \hline
      3 PCA, 2 LinApp & 6681.76 \\
      \hline
      4 PCA, 1 LinApp& 50505.10 \\
      \hline
      5 PCA  & 84191.15 \\
      \hline
      5 Static Diagonal (Sapo) & 66182.18  \\
      \hline
     \end{tabular}
     \caption{Jet Engine}
     \label{tab:enginevol}
  \end{subtable}

  \hspace{1em}
  \begin{subtable}[h]{0.45\textwidth}
       \centering
       \begin{tabular}{|c|c|}
       \hline
       Strategy & Total  Volume \\
       \hline
       5 LinApp  & 154.078\\
       \hline
       1 PCA, 4 LinApp  & 136.089\\
       \hline
       2 PCA, 3 LinApp  & 73.420\\
       \hline
       {\bf 3 PCA , 2 LinApp } & {\bf 73.126} \\
       \hline
       4 PCA, 1 LinApp  & 76.33 \\
       \hline
       5 PCA & 83.896 \\
       \hline
       5 Static Diagonal  (Sapo) & 202.406  \\
       \hline
      \end{tabular}
      \caption{FitzHugh-Nagumo}
      \end{subtable} \hspace{1em}
      \begin{subtable}[h]{0.45\textwidth}
        \centering
        \begin{tabular}{|c|c|}
        \hline
        Strategy & Total  Volume \\
        \hline
        {\bf 2 LinApp } & {\bf 0.001423} \\
        \hline
        1 PCA, 1 LinApp & 0.106546\\

        \hline
        2 PCA  & 0.117347\\
        \hline
        2 Static Diagonal (Sapo) & 0.020894\\
        \hline
       \end{tabular}
       \caption{SIR}
       \label{tab:sirvol}
    \end{subtable}

    \hspace{1em}
    \begin{subtable}[h]{0.45\textwidth}
         \centering
         \begin{tabular}{|c|c|}
         \hline
         Strategy & Total  Volume \\
         \hline
         5 LinApp & 5.5171 \\
         \hline
         {\bf 1 PCA, 4 LinApp } & {\bf 5.2536} \\
         \hline
         2 PCA, 3 LinApp  & 5.6670\\
         \hline
         3 PCA, 2 LinApp  & 5.5824\\
         \hline
         4 PCA, 1 LinApp  & 312.2108 \\
         \hline
         5 PCA  & 388.0513 \\
         \hline
         5 Static Diagonal (Best) & 3023.4463  \\
         \hline
        \end{tabular}
        \caption{Coupled Vanderpol}
        \label{tab:sirvol}
     \end{subtable}\hspace{1 em}
    \begin{subtable}[h]{0.45\textwidth}
         \centering
         \begin{tabular}{|c|c|}
         \hline
         Strategy & Total  Volume \\
         \hline
         3 LinApp & $2.95582227 * 10^{-10}$ \\
         \hline
         {\bf 1 PCA, 2 LinApp } & {\bf $2.33007583 * 10^{-10}$}\\
         \hline
         2 PCA, 1 LinApp &$ 4.02751770 * 10^{-9}$\\
         \hline
         3 PCA & $4.02749571 * 10^{-9}$\\
         \hline
         3 Static Diagonal (Sapo) & $4.02749571 * 10^{-9}$\\
         \hline
        \end{tabular}
        \caption{COVID}
        \label{tab:covidvol}
     \end{subtable}
 \caption{Tables presenting total reachable set volume by strategy. The static directions are retrieved and/or inspired from Sapo models of equal dimension for benchmarking. The best performing strategy is highlighted in bold.}
     \label{tab:voltable}
\end{table}
\newpage
\begin{figure}[h!]
    \hspace{-1.5em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVol.png}
    \caption{Vanderpol}
    \end{subfigure}%
    %\hspace{1em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVol.png}
    \caption{Jet Engine}
    \end{subfigure}

    \hspace{-1.5em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVol200steps.png}
    \caption{Neuron}
    \end{subfigure}%
    %\hspace{1em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVol.png}
    \caption{SIR}
    \end{subfigure}

    \hspace{-1.5em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVol.png}
    \caption{Coupled Vanderpol}
    \end{subfigure}%
    %\hspace{1em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVol.png}
    \caption{COVID}
    \end{subfigure}

    \caption{Comparison between the performance of diagonal static parallelotope bundles and that of the best performing dynamic parallelotope bundles as the volume of the initial set grows.}
    \label{fig:InitVolReachComp}
\end{figure}
\newpage

\begin{figure}[h!]
    \hspace{-1.5em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVolRanStrat.png}
    \caption{Vanderpol}
    \end{subfigure}%
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVolRanStrat.png}
    \caption{Jet Engine}
    \end{subfigure}

    \hspace{-1.5em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVolRanStrat.png}
    \caption{Neuron}
    \end{subfigure}%
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVolRanStrat.png}
    \caption{SIR}
    \end{subfigure}

    \hspace{-1.5em}
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVolRanStrat.png}
    \caption{Coupled Vanderpol}
    \end{subfigure}%
    \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVolRanStrat.png}
    \caption{COVID}
    \end{subfigure}

    \caption{Comparision between random static strategies and the best performing dynamic strategies as the volume of the initial set grows. The total reachable set volumes for random static strategies are averaged over ten trials for each system.}
    \label{fig:RanStaticStratComp}
\end{figure}
\newpage

### 2021-05-25 Nonlinear/formats2021/introduction.tex
Adding draft for formats2021 paper

\section{Introduction}
\label{sec:intro}

One of the widely used techniques for performing safety analysis of nonlinear dynamical systems is to perform reachable set computation.
%
Reachable set is an overapproximation of the set of states visited by all the trajectories of the system starting from an initial set.
%
Computing the reachable set for nonlinear systems is very challenging.
%
This is primarily because of two reasons:
%
First, the tools for performing nonlinear analysis are not very scalable.
%
Second, computing the reachable set using set representations involves wrapping error.
%
That is, the overapproximation acquired at a given step would increase the conservativeness of the reachable set for all future steps.

One of the techniques for computing reachable set for discrete time nonlinear systems is to use Parallelotope bundles.
%
Here, the reachable set is represented as a parallelotope bundle, an intersection of several parallelotopes.
%
One of the advantages of this technique is that it uses a special form of nonlinear optimization problem to compute the reachable set.
%
The usage of a specific form of nonlinear optimization mitigates the drawback involved with the scalability of nonlinear analysis.

However, wrapping error still remains to be a problem for reachability using parallelotope bundles.
%
The template directions for specifying these parallelotopes are provided as an input by the user.
%
Often, these template directions are selected to be either the cardinal axis directions or some directions from octahedral domains.
%
However, it is not clear that the axis directions and octagonal directions are optimal for computing reachable sets.
%
Also, even an expert user of reachable set computation tools might not be able to provide the most suitable set of template directions for computing the most accurate reachable set.
%
Picking an unsuitable template directions would only cause the wrapping error to increase, thus increasing the conservativeness of the safety analysis.

In this paper, we investigate techniques for generating template directions automatically and dynamically.
%
That is, instead of providing the template directions to compute the parallelotope, the user just specifies the number of templates and the algorithm automatically generates the template directions.
%
We study two techniques for generating the template directions.
%
First, we compute a local linear approximation of the nonlinear dynamics and use the linear approximation to compute the templates.
%
Second, we generate a specific set of sample trajectories from the set and use principal component analysis (PCA) over these trajectories.
%
We observe that the accuracy of the reachable set can be drastically improved by using templates generated using these two techniques.
%
For standard nonlinear benchmark systems, we show that generating templates in a dynamic fashion improves the accuracy of the reachable set by two orders of magnitude.
%
We demonstrate that even when the size of the initial set increases, our template generation technique returns much more accurate reachable sets than both manually specified and random template directions.

% [Introduction para]
% \begin{itemize}
% \item Reachable set computation.
% \item Nonlinear dynamics is challenging.
% \item Overapproximation, wrapping error.
% \item Parallelotope bundle reachability.
% \end{itemize}

% [Templates for reachability]
% \begin{itemize}
% \item Often the templates are generated statically.
% \item However, the most appropriate directions for templates that improve the accuracy is unknwon.
% \item Challenges: Wrapping error, cannot predict.
% \item Static template directions often aggrevate such errors.
% \item Generating templates dynamically is a challenge.
% \end{itemize}

% [Template generation]
% \begin{itemize}
% \item Template based overapproximation are used extensively in verification.
% \item More on this in the related work section.
% \item Dynamic templates using PCA have been investigated already.
% \item Not explicitly in the context of parallelotope bundles.
% \item In this paper we propose two techniques for  generating templates dynamically.
% \end{itemize}

\section{Related Work}

Reachable set computation of nonlinear systems using template polyhedra and Bernstein polynomials has been first proposed in~\cite{dang2009image}.
%
In~\cite{dang2009image}, Bernstein polynomial representation is used to compute an upper bound of a special type of nonlinear optimization problem~\cite{garloff2003bernstein}.
%
Several improvements to this algorithm were suggested in~\cite{dang2012reachability,sassi2021reachability} and~\cite{dang2014parameter} extends it for performing parameter synthesis.
%
The representation of parallelotope bundles for reachability was proposed in~\cite{dreossi2016parallelotope} and the effectiveness of using bundles for reachability was demonstrated in~\cite{dreossi2017sapo,dreossi2017reachability}.
%
However, all of these papers used static template directions for computing the reachable set.

Using template directions for reachable set has been proposed in~\cite{sankaranarayanan2008symbolic} and later improved in~\cite{dang2011template}.
%
Leveraging the principle component analysis of sample trajectories for computing reachable set has been proposed in~\cite{stursberg2003efficient,chen2011choice,seladji2017finding}.
%
More recently, connections between optimal template directions for reachability of linear dynamical systems and bilinear programming have been highlighted in~\cite{gronski2019template}.
%
For static template directions, Octahedral domain directions~\cite{clariso2004octahedron} remain a popular choice.

### 2021-05-25 Nonlinear/formats2021/prelims.tex
Adding draft for formats2021 paper

\section{Preliminaries}
\label{sec:prelims}

The state of a system, denoted as $x$, lies in a domain $D \subseteq \reals^n$.
%
A discrete-time nonlinear system is denoted as
\begin{equation}
  x^{+} = f(x)
\label{eq:sys}
\end{equation}
%
where $f: \reals^{n} \rightarrow \reals^{n}$ is a nonlinear function.
%
The trajectory of a system that evolves according to Equation~\ref{eq:sys}, denoted as $\xi(x_0)$ is a sequence $x_0, x_1, \ldots $ where $x_{i+1} = f(x_i)$.
%
The $k^{th}$ element in this sequence $x_k$ is denoted as $\xi(x_0,k)$.
%
Given an initial set $\Theta \subseteq \reals^n$, the \emph{reachable set} at step $k$, denoted as $\Theta_k$ is defined as
\begin{equation}
  \Theta_k = \{ \xi(x,k)\: | \: x \in \Theta\}
\label{eq:reachset}
\end{equation}

A parallelotope $P$, denoted as a tuple $\tup{a, G}$ where $a \in \reals^n$ is called the \emph{anchor} and $G$ is a set of vectors $\{g_1, g_2, \ldots, g_n\}$, $\forall_{1 \leq i \leq n} ~ g_i \in \reals^n$ called \emph{generators}, represents a set
\begin{equation}
P = \{ x\:|\: \exists \alpha_1, \ldots, \alpha_n, \mbox{ such that }  0 \leq \alpha_i \leq 1, x = a + \sum_{i=1}^n \alpha_i g_i \}.
\label{eq:ptope}
\end{equation}

We call this representation as the \emph{generator representation} of the parallelotope.
%
We refer to a generator of a specific parallelotope $P$ using dot notation, for example $P.g_1$.
%
For readers familiar with zonotopes~\cite{girard2005reachability,althoff2010computing}, a parallelotope is a special form of zonotope where the number of generators $n$ equals the dimensionality of the set.
%
One can also represent the parallelotope as a conjunction of half-space constraints. 
%
In half-space representation, a parallelotope is represented as a tuple $\tup{\T, c_{l}, c_{u}}$ where $\T \in \reals^{n \times n}$ are called \emph{template directions} and $c_{l}, c_{u} \in \reals^{n}$ such that $\forall_{1 \leq i \leq n} ~  c_{l}[i] \leq c_{u}[i]$ are called \emph{bounds}. The half-space representation defines the set of states
%
$$
P = \{\: x\: | \: c_{l} \leq \T x \leq c_{u} \}.
$$
%
% The rows in $\Lambda$ are called template directions in the parallelotope.
%
Intuitively, the $i^{th}$ constraint in the parallelotope corresponds to an upper and lower bound on the function $\T_{i} x$.
%
That is, $c_{l}[i] \leq \T_{i}x \leq c_{u}[i]$.
%
The half-plane representation of a parallelotope can be converted into the generator representation by computing $n+1$ vertices $v_1, v_2, \ldots, v_{n+1}$ of the parallelotope in the following way.
%
The vertex $v_1$ is obtained by solving the linear equation $\Lambda x = c_{l}$.
%
The $j+1$ vertex is obtained by solving the linear equation $\Lambda x = \mu_{j}$ where $\mu_{j}[i] = c_{l}[i]$ when $i \neq j$ and $\mu_{j}[j] = c_{u}[j]$.
%
The anchor $a$ of the parallelotope is the vertex $v_1$ and the generator $g_{i} = v_{i+1} - v_{1}$.

\begin{example}
  \label{ex:ptope}
  Consider the xy-plane and the parallelotope $P$ given in half-plane representation as $0 \leq x-y \leq 1$, $0 \leq y \leq 1$.
  %
  This is a parallelotope with vertices at $(0,0)$, $(1,0)$, $(2,1)$, and $(1,1)$.
  %
  In the half-space representation, the template directions of the parallelotope $P$ are given by the directions $[1, -1]$ and $[0, 1]$.
  %
  The half-space representation in matrix form is given as follows:

  \begin{equation}
    \begin{bmatrix} 0 \\ 0 \end{bmatrix} \leq \begin{bmatrix}  1 & -1 \\ 0 &  1 \end{bmatrix}  \begin{bmatrix} x \\ y \end{bmatrix} \leq \begin{bmatrix} 1 \\ 1 \end{bmatrix}. \label{eq:ptopeexample}
\end{equation}

  To compute the generator representation of $P$, we need to compute the \emph{anchor} and the \emph{generators}.
  %
  The anchor is obtained by solving the linear equations $x-y = 0, y = 0$.
  %
  Therefore, the anchor $a$ is the vertex at origin $(0,0)$
  %
  To compute the two generator of the parallelotope, we compute two vertices of the parallelotope.
  %
  Vertex $v_1$ is obtained by solving the linear equations $x - y = 1, y = 0$.
  %
  Therefore, vertex $v_1$ is the vertex $(1,0)$.
  %
  Similarly, vertex $v_2$ is obtained by solving the linear equations $x-y = 0, y = 1$.
  %
  Therefore, $v_2$ is the vertex $(1,1)$.
  %
  The generator $g_1$ is the vector $v_1 - a$, that is $(1,0)- (0,0) = (1,0)$
  %
  The generator $g_2$ is the vector $v_2 - a$, that is $(1,1) - (0,0) = (1,1)$.
  %
  Therefore, all the points in the paralellotope can be written as $(x,y) = (0,0) + \alpha_1 (1,0) + \alpha_2(1,1)$, $0 \leq \alpha_1, \alpha_2 \leq 1$.
  %
  % We can also convert a generator representation of a parallelotope into the half-plane representation.
  %
\end{example}

A parallelotope bundle $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$.
%
The set of states represented by a parallelotope bundle is given as the intersection
\begin{equation}
Q = \bigcap_{i=1}^m P_i.
\end{equation}
%
Often, the various parallelotopes in a bundle share common template directions.
%
In such cases, the conjunction of all the parallelotope constraints in a bundle $Q$ is written as $c_{l}^Q \leq T^Q x \leq c_{u}^Q$.
%
Notice that the number of upper and lower bound half-space constraints in this bundle are stricly more than $n$, i.e., $T^Q \in R^{m \times n}$ where $m>n$.
%
Each parallelotope in such a bundle is represented as a subset of constraints in $c_{l}^Q \leq T^Q x \leq c_{u}^Q$.
%
These types of bundles are often considered in the literature.

Alternatively, in this paper, we consider parallelotope bundles where the consisting parallelotopes do not share template directions.
%
We consider such bundles because our we generate template directions automatically.
%
We propose techniques to generate $n$ template directions at each instance.
%
% collectively generate one set of template directions.

The basic building block in this work is a conservative overapproximation to a constrained nonlinear optimization problem with a box domain.
%
Consider a nonlinear function $h : \reals^n \rightarrow \reals$ and the optimization problem denoted as $\mathsf{optBox(h)}$ as

\begin{eqnarray}
  \max ~ h(x) \label{eq:maxsup}\\
  s.t. ~~ x \in [0,1]^{n}.\nonumber
\end{eqnarray}

For computing the reachable set of a nonlinear system, we need an upper bound for the optimization problem.
%
Several techniques using interval arithmetic and Bernstein polynomials have been developed in the recent past~\cite{garloff2003bernstein,munoz2013formalization,kodiak,smith2009fast}.
%

### 2021-05-25 Nonlinear/formats2021/theory.tex
Adding draft for formats2021 paper

\section{Reachability Algorithm}
\label{sec:theory}

In this work, we develop parallelotope reachability algorithms that are \textbf{automatic} with \textbf{dynamic} parallelotopes.
%
The state of the art, in contrast, is \textbf{manual}, where the user specifies a set of parallelotope directions at the start of the problem.
%
Further, the parallelotopes are \textbf{static}, and do not change during the course of the computation.
%
In our evaluation in the next section, we will demonstrate the benefits of these change on a series of nonlinear example systems.
%
In this section, we detail the modifications to the algorithm present correctness arguments.

\subsection{Manual Static Algorithm}
\label{sec:manualstatic}

We first present the original algorithm where the user manually specifies the number of parallelotopes and a set of directions for each parallelotope that remain static throughout the computation.

Recall the system is $n$-dimensional with dynamics function $f: \reals^{n} \rightarrow \reals^{n}$.
%
The parallelotope bundle $Q$ is specified as a collection of $m$ template directions $\T^{Q} \in \reals^{m \times n} (m > n)$ and the set of constraints that define each of the consisting parallelotope.
%
% consisting of $m$ parallelotopes , the user should manually provide $n$ generator directions for each parallelotope.
%
% We write the collective set of generator directions as $\mathcal{G}$, which is a set of sets, each of which contains $n$ direction vectors.
%
Another input to the algorithm is the initial set, given as a parallelotope $P_0$.
%
When the initial set is a box, $P_0$ consists has axis-aligned template directions.
%
The output of the algorithm is, for each step $k$, the set $\overline\Theta_k$, which is a conservative overapproximation of the reachable set at step $k$, $\Theta_k \subseteq \overline\Theta_k$.

The high-level pseudo-code is written in Algorithm~\ref{alg:old}.
%
The algorithm simply calls \tbundle for each step, producing a new parallelotope bundle based on the parallelotope bundle from the previous step.
%
To compute the image of $Q$, the algorithm computes the upper and lower bounds of $f(x)$ with respect to each template direction.
%
Since computing the maximum value of $f(x)$ along each template direction on $Q$ is computationally very hard, the algorithm instead computes the maximum value over each of the consisting parallelotopes and uses the mimimum of all these maximum values.
%
The \tbundle operation works as follows.
%
Consider a parallelotope $P$ in the bundle $Q$.
%
From the definition, it follows that $Q \subseteq P$.
%
% and half-space representation as $\tup{\T, c_{l}, c_{u}$.
Given a template direction $\T_i$, the maximum value of $\T_{i} f(x)$ for all $x \in Q$ is less than or equal to the maximum value of $\T_{i} f(x)$ for all $x \in P$.
%
Similar argument holds for the mimum value of $T_{i} f(x)$ for all $x \in Q$.
%
% We would like to compute the image of the parallelotope $P$ with respect to the nonlinear transformation $f(x)$.
%
% That is, we would like to compute $P'$ such that $\forall x \in P, f(x) \in P'$.
%
% Given that current techniques use static template directions, the parallelotope $P'$ also uses the template directions $\T^{P}$.
%
To compute the upper and lower bounds of each template direction $\T_{i} f(x)$, for all $x \in P$, we perform the following optimization.
%
\begin{eqnarray}
  \max ~ \T_i^{P} \cdot f(x) \label{eq:maxf}\\
  s.t. ~~ x \in P.\nonumber
\end{eqnarray}

Given $P$ is a parallelotope, all the states in $P$ can be expressed as a vector summation of anchor and scaled generators.
%
Let  $\tup{a, G}$ be the generator representation of $P$.
%
The optimization problem given in Equation~\ref{eq:maxf} would then transform as follows.

\begin{eqnarray}
  \max ~ \T_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i) \label{eq:maxalpha}\\
  s.t. ~~ \overline\alpha \in [0,1]^{n}.\nonumber
\end{eqnarray}

Equation~\ref{eq:maxalpha} is a form of $\mathsf{optBox(\T_{i} \cdot f)}$ over $[0,1]^n$.
%
For computing an upper-bound to the constrained nonlinear optimization by invoking one of the Bernstein polynomial or interval arithmetic based methods.
%
Similarly, we compute the lowerbound of $\T_{i}f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \T_{i}f(x)$.
%

We iterate this process (i.e., computing the upper and lower bound of $T_{i}f(x)$) for each parallelotope in the bundle $Q$.
%
% Observe that the role of the parallelotopes in the bundle is to compute an upper bound on the maximum value of $T_{i}f(x)$ by expanding the domain from $Q$ to the corresponding parallelotope $P$.
%
Therefore, the tightest upper bound on $T_{i}f(x)$ over $Q$ is the least of the upper bounds computed from each of the parallelotopes.
%
Similar argument holds for lower bounds of $T_{i}f(x)$ over $Q$.
%
Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving several constrained nonlinear optimization problems.

\begin{lemma}
\label{lem:correctness}
The parallelotope bundle $Q'$ computed using \tbundle (Algorithm~\ref{alg:old}) is a sound overapproximation of the image of bundle $Q$ w.r.t the dynamics $x^{+} = f(x)$.
\end{lemma}
% \begin{proof}
%   Suppose that parallelotope $P$ is part of the bundle $Q$ and is provided in generator representation as $\tup{a, G}$.
%   %
%   By definition, we know that $Q \subseteq P$.
%   % in generator representation and as $\tup{\T, c_{l}, c_{u}}$ in half-space representation.
%   %
%   Consider an $x \in Q$ and $x' = f(x)$.
%   %
%   Given a template direction $\T_{i}$, we know that $\T_{i} x' \leq \mathsf{optBox(\T_{i} \cdot f)}$ (Equation~\ref{eq:maxalpha}) and $-1\times optBox(-1\times \T_{i} \cdot f) \leq \T_{i} x'$.
%   %
%   Therefore, for all $x \in Q$, $c_{l}'[i] \leq \T_{i} \cdot f(x) \leq c_{u}'[i]$.
%   %
%   Therefore, for all $x \in Q$, $f(x) \in Q'$.
% \end{proof}

% \textcolor{red}{\textbf{SRIDHAR: add this (as well as argument for correctness)}}

\input{algorithm_old}
\vspace{-1.5em}
\subsection{Automatic Dynamic Algorithm}

The proposed automatic dynamic algorithm does not require the user provide the set of template direction $\T$; instead it creates templates automatically at each step based.
%
We use two techniques to generate such template directions, first: computing local linear approximations of the dynamics and second, perform principle component analysis (PCA) over sample trajectories.
%
To do this, we first sample a set of points in the parallelotope bundle called \emph{support points}, and propagate them to the next step using the dynamics function $f$.
%
Support points are a subset of the vertices of the parallelotope that either maximize or minimize the template directions.

Intuitively, linear approximations can provide good approximations when the dynamics function is a time-discretization of a continuous system.
%
In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear function.
%
We use the support points as a data-driven approach to find the best-fit linear function to use.
%
If the dynamics of a system is linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \T x \leq c_{u}$, is the set $c_{l} \leq \T \cdot A^{-1} x \leq c_{u}$.
%
Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the nonlinear dynamics (from the support points) and change the template directions by multiplying it with the inverse of the approximate linear dynamics.
%
The second technique for generating template directions performs principle component analysis over the images of the support points.
%
Using PCA is a reasonable choice as it produces orthonormal directions that can construct a rotated box for bounding the points.
%

Observe that in general, the dynamics is nonlinear and therefore, the reachable set could non-convex.
%
However, a parallelotope bundle is always a convex set.
%
Therefore, one way to improve accuracy of this representation is to consider more template directions.
%
For this purpose, we use a notion of \emph{template lifespan}, where we use the linear approximation and/or PCA template directions not only from the current step, but also from the previous $L$ steps.
%
We will demonstrate the effectiveness and tune each of the options (PCA / linear approximation as well as lifespan option) in our evaluation.

The new approach is given in Algorithm~\ref{alg:new}.
%
In this algorithm, instead of fixing the set of templates, we compute one set of templates (that is, a collection of $n$ template directions), using linear approximation of the dynamics and PCA.
%
The algorithm makes use of helper function \texttt{hstack}, which converts a bunch of column vectors into a matrix (as shown in Equation~\ref{eq:ptopeexample} provided in Example~\ref{ex:ptope}).
%
The notation $M_{*,i}$ is used to refer to the $i^{th}$ column of matrix $M$.
%
The \texttt{Maximize} function takes in a parallelotope bundle $Q$ and direction vector $v$ (one of the template directions), and returns the point $p \in Q$ that maximizes the dot product $v \cdot p$ (for computing support points).
%
This can be computed efficiently using linear programming.
%
The \texttt{ApproxLinearTrans} function computes the best approximation of a linear transformation given a list of points before and after the one-step transformation $f$.
%
More specifically, given a matrix $X$ of points before applying the transformation $f$, a matrix of points after the transformation $X'$, we perform a least-squares fit for the linear transition matrix $A$ such that $X' \approx AX$.
%
This can be computed by $A = X' X^\dagger$, where $X^\dagger$ is the Moore-Penrose pseudoinverse of $X$.
%
The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points.
%
Finally, \texttt{TransformBundle} is the same as in Algorithm~\ref{alg:old}.

\input{algorithm_new}

Algorithm~\ref{alg:new} computes the dynamic templates for each time step $k$.
%
Line~\ref{ln:linearapprox} computes the linear approximation of the nonlinear dynamics and this linear approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
%
The PCA directions of the images of support points is computed in line~\ref{ln:pca}.
%
For the time step $k$, the linear and PCA templates are given as $\T_{k}^{lin}$ and $\T_{k}^{pca}$, respectively.
%
To improve the accuracy of the reachable set, we compute the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the lifespan $L$.
%
This is because, the nonlinear systems we deal with are often not chaotic in nature.
%
Therefore, the templates at time step $k-1$ are very similar to that of template directions at time step $k$.
%

### 2021-07-09 Nonlinear/formats2021/abstract.tex
Submitted version of formats2021 paper

@@
 %
-In this paper, we investigate parallelotope bundle based reachable set computation for discrete nonlinear systems.
+In this paper, we investigate reachable set computation for discrete nonlinear systems based on parallelotope bundles.
 %
-The algorithm relies on computing an upper bound on the supremum of a nonlinear function over a rectangular domain,
-traditionally done using Bernstein polynomials.
+The algorithm relies on computing an upper bound on the supremum of a nonlinear function over a rectangular domain, which has been traditionally done using Bernstein polynomials.
 %or nonlinear optimization tools.
@@
 %
-Further, we show that changing templates dynamically during computations can improve accuracy.
+Furthermore, we show that changing templates dynamically during computations cans improve accuracy.
 %
@@
 %
-We investigate two techniques for generating the template directions.
+To this end, we investigate two techniques for generating the template directions.
 %
@@
 %
-The second technique uses Principle Component Analysis (PCA) of sample trajectories for generating templates.
+The second technique uses Principal Component Analysis (PCA) of sample trajectories for generating templates.
 %
-We have implemented our approach in a Python based tool called Kaa and improve its performance by two main enhancements.
+We have implemented our approach in a Python-based tool called Kaa and improve its performance by two main enhancements.
 %

### 2021-07-09 Nonlinear/formats2021/evaluation.tex
Submitted version of formats2021 paper

@@
 To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize scipy.spatial library's Convex hull routine.
- These routines are esstentially Python wrappers around their corresponding \emph{QHull} procedures.
-For higher-dimensional systems, we employ the volume of the tightest enveloping box around the intersection of the parallelotopes.
- The dimensions of this box are easily calculated through linear programming. The total volume estimate of the reachable set will be the sum of all the computed bundles' volume estimates.
+For systems of dimension fewer than or equal to three, we utilize scipy's convex hull routine.
+ %
+For higher-dimensional systems, we employ the volume of the tightest enveloping box around the parallelotope bundle.
+ %
+The total volume estimate of the overapproximation will be the sum of all the bundles' volume estimates.
 
@@
 For benchmarking, we select six non-linear models with polynomial dynamics.
-Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability analysis with {\bf static} parallelotope bundles.
-In these cases, we directly compare the performance of our dynamic strategies with the static parallelotopes defined in Sapo.
+Benchmarks against more general dynamics can be found in the appendix of the \href{https://arxiv.org/abs/2105.11796}{expanded verison}.
+Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability with {\bf static} parallelotope bundles.
+In these cases, we directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes.
 To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
-Here, {\bf diagonal directions} are defined to be the vectors created by adding and subtracting distinct pairs of all unit axis-aligned vectors from each other.
-By {\bf diagonal parallelotopes}, we refer to parallelotopes which are defined only by axis-aligned and diagonal directions.
-Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles which solely consists of diagonal parallelotopes.
+Here, {\bf diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
+By {\bf diagonal parallelotopes}, we refer to parallelotopes defined only by axis-aligned and diagonal directions.
+Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles solely consisting of diagonal parallelotopes.
 Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
 Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
+%
 For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box.
@@
 \begin{example}
-In two dimensions, $\mathbb{R}^2$, we have the two unit axis-aligned directions, $[1,0]^T, [0,1]^T$. The diagonal directions will then
-
+In two dimensions, $\mathbb{R}^2$, we have the two unit axis-aligned directions, $[1,0]^T, [0,1]^T$. The diagonal directions will then be
 \[ [1,1]^T, \; [1,-1]^T\]
-Consequently, the diagonal parallelotopes will be precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
+Consequently, the diagonal parallelotopes will precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
 % Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions $[1,0,0]^T, [0,1,0]^T,[0,0,1]^T$. Calculating the diagonal directions yields the following directions:
@@
 \end{example}
-Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional model COVID supermodel is explained in the subsequent subsection below.
+Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional COVID supermodel is explained in the subsequent subsection below.
 
@@
   \centering
-\begin{tabular}{|p{1.5cm}|c|p{1.5cm}|c|c|p{5cm}|}
+\begin{tabular}{|p{1.5cm}|c|p{1.7cm}|c|c|p{5cm}|}
 \hline
@@
 \hline
-Vanderpol & 2 & - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
+Vanderpol & 2 & \quad \quad \; - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
 \hline
-Jet Engine& 2 & - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
+Jet Engine& 2 & \quad \quad \; - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
 \hline
-Neuron \cite{fitzhugh1961impulses}& 2 & - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
+Neuron \cite{fitzhugh1961impulses}& 2 & \quad \quad \; - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
 \hline
@@
 \hline
-Coupled \newline Vanderpol & 4 & - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
+Coupled \newline Vanderpol & 4 & \quad \quad \; - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
 \hline
-COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & Stated Below\\
+COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & \quad \quad \quad \; \; Stated Below\\
 \hline
@@
 %
-We further benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
+We benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
 This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients.
-These patients have the ability to infect susceptible members with a fixed probability.
-The following dynamics account for this new group and its interactions with the traditional groups found in the SIR model.
-The continuous dynamics given in \cite{indiansuper2020supermodel} are discretized through the Euler method.
+These patients can infect susceptible members with a fixed probability.
+The dynamics account for this new group and its interactions with the traditional SIR groups.
 
@@
   \begin{split}
-   S_A & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
-   S_I & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
-   A & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
-   I & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
-   R_A & = R_A + (\gamma A)\cdot \Delta \\
-   R_I & = R_I + (\gamma I)\cdot \Delta \\
-   D & = D + (\eta I)\cdot \Delta
+   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
+   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
+   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
+   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
+   R_A' & = R_A + (\gamma A)\cdot \Delta \\
+   R_I' & = R_I + (\gamma I)\cdot \Delta \\
+   D' & = D + (\eta I)\cdot \Delta
   \end{split}
 \end{align}
-where the variables denote the fraction of a population of individuals designated as \emph{Susceptible Transition to Asymptomatic $(S_A)$}, \emph{Susceptible Transition to Infected $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
-We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the average probablity of infection, $\gamma$ is the average removal rate, and $\eta$ is the average mortality rate.
-The parameters are set based on figures shown in \cite{ansumali2020modelling} for the USA.
+where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
+We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate.
+The parameters are set based on figures shown in \cite{ansumali2020modelling}.
 The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
 
+\vspace{-1em}
 \subsubsection{Accuracy of Dynamic Strategies}
@@
 %
-A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is impossible.
+A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is infeasible.
 
@@
 From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks.
-In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the PCA/LinApp ratio imparts differing effects on the reachable set.
-Observe that using parallelotopes defined by Linear Approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.
+In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the ratio imparts differing effects.
+Observe that using parallelotopes defined by linear approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.
 
@@
 We then plot the total flowpipe volume after running the benchmark.
-The same initial boxes are also fed as input into computation using only static parallelotopes borrowed from Sapo.
+The same initial boxes are also used in computations using Sapo's static parallelotopes.
 The number of parallelotopes defined by PCA and Linear Approximation directions were chosen based on best performance as seen in Table \ref{tab:voltable}.
@@
 
+\vspace{-1em}
 \subsubsection{Performance against Random Static Templates}
@@
 
-\vspace{-1em}
+\vspace{-0.8em}
 \section{Conclusions}
-In this paper we investigated two techniques for generating templates dynamically, first using linear approximation of the dynamics, and second using PCA.
+In this paper, we investigated two techniques for generating templates dynamically: first using linear approximation of the dynamics, and second using PCA.
 %
@@
 %
-In future, we intend to investigate Koopman linearization techniques for computing the template directions~\cite{bak2021reachability}.
+In future, we intend to investigate Koopman linearization techniques for computing alternative linear approximation template directions \cite{bak2021reachability}.
+%
+We also wish to investigate the use of a massively parallel implementation using HPC hardware such as GPUs for optimizing over an extremely large number of parallelotopes and their template directions.
+This is inspired by the approach behind the recent tool \emph{PIRK} \cite{devonport2020pirk}.
+
+\subsection{Acknowledgements}
+Parasara Sridhar Duggirala and Edward Kim acknowledge the support of the Air Force Office of Scientific Research under award number FA9550-19-1-0288 and FA9550-21-1-0121 and National Science Foundation (NSF) under grant numbers CNS 1935724 and CNS 2038960.
+Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the United States Air Force or the National Science Foundation.
 
@@
      \end{subtable}
- \caption{Tables presenting total reachable set volume by strategy.
+ \caption{Tables presenting upper bounds on the total reachable set volume by strategy.
 The static directions are retrieved and/or inspired from Sapo models of equal dimension for benchmarking.
@@
 \end{table}
-\newpage
+\clearpage
+
 \begin{figure}[h!]
@@
 \end{figure}
-\newpage
+\clearpage
 
@@
 \end{figure}
-\newpage
+\clearpage

### 2021-07-09 Nonlinear/formats2021/introduction.tex
Submitted version of formats2021 paper

@@
 
-One of the widely used techniques for performing safety analysis of nonlinear dynamical systems is to perform reachable set computation.
+One of the most widely-used techniques for performing safety analysis of nonlinear dynamical systems is reachable set computation.
 %
-Reachable set is an overapproximation of the set of states visited by all the trajectories of the system starting from an initial set.
+The reachable set is defined to be the set of states visited by at least one of the trajectories of the system starting from an initial set.
 %
-Computing the reachable set for nonlinear systems is very challenging.
-%
-This is primarily because of two reasons:
+Computing the reachable set for nonlinear systems is challenging primarily due to two reasons:
 %
@@
 %
-That is, the overapproximation acquired at a given step would increase the conservativeness of the reachable set for all future steps.
+That is, the overapproximation acquired at a given step would increase the conservativeness of the overapproximation for all future steps.
 
-One of the techniques for computing reachable set for discrete time nonlinear systems is to use Parallelotope bundles.
+One of the techniques for computing the overapproximation of reachable sets for discrete time nonlinear systems is to use parallelotope bundles.
 %
@@
 %
-One of the advantages of this technique is that it uses a special form of nonlinear optimization problem to compute the reachable set.
+One of the advantages of this technique is its utilization of a special form of nonlinear optimization problem to overapproximate the reachable set.
 %
@@
 %
-Also, even an expert user of reachable set computation tools might not be able to provide the most suitable set of template directions for computing the most accurate reachable set.
+Also, even an expert user of reachable set computation tools may not be able to provide a suitable set of template directions for computing reasonably accurate over-approximations of the reachable set.
 %
-Picking an unsuitable template directions would only cause the wrapping error to increase, thus increasing the conservativeness of the safety analysis.
+Picking unsuitable template directions would only cause the wrapping error to increase, thus increasing the conservativeness of the safety analysis.
 
@@
 %
-We demonstrate that even when the size of the initial set increases, our template generation technique returns much more accurate reachable sets than both manually specified and random template directions.
+We demonstrate that even when the size of the initial set increases, our template generation technique returns more accurate reachable sets than both manually-specified and random template directions.
 
@@
 %
-Several improvements to this algorithm were suggested in~\cite{dang2012reachability,sassi2021reachability} and~\cite{dang2014parameter} extends it for performing parameter synthesis.
+Several improvements to this algorithm were suggested in~\cite{dang2012reachability,sassi2012reachability} and~\cite{dang2014parameter} extends it for performing parameter synthesis.
 %
@@
 %
-Leveraging the principle component analysis of sample trajectories for computing reachable set has been proposed in~\cite{stursberg2003efficient,chen2011choice,seladji2017finding}.
+Leveraging the principal component analysis of sample trajectories for computing reachable set has been proposed in~\cite{stursberg2003efficient,chen2011choice,seladji2017finding}.
 %
@@
 %
-For static template directions, Octahedral domain directions~\cite{clariso2004octahedron} remain a popular choice.
+For static template directions, octahedral domain directions~\cite{clariso2004octahedron} remain a popular choice.

### 2021-07-09 Nonlinear/formats2021/prelims.tex
Submitted version of formats2021 paper

@@
 
-A parallelotope $P$, denoted as a tuple $\tup{a, G}$ where $a \in \reals^n$ is called the \emph{anchor} and $G$ is a set of vectors $\{g_1, g_2, \ldots, g_n\}$, $\forall_{1 \leq i \leq n} ~ g_i \in \reals^n$ called \emph{generators}, represents a set
+A parallelotope $P$, denoted as a tuple $\tup{a, G}$ where $a \in \reals^n$ is called the \emph{anchor} and $G$ is a set of vectors $\{g_1, g_2, \ldots, g_n\}$, $\forall_{1 \leq i \leq n} ~ g_i \in \reals^n$ called \emph{generators}, represents the set
 \begin{equation}
@@
   %
-  To compute the two generator of the parallelotope, we compute two vertices of the parallelotope.
+  To compute the two generators of the parallelotope, we compute two vertices of the parallelotope.
   %
@@
 %
-Notice that the number of upper and lower bound half-space constraints in this bundle are stricly more than $n$, i.e., $T^Q \in R^{m \times n}$ where $m>n$.
+Notice that the number of upper and lower bound half-space constraints in this bundle are stricly more than $n$ in these cases, i.e., $T^Q \in R^{m \times n}$ where $m>n$.
 %
@@
 %
-These types of bundles are often considered in the literature.
+These types of bundles are often considered in the literature. \cite{dang2009image,dang2012reachability,dreossi2017reachability}
 
-Alternatively, in this paper, we consider parallelotope bundles where the consisting parallelotopes do not share template directions.
+Alternatively, we consider parallelotope bundles where the consisting parallelotopes do not share template directions.
 %
-We consider such bundles because our we generate template directions automatically.
-%
-We propose techniques to generate $n$ template directions at each instance.
+We consider such bundles because we generate the $n$ template directions automatically at each step.
 %
@@
 %
-Several techniques using interval arithmetic and Bernstein polynomials have been developed in the recent past~\cite{garloff2003bernstein,munoz2013formalization,kodiak,smith2009fast}.
+Several techniques using interval arithmetic and Bernstein polynomials have been developed in the recent past \newline\cite{kodiak,garloff2003bernstein,munoz2013formalization,smith2009fast}.
 %

### 2021-07-09 Nonlinear/formats2021/theory.tex
Submitted version of formats2021 paper

@@
+\vspace{-1.2em}
 \section{Reachability Algorithm}
@@
 %
-The state of the art, in contrast, is \textbf{manual}, where the user specifies a set of parallelotope directions at the start of the problem.
+The state of the art, in contrast, is \textbf{manual}, where the user specifies a set of parallelotope directions at the beginning.
 %
-Further, the parallelotopes are \textbf{static}, and do not change during the course of the computation.
+The parallelotopes are also \textbf{static} and do not change during the course of the computation.
 %
-In our evaluation in the next section, we will demonstrate the benefits of these change on a series of nonlinear example systems.
-%
-In this section, we detail the modifications to the algorithm present correctness arguments.
+In this section, we detail the modifications to the algorithm and present their correctness arguments.
 
@@
 
-We first present the original algorithm where the user manually specifies the number of parallelotopes and a set of directions for each parallelotope that remain static throughout the computation.
+We first present the original algorithm\cite{dang2012reachability} where the user manually specifies the number of parallelotopes and a set of static directions for each parallelotope.
 
@@
 %
-The parallelotope bundle $Q$ is specified as a collection of $m$ template directions $\T^{Q} \in \reals^{m \times n} (m > n)$ and the set of constraints that define each of the consisting parallelotope.
+The parallelotope bundle $Q$ is specified as a collection of $m$ template directions $\T^{Q} \in \reals^{m \times n} (m > n)$ and the set of constraints that define each of the member parallelotopes.
 %
+
 % consisting of $m$ parallelotopes , the user should manually provide $n$ generator directions for each parallelotope.
@@
 %
-The output of the algorithm is, for each step $k$, the set $\overline\Theta_k$, which is a conservative overapproximation of the reachable set at step $k$, $\Theta_k \subseteq \overline\Theta_k$.
+The output of the algorithm is, for each step $k$, the set $\overline\Theta_k$, which is an overapproximation of the reachable set at step $k$, $\Theta_k \subseteq \overline\Theta_k$.
 
@@
 %
-The algorithm simply calls \tbundle for each step, producing a new parallelotope bundle based on the parallelotope bundle from the previous step.
+The algorithm simply calls \tbundle for each step, producing a new parallelotope bundle computed from the previous step's bundle.
 %
@@
 %
-Since computing the maximum value of $f(x)$ along each template direction on $Q$ is computationally very hard, the algorithm instead computes the maximum value over each of the consisting parallelotopes and uses the mimimum of all these maximum values.
+Since computing the maximum value of $f(x)$ along each template direction on $Q$ is computationally difficult, the algorithm instead computes the maximum value over each of the constituent parallelotopes and uses the minimum of all these maximum values.
 %
@@
 %
-Similar argument holds for the mimum value of $T_{i} f(x)$ for all $x \in Q$.
+Similar argument holds for the minimum value of $T_{i} f(x)$ for all $x \in Q$.
 %
@@
 
-Given $P$ is a parallelotope, all the states in $P$ can be expressed as a vector summation of anchor and scaled generators.
+Given that $P$ is a parallelotope, all the states in $P$ can be expressed as a vector summation of anchor and scaled generators.
 %
@@
 %
-For computing an upper-bound to the constrained nonlinear optimization by invoking one of the Bernstein polynomial or interval arithmetic based methods.
+One can compute an upper-bound to the constrained nonlinear optimization by invoking one of the Bernstein polynomial or interval-arithmetic-based methods.
 %
@@
 %
-Similar argument holds for lower bounds of $T_{i}f(x)$ over $Q$.
+A similar argument holds for lower bounds of $T_{i}f(x)$ over $Q$.
 %
@@
 
-The proposed automatic dynamic algorithm does not require the user provide the set of template direction $\T$; instead it creates templates automatically at each step based.
+The proposed automatic dynamic algorithm does not require the user to provide the set of template directions $\T$; instead it generates these templates directions automatically at each step.
 %
-We use two techniques to generate such template directions, first: computing local linear approximations of the dynamics and second, perform principle component analysis (PCA) over sample trajectories.
+We use two techniques to generate such template directions, first: computing local linear approximations of the dynamics and second, performing principal component analysis (PCA) over sample trajectories.
 %
-To do this, we first sample a set of points in the parallelotope bundle called \emph{support points}, and propagate them to the next step using the dynamics function $f$.
+To do this, we first sample a set of points in the parallelotope bundle called \emph{support points} and propagate them to the next step using the dynamics $f$.
 %
@@
 %
-In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear function.
+In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear transformation.
 %
@@
 %
-Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the nonlinear dynamics (from the support points) and change the template directions by multiplying it with the inverse of the approximate linear dynamics.
+Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the nonlinear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
 %
-The second technique for generating template directions performs principle component analysis over the images of the support points.
+The second technique for generating template directions performs principal component analysis over the images of the support points.
 %
@@
 
-Observe that in general, the dynamics is nonlinear and therefore, the reachable set could non-convex.
+Observe that in general, the dynamics is nonlinear and therefore, the reachable set could be non-convex.
 %
-However, a parallelotope bundle is always a convex set.
+On the other hand, a parallelotope bundle is always a convex set.
 %
-Therefore, one way to improve accuracy of this representation is to consider more template directions.
+To mitigate this discrepancy, we can improve accuracy of this representation by considering more template directions.
 %
@@
 %
-The algorithm makes use of helper function \texttt{hstack}, which converts a bunch of column vectors into a matrix (as shown in Equation~\ref{eq:ptopeexample} provided in Example~\ref{ex:ptope}).
+The algorithm makes use of helper function \texttt{hstack}, which converts column vectors into a matrix (as shown in Equation~\ref{eq:ptopeexample} provided in Example~\ref{ex:ptope}).
 %
@@
 %
-To improve the accuracy of the reachable set, we compute the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the lifespan $L$.
+To improve the accuracy of the reachable set, we compute the overapproximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the lifespan $L$.
 %
-This is because, the nonlinear systems we deal with are often not chaotic in nature.
+%This is because, the nonlinear systems we deal with are often %not chaotic in nature.
 %
-Therefore, the templates at time step $k-1$ are very similar to that of template directions at time step $k$.
+%Therefore, the templates at time step $k-1$ are very similar %to that of template directions at time step $k$.
 %

### 2021-08-05 Complexity/LMNTheorem/paper.tex
Re-compiled LMN document with some minor edits.

@@
 The above properties on characters can be shown through techniques in Representation Theory.
-For instance, we know representations of finite groups are group homomorphisms into the group of linear automorphisms of a suitable finite-dimensional complex vector space $\rho: G \rightarrow GL(\mathbb{C}^n)$.
-The \emph{characters} of $\rho$ are defined as $\chi(x) = Tr(\rho(x))$ where $Tr$ is the trace operator.
+For instance, we know representations of finite groups are group homomorphisms into the group of linear automorphisms of a suitable finite-dimensional complex vector space $\rho: G \rightarrow GL(V)$ where $GL(V)$ denotes the group of invertible linear operators on finite-dimensional complex vector space $V$.
+The \emph{characters} of $\rho$ are defined as $\chi_\rho(x) = Tr(\rho(x))$ where $Tr$ is the trace operator.
 It turns out that the representations of finite abelian groups are one-dimensional i.e isomorphic to $\mathbb{C}$.
@@
 $$ \chi(a+b) = Tr(\rho(a+b)) = Tr(\rho(a)\circ\rho(b)) = \lambda_a \lambda_b = Tr(\rho(a))Tr(\rho(b)) = \chi(a)\chi(b)$$ Note that this relation does not hold for arbitrary finite groups since the trace is generally not multiplicative.
-The genealizations of Fourier Analysis generally lie in \emph{Harmonic Analysis}.
-For more information on the Representation Theory of Finite Groups, see \cite{serre}, \cite{webb}.
+In addition, for arbitrary finite group $G$, there is an inner product defined on the vector space spanned by the characters of \emph{irreducible} representations i.e representations containing no subspace invariant under the action of $G$.
+This inner product is defined as:
+
+\[ \langle \chi_{\rho}, \chi_{\sigma} \rangle =
+\frac{1}{|G|}\sum_{g \in G} \chi_{\rho}(g)\overline{\chi_{\sigma}(g)}\]
+
+for irredudible representations $\rho: G \rightarrow GL(V_\rho)$ and $\sigma: G \rightarrow GL(V_\sigma)$.
+By an orthogonality argument similar to Theorem \ref{orthonormal}, we have that for irreducible representations $\rho, \sigma$, $ \langle \chi_{\rho}, \chi_{\sigma} \rangle = 1$ iff $\rho, \sigma$ denote the same representation else $\langle \chi_{\rho}, \chi_{\sigma} \rangle = 0$.
+For finite abelian group $G$, the one-dimensional representations above are irreducible.
+Thus, for non-trivial character $\chi$,
+%
+ \[ |G|\langle \chi, \chi_0 \rangle = \sum_{g \in G} \chi(g)\chi_0(g) = \sum_{g \in G} \chi(g)  = 0 \]
+%
+where $\chi_0$ is the trivial character of $G$. This gives another proof of Lemma \ref{totalsum}. For more information on the Representation Theory of Finite Groups, see \cite{serre}, \cite{webb}.
 \end{remark}
@@
 \end{remark}
+\newpage
 %    Bibliographies can be prepared with BibTeX using amsplain,

### 2021-08-15 Nonlinear/formats2021/appendix.tex
Started adding InvertPend data to extended arxiv version

\section{Benchmarks with Trigonometric Dynamics}
In this section, we show the efficacy of dynamic strategies on more general sets of benchmarks other than ones restricted to polynomial dynamics. In particular, we wish to demonstrate the improvements reaped by utilizing \emph{Kodiak} over exclusively using Bernstein polynomials to perform the optimization shown in Equation~\ref{eq:maxf} above.

\subsection{Inverted Pendulum Model}
We test our strategies on the well-known Inverted Pendulum model governed by the following dynamics:
\begin{align}
  x' & = y \\
  y' & = -0.2y - \sin{x}
\end{align}
The model is run for 80 timesteps with discretization step $\Delta=0.1$. The initial set is set to be
\[ x \in [0.25, 0.3], \quad y \in [0.25, 0.3] \]
%
We present the same data and plots shown for the benchmarks with polynomial dynamics above.

\begin{figure}[h!]
  \includegraphics[width=\textwidth, height=0.75\textwidth]{figures/InvertPend}
  \caption{Reachable set for the Inverted Pendulum model executed under the parameters presented above.}
\end{figure}

### 2021-08-15 Nonlinear/formats2021/evaluation.tex
Started adding InvertPend data to extended arxiv version

@@
 Benchmarks against more general dynamics can be found in the appendix of the \href{https://arxiv.org/abs/2105.11796}{expanded verison}.
+%
 Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability with {\bf static} parallelotope bundles.
-In these cases, we directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes.
+%
+We choose benchmarks with polynomial dynamics to directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes.
 To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
 Here, {\bf diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
+%
 By {\bf diagonal parallelotopes}, we refer to parallelotopes defined only by axis-aligned and diagonal directions.
@@
 Consequently, the diagonal parallelotopes will precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
-% Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions $[1,0,0]^T, [0,1,0]^T,[0,0,1]^T$. Calculating the diagonal directions yields the following directions:
-% \[ [1,1,0]^T, \; [1,-1,0]^T, \; [1,0,1]^T,\;  [1,0,-1]^T,\;  [0,1,1]^T, \; [0,1,-1]^T \]
-% Thus, there are ${9 \choose 3} = 84$ diagonal parallelotopes.
 %
-% Note that for both examples, we do not consider the directions' negative counterparts since we define parallelotopes use both the positive and negative counterparts of a its chosen directions.
+For two dimensions, the non-axis-aligned parallelotopes would be all parallelotopes excluding the one defined by directions $[0,1]^T, [1,0]^T$. \newline
+
+\noindent Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions
 %
-% For two dimensions, the non-axis-aligned parallelotopes would be all parallelotopes excluding the one defined by directions $[0,1]^T, [1,0]^T$.
-In particular, there are 5 non-axis-aligned diagonal parallelotopes.
+\[ [1,0,0]^T, [0,1,0]^T,[0,0,1]^T \]
+%
+Calculating the diagonal directions yields the following directions:
+%
+\[ [1,1,0]^T, \; [1,-1,0]^T, \; [1,0,1]^T,\;  [1,0,-1]^T,\;  [0,1,1]^T, \; [0,1,-1]^T \]
+%
+Thus, there are ${9 \choose 3} = 84$ diagonal parallelotopes.
+%
+Note that for both examples, we do not consider the directions' negative counterparts since we define parallelotopes use both the positive and negative counterparts of a its chosen directions.
 \end{example}
@@
 \vspace{-2em}
-\noindent \textbf{COVID Supermodel:}
+\noindent \textbf{COVID Supermodel}
 %
@@
 %
-We also observed that both these techniques improve the accuracy of the reachable sets for different benchamrks.
+We also observed that both these techniques improve the accuracy of the reachable sets for different benchmarks.
 %
@@
 %
-We also wish to investigate the use of a massively parallel implementation using HPC hardware such as GPUs for optimizing over an extremely large number of parallelotopes and their template directions.
+We also wish to investigate the use of a massively-parallel implementation using HPC hardware such as GPUs for optimizing over an extremely large number of parallelotopes and their template directions.
 This is inspired by the approach behind the recent tool \emph{PIRK} \cite{devonport2020pirk}.

### 2021-08-15 Nonlinear/formats2021/theory.tex
Started adding InvertPend data to extended arxiv version

@@
 \end{lemma}
-% \begin{proof}
-%   Suppose that parallelotope $P$ is part of the bundle $Q$ and is provided in generator representation as $\tup{a, G}$.
-%   %
-%   By definition, we know that $Q \subseteq P$.
-%   % in generator representation and as $\tup{\T, c_{l}, c_{u}}$ in half-space representation.
-%   %
-%   Consider an $x \in Q$ and $x' = f(x)$.
-%   %
-%   Given a template direction $\T_{i}$, we know that $\T_{i} x' \leq \mathsf{optBox(\T_{i} \cdot f)}$ (Equation~\ref{eq:maxalpha}) and $-1\times optBox(-1\times \T_{i} \cdot f) \leq \T_{i} x'$.
-%   %
-%   Therefore, for all $x \in Q$, $c_{l}'[i] \leq \T_{i} \cdot f(x) \leq c_{u}'[i]$.
-%   %
-%   Therefore, for all $x \in Q$, $f(x) \in Q'$.
-% \end{proof}
+\begin{proof}
+     Suppose that parallelotope $P$ is part of the bundle $Q$ and is provided in generator representation as $\tup{a, G}$.
+%
+By definition, we know that $Q \subseteq P$.
+%
+in generator representation and as $\tup{\T, c_{l}, c_{u}}$ in half-space representation.
+%
+Consider an $x \in Q$ and $x' = f(x)$.
+%
+Given a template direction $\T_{i}$, we know that $\T_{i} x' \leq \mathsf{optBox(\T_{i} \cdot f)}$ (Equation~\ref{eq:maxalpha}) and $-1\times optBox(-1\times \T_{i} \cdot f) \leq \T_{i} x'$.
+%
+Therefore, for all $x \in Q$, $c_{l}'[i] \leq \T_{i} \cdot f(x) \leq c_{u}'[i]$.
+%
+Therefore, for all $x \in Q$, $f(x) \in Q'$.
+\end{proof}
 
@@
 %
-
 Observe that in general, the dynamics is nonlinear and therefore, the reachable set could be non-convex.

### 2021-11-12 Complexity/LMNTheorem/paper.tex
Beginnings of UGC project

@@
 
-\subsection{Characters}
 We introduce some stepping stones defining the tools used in the Fourier Analysis of boolean functions.
@@
 
+\subsection{Characters}
 Let us first define some general concepts for any finite abelian group $G$.
@@
 This inner product is defined as:
-
+%
 \[ \langle \chi_{\rho}, \chi_{\sigma} \rangle =
 \frac{1}{|G|}\sum_{g \in G} \chi_{\rho}(g)\overline{\chi_{\sigma}(g)}\]
-
+%
 for irredudible representations $\rho: G \rightarrow GL(V_\rho)$ and $\sigma: G \rightarrow GL(V_\sigma)$.
@@
 \begin{definition}
-  A function $f:\{0,1\}^m \times \{0,1\}^n \rightarrow \{0,1\}$ is called a \emph{pseudorandom function generator} if there exists no polynomial-time oracle Turing machine which can distinguish between the outputs from a true random oracle versus $f(s,_)$ for some random seed $s \in \{0,1\}^m$.
+  A function $f:\{0,1\}^m \times \{0,1\}^n \rightarrow \{0,1\}$ is called a \emph{pseudorandom function generator} if there exists no polynomial-time oracle Turing machine which can distinguish between the outputs from a true random oracle versus $f(s,-)$ for some random seed $s \in \{0,1\}^m$.
 \end{definition}

### 2021-11-12 Nonlinear/formats2021/Slides/teaser/teaser.tex
Beginnings of UGC project

\documentclass{beamer}
\usepackage{pgfpages}
\usepackage{algorithm2e}
\usepackage{complexity}

\usetheme{Pittsburgh}
\useinnertheme{rectangles}
\usefonttheme{serif}
\newcommand{\T}{\mathcal{T}}

\newcommand{\tup}[1]
           {
             \relax\ifmmode
             \langle #1 \rangle
             \else $\langle$ #1 $\rangle$ \fi
           }

\title{\textbf{Automatic Dynamic Parallelotope Bundles for Reachability Analysis of Nonlinear Systems}}
\subtitle{
  19th International Conference on Formal Modeling and Analysis of Timed Systems}

\date{}

\author{\textbf{Edward Kim}\inst{1} \newline
        Stanley Bak\inst{2} \newline
        Parasara Sridhar Duggirala\inst{1}}
\institute{\inst{1}University of North Carolina at Chapel Hill \newline \inst{2}Stony Brook University}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Reachability with Template Polyhedra}
  \begin{enumerate}
    \item Reachable set computation is an instrumental tool in performing safety analysis over non-linear systems.
    \item One of many techniques in computing the overapproximation of the reachable sets for discrete non-linear systems is to use \textbf{template polyhedra} to bound the reachable set.
    \item We are particularly interested in \emph{parallelotopes} as our template polyhedra.
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Parallelotopes}
  \begin{definition}
    A \emph{parallelotope} in $\mathbb{R}^n$ is represented as a tuple $\langle \mathcal{T}, c_{l}, c_{u} \rangle$ where $\mathcal{T} \in \mathbb{R}^{n \times n}$ are called \emph{template directions} and $c_{l}, c_{u} \in \mathbb{R}^{n}$ such that $\forall_{1 \leq i \leq n} ~  c_{l}[i] \leq c_{u}[i]$ are called \emph{bounds}. The half-space representation defines the set of states
    %
    $$
    P = \{\: x \in \mathbb{R}^n \: | \: c_{l}[i] \leq \mathcal{T}_{i}x \leq c_{u}[i], \; 1 \leq i \leq n \}.
    $$
   \end{definition}
   \begin{definition}
     A \emph{parallelotope bundle} $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$. The set of states represented by a parallelotope bundle is given as the intersection
   \[ Q = \bigcap_{i=1}^m P_i \]
   \end{definition}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Drawbacks}
  \begin{enumerate}
    \item Hitherto only \emph{static} parallelotopes have been considered. In other words, the template directions specifying the parallelotopes are to be given as user input at the beginning of the reachable set computation.
    \item The template directions chosen are generally the axis-aligned, diagonal directions. However, it's not clear that these directions necessary yield good overapproximations.
    \item Since the template directions are set at the beginning, they cannot adapt to the behavior of the dynamics. This could yield overapproximations which are too conservative for any practical use.
  \end{enumerate}
  %Static, diagonal templates are used.
  %Bad for wrapping error as initial sets increase.
\end{frame}

\begin{frame}
  \frametitle{Contributions}
  \begin{enumerate}
    \item We present a method which is both \emph{dynamic} and \emph{automatic}. Our method utilizes the Principal Component Analysis (PCA) and Local Linear Approximations.
    \item  We extend our tool \emph{Kaa} to leverage NASA's \emph{Kodiak} to perform the optimization procedure.
    \item We parallelize our implementation to scale with an increasing number of parallelotopes in our bundles.
  \end{enumerate}
\end{frame}

\end{document}

### 2021-11-12 Nonlinear/formats2021/acknowledge.tex
Beginnings of UGC project

\newpage
\section*{Acknowledgements}
Parasara Sridhar Duggirala and Edward Kim acknowledge the support of the Air Force Office of Scientific Research under award number FA9550-19-1-0288 and FA9550-21-1-0121 and National Science Foundation (NSF) under grant numbers CNS 1935724 and CNS 2038960. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the United States Air Force or National Science Foundation. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the United States Air Force.

### 2021-11-12 QIT/2Designs/final.tex
Beginnings of UGC project

--- 2020-11-23 PHYS790/final.tex
@@
   The diamond norm of an operator $T$ is defined as
-  $$||T||_{\diamond} = \sup_d ||T \otimes I_d ||_{\infty} $$
+  $$||T||_{\diamond} = \sup_d ||T \otimes I_d ||_{\infty} = \sup_d \sup_{X \neq 0} \frac{||(T \otimes I_d)(X)||_1}{||X||_1}$$
 \end{definition}
@@
 Let us introduce some definitions for properties of the Markov chains we are interested in.
-Although we assume familiarity with the definition of a Markov chain along with the its associated transition matrix $P$, we refer the reader to the reference \cite{cover} for greater detail.
-Recall that if $q^{(0)}$ is some initial  distribution over the $\Omega$ then $q^{(t)} = q^{(0)}P^t$.
-The \emph{stationary distribution} of a Markov chain $\pi$ is a probability distribution over $\Omega$ such that $\pi P = \pi$.
+Although we assume familiarity with the definition of a Markov chain along with its associated transition matrix $P$, we refer the reader to the reference \cite{cover} for greater detail.
+Recall that if $q^{(0)}$ is some initial distribution over a finite set of states $\Omega$ then $q^{(t)} = q^{(0)}P^t$ describes the distribution over $\Omega$ after $t \geq 1$ steps.
+The \emph{stationary distribution} of a Markov chain $\pi$ is a distribution over $\Omega$ such that $\pi P = \pi$.
 %
@@
 
-%\subsection{Tying Some Loose Ends}
+\subsection{Tying Some Loose Ends}
+We dedicate this section to addressing some details glossed over in the previous two sections.
+In particular, the vanishing rate of the non-diagonal terms must be handled before considering the proof for Theorem \ref{maintheorem}.
+First, the proof from Lemma \ref{mainlemma} to the more general statement of Corollary \ref{maincorollary} must be given.
+The proof will demonstrate how the more restricted set of initial states used in Lemma \ref{mainlemma} can easily generalize to all initial states, even those with negative diagonal coefficients.
+%
+\begin{proof}[(Proof of Corollary \ref{maincorollary})]
+ Begin by first noting that Lemma \ref{mainlemma} immediately implies the corollary for coeffcients of non-diagonal terms $\gamma(p,q), \; p \neq q$.
+Thus, it suffices to only consider the diagonal terms $\gamma(p,p)$.
+Now from the observations made in Theorem \ref{haarevol} and linearity, we can view the expected coeffcients $\mathbb{E}_{W}\gamma_W(p,p)$ over all $t$-length random circuits $W$ to follow that of a random walk over the diagonal states.
+Let $g_t(p;q)$ be the probability of arriving at state $p,p$ in $t$-steps starting from state $q,q$ via the random walk defined in Theorem \ref{haarevol}.
+This means that $g_0(p,q) = \delta_{p,q}$.
+We can calculate the expected coeffcients as
+ \begin{equation}
+   \gamma_t(p,p) = \sum_{q \neq 0} g_t(p,q)\gamma_0(q,q)
+ \end{equation}
+\end{proof}
+By performing an effective normalization, we can generalize Lemma \ref{mainlemma} to the corollary as follows: Consider first the inequality for the 2-norm
+%
+\begin{align*}
+\sum_{p \neq 0} \left( \gamma_t(p,p) - \frac{\sum_{q \neq 0} \gamma_0(q,q)}{4^n -1}\right)^2
+& = \sum_{p \neq 0} \left(\sum_{q\neq 0} \gamma_0(q,q) \left( g_t(p,q) - \frac{1}{4^n -1} \right)\right)^2 \\
+& \leq \left( \sum_{q \neq 0} |\gamma_0(q,q)|^2 \right) \sum_{q' \neq 0} \sum_{p \neq 0 }\left( g_t(p,q') - \frac{1}{4^n -1} \right)^2 \\
+& \leq  \left( \sum_{q \neq 0} |\gamma_0(q,q)|^2 \right) (4^n -1)\epsilon \\
+& = 4^n\epsilon \cdot Tr(\rho^2) \\
+& \leq 4^n \epsilon
+\end{align*}
 
+The first inequality arises from the Cauchy-Schwartz inequality and rearranging terms.
+The second inequality is Lemma \ref{mainlemma}, and the final inequality is just $Tr(\rho^2) \leq 1$ for any state $\rho$.
+Invoking Lemma \ref{mainlemma} shows that the above inequality holds for $t \geq n(n + \log4^n/\epsilon) = Cn((\log 4 + 1)n + \log1/\epsilon)$ for sufficiently large $C$ chosen in the Lemma.
+Now choose a larger constant $C'$ dominating the righthand expression to prove the desired result.
+The 1-norm case proceeds similarly:
+%
+\begin{align*}
+ \sum_{p \neq 0} \left( \gamma_t(p,p) - \frac{\sum_{q \neq 0} \gamma_0(q,q)}{4^n -1}\right)
+ & = \sum_{p \neq 0} \left|\sum_{q\neq 0} \gamma_0(q,q) \left( g_t(p,q) - \frac{1}{4^n -1} \right) \right| \\
+ & \leq \sum_{q \neq 0}|\gamma_0(q,q)|\sum_{p \neq 0} \left| g_t(p,q) - \frac{1}{4^n -1} \right| \\
+ & \leq \sum_{q \neq 0}|\gamma_0(q,q)| 2^n \epsilon \\
+ & \leq 2^n \epsilon
+\end{align*}
+%
+\subsubsection{First and Second Moments Convergence for 1-designs}
+We show that our scheme acts an approximate 1-design in the sense below:
+%
+\begin{lemma} \label{onedesign}
+  \begin{equation}
+    \sum_{p \neq 0} (\mathbb{E}_{W_t}|\gamma_{W_t}(p)|)^2 \leq \epsilon
+  \end{equation}
+  after $t \geq n\log1/\epsilon$
+
+and
+\begin{equation}
+  \sum_{p \neq 0} \mathbb{E}_{W_t}|\gamma_{W_t}(p)| \leq \epsilon
+\end{equation}
+after $t \geq n\log n/\epsilon$
+\end{lemma}
+
+Note that since we are considering 1-designs here, we index through only one $n$-string $p$.
+From subsection \ref{eigenanalysis}, recall that applying a unitary $U \sim \mathbb{U}(4)$ completely randomizes the qubit pair if at least one of them are non-zero i.e take basis operators $\sigma_1,\sigma_2,\sigma_3$ in the Pauli basis.
+This means that if sites $i,j$ are picked, then coefficients of basis states $\gamma(p)$ where $p_i \neq 0$ or $p_j \neq 0$ vanish under $U$.
+In this case, we say that sites $i,j$ have been \emph{hit}.
+For an unitary sampled over a universal distribution $U \sim \mu$, the coefficients shrink by a factor $1 - \Delta$ where $\Delta$ is deemed as a gap for distribution $\mu$.
+Hence, sites which have been hit $r$ times shrink by a factor $(1- \Delta)^r$.
+The case for Haar distribution follows the analysis of the standard coupon collectors problem (see \cite{raghavan} for an introduction).
+However, the case for any universal distribution follows from a similar line of thinking as demonstrated in the proof of Lemma \ref{onedesign}.
+
+\begin{proof}[(Proof of Lemma \label{onedesign})]
+  Let us first focus on the case for the coefficient of a specific basis state $\mathbb{E}_{W_t} |\gamma_{W_t}(p)|$.
+From the discussion above, the coefficient shrinks by a factor of $1 -\Delta$ if a pair with at least one non-zero site is picked.
+The probablity of this event occuring is easily calculated to be
+  $$ 1 - \frac{(n-d)(n-d-1)}{n(n-1)} \geq 1 - \frac{n-d}{n} = \frac{d}{n} $$
+  if $d$ denotes the number of non-zeros contained in $n$-string $p$. This gives us the recurrence inequality:
+  \begin{equation*}
+    \mathbb{E}_{W_t}|\gamma_{W_t}(p)| \leq ((1- \Delta)(d/n) + (1 - d/n))\cdot |\gamma_{W_{t-1}}(p)| = \left(2 - \frac{\Delta d}{n} \right)\cdot |\gamma_{W_{t-1}}(p)|
+  \end{equation*}
+%
+Iterating this $t$ times gives the inequality:
+%
+\begin{align*}
+  \mathbb{E}_{W_t}|\gamma_{W_t}(p)| \leq \left(2 - \frac{\Delta d}{n}\right)^t |\gamma_0(p)| = \left(2 - \frac{\Delta d}{n}\right)^{n \cdot \frac{t}{n}} |\gamma_0(p)| \leq e^{-\Delta d  t/ n} |\gamma_0(p)|
+\end{align*}
+Summing over all $p$ then yields the total inequality
+\begin{equation*}
+  \sum_{p \neq 0} \mathbb{E}_{W_t}|\gamma_{W_t}(p)| \leq \sum_{d = 1}^n e^{-\Delta d t/n} \sum_{d(p) = d} |\gamma_0(p)|
+\end{equation*}
+where $d(p)$ is the number of non-zeros in string $p$. With th-- obvious bound $|\gamma_0(p)| \leq 1$, we give the total bound:
+$$ \sum_{d(p) = d} |\gamma_0(p)| \leq {n \choose d} 3^d $$
+This is easily seen as the number of basis states with exactly $d$ non-zero states since each non-zero site would take values in $\{\sigma_1, \sigma_2, \sigma_3\}$.
+Substituting this bound into the inequality above:
+\begin{align*}
+   \sum_{d = 1}^n e^{-\Delta d t/n} \sum_{d(p) = d} |\gamma_0(p)| \leq {n \choose d} (3e^{-\Delta t/n} )^d \leq (1 + 3e^{-\Delta t/n})^n -1
+\end{align*}
+where the last inequality is an invocation of the Binomial theorem. Setting $t \geq \frac{n}{\Delta} \log\frac{3n}{\epsilon}$ yields
+%
+\begin{equation*}
+  \sum_{p \neq 0} \mathbb{E}_{W_t}|\gamma_{W_t}(p)| \leq \left(1 + \frac{\epsilon}{n} \right)^n - 1 = \mathcal{O}(\epsilon)
+\end{equation*}
+%
+The bound for the 2-norm is dervied in a similar fashion. See \cite{harrlow}, Lemma 4.1 for the corresponding 2-norm bound.
+\end{proof}
+%
+The proof for Lemma \ref{onedesign} can be repurposed to prove rapid vanishing of the non-diagonal terms $\sigma_{p_1} \otimes \sigma_{p_2}$ as presented in the lemma below:
+%
+\begin{lemma}
+  \begin{equation}
+     \sum_{p \neq q} \left( \mathbb{E}_{W_t} |\gamma_{W_t}(p,q)|\right)^2 \leq \epsilon
+  \end{equation}
+  for $t \geq n\log\frac{1}{\epsilon}$
+  and
+  \begin{equation}
+    \sum_{p \neq q}  \mathbb{E}_{W_t} |\gamma_{W_t}(p,q)| \leq \epsilon
+  \end{equation}
+\end{lemma}
+%
+\begin{proof}
+  From the analysis for the second moments given in Section \ref{eigenanalysis}, the cofficients decay by a factor of $1- \Delta$ if $\sigma_{p_{i}} \otimes \sigma_{p_{j}}$ is different from the state $\sigma_{q_{i}} \otimes \sigma_{q_{j}}$.
+Thus, picking at least one differing site will trigger a shrinking of the coefficient $\gamma(p,q)$.
+This motivates the setting of the parameter $d$ in the previous proof to be the number of \emph{non-differing sites}.
+This results in the same inequality
+  $$ \mathbb{E}_{W_t}|\gamma_{W_t}(p,q)| \leq \left(2 - \frac{\Delta d}{n} \right)\cdot |\gamma_{W_{t-1}}(p,q)|$$
+  which yields the identical inequality:
+  \begin{equation*}
+    \sum_{\substack{p \neq q \\ pq \neq 00}} \mathbb{E}_{W_t}|\gamma_{W_t}(p,q)| \leq \sum_{d = 1}^n e^{-\Delta d t/n} \sum_{d(p,q) = d} |\gamma_0(p,q)|
+  \end{equation*}
+  where $d(p,q)$ here would count the number of non-differing sites instead. We bound $\sum_{d(p,q) = d} |\gamma_0(p,q)| \leq {n \choose d} 4^d12^{n-d}$ to arrive at
+  \begin{align*}
+    \sum_{\substack{p \neq q \\ pq \neq 00}} \mathbb{E}_{W_t}|\gamma_{W_t}(p,q)| \leq \sum_{d = 1}^n e^{-\Delta d t/n} {n \choose d} 4^d12^{n-d} =  4^n \sum_{d = 1}^n{n \choose d} 3^{n-d} = 4^n(1 + 3e^{-\Delta t/n})^n -1)
+  \end{align*}
+  Set $t \geq \frac{n}{\Delta}(n\log{4} + \log{1/\epsilon})$ to finish the proof.
+\end{proof}
+%
+
+\begin{proof}[(Proof of Theorem \ref{maintheorem})]
+  We are now in the position to prove the main result in respect to the \emph{diamond norm}. The following string of inequalities hold:
+  \begin{align}
+    ||\mathcal{G}_W - \mathcal{G}_H ||_{\diamond}^2
+    & = \sup_d|| \mathcal{G}_W \otimes I_d - \mathcal{G}_W \otimes I_d||_{\infty} = \sup_{\rho} ||(\mathcal{G}_W \otimes I_n)(\rho) - (\mathcal{G}_H \otimes I_n)(\rho) ||_1^2 \\
+    & \leq (2^{2(2n)})\sup_{\rho} ||(\mathcal{G}_W \otimes I_n)(\rho) - (\mathcal{G}_H \otimes I_n)(\rho) ||_2^2\\
+    & = 2^{4n} \sup_{\rho} ||\sum_{p_1,p_2,p_3,p_4}\gamma_{\rho}(p_1,p_2,p_3,p_4) ( \mathcal{G}_W(\sigma_{p_1} \otimes \sigma_{p_2}) \otimes \sigma_{p_3} \otimes \sigma_{p_4} \\
+    & - \mathcal{G}_H(\sigma_{p_1} \otimes \sigma_{p_2}) \otimes \sigma_{p_3} \otimes \sigma_{p_4} ) ||_2^2
+  \end{align}
+  The first inequality is just the standard inequality $||\rho||_1 \leq d ||\rho||_2$ where $d$ is the dimension of the system $\rho \in \mathbb{C}^{d}$.
+The last equality arises from linearity.
+Recall that
+  $$ \mathcal{G}_{W_t}(\sigma_{p_1}\otimes \sigma_{p_2}) = \sum_{\substack{q_1,q_2 \\ q_1q_2 \neq 00}} g_t(p_1,p_2 ; q_1,q_2) \sigma_{q_1} \otimes \sigma_{q_2} $$
+  over all random circuits of length $t$.
+Here $g_t$ represents the evolution of the coefficients as described by the combination of the analyses for the diagonal coefficients $g_t(p,p;q,q)$ and the non-diagonal ones presented above, culminating in Corollary \ref{maincorollary}.
+This yields
+  \begin{align*}
+     2^{4n} & \sup_{\rho} ||\gamma_{\rho}(p_1,p_2,p_3,p_4) \left( \mathcal{G}_W(\sigma_{p_1} \otimes \sigma_{p_2}) \otimes \sigma_{p_3} \otimes \sigma_{p_4} - \mathcal{G}_H(\sigma_{p_1} \otimes \sigma_{p_2}) \otimes \sigma_{p_3} \otimes \sigma_{p_4} \right) ||_2^2  \\
+     & = 2^{4n} \sup_{\rho} || \sum_{p_3,p_4} \sum_{\substack{p_1,p_2,q_1,q_2 \\ q_1,q_2 \neq 00}}\gamma_{\rho}(p_1,p_2,p_3,p_4) \left( g_t(p_1,p_2;q_1,q_2) - \frac{\delta_{p_1,p_2}\delta_{q_1,q_2}}{2^n(2^n+1)} \right) \sigma_{q_1} \otimes \sigma_{q_2} \otimes \sigma_{p_3} \otimes \sigma_{p_4}   ||^2_2
+  \end{align*}
+  Taking the Hilbert-Schmidt norm and squaring the result kills the non-aligning terms to output
+  \begin{align*}
+  &   2^{4n} \sup_{\rho} \sum_{\substack{p_1,p_2,p_3,p_4,q_1,q_2 \\ q_1,q_2 \neq 0}} \gamma_{\rho}(p_1,p_2,p_3,p_4)^2 \left( g_t(p_1,p_2;q_1,q_2) - \frac{\delta_{p_1,p_2}\delta_{q_1,q_2}}{2^n(2^n+1)} \right)^2 \\
+    & \leq 2^{4n} \sup_{\rho} \sum_{\substack{p_1,p_2,p_3,p_4,q_1,q_2 \\ q_1,q_2 \neq 0}} \gamma_{\rho}(p_1,p_2,p_3,p_4)^2 \epsilon^2 \\
+    & \leq 2^{4n}\epsilon^2
+  \end{align*}
+  The last equality is from Lemma \ref{mainlemma}. This finally shows that
+  \begin{equation}
+    ||\mathcal{G}_W - \mathcal{G}_H ||_{\diamond} \leq 2^{2n}\epsilon
+  \end{equation}
+  when $t \geq Cn(n + \log{1/\epsilon})$. Take the final constant $C'$ to be sufficiently large such that $C'n(n + \log{1/\epsilon}) \geq Cn(n + \log{2^{2n}/\epsilon})$ to finish the proof.
+\end{proof}
 %Finally, we state the proof of Theorem \ref{maintheorem}. \newline
@@
 
+\newpage
 \bibliographystyle{amsplain}

### 2021-11-17 Complexity/UGC/approx.tex
UGC notes half finished

\section{Approximation Algorithms and CSPs}
%3.2 Limits of Approximation – Introduction (TIFR)
%AB - Chapter 11
%Explain intuition.
\subsection{Definitions and Examples}
The theory of approximating $\NP$-hard problems roots itself in the following question: ``Is it possible to efficiently approximate $\NP$-complete problem to some arbitrary degree of accuracy?" Since Cook-Levin result demonstrated that the $\SAT$ decision problem is $\NP$-complete, the question could be rephrased as ``Can we find an efficient algorithhm for $\SAT$ which outputs an assignment which satisfies a $1-\delta$ fraction of the clauses for any constant $\delta >  0$?" To this end, let us first introduce a few motivating examples which will find utility in our analysis. \newline

Consider $\mathsf{Max3CNF}$, the problem of, given a 3CNF formula $\phi$ as input, outputting an assignment which maximizes the number of clauses satisfied in $\phi$. Given any assignment $x$ to the variables, say as an $n$-bit string while viewing $\phi:\{0,1\}^n \rightarrow \{0,1\}$, there is some $0 \leq \rho_{\phi,x} \leq 1$ representing the fraction of clauses satisfied in $\phi$. Take $\mathsf{Opt}(\phi)$ to be the maximum such value over all such assignments:

\[ \mathsf{Opt}(\phi) = \max_{x \in \{0,1\}^n}\rho_{\phi,x} \]

Naturally, $\mathsf{Max3CNF}$ is $\NP$-hard since its corresponding decision problem $\mathsf{3CNF}$ is $\NP$-complete i.e $\phi$ is satisfiable iff $\mathsf{Opt}(\phi) = 1$.
%
However, we can ask if there exists an algorithm $A$ which outputs an assignment which satisfies at least some $\beta\cdot\mathsf{Opt}(\phi)$ fraction of the clauses of $\phi$ for some $\beta \leq 1$. To formalize this: \newline
%
\begin{definition} \label{Opt3SATDef}
For $\beta \leq 1$, a polynomial-time algorithm $A$ is deemed as a $\beta$-approximation algorithm for $\mathsf{Max3CNF}$ if $A(\phi)$ outputs an assignment which satisifies at least $\beta\cdot\mathsf{Opt}(\phi)$ fraction of $\phi$'s clauses for every 3CNF instance $\phi$. More specifically, the algorithm $A$ outputs some value such that:
\begin{equation}
  \beta\cdot\mathsf{Opt}(\phi) \leq A(\phi) \leq \mathsf{Opt}(\phi)
\end{equation}
If $\mathsf{Opt}(\phi) = 1$, then $\phi$ is said to be \emph{satisfiable}.
\end{definition}

A canonical example of a simple approximation algorithm for $\mathsf{Max3CNF}$ is the following scheme: for every variable in $\phi$ choose with uniform probability an assignment from $\{0,1\}$. The probability of any given clause being satisfied by such a random assignment is $\frac{7}{8}$. Thus
%
\[ \mathbb{E}_{x \in \{0,1\}^n}[\text{\# satisfied clauses of $\phi(x)$}] = \frac{7m}{8} \]
%
where $m$ is the number of satisfiable clauses of $\phi$. This gives us a $\frac{7}{8}$-approximation algorithm for $\mathsf{Max3CNF}$. Here's another example:
%
\begin{example}
  Let $\mathsf{MaxE3Lin}$ define the problem of the following form: let $\xi$ be defined as a system of linear equations over the field $\mathbb{F}_2$ where every equation contains \emph{exactly} $3$ variables from $n$ variables $x_1,\cdots,x_n$. Find an assignment to the variables which maximizes the number of satisified linear equations in $\xi$. An example is the following system over variables $x_1,x_2,x_3,x_4$:
  %
  \begin{equation*}
    \begin{alignedat}{3}
      x_1 & +{} &  x_2 & +{} & x_3 & = 0 \\
      x_1 & +{} &  x_2 & +{} &  x_4 & = 1 \\
      x_1 & +{} & x_3 & +{} & x_4 & = 1
\end{alignedat}
\end{equation*}
\end{example}
%
The $\mathsf{E3Lin}$ represents a linear system of $\mathsf{E}$xactly $\mathsf{3}$ variables. We can extend the notation treated above for $\mathsf{Max3CNF}$ to this situation. In other words, if $\rho_{\xi,x}$ is the fraction of satisfied linear constraints of $\xi$ under assignment $x \in \mathbb{F}_2^n$.
%
\[\mathsf{Opt}(\xi) =\max_{x \in \{0,1\}^n}\rho_{\xi,x}\]
%
A comment regarding this problem. Firstly, if the given E3Lin instance has a guaranteed solution i.e $\mathsf{Opt}(\phi) = 1$, then Gaussian Elimination will always output an assignment which satisfies all linear constaints in polynomial-time. It's more interesting to consider instances where no such solution satisfing all of the constraints exists over $\mathbb{F}_2^n$. In these cases, $\mathsf{Opt}(\xi) = 1 - \epsilon$ for some $\epsilon \leq 1$. This problem also has a fairly simple $\frac{1}{2}$-approximation algorithm: Set all $x_1 = \cdots = x_n  = 0$ or $x_1 = \cdots = x_n  = 1$ depending on which satisfies the most constraints. This scheme must satisfy at least $\frac{1}{2}$ of the constraints for any instance $\xi$. \newline

\begin{example} There exists an Semi-definite Programming (SDP) relaxation for $\mathsf{MaxCut}$.

\end{example}

\subsection{Constraint Satisfaction Problems}

 Generally speaking, it's useful to concretize these problems into a unified framework. The examples presented in the last section have some common interpretation shared between them: they could all be seen as manifestations of Constaint-Satisfaction Problem (CSP):
%
\begin{definition}
  Let $\Omega$ be a finite set deemed as the \emph{domain}. A constraint satisfaction problem (CSP) $\Psi$ over domain $\Omega$ is a finite set of predicates $\psi:\Omega^r \rightarrow \{0,1\}$ where $r$ would be the \emph{arity} of predicate $\psi$. The predicates can be of different arities. The arity of the CSP $\Psi$ would be the maximum of all the arities of the predicates in $\Psi$. \newline

  \noindent An \emph{instance} $\mathcal{I}$ over CSP $\Psi$ is a set of tuples $(S,\psi)$ where if $r$ is the arity of predicate $\psi$, $S = (v_1,\cdots,v_r)$ is some ordered tuple of variables taken from finite set $V$ consisting of variables contained in CSP $\Psi$. These tuples are called the \emph{constaints} of $\mathcal{I}$. In addition, we add the conditon that every variable show up in at least one constraint. Variables which do not appear in any of the constraints can simply be removed. \newline

  \noindent Now given some instance $\mathcal{I}$, an \emph{assignment} $F:V \rightarrow \Omega$ is simply some  map between the variables and the domain. We say that $F$ satisfies a constaint $(S,\psi)$ if $\psi(F(S)) = 1$. For tuple $S = (v_1, \cdots, v_r)$, we define $F(S) = (F(v_1), \cdots, F(v_r))$. Consider the fraction of constraints of satisfied by $F$ in $\mathcal{I}$:
  %
  \begin{equation}
  \mathsf{Val}_{\I}(F) = \mathbb{E}_{(S, \psi) \sim \I}[\psi(F(S))]
  \end{equation}
  By taking the maximum fraction over all such assignments:
  %
  \begin{equation}
  \mathsf{Opt}(\I) = \max_{F: V \rightarrow \Omega} \mathsf{Val}_{\I}(F)
  \end{equation}
\end{definition}

  %
  \begin{example} \hfill{}
    \begin{itemize}
      \item For $\MaxCNF$, the domain would be $\Omega=\{0,1\}$ and the CSP $\Psi$ would be composed of the predicate $\vee_{3}:\{0,1\}^3 \rightarrow \{0,1\}$ just taking the logical ORs of the three input variables. Any instance $\phi$ would be composed of $(S,\vee_{3})$ where $S$ would be the variables inputted into $\vee_{3}$. Hence, an assignment $F$ would be an assignment into the variables found in 3CNF $\phi$, showing that $\mathsf{Opt}(\phi)$ aligns with the definition given in the last section. \newline

      \item For $\ELin$, the domain would be $\Omega = \mathbb{F}_2$ and  $\Psi$ consist of predicates of the form $(x_1,\cdots,x_3) = x_1 + x_2 + x_3$ and $(x_1,\cdots,x_3) = x_1 + x_2 + x_3 + 1$ representing both types of linear constraints found in an system of three-variable equations over $\mathbb{F}_2$. An instance $\xi$ would consist of constraints $(S,\psi)$ where $S$ would be a three variable tuple containing the variables showing up in the linear constraint $\psi$. \newline

      \item For $\mathsf{MaxCut}$, the domain can be defined as $\Omega = \{-1,1\}$ with $\Psi$ set to the simple predicate $\neq:\{-1,1\}^2 \rightarrow \{0,1\}$. This simply tests if the inputted values are not equal. The variable set $V$ of an instance $\I$ would be indexed by the vertices contained in a graph $G=(V,E)$. There is one constraint tuple, $((v_i,v_j), \neq)$ for every edge $(v_i,v_j) \in E$. Thus, an assignment $F:V \rightarrow \{-1,1\}$ would encode a partition of $V$ into a cut $C$ with a constraint becoming satisified if the corresponding edge is contained in $C$. \newline

      \item The CSP for $\mathsf{Max3Color}$, the maximization counterpart for the $\NP$-complete decision problem, $\mathsf{3Color}$, is similarly defined to that of $\mathsf{MaxCut}$ except our domain would be $\Omega = \{0,1,2\}$. This signifies the three possible colors to color any vertex $v \in V$ in an input graph $G = (V,E)$.

    \end{itemize}
  \end{example}

  As implied in the examples above, there is a generic method to formulate a maximization problem in respect to a given CSP $\Psi$:
  %
  \begin{definition}
    For a given CSP $\Psi$, formulate $\mathsf{MaxCSP}(\Psi)$ as the problem: given an instance $\I$, output an assignment $F$ which satisfies the largest number of constraints in $\I$.
  \end{definition}

\subsection{Gap Problems}

The $\NP$-hardness theory frequently relies on Karp reductions from decision problems to decision problems. In light of this, we can tailor optimization problems to related promise problems in the form of so-called \emph{gap problems}.

\begin{definition} \label{promiseDef}
  A \emph{promise problem} is defined as a tuple $(\mathsf{YES}, \mathsf{NO})$ where $\mathsf{YES},\mathsf{NO} \subseteq \Sigma^*$ in respect to some alphabet $\Sigma$. Furthermore, we require that $\mathsf{YES} \cap \mathsf{NO} = \emptyset$ but not necessarily that $\mathsf{YES} \cup \mathsf{NO} = \Sigma^*$.
\end{definition}

\begin{definition} \label{gapDef}
  Given a $\MaxCSP$ problem, we define $\GapMaxCSP{\alpha}{\beta}$ for $\alpha < \beta$ as the promise problem: given an instance $\I$:
  \begin{align}
      \I \in \mathsf{YES} & \iff \mathsf{Opt}(\I) \geq \beta \\
        \I \in \mathsf{NO} & \iff \mathsf{Opt}(\I) < \alpha
  \end{align}
\end{definition}

The $\NP$-hardness of approximation algorithms reduces to that of gap problems as shown in the below observation:

\begin{theorem} \label{GapCSPtoAlgHard}
Suppose $\GapMaxCSP{\alpha}{\beta}$ is $\NP$-hard for CSP $\Psi$, then approximating $\MaxCSP$ to at least an $\frac{\alpha}{\beta}$ factor is $\NP$-hard.
\end{theorem}
%
\begin{proof}
Suppose there exists an algorithm $A$ which can $\frac{\alpha}{\beta}$-approximate $\MaxCSP$. For an instance $\I$ such that $\mathsf{Opt}(\I) \geq \beta$:

\begin{gather*}
A(\I) \geq  \frac{\alpha}{\beta} \cdot \mathsf{Opt}(\I) = \frac{\alpha}{\beta} \cdot \beta =  \alpha \\
\end{gather*}

Else if $\mathsf{Opt}(\I) < \alpha$
\begin{gather*}
A(\I) \leq \mathsf{Opt}(\I) < \alpha
\end{gather*}
by Definition \ref{Opt3SATDef} adapted to $\MaxCSP$ instances. Hence, the algorithm can decide  $\GapMaxCSP{\alpha}{\beta}$ by checking it's outputted value in respect to $\alpha$.
\end{proof}
%
This in particular demonstrates that showing the hardness of approximating a particular problem is equivalent to showing the hardness of its corresponding gap problem.

%It was shown early on that traditional Karp reductions were unlikely to produce the gaps required in Definition \ref{gapDef} in reducing $\SAT$ to such a promise problem.

%Traditional Karp reductions don't work

### 2021-11-17 Complexity/UGC/labelcover.tex
UGC notes half finished

\section{Label-Cover and Projection Games}
%Definition of Label-Cover
We now introduce a problem which manages to provide a natural paradigm for capturing the essence of CSPs and proving inapproximability results. These ``projection games" were introduced by Bellare, Goldreich, and Sudan \cite{bellare1998free}. The $NP$-hardness of the gap problem version of Label Cover was used by H\aa stad to show tight inapproximability results for $\mathsf{Max3SAT}$ and $\ELin$ \cite{haastad2001some}.

\begin{definition}
A \emph{Label Cover (LC) Problem} instance $\mathcal{G}$ is defined by a bipartite graph $(A \sqcup B,E)$, finite alphabets $\Sigma_A, \Sigma_B$, and a set of projections $\pi_e:\Sigma_A \rightarrow \Sigma_B$ for every edge $e \in E$. Define an \emph{assignment} as consisting of two maps $\mathfrak{A}: A \rightarrow \Sigma_A$, $\mathfrak{B}: B \rightarrow \Sigma_B$. An edge $e = (a,b) \in E$ is said to be satisfied by this assignment if the assignment is compatible with projection $\pi_e$:

\begin{equation}
  \pi_e(\mathfrak{A}(a)) = \mathfrak{B}(b)
\end{equation}

The value of this game will be

\begin{equation} \label{optvalLC}
  \mathsf{Opt}(\mathcal{G}) = \max_{(\mathfrak{A},\mathfrak{B})} \mathbb{E}_{e \sim E}[e \text{ satisfied}]
\end{equation}
In other words, the value will be the largest fraction of edges satisfied by any assignment to the vertices. The corresponding gap problem for Label Cover, $\mathsf{Gap}_{\alpha,\beta}\mathsf{LC}$, is defined as the promise problem:
%
\begin{align*}
    \mathsf{YES} & = \{\mathcal{G} \mid \mathsf{Opt}(\mathcal{G}) \geq \beta\} \\
    \mathsf{NO} & = \{\mathcal{G} \mid \mathsf{Opt}(\mathcal{G}) < \alpha \}
\end{align*}

In the case of perfect completeness, we abbreviate $\mathsf{Gap}_{\alpha,1}\mathsf{LC}$ as simply $\mathsf{Gap}_{\alpha}\mathsf{LC}$.
\end{definition}

There are a few observations worthy of mentioning here. The first regards a type of equivalence between CSP instances and Label Cover instances. Specifically, let $\I$ be an instance of a given CSP $\Psi$ over domain $\Omega$. We can translate this CSP instance into a Label Cover instance as follows: Let the left-hand partition $A$ of our bipartite graph be indexed by the set of constraint tuples $(S,\psi)$ and the right-hand partition $B$ be indexed by the variables of the CSP $V$. Draw an edge from a constraint tuple $(S,\psi)$ to a variable $v$ if that variable appears in $S$. Set $\Sigma_A = \Omega^r, \Sigma_B = \Omega$ where $r$ is the arity of the CSP, and for every edge $e = ((v_1,\cdots,v_r), \psi), v)$ define the projection $\pi_e:\Sigma_A \rightarrow \Sigma_B$ to be

\[ \pi_e(\omega_1, \cdots, \omega_r) =  \omega_i \text{ if } v_i = v\]\newline

%Finish this.

On the other hand, every Label Cover instance can be seen as a $2$CSP over a sufficiently large domain: the predicates of the CSP would be all $2$-ary predicates $\pi:\Sigma_A \times \Sigma_B \rightarrow \{0,1\}$ representing every possible map from $\Sigma_A \rightarrow \Sigma_B$. Thus, the domain of our CSP can be defined as $\Omega = \Sigma_A \cup \Sigma_B$. The corresponding instance of this CSP would be $(S,\pi_e)$ where $\pi_e$ represents the predicate corresponding to the edge $e$'s projection map $\pi_e$ and $S = (a,b)$ would be the vertices of $e$ between $A$ and $B$ respectively.

\begin{theorem} (Weak Projection Games Theorem)
Label Cover is $\NP$-hard to approximate within some constant.
\end{theorem}

%alphabet size.
%examples.
%trivial NP hardness proof.

%\subsection{}
%Arora, Barak
\subsection{Raz's Parallel Repetition Theorem}
To be written...

\begin{theorem} (Projection Games Theorem)
  For every $\epsilon > 0$, there exist alphabets $\Sigma_A, \Sigma_B$ where $|\Sigma_A|,|\Sigma_B| \leq \mathsf{poly}(\frac{1}{\epsilon})$ such that $\mathsf{Gap}_\epsilon\mathsf{LC}$ is $\NP$-hard.
\end{theorem}

### 2021-11-17 Complexity/UGC/pcp.tex
UGC notes half finished

\section{The PCP Theorem}
\subsection{Intuitions}
This section will be centered around the seminal PCP Theorem \cite{arora1998proof}, \cite{arora1998probabilistic}, which characterized $\NP$ in a framework considered unconventional at the time. PCPs, or Probabilistically Checkable Proofs, represent a twist on the idea of $\NP$. Recall that $\NP$ roughly represents the languages which have verifiers which can check proofs of membership in polynomial time. PCPs represent an extension of this definition where the verifier can be \emph{probabilistic} and is granted \emph{random access} to the proof string $\pi$. If we allow the verfier to simply query $\pi$ by outputting a index $i$, it has access to $\pi[i]$. Since we can express an index in $\log{n}$ bits, this in theory gives the verifier access to proof strings of exponential length. To formalize these notions, we begin with definitions:

\begin{definition}
  Given a language $L$ and $r,q: \mathbb{N} \rightarrow \mathbb{N}$, a \emph{(r(n),q(n))-$\PCP$-verifier} for $L$ consists of a polynomial-time algorithm $V$ with the following properties: \newline

  \begin{itemize}
    \item For input strings $x \in \{0,1\}^n, \; \pi \in \{0,1\}^{\leq N}$ for $N = q(n)2^{r(n)}$, $V$ makes $r(n)$ coin flips and decides $q(n)$ queries addresses $i_1, \cdots, i_{q(n)}$ of the proof $\pi$. Based on these queries, it outputs $1$ for ``accept" or $0$ for ``reject". \newline

    \item (Completeness) For $x \in L$, there exists some proof $\pi$ such that $V(x,\pi,r) = 1$ for all random coin tosses $r$. In other words:
    %
    \begin{equation}
      \mathbb{P}_{r}[V(x,\pi,r) = 1] = 1
    \end{equation}

    \item (Soundness) For $x \not\in L$, for all proofs $\pi$:
    %
    \begin{equation}
      \mathbb{P}_{r}[V(x,\pi,r) = 1] \leq \frac{1}{2}
    \end{equation}
  \end{itemize}
  Define the class $\PCP(r(n),q(n))$ as the set of languages $L$ which has a $(c\cdot r(n),d\cdot q(n))$-$\PCP$-verifier for some $c,d > 0$.
\end{definition}

\begin{remark}
  Sometimes the completeness criterion is too strong for our purposes (see the comments on $\ELin$ problem). In these cases, we like to denote the class $\PCP_{\beta, \alpha}(r(n),q(n))$ as the languages $L$ which have a $(r(n),q(n))$-$\NP$-verifier such that the completeness and soundness criteria are amended as below: \newline

  \begin{itemize}
    \item (Completeness) For $x \in L$, there exists some proof $\pi$ such that $V(x,\pi,r) = 1$ for all random coin tosses $r$. In other words:
    %
    \begin{equation}
      \mathbb{P}_{r}[V(x,\pi,r) = 1] \geq \beta
    \end{equation}

    \item (Soundness) For $x \not\in L$, for all proofs $\pi$:
    %
    \begin{equation}
      \mathbb{P}_{r}[V(x,\pi,r) = 1] \leq \alpha
    \end{equation}
  \end{itemize}
The class introduced in the original definition would thus be denoted as $\PCP_{1,\frac{1}{2}}(r(n),q(n))$. PCP verifiers whose completeness constant is one ($\beta = 1$) is deemed as \emph{perfectly complete}.
\end{remark}

The PCP Theorem says that $\NP$ is \emph{exactly} the class of PCPs which uses a \emph{logarithmic} number of random bits and a \emph{constant} number of queries.
%
\begin{theorem} \label{pcptheorem} (The $\PCP$ Theorem \cite{arora1998proof}, \cite{arora1998probabilistic})
%
\begin{equation}
\NP = \PCP_{1,\frac{1}{2}}(O(\log{n}), O(1))
\end{equation}
\end{theorem}

Actually, one direction of this theorem is not too difficult to see:

\begin{proposition}
For every constants $Q \in \mathbb{N}, c > 0$, $\PCP_{1,\frac{1}{2}}(c\cdot\log{n}, Q) \subseteq \NP$
\end{proposition}

\begin{proof}
Begin with the observation that $\PCP_{1,\frac{1}{2}}(r(n), q(n)) \subseteq \NTIME(q(n)2^{r(n)})$. This is justified by the view of an $\NTIME$ machine simulating the verifier by trying all possible coin tosses and queries to the input string $x$ and proof string $\pi$. It can then count all of the accepting paths to determine the probability of acceptance. If $q = O(1)$ and $r = O(\log{n})$, then the right side of the inclusion will be $\NTIME(2^{O(\log{n})}) = \NP$.
\end{proof}

\begin{remark}
  The queries a PCP verifier makes could be \emph{adaptive} or \emph{non-adaptive}. Adaptive queries can be dependent on the outcome of previous queries while non-adaptive queries are independent of one another. The verifiers in these notes will all be non-adaptive for the sake of presentation. The $\PCP$ Theorem still holds when the verifier makes adaptive queries. The only change would be that the proof length would be at most $2^{r(n) + q(n)}$ rather than at most $q(n)2^{r(n)}$.
\end{remark}

%
\subsection{Equivalence of PCP Theorems}
It may be difficult to understand the importance of the PCP Theorem in its form presented in Theorem \ref{pcptheorem}. It turns out there are other equivalent forms of the PCP Theorem more palatable in the context of our goal to prove hardness of approximation results.

\begin{theorem} \label{pcpgapsat}($\PCP$ Theorem: $\mathsf{Gap3SAT}$-hardness)
The problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max3SAT}$ is $\NP$-hard. In other words, for every $\NP$ language $L$, there exists a polynomial-time reduction $f$ mapping $L$ to 3CNF formulas such that:

\begin{align*}
  x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
  x \not\in L & \implies \mathsf{Opt}(f(x)) < \alpha
\end{align*}
\end{theorem}

An immediate consequence of Theorem \ref{pcpgapsat} and Theorem \ref{GapCSPtoAlgHard} is that if there exists an $\alpha$-approximation algorithm for $\mathsf{Max3SAT}$, then $\P = \NP$. With this, we have the first steps towards an inapproximability result: if $\P \neq \NP$, there exists no efficient $\mathsf{Max3SAT}$ algorithm which can approximate better than an $\alpha$ factor. Note that we haven't actually found a concrete value for $\alpha$ yet. This will be addressed once we prove H\aa stad's 3-bit PCP for $\NP$ in a future section.

\begin{theorem} ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) For some constants $q \in \mathbb{N}, \alpha \in (0,1)$, the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard. To elaborate, for every $\NP$ language $L$, there exists a polynomial time reduction mapping an $L$ to a instance $f(x)$ of some CSP $\Psi$ where $\Psi$ consists of $q$-ary predicates, such that

\begin{align*}
  x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
  x \not\in L & \implies \mathsf{Opt}(f(x)) < \alpha
\end{align*}
\end{theorem}

\begin{theorem}
  All the PCP Theorems above are equivalent to each other.
\end{theorem}

Before we embark on the proof, let us establish an equivalence between PCPs and CSPs:

\begin{lemma} (Equivalence between PCPs and CSPs)

\end{lemma}

\begin{proof}
  To be written...
\end{proof}

%2 The PCP Theorem: An Introduction (DIMACS)
%\subsection{Examples of PCPs}

%PCPs and CSPs.
%\subsection{``Naive" PCP for NP}

### 2021-11-17 Complexity/UGC/ugcmaxcut.tex
UGC notes half finished

\section{UG-hardness of MAXCUT}
\subsection{Intuitions}
\subsection{Goemans-Williams Algorithm for MAXCUT}

### 2021-11-17 Complexity/UGC/uniquegames.tex
UGC notes half finished

\section{Unique Games}
\subsection{Definitions}
The PCP Theorem culminated in a proof of the $\NP$-hardness of Label Cover by H\aa stad. Although these results gave proofs of the $\NP$-hardness of $\mathsf{Gap}_{\frac{7}{8} + \epsilon, 1- \epsilon}$-$\mathsf{Max3SAT}$ and $\mathsf{Gap}_{\frac{1}{2} + \epsilon, 1- \epsilon}$-$\mathsf{MaxE3Lin}$, similar hardness proofs for other canonical problems such as $\mathsf{MaxCut}$ didn't seem to follow from these ideas. In his seminal paper, Khot proposed a relaxation of the Label Cover Problem \cite{khot2002power}. The instances of this relaxed version are called \emph{Unique Games}:

\begin{definition}
  A \emph{Unique Label Cover Problem} with $m$ labels (UniqueLC($m$)) instance $\mathcal{U}$ is defined by a bipartite graph $(A \sqcup B,E)$, finite alphabet $\Sigma_A = \Sigma_B = \Sigma$ such that $|\Sigma| = m$, and a set of \emph{permutations} $\pi_e:[m] \rightarrow [m]$ for every edge $e \in E$. Define an \emph{assignment} as consisting of a map $\sigma: A \sqcup B \rightarrow [m]$. An edge $e = (a,b) \in E$ is said to be satisfied by this assignment if the assignment is compatible with projection $\pi_e$:

  \begin{equation}
    \pi_e(\sigma(a)) = \sigma(b)
  \end{equation}

  The value of this game will be

  \begin{equation} \label{optvalLC}
    \mathsf{Opt}(\mathcal{G}) = \max_{\sigma} \mathbb{E}_{e \sim E}[e \text{ satisfied}]
  \end{equation}
  In other words, the value will be the largest fraction of edges satisfied by any assignment to the vertices. The corresponding gap problem for Label Cover, $\mathsf{Gap}_{\alpha,\beta}\mathsf{UniqueLC}(m)$, is defined as the promise problem:
  %
  \begin{align*}
      \mathsf{YES} & = \{\mathcal{G} \mid \mathsf{Opt}(\mathcal{U}) \geq \beta\} \\
      \mathsf{NO} & = \{\mathcal{G} \mid \mathsf{Opt}(\mathcal{U}) < \alpha \}
  \end{align*}

  In the case of perfect completeness, we abbreviate $\mathsf{Gap}_{\alpha,1}\mathsf{UniqueLC}(m)$ as simply $\mathsf{Gap}_{\alpha}\mathsf{UniqueLC}(m)$.
\end{definition}

\begin{theorem} (Unique Games Conjecture \cite{khot2002power})
  For any constant $\delta > 0$, there exists sufficiently large $m \in \mathbb{N}$ such that $\mathsf{Gap}_{\delta,1-\delta}\mathsf{UniqueLC}(m)$ is $\NP$-hard.
\end{theorem}

%Polynomial time algorithm for guaranteed solutions.

### 2021-11-18 Complexity/UGC/hastad.tex
First draft of UGC paper

\section{H\aa stad's 3-bit PCP}
%Arora, Barak
\subsection{Aside on Dictatorship Testing}
To be written...
\subsection{Long Code}
To be written...

### 2021-11-18 Complexity/UGC/pcp.tex
First draft of UGC paper

@@
   \end{itemize}
+Here, $\beta$ is the \emph{completeness parameter} while $\alpha$ is the \emph{soundness parameter}.
 The class introduced in the original definition would thus be denoted as $\PCP_{1,\frac{1}{2}}(r(n),q(n))$.
-PCP verifiers whose completeness constant is one ($\beta = 1$) is deemed as \emph{perfectly complete}.
+PCP verifiers whose completeness parameter is one ($\beta = 1$) is deemed as \emph{perfectly complete}.
 \end{remark}

### 2021-11-18 Complexity/UGC/ugcmaxcut.tex
First draft of UGC paper

@@
+\newcommand{\GapMaxCut}{\mathsf{Gap}_{\GWConstant + \epsilon, \frac{1-\rho}{2} - \epsilon}\mathsf{MaxCut}}
+\newcommand{\ULC}{\mathsf{UniqueLC}(m)}
+\newcommand{\GapULC}{\mathsf{GapUniqueLC}(m)}
+\newcommand{\GapDeltaULC}{\mathsf{Gap}_{\delta, 1-\delta}\ULC}
+
 \section{UG-hardness of MAXCUT}
 \subsection{Intuitions}
-\subsection{Goemans-Williams Algorithm for MAXCUT}
+Recall that the keen insight behind H\aa stad's 3-bit PCP for $\NP$ is the embedding of a dictatorship test within a Label Cover instance.
+In a similar spirit, the proof for the optimality of the Goemans-Williamson algorithm will hinge on crafting a clever dictatorship test embedded within a $\MaxCut$ instance.
+By composing this test with
+
+\subsection{Goemans-Williamson Algorithm for $\MaxCut$}
+Recall that Example \ref{maxcutexample} presented an LP-based approximation algorithm for $\MaxCut$.
+Goemans and Williamson \cite{goemans1995improved} designed an SDP to give an $\alpha_{GW}$-approximation algorithm for $\MaxCut$ where:
+
+\begin{equation}
+  \alpha_{GW} = \min_{-1 \leq \rho \leq 1} \frac{2}{\pi}\frac{\cos^{-1}(\rho)}{1 - \rho} \approx 0.87856
+\end{equation}
+
+\subsection{$\MaxCut$ is $\UG$-hard}
+After introducing the basic background, we are finally ready to show a non-trivial inapproximability result assuming the UGC:
+
+\begin{theorem} \label{maxcutughard} ($\MaxCut$ is $\UG$-hard) The problem $\GapMaxCut$ is $\UG$-hard]
+\end{theorem}
+Note that an immediate corollary of Theorem and the UGC is that the Goemans-Williamson algorithm is tight:
+
+\begin{corollary}
+Assuming the UGC, if an $\alpha_{GW}$-approximation algorithm for $\MaxCut$ exists, then $\P = \NP$.
+\end{corollary}
+%
+\begin{proof}
+Using Theorem \ref{GapCSPtoAlgHard}, we see that for small $\epsilon$:
+\[ \frac{\GWConstant + \epsilon}{\frac{1-\rho}{2} - \epsilon} \approx \frac{2}{\pi}\frac{\cos^{-1}(\rho)}{1 - \rho} \]
+Setting $\rho \approx -0.6934$ yields the desired result.
+\end{proof}
+
+Onwards to the proof of Theorem \ref{maxcutughard}.
+We first reason about the relationship between $\ULC$ instances and $\MaxCut$ instances.
+In particular, we wish to demonstrate that proving $\UG$-hardness for $\MaxCut$ is equivalent to constructing a $2$-query PCP for $\GapULC$ for all $m$.
+
+\begin{theorem}
+  $\mathsf{Gap}_{\alpha,\beta}\mathsf{MaxCut}$ is $\UG$-hard iff there exists a constant $\delta > 0$ such that for all $m$ there exists a 2-query PCP for $\mathsf{Gap}_{\delta, 1-\delta}\ULC$ with completeness $\beta$ and soundness $\alpha$.
+\end{theorem}
+%
+\begin{proof}
+  First assume that for some $\delta > 0$ and all $m$, there exists a polynomial-time reduction from $\GapDeltaULC$ to $\mathsf{Gap}_{\alpha,\beta}\mathsf{MaxCut}$.
+Let $G=(V,E), \mathcal{C}$ be the graph outputted by reduction and the cut outputted by the cut approximation algorithm.
+Construct a PCP samples two vertices from $v_1, v_2 \sim V$ by querying the proof tape encoding $G,\mathcal{C}$ and outputs ``accept" if $(v_1,v_2) \in \mathcal{C}$.
+The completeness and soundness parameters immediately follow from definition of $\mathsf{Gap}_{\alpha,\beta}\mathsf{MaxCut}$.
+Conversely, assume the existance of a 2-query PCP for $\GapDeltaULC$ with the above properties.
+We will calculate the acceptance probabilities of this PCP given a proof string and an input instance by finding the max cut.
+Let the vertices be the proof locations of the input proof string $\pi$.
+It suffcies to draw an edge between two vertices weighted by the probability their corresponding proof locations are queried by the verifier.
+\end{proof}
+
+Thus, we can proceed by constructing a 2-query PCP for $\GapDeltaULC$ with the completeness $\frac{1-\rho}{2} - \epsilon$ and soundness $\GWConstant + \epsilon$.
+As with H\aa stad's 3-bit PCP, we aim to construct a dictatorship test which generates the parameters needed.
+Once again, a proof string $\pi$ containing an assignment of labels to the vertices of a $\ULC$ instance would be encoded into a truth table of a dictator i.e into a \emph{long code}.
+\newline
+
+%
+%
+%
+\subsection{Majority is the Stablest (MIS)}
+%
+Before we craft our dictator test, we require a tool bounding the noise sensitivity of a boolean function with small low-degree influence. This is captured by the ``Majority is the Stablest" theorem:
+
+\begin{theorem} (``Majority is the Stablest" (MIS) \cite{mossel2005noise})
+For every $\epsilon > 0$, $\rho \in (-1,0)$, there exists $\tau > 0$ such that if for all $i \in [n]$, $\mathsf{Inf}_i(f) < \tau$ for function $f:\{-1,1\}^n \rightarrow [-1,1]$, then
+
+\begin{equation}
+  \mathsf{NS}_{\rho}(f) < \GWConstant + \epsilon
+\end{equation}
+\end{theorem}
+
+Recall that the noise sensitivity of a boolean function $f:\{-1,1\}^n \rightarrow \{-1,1\}$ is defined as:
+
+\begin{equation} \label{NSdef}
+  \mathsf{NS}_\rho(f)  = \mathbb{P}_{y \sim N_{\rho}(x), x} [f(x) \neq f(y)]
+\end{equation}
+
+where $y \sim N_\rho(x)$ refers to sampling a string $y$ under the procedure:
+
+\begin{equation*}
+  y_i = \begin{cases}
+        x_i \quad \text{with probabilty} \frac{1 + \rho}{2} \\
+        -x_i \quad \text{with probabilty} \frac{1 - \rho}{2}
+        \end{cases}
+\end{equation*}
+\newline
+
+Now equation \ref{NSdef} can be re-expressed in terms of the noise stability of $f$:
+
+\begin{equation}
+  \mathsf{NS}_\rho(f) = \frac{1}{2} - \frac{1}{2}\mathsf{Stab}_\rho(f)
+\end{equation}
+where
+
+\begin{equation*}
+  \mathsf{Stab}_\rho(f) = \mathbb{E}_{y \sim N_{\rho}(x), x}[f(x)f(y)]
+\end{equation*}
+Through Fourier-analytic techniques, we derive that:
+
+\begin{equation}
+  \mathsf{NS}_\rho(f) = \frac{1}{2} - \frac{1}{2}\sum_{S\subseteq [n]}\hat{f}(S)^2\rho^{|S|}
+\end{equation}
+
+A generalization of the MIS theorem will serve our purposes:
+
+\begin{theorem} \label{generalMIS} (Generalized MIS \cite{khot2007optimal},\cite{mossel2005noise})
+  For all $\epsilon > 0$, $\rho \in (0,1)$, there exists some $\tau > 0$ and finite $d$ such that if $f:\{-1,1\}^n \rightarrow [-1,1]$ and for all $i \in [n]$, $\mathsf{Inf}^{\leq d}_i(f) \leq \tau$:
+  \begin{equation}
+    \frac{1}{2} - \frac{1}{2}\sum_{S\subseteq [n]}\hat{f}(S)^2\rho^{|S|} < \GWConstant + \epsilon
+  \end{equation}
+\end{theorem}
+
+%
+%
+%
+\subsection{The 2-query PCP}
+With the MIS theorem, we can begin our analysis of the promised 2-query PCP. Before the procedure, define $(x \circ \pi)_i = x_{\pi(i)}$ for $x \in \{-1,1\}^n, \; i \in [n], \; \pi \in S_n$.  \newline
+
+\begin{enumerate}
+  \item Sample a vertex $a \in A$ uniformly.
+  \itemsep1em
+  \item Sample two of its neighbors $b,b' \in B$ uniformly. Let $\pi_{b,a}$ and $\pi_{b',a}$ denote the \emph{inverses} of the constraints associated to $(a,b), (a,b')$ respectively.
+  \item Sample $x \sim \{-1,1\}^m$ and $y \sim N_\rho(x)$
+  \item Accept if $f_b(x \circ \pi_{b,a}) \neq f_b'(y \circ \pi_{b',a})$
+\end{enumerate}
+
+For the proof of Theorem \ref{maxcutughard}, invoke Theorem \ref{generalMIS} for $\frac{\epsilon}{2}$ and $\rho$ assumed.
+This will yield a $\tau$ and a degree upper bound $d$.
+Set parameter $$ \delta = \frac{\epsilon \tau^2}{8d} $$
+
+\subsubsection{Completeness}
+%
+Suppose we have a $\ULC$ instance $\mathcal{U}$ such that $\mathsf{Opt}(\mathcal{U}) \geq 1 - \delta$.
+Through a simple union bound argument, the probability that both $\pi_{b,a},\pi_{b',a}$ are satisfied by the assignment is at least $1- 2\delta$.
+Now if both are indeed satisfied and the test accepts, it must be true that:
+%
+\begin{align*}
+    f_b(x \circ \pi_{b,a}) \neq f_b'(y \circ \pi_{b',a}) & \iff
+    (x \circ \pi_{b,a})_{\sigma(b)} \neq (x \circ \pi_{b',a})_{\sigma(b')} \\
+    & \iff x_ {\pi_{b,a}(\sigma(b))} \neq x_ {\pi_{b',a}(\sigma(b'))} \\
+    & \iff x_{\sigma(a)} \neq y_{\sigma(a)}
+\end{align*}
+where for the last equivalence, we invoked the assumption that $\sigma$ satisfies both $\pi_{b,a},\pi_{b',a}$.
+Recall that $\sigma$ is the assignment of labels to the vertices of the biparitite graph.
+The last expression occurs with probabilty $\frac{1 - \rho}{2}$ by step three of the verification algorithm.
+Hence,
+\[ \mathbb{P}[\text{Test accepts}] \geq \frac{(1-2\delta)(1 - \rho)}{2}\]
+
+By observing that $\delta < \frac{\epsilon}{2}$, the inequality above reduces to:
+
+\begin{equation}
+  \mathbb{P}[\text{Test accepts}] \geq \frac{(1-\epsilon)(1 - \rho)}{2} \geq \frac{1 - \rho}{2} - \epsilon
+\end{equation}
+as desired.
+
+\subsubsection{Soundness}
+To show soundness, we prove the contrapositive, namely if
+$$ \mathbb{P}[\text{Test accepts}] \geq \frac{\cos^{-1}(\rho)}{\pi} + \epsilon $$
+then there exists a labeling which satisfies more than a $\delta$ fraction of constraints.
+%
+\begin{proof}
+First:
+\begin{align} \label{step1}
+  \mathbb{P}[\text{Test accepts}] = \mathbb{E}_{a,b,b'}\left[
+  \mathbb{E}_{x,y\sim N_{\rho}(x)}\left[
+  \frac{1}{2} - \frac{1}{2}f_b(x \circ \pi_{b,a})f_{b'}(y \circ \pi_{b',a})\right]\right] \geq \frac{\cos^{-1}(\rho)}{\pi} + \epsilon
+\end{align}
+An averaging argument on the vertex $a \in A$ tells us that, since the test passes with at least $\GWConstant + \epsilon$ probability, there must exist at least $\epsilon/2$ fraction of the vertices in $A$ such that the test conditioned on picking $a$ from this fraction passes with probability at least $\GWConstant + \epsilon/2$.
+Otherwise, if there existed fewer than a $\epsilon/2$ fraction such that the test passed with at least this probability, then the total probability of the test passing is \emph{at most} $(\epsilon/2)\cdot 1 + (1 - \epsilon/2)(\GWConstant + \epsilon/2) < \GWConstant + \epsilon$, which is a contradiction.
+Let us label the vertices picked from this $\epsilon/2$ fraction as \emph{good} vertices.
+\newline
+
+So say we picked one of these good vertices say $a$. Let us define the below function:
+
+\begin{definition}
+  Define $g_a: \{-1,1\}^m \rightarrow [-1,1]$ as
+  \begin{equation}
+    g_a(x) = \mathbb{E}_{b}\left[f_b(x \circ \pi_{b,a}) \right]
+  \end{equation}
+  where the expectation is drawn uniformly over the neighbors $b$ of $a$.
+\end{definition}
+%
+This allows us to re-express the inequality \ref{step1}:
+
+\begin{align}
+  & \mathbb{E}_{b,b'}\left[
+  \mathbb{E}_{x,y\sim N_{\rho}(x)}\left[
+  \frac{1}{2} - \frac{1}{2}f_b(x \circ \pi_{b,a})f_{b'}(y \circ \pi_{b',a})\right]\right] \\
+  =  & \mathbb{E}_{x,y\sim N_{\rho}(x)}\left[
+    \frac{1}{2} - \frac{1}{2}\mathbb{E}_{b}\left[f_b(x \circ \pi_{b,a})\right]\mathbb{E}_{b'}\left[f_{b'}(y \circ \pi_{b',a})\right]\right] \\
+     = &  \frac{1}{2} - \frac{1}{2}\mathbb{E}_{x,y\sim N_{\rho}(x)}\left[g_a(x)g_a(y)\right]\\
+     = & \frac{1}{2} - \frac{1}{2}\mathsf{Stab}_{\rho}(g_a) \\
+   \geq & \frac{\cos^{-1}(\rho)}{\pi} + \epsilon/2
+\end{align}
+By invoking the contrapositive of the generalized MIS theorem (Theorem \ref{generalMIS}) on $g_a$, we deduce that there must exist some index $i_a \in [m]$ such that:
+%
+\begin{equation} \label{contra}
+\mathsf{Inf}_{i_a}^{\leq d}(g_a) \geq \tau
+\end{equation}
+%
+Fortunately, there is a method to equate the Fourier coefficients $g_a$ with those of $f_b$:
+
+\begin{align*}
+  g_a(x) & = \mathbb{E}_{b}\left[ f_b(x \circ \pi_{b,a})\right] \\
+         & = \mathbb{E}_{b}\left[ \sum_{S \subseteq [n]} \hat{f_b}(S)\chi_S(x \circ \pi_{b,a})\right] \\
+         & = \mathbb{E}_{b}\left[\sum_{S \subseteq [n]} \hat{f_b}(S)\chi_{\pi_{b,a}(S)}(x) \right] \\
+         & =  \mathbb{E}_{b}\left[\sum_{S \subseteq [n]} \hat{f_b}(S)\chi_{\pi_{b,a}(S)}(x) \right] \\
+         & = \sum_{S \subseteq [n]}\mathbb{E}_{b}\left[ \hat{f_b}(\pi_{b,a}^{-1}(S))\right]\chi_{S}(x)
+\end{align*}
+where the last equality follows from reparameterizing the sum. By expanding $g_a$ on the left-hand side of the equality into its Fourier expansion, from the orthogonality of characters:
+%
+\begin{equation}
+  \hat{g_a}(S) = \mathbb{E}_{b}\left[ \hat{f_b}(\pi_{b,a}^{-1}(S))\right]
+\end{equation}
+Starting from inequality \ref{contra},
+%
+\begin{align*}
+  \tau \leq \mathsf{Inf}_{i_a}^{\leq d}(g_a)
+  & = \sum_{\overset{S \subseteq [n], i_a \in S}{|S| \leq d}}  \hat{g_a}^2(S) \\
+  & = \sum_{\overset{S \subseteq [n], i_a \in S}{|S| \leq d}}  \left(\mathbb{E}_{b}\left[ \hat{f_b}(\pi_{b,a}^{-1}(S))\right] \right)^2 \\
+ & \leq \sum_{\overset{S \subseteq [n], i_a \in S}{|S| \leq d}}  \mathbb{E}_{b}\left[ \hat{f_b}^2(\pi_{b,a}^{-1}(S))\right] \\
+ & \mathbb{E}_{b}\left[ \sum_{\overset{S \subseteq [n], i_a \in   S}{|S| \leq d}}  \hat{f_b}^2(\pi_{b,a}^{-1}(S))\right] \\
+ & = \mathbb{E}_{b}\left[ \mathsf{Inf}^{\leq d}_{\pi_{b,a}^{-1}(i_a)}(f_b) \right]
+\end{align*}
+The inequality uses Cauchy-Schwarz.
+We use another averaging argument to see that there must exist at least a $\tau/2$ fraction of $a$'s neighbors $b$ such that $$\mathsf{Inf}^{\leq d}_{\pi_{b,a}^{-1}(i_a)}(f_b) \geq \tau/2$$
+otherwise the total influence term would be at most $\tau/2 \cdot 1 + (1 -\tau/2)(\tau/2) < \tau$.
+For each $b$ in that $\tau/2$ fraction, we pick a label uniformly at random from the set:
+
+\[ S_b = \{\ell \mid \mathsf{Inf}^{\leq d}_{\ell}(f_b) \geq \tau/2 \} \]
+which must be \emph{non-empty} by the averaging argument made above.
+Notice that the label picked will satisfy $\pi_{b,a}$ by construction.
+We can thus lower bound the probability of constraint $\pi_{b,a}$ being satisfied:
+%
+\begin{equation} \label{probbound}
+\mathbb{P}[\pi_{(b,a)} \text{ is satisfied}] \geq \frac{\epsilon}{2}\frac{\tau}{2}\frac{1}{|S_b|}
+\end{equation}
+
+Through a Fourier-analytic argument, we can upper-bound $|S_b|$:
+\begin{align*}
+  \frac{|S_b|\tau}{2} \leq \sum_{i = 1}^{|S_b|} \mathsf{Inf}_i^{\leq d}(f_b) \leq \sum_{i = 1}^m \mathsf{Inf}_i^{\leq d}(f_b) = \sum_{\overset{S \subseteq [m]}{|S| \leq d}} |S| \hat{f}^2(S) \leq d \sum_{S \subseteq [m]}  \hat{f}^2(S) = d
+\end{align*}
+This yields that $|S_b| \leq \frac{2d}{\tau}$. So the probability bound of \ref{probbound} would become:
+%
+\begin{equation}
+  \mathbb{P}[\pi_{(b,a)} \text{ is satisfied}] \geq \frac{\epsilon}{2}\frac{\tau}{2}\frac{\tau}{2d} = \frac{\epsilon\tau^2}{8d} = \delta
+\end{equation}
+as desired. This completes the proof of Theorem \ref{maxcutughard}.
+
+\end{proof}
+
+%add a big picture composing both parts.

### 2021-11-18 Complexity/UGC/uniquegames.tex
First draft of UGC paper

@@
 \begin{definition}
-  A \emph{Unique Label Cover Problem} with $m$ labels (UniqueLC($m$)) instance $\mathcal{U}$ is defined by a bipartite graph $(A \sqcup B,E)$, finite alphabet $\Sigma_A = \Sigma_B = \Sigma$ such that $|\Sigma| = m$, and a set of \emph{permutations} $\pi_e:[m] \rightarrow [m]$ for every edge $e \in E$.
+  A \emph{Unique Label Cover Problem} with $m$ labels (UniqueLC($m$)) instance $\mathcal{U}$ is defined by a bipartite graph $(A \sqcup B,E)$ where $|A| = |B| = n$ for some $n \in \mathbb{N}$, finite alphabet $\Sigma_A = \Sigma_B = \Sigma$ such that $|\Sigma| = m$, and a set of \emph{permutations} $\pi_e:[m] \rightarrow [m]$ for every edge $e \in E$.
 Define an \emph{assignment} as consisting of a map $\sigma: A \sqcup B \rightarrow [m]$.
@@
   \begin{align*}
-      \mathsf{YES} & = \{\mathcal{G} \mid \mathsf{Opt}(\mathcal{U}) \geq \beta\} \\
-      \mathsf{NO} & = \{\mathcal{G} \mid \mathsf{Opt}(\mathcal{U}) < \alpha \}
+      \mathsf{YES} & = \{\mathcal{U} \mid \mathsf{Opt}(\mathcal{U}) \geq \beta\} \\
+      \mathsf{NO} & = \{\mathcal{U} \mid \mathsf{Opt}(\mathcal{U}) < \alpha \}
   \end{align*}
@@
 
-\begin{theorem} (Unique Games Conjecture \cite{khot2002power})
+In addition, Khot formulated the \emph{Unique Games Conjecture} and utilized it to prove several inapproxability results assuming the conjecture is true.
+
+\begin{conjecture} (Unique Games Conjecture \cite{khot2002power})
   For any constant $\delta > 0$, there exists sufficiently large $m \in \mathbb{N}$ such that $\mathsf{Gap}_{\delta,1-\delta}\mathsf{UniqueLC}(m)$ is $\NP$-hard.
-\end{theorem}
+\end{conjecture}
 
-%Polynomial time algorithm for guaranteed solutions.
+\begin{remark}
+  The $1-\delta$ constant is crucial to the validity of the conjecture.
+For an instance of $\mathsf{UniqueLC}(m)$ for any $m$ with a guaranteed solution, there exists a polynomial-time algorithm finding an assignment which satisfies all projection constraints: Start with a vertex and set it to a label.
+If it's a part of an edge, follow the edge to the other side and find a label which satisfies as many edges as possible.
+Repeat in a breadth-first search fashion.
+In virtue of the projection maps being permutations, this amounts to searching for the guaranteed solution in time $O(mn^2)$.
+\end{remark}
+
+\begin{example}
+  The $\mathsf{MaxCut}$ problem for an input graph $G = (V,E)$ can be cast as a $\mathsf{UniqueLC}(|V|)$ instance.
+The two partitions of the bipartite graph $A,B$ will be indexed by the vertices $V$.
+Draw an edge between two vertices $v_1,v_2$ if $(v_1,v_2) \in E$.
+Set the alphabet to be $\Sigma =\{-1,1\}$ and the projection maps simply be the ``swap" map $-1 \mapsto 1, 1 \mapsto -1$.
+\end{example}
+
+\begin{example}
+  $\mathsf{MaxE2LinModp}$ for prime $p$ denotes the problem of finding an assignment which maximizes the number of satisfied linear constraints consisting of exactly two variables over field $\mathbb{F}_p$.
+An example of an instance of this problem over variables $x_1,x_2,x_3,x_4$ is shown below:
+
+  \begin{equation*}
+    \begin{alignedat}{3}
+      x_1 & +{} & x_3 & = 3 \\
+      x_2 & +{}  & x_4 & = 2 \\
+      x_1 & +{} & x_4 & = 1
+    \end{alignedat}
+  \end{equation*}
+
+  An instance of this problem can also be translated as an instance of $\mathsf{UniqueLC}(m)$.
+\end{example}
+
+\begin{definition}
+  A problem is said to be \emph{$\UG$-hard}
+\end{definition}
+
+%\subsection{Is the Unique Games Conjecture True?}

### 2021-11-18 Complexity/UGC/approx.tex
First draft of UGC paper
SDP and Goemans written

@@
 
-\begin{example} There exists an Semi-definite Programming (SDP) relaxation for $\mathsf{MaxCut}$.
+\begin{example} \label{maxcutexample}
+The problem $\MaxCut$ will be casted as follows: given an undirected graph $G=(V,E)$, find the largest cut of $G$. There is a straightforward LP formulation for this problem:
+
+\begin{align}
+  &  \min \sum_{(u,v) \in E} e_{u,v} \\
+  & e_{u,v} \geq
+  \max \begin{cases}
+      x_1 + x_2 \\
+      2 - (x_1 + x_2)
+  \end{cases} \\
+  & e_{u,v} \in [1,2] \\
+  & x_v \in [0,1]
+\end{align}
+%
+where the $e_{u,v}$ variables represent the edges $e \in E$ of the input graph and the $x_v$ represent the vertex variables $v \in V$.
+This will actually be a relaxation from its respective integer program in which the variables can take the integer values:
+
+\begin{align}
+  e_{u,v} = \begin{cases}
+              1 \text{ if } e_{u,v} \in \mathcal{C} \\
+              2 \text{ otherwise }
+            \end{cases}
+  x_v = \begin{cases}
+              0 \text{ if } v \text{ is in partition } S \\
+              1 \text{ otherwise }
+        \end{cases}
+\end{align}
+
+where $\mathcal{C} \subseteq E$ is the edges constituting the cut and $(S,\bar{S})$ is the partition defining the cut.
+Naturally, a solution to the Integer program will also be a solution to the LP above, which implies that
+
+\[ \mathsf{Opt}_{LP} \geq \MaxCut(G)\]
+
+There actually exists an Semi-definite Programming (SDP) relaxation for $\MaxCut$ called the \emph{Goemans-Williamson} algorithm which will be treated in Section \ref{goemans}.
 

### 2021-11-18 Complexity/UGC/labelcover.tex
SDP and Goemans written

@@
 These ``projection games" were introduced by Bellare, Goldreich, and Sudan \cite{bellare1998free}.
-The $NP$-hardness of the gap problem version of Label Cover was used by H\aa stad to show tight inapproximability results for $\mathsf{Max3SAT}$ and $\ELin$ \cite{haastad2001some}.
+The $\NP$-hardness of the gap problem version of Label Cover was used by H\aa stad to show tight inapproximability results for $\mathsf{Max3SAT}$ and $\ELin$ \cite{haastad2001some}.
 
@@
 %Arora, Barak
+\subsection{Some Structural Results of PCPs}
+
 \subsection{Raz's Parallel Repetition Theorem}
-To be written...
 

### 2021-11-18 Complexity/UGC/ugcmaxcut.tex
SDP and Goemans written

@@
-\newcommand{\GapMaxCut}{\mathsf{Gap}_{\GWConstant + \epsilon, \frac{1-\rho}{2} - \epsilon}\mathsf{MaxCut}}
-\newcommand{\ULC}{\mathsf{UniqueLC}(m)}
-\newcommand{\GapULC}{\mathsf{GapUniqueLC}(m)}
-\newcommand{\GapDeltaULC}{\mathsf{Gap}_{\delta, 1-\delta}\ULC}
-
 \section{UG-hardness of MAXCUT}
@@
 
-\subsection{Goemans-Williamson Algorithm for $\MaxCut$}
-Recall that Example \ref{maxcutexample} presented an LP-based approximation algorithm for $\MaxCut$.
+\subsection{Goemans-Williamson Algorithm for $\MaxCut$} \label{goemans}
+ Example \ref{maxcutexample} presented an LP-based approximation algorithm for $\MaxCut$.
 Goemans and Williamson \cite{goemans1995improved} designed an SDP to give an $\alpha_{GW}$-approximation algorithm for $\MaxCut$ where:
@@
 
+To begin, a semi-definite program is a generalization of a linear program where instead of optimizing over a vector of variables $\vec{x}$, the program considers a positive semi-definite matrix of variables i.e a matrix whose eigenvalues are non-negative.
+An equivalent formulation considers inner products between pairs of vectors:
+
+\begin{align*}
+     \max & \sum_{i,j} c_{ij}\langle v_i, v_j \rangle \\
+     \text{ under constraints }& a_{ij}^k\langle v_i, v_j \rangle \leq b_{ij}^k \\
+     v_1, \cdots v_n \in \mathbb{R}, \quad & k \in [C] \text{ constraints }
+\end{align*}
+
+Note that this form also subsumes quadratic programming by setting $c_{ij} = a_{ij}^k = b_{ij}^k = 0$ for all $i \neq j$ and $k \in [C]$.
+The Goemans-Williamson algorithm concerns the solution to the semi-definite program given some graph $G = (V,E)$: \newline
+%
+\begin{align}
+  & \max \sum_{i,j} \frac{1 - \langle v_i, v_j \rangle}{2} \\
+  & \langle v_i, v_i \rangle = 1 \text{ for all } i \in V
+\end{align}
+
+Let us first show that indeed the program is a relaxation of the integer program crafted in Example \ref{maxcutexample}.
+Indeed, if we set any unit vector $\vec{u}$ such that for a cut defined by $(S,\bar{S})$:
+%
+\begin{align*}
+  \vec{v_i} = \begin{cases}
+                u \text{ if } i \in S \\
+                -u \text{ if } i \not\in S
+              \end{cases}
+\end{align*}
+
+A direct calculation yields that the term $\frac{1 - \langle v_i, v_j \rangle}{2} = 0$ if $v_i = v_j = u$ else it is equal to 1. Thus, by applying this observation to the maximum cut,
+%
+ \[\mathsf{Opt}_{GW}(G) \geq \MaxCut(G)\]
+
+So far the semi-definite program seems to be rather simple.
+The insight made by Goemans and Williamson lies in the rounding procedure of the solution outputted by the program.
+The procedure proceeds by first drawing a random hyperplane passing through the origin and taking the two partitions of the cut $S_+, S_-$ to be the solution $v_i$ which lie on positive and negative sides of the hyperplane respectively.
+We can calculate the expected size of the cut:
+%
+\begin{align*}
+\mathbb{E}[|E(S_+,S_-)|] = \sum_{(i,j) \in E} \mathbb{P}[v_i,v_j \text{ lie on different sides of the hyperplane}]
+\end{align*}
+
+Now if $\theta_{ij}$ denotes the angle between $v_i,v_j$, a simple geometric argument shows that:
+
+\[\mathbb{P}[v_i,v_j \text{ lie on different sides of the hyperplane}] = \frac{\theta_{ij}}{\pi} \]
+
+By definition of the dot product:
+
+\[\frac{1 - \langle v_i, v_j \rangle}{2} =  \frac{1 - \cos(\theta_{ij})}{2} \]
+
+and the magical inequality:
+
+\[ \frac{\theta_{ij}}{\pi} \geq \alpha_{GW} \cdot \frac{1 - \cos(\theta_{ij})}{2}\]
+
+the two can be combined to finally yield that:
+
+\begin{equation}
+  \frac{\mathbb{E}[|E(S_+,S_-)|]}{\mathsf{Opt}_{GW}(f)} \approx \alpha_{GW}
+\end{equation}
+
 \subsection{$\MaxCut$ is $\UG$-hard}
@@
 
-\begin{theorem} \label{maxcutughard} ($\MaxCut$ is $\UG$-hard) The problem $\GapMaxCut$ is $\UG$-hard]
+\begin{theorem} \label{maxcutughard} ($\MaxCut$ is $\UG$-hard) The problem $\GapMaxCut$ is $\UG$-hard
 \end{theorem}

### 2021-11-18 Complexity/UGC/uniquegames.tex
SDP and Goemans written

@@
 For an instance of $\mathsf{UniqueLC}(m)$ for any $m$ with a guaranteed solution, there exists a polynomial-time algorithm finding an assignment which satisfies all projection constraints: Start with a vertex and set it to a label.
-If it's a part of an edge, follow the edge to the other side and find a label which satisfies as many edges as possible.
+If it is an endpoint of an edge $e$, follow $e$ to the other side and find a label which satisfies as many neighbors as possible.
 Repeat in a breadth-first search fashion.
@@
 \begin{definition}
-  A problem is said to be \emph{$\UG$-hard}
+  A problem $P$ is said to be \emph{$\UG$-hard} if there is some constant $\delta > 0$ such that for all $m \in \mathbb{N}$, there exists a polynomial-time reduction from $\GapDeltaULC$ to $P$.
 \end{definition}

### 2021-11-19 Complexity/UGC/hastad.tex
Started on Hastad's 3 bit PCP write-up

@@
 %Arora, Barak
+\subsection{Inapproximability Results for $\ELin$ and $\mathsf{Max3Sat}$}
+The significance of H\aa stad's PCP for $\NP$ is its use in showing tight inapproximability results for $\ELin$ and hence $\mathsf{Max3Sat}$. The theorem is first stated below:
+
+\begin{theorem}(H\aa stad's 3-bit PCP, \cite{haastad2001some})
+For every $\delta > 0$ and $L \in \NP$, there exists a PCP verifier for $L$ over the boolean alphabet such that for every
+\begin{itemize}
+  \item The verifier $V$ queries 3 bits of the proof $x_{q_1},x_{q_2}, x_{q_3} \in \pi$ such that verification predicate is a three variable linear equation over $\mathbb{F}_2$ depending on the queried bits $x_{q_1},x_{q_2}, x_{q_3}$.
+  \item If $x \in L$, then there exists a proof $\pi$:
+        \begin{equation}
+          \mathbb{P}[V(x,\pi) = 1] \geq 1 - \delta
+        \end{equation}
+  \item If $x \not\in L$, then for all proofs $\pi$:
+        \begin{equation}
+          \mathbb{P}[V(x,\pi) = 1] \geq \frac{1}{2} + \delta
+        \end{equation}
+\end{itemize}
+
+\end{theorem}
+
+\subsection{The Long Code}
+To be written...
+
 \subsection{Aside on Dictatorship Testing}
-To be written...
-\subsection{Long Code}
-To be written...
+The first step towards testing if an input boolean function $f:\{-1,1\}^n \rightarrow \{-1.1\}$ is a dictator arises from the Blum-Luby-Rubinfeld Test (BLR) which we restate below: \newline
+
+\begin{enumerate}
+  \item Sample $x,y \sim_R \{-1,1\}^n$
+  \item Accept iff $f(x)f(y) = f(xy)$
+\end{enumerate}
+
+\begin{theorem}
+  Suppose the BLR test accepts $f:\{-1,1\}^n \rightarrow \{-1,1\}$ with probability $1 - \epsilon$, then $f$ is $\epsilon$-close to a linear function $\chi_{S*}$ for some $S^* \subseteq [n]$.
+\end{theorem}
+%
+Certainly the dictators $\chi_i, \; i \in [n]$ are linear functions.
+Hence, if $f = \chi_i$ for some $i$, then the BLR test accepts with probability one.
+However, we have the rest of the parity functions $\chi_{S}, |S| \geq 2$ which the BLR Test cannot distinguish.
+We need to amend the test to ensure that parity functions of higher weight are rejected with high probability.
+H\aa stad proposed modifiying the vanilla BLR test to add noise to the sampled product $xy$.
+Although this sacrifices perfect completeness, it penalizes large parity functions: \newline
+%
+\begin{enumerate}
+  \item Sample $x,y \sim \{-1,1\}^n$
+  \item Sample $z \sim N_{1-2\epsilon}(xy)$
+  \item Accept iff $f(x)f(y)=f(z)$
+\end{enumerate}
+%
+
+\begin{lemma} (Completeness of the Noisy BLR Test)
+ If $f = \chi_{i}$ is a dictator for some $i$, then
+  \[ \mathbb{P}[\text{ Noisy BLR test accepts }] \geq 1 - \epsilon \]
+\end{lemma}
+\begin{proof}
+  Note that if $z \sim N_{1-2\epsilon}(xy)$ for some $x \in \{-1,1\}^n$, then $y$ can be expressed as below:
+  \begin{equation}
+    z_i = \begin{cases}
+             \phantom{-} x_iy_i \text{ with probability } 1- \epsilon \\
+             - x_iy_i \text{ with probability } \epsilon
+          \end{cases}
+  \end{equation}
+  By the acceptance criterion,
+  \[ f(z) = f(x)f(y) \implies z_i = x_iy_i \]
+  This occurs with probability $1- \epsilon$ by the observation above, yielding completeness as claimed.
+\end{proof}
+%
+%
+\begin{lemma} Suppose that for some constant $\nu > 0$:
+\begin{equation*}
+  \mathbb{P}[\text{ Noisy BLR test accepts }] \geq \frac{1}{2} + \nu
+\end{equation*} then
+\begin{equation}
+  2\nu \leq \sum_{S \subseteq [n]} \hat{f}(S)^3(1-2\epsilon)^{|S|}
+\end{equation}
+\end{lemma}
+%
+\begin{proof}
+  The proof begins by noticing the accepting probability can be expressed as:
+  \[ \mathbb{P}[\text{ Noisy BLR test accepts }] = \frac{1}{2} + \frac{1}{2}\mathbb{E}_{x,y,z} \left[f(x)f(y)f(z) \right] \]
+  By our assumption, we prove that:
+  \begin{align*}
+    \frac{1}{2} + \nu & \leq \frac{1}{2} + \frac{1}{2}\mathbb{E}_{x,y,z} \left[f(x)f(y)f(z) \right] \implies \\[0.7ex]
+    2\nu & \leq \mathbb{E}_{x,y,z} \left[f(x)f(y)f(z) \right] \implies \\[0.7ex]
+    2\nu & \leq\mathbb{E}_{x,y,z} \left[f(x)f(y) \mathbb{E}_{z \sim N_{1 - 2\epsilon}(xy)} \left[f(z)\right]\right] \implies \\[0.7ex]
+    2\nu & \leq \mathbb{E}_{x}\left[ f(x)\mathbb{E}_{y}\left[ f(y)\mathcal{T}_{1-2\epsilon}f(xy)\right]\right] \\[0.7ex]
+    2\nu & \leq \mathbb{E}_{x}\left[ f(x) (f * \mathcal{T}_{1-2\epsilon}f)(x)\right] \implies\\[0.7ex]
+    2\nu & \leq \sum_{S \subseteq} \widehat{f}(S)\widehat{f}(S)\widehat{\mathcal{T}_{1-2\epsilon}f}(S) \implies \\[0.7ex]
+    2\nu & \leq  \sum_{S \subseteq} \widehat{f}(S)^3 (1-2\epsilon)^{|S|}
+  \end{align*}
+  as claimed.
+\end{proof}
+%
+\begin{corollary}
+  (Soundness of the Noisy BLR Test) There exists some $S^* \subseteq [n]$ such that
+  \begin{equation}
+    |\widehat{f}(S^*)| \geq 2\nu \quad |S^*| \leq O\left(\frac{1}{\epsilon}\log{\frac{1}{\nu}}\right)
+  \end{equation}
+\end{corollary}

### 2021-11-21 Complexity/UGC/approx.tex
finished up Hastad's 3bit PCP section

@@
 %
-\begin{example}
+\begin{example} \label{e3linexample}
   Let $\mathsf{MaxE3Lin}$ define the problem of the following form: let $\xi$ be defined as a system of linear equations over the field $\mathbb{F}_2$ where every equation contains \emph{exactly} $3$ variables from $n$ variables $x_1,\cdots,x_n$.

### 2021-11-21 Complexity/UGC/hastad.tex
finished up Hastad's 3bit PCP section

@@
 
-\begin{theorem}(H\aa stad's 3-bit PCP, \cite{haastad2001some})
+\begin{theorem} \label{hastadpcp} (H\aa stad's 3-bit PCP, \cite{haastad2001some})
 For every $\delta > 0$ and $L \in \NP$, there exists a PCP verifier for $L$ over the boolean alphabet such that for every
@@
         \begin{equation}
-          \mathbb{P}[V(x,\pi) = 1] \geq \frac{1}{2} + \delta
+          \mathbb{P}[V(x,\pi) = 1] \leq \frac{1}{2} + \delta
         \end{equation}
 \end{itemize}
-
 \end{theorem}
 
+On closer inspection, Theorem \ref{hastadpcp} reveals that a gap problem based on $\ELin$ is actually $\NP$-hard to decide:
+
+\begin{corollary} \label{e3linhard}
+For every $\delta >0$, the promise problem $\mathsf{Gap}_{\frac{1}{2} + \delta, 1 - \delta}\ELin$ is $\NP$-hard.
+\end{corollary}
+
+This gives our first threshold result: Corollary \ref{e3linhard} states that the simple $\frac{1}{2}$-approximation algorithm introduced in Example \ref{e3linexample} is optimal in the sense that the existence of an efficient algorithm improving on the $\frac{1}{2}$ constant factor would show that $\P = \NP$.
+From this result, by reducing $\ELin$ to $\mathsf{Max3Sat}$, another threshold result appears:
+
+\begin{corollary}
+For every $\delta >0$, the promise problem $\mathsf{Gap}_{\frac{7}{8}+\delta,1-\delta}\mathsf{Max3Sat}$ is $\NP$-hard.
+\end{corollary}
+%
+\begin{proof}
+It suffices to transform $\ELin$ CSP instances to $\mathsf{MaxSat}$ CSP instances. For predicates of the form $x + y + z = 1$, consider the conjunction of the clauses:
+\begin{align*}
+   x \vee y \vee z \\
+   \bar{x} \vee y \vee \bar{z} \\
+   \bar{x} \vee y \vee \bar{z} \\
+   x \vee \bar{y} \vee \bar{z}
+\end{align*}
+The case for predicates of the form $x + y + z = 0$ are handled with a similar idea.
+For an assignment of a $\ELin$ instance which satisfies a linear constraint, it must satisfy all four of the associated $\mathsf{MaxSat}$ constraints.
+In the case where an assignment does not satisfy a linear constraint, at most three of the four clauses will be satisified.
+Hence if an assignment satifies $\frac{1}{2} + \delta$ of the constraints, then it can satisify at most:
+\[ (\frac{1}{2} + \delta)4 + (\frac{1}{2} - \delta)\frac{3}{4} = \frac{7}{8} + \frac{3\delta}{4}\]
+
+and if at least $1 - \delta$ of the linear constaints are satisfied then at least $1 - \delta$ of the clauses must also be satisfied.
+\end{proof}
+
+The proof of Theorem \ref{hastadpcp} leverages the $\NP$-hardness of $\mathsf{Gap}_\epsilon\mathsf{LC}$ mentioned in Theorem \ref{labelcoverhard}.
+H \aa stad devised a dictatorship test based on $\mathsf{E3Lin}$ predicates then composed this test with checking the validity of a $\mathsf{Gap}_\epsilon\mathsf{LC}$ proof.
+In fact, the three queries to the proof arise from a modified Blum-Luby-Rubinfeld Linearity Test.
+The details will be treated in the upcoming subsections.
+
 \subsection{The Long Code}
-To be written...
+The workhorse for H\aa stad's proof will be the use of the \emph{long code}.
+A long code is esstentially the truth table of a dictator function.
+The idea roughly revolve around the idea that dictator functions are locally-testable as linear functions (\`a la BLR Test) and have an useful error-correction property.
+Correct proofs will be expected to encode labels to vertices in a Label Cover instance as long codes of dictators corresponding to the indices of the labels.
+In other words, for a label $\ell \in \Sigma$, a corresponding index $i(\ell)$ can be computed with $\log{|\Sigma|}$ bits.
+The total number of such bitstrings will be, of course, length $2^{\log{|\Sigma|}} = |\Sigma|$.
+If we additionally encode this bitstring into the dictator function $\chi_{i(\ell)}(x_1,\cdots,x_{\Sigma}) = x_{i(\ell)}$, the truth table for this function will be of total length $2^{|\Sigma|}$.
+This reveals that the long code requires a \emph{double-exponential} proof length.
+The truth table will serve as query access to the modified BLR test mentioned above.
 
@@
 %
-\begin{lemma} Suppose that for some constant $\nu > 0$:
+\begin{lemma} Suppose that for some constant $\nu > 0$: \label{noisyblrsoundness}
 \begin{equation*}
@@
     2\nu & \leq \mathbb{E}_{x}\left[ f(x) (f * \mathcal{T}_{1-2\epsilon}f)(x)\right] \implies\\[0.7ex]
-    2\nu & \leq \sum_{S \subseteq} \widehat{f}(S)\widehat{f}(S)\widehat{\mathcal{T}_{1-2\epsilon}f}(S) \implies \\[0.7ex]
-    2\nu & \leq  \sum_{S \subseteq} \widehat{f}(S)^3 (1-2\epsilon)^{|S|}
+    2\nu & \leq \sum_{S \subseteq [n]} \widehat{f}(S)\widehat{f}(S)\widehat{\mathcal{T}_{1-2\epsilon}f}(S) \implies \\[0.7ex]
+    2\nu & \leq  \sum_{S \subseteq [n]} \widehat{f}(S)^3 (1-2\epsilon)^{|S|}
   \end{align*}
@@
 \end{proof}
+The operator $\mathcal{T}_{\rho}$ is the \emph{noise operator} which, given boolean function $f:\{-1,1\}^n \rightarrow \{-1,1\}$, maps $f$ to another boolean function $\mathcal{T}_\rho(f)$ with the following Fourier decomposition:
+
+\begin{equation}
+  \mathcal{T}_\rho(f) = \sum_{S \subseteq [n]} \rho^{|S|}\widehat{f}(S)\chi_S
+\end{equation}
+
 %
@@
 \end{corollary}
+
+\subsection{The Code Consistency Test}
+The consistency test will compose the dictatorship test formulated above with checking the validity of an alleged long code as well as satisfiablity of the labels encoded.
+More formally, the input will be a $\LC$ instance $((A\sqcup B,E),\Sigma_A,\Sigma_B,\{\pi_e\}_{e \in E})$ and the proof will be the long codes encoding the assignments to the vertices, $\mathfrak{A}: A \rightarrow \Sigma_A$, $\mathfrak{B}: B \rightarrow \Sigma_B$.
+The consistency check will check validity of the long code as well as checking the labels encoded by the long codes satisfy the projection constraints $\{\pi_e\}_{e \in E}$: First some notation is in order.
+Let $(x \circ \pi)_i = x_{\pi(i)}, \; \forall i \in \Sigma_A$ for $x \in \{-1,1\}^{|\Sigma_B|}$.
+It worth keeping in mind that $x \circ \pi \in \{-1,1\}^{|\Sigma_A|}$.
+
+\begin{enumerate}
+  \item Sample an edge $e = (a,b) \in E$. Let $\pi$ be the associated projection constraint.
+
+  \noindent Let $f_a:\{-1,1\}^{|\Sigma_A|} \rightarrow \{-1,1\}, \; f_b:\{-1,1\}^{|\Sigma_B|} \rightarrow \{-1,1\}$ be the functions claimed to be long codes for the labels assigned to vertices $a,b$: $\mathfrak{A}(a),\mathfrak{B}(b)$.
+Recall these will be encoded as truth table strings on the proof string $\pi$.
+  \item Sample $x \sim \{-1,1\}^{|\Sigma_B|}, \; y \sim \{-1,1\}^{|\Sigma_A|}$
+  \item Sample $z \sim N_{1 - 2\delta}((x \circ \pi)y)$
+  \item Accept iff $f_a(z) = f_b(x)f_a(y)$
+\end{enumerate}
+
+Note that if the labels assigned $a,b$ satisfied $\pi$ and $f_a,f_b$ are valid long codes: $f_a(x \circ \pi) = f_b(x )$ for all $x \in \{-1,1\}^{|\Sigma_B|}$.
+
+\begin{proposition} (Completeness)
+  If $\mathsf{Gap}_\delta\mathsf{LC}$ instance $((A\sqcup B,E),\Sigma_A,\Sigma_B,\{\pi_e\}_{e \in E})$ satisfies all constraints, then there exists code words $f_a, f_b$ such that
+  \begin{equation}
+    \mathbb{P}[ \text{ Consistency Test Accepts } ] \geq 1 - \delta
+  \end{equation}
+\end{proposition}
+\begin{proof}
+  Suppose the assignment atisfies all constaints of the $\LC$ instance above.
+Set $f_a(x_1,\cdots, x_{\Sigma_A}) = x_{\mathfrak{A}(a)}, \; f_b(x_1,\cdots, x_{\Sigma_B}) = x_{\mathfrak{B}(b)}$.
+The following then hold:
+
+  \begin{align*}
+      \mathbb{P}[ \text{ Consistency Test Accepts } ] & = \mathbb{P}_{x,y,z}\left[f_a(z) = f_b(x)f_a(y) \right] \\
+      & = \mathbb{P}_{x,y, \mu}\left[ (x \circ \pi)_{\mathfrak{A}(a)}y_{\mathfrak{A}(a)}\mu_{\mathfrak{A}(a)} = x_{\mathfrak{B}(b)}y_{\mathfrak{A}(a)}  \right] \quad (\mu \sim N_{1 - 2\delta}(1^{|\Sigma_A|})) \\
+      & = \mathbb{P}_{x,\mu} \left[(x \circ \pi)_{\mathfrak{A}(a)}\mu_{\mathfrak{A}(a)} = x_{\mathfrak{B}(b)} \right] \\
+      & =  \mathbb{P}_{x,\mu} \left[x_{\pi(\mathfrak{A}(a))}\mu_{\mathfrak{A}(a)} = x_{\mathfrak{B}(b)} \right]\\
+      & =  \mathbb{P}_{x,\mu} \left[x_{\mathfrak{B}(b)}\mu_{\mathfrak{A}(a)} = x_{\mathfrak{B}(b)} \right] \\
+      & = 1 - \delta
+  \end{align*}
+\end{proof}
+%
+%
+
+\subsection{Folded Functions}
+
+\subsection{Soundness and Decoding a Labeling}
+
+\begin{proposition}
+  Suppose that for a $\LC$ instance, the consistency test above accepts with at least $\frac{1}{2} + \delta$,  then there exists a labeling which satisfies at least $\delta$ of the projection constraints.
+\end{proposition}
+Note that the soundness criterion intuitively implies the use of a method to extract out a labeling from a high enough acceptance probability.
+This is the ``local-correctability" property of linear functions coming into play.
+We first prove the following lemma:
+
+\begin{definition}
+  For $S \subseteq \Sigma_A$ and $\pi:\Sigma_A \rightarrow \Sigma_B$, define $\pi_2(S) = \{j \in \Sigma_B \mid |\pi^{-1}(j) \cap S| \text{ is odd}\}$
+\end{definition}
+
+\begin{lemma} \label{soundnessbound}
+  Assume the test accepts with at least $\frac{1}{2} + \delta$ probability, then for at least $\delta$ fraction of the edges $(a,b)$, the associated codes $f_a:\{-1,1\}^{|\Sigma_A|} \rightarrow \{-1,1\}, \; f_b:\{-1,1\}^{|\Sigma_B|} \rightarrow \{-1,1\}$ satisfy the following inequality:
+  %
+  \begin{equation}
+  \sum_{S \subseteq [n]} \widehat{f_a}(S)^2\widehat{f_b}(\pi_2(S))(1-2\delta)^{|S|} \geq \delta
+  \end{equation}
+\end{lemma}
+%
+\begin{proof}
+  The proof is remarkably similar to that of Lemma \ref{noisyblrsoundness}. From our assumption, it is immediate that:
+  \begin{equation}
+    \frac{1}{2} + \frac{1}{2}\mathbb{E}_{(u,v),x,y,\mu}\left[f_b(x)f_a(y)f_a((x \circ \pi)y\mu) \right] \geq \frac{1}{2} + \delta  \quad (\mu \sim N_{1 - 2\delta}(1^{|\Sigma_A|}))
+  \end{equation}
+  which implies that:
+  \begin{equation}
+    \mathbb{E}_{(u,v),x,y,\mu}\left[f_b(x)f_a(y)f_a((x \circ \pi)y\mu) \right] \geq 2\delta
+  \end{equation}
+  By an averaging argument, there must exist at least $\delta$ fraction of the edges which satisfy:
+  \begin{equation}
+      \mathbb{E}_{x,y,\mu}\left[f_b(x)f_a(y)f_a((x \circ \pi)y\mu) \right] \geq \delta
+  \end{equation}
+  Hence, through a Fourier-analytic argument:
+  \begin{align*}
+    \mathbb{E}_{x,y,\mu}\left[f_b(x)f_a(y)f_a((x \circ \pi)y\mu) \right] & \geq \delta \implies \\[0.7ex]
+    \mathbb{E}_{x,y}\left[f_b(x)f_a(y) \mathbb{E}_{\mu} \left[f_a((x \circ \pi)y\mu) \right] \right] & \geq \delta \implies \\[0.7ex]
+    \mathbb{E}_{x,y}\left[f_b(x)f_a(y) \mathcal{T}_{1-2\delta}(f_a)(x \circ \pi)y) \right] & \geq \delta \implies \\[0.7ex]
+    \mathbb{E}_{x}\left[ f_b(x)(f * \mathcal{T}_{1 - 2\delta}(f_a))(x \circ \pi) \right] & \geq \delta \implies \\
+    \sum_{S} \widehat{f_b}(S) \widehat{f_a}(\pi^{-1}(S))^2(1 - 2\delta)^{|S|} & \geq \delta  \implies \\
+    \sum_{S} \widehat{f_b}(\pi(S))\widehat{f_a}(S)^2 (1 - 2\delta)^{|S|} & \geq \delta \implies \\
+    \sum_{S} \widehat{f_b}(\pi_2(S))\widehat{f_a}(S)^2 (1 - 2\delta)^{|S|} & \geq \delta
+    \end{align*}
+    The last implication follows from the observation that:
+
+    \[ \chi_S(x \circ \pi) = \prod_{i \in S} x_{\pi(i)} = \prod_{i \in \pi(S)} x_i = \prod_{i \in \pi_2(S)} x_i = \chi_{\pi_2(S)}(x) \]
+\end{proof}
+
+We can now finish the soundness proof by decoding the labeling from our supplied proof string $\pi$. The actual procedure is not complicated:
+
+\begin{enumerate}
+    \item For each $a \in A$, do the following steps:
+    \item Sample $S \sim (\widehat{f}_a(S))^2$
+    \item Set $\mathfrak{A}(a) \xleftarrow{R} S$ \quad (\text{uniformly sample from } $S$)
+\end{enumerate}
+
+Do the same for each $b \in B$ except sample the subset $T \sim \widehat{f}_b^2(T)$ from the Fourier spectrum of $f_b$:
+
+\begin{enumerate}
+    \item For each $b \in A$, do the following steps:
+    \item Sample $T \sim (\widehat{f}_b(T))^2$
+    \item Set $\mathfrak{B}(b) \xleftarrow{R} T$ \quad (\text{uniformly sample from } $T$)
+\end{enumerate}
+
+ Since $f_a,f_b$ are assumed to be folded (odd) functions, both $\widehat{f_a}(\emptyset) = \widehat{f_b}(\emptyset) = 0$, so we will never sample the empty set in any of the procedures above.
+We now aim to show the inequality:
+
+\begin{lemma} \label{firststepsoundness}
+  \begin{equation}
+      \mathbb{P}_{(a,b) \in E}[ \text{(a,b) is satisfied}] \geq \sum_{S} \sum_{T \subseteq \pi(S)} \widehat{f}_a(S)^2\widehat{f}_b(T)^2 \cdot \frac{1}{|S|}
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  Here is one way to get a label which satisfies a chosen edge $(a,b)$.
+The first procedure samples a subset from the Fourier spectrum of $f_a$.
+The second procedure then samples from the Fourier spectrum of $f_b$ and receives a subset $T \subseteq \pi(S)$.
+The second procedure then uniformly samples from $T$ to receive some label $t \in T$.
+Since $T$ is contained in the image of $S$ under $\pi$, there must exist at least one element in $s \in S$ such that $\pi(s) = t$, showing that this event occurs with probability at least $\frac{1}{|S|}$
+\end{proof}
+
+\begin{proposition} (Soundness)
+    For the labeling scheme constructed above, the probability:
+    \begin{equation}
+      \mathbb{P}_{(a,b) \in E}[ (a,b) \text{ is satisfied}] \geq 4\delta^4
+    \end{equation}
+\end{proposition}
+%
+\begin{proof}
+  By Lemma \ref{firststepsoundness}:
+  \begin{align*}
+    \mathbb{P}_{(a,b) \in E}[ \text{(a,b) is satisfied}] & \geq \sum_{S} \sum_{T \subseteq \pi(S)} \widehat{f}_a(S)^2\widehat{f}_b(T)^2 \cdot \frac{1}{|S|} \\
+    & \geq \sum_{S} \widehat{f}_a(S)^2\widehat{f}_b(\pi_2(S))^2  \frac{1}{|S|} \\
+    & =  \sum_{S} \left(\widehat{f}_a(S)\widehat{f}_b(\pi_2(S))  \frac{1}{\sqrt{|S|}}\right)^2 \cdot \left(\sum_{S}\widehat{f}(S)^2 \right) \\
+    & \geq \left(\sum_{S} \widehat{f}_a(S)^2\widehat{f}_b(\pi_2(S))  \frac{1}{\sqrt{|S|}} \right)^2 \quad (\text{Cauchy-Schwarz}) \\
+    & \geq 4\delta\left(\sum_{S} \widehat{f}_a(S)^2\widehat{f}_b(\pi_2(S))  (1 - 2\delta)^{|S|} \right)^2
+  \end{align*}
+  The last inequality holds since $\frac{1}{\sqrt{x}} \geq 2\sqrt{\delta}(1-2\delta)^x$ for all $x, \delta > 0$.
+By virtue our bound derived in Lemma \ref{soundnessbound}, we find that for at least an $\delta$ of the edges,
+  $\sum_{S} \widehat{f}_a(S)^2\widehat{f}_b(\pi_2(S))  (1 - 2\delta)^{|S|} \geq \delta$. Hence:
+
+  \begin{equation}
+          \mathbb{P}_{(a,b) \in E}[ (a,b) \text{ is satisfied}] \geq 4\delta\cdot\delta^3 = 4\delta^4
+  \end{equation}
+\end{proof}
+
+By Theorem \ref{labelcoverhard}, we know that $\mathsf{Gap}_{4\delta^4}\mathsf{LC}$ is $\NP$-hard for every $\delta \in [0,\frac{1}{2}]$. Since $\delta
+\geq 4\delta^4$ for $\delta \in [0,\frac{1}{2}]$.
+The argument for the completeness of this PCP verifier still holds if we set our $\delta$ to be $\delta' = 4\delta^4$.
+This demonstrates that we gave a PCP system for the $\NP$-hard $\mathsf{Gap}_{4\delta^4}\mathsf{LC}$:
+
+\begin{proposition}
+  Suppose for an input $\mathsf{Gap}_{\delta}\mathsf{LC}$ instance say $\mathcal{I}$ and associated proof string $\pi$  causes the test to pass with at least $\frac{1}{2} + \delta$:
+
+  \begin{equation}
+    \mathbb{P}[\text{Consistency Test Passes}] \geq \frac{1}{2} + \delta
+  \end{equation}
+  then there exists a assignment for $\I$ such that it satisfies at least $\delta' = 4\delta^4$ fraction of edges in $\I$.
+\end{proposition}
+This gives a PCP system for $\NP$ as desired.

### 2021-11-21 Complexity/UGC/labelcover.tex
finished up Hastad's 3bit PCP section

@@
 
-\begin{theorem} (Projection Games Theorem)
+\begin{theorem} (Projection Games Theorem) \label{labelcoverhard}
   For every $\epsilon > 0$, there exist alphabets $\Sigma_A, \Sigma_B$ where $|\Sigma_A|,|\Sigma_B| \leq \mathsf{poly}(\frac{1}{\epsilon})$ such that $\mathsf{Gap}_\epsilon\mathsf{LC}$ is $\NP$-hard.

### 2021-11-21 Complexity/UGC/ugcmaxcut.tex
Started on Hastad's 3 bit PCP write-up
finished up Hastad's 3bit PCP section

@@
 
-So far the semi-definite program seems to be rather simple.
+So far the semi-definite program appears to be rather simple.
 The insight made by Goemans and Williamson lies in the rounding procedure of the solution outputted by the program.
@@
 
-the two can be combined to finally yield that:
+the two can be combined to yield that:
 
 \begin{equation}
-  \frac{\mathbb{E}[|E(S_+,S_-)|]}{\mathsf{Opt}_{GW}(f)} \approx \alpha_{GW}
+  \frac{\mathbb{E}[|E(S_+,S_-)|]}{\mathsf{Opt}_{GW}(f)} \geq \alpha_{GW}
+\end{equation}
+
+Furthermore, the first inequality above implies that the Goemans-Williamson algorithm is indeed an $\alpha_{GW}$ approximation algorithm:
+
+\begin{equation}
+  \frac{\mathbb{E}[|E(S_+,S_-)|]}{\MaxCut(G)} \geq \alpha_{GW}
 \end{equation}
@@
 \newline
+
+%
+%
+%
+\subsection{Dictatorship Testing Redux}
 

### 2021-11-23 Complexity/UGC/pcp.tex
More touches to UGC paper

@@
   x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
-  x \not\in L & \implies \mathsf{Opt}(f(x)) < \alpha
+  x \not\in L & \implies \mathsf{Opt}(f(x)) \leq \frac{1}{2}
 \end{align*}
@@
 
-\begin{theorem} ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) For some constants $q \in \mathbb{N}, \alpha \in (0,1)$, the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard.
+\begin{theorem} \label{pcptheoremgapcsp}  ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) For some constants $q \in \mathbb{N}$, the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard.
 To elaborate, for every $\NP$ language $L$, there exists a polynomial time reduction mapping an $L$ to a instance $f(x)$ of some CSP $\Psi$ where $\Psi$ consists of $q$-ary predicates, such that
@@
 \begin{lemma} (Equivalence between PCPs and CSPs)
-
+  Theorem \ref{pcptheorem} and  Theorem \ref{pcptheoremgapcsp} are equivalent.
 \end{lemma}
 
+\begin{proof}
+First, assume $\NP = \PCP_{1,\frac{1}{2}}(O(\log{n}), O(1))$.
+We will outline a procedure to convert the verifier $V$ into an instance $\I$ for a $q$-ary CSP $\Psi$ for some constant $q$.
+For some input string $x \in \{0,1\}^n$ and proof string $\pi$, let $r \in \{0,1\}^{c \cdot \log{n}}$ be the coin flips made by $V$ and $V_{x,r}$ be the deterministic procedure which is executed on input $x$ and coin flip $r$ such that $V_{x,r} = 1$ iff $V$ accepts proof $\pi$ on input $x$ and coin flip $r$.
+We can define the domain of our constructed CSP $\Psi$ to be $\Omega=\{0,1\}$ and the predicates to be $\{V_{x,r}\}_{r}$.
+Now our instance $\I$ of $\Psi$ is casted as the tuples $(S, V_{x,r})$ where $S$ will be at most a $q$-sized tuple indicating which indices of the proof $\pi$ are queried when conditioned on $r$.
+This yields a polynomially-sized $\mathsf{qCSP}$ instance $\I$.
+Furthermore, since the verifier $V$ runs in polynomial time, it's execution can be simulated on all $r$ to output the instance $\I$ in polynomial-time.
+Thus, we have given a polynomial-time reduction from an input $x$ to its corresponding CSP instance $\I$, so Theorem \ref{pcptheoremgapcsp}.
+\newline
+
+Conversely, suppose we had a reduction from $\NP$ to $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ as stated in Theorem \ref{pcptheoremgapcsp}.
+We devise polynomial-time reduction taking an instance $\mathsf{Max}$-$\mathsf{qCSP}$ to a polynomial-time PCP verifier $V$ using logarithmic number of random bits and a constant number of queries to the supplied proof $\pi$.
+For an input $x \in \{0,1\}^n$, the proof will be expected to be an assignment to its respective instance $f(x)$ in the notation utilized in Theorem \ref{pcptheoremgapcsp}.
+Verifier $V$ makes coin flips $r \in \{0,1\}^{c\log{n}}$ to choose one constraint tuple $(S,\psi)$ where $\psi$ is some $q$-ary predicate.
+Only a logarithmic number of random bits are required to query any constraint in instance $f(x)$ as the polynomial-time $\NP$ reduction can only generate a polynomial number of such constraints.
+The PCP only has to make $q$-queries to the proof $\pi$ to find the assignments to the variables listed in $S = (v_1,\cdots,v_q)$.
+By the properties listed in Theorem \ref{pcptheoremgapcsp}, the PCP verifier $V$ must have completeness $1$ and soundness $\leq \frac{1}{2}$ as claimed.
+\end{proof}
+
+\begin{lemma} (Equivalence between $\mathsf{Gap}$-$\mathsf{3SAT}$ and $\mathsf{Max}$-$\mathsf{qCSP}$). Theorem \ref{pcpgapsat} and Theorem \ref{pcptheoremgapcsp} are equivalent.
+\end{lemma}
+%
 \begin{proof}
@@
 %2 The PCP Theorem: An Introduction (DIMACS)
+
 %\subsection{Examples of PCPs}

### 2021-11-23 Complexity/UGC/hastad.tex
Yet more touches to UGC paper

@@
 \end{corollary}
+%
+\begin{proof}
+  From Theorem \ref{noisyblrsoundness}, we deduce that:
+  \begin{align*}
+      2\nu \leq  \sum_{S \subseteq [n]} \widehat{f}(S)^3 (1-2\epsilon)^{|S|} & \leq \max_{S}{\widehat{f}(S)}(1-2\epsilon)^{|S|} \sum_{S \subseteq [n]} \widehat{f}(S)^2 \\
+      & = \max_{S}{\widehat{f}(S)}(1-2\epsilon)^{|S|}
+  \end{align*}
+  By taking the maximum attained as $S^*$,
+  \[2\nu \leq \widehat{f}(S^*)(1-2\epsilon)^{|S^*|} \]
+  Since both $\widehat{f}(S^*) \geq 2\nu$ and $(1-2\epsilon)^{|S^*|} \geq 2\nu$ (as $f$ is boolean):
+
+  \[e^{-2\epsilon|S^*|} \geq 2\nu \]
+  which shows that $|S^*| \leq O\left(\frac{1}{\epsilon}\log{\frac{1}{\nu}}\right)$.
+\end{proof}
 
@@
 More formally, the input will be a $\LC$ instance $((A\sqcup B,E),\Sigma_A,\Sigma_B,\{\pi_e\}_{e \in E})$ and the proof will be the long codes encoding the assignments to the vertices, $\mathfrak{A}: A \rightarrow \Sigma_A$, $\mathfrak{B}: B \rightarrow \Sigma_B$.
-The consistency check will check validity of the long code as well as checking the labels encoded by the long codes satisfy the projection constraints $\{\pi_e\}_{e \in E}$: First some notation is in order.
+The consistency check will aim to perform two tasks simultaneously given an edge $(a,b)$ and its associated projection constraint $\pi$: Let $f_a:\{-1,1\}^{|\Sigma_A|} \rightarrow \{-1,1\}, \; f_b:\{-1,1\}^{|\Sigma_B|} \rightarrow \{-1,1\}$ be the functions claimed to be long codes for the labels assigned to vertices $a,b$: $\mathfrak{A}(a),\mathfrak{B}(b)$.
+Recall these will be encoded as truth table strings on the proof string $\pi$.
+
+\begin{description}
+  \item[Validity Check] Check that the functions $f_a,f_b$ are indeed valid long codes for some labels $\mathfrak{A}(a),\mathfrak{B}(b)$.
+  \item[Consistency Check] Verify that the assigned labels satisfy $\pi$
+  \[ \pi(\mathfrak{A}(a)) =  \mathfrak{B}(b)\]
+\end{description}
+
+Before presenting the test, some notation is in order.
 Let $(x \circ \pi)_i = x_{\pi(i)}, \; \forall i \in \Sigma_A$ for $x \in \{-1,1\}^{|\Sigma_B|}$.
@@
 
-  \noindent Let $f_a:\{-1,1\}^{|\Sigma_A|} \rightarrow \{-1,1\}, \; f_b:\{-1,1\}^{|\Sigma_B|} \rightarrow \{-1,1\}$ be the functions claimed to be long codes for the labels assigned to vertices $a,b$: $\mathfrak{A}(a),\mathfrak{B}(b)$.
-Recall these will be encoded as truth table strings on the proof string $\pi$.
   \item Sample $x \sim \{-1,1\}^{|\Sigma_B|}, \; y \sim \{-1,1\}^{|\Sigma_A|}$

### 2021-11-24 Complexity/UGC/approx.tex
Grammar and additions to pcp section

@@
 \subsection{Definitions and Examples}
-The theory of approximating $\NP$-hard problems roots itself in the following question: ``Is it possible to efficiently approximate $\NP$-complete problem to some arbitrary degree of accuracy?" Since Cook-Levin result demonstrated that the $\SAT$ decision problem is $\NP$-complete, the question could be rephrased as ``Can we find an efficient algorithhm for $\SAT$ which outputs an assignment which satisfies a $1-\delta$ fraction of the clauses for any constant $\delta >  0$?" To this end, let us first introduce a few motivating examples which will find utility in our analysis.
+The theory of approximating $\NP$-hard problems roots itself in the following question: ``Is it possible to efficiently approximate $\NP$-complete problem to some arbitrary degree of accuracy?" Since Cook-Levin result demonstrated that the $\SAT$ decision problem is $\NP$-complete, the question could be rephrased as ``Can we find an efficient algorithhm for $\SAT$ which outputs an assignment which satisfies a $1-\delta$ fraction of the clauses for any constant $\delta >  0$?" To this end, let us first introduce a few motivating examples which will find utility in our forthcoming analysis.
 \newline
@@
 
+An polynomial-time algorithm which solves $\MaxCNF$ is one which takes a 3CNF $\phi$ as input and outputs $\Opt(\phi)$.
 Naturally, $\mathsf{Max3CNF}$ is $\NP$-hard since its corresponding decision problem $\mathsf{3CNF}$ is $\NP$-complete i.e $\phi$ is satisfiable iff $\mathsf{Opt}(\phi) = 1$.
@@
 \begin{align}
-  &  \min \sum_{(u,v) \in E} e_{u,v} \\
-  & e_{u,v} \geq
-  \max \begin{cases}
+  &  \max \sum_{(u,v) \in E} e_{u,v} \\
+  & e_{u,v} \leq
+  \min \begin{cases}
       x_1 + x_2 \\
@@
   \end{cases} \\
-  & e_{u,v} \in [1,2] \\
+  & e_{u,v} \in [0,1] \\
   & x_v \in [0,1]
@@
               1 \text{ if } e_{u,v} \in \mathcal{C} \\
-              2 \text{ otherwise }
+              0 \text{ otherwise }
             \end{cases}
@@
   \end{align}
+  Furthermore, an algorithm $A$ decides $\MaxCSP$ if for input instance $\I$ it accepts iff $\I \in \mathsf{Yes}$ and rejects iff $\I \in \mathsf{No}$.
+If $\I$ is in neither $\mathsf{Yes}, \mathsf{No}$, we do not care what the algorithm outputs.
 \end{definition}
+
+In particular, we deem a $\GapMaxCSP{\alpha}{\beta}$ problem as $\NP$-hard if for every language $L \in \NP$, there exists a polynomial-time reduction $f$ taking $x$ to CSP $\Psi$ instances such that:
+\begin{align*}
+  x \in L \implies \mathsf{Opt}(f(x)) & \geq \beta \\
+  x \not\in L \implies \mathsf{Opt}(f(x))&  < \alpha
+\end{align*}
 

### 2021-11-24 Complexity/UGC/pcp.tex
Grammar and additions to pcp section

@@
 \begin{theorem} \label{pcpgapsat}($\PCP$ Theorem: $\mathsf{Gap3SAT}$-hardness)
-The problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max3SAT}$ is $\NP$-hard. In other words, for every $\NP$ language $L$, there exists a polynomial-time reduction $f$ mapping $L$ to 3CNF formulas such that:
+The problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max3SAT}$ is $\NP$-hard.
+In other words there exists a constant $\alpha  \leq 1$ such that, for every $\NP$ language $L$, there exists a polynomial-time reduction $f$ mapping $L$ to 3CNF formulas such that:
 
@@
   x \in L & \implies \mathsf{Opt}(f(x)) = 1 \\
-  x \not\in L & \implies \mathsf{Opt}(f(x)) \leq \frac{1}{2}
+  x \not\in L & \implies \mathsf{Opt}(f(x)) \leq \alpha
 \end{align*}
@@
 
-\begin{theorem} \label{pcptheoremgapcsp}  ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) For some constants $q \in \mathbb{N}$, the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard.
-To elaborate, for every $\NP$ language $L$, there exists a polynomial time reduction mapping an $L$ to a instance $f(x)$ of some CSP $\Psi$ where $\Psi$ consists of $q$-ary predicates, such that
+\begin{theorem} \label{pcptheoremgapcsp}  ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) There exists constants $q \in \mathbb{N}$ and $\alpha \leq 1$ where the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard.
+To elaborate, for every $\NP$ language $L$, there exists a polynomial time reduction mapping an $L$ to a instance $f(x)$ of some CSP $\Psi$ over domain $\Omega=\{0,1\}$ where $\Psi$ consists of $q$-ary predicates, such that
 
@@
 
-\begin{lemma} (Equivalence between $\mathsf{Gap}$-$\mathsf{3SAT}$ and $\mathsf{Max}$-$\mathsf{qCSP}$). Theorem \ref{pcpgapsat} and Theorem \ref{pcptheoremgapcsp} are equivalent.
+\begin{lemma} (Equivalnce between $\mathsf{GapMaxCSP}$and $\mathsf{GapMax3SAT}$ Theorem \ref{pcpgapsat} and Theorem \ref{pcptheoremgapcsp} are equivalent.
 \end{lemma}
@@
 \begin{proof}
-To be written...
+Since any $\mathsf{3Sat}$ instance can be seen as a particular type of $\mathsf{3CSP}$ instance, one direction is immediate.
+Conversely, if we assume Theorem \ref{pcptheoremgapcsp}, we aim to find some constant $\alpha'$ such that there exist a reduction from an instance of $\mathsf{Gap}_{1-\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ for $q \in \mathbb{N}$ claimed in Theorem \ref{pcptheoremgapcsp} to an instance of $\mathsf{Gap}_{1-\alpha',1}\mathsf{Max3Sat}$.
+%
+Let the CSP $\Psi$ be comprised $q$-ary predicates $\psi$.
+For an instance $\I$ of $\Psi$, an constraint tuple $(S,\psi)$ can be expressed as a logical AND of $2^q$ clauses where each clause is a logical OR of $q$ variables or their negations.
+In other words, $(S,\psi)$ is essentially a CNF of width $q$ and of size at most $2^q$.
+We can then construct an ``equivalent" 3CNF as follows: add extra symbols $\Pi_{1}, \cdots, \Pi_{(q-3)2^{q}}$.
+It can be shown that there exists a 3CNF $\psi'_S$ of size at most $(q-2)2^{q}$ such that:
+
+\begin{enumerate}
+  \item For every $x \in \{0,1\}^q$ which causes $\psi(x) = 1$, there exists an assignment $\Pi$ such that $\psi'(x,\Pi) = 1$.
+  \item Else if $\psi(x) = 0$, then for all assignments $\Pi$, $\psi'(x,\Pi) = 0$
+\end{enumerate}
+
+Now take the total 3CNF defined by a conjection of all such $\psi'_S$:
+\[ \psi_{\I} = \bigwedge_{(S,\psi) \in \I}  \psi_S' \]
+
+The total formula $\psi_\I$ is determined by at most $mq2^q$ number of clauses and at most $n + m(q-3)2^q$ number of variables.
+%
+If an instance $\I$ has all of its constraints satisfiable by some assignment, then by proprety (1) listed above, there must exist a assignment to the variables of the $\psi_{\I}$ such that it is satisifed.
+%
+On the other hand, if for all assignments to instance $\I$ only satisfy at most an $1 - \alpha$ fraction of constraints, then fraction of clauses satisfied can be at most $1 - \alpha + \alpha(1 - \frac{1}{(q-2)2^qa}) = 1 - \frac{\alpha}{(q-2)2^q}$.
+This is due to the fact that for each unsatisfied $\psi_S'$, the most number of its clauses which can be satisfied by any assignment will be $(q-2)2^q - 1$.
+Taking $\alpha' = \frac{\alpha}{(q-2)2^q}$
+yields the required constant.
 \end{proof}

### 2021-11-24 Complexity/UGC/uniquegames.tex
Grammar and additions to pcp section

@@
 \begin{definition}
-  A problem $P$ is said to be \emph{$\UG$-hard} if there is some constant $\delta > 0$ such that for all $m \in \mathbb{N}$, there exists a polynomial-time reduction from $\GapDeltaULC$ to $P$.
+  A promise problem $P$ is said to be \emph{$\UG$-hard} if there is some constant $\delta > 0$ such that for all $m \in \mathbb{N}$, there exists a polynomial-time reduction $f$ from $\GapDeltaULC$ to $P$, in the sense that for a $\GapDeltaULC$ instance $\I$:
+  \begin{align*}
+    \mathsf{Opt}(\I)  \geq 1 - \delta & \implies f(x) \in \mathsf{Yes} \\
+    \mathsf{Opt}(\I)  < \delta & \implies f(x) \in \mathsf{No}
+  \end{align*}
 \end{definition}

### 2021-11-26 Complexity/UGC/hastad.tex
pretty much finished draft of ugc writeup

@@
 
-Note that if the labels assigned $a,b$ satisfied $\pi$ and $f_a,f_b$ are valid long codes: $f_a(x \circ \pi) = f_b(x )$ for all $x \in \{-1,1\}^{|\Sigma_B|}$.
+First, note that if the labels assigned $a,b$ satisfied $\pi$ and $f_a,f_b$ are valid long codes: $f_a(x \circ \pi) = f_b(x )$ for all $x \in \{-1,1\}^{|\Sigma_B|}$.
+Second, Step 4 of the verfier can be expressed as a three variable linear equation:
+
+\begin{equation} \label{hastadpred}
+ f_a(z) f_b(x)f_a(y) = 1
+\end{equation}
 
@@
 \subsection{Folded Functions}
+A technical detail requires that functions queried $f_a,f_b$ be \emph{folded} in the sense that:
+%
+\begin{definition}
+A boolean function $f:\{-1,1\}^n \rightarrow \{-1,1\}$ is said to be folded if for every $x \in \{-1,1\}$, $f(-x) = -f(x)$.
+\end{definition}
+
+These functions are also said to \emph{odd}.
+It can be shown that \emph{folded} functions have their Fourier weights concentrated on subsets of odd cardinality.
+In particular, for boolean function $f$ folded: $\widehat{f}(\emptyset) = 0$.
+Why would we need this property?
+One issue that arises with the predicates shown in Equation \ref{hastadpred} is when all functions are simply the constant function $f_a = f_b = 1$.
+This passes the test with probability one.
+This can be avoided by restricting the functions to be folded.
+Of course, dictator functions are folded.
+Furthermore, we do not need to increase the number of queries to enforce this requirement: for any two pair of queries $(x,-x)$ and a queried function $f$, if the test queries $f(x)$, the value $f(-x)$ can be subsequently computed as $-f(x)$.
 

### 2021-11-26 Complexity/UGC/labelcover.tex
Grammar and additions to pcp section
pretty much finished draft of ugc writeup

@@
 %Arora, Barak
-\subsection{Some Structural Results of PCPs}
 
 \subsection{Raz's Parallel Repetition Theorem}
+In this section, we give a cursory outline of Raz's Parallel Repetition Theorem and its consequences. First, we can extend Definition \ref{pcpdef} to include proof strings over non-boolean alphabets.
+%
+To accomdate such verifiers, we can extend the definition to $\PCP^\Sigma_{\beta, \alpha}(r(n),q(n))$, the class of languages with polynomial-time verifiers taking proof strings $\pi$ over non-boolean alphabet $\Sigma$ and uses $r(n)$ random bits and at most $q(n)$ queries to $\pi$.
+%
+The following theorem shows that one can reduce the number of queries to two at the expense of a weaker soundness constant:
+%
+\begin{theorem} \label{2queryinclusion}
+  $\PCP^\Sigma_{1 - \alpha, 1}(r(n),q(n)) \subseteq \PCP^{\Sigma^q}_{1 - \frac{\alpha}{q},1}(r(n) + \log{q(n)}, 2)$
+\end{theorem}
+%
+\begin{proof}
+Given a verifier for a language $L \in \PCP^\Sigma_{\beta, 1 - \alpha}(r(n),q(n))$, it behaves as such on input $x \in \{0,1\}^n$:
+%
+\begin{enumerate}
+  \item Flips a $r(n)$ coins which we denote as $R$.
+  \item Using $R$, it computes a series of indices $i_1, \cdots i_q$ where $q = q(n)$.
+  \item It queries $\pi_{i_1}, \cdots \pi_{i_q}\in \Sigma$ from the proof $\pi$.
+  \item Finally, it feeds the symbols in a predicate $V_{x,R}(\pi_{i_1}, \cdots \pi_{i_q})$ which outputs ``accept" or ``reject".
+\end{enumerate}
+We extend this to a modified verfier $V'$ to accomodate the concatenation of two proof strings: if $m = |\Sigma|$ $\pi_1:[m]^q$ then $\rightarrow \Sigma^{q}$ and $\pi_2:[m]\rightarrow \Sigma$:
+%
+\begin{enumerate}
+  \item Flips a $r(n)$ coins which we denote as $R$.
+  \item Using $R$, it computes a series of indices $i_1, \cdots i_q$ where $q = q(n)$
+  \item Queries $\pi' = \pi_1(i_1, \cdots, i_q)$
+  \item Computes the predicate:
+  \[ V_{x,R}'(\pi') = V_{x,R}(\pi'_1,\cdots,\pi'_q) \]
+  \item Finally, chooses random $\ell \in [q]$ and checks if $\pi_2(i_\ell) = \pi'_\ell$.
+  \item The verifier accepts iff both checks pass.
+\end{enumerate}
+
+For completeness, if there exists a proof $\pi$ which the original verifier $V$ accepts with probability one, then setting $\pi_1(i_1, \cdots, i_q) = (\pi_{i_1},\cdots,\pi_{i_q})$ and $\pi_2 = \pi$ will induce $V'$ to accept with probability one by construction.
+%
+As for soundness, suppose  at least $\alpha$ fraction of the predicates $V_{x,R}$ reject in respect to some proof string $\pi$.
+Each one of the predicates contained in this fraction must fail the check at Step $4$ if $\pi' = \pi_1(i_1, \cdots, i_q) = (\pi_{i_1},\cdots,\pi_{i_q})$.
+However, $\pi_1$ does not have to respect this rule for any $q$-tuple of queries $(i_1, \cdots, i_q)$ given as input.
+If so, there must exist at least one index $\ell$ such that $\pi'_\ell \neq \pi_2(i_\ell) = \pi_\ell$.
+Since such an index is chosen with probability $\frac{1}{q}$, we have that $V'$ must reject with probability at least $\frac{\alpha}{q}$.
+This completes the proof.
+\end{proof}
+
+%
+Raz's Parallel Repetition Theorem constructed a verifier which allowed the soundness constant to drop exponentially with the cost of bloating the alphabet size:
+\begin{theorem} (Raz's Parallel Repetition Theorem \ref{raz1998parallel}) \label{razparallelrep}
+  For all $s \in (0,1)$, there exists $c_s \in (0,s)$ such that:
+    \[ \PCP^{\Sigma}_{s,1}(r,2) \subseteq \PCP^{\Sigma^t}_{c_s^t,1}(rt,2)  \]
+\end{theorem}
+
+\begin{corollary} \label{razcorollary} For all $\epsilon > 0$, there exists an alphabet $\Sigma$ such that $|\Sigma| \leq \mathsf{poly}(\frac{1}{\epsilon})$:
+  $$ \NP \subseteq \PCP^{\Sigma}_{\epsilon,1}(O(\log{n}\cdot\log{1/\epsilon}),2) $$
+\end{corollary}
+%
+\begin{proof}
+  For some constant $Q > 0$,
+  \begin{align*}
+    \NP & \subseteq \PCP_{\frac{1}{2},1}^{\{0,1\}}(O(\log{n}), Q) \\
+    & \subseteq \PCP_{1- \frac{1}{2Q},1}^{\{0,1\}^Q}(O(\log{n}), 2) \\
+    & \subseteq \PCP_{c_Q^t,1}^{\{0,1\}^{Qt}}(O(t\log{n}), 2) \\
+    & \subseteq   \PCP_{\epsilon,1}^{\Sigma}(O(\log{n}\cdot\log{1/\epsilon}), 2)
+  \end{align*}
+  where $|\Sigma| \leq \mathsf{poly}(\frac{1}{\epsilon})$. The second inclusion follows from Theorem \ref{2queryinclusion} and the third inclusion follows from Theorem \ref{razparallelrep}.
+\end{proof}
+
+By combining Corollary \ref{razcorollary}, the $\NP$-hardness of Label Cover comes to fruition:
 
@@
 \end{theorem}
+
+This is shown by embedding the query pairs of the verifier into the projection constraints of a Label Cover instance graph.

### 2021-11-26 Complexity/UGC/pcp.tex
pretty much finished draft of ugc writeup

@@
 
-\begin{definition}
+\begin{definition} \ref{pcpdef}
   Given a language $L$ and $r,q: \mathbb{N} \rightarrow \mathbb{N}$, a \emph{(r(n),q(n))-$\PCP$-verifier} for $L$ consists of a polynomial-time algorithm $V$ with the following properties: \newline
@@
   Sometimes the completeness criterion is too strong for our purposes (see the comments on $\ELin$ problem).
-In these cases, we like to denote the class $\PCP_{\beta, \alpha}(r(n),q(n))$ as the languages $L$ which have a $(r(n),q(n))$-$\NP$-verifier such that the completeness and soundness criteria are amended as below: \newline
+In these cases, we like to denote the class $\PCP_{\alpha, \beta }(r(n),q(n))$ as the languages $L$ which have a $(r(n),q(n))$-$\NP$-verifier such that the completeness and soundness criteria are amended as below: \newline
 
@@
 Here, $\beta$ is the \emph{completeness parameter} while $\alpha$ is the \emph{soundness parameter}.
-The class introduced in the original definition would thus be denoted as $\PCP_{1,\frac{1}{2}}(r(n),q(n))$.
+The class introduced in the original definition would thus be denoted as $\PCP_{\frac{1}{2},1}(r(n),q(n))$.
 PCP verifiers whose completeness parameter is one ($\beta = 1$) is deemed as \emph{perfectly complete}.
@@
 \begin{equation}
-\NP = \PCP_{1,\frac{1}{2}}(O(\log{n}), O(1))
+\NP = \PCP_{frac{1}{2},1}(O(\log{n}), O(1))
 \end{equation}
@@
 \begin{proposition}
-For every constants $Q \in \mathbb{N}, c > 0$, $\PCP_{1,\frac{1}{2}}(c\cdot\log{n}, Q) \subseteq \NP$
+For every constants $Q \in \mathbb{N}, c > 0$, $\PCP_{\frac{1}{2}, 1}(c\cdot\log{n}, Q) \subseteq \NP$
 \end{proposition}
@@
 \begin{proof}
-Begin with the observation that $\PCP_{1,\frac{1}{2}}(r(n), q(n)) \subseteq \NTIME(q(n)2^{r(n)})$.
+Begin with the observation that $\PCP_{\frac{1}{2},1}(r(n), q(n)) \subseteq \NTIME(q(n)2^{r(n)})$.
 This is justified by the view of an $\NTIME$ machine simulating the verifier by trying all possible coin tosses and queries to the input string $x$ and proof string $\pi$.
@@
 
-\begin{theorem} \label{pcptheoremgapcsp}  ($\PCP$-Theorem: $\mathsf{GapMaxCSP}$- hardness) There exists constants $q \in \mathbb{N}$ and $\alpha \leq 1$ where the problem $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ is $\NP$-hard.
+\begin{theorem} \label{pcptheoremgapcsp}  ($\PCP$-Theorem: $\mathsf{GapMaxqCSP}$- hardness) There exists constants $q \in \mathbb{N}$ and $\alpha \leq 1$ where the problem $\mathsf{Gap}_{\alpha,1}\mathsf{MaxqCSP}$ is $\NP$-hard.
 To elaborate, for every $\NP$ language $L$, there exists a polynomial time reduction mapping an $L$ to a instance $f(x)$ of some CSP $\Psi$ over domain $\Omega=\{0,1\}$ where $\Psi$ consists of $q$-ary predicates, such that
@@
 \begin{proof}
-First, assume $\NP = \PCP_{1,\frac{1}{2}}(O(\log{n}), O(1))$.
+First, assume $\NP = \PCP_{\frac{1}{2},1}(O(\log{n}), O(1))$.
 We will outline a procedure to convert the verifier $V$ into an instance $\I$ for a $q$-ary CSP $\Psi$ for some constant $q$.
@@
 
-Conversely, suppose we had a reduction from $\NP$ to $\mathsf{Gap}_{\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ as stated in Theorem \ref{pcptheoremgapcsp}.
-We devise polynomial-time reduction taking an instance $\mathsf{Max}$-$\mathsf{qCSP}$ to a polynomial-time PCP verifier $V$ using logarithmic number of random bits and a constant number of queries to the supplied proof $\pi$.
+Conversely, suppose we had a reduction from $\NP$ to $\mathsf{Gap}_{\alpha,1}\mathsf{MaxqCSP}$ as stated in Theorem \ref{pcptheoremgapcsp}.
+We devise polynomial-time reduction taking an instance $\mathsf{MaxqCSP}$ to a polynomial-time PCP verifier $V$ using logarithmic number of random bits and a constant number of queries to the supplied proof $\pi$.
 For an input $x \in \{0,1\}^n$, the proof will be expected to be an assignment to its respective instance $f(x)$ in the notation utilized in Theorem \ref{pcptheoremgapcsp}.
@@
 
-\begin{lemma} (Equivalnce between $\mathsf{GapMaxCSP}$and $\mathsf{GapMax3SAT}$ Theorem \ref{pcpgapsat} and Theorem \ref{pcptheoremgapcsp} are equivalent.
+\begin{lemma} (Equivalnce between $\mathsf{GapMaxqCSP}$and $\mathsf{GapMax3SAT}$ Theorem \ref{pcpgapsat} and Theorem \ref{pcptheoremgapcsp} are equivalent.
 \end{lemma}
@@
 Since any $\mathsf{3Sat}$ instance can be seen as a particular type of $\mathsf{3CSP}$ instance, one direction is immediate.
-Conversely, if we assume Theorem \ref{pcptheoremgapcsp}, we aim to find some constant $\alpha'$ such that there exist a reduction from an instance of $\mathsf{Gap}_{1-\alpha,1}\mathsf{Max}$-$\mathsf{qCSP}$ for $q \in \mathbb{N}$ claimed in Theorem \ref{pcptheoremgapcsp} to an instance of $\mathsf{Gap}_{1-\alpha',1}\mathsf{Max3Sat}$.
+Conversely, if we assume Theorem \ref{pcptheoremgapcsp}, we aim to find some constant $\alpha'$ such that there exist a reduction from an instance of $\mathsf{Gap}_{1-\alpha,1}\mathsf{MaxqCSP}$ for $q \in \mathbb{N}$ claimed in Theorem \ref{pcptheoremgapcsp} to an instance of $\mathsf{Gap}_{1-\alpha',1}\mathsf{Max3Sat}$.
 %
@@
 In other words, $(S,\psi)$ is essentially a CNF of width $q$ and of size at most $2^q$.
-We can then construct an ``equivalent" 3CNF as follows: add extra symbols $\Pi_{1}, \cdots, \Pi_{(q-3)2^{q}}$.
+We can then construct an ``equivalent" 3CNF $\psi'_S$ as follows: add extra symbols $\Pi_{1}, \cdots, \Pi_{(q-3)2^{q}}$.
 It can be shown that there exists a 3CNF $\psi'_S$ of size at most $(q-2)2^{q}$ such that:
@@
 
+\begin{remark}
+  The existence of the ``equivalent" 3CNF is outlined in Problem 7.11 of the O'Donnell textbook \cite{o2014analysis}.
+ \end{remark}
+
+\begin{remark}
+The proof of the PCP Theorem (Theorem \ref{pcptheorem}) is quite involved and out of the scope of these notes.
+The original proof hinged on clever algebraic techniques such as low-degree testing \cite{arora1998probabilistic}.
+A simpler combinatorial proof by Dinur arrived years later  \cite{dinur2007pcp}.
+For more information, refer to the cited publications or Chapter 22 of the Arora, Barak textbook \cite{arora2009computational}.
+\end{remark}
+
 %2 The PCP Theorem: An Introduction (DIMACS)

### 2021-11-26 Complexity/UGC/ugcmaxcut.tex
Yet more touches to UGC paper
Grammar and additions to pcp section
pretty much finished draft of ugc writeup

@@
 The completeness and soundness parameters immediately follow from definition of $\mathsf{Gap}_{\alpha,\beta}\mathsf{MaxCut}$.
-Conversely, assume the existance of a 2-query PCP for $\GapDeltaULC$ with the above properties.
+Conversely, assume the existence of a 2-query PCP for $\GapDeltaULC$ with the above properties.
 We will calculate the acceptance probabilities of this PCP given a proof string and an input instance by finding the max cut.
 Let the vertices be the proof locations of the input proof string $\pi$.
-It suffcies to draw an edge between two vertices weighted by the probability their corresponding proof locations are queried by the verifier.
+It suffices to draw an edge between two vertices weighted by the probability their corresponding proof locations are queried and accepted by the verifier.
 \end{proof}
@@
 %
-\subsection{Dictatorship Testing Redux}
+\subsection{Dictatorship Testing and Majority is the Stablest (MIS)}
+A key difference between dictatorship test crafted in H\aa stad's 3-bit PCP for $\NP$ and the one will aim to utilize here is the number of queries the PCP can make.
+Namely, the reduction here can only use \emph{two} queries to the proof rather than the three queries afforded to H\aa satd's verifier.
+The dictatorship test we will focus on for this reduction utilizes only two queries to the functions: given input boolean function $f:\{-1,1\}^n \rightarrow \{-1,1\}$ and $\rho \in (-1,0)$:
 
+\begin{enumerate}
+  \item Sample $x \sim \{-1,1\}^n$
+  \item Sample $y \sim N_{\rho}(x)$
+  \item Accept iff $f(x) \neq f(y)$
+\end{enumerate}
 %
-%
-%
-\subsection{Majority is the Stablest (MIS)}
-%
-Before we craft our dictator test, we require a tool bounding the noise sensitivity of a boolean function with small low-degree influence. This is captured by the ``Majority is the Stablest" theorem:
+Before we analyze the acceptance probability of our dictatorship test, we require a tool bounding the noise sensitivity of a boolean function with small low-degree influence.
+This is captured by the ``Majority is the Stablest" theorem:
 
@@
 
-A generalization of the MIS theorem will serve our purposes:
+To see the significance of the MIS Theorem, we compute the acceptance rate of the dictatorship test above:
+
+\begin{align*}
+\mathbb{P}\left[ \text{ Test Accepts }\right] & = \mathbb{E}_{x,y}\left[\frac{1}{2} - \frac{1}{2}f(x)f(y) \right] \\
+& = \frac{1}{2} - \frac{1}{2}\mathbb{E}_{x,y}\left[f(x)f(y) \right] \\
+& = \frac{1}{2} - \frac{1}{2}\mathsf{Stab}_\rho(f)
+\end{align*}
+
+We can now reason about acceptance probability in the context of several classes of functions:
+
+\begin{itemize}
+  \item Dictators pass the test with probabilty $\frac{1 - \rho}{2}$
+  \item The parity functions $\chi_S,\; S \subseteq [n]$ pass with probability $\frac{1 - \rho^{|S|}}{2}$. As $|S|$ grows, the acceptance probability tends to $\frac{1}{2}$.
+  \item Majority functions passes the test with probability approaching in the limit $\frac{\cos^{-1}(\rho)}{\pi}$.
+\end{itemize}
+
+What the MIS Theorem roughly tells us is that functions with no ``notable" coordinates, i.e all coordinates have small influence, can pass the test with probability at most $\frac{\cos^{-1}(\rho)}{\pi}$.
+The proof of the soundness of our verifier will exploit the observation that functions with a sufficiently large $i^{th }$ influence for some coordinate $i$ will be ``close enough" to a dictator.
+This property will be used to extract a labeling to the $\mathsf{UniqueLC}(m)$ instance satisfying a large fraction of constraints.
+
+For the analysis of constructed PCP verifier, a generalization of the MIS theorem will serve our purposes:
 
@@
 
+where $\mathsf{Inf}^{\leq d}_i(f)$ refers to the $i^{th}$ influence of $f$ restricted to degrees of at most $d$:
+
+\begin{equation}
+  \mathsf{Inf}^{\leq d}_i(f) = \sum_{\substack{S \subseteq [n], \; i \in S \\ |S| \leq d}} |S|
+  \cdot\widehat{f}(S)^2
+\end{equation}
+
+Note that the image of our function $f$ is the contained in the closed interval $[-1,1]$ rather than the finite set $\{-1, 1\}$ as stated in the ordinary MIS theorem.
 %
@@
 as desired. This completes the proof of Theorem \ref{maxcutughard}.
-
 \end{proof}
 
+%We showed that the test to accept with probability %larger than $\GWConstant + \epsilon$ must
+
 %add a big picture composing both parts.

### 2022-07-02 thesis/frontmatter/acknowledgement.tex
Prelim thesis commit

%Acknowledgements are the author's statement of gratitude to and
%recognition of the people and institutions who helped the author's
%research and writing.

\begin{center}
\vspace*{52pt}
{\Large \textbf{ACKNOWLEDGEMENTS}}
\end{center}

I would like to thank the UNC Computer Science community for supporting my work and providing me a warm community for pursuing this research. In particular, I would like to thank my advisor, Parasara Sridhar Duggirala, for the opporunity to perform research and participate in his group.
%
%His choice to give me more autonomy to take classes at Duke University really helped me understand my academic interests better.
%
Bineet Ghosh, Manish Goyal, Abel Karimi, and Meghan Stuart were all great colleagues and I enjoyed my time exchanging ideas, proofreading drafts, and socializing over group lunches. I would also like to thank Stanley Bak for helping me improve the implementation of Kaa and providing valuable feedback during this research. Juan Garcia made a wonderful collegue and an equally wonderful friend. The nights we raided Peabody and Phillips hall was certainly a memorable time. Finally, I would like to thank my parents and my family for their ardent, un-ending support.

\clearpage

### 2022-07-02 thesis/intro.tex
Prelim thesis commit

\chapter{Introduction}

% Introduction from previous works, sources (ARCH, Formats)_
% Related works from same papers.

One of the most widely-used techniques for performing safety analysis of non-linear dynamical systems is reachable set computation
%
For example, reachability analysis has found many applications in formally verifying the safety properties of Cyber-physical Systems governed by Neural Network Controllers \cite{tran2019star, fan2020reachnn, bak2021nnenum}.
%
The reachable set is defined to be the set of states visited by at least one of the trajectories of the system starting from an initial set and propagated forward in time by a finite fixed number of steps.
%
Computing the exact reachable set for non-linear systems is challenging due to several reasons:
%
First, unlike linear dynamical systems whose solutions can be expressed as closed form, non-linear dynamical systems generally do not admit such a nice form.
%
Second, computationally speaking, current tools for performing non-linear reachability analysis are not very scalable. This is also in stark contrast to several scalable approaches developed for linear dynamical systems \cite{duggirala2016parsimonious, bak2017simulation}.
%
Finally, computing the reachable set using various set representations involves wrapping error which may be too conservative for practical use.
%
That is, the overapproximation acquired at a given step would increase the conservativeness of the overapproximation for all future steps.

One of the several techniques for computing the overapproximation of reachable sets for discrete non-linear systems is to encode the reachable set through parallelotope bundles.
%
Here, the reachable set is represented as a parallelotope bundle, an geometric data structure representing an intersection of several simpler objects called parallelotopes.
%
One of the advantages of this technique is its exploitation of a special form of non-linear optimization problem to overapproximate the reachable set.
%
The usage of a specific form of non-linear optimization mitigates many drawbacks involved with the scalability of non-linear analysis.

However, wrapping error still remains to be a problem for reachability using parallelotope bundles.
%
An immediate reason stems from the responsbility of the practitioner to define the template directions specifiying the parallelotopes.
%
Often, these template directions are selected to be either the cardinal axis directions or some directions from octahedral domains.
%
However, it is not certain that the axis-aligned and octagonal directions are optimal for computing reachable sets over general non-linear dynamics.
%
Additionally, even an expert user of reachable set computation tools may not be able to ascertain a suitable set of template directions for computing reasonably accurate over-approximations of the reachable set.
%
Picking unsuitable template directions would only cause the wrapping error to grow, leading to the aforementioned issue of overly conservative reachable sets.

In this thesis, we investigate techniques for generating template directions automatically and dynamically, which is the culmination of several publications in different venues \cite{kim2020kaa,kim2021automatic,geretti2021arch}.
%
Specifically, we propose a method where instead of the user providing the template directions to define the parallelotope bundle, he or she specifies the number of templates whose template directions are to be generated by our algorithm automatically.

To this end, we study two techniques for generating the said template directions.
%
First, we compute a local linear approximation of the non-linear dynamics and use the linear approximation to compute the template directions.
%
Second, we generate a set of trajectories sampled from within the reachable set and use Principal Component Analysis (PCA) over these trajectories.
%
We observe that the accuracy of the reachable set can be drastically improved by using templates generated using these two techniques.
%
To address scalability, we demonstrate that even when the size of the initial set increases, our template generation algorithm returns more accurate reachable sets than both manually-specified and random template directions.
%
Finally, we experiment with our dynamic template generation algorithm's effectiveness on approximating the reachable set of high-dimensional COVID19 dynamics proposed by the Indian Supermodel Committee \cite{indiansuper2020supermodel}. The results were published in an ACM blogpost detailing the utility of reachable set computation in modeling disease dynamics \cite{bak2021covid}.

% [Introduction para]asdf
% \begin{itemize}
% \item Reachable set computation.
% \item non-linear dynamics is challenging.
% \item Overapproximation, wrapping error.
% \item Parallelotope bundle reachability.
% \end{itemize}

% [Templates for reachability]
% \begin{itemize}
% \item Often the templates are generated statically.
% \item However, the most appropriate directions for templates that improve the accuracy is unknwon.
% \item Challenges: Wrapping error, cannot predict.
% \item Static template directions often aggrevate such errors.
% \item Generating templates dynamically is a challenge.
% \end{itemize}

% [Template generation]
% \begin{itemize}
% \item Template based overapproximation are used extensively in verification.
% \item More on this in the related work section.
% \item Dynamic templates using PCA have been investigated already.
% \item Not explicitly in the context of parallelotope bundles.
% \item In this paper we propose two techniques for  generating templates dynamically.
% \end{itemize}

\section{Related Work}
\label{sec:related}

Reachable set computation of non-linear systems using template polyhedra and Bernstein polynomials has been first proposed in~\cite{dang2009image}.
%
In~\cite{dang2009image}, Bernstein polynomial representation is used to compute an upper bound of a special type of non-linear optimization problem.
%
This enclosing property of Bernstein polynomials has been actively studied in the area of global optimization \cite{nataray2002algorithm, garloff2003bernstein, nataraj2007new}.
%
Furthermore, several heuristics have been proposed for improving the computational performance of optimization using Bernstein polynomials~\cite{smith2009fast,munoz2013formalization}.

Several improvements to this algorithm were suggested in~\cite{dang2012reachability, sassi2012reachability} and~\cite{dang2014parameter} extends it for performing parameter synthesis.
%Hey
The representation of parallelotope bundles for reachability was proposed in~\cite{dreossi2016parallelotope} and the effectiveness of using bundles for reachability was demonstrated in~\cite{dreossi2017sapo, dreossi2017reachability}.
%
However, all of these papers used static template directions for computing the reachable set.
%
In other words, the user must specify the template directions before the reachable set computation proceeds.

Using template directions for reachable set has been proposed in~\cite{sankaranarayanan2008symbolic} and later improved in~\cite{dang2011template}.
%
Leveraging the principal component analysis of sample trajectories for computing reachable set has been proposed in~\cite{stursberg2003efficient,chen2011choice,seladji2017finding}.
%
More recently, connections between optimal template directions for reachability of linear dynamical systems and bilinear programming have been highlighted in~\cite{gronski2019template}.
%
For static template directions, octahedral domain directions~\cite{clariso2004octahedron} remain a popular choice.

### 2022-07-02 thesis/prelim.tex
Prelim thesis commit

\chapter{Preliminaries}
\label{chap:prelim}

We begin with some basic definitions pertaining to reachability and parallelotopes. The definition of Bernstein polynomials and the reachable set computaion algorithm will be defined. Finally, an outline of the reachability algorithm given by \cite{dreossi2016parallelotope} for polynomial dynamical systems will be presented.

\section{Basic Definitions}
\label{sec:definitions}

As stated in the previous sections, this thesis pertains to the reachability analysis of dynamical systems. Roughly speaking, a dynamical system is governed by a set of differential equations such that the states of the system evolve according the solutions of the said differential equations.
%
The state of a system, denoted as $x$, lies in a domain $D \subseteq \reals^n$ where the solution to the differential equations is defined.
%
We restrict our attention to a specific definition of these dynamical systems:
%
\begin{definition}
A discrete-time nonlinear system is denoted as
\begin{equation}
  x^{+} = f(x)
\label{eq:sys}
\end{equation}
where $f: \reals^{n} \rightarrow \reals^{n}$ is a nonlinear function.
\end{definition}
%
 Intuitively, the function $f$ takes input a state of the system and outputs the next step of the system evolved according to the non-linear dynamics.
%
Here, the function $f$ generally represents some discretized version of some specified continuous non-linear dynamical systems. Recall that a dynamical system is considered \emph{linear} if its dynamics can be expressed as

$$
x' = Ax, \quad A \in \reals^{n \times n}
$$
%
Otherwise, we deem the system to be \emph{nonlinear}.
%
Hence, in particular, the function $f$ cannot be expressed as some matrix $A \in \reals^{n \times n}$.

Examples of prominent non-linear dynamical systems include the Lotka-Volterra predator-prey model \cite{wangersky1978lotka}, FitzHugh Neuron model \cite{fitzhugh1961impulses}, and recently introduced COVID19 disease model \cite{indiansuper2020supermodel}. Throughout this thesis, we discretize any continuous dynamics through the well-known Euler method. Thus, up to some error term of bounded degree, we can turn any non-linear system into the form given by Equation~\ref{eq:sys}.
%
\begin{example}
The SIR Epidemic model is a 3-dimensional dynamical system governed by the following continuous dynamics:
\begin{align} \label{eqn:sir}
    \begin{split}
        s' &= \beta \cdot s_k i_k \\
        i' &= \beta \cdot s_k i_k - \gamma \cdot i_k \\
        r' &= \gamma \cdot i_k
    \end{split}
\end{align}
where $s,i,r$ represent the fractions of a population of individuals designated as \textit{susceptible}, \textit{infected}, and \textit{recovered} respectively. There are two parameters, namely $\beta$ and $\gamma$, which influence the evolution of the system. $\beta$ is labeled as the contraction rate and $1/\gamma$ is the mean infective period.
%
Discretizing Equation~\ref{eqn:sir} according the Euler method yields the dynamics:
%
\begin{align} \label{eqn:discrete_sir}
    \begin{split}
        s_{k+1} &= s_k - (\beta \cdot s_k i_k)\cdot\Delta \\
        i_{k+1} &= i_k + (\beta \cdot s_k i_k - \gamma \cdot i_k)\cdot\Delta \\
        r_{k+1} &= r_k + (\gamma\cdot i_k)\cdot\Delta
    \end{split}
\end{align}
 Here, $\Delta$ is the discretization step and the index $k \in \mathbb{N}$ simply represents the current step.
%For the benchmarks, we set $\beta = 0.34$, $\gamma=0.05$, and $\Delta=0.5$.
%
Note the non-linear terms $s_ki_k$ which precludes the expression of the dynamics as a linear transformation.
\end{example}

The trajectory of a system that evolves according to Equation~\ref{eq:sys}, denoted as $\xi(x_0)$ is a sequence $x_0, x_1, \ldots $ where $x_{i+1} = f(x_i)$.
%
The $k^{th}$ element in this sequence $x_k$ is denoted as $\xi(x_0,k)$.

\begin{definition}
Given an initial set $\Theta \subseteq \reals^n$, the \emph{reachable set at step $k$}, denoted as $\Theta_k$ is defined as
\begin{equation}
  \Theta_k = \{ \xi(x,k)\: | \: x \in \Theta\}
\label{eq:reachset}
\end{equation}
If we set the number of steps to be some $n \in \mathbb{N}$, we say the \emph{reachable set} is
\begin{equation}
  \Theta = \bigcup_{i=1}^n \Theta_i
\end{equation}
\end{definition}

%
We will see in a future section an example of the reachable set of the discretized SIR model presented in Equation~\ref{eqn:discrete_sir}.

\section{Parallelotope-based Reachability}

\subsection{Parallelotopes}
\label{sec:parallelotope}
A parallelotope $P$ is a set of states in $\mathbb{R}^{n}$ captured by the tuple $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ is a matrix and $c \in \mathbb{R}^{2n}$ is a column vector. We impose the condition that $\Lambda_{i+n} = -\Lambda_{i}$ for all $i \in \{1, \ldots, n\}$ such that

$$
x \in P \mbox{ if and only if } \Lambda x \leq c.
$$
We deem $\Lambda$ as the \emph{direction matrix} where $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$. The column vector $c$ is called the \emph{offset vector}.
%
Alternatively, a parallelotope can also be represented in vertex-generator representation as $\langle v, g_1, \ldots, g_n\rangle$. Here $v \in \mathbb{R}^n$ is called the \emph{anchor} and $g_1, \ldots, g_n$, $g_i \in \mathbb{R}^n$, are called the \emph{generators}. The parallelotope is defined as the set:
$$
P := \{ x ~|~ \exists \alpha_1, \ldots, \alpha_n \in [0,1], \; x = v + \sum_{i=1}^n \alpha_i g_i \}
$$
This representation is very similar to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.

The vertex $v_1$ is obtained by solving the linear equation $\Lambda x = c_{l}$.
%
The $j+1$ vertex is obtained by solving the linear equation $\Lambda x = \mu_{j}$ where $\mu_{j}[i] = c_{l}[i]$ when $i \neq j$ and $\mu_{j}[j] = c_{u}[j]$.
%
The anchor $a$ of the parallelotope is the vertex $v_1$ and the generator $g_{i} = v_{i+1} - v_{1}$.

%
Notice that for a parallelotope $P$, the vertex-generator representation also defines the affine transformation that maps $[0,1]^{n}$ to $P$.
%
We denote this affine transformation as $T_{p}$.
%
\begin{definition}
A parallelotope bundle $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$ such that
  $$ Q = \bigcap_{i=1}^{m}P_i $$.
\end{definition}

\subsection{Bernstein Polynomials}
\label{sec:bernstein}

Given two multi-indices $i$ and $d$ of size $n$, where $i \leq d$, the Bernstein polynomial of degree $d$ and index $i$ is
$$
\mathcal{B}_{i,d} = \beta_{i_1,d_1}(x_1) \beta_{i_2,d_2}(x_2)\ldots \beta_{i_n,d_n}(x_n).
$$
%
where $\beta_{i_m,d_m}(x_m) = \binom{d_m}{i_m}x_{m}^{i_m}(1-x_m)^{d_m - i_m}$.
%
Any polynomial function can be expressed in the Bernstein basis.
%
The primary advantage of the Bernstein representation of a polynomial $h(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $h(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis.

In other words, given a polynomial $h(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $h(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $h(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
%
The upper and lower bounds of $h(x_1,\ldots,x_n)$ over $[0,1]^n$ are bounded by the Bernstein coefficients:
$$
min \{b_1, \ldots, b_m\} ~~\leq~~ inf_{x \in [0,1]^n} h(x) ~~~~~\leq~~~~ sup_{x \in [0,1]^{n}} h(x) ~~\leq~~ max \{b, \ldots, b_m\}.
$$

As mentioned earlier, a parallelotope $P$ can also be represented as an affine transformation $T_{p}$ from $[0,1]^{n}$ to $P$.
%
Therefore, upper bounds on the suprenum of a function $h$ over $P$ is equivalent to upper bound of $h \circ T_{p}$ over $[0,1]^{n}$.
%
A similar argument follows for the lower bound on the infimum.

\subsection{The Static Algorithm}
\label{sec:static}

### 2022-07-02 thesis/results.tex
Prelim thesis commit

\chapter{Experimental Results}

\subsection{Kaa}
\label{sec:kaa}

\subsection{Benchmarks}
\label{sec:benchmarks}

\subsection{COVID19 Supermodel}
\label{sec:covid}

\subsection{Comparison of Template Generation Techniques}
\label{sec:compare}

### 2022-07-02 thesis/prelim.tex
Static Algorithm section started

@@
 \label{sec:parallelotope}
-A parallelotope $P$ is a set of states in $\mathbb{R}^{n}$ captured by the tuple $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ is a matrix and $c \in \mathbb{R}^{2n}$ is a column vector.
+A parallelotope $P$ is a set of states in $\mathbb{R}^{n}$ captured by the tuple $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ is a matrix and $ç$ is a column vector.
 We impose the condition that $\Lambda_{i+n} = -\Lambda_{i}$ for all $i \in \{1, \ldots, n\}$ such that
-
-$$
+%
+\begin{equation}
+\label{eq:halfspace}
 x \in P \mbox{ if and only if } \Lambda x \leq c.
-$$
-We deem $\Lambda$ as the \emph{direction matrix} where $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$. The column vector $c$ is called the \emph{offset vector}.
+\end{equation}
+We deem $\Lambda$ as the \emph{template direction matrix} where $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$ called the \emph{$i^{th}$ template direction}.
+The column vector $c$ is called the \emph{offset vector} with $c_i$ denoting the $i^{th}$ element of $c$.
 %
-Alternatively, a parallelotope can also be represented in vertex-generator representation as $\langle v, g_1, \ldots, g_n\rangle$.
-Here $v \in \mathbb{R}^n$ is called the \emph{anchor} and $g_1, \ldots, g_n$, $g_i \in \mathbb{R}^n$, are called the \emph{generators}.
+If we unpack Equation~\ref{eq:halfspace}, we can re-express the inequalities as a conjection of half-space constraints.
+If we define $c_{u} = [c_1, c_2, \cdots, c_n]^T$ and $c_{l} = [c_{n+1}, c_{n+2}, \cdots, c_{2n}]^T$, then Equation~\ref{eq:halfspace} tells us that:
+%
+\begin{align*}
+  & \Lambda_i x \leq (c_u)_i \\
+  -& \Lambda_i x \leq (c_l)_i
+\end{align*}
+%
+Combining these two observation yields the \emph{half-space representation} of parallelotope $P$.
+%
+\begin{definition}
+The half-space representation of parallelotope $P$ is tuple  $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ and $\Lambda \in \mathbb{R}^{2n \times n}$ such that if $\overline{\Lambda}$ represents the submatrix of $\Lambda$ consisting of the first $n$ rows of $\Lambda$, then
+\begin{equation}
+\label{eq:halfspaceconst}
+P =\{x ~|~ -c_l \leq \overline{\Lambda} x \leq c_u\}
+\end{equation}
+\end{definition}
+%
+\noindent Alternatively, a parallelotope can also be represented in a \emph{generator representation}
+%
+\begin{definition}
+The generator representation of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$.
+The vector  $v \in \mathbb{R}^n$ is called the \emph{anchor} and the $g_i \in \mathbb{R}^n$, are called the \emph{generators}.
 The parallelotope is defined as the set:
@@
 $$
-This representation is very similar to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
+\end{definition}
+%
+\noindent This representation is very similar to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
+%
+In particular, a parallelotope is a special case of a zonotope where the number of generators is exactly the dimension of the system $n$.
 
-The vertex $v_1$ is obtained by solving the linear equation $\Lambda x = c_{l}$.
+There is a simple method to convert from the half-space representation of $P$ to its generator representation.
+%
+Obtain vertex $v_1$ by solving the linear equation $\Lambda x = c_{l}$.
 %
@@
 %
-The anchor $a$ of the parallelotope is the vertex $v_1$ and the generator $g_{i} = v_{i+1} - v_{1}$.
+The anchor $v$ of the parallelotope is the vertex $v_1$ and the generators will be $g_{i} = v_{i+1} - v_{1}$. Refer to \cite{dang2014parameter} for a more detailed exposition of this procedure.
 
 %
-Notice that for a parallelotope $P$, the vertex-generator representation also defines the affine transformation that maps $[0,1]^{n}$ to $P$.
+As a final remark, notice that for a parallelotope $P$, the generator representation also defines an affine transformation that maps $[0,1]^{n}$ to $P$.
 %
-We denote this affine transformation as $T_{p}$.
+Let us denote this affine transformation associated to $P$ as $T_P$.
+
+\begin{example}
+\label{ex:simple_ptope}
+Consider the 2D plane, namely $\reals^2$. We can construct a couple of simple examples of parallelotopes. First, if we define our parallelotope's template direction matrix to be
+%come back to this
+\end{example}
+
+\begin{example}
+\label{ex:ptope}
+Once again, consider the space $\reals^2$ and the parallelotope $P$ given in half-plane representation as $0 \leq x-y \leq 1$, $0 \leq y \leq 1$.
+%
+This is a parallelotope with vertices at $(0,0)$, $(1,0)$, $(2,1)$, and $(1,1)$.
+%
+In the half-space representation, the template directions of the parallelotope $P$ are given by the directions $[1, -1]$ and $[0, 1]$.
+%
+The half-space representation in matrix form is given as follows:
+
+\begin{equation}
+  \begin{bmatrix} 0 \\ 0 \end{bmatrix} \leq \begin{bmatrix}  1 & -1 \\ 0 &  1 \end{bmatrix}  \begin{bmatrix} x \\ y \end{bmatrix} \leq \begin{bmatrix} 1 \\ 1 \end{bmatrix}. \label{eq:ptopeexample}
+\end{equation}
+
+To compute the generator representation of $P$, we need to compute the \emph{anchor} and the \emph{generators}.
+%
+The anchor is obtained by solving the linear equations $x-y = 0, y = 0$.
+%
+Therefore, the anchor $a$ is the vertex at origin $(0,0)$
+%
+To compute the two generators of the parallelotope, we compute two vertices of the parallelotope.
+%
+Vertex $v_1$ is obtained by solving the linear equations $x - y = 1, y = 0$.
+%
+Therefore, vertex $v_1$ is the vertex $(1,0)$.
+%
+Similarly, vertex $v_2$ is obtained by solving the linear equations $x-y = 0, y = 1$.
+%
+Therefore, $v_2$ is the vertex $(1,1)$.
+%
+The generator $g_1$ is the vector $v_1 - a$, that is $(1,0)- (0,0) = (1,0)$
+%
+The generator $g_2$ is the vector $v_2 - a$, that is $(1,1) - (0,0) = (1,1)$.
+%
+Therefore, all the points in the paralellotope can be written as $(x,y) = (0,0) + \alpha_1 (1,0) + \alpha_2(1,1)$, $\alpha_1, \alpha_2 \in [0,1]$.
+\end{example}
+%
+%We denote this affine transformation as $T_{p}$.
 %
@@
 \end{definition}
+%
+\noindent This parallelotope bundle will be the geometric data structure which will enclose the region we compute to be the approximation of the exact reachable set.
 
@@
 \label{sec:bernstein}
+In this section, we define Bernstein polynomials and state some of their enclosure properties.
+A \emph{multi-index} ${\bf i}$ of length $n$ is defined as tuple of $n$ elements ${\bf i} = (i_1,\cdots, i_n)$ such that each $i_k \in \mathbb{N}$.
+Furthermore, we order the multi-indices as follows: if ${\bf i}$ and ${\bf j}$ are two multi-indices of length $n$, then
+\[
+  {\bf i} \leq {\bf j} \iff i_k \leq j_k, \quad 1 \leq  k \leq n
+\]
+Finally, we generalize the product of binomial coefficients as:
+\[
+  \binom{\boldi}{\boldj} := \prod_{k=1}^n \binom{i_k}{j_k}
+\]
+%
+Given two multi-indices $\boldi$ and $\boldd$ of size $n$, where $\boldi \leq \boldd$, the Bernstein basis polynomial of degree $\boldd$ and index $\boldi$ is defined as:
+\begin{equation}
+\label{eq:bernstein_basis_poly}
+\mathcal{B}_{(\boldi,\boldd)}(\boldx) = \beta_{i_1,d_1}(x_1) \beta_{i_2,d_2}(x_2)\ldots \beta_{i_n,d_n}(x_n).
+\end{equation}
+%
+where for $i,d,x \in \reals$:
+%
+\begin{equation}
+\beta_{i,d}(x) = \binom{d}{i}x^{i}(1-x)^{d - i}
+\end{equation}
+%
+Let $p: \reals^n \rightarrow \reals$ of be a real polynomial of degree at most $\boldd$. We can express $p$ as a linear combination of monomials of degree at most $\boldd$:
+\[
+p(\boldx) = \sum_{\boldi \leq \boldd} a_\boldi \cdot \boldx^\boldi
+\]
+where $\boldx^\boldi$ represents the monomial $x_1^{i_1}x_2^{i_2}\cdots x_n^{i_n}$.
+Every such real polynomial $p$ can be represented as linear combination of Bernstein basis polynomials of of degree $\boldd$:
+%
+\begin{equation}
+  p(\boldx) = \sum_{\boldi \leq \boldd} b_{\boldi} \cdot \mathcal{B}_{(\boldi,\boldd)}(\boldx)
+\end{equation}
+where $b_\boldi$ denotes the \emph{$\boldi^{th}$ Bernstein Coefficient}:
 
-Given two multi-indices $i$ and $d$ of size $n$, where $i \leq d$, the Bernstein polynomial of degree $d$ and index $i$ is
+\begin{equation}
+b_{\boldi} = \sum_{\boldj \leq \boldi} \frac{\binom{\boldi}{\boldj}}{\binom{\boldd}{\boldj}}\cdot a_{\boldj}
+\end{equation}
+%
+\noindent In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
+
+The primary advantage of the Bernstein representation of a polynomial $p(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $p(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis.
+Specifically, the upper and lower bounds of $p(x_1,\ldots,x_n)$ over $[0,1]^n$ are bounded by the Bernstein coefficients:
 $$
-\mathcal{B}_{i,d} = \beta_{i_1,d_1}(x_1) \beta_{i_2,d_2}(x_2)\ldots \beta_{i_n,d_n}(x_n).
-$$
-%
-where $\beta_{i_m,d_m}(x_m) = \binom{d_m}{i_m}x_{m}^{i_m}(1-x_m)^{d_m - i_m}$.
-%
-Any polynomial function can be expressed in the Bernstein basis.
-%
-The primary advantage of the Bernstein representation of a polynomial $h(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $h(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis.
-
-In other words, given a polynomial $h(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $h(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $h(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
-%
-The upper and lower bounds of $h(x_1,\ldots,x_n)$ over $[0,1]^n$ are bounded by the Bernstein coefficients:
-$$
-min \{b_1, \ldots, b_m\} ~~\leq~~ inf_{x \in [0,1]^n} h(x) ~~~~~\leq~~~~ sup_{x \in [0,1]^{n}} h(x) ~~\leq~~ max \{b, \ldots, b_m\}.
+\min \{b_1, \ldots, b_m\} ~~\leq~~ \inf_{x \in [0,1]^n} p(x) ~~~~~\leq~~~~ \sup_{x \in [0,1]^{n}} p(x) ~~\leq~~ \max \{b, \ldots, b_m\}.
 $$
@@
 %
-Therefore, upper bounds on the suprenum of a function $h$ over $P$ is equivalent to upper bound of $h \circ T_{p}$ over $[0,1]^{n}$.
+Therefore, upper bounds on the suprenum of a polynomial function $p$ over $P$ is equivalent to upper bound of $p \circ T_{p}$ over $[0,1]^{n}$.
 %
 A similar argument follows for the lower bound on the infimum.
+%
+The crux of the reachability algorithm involves exploiting this property of Bernstein polynomials to approximate the solution of certain non-linear optimization problem invovilng polynomial predicates over the unitbox, $[0,1]^n$.
+%
+We will cover this algorithm in the upcoming section.
+%
+For a more rigorous exposition on Bernstein polynomials and their properties, refer to \cite{garloff2003bernstein}.
 
-\subsection{The Static Algorithm}
-\label{sec:static}
+\input{staticalg}

### 2022-07-02 thesis/staticalg.tex
Static Algorithm section started

--- 2021-08-15 Nonlinear/formats2021/theory.tex
@@
-\vspace{-1.2em}
-\section{Reachability Algorithm}
-\label{sec:theory}
+\subsection{The Static Algorithm}
+\label{sec:static}
 
-In this work, we develop parallelotope reachability algorithms that are \textbf{automatic} with \textbf{dynamic} parallelotopes.
+We will end with an outline of the static algorithm first investigated in works \cite{dang2012reachability, dreossi2016parallelotope}.
+As mentioned in the previous section, the building block of the reachability algorithm relies on approximate solutions to a non-linear optimization problem over the unitbox domain.
+Consider a nonlinear function $h : \reals^n \rightarrow \reals$.
+The most general form of this optimization problem can be expressed as:
 %
-The state of the art, in contrast, is \textbf{manual}, where the user specifies a set of parallelotope directions at the beginning.
-%
-The parallelotopes are also \textbf{static} and do not change during the course of the computation.
-%
-In this section, we detail the modifications to the algorithm and present their correctness arguments.
+\begin{eqnarray}
+  \max ~ h(x) \label{eq:maxsup}\\
+  s.t. ~~ x \in [0,1]^{n}.\nonumber
+\end{eqnarray}
 
-\subsection{Manual Static Algorithm}
-\label{sec:manualstatic}
+In the static algorithm, the user manually specifies the number of parallelotopes and a set of static directions for each parallelotope.
+In other words, the user must specify the template matrix $\Lambda$ and its corresponding offset vector $c$ for each parallelotope $P = \langle \Lambda, c\rangle$ contained in the bundle \emph{before} the computation begins.
 
-We first present the original algorithm\cite{dang2012reachability} where the user manually specifies the number of parallelotopes and a set of static directions for each parallelotope.
+We now proceed to formally describe the static algorithm.
+First, the parallelotope bundle $Q$ is specified as a collection of $m$ template directions $\Lambda^{Q} \in \reals^{m \times n} \; (m > n)$ and the set of constraints that define each of the member parallelotope.
 
-Recall the system is $n$-dimensional with dynamics function $f: \reals^{n} \rightarrow \reals^{n}$.
-%
-The parallelotope bundle $Q$ is specified as a collection of $m$ template directions $\T^{Q} \in \reals^{m \times n} (m > n)$ and the set of constraints that define each of the member parallelotopes.
-%
+Another input to the algorithm is the initial set, given as a parallelotope $P_0$. When the initial set is a box, $P_0$ consists has axis-aligned template directions.
 
-% consisting of $m$ parallelotopes , the user should manually provide $n$ generator directions for each parallelotope.
-%
-% We write the collective set of generator directions as $\mathcal{G}$, which is a set of sets, each of which contains $n$ direction vectors.
-%
-Another input to the algorithm is the initial set, given as a parallelotope $P_0$.
-%
-When the initial set is a box, $P_0$ consists has axis-aligned template directions.
-%
 The output of the algorithm is, for each step $k$, the set $\overline\Theta_k$, which is an overapproximation of the reachable set at step $k$, $\Theta_k \subseteq \overline\Theta_k$.
@@
 % and half-space representation as $\tup{\T, c_{l}, c_{u}$.
-Given a template direction $\T_i$, the maximum value of $\T_{i} f(x)$ for all $x \in Q$ is less than or equal to the maximum value of $\T_{i} f(x)$ for all $x \in P$.
+Given a template direction $\Lambda_i$, the maximum value of $\Lambda_{i} f(x)$ for all $x \in Q$ is less than or equal to the maximum value of $\Lambda_{i} f(x)$ for all $x \in P$.
 %
@@
 %
-To compute the upper and lower bounds of each template direction $\T_{i} f(x)$, for all $x \in P$, we perform the following optimization.
+To compute the upper and lower bounds of each template direction $\Lambda_{i} f(x)$, for all $x \in P$, we perform the following optimization.
 %
 \begin{eqnarray}
-  \max ~ \T_i^{P} \cdot f(x) \label{eq:maxf}\\
+  \max ~ \Lambda_i^{P} \cdot f(x) \label{eq:maxf}\\
   s.t. ~~ x \in P.\nonumber
@@
 %
-Let  $\tup{a, G}$ be the generator representation of $P$.
+Let  $\langle v, g_1, \cdots, g_n \rangle$ be the generator representation of $P$.
 %
@@
 \begin{eqnarray}
-  \max ~ \T_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i) \label{eq:maxalpha}\\
+  \max ~ \Lambda_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i) \label{eq:maxalpha}\\
   s.t. ~~ \overline\alpha \in [0,1]^{n}.\nonumber
@@
 
-Equation~\ref{eq:maxalpha} is a form of $\mathsf{optBox(\T_{i} \cdot f)}$ over $[0,1]^n$.
+Equation~\ref{eq:maxalpha} is a form of $\mathsf{optBox(\Lambda_{i} \cdot f)}$ over $[0,1]^n$.
 %
@@
 %
-Similarly, we compute the lowerbound of $\T_{i}f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \T_{i}f(x)$.
+Similarly, we compute the lowerbound of $\Lambda_{i}f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \Lambda_{i}f(x)$.
 %
@@
 Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving several constrained nonlinear optimization problems.
-
-\begin{lemma}
-\label{lem:correctness}
-The parallelotope bundle $Q'$ computed using \tbundle (Algorithm~\ref{alg:old}) is a sound overapproximation of the image of bundle $Q$ w.r.t the dynamics $x^{+} = f(x)$.
-\end{lemma}
-\begin{proof}
-     Suppose that parallelotope $P$ is part of the bundle $Q$ and is provided in generator representation as $\tup{a, G}$.
-%
-By definition, we know that $Q \subseteq P$.
-%
-in generator representation and as $\tup{\T, c_{l}, c_{u}}$ in half-space representation.
-%
-Consider an $x \in Q$ and $x' = f(x)$.
-%
-Given a template direction $\T_{i}$, we know that $\T_{i} x' \leq \mathsf{optBox(\T_{i} \cdot f)}$ (Equation~\ref{eq:maxalpha}) and $-1\times optBox(-1\times \T_{i} \cdot f) \leq \T_{i} x'$.
-%
-Therefore, for all $x \in Q$, $c_{l}'[i] \leq \T_{i} \cdot f(x) \leq c_{u}'[i]$.
-%
-Therefore, for all $x \in Q$, $f(x) \in Q'$.
-\end{proof}
-
-% \textcolor{red}{\textbf{SRIDHAR: add this (as well as argument for correctness)}}
-
-\input{algorithm_old}
-\vspace{-1.5em}
-\subsection{Automatic Dynamic Algorithm}
-
-The proposed automatic dynamic algorithm does not require the user to provide the set of template directions $\T$; instead it generates these templates directions automatically at each step.
-%
-We use two techniques to generate such template directions, first: computing local linear approximations of the dynamics and second, performing principal component analysis (PCA) over sample trajectories.
-%
-To do this, we first sample a set of points in the parallelotope bundle called \emph{support points} and propagate them to the next step using the dynamics $f$.
-%
-Support points are a subset of the vertices of the parallelotope that either maximize or minimize the template directions.
-
-Intuitively, linear approximations can provide good approximations when the dynamics function is a time-discretization of a continuous system.
-%
-In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear transformation.
-%
-We use the support points as a data-driven approach to find the best-fit linear function to use.
-%
-If the dynamics of a system is linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \T x \leq c_{u}$, is the set $c_{l} \leq \T \cdot A^{-1} x \leq c_{u}$.
-%
-Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the nonlinear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
-%
-The second technique for generating template directions performs principal component analysis over the images of the support points.
-%
-Using PCA is a reasonable choice as it produces orthonormal directions that can construct a rotated box for bounding the points.
-%
-Observe that in general, the dynamics is nonlinear and therefore, the reachable set could be non-convex.
-%
-On the other hand, a parallelotope bundle is always a convex set.
-%
-To mitigate this discrepancy, we can improve accuracy of this representation by considering more template directions.
-%
-For this purpose, we use a notion of \emph{template lifespan}, where we use the linear approximation and/or PCA template directions not only from the current step, but also from the previous $L$ steps.
-%
-We will demonstrate the effectiveness and tune each of the options (PCA / linear approximation as well as lifespan option) in our evaluation.
-
-The new approach is given in Algorithm~\ref{alg:new}.
-%
-In this algorithm, instead of fixing the set of templates, we compute one set of templates (that is, a collection of $n$ template directions), using linear approximation of the dynamics and PCA.
-%
-The algorithm makes use of helper function \texttt{hstack}, which converts column vectors into a matrix (as shown in Equation~\ref{eq:ptopeexample} provided in Example~\ref{ex:ptope}).
-%
-The notation $M_{*,i}$ is used to refer to the $i^{th}$ column of matrix $M$.
-%
-The \texttt{Maximize} function takes in a parallelotope bundle $Q$ and direction vector $v$ (one of the template directions), and returns the point $p \in Q$ that maximizes the dot product $v \cdot p$ (for computing support points).
-%
-This can be computed efficiently using linear programming.
-%
-The \texttt{ApproxLinearTrans} function computes the best approximation of a linear transformation given a list of points before and after the one-step transformation $f$.
-%
-More specifically, given a matrix $X$ of points before applying the transformation $f$, a matrix of points after the transformation $X'$, we perform a least-squares fit for the linear transition matrix $A$ such that $X' \approx AX$.
-%
-This can be computed by $A = X' X^\dagger$, where $X^\dagger$ is the Moore-Penrose pseudoinverse of $X$.
-%
-The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points.
-%
-Finally, \texttt{TransformBundle} is the same as in Algorithm~\ref{alg:old}.
-
-\input{algorithm_new}
-
-Algorithm~\ref{alg:new} computes the dynamic templates for each time step $k$.
-%
-Line~\ref{ln:linearapprox} computes the linear approximation of the nonlinear dynamics and this linear approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
-%
-The PCA directions of the images of support points is computed in line~\ref{ln:pca}.
-%
-For the time step $k$, the linear and PCA templates are given as $\T_{k}^{lin}$ and $\T_{k}^{pca}$, respectively.
-%
-To improve the accuracy of the reachable set, we compute the overapproximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the lifespan $L$.
-%
-%This is because, the nonlinear systems we deal with are often %not chaotic in nature.
-%
-%Therefore, the templates at time step $k-1$ are very similar %to that of template directions at time step $k$.
-%

### 2022-07-03 thesis/formats.tex
More progress on static algorithm. Starrting on dynamic part

\chapter{Dynamic Paralleotope Bundles}

\subsection{Local Linear Approximations}
\label{sec:lin_app}

\subsection{Principal Component Analysis}
\label{sec:pca}

\subsection{The Dynamic Algorithm}
\label{sec:dyna_algo}

### 2022-07-03 thesis/prelim.tex
More progress on static algorithm. Starrting on dynamic part

@@
 %
-Combining these two observation yields the \emph{half-space representation} of parallelotope $P$.
+Additionally, the definition of the paralleotope above requires that for each of $n$ ``postive" directions, there must exist a corresponding ``negative" direction.
+This is encoded into the template matrix $\Lambda$ by the condition $\Lambda_{i+n} = -\Lambda_{i}$.
+However, by the observation made above, we only need to keep the positive directions and divide our offset vector into equal components with the top half encoding the offsets for the positive directions and the bottom half encoding the offsets for the negative directions.
+%
+Combining these remarks yields the \emph{half-space representation} of parallelotope $P$.
 %
 \begin{definition}
-The half-space representation of parallelotope $P$ is tuple  $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ and $\Lambda \in \mathbb{R}^{2n \times n}$ such that if $\overline{\Lambda}$ represents the submatrix of $\Lambda$ consisting of the first $n$ rows of $\Lambda$, then
+\label{def:halfspace_def}
+The half-space representation of parallelotope $P$ is tuple  $\langle \Lambda, c_l, c_u \rangle$ where $\Lambda \in \mathbb{R}^{n \times n}$ and $c_l, c_u \in \reals^n$
+such that
 \begin{equation}
 \label{eq:halfspaceconst}
-P =\{x ~|~ -c_l \leq \overline{\Lambda} x \leq c_u\}
+P =\{x ~|~ c_l \leq \Lambda x \leq c_u\}
 \end{equation}
 \end{definition}
-%
+
+\begin{example}
+\label{ex:simple_ptope}
+Consider the 2D plane, namely $\reals^2$. We can construct a couple of simple examples of parallelotopes. First, if we define our parallelotope's template direction matrix to be
+%come back to this
+\end{example}
+
 \noindent Alternatively, a parallelotope can also be represented in a \emph{generator representation}
@@
 Let us denote this affine transformation associated to $P$ as $T_P$.
-
-\begin{example}
-\label{ex:simple_ptope}
-Consider the 2D plane, namely $\reals^2$. We can construct a couple of simple examples of parallelotopes. First, if we define our parallelotope's template direction matrix to be
-%come back to this
-\end{example}
 
@@
 \begin{definition}
-A parallelotope bundle $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$ such that
+A parallelotope bundle $Q$ is a set of parallelotopes $\{P_0, \ldots, P_m\}$ such that
   $$ Q = \bigcap_{i=1}^{m}P_i $$.
 \end{definition}
+
+\begin{remark}
+There is a slight abuse of notation above where we refer to the parallelotope bundle $Q$ as both the set of parallelotopes and the region in $\reals^n$ of the intersection of all the parallelotopes $P_i$.
+To specify the \emph{set of parallelotopes} which consists the bundle, we will write
+$$
+\mathcal{P}(Q) = \{P_0, \ldots, P_m\}
+$$
+\end{remark}
+
 %
@@
 The primary advantage of the Bernstein representation of a polynomial $p(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $p(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis.
-Specifically, the upper and lower bounds of $p(x_1,\ldots,x_n)$ over $[0,1]^n$ are bounded by the Bernstein coefficients:
+Specifically, the upper and lower bounds of $p(x_1,\ldots,x_n)$ over $[0,1]^n$ are bounded by the Bernstein coefficients.
+We state this as a property without proof.
+%
+\begin{property} (Enclosure Property)
+  \label{prop:bern_enclosure}
+  Let $p: \reals^n \rightarrow \reals$ be a real mutlivarite polynomial of degree $\boldd$, and let $p(\boldx) = \sum_{\boldi \leq \boldd} b_\boldi \cdot \mathcal{B}_{(\boldi,\boldd)}(\boldx)$ be the Bernstein expansion of $p$, then
   $$
-\min \{b_1, \ldots, b_m\} ~~\leq~~ \inf_{x \in [0,1]^n} p(x) ~~~~~\leq~~~~ \sup_{x \in [0,1]^{n}} p(x) ~~\leq~~ \max \{b, \ldots, b_m\}.
+  \min_{\boldi \leq \boldd} \{b_\boldi\} ~~\leq~~ \inf_{x \in [0,1]^n} p(x) ~~~~~\leq~~~~ \sup_{x \in [0,1]^{n}} p(x) ~~\leq~~ \max_{\boldi \leq \boldd} \{b_\boldi\}
   $$
+\end{property}
 
@@
 %
-For a more rigorous exposition on Bernstein polynomials and their properties, refer to \cite{garloff2003bernstein}.
+For a more rigorous exposition on Bernstein polynomials and Property~\ref{prop:bern_enclosure}, refer to \cite{garloff2003bernstein}.
 

### 2022-07-03 thesis/staticalg.tex
More progress on static algorithm. Starrting on dynamic part

@@
 We now proceed to formally describe the static algorithm.
-First, the parallelotope bundle $Q$ is specified as a collection of $m$ template directions $\Lambda^{Q} \in \reals^{m \times n} \; (m > n)$ and the set of constraints that define each of the member parallelotope.
+%
+First, a small remark on the template matrix of the parallelotopes $P_i$ contained in some bundle $Q$. It is possible that some of the parallelotopes share the same template matrix directions.
+%
+ In other words, for $P_i = \langle \Lambda^{P_i}, c^{P_i} \rangle, P_j = \langle \Lambda^{P_j}, c^{P_j} \rangle$ such that $P_i, P_j \in \ptopeset{Q}$, there could exist some $k$ such that $\Lambda^{P_i}_k = \Lambda^{P_j}_k$ as row vectors.
+%
+Thus, a more compact method of encoding the bundle is by taking the \emph{distinct} template directions as rows of a new template matrix $\Lambda^Q$ along with its corresponding offset vector $c^Q$.
+%
+To distinguish between the distinct parallelotopes contained in the bundle, we add a new matrix called $\T^Q \in \mathbb{N}^{p \times n}$ such that $\T^Q_i$ is a vector of row indices of $\Lambda^Q$ which specify the template directions defining parallelotope $P_i \in \ptopeset{Q}$.
+%
+\begin{remark}
+If none of the paralleotopes $P_i \in \ptopeset{Q}$ share common template directions, then $\Lambda^Q$ will simply be the template direction matrices $\{\Lambda^P\}_{P \in \ptopeset{Q}}$ concatented along their rows.
+This will generally be the matrix generated by the dynamic algorithm we will outline in a future section.
+\end{remark}
+%
+\begin{example}
 
-Another input to the algorithm is the initial set, given as a parallelotope $P_0$. When the initial set is a box, $P_0$ consists has axis-aligned template directions.
+\end{example}
+
+Another input to the algorithm is the initial set, given as a parallelotope $P_0$. When the initial set is a box, $P_0$ will be defined by the axis-aligned template directions.
 
 The output of the algorithm is, for each step $k$, the set $\overline\Theta_k$, which is an overapproximation of the reachable set at step $k$, $\Theta_k \subseteq \overline\Theta_k$.
+The total overapproximation of the reachable set for a finite number of steps $n$ will be $\Theta = \cup_{k=1}^n \Theta_k$.
+%
+The high-level pseudo-code is written in Algorithm~\ref{alg:old}.
 
-The high-level pseudo-code is written in Algorithm~\ref{alg:old}.
+The algorithm simply calls $\tbundle$ for each step, producing a new parallelotope bundle computed from the previous step's bundle.
 %
-The algorithm simply calls \tbundle for each step, producing a new parallelotope bundle computed from the previous step's bundle.
+To compute the image of $Q$, the algorithm computes the upper and lower bounds of $f(x)$ with respect to each template direction $\Lambda_i^Q$.
 %
-To compute the image of $Q$, the algorithm computes the upper and lower bounds of $f(x)$ with respect to each template direction.
-%
-Since computing the maximum value of $f(x)$ along each template direction on $Q$ is computationally difficult, the algorithm instead computes the maximum value over each of the constituent parallelotopes and uses the minimum of all these maximum values.
-%
-The \tbundle operation works as follows.
+Since computing the maximum value of $f(x)$ along each template direction on the $Q$ in one-shot is computationally difficult, the algorithm instead computes the maximum value over each of the constituent parallelotopes and uses the minimum of all these maximum values.
+
+The $\tbundle$  operation works as follows.
 %
@@
 %
-From the definition, it follows that $Q \subseteq P$.
+% and half-space representation as $\tup{\T, c_{l}, c_{u}$.
+Given a template direction $\Lambda_i^Q$, the maximum value of $\Lambda_{i}^Q f(x)$ for all $x \in Q$ is less than or equal to the maximum value of $\Lambda_{i}^P \cdot f(x)$ for all $x \in P$ such that $\Lambda_i^Q$ is a row in $\Lambda^P$.
 %
-% and half-space representation as $\tup{\T, c_{l}, c_{u}$.
-Given a template direction $\Lambda_i$, the maximum value of $\Lambda_{i} f(x)$ for all $x \in Q$ is less than or equal to the maximum value of $\Lambda_{i} f(x)$ for all $x \in P$.
+Similar argument holds for the minimum value of $\Lambda_{i}^Q \cdot f(x)$ for all $x \in Q$.
 %
-Similar argument holds for the minimum value of $T_{i} f(x)$ for all $x \in Q$.
+Observe that these inequalities hold by virtue of the fact that $Q \subseteq P$ by definition.
+%
+To describe this more formally: if $\lambda^Q_i = \{ P \in \ptopeset{Q} ~|~ \Lambda_i^Q = \Lambda_k^P \text{ for some } k \}$, then
+%
+\begin{align}
+    & \max_{x \in Q}{\Lambda_{i}^Q \cdot f(x)} \leq \min_{P \in \lambda^Q_i} {\max_{x \in P}{\Lambda_{i}^P \cdot f(x)}}  \label{eq:max_over_ptopes}
+ \\
+    & \min_{P \in \lambda^Q_i} {\min_{x \in P}{\Lambda_{i}^P \cdot f(x)}} \leq \min_{x \in Q}{\Lambda_{i}^Q \cdot f(x)} \label{eq:min_over_ptopes}
+\end{align}
+
 %
@@
 \end{eqnarray}
+%
+\noindent Note that $\Lambda_i^{P} \cdot f(x)$ is a dot product between the row vector $\Lambda_i^{P}$ and the component-wise dynamics of $f(x)$.
+This is similar to the method of computing support functions over convex sets \cite{boyd2004convex}.
 
@@
 \begin{eqnarray}
-  \max ~ \Lambda_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i) \label{eq:maxalpha}\\
+  \max ~ \Lambda_i^P \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i) \label{eq:maxalpha}\\
   s.t. ~~ \overline\alpha \in [0,1]^{n}.\nonumber
@@
 %
-One can compute an upper-bound to the constrained nonlinear optimization by invoking one of the Bernstein polynomial or interval-arithmetic-based methods.
+One can compute an upper-bound to this nonlinear optimization by computing the Bernstein coefficients of $\Lambda_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i)$ and taking the maximum and minimum coefficients as shown in Property~\ref{prop:bern_enclosure}.
 %
-Similarly, we compute the lowerbound of $\Lambda_{i}f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \Lambda_{i}f(x)$.
+Similarly, we compute the lower-bound of $\Lambda_{i}\cdot f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \Lambda_{i}\cdot f(x)$.
 %
 
-We iterate this process (i.e., computing the upper and lower bound of $T_{i}f(x)$) for each parallelotope in the bundle $Q$.
+We iterate this process (i.e., computing the upper and lower bound of $\Lambda^Q_{i}\cdot f(x)$) for each parallelotope in the bundle $Q$ according to Equation \ref{eq:max_over_ptopes} and Equation \ref{eq:min_over_ptopes}).
 %
@@
 %
-Therefore, the tightest upper bound on $T_{i}f(x)$ over $Q$ is the least of the upper bounds computed from each of the parallelotopes.
+Therefore, the tightest upper bound on $\Lambda^Q_{i}\cdot f(x)$ over $Q$ is the least of the upper bounds computed from each of the parallelotopes.
 %
-A similar argument holds for lower bounds of $T_{i}f(x)$ over $Q$.
+A similar argument holds for lower bounds of $\Lambda^Q_{i}\cdot f(x)$ over $Q$.
 %
-Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving several constrained nonlinear optimization problems.
+Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of nonlinear optimization problems of the form presented in Equation~\ref{eq:maxf}.
+
+Finally, once the loop on step two of Algorithm~\ref{alg:old} halts at step $S$, the outputted reachable set will be the computed over-approximations $\overline\Theta_1, \cdots, \overline\Theta_S$.
+As step four within the loop implies, this is simply the image bundles $Q'$ returned by our $\tbundle$ procedure.
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+%\input{static_alg_fig}
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\begin{example}
+  We return to the SIR model briefly treated in Section \ref{sec:definitions}.
+Figure~\ref{fig:kaa_sir} shows the reachable set computed with the static algorithm and plotted using the following parameters:
+  \begin{itemize}
+    \item The parameters of the model are set to $\beta = 0.34$ and $\gamma = 0.05$. The discretization step is set to $\Delta = 0.1$.
+    \item The parallelotope only has one static parallelotope, namely the initial box. This shows that our template matrix for $P$ is
+        \begin{equation*}
+          \Lambda^Q = \begin{bmatrix}
+                      1 & 0 & 0 \\
+                      0 & 1 & 0 \\
+                      0 & 0 & 1
+                      \end{bmatrix}
+                      \quad
+          \T^Q = \begin{bmatrix}
+                      0 & 1 & 2 \\
+                      \end{bmatrix}
+        \end{equation*}
+        %
+        \begin{equation*}
+            c_l^P = \begin{bmatrix} -0.79 & -0.19 & 0 \end{bmatrix}^T \quad c_u^P = \begin{bmatrix} 0.8 & 0.2 & 0 \end{bmatrix}^T
+        \end{equation*}
+    \item We set the number of time steps $S = 300$.
+  \end{itemize}
+
+  There a few points worth noting here.
+First, by the discussion leading to Definition~\ref{def:halfspace_def}, the initial set would be the box $[0.79,0.8] \times [0.19, 0.2] \times 0$.
+This can be interpreted as initializing the model such that $79-80\%$ of the population is susceptible (not yet infected) with $19-20\%$ of the population is infected.
+As the simulation is beginning, no percentage of the population has recovered from the disease.
+Hence, the third parameter $r$ is set to zero.
+
+  Second, since we only have the axis-aligned parallelotope in our initial bundle, the matrix $\T^Q$ will consist of only one row indicing the axis-aligned directions expressed as distinct rows in $\Lambda^Q$.
+
+  \begin{figure}[ht]
+  \label{fig:kaa_sir}
+    \hspace*{-2.3cm}
+    \includegraphics[width=1.3\textwidth]{figures/SIRProj.png}
+    \caption{Projection of Reachable Set of SIR propagated 300 steps in time.}
+  \end{figure}
+\end{example}
+
+\subsection{Optimization Modes}

### 2022-07-04 thesis/comp_temp_fig.tex
Started dynamic algo and results section

--- 2021-05-25 Nonlinear/formats2021/evaluation.tex
@@
-\vspace{-1em}
-\section{Evaluation}
-\label{sec:eval}
-%\vspace{-0.5em}
- We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} \cite{kim2020kaa}.
-Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems.
-The optimization procedure for finding the direction offets is performed through the \emph{Kodiak} library.
-Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module.
-To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize scipy.spatial library's Convex hull routine.
- These routines are esstentially Python wrappers around their corresponding \emph{QHull} procedures.
-For higher-dimensional systems, we employ the volume of the tightest enveloping box around the intersection of the parallelotopes.
- The dimensions of this box are easily calculated through linear programming. The total volume estimate of the reachable set will be the sum of all the computed bundles' volume estimates.
-
- %For two-dimensional models, note that there are only five non-trivial templates.
-The trivial template here refers to the initial box which is defined exactly the axis-aligned directions.
-Refer to the non-axis aligned templates.
-At each every time, pca.
-lin app directions to
-\vspace{-1em}
-\subsubsection{Model Dynamics}
-For benchmarking, we select six non-linear models with polynomial dynamics.
-Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability analysis with {\bf static} parallelotope bundles.
-In these cases, we directly compare the performance of our dynamic strategies with the static parallelotopes defined in Sapo.
-To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
-Here, {\bf diagonal directions} are defined to be the vectors created by adding and subtracting distinct pairs of all unit axis-aligned vectors from each other.
-By {\bf diagonal parallelotopes}, we refer to parallelotopes which are defined only by axis-aligned and diagonal directions.
-Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles which solely consists of diagonal parallelotopes.
-Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
-Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
-For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box.
-We refer to these parallelotopes as {\bf non-axis-aligned parallelotopes}.
-
-%Maybe mention something about role of diagonal templates in program verfication here.
-\begin{example}
-In two dimensions, $\mathbb{R}^2$, we have the two unit axis-aligned directions, $[1,0]^T, [0,1]^T$. The diagonal directions will then
-
-\[ [1,1]^T, \; [1,-1]^T\]
-Consequently, the diagonal parallelotopes will be precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
-% Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions $[1,0,0]^T, [0,1,0]^T,[0,0,1]^T$. Calculating the diagonal directions yields the following directions:
-% \[ [1,1,0]^T, \; [1,-1,0]^T, \; [1,0,1]^T,\;  [1,0,-1]^T,\;  [0,1,1]^T, \; [0,1,-1]^T \]
-% Thus, there are ${9 \choose 3} = 84$ diagonal parallelotopes.
-%
-% Note that for both examples, we do not consider the directions' negative counterparts since we define parallelotopes use both the positive and negative counterparts of a its chosen directions.
-%
-% For two dimensions, the non-axis-aligned parallelotopes would be all parallelotopes excluding the one defined by directions $[0,1]^T, [1,0]^T$.
-In particular, there are 5 non-axis-aligned diagonal parallelotopes.
-\end{example}
-Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional model COVID supermodel is explained in the subsequent subsection below.
-
-\vspace{-1em}
-\begin{table}[h!]
-%\hspace{-5em}
-  \centering
-\begin{tabular}{|p{1.5cm}|c|p{1.5cm}|c|c|p{5cm}|}
-\hline
-Model & Dimension & Parameters & \# steps & $\Delta$ & \hspace{1.5cm}Initial Box \\
-\hline
-Vanderpol & 2 & - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
-\hline
-Jet Engine& 2 & - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
-\hline
-Neuron \cite{fitzhugh1961impulses}& 2 & - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
-\hline
-SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 & $s \in [0.79,0.8], i \in [0.19,0.2], r = 0$ \\
-\hline
-Coupled \newline Vanderpol & 4 & - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
-\hline
-COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & Stated Below\\
-\hline
-\end{tabular}
-\caption{Benchmark models and relevant information}
-\label{tab:modeldyns}
-\end{table}
-
-\vspace{-2em}
-\noindent \textbf{COVID Supermodel:}
-%
-We further benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
-This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients.
-These patients have the ability to infect susceptible members with a fixed probability.
-The following dynamics account for this new group and its interactions with the traditional groups found in the SIR model.
-The continuous dynamics given in \cite{indiansuper2020supermodel} are discretized through the Euler method.
-
-\begin{align}
-  \begin{split}
-   S_A & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
-   S_I & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
-   A & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
-   I & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
-   R_A & = R_A + (\gamma A)\cdot \Delta \\
-   R_I & = R_I + (\gamma I)\cdot \Delta \\
-   D & = D + (\eta I)\cdot \Delta
-  \end{split}
-\end{align}
-where the variables denote the fraction of a population of individuals designated as \emph{Susceptible Transition to Asymptomatic $(S_A)$}, \emph{Susceptible Transition to Infected $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
-We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the average probablity of infection, $\gamma$ is the average removal rate, and $\eta$ is the average mortality rate.
-The parameters are set based on figures shown in \cite{ansumali2020modelling} for the USA.
-The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
-
-\subsubsection{Accuracy of Dynamic Strategies}
-The results of testing our dynamic strategies against static ones are summarized in Table~\ref{tab:voltable}.
-For models previously defined in Sapo, we set the static parallelotopes to be exactly those found in Sapo.
-If a model is not implemented in Sapo, we simply use the static parallelotopes defined in a model of equal dimension.
-To address the unavailability of a four-dimensional model implemented in Sapo, we sampled random subsets of five static non-axis-aligned parallelotopes and chose the flowpipe with smallest volume.
-%
-% As justifiction for this scheme, we wish to remark the time required in testing all possible subsets of diagonal parallelotopes.
-%
-A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is impossible.
-
-% are $2{n \choose 2}$ diagonal directions for an $n$-dimensional system, giving a total of $n + 2 {n \choose 2}$ axis-aligned and diagonal directions.
-Thus, there are ${n + 2 {n \choose 2} \choose n} = {n^2 \choose n}$ possible diagonal parallelotopes for an $n$-dimensional system.
-This explosive growth in search space size compelled us to proceed by sampling the search space instead.
-
-From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks.
-In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the PCA/LinApp ratio imparts differing effects on the reachable set.
-Observe that using parallelotopes defined by Linear Approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.
-
 \begin{figure}[h!]
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}
 \includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5Lin_.png}
@@
 \end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}
 \includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_1PCA5Lin_.png}
@@
 
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}
 \includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5PCA_.png}
@@
 \end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}
 \includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_5PCA1Lin_.png}
@@
 
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}
 \includegraphics[width=\textwidth]{figures/PhasePlots/VDP_Sapo_.png}
@@
 \end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}
 \includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_Sapo_.png}
@@
 \end{figure}
-
-\vspace{-1em}
-\subsubsection{Performance under Increasing Initial Sets}
-A key advantage of our dynamic strategies is the improved ability to control the wrapping error naturally arising from larger initial sets.
-Figure \ref{fig:InitVolReachComp} presents charts showcasing the effect of increasing initial sets on the total flowpipe volume.
-We vary the initial box dimensions to gradually increase the box's volume.
-We then plot the total flowpipe volume after running the benchmark.
-The same initial boxes are also fed as input into computation using only static parallelotopes borrowed from Sapo.
-The number of parallelotopes defined by PCA and Linear Approximation directions were chosen based on best performance as seen in Table \ref{tab:voltable}.
-We remark that our dynamic strategies perform better than static ones in controlling the total flowpipe volume as the initial set becomes larger.
-On the other hand, the performance of static parallelotopes tends to degrade rapidly as we increase the volume of the initial box.
-
-\subsubsection{Performance against Random Static Templates}
-We additionally benchmark our dynamic strategies against static random parallelotope bundles.
-We sample such parallelotopes in $n$ dimensions by first sampling a set of $n$ directions uniformly on the surface of the unit $(n-1)$-sphere, then defining our parallelotope using these sampled directions.
-We sample twenty of these parallelotopes for each trial and average the total flowpipe volumes.
-As shown in Figure \ref{fig:RanStaticStratComp}, our best-performing dynamic strategies consistently outperform static random strategies for all tested benchmarks.
-
-\vspace{-1em}
-\section{Conclusions}
-In this paper we investigated two techniques for generating templates dynamically, first using linear approximation of the dynamics, and second using PCA.
-%
-We demonstrated that these techniques improve the accuracy of reachable set by an order of magnitude when compared to static or random template directions.
-%
-We also observed that both these techniques improve the accuracy of the reachable sets for different benchamrks.
-%
-In future, we intend to investigate Koopman linearization techniques for computing the template directions~\cite{bak2021reachability}.
-
-\begin{table}[h!]
-\hspace{1em}
-\begin{subtable}[h]{0.45\textwidth}
-     \centering
-     \begin{tabular}{|c|c|}
-     \hline
-     Strategy & Total  Volume \\
-     \hline\
-     5 LinApp & 0.227911 \\
-     \hline
-     1 PCA, 4 LinApp& 0.225917 \\
-     \hline
-     2 PCA, 3 LinApp & 0.195573 \\
-     \hline
-     {\bf 3 PCA, 2 LinApp} & {\bf 0.188873} \\
-     \hline
-     4 PCA, 1 LinApp & 1.227753\\
-     \hline
-     5 PCA & 1.509897 \\
-     \hline
-     5 Static Diagonal(Sapo) & 2.863307  \\
-     \hline
-    \end{tabular}
-    \caption{Vanderpol}
-    \label{tab:vdpvol}
- \end{subtable}\hspace{1em}
- \begin{subtable}[h]{0.45\textwidth}
-      \centering
-      \begin{tabular}{|c|c|}
-      \hline
-      Strategy & Total  Volume \\
-      \hline
-      5 LinApp & 58199.62 \\
-      \hline
-      1 PCA, 4 LinApp & 31486.16 \\
-      \hline
-      {\bf 2 PCA, 3 LinApp} & {\bf 5204.09}\\
-      \hline
-      3 PCA, 2 LinApp & 6681.76 \\
-      \hline
-      4 PCA, 1 LinApp& 50505.10 \\
-      \hline
-      5 PCA  & 84191.15 \\
-      \hline
-      5 Static Diagonal (Sapo) & 66182.18  \\
-      \hline
-     \end{tabular}
-     \caption{Jet Engine}
-     \label{tab:enginevol}
-  \end{subtable}
-
-  \hspace{1em}
-  \begin{subtable}[h]{0.45\textwidth}
-       \centering
-       \begin{tabular}{|c|c|}
-       \hline
-       Strategy & Total  Volume \\
-       \hline
-       5 LinApp  & 154.078\\
-       \hline
-       1 PCA, 4 LinApp  & 136.089\\
-       \hline
-       2 PCA, 3 LinApp  & 73.420\\
-       \hline
-       {\bf 3 PCA , 2 LinApp } & {\bf 73.126} \\
-       \hline
-       4 PCA, 1 LinApp  & 76.33 \\
-       \hline
-       5 PCA & 83.896 \\
-       \hline
-       5 Static Diagonal  (Sapo) & 202.406  \\
-       \hline
-      \end{tabular}
-      \caption{FitzHugh-Nagumo}
-      \end{subtable} \hspace{1em}
-      \begin{subtable}[h]{0.45\textwidth}
-        \centering
-        \begin{tabular}{|c|c|}
-        \hline
-        Strategy & Total  Volume \\
-        \hline
-        {\bf 2 LinApp } & {\bf 0.001423} \\
-        \hline
-        1 PCA, 1 LinApp & 0.106546\\
-
-        \hline
-        2 PCA  & 0.117347\\
-        \hline
-        2 Static Diagonal (Sapo) & 0.020894\\
-        \hline
-       \end{tabular}
-       \caption{SIR}
-       \label{tab:sirvol}
-    \end{subtable}
-
-    \hspace{1em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         5 LinApp & 5.5171 \\
-         \hline
-         {\bf 1 PCA, 4 LinApp } & {\bf 5.2536} \\
-         \hline
-         2 PCA, 3 LinApp  & 5.6670\\
-         \hline
-         3 PCA, 2 LinApp  & 5.5824\\
-         \hline
-         4 PCA, 1 LinApp  & 312.2108 \\
-         \hline
-         5 PCA  & 388.0513 \\
-         \hline
-         5 Static Diagonal (Best) & 3023.4463  \\
-         \hline
-        \end{tabular}
-        \caption{Coupled Vanderpol}
-        \label{tab:sirvol}
-     \end{subtable}\hspace{1 em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         3 LinApp & $2.95582227 * 10^{-10}$ \\
-         \hline
-         {\bf 1 PCA, 2 LinApp } & {\bf $2.33007583 * 10^{-10}$}\\
-         \hline
-         2 PCA, 1 LinApp &$ 4.02751770 * 10^{-9}$\\
-         \hline
-         3 PCA & $4.02749571 * 10^{-9}$\\
-         \hline
-         3 Static Diagonal (Sapo) & $4.02749571 * 10^{-9}$\\
-         \hline
-        \end{tabular}
-        \caption{COVID}
-        \label{tab:covidvol}
-     \end{subtable}
- \caption{Tables presenting total reachable set volume by strategy.
-The static directions are retrieved and/or inspired from Sapo models of equal dimension for benchmarking.
-The best performing strategy is highlighted in bold.}
-     \label{tab:voltable}
-\end{table}
-\newpage
-\begin{figure}[h!]
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVol.png}
-    \caption{Vanderpol}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVol.png}
-    \caption{Jet Engine}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVol200steps.png}
-    \caption{Neuron}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVol.png}
-    \caption{SIR}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVol.png}
-    \caption{Coupled Vanderpol}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVol.png}
-    \caption{COVID}
-    \end{subfigure}
-
-    \caption{Comparison between the performance of diagonal static parallelotope bundles and that of the best performing dynamic parallelotope bundles as the volume of the initial set grows.}
-    \label{fig:InitVolReachComp}
-\end{figure}
-\newpage
-
-\begin{figure}[h!]
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVolRanStrat.png}
-    \caption{Vanderpol}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVolRanStrat.png}
-    \caption{Jet Engine}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVolRanStrat.png}
-    \caption{Neuron}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVolRanStrat.png}
-    \caption{SIR}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVolRanStrat.png}
-    \caption{Coupled Vanderpol}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVolRanStrat.png}
-    \caption{COVID}
-    \end{subfigure}
-
-    \caption{Comparision between random static strategies and the best performing dynamic strategies as the volume of the initial set grows.
-The total reachable set volumes for random static strategies are averaged over ten trials for each system.}
-    \label{fig:RanStaticStratComp}
-\end{figure}
-\newpage

### 2022-07-04 thesis/formats.tex
Started dynamic algo and results section

@@
 \chapter{Dynamic Paralleotope Bundles}
+
+In this chapter, we cover a method of generating template directions dynamically and automatically.
+By dynamic, we mean that the template directions must be generated adaptively based on sampled trajectories and/or data from the state of the system.
+%
+ By automatic, we mean that the template directions require no consideration from the user to proceed with the reachable set computation.
+This is in contrast to the original static algorithm treated in Section \ref{sec:static} where the user must input his or her own template directions to specify the parallelotopes before the computation starts.
+%
+To briefly outline the structure of this chapter, we first expound on the two techniques we utilize to dynamically generate template directions at each step.
+%
+The first method is based on local linear approximations where the algorithm approximates the dynamics as a linear transformation based on sample trajectories.
+%
+The second method is based on Principal Compoenent Analysis (PCA) where the algorithm runs PCA on the end points of the sample trajectories.
+%
+Finally, we cover the high-level pseudo-code of the dynamic algorithm and explain a set of parameters we feed into the algorithm in order to improve performance and the accuracy of the outputted reachable set.
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 
@@
 \label{sec:lin_app}
+%
+Intuitively speaking, if time step is discretized to be sufficiently small, propagating trajectories according to the nonlinear dynamics $f$ for one step could lead to good lienar approximations of the dynamics within a small region.
+%
+%
+To do this, we first sample a set of points in the parallelotope bundle called \emph{support points} and propagate them to the next step using the dynamics $f$.
+%
+Support points are a subset of the vertices of the parallelotope that either maximize or minimize the template directions over the parallelotope bundle.
+%
+That is the support points are the set of points $x_i$ such that:
+\begin{equation}
+\label{eq:support}
+x_i = \max_{x \in Q} \Lambda^Q_i\cdot x
+\end{equation}
+%
+for all tempalte direcitons of the bundle $\Lambda^Q_i$. These are all found by a straightforward linear program.
+%
+%In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear transformation.
+%
+We use the support points as a data-driven approach to find the best-fit linear function to use.
+%
+If the dynamics of a system is linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \T x \leq c_{u}$, is the set $c_{l} \leq \T \cdot A^{-1} x \leq c_{u}$.
+%
+Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the nonlinear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
+%
+To find the approximate linear transformation, let $x_i$ denote the support points defined in Equation \ref{eq:support}, then we perform the following least-squares procedure: the objective would be to find an linear transformation $A$ such that it minimizes the following objective function:
+\begin{equation}
+  \min_{A} \sum_{x_i} |f(x_i) - Ax_i|^2
+\end{equation}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 
@@
 \label{sec:pca}
+The second technique for generating template directions performs Principal cCmponent Analysis (PCA) over the images of the support points.
+%
+Using PCA is a reasonable choice as it produces orthonormal directions that can construct a rotated box for bounding the points.
+%
 
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \subsection{The Dynamic Algorithm}
 \label{sec:dyna_algo}
+Observe that in general, the dynamics is nonlinear and therefore, the reachable set could be non-convex.
+%
+On the other hand, a parallelotope bundle is always a convex set.
+%
+To mitigate this discrepancy, we can improve accuracy of this representation by considering more template directions.
+%
+For this purpose, we use a notion of \emph{template lifespan}, where we use the linear approximation and/or PCA template directions not only from the current step, but also from the previous $L$ steps.
+%
+We will demonstrate the effectiveness and tune each of the options (PCA / linear approximation as well as lifespan option) in our benchmarks.
+
+The new approach is given in Algorithm~\ref{alg:new}.
+%
+In this algorithm, instead of fixing the set of templates, we compute one set of templates (that is, a collection of $n$ template directions), using linear approximation of the dynamics and PCA.
+%
+The algorithm makes use of helper function \texttt{hstack}, which converts column vectors into a matrix (as shown in Equation~\ref{eq:ptopeexample} provided in Example~\ref{ex:ptope}).
+%
+The notation $M_{*,i}$ is used to refer to the $i^{th}$ column of matrix $M$.
+%
+The \texttt{ApproxLinearTrans} function computes the best approximation of a linear transformation given a list of points before and after the one-step transformation $f$.
+%
+%More specifically, given a matrix $X$ of points before applying the transformation $f$, a matrix of points after the transformation $X'$, we perform a least-squares fit for the linear transition matrix $A$ such that $X' \approx AX$.
+%
+%This can be computed by $A = X' X^\dagger$, where $X^\dagger$ is the Moore-Penrose pseudoinverse of $X$.
+%
+The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points.
+%
+Finally, \texttt{TransformBundle} is the same as in Algorithm~\ref{alg:old}.
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\input{dynamic_alg_fig}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+lgorithm~\ref{alg:new} computes the dynamic templates for each time step $k$.
+%
+Line~\ref{ln:linearapprox} computes the linear approximation of the nonlinear dynamics and this linear approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
+%
+The PCA directions of the images of support points is computed in line~\ref{ln:pca}.
+%
+For the time step $k$, the linear and PCA templates are given as $\T_{k}^{lin}$ and $\T_{k}^{pca}$, respectively.
+%
+To improve the accuracy of the reachable set, we compute the overapproximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the lifespan $L$.
+%

### 2022-07-04 thesis/prelim.tex
Started dynamic algo and results section

@@
 \end{definition}
+%
+\noindent We will see in a future section an example of the reachable set of the discretized SIR model presented in Equation~\ref{eqn:discrete_sir}.
 
-%
-We will see in a future section an example of the reachable set of the discretized SIR model presented in Equation~\ref{eqn:discrete_sir}.
-
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \section{Parallelotope-based Reachability}
@@
 We deem $\Lambda$ as the \emph{template direction matrix} where $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$ called the \emph{$i^{th}$ template direction}.
-The column vector $c$ is called the \emph{offset vector} with $c_i$ denoting the $i^{th}$ element of $c$.
+The column vector $c$ is called the \emph{offset vector} with $c(i)$ denoting the $i^{th}$ element of $c$.
 %
-If we unpack Equation~\ref{eq:halfspace}, we can re-express the inequalities as a conjection of half-space constraints.
-If we define $c_{u} = [c_1, c_2, \cdots, c_n]^T$ and $c_{l} = [c_{n+1}, c_{n+2}, \cdots, c_{2n}]^T$, then Equation~\ref{eq:halfspace} tells us that:
+If we unpack Equation~\ref{eq:halfspace}, we can re-express the inequalities as a conjunction of half-space constraints.
+If we define $c_{u} = [c(1), c(2), \cdots, c(n)]^T$ and $c_{l} = [c(n+1), c(n+2), \cdots, c(2n)]^T$, then Equation~\ref{eq:halfspace} tells us that:
 %
-\begin{align*}
-  & \Lambda_i x \leq (c_u)_i \\
-  -& \Lambda_i x \leq (c_l)_i
-\end{align*}
+\begin{align}
+  & \Lambda_i x \leq c_u(i)  \label{eq:upper_ineq}\\
+  -& \Lambda_i x \leq c_l(i) \label{eq:lower_ineq}
+\end{align}
 %
@@
 However, by the observation made above, we only need to keep the positive directions and divide our offset vector into equal components with the top half encoding the offsets for the positive directions and the bottom half encoding the offsets for the negative directions.
+The bottom half must be multiplied by a negative sign to account for Inequality \ref{eq:lower_ineq}.
 %
@@
 \end{definition}
-
+%
+\noindent In particular, as a bounded intersection of pairs of parallel half-spaces, it is convex.
+%
 \begin{example}
 \label{ex:simple_ptope}
-Consider the 2D plane, namely $\reals^2$. We can construct a couple of simple examples of parallelotopes. First, if we define our parallelotope's template direction matrix to be
-%come back to this
+Consider the 2D plane, namely $\reals^2$.
+We can construct a couple of simple examples of parallelotopes.
+First, if we define our parallelotope's template direction matrix to be the rows of the matrix:
+%
+\begin{equation}
+  \Lambda = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
+\end{equation}
+%
+We see that our template directions will be the vectors $[1,0]^T, [0,1]^T$. Suppose now we set our upper and lower offsets to be:
+%
+\begin{equation}
+  c_l = [1,1]^T, \quad c_u = [2,2]^T
+\end{equation}
+Then by Definition \ref{def:halfspace_def}, the bounded region in space will be the intersection of the following linear constraints:
+%
+\begin{align}
+  1 \leq x \leq 2 \\
+  1 \leq y \leq 2
+\end{align}
+This is exactly the shifted unitbox $[1,2] \times [1,2]$.
+In fact, we can easily generalize this a general $n$-dimensional system by considering the template direction matrix $\Lambda = I_n$ where $I_n$ is the $n \times n$ identity matrix and two offset vectors $c_l, c_u$ of length $n$.
+This would yield the shifted $n$-dimensional unitbox:
+
+\begin{equation}
+[c_l(1), \;c_u(1)] \times [c_l(2),\; c_u(2)] \times \cdots \times [c_l(n), \; c_u(n)]
+\end{equation}
+%
+It is worth noting that axis-aligned box on the 2D plane above would give the representation:
+\begin{align}
+  \Lambda = \begin{bmatrix}
+            1 & 0 \\
+            0 & 1 \\
+            -1 & 0 \\
+            0 & -1 \\
+            \end{bmatrix} , \quad c = [2,2,-1,-1]^T
+\end{align}
+if we were to convert the half-space representation above into the form defined in Equation \ref{eq:halfspace}.
 \end{example}
 
-\noindent Alternatively, a parallelotope can also be represented in a \emph{generator representation}
+\noindent Alternatively, a parallelotope can also be represented in a \emph{generator representation}.
 %
 \begin{definition}
+\label{def:generator_def}
 The generator representation of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$.
@@
 %
-\noindent This representation is very similar to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
+\noindent This is esstentially a convex reprsentation of the parallelotope, which shares many similarities to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
 %
@@
 
-There is a simple method to convert from the half-space representation of $P$ to its generator representation.
+There is a simple method to convert from the half-space representation of $P$ to its generator representation:
+
+\hspace{1.5cm} \textbf{Half-Space Representation to Generator Representation Conversion}
+\begin{enumerate}
+
+  \item Obtain vertex $v_1$ by solving the linear equation $\Lambda x = c_{l}$.
+  \item The $j+1$ vertex is obtained by solving the linear equation $\Lambda x = \mu_{j}$ where $\mu_{j}[i] = c_{l}[i]$ when $i \neq j$ and $\mu_{j}[j] = c_{u}[j]$.
+  \item The anchor $v$ of the parallelotope is the vertex $v_1$ and the generators will be $g_{i} = v_{i+1} - v_{1}$.
+\end{enumerate}
 %
-Obtain vertex $v_1$ by solving the linear equation $\Lambda x = c_{l}$.
-%
-The $j+1$ vertex is obtained by solving the linear equation $\Lambda x = \mu_{j}$ where $\mu_{j}[i] = c_{l}[i]$ when $i \neq j$ and $\mu_{j}[j] = c_{u}[j]$.
-%
-The anchor $v$ of the parallelotope is the vertex $v_1$ and the generators will be $g_{i} = v_{i+1} - v_{1}$. Refer to \cite{dang2014parameter} for a more detailed exposition of this procedure.
+There is a procedure to perform the reverse direction, namely to convert from the generator representation to the half-space representation.
+However, we will not need this procedure for this thesis.
+Refer to \cite{dang2014parameter} for a more detailed exposition.
 
-%
 As a final remark, notice that for a parallelotope $P$, the generator representation also defines an affine transformation that maps $[0,1]^{n}$ to $P$.
 %
-Let us denote this affine transformation associated to $P$ as $T_P$.
-
+We refer to this affine transformation associated to $P$ as $T_P:[0,1]^n \rightarrow P$ when necessary.
+%
 \begin{example}
-\label{ex:ptope}
+Let us return to the axis-aligned box considered in Example \ref{ex:simple_ptope}.
+%
+To obtain the anchor, we end up with the trivial solution $x = 1, y=1$ by adhering to Step 1. Hence, the anchor is set to $v = (1,1)$
+%
+Subsequently, we solve for the two other vertices by following Step 2 to obtain $x=2, y= 1$ and $x=1,y=2$.
+%
+By Step 3, this would imply that the two generators are $g_1 = (2,1)-(1,1) = (1,0)$ and $g_2 = (1,2)-(1,1) = (0,1)$.
+%
+Now combine the anchor and generators to get the generator representation for this paralellotope:
+%
+\begin{equation}
+ P = (1,1) + \alpha_1\cdot(1,0) + \alpha_2\cdot(0,1) \quad \alpha_1, \alpha_2 \in [0,1]
+\end{equation}
+%
+This is exactly the unit box $[0,1]^2$ with its corner at the origin shifted to $(1,1)$.
+\end{example}
+%
+\begin{example}
+\label{ex:diag_ptope}
 Once again, consider the space $\reals^2$ and the parallelotope $P$ given in half-plane representation as $0 \leq x-y \leq 1$, $0 \leq y \leq 1$.
@@
 %
-Therefore, all the points in the paralellotope can be written as $(x,y) = (0,0) + \alpha_1 (1,0) + \alpha_2(1,1)$, $\alpha_1, \alpha_2 \in [0,1]$.
+Therefore, all the points in the paralellotope can be written as $(x,y) = (0,0) + \alpha_1 \cdot (1,0) + \alpha_2\cdot (1,1)$, $\alpha_1, \alpha_2 \in [0,1]$.
 \end{example}
@@
 \noindent This parallelotope bundle will be the geometric data structure which will enclose the region we compute to be the approximation of the exact reachable set.
+Observe that $Q$ can be expressed as the conjunction of all the linear constraints defining each parallelotope $P_i \in \ptopeset{Q}$.
 
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \subsection{Bernstein Polynomials}
@@
 
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \input{staticalg}

### 2022-07-04 thesis/results.tex
Started dynamic algo and results section

@@
 \chapter{Experimental Results}
-
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \subsection{Kaa}
 \label{sec:kaa}
-
+We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} \cite{kim2020kaa}.
+Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems.
+The optimization procedure for finding the direction offets is performed through the \emph{Kodiak} library.
+Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module.
+To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
+For systems of dimension fewer than or equal to three, we utilize scipy's convex hull routine.
+%
+For higher-dimensional systems, we employ the volume of the tightest enveloping box around the parallelotope bundle.
+%
+The total volume estimate of the overapproximation will be the sum of all the bundles' volume estimates.
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \subsection{Benchmarks}
 \label{sec:benchmarks}
+For benchmarking, we select six non-linear models with polynomial dynamics.
+%
+Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability with {\bf static} parallelotope bundles.
+%
+We choose benchmarks with polynomial dynamics to directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes.
+To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
+Here, {\bf diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
+%
+By {\bf diagonal parallelotopes}, we refer to parallelotopes defined only by axis-aligned and diagonal directions.
+Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles solely consisting of diagonal parallelotopes.
+Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
+Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
+%
+For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box.
+We refer to these parallelotopes as {\bf non-axis-aligned parallelotopes}.
 
+Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional COVID supermodel is explained in the subsequent subsection below.
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \subsection{COVID19 Supermodel}
 \label{sec:covid}
+We benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
+This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients.
+These patients can infect susceptible members with a fixed probability.
+The dynamics account for this new group and its interactions with the traditional SIR groups.
 
+\begin{align}
+  \begin{split}
+   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
+   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
+   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
+   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
+   R_A' & = R_A + (\gamma A)\cdot \Delta \\
+   R_I' & = R_I + (\gamma I)\cdot \Delta \\
+   D' & = D + (\eta I)\cdot \Delta
+  \end{split}
+\end{align}
+where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
+We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate.
+The parameters are set based on figures shown in \cite{ansumali2020modelling}.
+The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
+
+\begin{table}
+%\hspace{-5em}
+  \centering
+\begin{tabular}{|p{1.5cm}|c|p{1.7cm}|c|c|p{5cm}|}
+\hline
+Model & Dimension & Parameters & \# steps & $\Delta$ & \hspace{1.5cm}Initial Box \\
+\hline
+Vanderpol & 2 & \quad \quad \; - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
+\hline
+Jet Engine& 2 & \quad \quad \; - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
+\hline
+Neuron \cite{fitzhugh1961impulses}& 2 & \quad \quad \; - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
+\hline
+SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 & $s \in [0.79,0.8], i \in [0.19,0.2], r = 0$ \\
+\hline
+Coupled \newline Vanderpol & 4 & \quad \quad \; - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
+\hline
+COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & \quad \quad \quad \; \; Stated Below\\
+\hline
+\end{tabular}
+\caption{Benchmark models and relevant information}
+\label{tab:modeldyns}
+\end{table}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \subsection{Comparison of Template Generation Techniques}
 \label{sec:compare}
+
+The results of testing our dynamic strategies against static ones are summarized in Table~\ref{tab:voltable}.
+For models previously defined in Sapo, we set the static parallelotopes to be exactly those found in Sapo.
+If a model is not implemented in Sapo, we simply use the static parallelotopes defined in a model of equal dimension.
+To address the unavailability of a four-dimensional model implemented in Sapo, we sampled random subsets of five static non-axis-aligned parallelotopes and chose the flowpipe with smallest volume.
+%
+% As justifiction for this scheme, we wish to remark the time required in testing all possible subsets of diagonal parallelotopes.
+%
+A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is infeasible.
+
+% are $2{n \choose 2}$ diagonal directions for an $n$-dimensional system, giving a total of $n + 2 {n \choose 2}$ axis-aligned and diagonal directions.
+Thus, there are ${n + 2 {n \choose 2} \choose n} = {n^2 \choose n}$ possible diagonal parallelotopes for an $n$-dimensional system.
+This explosive growth in search space size compelled us to proceed by sampling the search space instead.
+
+From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks.
+In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the ratio imparts differing effects.
+Observe that using parallelotopes defined by linear approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+%\input{comp_temp_fig}

### 2022-07-04 thesis/staticalg.tex
Started dynamic algo and results section

@@
  \\
-    & \min_{P \in \lambda^Q_i} {\min_{x \in P}{\Lambda_{i}^P \cdot f(x)}} \leq \min_{x \in Q}{\Lambda_{i}^Q \cdot f(x)} \label{eq:min_over_ptopes}
+    & \max_{P \in \lambda^Q_i} {\min_{x \in P}{\Lambda_{i}^P \cdot f(x)}} \leq \min_{x \in Q}{\Lambda_{i}^Q \cdot f(x)} \label{eq:min_over_ptopes}
 \end{align}
@@
 
-  \begin{figure}[ht]
+  \begin{figure}
   \label{fig:kaa_sir}
@@
 
-\subsection{Optimization Modes}
+\begin{example}
+To include an example of a higher-dimensional non-linear system, we introduce the Phosporaley model.
+The Phosphoraley model describes a certain cellular regulatory system.
+It is captured by seven variables governed by the following discretized dynamics:
+\begin{align}
+  \begin{split}
+      x^1_{k+1} &= x^1_k + ( -\alpha \cdot x^1_k + \beta \cdot x^3_k x^4_k)\cdot \Delta \\
+      x^2_{k+1} &= x^2_k + (  \alpha\cdot  x^1_k - x^2_k)\cdot \Delta \\
+      x^3_{k+1} &= x^3_k + ( x^2_k - \beta \cdot x^3_k x^4_k)\cdot \Delta \\
+      x^4_{k+1} &= x^4_k + ( \beta \cdot x^5_k x^6_k - \beta \cdot x^3_k x^4_k)\cdot \Delta \\
+      x^5_{k+1} &= x^5_k + ( -\beta \cdot x^5_k x^6_k + \beta \cdot x^3_k x^4_k)\cdot \Delta \\
+      x^6_{k+1} &= x^6_k + ( \alpha\cdot  x^7_k - \beta \cdot x^5_k x^6_k)\cdot \Delta \\
+      x^7_{k+1} &= x^7_k + ( -\alpha\cdot  x^7_k + \beta \cdot  x^5_k x^6_k)\cdot \Delta
+  \end{split}
+\end{align}
+Here, we set the two parameters $\alpha,\beta$ as $\alpha=0.5$ and $\beta=5$.
+The discretization step is set to $\Delta =0.01$ and we propagate the reachble set for $S=300$ time steps.
+Additionally, the initial box is set to be $[1.00,1.01]^7$.
+%
+Figure \ref{fig:kaa_phos} depicts the projection of the reachable set on the first three variables $x_1, x_2, x_3$.
+The paralleotopes are the axis-aligned parallotope and another with the following template direction matrix:
+
+\begin{equation}
+  \Lambda = \begin{bmatrix}
+            1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+            0 & 1 & 0 & 0 & 0 & 0 & 0 \\
+            0 & 0 & 1 & 0 & 0 & 0 & 0 \\
+            0 & 0 & 0 & 0 & 1 & 0 & 0 \\
+            0 & 0 & 0 & 0 & 0 & 1 & 0 \\
+            0 & 0 & 0 & 0 & 0 & 0 & 1 \\
+            0 & 0 & 1 & 1 & 0 & 0 & 0 \\
+            \end{bmatrix}
+\end{equation}
+
+\begin{figure}
+\label{fig:kaa_phos}
+  \hspace*{-2.3cm}
+  \includegraphics[width=1.3\textwidth]{figures/PhosporlayProjOn X_1, X_2, X_3.png}
+  \caption{Projection of Reachable Set of the Phosporaley model propagated 300 steps in time.}
+\end{figure}
+\end{example}

### 2022-07-05 thesis/conclusion.tex
Adding evaluation tables and plots

--- 2021-08-15 Nonlinear/formats2021/evaluation.tex
@@
-\vspace{-1em}
-\section{Evaluation}
-\label{sec:eval}
-%\vspace{-0.5em}
- We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} \cite{kim2020kaa}.
-Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems.
-The optimization procedure for finding the direction offets is performed through the \emph{Kodiak} library.
-Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module.
-To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize scipy's convex hull routine.
- %
-For higher-dimensional systems, we employ the volume of the tightest enveloping box around the parallelotope bundle.
- %
-The total volume estimate of the overapproximation will be the sum of all the bundles' volume estimates.
-
- %For two-dimensional models, note that there are only five non-trivial templates.
-The trivial template here refers to the initial box which is defined exactly the axis-aligned directions.
-Refer to the non-axis aligned templates.
-At each every time, pca.
-lin app directions to
-\vspace{-1em}
-\subsubsection{Model Dynamics}
-For benchmarking, we select six non-linear models with polynomial dynamics.
-Benchmarks against more general dynamics can be found in the appendix of the \href{https://arxiv.org/abs/2105.11796}{expanded verison}.
-%
-Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability with {\bf static} parallelotope bundles.
-%
-We choose benchmarks with polynomial dynamics to directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes.
-To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
-Here, {\bf diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
-%
-By {\bf diagonal parallelotopes}, we refer to parallelotopes defined only by axis-aligned and diagonal directions.
-Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles solely consisting of diagonal parallelotopes.
-Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
-Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
-%
-For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box.
-We refer to these parallelotopes as {\bf non-axis-aligned parallelotopes}.
-
-%Maybe mention something about role of diagonal templates in program verfication here.
-\begin{example}
-In two dimensions, $\mathbb{R}^2$, we have the two unit axis-aligned directions, $[1,0]^T, [0,1]^T$. The diagonal directions will then be
-\[ [1,1]^T, \; [1,-1]^T\]
-Consequently, the diagonal parallelotopes will precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
-%
-For two dimensions, the non-axis-aligned parallelotopes would be all parallelotopes excluding the one defined by directions $[0,1]^T, [1,0]^T$. \newline
-
-\noindent Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions
-%
-\[ [1,0,0]^T, [0,1,0]^T,[0,0,1]^T \]
-%
-Calculating the diagonal directions yields the following directions:
-%
-\[ [1,1,0]^T, \; [1,-1,0]^T, \; [1,0,1]^T,\;  [1,0,-1]^T,\;  [0,1,1]^T, \; [0,1,-1]^T \]
-%
-Thus, there are ${9 \choose 3} = 84$ diagonal parallelotopes.
-%
-Note that for both examples, we do not consider the directions' negative counterparts since we define parallelotopes use both the positive and negative counterparts of a its chosen directions.
-\end{example}
-Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional COVID supermodel is explained in the subsequent subsection below.
-
-\vspace{-1em}
-\begin{table}[h!]
-%\hspace{-5em}
-  \centering
-\begin{tabular}{|p{1.5cm}|c|p{1.7cm}|c|c|p{5cm}|}
-\hline
-Model & Dimension & Parameters & \# steps & $\Delta$ & \hspace{1.5cm}Initial Box \\
-\hline
-Vanderpol & 2 & \quad \quad \; - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
-\hline
-Jet Engine& 2 & \quad \quad \; - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
-\hline
-Neuron \cite{fitzhugh1961impulses}& 2 & \quad \quad \; - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
-\hline
-SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 & $s \in [0.79,0.8], i \in [0.19,0.2], r = 0$ \\
-\hline
-Coupled \newline Vanderpol & 4 & \quad \quad \; - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
-\hline
-COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & \quad \quad \quad \; \; Stated Below\\
-\hline
-\end{tabular}
-\caption{Benchmark models and relevant information}
-\label{tab:modeldyns}
-\end{table}
-
-\vspace{-2em}
-\noindent \textbf{COVID Supermodel}
-%
-We benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
-This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients.
-These patients can infect susceptible members with a fixed probability.
-The dynamics account for this new group and its interactions with the traditional SIR groups.
-
-\begin{align}
-  \begin{split}
-   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
-   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
-   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
-   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
-   R_A' & = R_A + (\gamma A)\cdot \Delta \\
-   R_I' & = R_I + (\gamma I)\cdot \Delta \\
-   D' & = D + (\eta I)\cdot \Delta
-  \end{split}
-\end{align}
-where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
-We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate.
-The parameters are set based on figures shown in \cite{ansumali2020modelling}.
-The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
-
-\vspace{-1em}
-\subsubsection{Accuracy of Dynamic Strategies}
-The results of testing our dynamic strategies against static ones are summarized in Table~\ref{tab:voltable}.
-For models previously defined in Sapo, we set the static parallelotopes to be exactly those found in Sapo.
-If a model is not implemented in Sapo, we simply use the static parallelotopes defined in a model of equal dimension.
-To address the unavailability of a four-dimensional model implemented in Sapo, we sampled random subsets of five static non-axis-aligned parallelotopes and chose the flowpipe with smallest volume.
-%
-% As justifiction for this scheme, we wish to remark the time required in testing all possible subsets of diagonal parallelotopes.
-%
-A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is infeasible.
-
-% are $2{n \choose 2}$ diagonal directions for an $n$-dimensional system, giving a total of $n + 2 {n \choose 2}$ axis-aligned and diagonal directions.
-Thus, there are ${n + 2 {n \choose 2} \choose n} = {n^2 \choose n}$ possible diagonal parallelotopes for an $n$-dimensional system.
-This explosive growth in search space size compelled us to proceed by sampling the search space instead.
-
-From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks.
-In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the ratio imparts differing effects.
-Observe that using parallelotopes defined by linear approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.
-
-\begin{figure}[h!]
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5Lin_.png}
-\caption{5 Lin}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_1PCA5Lin_.png}
-\caption{1 PCA 5 Lin}
-\end{subfigure}%
-
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5PCA_.png}
-\caption{5 PCA}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_5PCA1Lin_.png}
-\caption{5 PCA 1 Lin}
-\end{subfigure}
-
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_Sapo_.png}
-\caption{Sapo}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_Sapo_.png}
-\caption{Sapo}
-\end{subfigure}
-\caption{Effect of varying ratio between the number of PCA and Linear Approximation parallelotopes.
-The Vanderpol (left) and the FitzHugh-Nagumo Neuron (right) phase plots are shown to illustrate differing effects of varying the PCA/LinApp ratio.
-The initial set for the Vanderpol model is set to $x \in [0,0.05], \, y \in [1.95,2]$}
-\label{fig:PCALinAppRatio}
-\end{figure}
-
-\vspace{-1em}
-\subsubsection{Performance under Increasing Initial Sets}
-A key advantage of our dynamic strategies is the improved ability to control the wrapping error naturally arising from larger initial sets.
-Figure \ref{fig:InitVolReachComp} presents charts showcasing the effect of increasing initial sets on the total flowpipe volume.
-We vary the initial box dimensions to gradually increase the box's volume.
-We then plot the total flowpipe volume after running the benchmark.
-The same initial boxes are also used in computations using Sapo's static parallelotopes.
-The number of parallelotopes defined by PCA and Linear Approximation directions were chosen based on best performance as seen in Table \ref{tab:voltable}.
-We remark that our dynamic strategies perform better than static ones in controlling the total flowpipe volume as the initial set becomes larger.
-On the other hand, the performance of static parallelotopes tends to degrade rapidly as we increase the volume of the initial box.
-
-\vspace{-1em}
-\subsubsection{Performance against Random Static Templates}
-We additionally benchmark our dynamic strategies against static random parallelotope bundles.
-We sample such parallelotopes in $n$ dimensions by first sampling a set of $n$ directions uniformly on the surface of the unit $(n-1)$-sphere, then defining our parallelotope using these sampled directions.
-We sample twenty of these parallelotopes for each trial and average the total flowpipe volumes.
-As shown in Figure \ref{fig:RanStaticStratComp}, our best-performing dynamic strategies consistently outperform static random strategies for all tested benchmarks.
-
-\vspace{-0.8em}
-\section{Conclusions}
+\chapter{Conclusion}
+\label{chap:conclusion}
 In this paper, we investigated two techniques for generating templates dynamically: first using linear approximation of the dynamics, and second using PCA.
@@
 This is inspired by the approach behind the recent tool \emph{PIRK} \cite{devonport2020pirk}.
-
-\subsection{Acknowledgements}
-Parasara Sridhar Duggirala and Edward Kim acknowledge the support of the Air Force Office of Scientific Research under award number FA9550-19-1-0288 and FA9550-21-1-0121 and National Science Foundation (NSF) under grant numbers CNS 1935724 and CNS 2038960.
-Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the United States Air Force or the National Science Foundation.
-
-\begin{table}[h!]
-\hspace{1em}
-\begin{subtable}[h]{0.45\textwidth}
-     \centering
-     \begin{tabular}{|c|c|}
-     \hline
-     Strategy & Total  Volume \\
-     \hline\
-     5 LinApp & 0.227911 \\
-     \hline
-     1 PCA, 4 LinApp& 0.225917 \\
-     \hline
-     2 PCA, 3 LinApp & 0.195573 \\
-     \hline
-     {\bf 3 PCA, 2 LinApp} & {\bf 0.188873} \\
-     \hline
-     4 PCA, 1 LinApp & 1.227753\\
-     \hline
-     5 PCA & 1.509897 \\
-     \hline
-     5 Static Diagonal(Sapo) & 2.863307  \\
-     \hline
-    \end{tabular}
-    \caption{Vanderpol}
-    \label{tab:vdpvol}
- \end{subtable}\hspace{1em}
- \begin{subtable}[h]{0.45\textwidth}
-      \centering
-      \begin{tabular}{|c|c|}
-      \hline
-      Strategy & Total  Volume \\
-      \hline
-      5 LinApp & 58199.62 \\
-      \hline
-      1 PCA, 4 LinApp & 31486.16 \\
-      \hline
-      {\bf 2 PCA, 3 LinApp} & {\bf 5204.09}\\
-      \hline
-      3 PCA, 2 LinApp & 6681.76 \\
-      \hline
-      4 PCA, 1 LinApp& 50505.10 \\
-      \hline
-      5 PCA  & 84191.15 \\
-      \hline
-      5 Static Diagonal (Sapo) & 66182.18  \\
-      \hline
-     \end{tabular}
-     \caption{Jet Engine}
-     \label{tab:enginevol}
-  \end{subtable}
-
-  \hspace{1em}
-  \begin{subtable}[h]{0.45\textwidth}
-       \centering
-       \begin{tabular}{|c|c|}
-       \hline
-       Strategy & Total  Volume \\
-       \hline
-       5 LinApp  & 154.078\\
-       \hline
-       1 PCA, 4 LinApp  & 136.089\\
-       \hline
-       2 PCA, 3 LinApp  & 73.420\\
-       \hline
-       {\bf 3 PCA , 2 LinApp } & {\bf 73.126} \\
-       \hline
-       4 PCA, 1 LinApp  & 76.33 \\
-       \hline
-       5 PCA & 83.896 \\
-       \hline
-       5 Static Diagonal  (Sapo) & 202.406  \\
-       \hline
-      \end{tabular}
-      \caption{FitzHugh-Nagumo}
-      \end{subtable} \hspace{1em}
-      \begin{subtable}[h]{0.45\textwidth}
-        \centering
-        \begin{tabular}{|c|c|}
-        \hline
-        Strategy & Total  Volume \\
-        \hline
-        {\bf 2 LinApp } & {\bf 0.001423} \\
-        \hline
-        1 PCA, 1 LinApp & 0.106546\\
-
-        \hline
-        2 PCA  & 0.117347\\
-        \hline
-        2 Static Diagonal (Sapo) & 0.020894\\
-        \hline
-       \end{tabular}
-       \caption{SIR}
-       \label{tab:sirvol}
-    \end{subtable}
-
-    \hspace{1em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         5 LinApp & 5.5171 \\
-         \hline
-         {\bf 1 PCA, 4 LinApp } & {\bf 5.2536} \\
-         \hline
-         2 PCA, 3 LinApp  & 5.6670\\
-         \hline
-         3 PCA, 2 LinApp  & 5.5824\\
-         \hline
-         4 PCA, 1 LinApp  & 312.2108 \\
-         \hline
-         5 PCA  & 388.0513 \\
-         \hline
-         5 Static Diagonal (Best) & 3023.4463  \\
-         \hline
-        \end{tabular}
-        \caption{Coupled Vanderpol}
-        \label{tab:sirvol}
-     \end{subtable}\hspace{1 em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         3 LinApp & $2.95582227 * 10^{-10}$ \\
-         \hline
-         {\bf 1 PCA, 2 LinApp } & {\bf $2.33007583 * 10^{-10}$}\\
-         \hline
-         2 PCA, 1 LinApp &$ 4.02751770 * 10^{-9}$\\
-         \hline
-         3 PCA & $4.02749571 * 10^{-9}$\\
-         \hline
-         3 Static Diagonal (Sapo) & $4.02749571 * 10^{-9}$\\
-         \hline
-        \end{tabular}
-        \caption{COVID}
-        \label{tab:covidvol}
-     \end{subtable}
- \caption{Tables presenting upper bounds on the total reachable set volume by strategy.
-The static directions are retrieved and/or inspired from Sapo models of equal dimension for benchmarking.
-The best performing strategy is highlighted in bold.}
-     \label{tab:voltable}
-\end{table}
-\clearpage
-
-\begin{figure}[h!]
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVol.png}
-    \caption{Vanderpol}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVol.png}
-    \caption{Jet Engine}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVol200steps.png}
-    \caption{Neuron}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVol.png}
-    \caption{SIR}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVol.png}
-    \caption{Coupled Vanderpol}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVol.png}
-    \caption{COVID}
-    \end{subfigure}
-
-    \caption{Comparison between the performance of diagonal static parallelotope bundles and that of the best performing dynamic parallelotope bundles as the volume of the initial set grows.}
-    \label{fig:InitVolReachComp}
-\end{figure}
-\clearpage
-
-\begin{figure}[h!]
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVolRanStrat.png}
-    \caption{Vanderpol}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVolRanStrat.png}
-    \caption{Jet Engine}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVolRanStrat.png}
-    \caption{Neuron}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVolRanStrat.png}
-    \caption{SIR}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVolRanStrat.png}
-    \caption{Coupled Vanderpol}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVolRanStrat.png}
-    \caption{COVID}
-    \end{subfigure}
-
-    \caption{Comparision between random static strategies and the best performing dynamic strategies as the volume of the initial set grows.
-The total reachable set volumes for random static strategies are averaged over ten trials for each system.}
-    \label{fig:RanStaticStratComp}
-\end{figure}
-\clearpage

### 2022-07-05 thesis/formats.tex
Adding evaluation tables and plots

--- 2022-07-03 thesis/formats.tex
@@
 
-\subsection{Local Linear Approximations}
+In this chapter, we cover a method of generating template directions dynamically and automatically.
+%
+By dynamic, we mean that the template directions must be generated adaptively based on sampled trajectories and/or data from the state of the system.
+%
+By automatic, we mean that the template directions require no consideration from the user to proceed with the reachable set computation.
+This is in contrast to the original static algorithm treated in Section \ref{sec:static} where the user must input his or her own template directions to specify the parallelotopes before the computation starts.
+%
+To briefly outline the structure of this chapter, we first expound on two techniques we utilize to dynamically generate template directions at each step.
+%
+The first method is based on local linear approximations where the algorithm approximates the dynamics as a linear transformation based on sample trajectories.
+%
+The second method is based on Principal Compoenent Analysis (PCA) where the algorithm runs PCA on the end points of the sample trajectories.
+%
+Finally, we cover the high-level pseudo-code of the dynamic algorithm and explain a set of parameters we feed into the algorithm in order to improve performance and the accuracy of the outputted reachable set.
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\section{Local Linear Approximations}
 \label{sec:lin_app}
+%
+Intuitively speaking, if time step is discretized to be sufficiently small, propagating trajectories according to the non-linear dynamics $f$ for one step could lead to good lienar approximations of the dynamics within a small region.
+%
+%
+To do this, we first sample a set of points in the parallelotope bundle called \emph{support points} and propagate them to the next step using the dynamics $f$.
+%
+Support points are a subset of the vertices of the parallelotope that either maximize or minimize the template directions over the parallelotope bundle.
+%
+That is the support points are the set of points $x_i$ such that:
+\begin{equation}
+\label{eq:support}
+x_i = \max_{x \in Q} \Lambda^Q_i\cdot x
+\end{equation}
+%
+for all tempalte direcitons of the bundle $\Lambda^Q_i$. These are all found by a straightforward linear program.
+%
+%In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear transformation.
+%
+We use the support points as a data-driven approach to find the best-fit linear function to use.
+%
+To find the approximate linear transformation, let $x_i$ denote the support points calculated by Equation \ref{eq:support}.
+We perform the following least-squares procedure: the objective would be to find an linear transformation $A$ such that it minimizes the following objective function:
+\begin{equation}
+  \min_{A} \sum_{x_i} ||f(x_i) - Ax_i||^2_2
+\end{equation}
+%
+where $||\cdot||_2$ is the standard $2$-norm on $\reals^n$.
+%
+If the dynamics of a system is linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \T x \leq c_{u}$, is the set $c_{l} \leq \T \cdot A^{-1} x \leq c_{u}$.
+%
+Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the non-linear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
+%
 
-\subsection{Principal Component Analysis}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\section{Principal Component Analysis}
 \label{sec:pca}
+The second technique for generating template directions performs Principal Component Analysis (PCA) over the images of the support points.
+%
+PCA is a standard technique in Statistical Machine Learning used for reduction of dimensionality by performing Singular Value Decomposition (SVD) on the covariance matrix generated by a set of data points.
+%
+Since the covariance matrix is symmetric (i.e $A = A^T$), the eigenvectors of this matrix will always be a orthonormal basis of the system.
+%
+Using PCA is a reasonable choice as it produces orthonormal directions that can construct a rotated box for bounding the points.
+%
 
-\subsection{The Dynamic Algorithm}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+\section{The Dynamic Algorithm}
 \label{sec:dyna_algo}
+Observe that in general, our input dynamics are non-linear and therefore, the reachable set is generally non-convex.
+%
+On the other hand, a parallelotope bundle is always a convex set.
+%
+To mitigate the drawbacks of this discrepancy, we can improve accuracy of this representation by considering more template directions and more parallelotopes.
+%
+To this end, we add a parameter called \emph{template lifespan}, where we use the generated linear approximation and/or PCA template directions not only from the current step but also from previous steps.
+%
+ In our benchmarks, we tune each of the options (PCA / linear approximation as well as lifespan option) to demonstrate that specific parameters generate more accurate reachable sets than those generated by the static algorithm.
+
+The new approach is given in Algorithm~\ref{alg:new}.
+%
+During each step, the algorithm computes a collection of template directions from the two techniques outlined in Sections \ref{sec:lin_app} and \ref{sec:pca}.
+Each technique will be encoded as a subroutine labeled \texttt{ApproxLinearTrans} and \texttt{PCA}.
+%
+The \texttt{ApproxLinearTrans} function computes the best approximation of a linear transformation given a list of points before and after the one-step transformation $f$.
+%
+The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points.
+%
+Now each subroutine will return a collection of $n$ template directions, which in turn will specifiy exactly one parallelotope.
+%
+Hence, two parallelotopes, one generated by \texttt{ApproxLinearTrans} and the other by \texttt{PCA}, can be added to the parallelotope bundle at each step.
+%
+
+There are a few subtle points to be made about the sub-routines used in Algorithm \ref{alg:new}.
+%
+First, The algorithm makes use of helper function \texttt{hstack}, which attaches two matrices along their rows.
+In other words, \texttt{hstack} can be visualized as two matrices with the same number of columns stacked on top of one another.
+%
+Second, we assume that the sub-routine  \texttt{PCA} returns the orthonormal eigenvectors as rows. This assures that the eigenvectors are rows of a template direction matrix of a parallelotope.
+%
+Third, the \texttt{Maximize} and \texttt{Minimize} sub-routines encapsulate the linear programming procedures required to compute the support points as discussed in Equation \ref{eq:support}.
+%
+Both sub-routines take the feasible region as the first parameter and the objective function as the second parameter.
+%
+%The notation $M_{*,i}$ is used to refer to the $i^{th}$ column of matrix $M$.
+ Finally, the subroutine \texttt{TransformBundle} is the same as specified in Algorithm~\ref{alg:old}.
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\input{dynamic_alg_fig}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+Algorithm~\ref{alg:new} computes the dynamic templates for each time step $k$.
+%
+Line~\ref{ln:linearapprox} computes the linear approximation of the non-linear dynamics and this linear approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
+%
+The PCA directions of the images of support points is computed in line~\ref{ln:pca}.
+%
+For the time step $k$, the linear approximation and PCA templates direction matrices are given as $\Lambda_{k}^{lin}$ and $\Lambda_{k}^{pca}$, respectively.
+%
+To improve the accuracy of the reachable set, we compute the overapproximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the template lifespan parameter $L$.
+%
+Alternatively, we assign each parallelotope a parameter $L$ which dictates the number of steps we keep the parallelotope in the bundle after adding it in the current step.

### 2022-07-05 thesis/initvolreachtab.tex
Adding evaluation tables and plots

--- 2021-05-25 Nonlinear/formats2021/evaluation.tex
@@
-\vspace{-1em}
-\section{Evaluation}
-\label{sec:eval}
-%\vspace{-0.5em}
- We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} \cite{kim2020kaa}.
-Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems.
-The optimization procedure for finding the direction offets is performed through the \emph{Kodiak} library.
-Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module.
-To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize scipy.spatial library's Convex hull routine.
- These routines are esstentially Python wrappers around their corresponding \emph{QHull} procedures.
-For higher-dimensional systems, we employ the volume of the tightest enveloping box around the intersection of the parallelotopes.
- The dimensions of this box are easily calculated through linear programming. The total volume estimate of the reachable set will be the sum of all the computed bundles' volume estimates.
-
- %For two-dimensional models, note that there are only five non-trivial templates.
-The trivial template here refers to the initial box which is defined exactly the axis-aligned directions.
-Refer to the non-axis aligned templates.
-At each every time, pca.
-lin app directions to
-\vspace{-1em}
-\subsubsection{Model Dynamics}
-For benchmarking, we select six non-linear models with polynomial dynamics.
-Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability analysis with {\bf static} parallelotope bundles.
-In these cases, we directly compare the performance of our dynamic strategies with the static parallelotopes defined in Sapo.
-To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
-Here, {\bf diagonal directions} are defined to be the vectors created by adding and subtracting distinct pairs of all unit axis-aligned vectors from each other.
-By {\bf diagonal parallelotopes}, we refer to parallelotopes which are defined only by axis-aligned and diagonal directions.
-Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles which solely consists of diagonal parallelotopes.
-Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
-Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
-For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box.
-We refer to these parallelotopes as {\bf non-axis-aligned parallelotopes}.
-
-%Maybe mention something about role of diagonal templates in program verfication here.
-\begin{example}
-In two dimensions, $\mathbb{R}^2$, we have the two unit axis-aligned directions, $[1,0]^T, [0,1]^T$. The diagonal directions will then
-
-\[ [1,1]^T, \; [1,-1]^T\]
-Consequently, the diagonal parallelotopes will be precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
-% Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions $[1,0,0]^T, [0,1,0]^T,[0,0,1]^T$. Calculating the diagonal directions yields the following directions:
-% \[ [1,1,0]^T, \; [1,-1,0]^T, \; [1,0,1]^T,\;  [1,0,-1]^T,\;  [0,1,1]^T, \; [0,1,-1]^T \]
-% Thus, there are ${9 \choose 3} = 84$ diagonal parallelotopes.
-%
-% Note that for both examples, we do not consider the directions' negative counterparts since we define parallelotopes use both the positive and negative counterparts of a its chosen directions.
-%
-% For two dimensions, the non-axis-aligned parallelotopes would be all parallelotopes excluding the one defined by directions $[0,1]^T, [1,0]^T$.
-In particular, there are 5 non-axis-aligned diagonal parallelotopes.
-\end{example}
-Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional model COVID supermodel is explained in the subsequent subsection below.
-
-\vspace{-1em}
-\begin{table}[h!]
-%\hspace{-5em}
-  \centering
-\begin{tabular}{|p{1.5cm}|c|p{1.5cm}|c|c|p{5cm}|}
-\hline
-Model & Dimension & Parameters & \# steps & $\Delta$ & \hspace{1.5cm}Initial Box \\
-\hline
-Vanderpol & 2 & - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
-\hline
-Jet Engine& 2 & - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
-\hline
-Neuron \cite{fitzhugh1961impulses}& 2 & - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
-\hline
-SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 & $s \in [0.79,0.8], i \in [0.19,0.2], r = 0$ \\
-\hline
-Coupled \newline Vanderpol & 4 & - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
-\hline
-COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & Stated Below\\
-\hline
-\end{tabular}
-\caption{Benchmark models and relevant information}
-\label{tab:modeldyns}
-\end{table}
-
-\vspace{-2em}
-\noindent \textbf{COVID Supermodel:}
-%
-We further benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
-This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients.
-These patients have the ability to infect susceptible members with a fixed probability.
-The following dynamics account for this new group and its interactions with the traditional groups found in the SIR model.
-The continuous dynamics given in \cite{indiansuper2020supermodel} are discretized through the Euler method.
-
-\begin{align}
-  \begin{split}
-   S_A & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
-   S_I & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
-   A & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
-   I & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
-   R_A & = R_A + (\gamma A)\cdot \Delta \\
-   R_I & = R_I + (\gamma I)\cdot \Delta \\
-   D & = D + (\eta I)\cdot \Delta
-  \end{split}
-\end{align}
-where the variables denote the fraction of a population of individuals designated as \emph{Susceptible Transition to Asymptomatic $(S_A)$}, \emph{Susceptible Transition to Infected $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
-We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the average probablity of infection, $\gamma$ is the average removal rate, and $\eta$ is the average mortality rate.
-The parameters are set based on figures shown in \cite{ansumali2020modelling} for the USA.
-The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
-
-\subsubsection{Accuracy of Dynamic Strategies}
-The results of testing our dynamic strategies against static ones are summarized in Table~\ref{tab:voltable}.
-For models previously defined in Sapo, we set the static parallelotopes to be exactly those found in Sapo.
-If a model is not implemented in Sapo, we simply use the static parallelotopes defined in a model of equal dimension.
-To address the unavailability of a four-dimensional model implemented in Sapo, we sampled random subsets of five static non-axis-aligned parallelotopes and chose the flowpipe with smallest volume.
-%
-% As justifiction for this scheme, we wish to remark the time required in testing all possible subsets of diagonal parallelotopes.
-%
-A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is impossible.
-
-% are $2{n \choose 2}$ diagonal directions for an $n$-dimensional system, giving a total of $n + 2 {n \choose 2}$ axis-aligned and diagonal directions.
-Thus, there are ${n + 2 {n \choose 2} \choose n} = {n^2 \choose n}$ possible diagonal parallelotopes for an $n$-dimensional system.
-This explosive growth in search space size compelled us to proceed by sampling the search space instead.
-
-From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks.
-In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the PCA/LinApp ratio imparts differing effects on the reachable set.
-Observe that using parallelotopes defined by Linear Approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.
-
-\begin{figure}[h!]
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5Lin_.png}
-\caption{5 Lin}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_1PCA5Lin_.png}
-\caption{1 PCA 5 Lin}
-\end{subfigure}%
-
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5PCA_.png}
-\caption{5 PCA}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_5PCA1Lin_.png}
-\caption{5 PCA 1 Lin}
-\end{subfigure}
-
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_Sapo_.png}
-\caption{Sapo}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_Sapo_.png}
-\caption{Sapo}
-\end{subfigure}
-\caption{Effect of varying ratio between the number of PCA and Linear Approximation parallelotopes.
-The Vanderpol (left) and the FitzHugh-Nagumo Neuron (right) phase plots are shown to illustrate differing effects of varying the PCA/LinApp ratio.
-The initial set for the Vanderpol model is set to $x \in [0,0.05], \, y \in [1.95,2]$}
-\label{fig:PCALinAppRatio}
-\end{figure}
-
-\vspace{-1em}
-\subsubsection{Performance under Increasing Initial Sets}
-A key advantage of our dynamic strategies is the improved ability to control the wrapping error naturally arising from larger initial sets.
-Figure \ref{fig:InitVolReachComp} presents charts showcasing the effect of increasing initial sets on the total flowpipe volume.
-We vary the initial box dimensions to gradually increase the box's volume.
-We then plot the total flowpipe volume after running the benchmark.
-The same initial boxes are also fed as input into computation using only static parallelotopes borrowed from Sapo.
-The number of parallelotopes defined by PCA and Linear Approximation directions were chosen based on best performance as seen in Table \ref{tab:voltable}.
-We remark that our dynamic strategies perform better than static ones in controlling the total flowpipe volume as the initial set becomes larger.
-On the other hand, the performance of static parallelotopes tends to degrade rapidly as we increase the volume of the initial box.
-
-\subsubsection{Performance against Random Static Templates}
-We additionally benchmark our dynamic strategies against static random parallelotope bundles.
-We sample such parallelotopes in $n$ dimensions by first sampling a set of $n$ directions uniformly on the surface of the unit $(n-1)$-sphere, then defining our parallelotope using these sampled directions.
-We sample twenty of these parallelotopes for each trial and average the total flowpipe volumes.
-As shown in Figure \ref{fig:RanStaticStratComp}, our best-performing dynamic strategies consistently outperform static random strategies for all tested benchmarks.
-
-\vspace{-1em}
-\section{Conclusions}
-In this paper we investigated two techniques for generating templates dynamically, first using linear approximation of the dynamics, and second using PCA.
-%
-We demonstrated that these techniques improve the accuracy of reachable set by an order of magnitude when compared to static or random template directions.
-%
-We also observed that both these techniques improve the accuracy of the reachable sets for different benchamrks.
-%
-In future, we intend to investigate Koopman linearization techniques for computing the template directions~\cite{bak2021reachability}.
-
-\begin{table}[h!]
-\hspace{1em}
-\begin{subtable}[h]{0.45\textwidth}
-     \centering
-     \begin{tabular}{|c|c|}
-     \hline
-     Strategy & Total  Volume \\
-     \hline\
-     5 LinApp & 0.227911 \\
-     \hline
-     1 PCA, 4 LinApp& 0.225917 \\
-     \hline
-     2 PCA, 3 LinApp & 0.195573 \\
-     \hline
-     {\bf 3 PCA, 2 LinApp} & {\bf 0.188873} \\
-     \hline
-     4 PCA, 1 LinApp & 1.227753\\
-     \hline
-     5 PCA & 1.509897 \\
-     \hline
-     5 Static Diagonal(Sapo) & 2.863307  \\
-     \hline
-    \end{tabular}
-    \caption{Vanderpol}
-    \label{tab:vdpvol}
- \end{subtable}\hspace{1em}
- \begin{subtable}[h]{0.45\textwidth}
-      \centering
-      \begin{tabular}{|c|c|}
-      \hline
-      Strategy & Total  Volume \\
-      \hline
-      5 LinApp & 58199.62 \\
-      \hline
-      1 PCA, 4 LinApp & 31486.16 \\
-      \hline
-      {\bf 2 PCA, 3 LinApp} & {\bf 5204.09}\\
-      \hline
-      3 PCA, 2 LinApp & 6681.76 \\
-      \hline
-      4 PCA, 1 LinApp& 50505.10 \\
-      \hline
-      5 PCA  & 84191.15 \\
-      \hline
-      5 Static Diagonal (Sapo) & 66182.18  \\
-      \hline
-     \end{tabular}
-     \caption{Jet Engine}
-     \label{tab:enginevol}
-  \end{subtable}
-
-  \hspace{1em}
-  \begin{subtable}[h]{0.45\textwidth}
-       \centering
-       \begin{tabular}{|c|c|}
-       \hline
-       Strategy & Total  Volume \\
-       \hline
-       5 LinApp  & 154.078\\
-       \hline
-       1 PCA, 4 LinApp  & 136.089\\
-       \hline
-       2 PCA, 3 LinApp  & 73.420\\
-       \hline
-       {\bf 3 PCA , 2 LinApp } & {\bf 73.126} \\
-       \hline
-       4 PCA, 1 LinApp  & 76.33 \\
-       \hline
-       5 PCA & 83.896 \\
-       \hline
-       5 Static Diagonal  (Sapo) & 202.406  \\
-       \hline
-      \end{tabular}
-      \caption{FitzHugh-Nagumo}
-      \end{subtable} \hspace{1em}
-      \begin{subtable}[h]{0.45\textwidth}
-        \centering
-        \begin{tabular}{|c|c|}
-        \hline
-        Strategy & Total  Volume \\
-        \hline
-        {\bf 2 LinApp } & {\bf 0.001423} \\
-        \hline
-        1 PCA, 1 LinApp & 0.106546\\
-
-        \hline
-        2 PCA  & 0.117347\\
-        \hline
-        2 Static Diagonal (Sapo) & 0.020894\\
-        \hline
-       \end{tabular}
-       \caption{SIR}
-       \label{tab:sirvol}
-    \end{subtable}
-
-    \hspace{1em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         5 LinApp & 5.5171 \\
-         \hline
-         {\bf 1 PCA, 4 LinApp } & {\bf 5.2536} \\
-         \hline
-         2 PCA, 3 LinApp  & 5.6670\\
-         \hline
-         3 PCA, 2 LinApp  & 5.5824\\
-         \hline
-         4 PCA, 1 LinApp  & 312.2108 \\
-         \hline
-         5 PCA  & 388.0513 \\
-         \hline
-         5 Static Diagonal (Best) & 3023.4463  \\
-         \hline
-        \end{tabular}
-        \caption{Coupled Vanderpol}
-        \label{tab:sirvol}
-     \end{subtable}\hspace{1 em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         3 LinApp & $2.95582227 * 10^{-10}$ \\
-         \hline
-         {\bf 1 PCA, 2 LinApp } & {\bf $2.33007583 * 10^{-10}$}\\
-         \hline
-         2 PCA, 1 LinApp &$ 4.02751770 * 10^{-9}$\\
-         \hline
-         3 PCA & $4.02749571 * 10^{-9}$\\
-         \hline
-         3 Static Diagonal (Sapo) & $4.02749571 * 10^{-9}$\\
-         \hline
-        \end{tabular}
-        \caption{COVID}
-        \label{tab:covidvol}
-     \end{subtable}
- \caption{Tables presenting total reachable set volume by strategy.
-The static directions are retrieved and/or inspired from Sapo models of equal dimension for benchmarking.
-The best performing strategy is highlighted in bold.}
-     \label{tab:voltable}
-\end{table}
-\newpage
-\begin{figure}[h!]
+begin{figure}[h!]
     \hspace{-1.5em}
@@
 \end{figure}
-\newpage
-
-\begin{figure}[h!]
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVolRanStrat.png}
-    \caption{Vanderpol}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVolRanStrat.png}
-    \caption{Jet Engine}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVolRanStrat.png}
-    \caption{Neuron}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVolRanStrat.png}
-    \caption{SIR}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVolRanStrat.png}
-    \caption{Coupled Vanderpol}
-    \end{subfigure}%
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVolRanStrat.png}
-    \caption{COVID}
-    \end{subfigure}
-
-    \caption{Comparision between random static strategies and the best performing dynamic strategies as the volume of the initial set grows.
-The total reachable set volumes for random static strategies are averaged over ten trials for each system.}
-    \label{fig:RanStaticStratComp}
-\end{figure}
-\newpage
+\clearpage

### 2022-07-05 thesis/prelim.tex
Adding evaluation tables and plots

@@
 if we were to convert the half-space representation above into the form defined in Equation \ref{eq:halfspace}.
+%
+From here on out, we refer to parallelotopes defined by the $n$ axis-aligned directions as the \emph{axis-aligned} parallelotopes.
+\end{example}
+
+\begin{example}
+  \label{ex:simple_diag_ptope}
+  We can also consider the axis-aligned directions rotated $45^\circ$ counter-clockwise.
+This would yield the two diagonal directions $[1,1]^T, [-1,1]$.
+Suppose we set the upper and lower offsets to be $c_u = [1,1]^T$ and $c_l = [-1,-1]^T$ respectively, then once again by Definition \ref{def:halfspace_def}, the bounded region in $\reals^2$ will be the conjunction of the linear inequalities:
+  \begin{align}
+    -1 \leq x - y \leq 1 \\
+    -1 \leq y - x \leq 1
+  \end{align}
+  In general, we define \emph{diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
+As a matter of convenience however, we refer to \emph{diagonal parallelotopes} as those defined by a combination of axis-aligned and diagonal directions.
+This definition will be useful when we consider parallelotopes defined by unconventional directions (i.e those which are neither axis-aligned nor diagonal).
 \end{example}

### 2022-07-05 thesis/randstaticstrattab.tex
Adding evaluation tables and plots

--- 2021-05-25 Nonlinear/formats2021/evaluation.tex
@@
-\vspace{-1em}
-\section{Evaluation}
-\label{sec:eval}
-%\vspace{-0.5em}
- We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} \cite{kim2020kaa}.
-Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems.
-The optimization procedure for finding the direction offets is performed through the \emph{Kodiak} library.
-Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module.
-To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize scipy.spatial library's Convex hull routine.
- These routines are esstentially Python wrappers around their corresponding \emph{QHull} procedures.
-For higher-dimensional systems, we employ the volume of the tightest enveloping box around the intersection of the parallelotopes.
- The dimensions of this box are easily calculated through linear programming. The total volume estimate of the reachable set will be the sum of all the computed bundles' volume estimates.
-
- %For two-dimensional models, note that there are only five non-trivial templates.
-The trivial template here refers to the initial box which is defined exactly the axis-aligned directions.
-Refer to the non-axis aligned templates.
-At each every time, pca.
-lin app directions to
-\vspace{-1em}
-\subsubsection{Model Dynamics}
-For benchmarking, we select six non-linear models with polynomial dynamics.
-Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability analysis with {\bf static} parallelotope bundles.
-In these cases, we directly compare the performance of our dynamic strategies with the static parallelotopes defined in Sapo.
-To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
-Here, {\bf diagonal directions} are defined to be the vectors created by adding and subtracting distinct pairs of all unit axis-aligned vectors from each other.
-By {\bf diagonal parallelotopes}, we refer to parallelotopes which are defined only by axis-aligned and diagonal directions.
-Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles which solely consists of diagonal parallelotopes.
-Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
-Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
-For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box.
-We refer to these parallelotopes as {\bf non-axis-aligned parallelotopes}.
-
-%Maybe mention something about role of diagonal templates in program verfication here.
-\begin{example}
-In two dimensions, $\mathbb{R}^2$, we have the two unit axis-aligned directions, $[1,0]^T, [0,1]^T$. The diagonal directions will then
-
-\[ [1,1]^T, \; [1,-1]^T\]
-Consequently, the diagonal parallelotopes will be precisely be defined by unique pairs of these directions, giving us a total ${4 \choose 2} = 6$ diagonal parallelotopes.
-% Similarly, in $\mathbb{R}^3$, we have three unit axis-aligned directions $[1,0,0]^T, [0,1,0]^T,[0,0,1]^T$. Calculating the diagonal directions yields the following directions:
-% \[ [1,1,0]^T, \; [1,-1,0]^T, \; [1,0,1]^T,\;  [1,0,-1]^T,\;  [0,1,1]^T, \; [0,1,-1]^T \]
-% Thus, there are ${9 \choose 3} = 84$ diagonal parallelotopes.
-%
-% Note that for both examples, we do not consider the directions' negative counterparts since we define parallelotopes use both the positive and negative counterparts of a its chosen directions.
-%
-% For two dimensions, the non-axis-aligned parallelotopes would be all parallelotopes excluding the one defined by directions $[0,1]^T, [1,0]^T$.
-In particular, there are 5 non-axis-aligned diagonal parallelotopes.
-\end{example}
-Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional model COVID supermodel is explained in the subsequent subsection below.
-
-\vspace{-1em}
-\begin{table}[h!]
-%\hspace{-5em}
-  \centering
-\begin{tabular}{|p{1.5cm}|c|p{1.5cm}|c|c|p{5cm}|}
-\hline
-Model & Dimension & Parameters & \# steps & $\Delta$ & \hspace{1.5cm}Initial Box \\
-\hline
-Vanderpol & 2 & - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
-\hline
-Jet Engine& 2 & - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
-\hline
-Neuron \cite{fitzhugh1961impulses}& 2 & - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
-\hline
-SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 & $s \in [0.79,0.8], i \in [0.19,0.2], r = 0$ \\
-\hline
-Coupled \newline Vanderpol & 4 & - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
-\hline
-COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & Stated Below\\
-\hline
-\end{tabular}
-\caption{Benchmark models and relevant information}
-\label{tab:modeldyns}
-\end{table}
-
-\vspace{-2em}
-\noindent \textbf{COVID Supermodel:}
-%
-We further benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
-This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients.
-These patients have the ability to infect susceptible members with a fixed probability.
-The following dynamics account for this new group and its interactions with the traditional groups found in the SIR model.
-The continuous dynamics given in \cite{indiansuper2020supermodel} are discretized through the Euler method.
-
-\begin{align}
-  \begin{split}
-   S_A & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
-   S_I & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
-   A & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
-   I & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
-   R_A & = R_A + (\gamma A)\cdot \Delta \\
-   R_I & = R_I + (\gamma I)\cdot \Delta \\
-   D & = D + (\eta I)\cdot \Delta
-  \end{split}
-\end{align}
-where the variables denote the fraction of a population of individuals designated as \emph{Susceptible Transition to Asymptomatic $(S_A)$}, \emph{Susceptible Transition to Infected $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
-We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the average probablity of infection, $\gamma$ is the average removal rate, and $\eta$ is the average mortality rate.
-The parameters are set based on figures shown in \cite{ansumali2020modelling} for the USA.
-The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
-
-\subsubsection{Accuracy of Dynamic Strategies}
-The results of testing our dynamic strategies against static ones are summarized in Table~\ref{tab:voltable}.
-For models previously defined in Sapo, we set the static parallelotopes to be exactly those found in Sapo.
-If a model is not implemented in Sapo, we simply use the static parallelotopes defined in a model of equal dimension.
-To address the unavailability of a four-dimensional model implemented in Sapo, we sampled random subsets of five static non-axis-aligned parallelotopes and chose the flowpipe with smallest volume.
-%
-% As justifiction for this scheme, we wish to remark the time required in testing all possible subsets of diagonal parallelotopes.
-%
-A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is impossible.
-
-% are $2{n \choose 2}$ diagonal directions for an $n$-dimensional system, giving a total of $n + 2 {n \choose 2}$ axis-aligned and diagonal directions.
-Thus, there are ${n + 2 {n \choose 2} \choose n} = {n^2 \choose n}$ possible diagonal parallelotopes for an $n$-dimensional system.
-This explosive growth in search space size compelled us to proceed by sampling the search space instead.
-
-From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks.
-In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the PCA/LinApp ratio imparts differing effects on the reachable set.
-Observe that using parallelotopes defined by Linear Approximation directions is more effective than those defined by PCA directions in the Vanderpol model whereas the Neuron model shows the opposite trend.
-
-\begin{figure}[h!]
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5Lin_.png}
-\caption{5 Lin}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_1PCA5Lin_.png}
-\caption{1 PCA 5 Lin}
-\end{subfigure}%
-
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_5PCA_.png}
-\caption{5 PCA}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_5PCA1Lin_.png}
-\caption{5 PCA 1 Lin}
-\end{subfigure}
-
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/VDP_Sapo_.png}
-\caption{Sapo}
-\end{subfigure}%
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\textwidth]{figures/PhasePlots/Neuron_Sapo_.png}
-\caption{Sapo}
-\end{subfigure}
-\caption{Effect of varying ratio between the number of PCA and Linear Approximation parallelotopes.
-The Vanderpol (left) and the FitzHugh-Nagumo Neuron (right) phase plots are shown to illustrate differing effects of varying the PCA/LinApp ratio.
-The initial set for the Vanderpol model is set to $x \in [0,0.05], \, y \in [1.95,2]$}
-\label{fig:PCALinAppRatio}
-\end{figure}
-
-\vspace{-1em}
-\subsubsection{Performance under Increasing Initial Sets}
-A key advantage of our dynamic strategies is the improved ability to control the wrapping error naturally arising from larger initial sets.
-Figure \ref{fig:InitVolReachComp} presents charts showcasing the effect of increasing initial sets on the total flowpipe volume.
-We vary the initial box dimensions to gradually increase the box's volume.
-We then plot the total flowpipe volume after running the benchmark.
-The same initial boxes are also fed as input into computation using only static parallelotopes borrowed from Sapo.
-The number of parallelotopes defined by PCA and Linear Approximation directions were chosen based on best performance as seen in Table \ref{tab:voltable}.
-We remark that our dynamic strategies perform better than static ones in controlling the total flowpipe volume as the initial set becomes larger.
-On the other hand, the performance of static parallelotopes tends to degrade rapidly as we increase the volume of the initial box.
-
-\subsubsection{Performance against Random Static Templates}
-We additionally benchmark our dynamic strategies against static random parallelotope bundles.
-We sample such parallelotopes in $n$ dimensions by first sampling a set of $n$ directions uniformly on the surface of the unit $(n-1)$-sphere, then defining our parallelotope using these sampled directions.
-We sample twenty of these parallelotopes for each trial and average the total flowpipe volumes.
-As shown in Figure \ref{fig:RanStaticStratComp}, our best-performing dynamic strategies consistently outperform static random strategies for all tested benchmarks.
-
-\vspace{-1em}
-\section{Conclusions}
-In this paper we investigated two techniques for generating templates dynamically, first using linear approximation of the dynamics, and second using PCA.
-%
-We demonstrated that these techniques improve the accuracy of reachable set by an order of magnitude when compared to static or random template directions.
-%
-We also observed that both these techniques improve the accuracy of the reachable sets for different benchamrks.
-%
-In future, we intend to investigate Koopman linearization techniques for computing the template directions~\cite{bak2021reachability}.
-
-\begin{table}[h!]
-\hspace{1em}
-\begin{subtable}[h]{0.45\textwidth}
-     \centering
-     \begin{tabular}{|c|c|}
-     \hline
-     Strategy & Total  Volume \\
-     \hline\
-     5 LinApp & 0.227911 \\
-     \hline
-     1 PCA, 4 LinApp& 0.225917 \\
-     \hline
-     2 PCA, 3 LinApp & 0.195573 \\
-     \hline
-     {\bf 3 PCA, 2 LinApp} & {\bf 0.188873} \\
-     \hline
-     4 PCA, 1 LinApp & 1.227753\\
-     \hline
-     5 PCA & 1.509897 \\
-     \hline
-     5 Static Diagonal(Sapo) & 2.863307  \\
-     \hline
-    \end{tabular}
-    \caption{Vanderpol}
-    \label{tab:vdpvol}
- \end{subtable}\hspace{1em}
- \begin{subtable}[h]{0.45\textwidth}
-      \centering
-      \begin{tabular}{|c|c|}
-      \hline
-      Strategy & Total  Volume \\
-      \hline
-      5 LinApp & 58199.62 \\
-      \hline
-      1 PCA, 4 LinApp & 31486.16 \\
-      \hline
-      {\bf 2 PCA, 3 LinApp} & {\bf 5204.09}\\
-      \hline
-      3 PCA, 2 LinApp & 6681.76 \\
-      \hline
-      4 PCA, 1 LinApp& 50505.10 \\
-      \hline
-      5 PCA  & 84191.15 \\
-      \hline
-      5 Static Diagonal (Sapo) & 66182.18  \\
-      \hline
-     \end{tabular}
-     \caption{Jet Engine}
-     \label{tab:enginevol}
-  \end{subtable}
-
-  \hspace{1em}
-  \begin{subtable}[h]{0.45\textwidth}
-       \centering
-       \begin{tabular}{|c|c|}
-       \hline
-       Strategy & Total  Volume \\
-       \hline
-       5 LinApp  & 154.078\\
-       \hline
-       1 PCA, 4 LinApp  & 136.089\\
-       \hline
-       2 PCA, 3 LinApp  & 73.420\\
-       \hline
-       {\bf 3 PCA , 2 LinApp } & {\bf 73.126} \\
-       \hline
-       4 PCA, 1 LinApp  & 76.33 \\
-       \hline
-       5 PCA & 83.896 \\
-       \hline
-       5 Static Diagonal  (Sapo) & 202.406  \\
-       \hline
-      \end{tabular}
-      \caption{FitzHugh-Nagumo}
-      \end{subtable} \hspace{1em}
-      \begin{subtable}[h]{0.45\textwidth}
-        \centering
-        \begin{tabular}{|c|c|}
-        \hline
-        Strategy & Total  Volume \\
-        \hline
-        {\bf 2 LinApp } & {\bf 0.001423} \\
-        \hline
-        1 PCA, 1 LinApp & 0.106546\\
-
-        \hline
-        2 PCA  & 0.117347\\
-        \hline
-        2 Static Diagonal (Sapo) & 0.020894\\
-        \hline
-       \end{tabular}
-       \caption{SIR}
-       \label{tab:sirvol}
-    \end{subtable}
-
-    \hspace{1em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         5 LinApp & 5.5171 \\
-         \hline
-         {\bf 1 PCA, 4 LinApp } & {\bf 5.2536} \\
-         \hline
-         2 PCA, 3 LinApp  & 5.6670\\
-         \hline
-         3 PCA, 2 LinApp  & 5.5824\\
-         \hline
-         4 PCA, 1 LinApp  & 312.2108 \\
-         \hline
-         5 PCA  & 388.0513 \\
-         \hline
-         5 Static Diagonal (Best) & 3023.4463  \\
-         \hline
-        \end{tabular}
-        \caption{Coupled Vanderpol}
-        \label{tab:sirvol}
-     \end{subtable}\hspace{1 em}
-    \begin{subtable}[h]{0.45\textwidth}
-         \centering
-         \begin{tabular}{|c|c|}
-         \hline
-         Strategy & Total  Volume \\
-         \hline
-         3 LinApp & $2.95582227 * 10^{-10}$ \\
-         \hline
-         {\bf 1 PCA, 2 LinApp } & {\bf $2.33007583 * 10^{-10}$}\\
-         \hline
-         2 PCA, 1 LinApp &$ 4.02751770 * 10^{-9}$\\
-         \hline
-         3 PCA & $4.02749571 * 10^{-9}$\\
-         \hline
-         3 Static Diagonal (Sapo) & $4.02749571 * 10^{-9}$\\
-         \hline
-        \end{tabular}
-        \caption{COVID}
-        \label{tab:covidvol}
-     \end{subtable}
- \caption{Tables presenting total reachable set volume by strategy.
-The static directions are retrieved and/or inspired from Sapo models of equal dimension for benchmarking.
-The best performing strategy is highlighted in bold.}
-     \label{tab:voltable}
-\end{table}
-\newpage
-\begin{figure}[h!]
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/VDPInitReachVol.png}
-    \caption{Vanderpol}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVol.png}
-    \caption{Jet Engine}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/NeuronInitReachVol200steps.png}
-    \caption{Neuron}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/SIRInitReachVol.png}
-    \caption{SIR}
-    \end{subfigure}
-
-    \hspace{-1.5em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CVDPInitReachVol.png}
-    \caption{Coupled Vanderpol}
-    \end{subfigure}%
-    %\hspace{1em}
-    \begin{subfigure}{0.5\textwidth}
-    \centering
-    \includegraphics[width=1.1\textwidth, height=0.75\textwidth]{figures/InitVolVSReachVol/CovidInitReachVol.png}
-    \caption{COVID}
-    \end{subfigure}
-
-    \caption{Comparison between the performance of diagonal static parallelotope bundles and that of the best performing dynamic parallelotope bundles as the volume of the initial set grows.}
-    \label{fig:InitVolReachComp}
-\end{figure}
-\newpage
-
 \begin{figure}[h!]
@@
 \end{figure}
-\newpage

### 2022-07-05 thesis/results.tex
Adding evaluation tables and plots

@@
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-\subsection{Kaa}
+\section{Kaa}
 \label{sec:kaa}
@@
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-\subsection{Benchmarks}
+\section{Benchmarks}
 \label{sec:benchmarks}
@@
 
-Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional COVID supermodel is explained in the subsequent subsection below.
+Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional COVID supermodel is explained in the subsequent section below.
 
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-\subsection{COVID19 Supermodel}
+\section{COVID19 Supermodel}
 \label{sec:covid}
@@
 
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \begin{table}
@@
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-\subsection{Comparison of Template Generation Techniques}
+\section{Comparison of Template Generation Techniques}
 \label{sec:compare}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 
+\subsection{Accuracy of Dynamic Strategies}
+\label{sec:dynamic_accuracy}
 The results of testing our dynamic strategies against static ones are summarized in Table~\ref{tab:voltable}.
@@
 
+%\input{voltable}
+
+%\input{comp_temp_fig}
+
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-%\input{comp_temp_fig}
+\subsection{Performance under Increasing Initial Sets}
+\label{sec:increasing_initial}
+
+A key advantage of our dynamic strategies is the improved ability to control the wrapping error naturally arising from larger initial sets.
+Figure \ref{fig:InitVolReachComp} presents charts showcasing the effect of increasing initial sets on the total flowpipe volume.
+We vary the initial box dimensions to gradually increase the box's volume.
+We then plot the total flowpipe volume after running the benchmark.
+The same initial boxes are also used in computations using Sapo's static parallelotopes.
+The number of parallelotopes defined by PCA and Linear Approximation directions were chosen based on best performance as seen in Table \ref{tab:voltable}.
+We remark that our dynamic strategies perform better than static ones in controlling the total flowpipe volume as the initial set becomes larger.
+On the other hand, the performance of static parallelotopes tends to degrade rapidly as we increase the volume of the initial box.
+
+%\input{initvolreachtab}
+
+\subsection{Performance against Random Static Templates}
+\label{sec:random_static}
+
+We additionally benchmark our dynamic strategies against static random parallelotope bundles.
+We sample such parallelotopes in $n$ dimensions by first sampling a set of $n$ directions uniformly on the surface of the unit $(n-1)$-sphere, then defining our parallelotope using these sampled directions.
+We sample twenty of these parallelotopes for each trial and average the total flowpipe volumes.
+As shown in Figure \ref{fig:RanStaticStratComp}, our best-performing dynamic strategies consistently outperform static random strategies for all tested benchmarks.
+
+%\input{randomstaticstrattab}

### 2022-07-07 thesis/covid19.tex
First draft of thesis more or less completed

\section{COVID19 Supermodel}
\label{sec:covid}
We benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}. This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients. These patients can infect susceptible members with a fixed probability. The dynamics account for this new group and its interactions with the traditional SIR groups.

\begin{align}
  \label{eq:covid19_dynamics}
  \begin{split}
   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
   R_A' & = R_A + (\gamma A)\cdot \Delta \\
   R_I' & = R_I + (\gamma I)\cdot \Delta \\
   D' & = D + (\eta I)\cdot \Delta
  \end{split}
\end{align}
where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}. We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate. The parameters are set based on figures shown in \cite{ansumali2020modelling}. The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.

Plots of the reachable set for this model tuned to specific values of parameters $\beta, \gamma, \eta$ were published in an ACM Sigbed Blogpost detailing applications of Formal methods for simulating disease dynamics \cite{bak2021covid}.
%
The main theme of the blogpost revolved around the difficulty of extracting accurate parameters from real-world data samples.
%
Certainly we can estimate the parameters by analyzing real-world data. However, minute changes in the parameters could yield estimates which would vastly overestimate or underestimate the true population of infected or asymptomatic patients.
%
This error is further compounded as our time horizon increases due to many of the issues pertaining to wrapping error, as discussed in the Introduction of this thesis (see \ref{sec:intro}).
%
Reachability analysis provides not only a method of simulating the disease dynamics in order to provide meaningful information for policy decisions, but also to demonstrate the effect of slight changes in the parameters on the conservativeness of the outputted reachable set.
%
See Figure \ref{fig:covid_plots} for the two plots published in the blogpost. The Confirmed population is the sum of the number of Symptomatic ($I$) and Asymptomatic ($A$) populations of the dynamics presented in Equation \ref{eq:covid19_dynamics}. The plots were created by dividing the total period into two separate periods.
%
We decided to separate the periods as the parameters presented in the original paper \cite{ansumali2020modelling} were estimated separately according to these exact periods. Furthermore, we wished to control the conservativeness of the over-approximation of the reachable set.
%
The red line represents the real data gathered from India during the prescribed time periods with the light-blue region representing the predicted region based on the parameters given in \cite{ansumali2020modelling}.
%
Several lines representing trajectories generated according to specific parameters are also plotted in order to convey the effect of changes in the population under slight changes of the underlying parameters.
%
\begin{figure}
  \centering
  % First COVID plot.
  \subfloat[Confirmed population from 06/21/20-08/22/20, ]{
    \includegraphics[width= 0.9\textwidth]{figures/IndianRecoveredPlotJun21toAug22-2-2048x813.png}
  }

  \subfloat[Confirmed population from 08/22/20-10/01/20]{
    \includegraphics[width= 0.9\textwidth]{figures/IndianRecoveredPlotAug22toOct01-2-2048x813.png}
  }
\caption{Reachable Sets for India's COVID19 Confirmed Population for the period 06/21/20-10/01/20}
\label{fig:covid_plots}
\end{figure}

### 2022-07-07 thesis/initvolreachcomp.tex
First draft of thesis more or less completed

\def \initvolwidth {0.6}
\def \initvolheight {0.5}
\def \initvolhorshift {-1.5cm}
\def \initvolvertshift {-2cm}

\begin{figure}
  \centering
  \vspace*{\initvolvertshift}
    \hspace*{\initvolhorshift}\subfloat[Vanderpol]{
    \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/VDPInitReachVol.png}
  }
  \subfloat[Jet Engine]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVol.png}
  }

  \hspace*{\initvolhorshift}\subfloat[Neuron]{
    \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/NeuronInitReachVol200steps.png}
  }
  \subfloat[Coupled Vanderpol]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/CVDPInitReachVol.png}
  }

  \hspace*{\initvolhorshift}\subfloat[SIR]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/SIRInitReachVol.png}
  }
  \subfloat[COVID]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/CovidInitReachVol.png}
  }

  \caption{Comparison between the performance of diagonal static parallelotope bundles and that of the best performing dynamic parallelotope bundles as the volume of the initial set grows.}
  \label{fig:InitVolReachComp}
\end{figure}

### 2022-07-07 thesis/intro.tex
First draft of thesis more or less completed

@@
 \chapter{Introduction}
+\label{sec:intro}
 
@@
 %
-For example, reachability analysis has found many applications in formally verifying the safety properties of Cyber-physical Systems governed by Neural Network Controllers \cite{tran2019star, fan2020reachnn, bak2021nnenum}.
+For example, reachability analysis has found many applications in formally verifying the safety properties of Cyber-physical Systems, such as autonomous vehicles \cite{althoff2010reachability}, F-16 aircraft \cite{heidlauf2018verification}, and those governed by Neural Network Controllers \cite{tran2019star, fan2020reachnn, bak2021nnenum}.
 %

### 2022-07-07 thesis/prelim.tex
First draft of thesis more or less completed

@@
 From here on out, we refer to parallelotopes defined by the $n$ axis-aligned directions as the \emph{axis-aligned} parallelotopes.
+%
+For a visual plot of above axis-aligned parallelotope, see Figure \ref{fig:axis_align_ptope}
+
+\begin{figure}
+  \centering
+  \includegraphics{example-image-a}
+  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}.}
+  \label{fig:axis_align_ptope}
+\end{figure}
+
 \end{example}
@@
 This definition will be useful when we consider parallelotopes defined by unconventional directions (i.e those which are neither axis-aligned nor diagonal).
+To see a visual plot of the diagonal parallelotope, see Figure \ref{fig:diagonal_ptope}.
+
+  \begin{figure}
+    \centering
+    \includegraphics{example-image-b}
+    \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.}
+    \label{fig:diagonal_ptope}
+  \end{figure}
 \end{example}
@@
 
+\begin{tcolorbox}
 \hspace{1.5cm} \textbf{Half-Space Representation to Generator Representation Conversion}
@@
 \end{enumerate}
+\end{tcolorbox}
 %

### 2022-07-07 thesis/ranstaticstratcomp.tex
First draft of thesis more or less completed

\def \randvolwidth {0.6}
\def \randvolheight {0.5}
\def \randvolhorshift {-1.5cm}
\def \randvolvertshift {-2cm}

\begin{figure}
  \centering
  \vspace*{\randvolvertshift}
    \hspace*{\randvolhorshift}\subfloat[Vanderpol]{
    \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/VDPInitReachVolRanStrat.png}
  }
  \subfloat[Jet Engine]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/JetEngineInitReachVolRanStrat.png}
  }

  \hspace*{\randvolhorshift}\subfloat[Neuron]{
    \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/NeuronInitReachVolRanStrat.png}
  }
  \subfloat[Coupled Vanderpol]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/CVDPInitReachVolRanStrat.png}
  }

  \hspace*{\randvolhorshift}\subfloat[SIR]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/SIRInitReachVolRanStrat.png}
  }
  \subfloat[COVID]{
  \includegraphics[width=\initvolwidth\textwidth, height=\initvolheight \textwidth]{figures/InitVolVSReachVol/CovidInitReachVolRanStrat.png}
  }

  \caption{Comparision between random static strategies and the best performing dynamic strategies as the volume of the initial set grows. The total reachable set volumes for random static strategies are averaged over ten trials for each system.}
  \label{fig:RanStaticStratComp}s
\end{figure}

### 2022-07-07 thesis/results.tex
First draft of thesis more or less completed

@@
-\chapter{Experimental Results}
+\chapter{Evaluations}
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
@@
 \label{sec:kaa}
-We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} \cite{kim2020kaa}.
-Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems.
-The optimization procedure for finding the direction offets is performed through the \emph{Kodiak} library.
-Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module.
+%Mention kodiak, general structure of program, difference between kaa and dynamic.
+%
+We evaluate the efficacy of our dynamic parallelotope bundle strategies with our tool, \emph{Kaa} .
+%
+Kaa is written in Python and relies on several modules to perform reachable set computation.
+\begin{description}
+  \item[Numpy:] The \emph{Numpy} module is used to do all matrix computations, such as matrix multiplication and matrix inversions.
+It is also used to efficiently solve systems of linear equations, especially those which arise from converting from half-space representation to generator representation (See Section \ref{sec:parallelotope}), and execute the Least Squares routine required to compute the solution to Problem \ref{eq:least_squares}.
+  %
+  \item[Sympy:] The \emph{Sympy} module is used to do all symbolic computations.
+The polynomials which result from performing the $\Lambda_i \cdot f(x)$ in Equation \ref{eq:maxalpha} are all simplified and encapsulated by the $\texttt{sp.Poly}$ object.
+This allows extraction of coefficients of monomial terms to become a simple call to
+  $\texttt{sp.Poly.coeff\_monomial}$.
+  %
+  \item[Sklearn:] The \emph{Sklearn} module called to perform PCA on the end-points of the sample trajectories as described in Section \ref{sec:pca}.
+The exact method is $\texttt{sklearn.decomposition.PCA}$.
+  %
+  \item[Scipy:] The \emph{Scipy} module offers several auxiliary routines important to the our analysis of reachable sets, especially \texttt{scipy.spatial.ConvexHull}.
+  %
+  \item[Matplotlib:] The \emph{Matplotlib} module is for all plotting of the computed reachable sets.
+Kaa utilizes the library's animation features to even animate the evolution of the parallelotope bundle as the reachable set computation proceeds.
+  %
+  \item[Multiprocessing:] The \emph{multiprocessing} module parallelizes all the non-linear optimization procedures required to compute the new upper and lower offsets of all the template directions of the parallelotope bundle.
+This is expressed by lines 13,14 in Algorithm \ref{alg:old}.
+  %
+  \item[Swiglpk:] The \emph{swiglpk} module is a simple Python wrapper over the C library, \emph{GLPK} (GNU Linear Programming Kit).
+Kaa uses swiglpk for all LP problems, such as those which arise from computing the support points over a bundle (see Equation \ref{eq:support}).
+\end{description}
+
+The original version of Kaa was created to compactify and simplify \emph{Sapo}, a previous tool exploring reachability computation with static parallelotope bundles \cite{dreossi2017sapo}.
+%
+Through the expressiveness and terseness of Python, Algorithm \ref{alg:old} was implemented in only $650$ lines of code.
+%
+We released as a pedagogical tool to allow practitioners and students to easily experiment with parallelotopes-based reachability and understand the effects of choosing different template directions \cite{kim2020kaa}.
+
+To extend Kaa to handle dynamic parallelotope bundles, we replaced the original optimization procedure leveraging Bernstein polynomials (see Section \ref{sec:bernstein}) to \emph{Kodiak}.
+%
+% FIX KODIAK CITATIONS IN THESIS.BIB
+%
+Kodiak is an optimization library implemented in C\texttt{++} that implements a branch-and-bound algorithm for numerical approximations.
+It uses a combination of interval arithmetric and Bernstein enclosure to approximate solutions to optimization problems of the form shown in Equation \ref{eq:maxsup}.
+%
+The optimization procedure for finding the direction offets is performed through Kodiak.
+%
+We decided to use Kodiak primarily for two reasons:
+%
+\begin{enumerate}
+  \item Kodiak is very fast as a Python wrapper over the original C\texttt{++} implementation. It is much faster than Kaa's original procedure of computing all Bernstein coefficients.
+  \item Kodiak can handle a wider variety of dynamics, including those which feature trigonometric terms and square root terms.
+This allows us to generalize beyond the polynomial dynamcis first considered by \cite{dreossi2016parallelotope}.
+\end{enumerate}
+%
+
 To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize scipy's convex hull routine.
+For systems of dimension fewer than or equal to three, we utilize Scipy's convex hull routine.
 %
@@
 %
-The total volume estimate of the overapproximation will be the sum of all the bundles' volume estimates.
+The total volume estimate of the over-approximation will be the sum of all the computed bundles' volume estimates.
+%
+To be specific, if \texttt{ApproxVol} is the routine used to approximate the volume of a bundle, then by utilizing the notion of line 22 of Algorithm \ref{algo:new}:
+%
+\begin{equation}
+  \texttt{ApproxVol}(\overline\Theta) = \sum_{k} \texttt{ApproxVol}(\overline\Theta_k)
+\end{equation}
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
@@
 %
-Many of these models are also implemented in \emph{Sapo} \cite{dreossi2017sapo}, a previous tool exploring reachability with {\bf static} parallelotope bundles.
+Since many of these models are also implemented in Sapo, we choose benchmarks with polynomial dynamics to directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes.
+To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
 %
-We choose benchmarks with polynomial dynamics to directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes.
-To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
-Here, {\bf diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
+Recall through the discussion in Example \ref{ex:simple_diag_ptope} that we refer to parallelotopes defined only by axis-aligned and diagonal directions as \emph{diagonal parallelotopes}.
 %
-By {\bf diagonal parallelotopes}, we refer to parallelotopes defined only by axis-aligned and diagonal directions.
-Similarly, {\bf diagonal parallelotope bundles} are parallelotope bundles solely consisting of diagonal parallelotopes.
-Sapo primarily utilizes {\bf static diagonal parallelotope bundles} to perform its reachability computation.
+Similarly, \emph{diagonal parallelotope bundles} are parallelotope bundles solely consisting of diagonal parallelotopes.
+Sapo primarily utilizes \emph{static diagonal parallelotope bundles} to perform its reachability computation.
+%
 Note that the initial box, which is defined only through the axis-aligned directions, is contained in every bundle.
@@
 For our experiments, we are concerned with the effects of additional static or dynamic parallelotopes added alongside the initial box.
-We refer to these parallelotopes as {\bf non-axis-aligned parallelotopes}.
+We refer to these parallelotopes \emph{non-axis-aligned parallelotopes}.
 
@@
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-\section{COVID19 Supermodel}
-\label{sec:covid}
-We benchmark our dynamic strategies with the recently introduced COVID supermodel \cite{ansumali2020modelling}, \cite{indiansuper2020supermodel}.
-This model is a modified SIR model accounting for the possibility of \emph{asymptomatic} patients.
-These patients can infect susceptible members with a fixed probability.
-The dynamics account for this new group and its interactions with the traditional SIR groups.
 
-\begin{align}
-  \begin{split}
-   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
-   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
-   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
-   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
-   R_A' & = R_A + (\gamma A)\cdot \Delta \\
-   R_I' & = R_I + (\gamma I)\cdot \Delta \\
-   D' & = D + (\eta I)\cdot \Delta
-  \end{split}
-\end{align}
-where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
-We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate.
-The parameters are set based on figures shown in \cite{ansumali2020modelling}.
-The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
+\input{covid19}
 
@@
 
-%\input{voltable}
+\input{voltable}
 
@@
 
-%\input{initvolreachtab}
+\input{initvolreachcomp}
 
@@
 
-%\input{randomstaticstrattab}
+\input{ranstaticstratcomp}

### 2022-07-14 thesis/covid19.tex
slight modifications and corrections in prelims

@@
 %
-This error is further compounded as our time horizon increases due to many of the issues pertaining to wrapping error, as discussed in the Introduction of this thesis (see \ref{sec:intro}).
+This error is further compounded as our time horizon increases due to many of the issues pertaining to wrapping error, as discussed in the Introduction of this thesis (see Section \ref{sec:intro}).
 %
-Reachability analysis provides not only a method of simulating the disease dynamics in order to provide meaningful information for policy decisions, but also to demonstrate the effect of slight changes in the parameters on the conservativeness of the outputted reachable set.
+Reachability analysis provides not only a method of simulating the disease dynamics in order to provide illuminating information for policy decisions but also to demonstrate the effect of slight changes in the parameters on the conservativeness of the outputted reachable set.
 %
@@
 %
-Several lines representing trajectories generated according to specific parameters are also plotted in order to convey the effect of changes in the population under slight changes of the underlying parameters.
+Several lines representing trajectories generated according to specific parameters are also plotted in order to convey the effect of changes in the population under slight perturbations of the underlying parameters.
 %

### 2022-07-14 thesis/prelim.tex
slight modifications and corrections in prelims

@@
 We begin with some basic definitions pertaining to reachability and parallelotopes.
-The definition of Bernstein polynomials and the reachable set computaion algorithm will be defined.
+The definition of Bernstein polynomials and their enclosure properties will be stated.
 Finally, an outline of the reachability algorithm given by \cite{dreossi2016parallelotope} for polynomial dynamical systems will be presented.
@@
 
-As stated in the previous sections, this thesis pertains to the reachability analysis of dynamical systems.
-Roughly speaking, a dynamical system is governed by a set of differential equations such that the states of the system evolve according the solutions of the said differential equations.
+As stated in the previous section, this thesis pertains to the reachability analysis of dynamical systems.
+Roughly speaking, a dynamical system describes the behavior of states in space which is governed by a set of differential equations.
+The states of the system evolve according to the flow of time and the vector field induced by the governing differential equations.
+
+Towards introducing more terminology, the \emph{state} of a system, denoted as $x$, lies in a domain $D \subseteq \reals^n$ where the solutions to the differential equations are defined.
 %
-The state of a system, denoted as $x$, lies in a domain $D \subseteq \reals^n$ where the solution to the differential equations is defined.
+In the reachability analysis literature, there are several definitions for dynamical systems which appear depending on the authors' taste of formalizations.
+Here is a more rigorous definition appearing in \cite{dang1998reachability}:
+\begin{definition}
+\label{def:diff_dyna_sys}
+A \emph{continuous dynamical system} is a tuple $\langle X, f \rangle$ where $X = \reals^n$ is finite-dimensional Euclidean space and $f$ is a continuous function on $X$.
+A point (state) of $x_0 \in X$ evolves according to a tracjectory $\xi(t): \reals^+ \rightarrow X$ such that the following hold:
 %
-We restrict our attention to a specific definition of these dynamical systems:
+\begin{equation}
+  \label{eq:diff_dyna_sys}
+  \begin{split}
+    \xi(0) & = x_0 \\
+    \frac{d \xi(t)}{dt} & = f(\xi(t)) \quad \forall t \in \reals^+
+  \end{split}
+\end{equation}
+\end{definition}
+%
+\noindent Note that, by a simple separation of variables argument and an initial value of $\xi(0) = x_0$, Equation \ref{eq:diff_dyna_sys} yields the following form for $\xi$:
+\begin{equation}
+\xi(t) = x_0 + \int_0^t f(\xi(\tau)) d\tau
+\end{equation}
+%
+\noindent However, many members of the research community choose to simply express the system as:
+\begin{equation}
+\label{eq:cont_dynas}
+x' = f(x)
+\end{equation}
+for a continuous function $f:\reals^n \rightarrow \reals^n$. We restrict our attention to a discretized version of this definition for this thesis:
 %
 \begin{definition}
-A discrete-time nonlinear system is denoted as
+A \emph{discrete-time dynamical system} is denoted as
 \begin{equation}
@@
 \end{equation}
-where $f: \reals^{n} \rightarrow \reals^{n}$ is a nonlinear function.
+where $f: \reals^{n} \rightarrow \reals^{n}$ is a function and $x^{+}$ denotes the evolved output state.
 \end{definition}
 %
- Intuitively, the function $f$ takes input a state of the system and outputs the next step of the system evolved according to the non-linear dynamics.
+ In other words, the function $f$ takes input a state of the system and outputs the next step of the system evolved according to the dynamics.
 %
-Here, the function $f$ generally represents some discretized version of some specified continuous non-linear dynamical systems.
-Recall that a dynamical system is considered \emph{linear} if its dynamics can be expressed as
+Intuitively speaking, the function $f$ generally represents a discretized version of some specified continuous dynamical system.
+Here, a discrete-time dynamical system is considered \emph{linear} if its dynamics can be expressed as
 
 $$
-x' = Ax, \quad A \in \reals^{n \times n}
+x^{+} = Ax, \quad A \in \reals^{n \times n}
 $$
 %
-Otherwise, we deem the system to be \emph{nonlinear}.
+Otherwise, we deem the system to be \emph{non-linear}.
 %
-Hence, in particular, the function $f$ cannot be expressed as some matrix $A \in \reals^{n \times n}$.
+Hence, in particular, a non-linear function $f$ cannot be expressed as some matrix $A \in \reals^{n \times n}$.
 
-Examples of prominent non-linear dynamical systems include the Lotka-Volterra predator-prey model \cite{wangersky1978lotka}, FitzHugh Neuron model \cite{fitzhugh1961impulses}, and recently introduced COVID19 disease model \cite{indiansuper2020supermodel}.
+Examples of prominent non-linear dynamical systems include the Lotka-Volterra predator-prey model \cite{wangersky1978lotka}, the Fitz-Hugh Neuron model \cite{fitzhugh1961impulses}, and the recently introduced COVID19 disease model \cite{indiansuper2020supermodel}.
 Throughout this thesis, we discretize any continuous dynamics through the well-known Euler method.
@@
     \begin{split}
-        s' &= \beta \cdot s_k i_k \\
-        i' &= \beta \cdot s_k i_k - \gamma \cdot i_k \\
-        r' &= \gamma \cdot i_k
+        s' &= \beta \cdot s i \\
+        i' &= \beta \cdot s i - \gamma \cdot i \\
+        r' &= \gamma \cdot i
     \end{split}
@@
 There are two parameters, namely $\beta$ and $\gamma$, which influence the evolution of the system.
-$\beta$ is labeled as the contraction rate and $1/\gamma$ is the mean infective period.
+$\beta$ is labeled as the \emph{contraction rate} and $1/\gamma$ is the \emph{mean infective period}.
 %
@@
 
-The trajectory of a system that evolves according to Equation~\ref{eq:sys}, denoted as $\xi(x_0)$ is a sequence $x_0, x_1, \ldots $ where $x_{i+1} = f(x_i)$.
+We can now give a discretized version of a trajectory as presented in Definition \ref{def:diff_dyna_sys}.
+The trajectory of a system that evolves according to Equation~\ref{eq:sys}, denoted as $\xi_{x_0}$, is a sequence $x_0, x_1, \ldots $ where $x_{i+1} = f(x_i)$.
 %
-The $k^{th}$ element in this sequence $x_k$ is denoted as $\xi(x_0,k)$.
+The $k^{th}$ element in this sequence $x_k$ is denoted as $\xi_{x_0}(k)$.
 
@@
 \begin{equation}
-  \Theta_k = \{ \xi(x,k)\: | \: x \in \Theta\}
+  \Theta_k = \{ \xi_x(k)\: | \: x \in \Theta\}
 \label{eq:reachset}
@@
 \section{Parallelotope-based Reachability}
-
 \subsection{Parallelotopes}
 \label{sec:parallelotope}
-A parallelotope $P$ is a set of states in $\mathbb{R}^{n}$ captured by the tuple $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ is a matrix and $ç$ is a column vector.
+The heart of our reachability algorithm relies on geometric objects called \emph{parallelotopes}. In this section, we define parallelotopes and two representations for them.
+
+\begin{definition}
+\label{def:ptope}
+A \emph{parallelotope} $P \subset \mathbb{R}^{n}$ is captured by the tuple $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ is a matrix and $c \in \reals^{2n}$ is a column vector.
 We impose the condition that $\Lambda_{i+n} = -\Lambda_{i}$ for all $i \in \{1, \ldots, n\}$ such that
@@
 \end{equation}
+\end{definition}
+%
 We deem $\Lambda$ as the \emph{template direction matrix} where $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$ called the \emph{$i^{th}$ template direction}.
@@
 %
-Additionally, the definition of the paralleotope above requires that for each of $n$ ``postive" directions, there must exist a corresponding ``negative" direction.
+Additionally, the definition of the paralleotope above requires that for each of $n$ ``positive" directions, there must exist a corresponding ``negative" direction.
 This is encoded into the template matrix $\Lambda$ by the condition $\Lambda_{i+n} = -\Lambda_{i}$.
@@
 %
-Combining these remarks yields the \emph{half-space representation} of parallelotope $P$.
+Combining these remarks yields the half-space representation of parallelotope $P$.
 %
@@
 \label{def:halfspace_def}
-The half-space representation of parallelotope $P$ is tuple  $\langle \Lambda, c_l, c_u \rangle$ where $\Lambda \in \mathbb{R}^{n \times n}$ and $c_l, c_u \in \reals^n$
+The \emph{half-space representation} of parallelotope $P$ is tuple  $\langle \Lambda, c_l, c_u \rangle$ where $\Lambda \in \mathbb{R}^{n \times n}$ and $c_l, c_u \in \reals^n$
 such that
@@
 \end{equation}
-Then by Definition \ref{def:halfspace_def}, the bounded region in space will be the intersection of the following linear constraints:
+Then by Definition \ref{def:halfspace_def}, the half-space representation is given by:
+\begin{equation}
+\begin{bmatrix}
+  1 \\ 1
+\end{bmatrix} \leq \begin{bmatrix}
+    1 & 0 \\ 0 & 1
+  \end{bmatrix}\begin{bmatrix}
+   x \\ y
+  \end{bmatrix} \leq \begin{bmatrix}
+    2 \\ 2
+  \end{bmatrix}
+\end{equation}
+%
+Accordingly, the bounded region in space will be the intersection of the following linear constraints:
 %
 \begin{align}
+  \begin{split}
   1 \leq x \leq 2 \\
   1 \leq y \leq 2
+  \end{split}
 \end{align}
@@
   We can also consider the axis-aligned directions rotated $45^\circ$ counter-clockwise.
-This would yield the two diagonal directions $[1,1]^T, [-1,1]$.
-Suppose we set the upper and lower offsets to be $c_u = [1,1]^T$ and $c_l = [-1,-1]^T$ respectively, then once again by Definition \ref{def:halfspace_def}, the bounded region in $\reals^2$ will be the conjunction of the linear inequalities:
+This would yield the two diagonal directions $[1,1]^T, [-1,1]^T$.
+Suppose we set the upper and lower offsets to be $c_u = [1,1]^T$ and $c_l = [-1,-1]^T$ respectively, then once again by Definition \ref{def:halfspace_def}:
+  %
+  \begin{equation}
+  \begin{bmatrix}
+    -1 \\ -1
+  \end{bmatrix} \leq \begin{bmatrix}
+      1 & 1 \\ -1 & 1
+    \end{bmatrix}\begin{bmatrix}
+     x \\ y
+    \end{bmatrix} \leq \begin{bmatrix}
+      2 \\ 2
+    \end{bmatrix}
+  \end{equation}
+  %
+  The bounded region in $\reals^2$ will be the conjunction of the linear inequalities:
+  %
   \begin{align}
-    -1 \leq x - y \leq 1 \\
+    \begin{split}
+    -1 \leq x + y \leq 1 \\
     -1 \leq y - x \leq 1
+    \end{split}
   \end{align}
+  %
   In general, we define \emph{diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
 As a matter of convenience however, we refer to \emph{diagonal parallelotopes} as those defined by a combination of axis-aligned and diagonal directions.
-This definition will be useful when we consider parallelotopes defined by unconventional directions (i.e those which are neither axis-aligned nor diagonal).
+This definition will be useful when we consider parallelotopes defined by unconventional directions (i.e those template directions which are neither axis-aligned nor diagonal).
 To see a visual plot of the diagonal parallelotope, see Figure \ref{fig:diagonal_ptope}.
@@
 
-\noindent Alternatively, a parallelotope can also be represented in a \emph{generator representation}.
+\noindent Alternatively, a parallelotope can also be represented in a generator representation.
 %
@@
 \label{def:generator_def}
-The generator representation of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$.
+The \emph{generator representation}of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$.
 The vector  $v \in \mathbb{R}^n$ is called the \emph{anchor} and the $g_i \in \mathbb{R}^n$, are called the \emph{generators}.
@@
 %
-There is a procedure to perform the reverse direction, namely to convert from the generator representation to the half-space representation.
+\noindent There also exists a procedure to perform the reverse direction, namely to convert from the generator representation to the half-space representation.
 However, we will not need this procedure for this thesis.
@@
 where $\boldx^\boldi$ represents the monomial $x_1^{i_1}x_2^{i_2}\cdots x_n^{i_n}$.
-Every such real polynomial $p$ can be represented as linear combination of Bernstein basis polynomials of of degree $\boldd$:
+Every such real polynomial $p$ of degree at most $\boldd$ can be represented as linear combination of Bernstein basis polynomials of of degree $\boldd$:
 %
@@
 \noindent In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
+%
+%\begin{property}
+%\label{prop:basis_property}
+
+%The Bernstein basis polynomials $\$
+%\end{property}
 

### 2022-07-14 thesis/staticalg.tex
Adding evaluation tables and plots
First draft of thesis more or less completed
slight modifications and corrections in prelims

@@
 As mentioned in the previous section, the building block of the reachability algorithm relies on approximate solutions to a non-linear optimization problem over the unitbox domain.
-Consider a nonlinear function $h : \reals^n \rightarrow \reals$.
+Consider a non-linear function $h : \reals^n \rightarrow \reals$.
 The most general form of this optimization problem can be expressed as:
@@
 To distinguish between the distinct parallelotopes contained in the bundle, we add a new matrix called $\T^Q \in \mathbb{N}^{p \times n}$ such that $\T^Q_i$ is a vector of row indices of $\Lambda^Q$ which specify the template directions defining parallelotope $P_i \in \ptopeset{Q}$.
+%
+The number of rows will thus be the number of distinct parallelotopes contained in the bundle $p = |\ptopeset{Q}|$.
 %
@@
 %
-% We would like to compute the image of the parallelotope $P$ with respect to the nonlinear transformation $f(x)$.
+% We would like to compute the image of the parallelotope $P$ with respect to the non-linear transformation $f(x)$.
 %
@@
 %
-One can compute an upper-bound to this nonlinear optimization by computing the Bernstein coefficients of $\Lambda_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i)$ and taking the maximum and minimum coefficients as shown in Property~\ref{prop:bern_enclosure}.
+One can compute an upper-bound to this non-linear optimization by computing the Bernstein coefficients of $\Lambda_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i)$ and taking the maximum and minimum coefficients as shown in Property~\ref{prop:bern_enclosure}.
 %
@@
 %
-Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of nonlinear optimization problems of the form presented in Equation~\ref{eq:maxf}.
+Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of non-linear optimization problems of the form presented in Equation~\ref{eq:maxf}.
 
@@
 
-%\input{static_alg_fig}
+\input{static_alg_fig}
 
@@
 \begin{example}
+
 To include an example of a higher-dimensional non-linear system, we introduce the Phosporaley model.
@@
 The discretization step is set to $\Delta =0.01$ and we propagate the reachble set for $S=300$ time steps.
-Additionally, the initial box is set to be $[1.00,1.01]^7$.
+Additionally, the initial box is set to be $[1.00,1.01]^7 \times [-100, 100]$.
 %
-Figure \ref{fig:kaa_phos} depicts the projection of the reachable set on the first three variables $x_1, x_2, x_3$.
-The paralleotopes are the axis-aligned parallotope and another with the following template direction matrix:
-
-\begin{equation}
-  \Lambda = \begin{bmatrix}
-            1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-            0 & 1 & 0 & 0 & 0 & 0 & 0 \\
-            0 & 0 & 1 & 0 & 0 & 0 & 0 \\
-            0 & 0 & 0 & 0 & 1 & 0 & 0 \\
-            0 & 0 & 0 & 0 & 0 & 1 & 0 \\
-            0 & 0 & 0 & 0 & 0 & 0 & 1 \\
-            0 & 0 & 1 & 1 & 0 & 0 & 0 \\
-            \end{bmatrix}
-\end{equation}
-
-\begin{figure}
+\begin{figure}[t!]
 \label{fig:kaa_phos}
@@
 \end{figure}
+%
+Figure \ref{fig:kaa_phos} depicts the projection of the reachable set on the first three variables $x_1, x_2, x_3$. The relevant matrices are defined as:
+
+\begin{equation}
+  \Lambda^Q = \begin{bmatrix}
+            1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+            0 & 1 & 0 & 0 & 0 & 0 & 0 \\
+            0 & 0 & 1 & 0 & 0 & 0 & 0 \\
+            0 & 0 & 0 & 1 & 0 & 0 & 0 \\
+            0 & 0 & 0 & 0 & 1 & 0 & 0 \\
+            0 & 0 & 0 & 0 & 0 & 1 & 0 \\
+            0 & 0 & 0 & 0 & 0 & 0 & 1 \\
+            0 & 0 & 1 & 1 & 0 & 0 & 0 \\
+            \end{bmatrix}, \quad
+  \T^Q = \begin{bmatrix}
+        0 & 1 & 2 & 3 & 4 & 5 & 6 \\
+        0 & 1 & 2 & 4 & 5 & 6 & 7
+        \end{bmatrix}
+\end{equation}
+$\T^Q$ tells us that the bundle consists of the axis-aligned parallelotope (the first row $\T^Q_1$) and another diagonal parallelotope (the second row $\T^Q_2$). 
 \end{example}

### 2022-07-26 thesis/covid19.tex
Grammar, improvements in figures, presentation

@@
 
-\begin{align}
-  \label{eq:covid19_dynamics}
+\begin{center}
+\begin{equation*}
   \begin{split}
@@
   \end{split}
-\end{align}
+\end{equation*}
+\captionof{figure}{Dynamics for the COVID19 Supermodel.}
+\label{fig:covid19_dynamics}
+\end{center}
+
 where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
@@
 The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
+The discretized dynamics are given in Figure \ref{fig:covid19_dynamics}.
 

### 2022-07-26 thesis/formats.tex
First draft of thesis more or less completed
Grammar, improvements in figures, presentation

@@
 %
-for all tempalte direcitons of the bundle $\Lambda^Q_i$. These are all found by a straightforward linear program.
+for all template directions of the bundle $\Lambda^Q_i$.
+We use the support points as a data-driven approach to find the best-fit linear function to use.
+These are all found by a straightforward linear program.
 %
 %In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear transformation.
-%
-We use the support points as a data-driven approach to find the best-fit linear function to use.
 %
@@
 \begin{equation}
+  \label{eq:least_squares}
   \min_{A} \sum_{x_i} ||f(x_i) - Ax_i||^2_2
@@
 %
+% Give a very light sketch of the procedure here.
 
@@
 %
-First, The algorithm makes use of helper function \texttt{hstack}, which attaches two matrices along their rows.
+First, the algorithm makes use of helper function \texttt{hstack}, which attaches two matrices along their rows.
 In other words, \texttt{hstack} can be visualized as two matrices with the same number of columns stacked on top of one another.

### 2022-07-26 thesis/pcalinratio.tex
Grammar, improvements in figures, presentation

\def \ratiovolwidth {0.5}
\def \ratiovolheight {0.4}
\def \ratiovolhorshift {0cm}
\def \ratiovolvertshift {-2cm}

\begin{figure}
  \vspace*{\ratiovolvertshift}
  \hspace*{\ratiovolhorshift}\subfloat[5 Lin]{
  \includegraphics[width=\ratiovolwidth\textwidth, height=\ratiovolheight \textwidth]{figures/PhasePlots/VDP_5Lin_.png}
}
\subfloat[1 PCA 5 Lin]{
\includegraphics[width=\ratiovolwidth\textwidth, height=\ratiovolheight \textwidth]{figures/PhasePlots/Neuron_1PCA5Lin_.png}
}

\hspace*{\ratiovolhorshift}\subfloat[5 PCA]{
  \includegraphics[width=\ratiovolwidth\textwidth, height=\ratiovolheight \textwidth]{figures/PhasePlots/VDP_5PCA_.png}
}
\subfloat[5 PCA 1 Lin]{
\includegraphics[width=\ratiovolwidth\textwidth, height=\ratiovolheight \textwidth]{figures/PhasePlots/Neuron_5PCA1Lin_.png}
}

\hspace*{\ratiovolhorshift}\subfloat[Sapo]{
\includegraphics[width=\ratiovolwidth\textwidth, height=\ratiovolheight \textwidth]{figures/PhasePlots/VDP_Sapo_.png}}
\subfloat[Sapo]{
\includegraphics[width=\ratiovolwidth\textwidth, height=\ratiovolheight \textwidth]{figures/PhasePlots/Neuron_Sapo_.png}
}

\caption{Effect of varying ratio between the number of PCA and Linear Approximation parallelotopes. The Vanderpol (left) and the FitzHugh-Nagumo Neuron (right) phase plots are shown to illustrate differing effects of varying the PCA/LinApp ratio. The initial set for the Vanderpol model is set to $x \in [0,0.05], \, y \in [1.95,2]$, and the initial set Neuron model is set to $x \in [0.9, 1.1], \, y \in [2.4,2.6]$}
\label{fig:PCALinAppRatio}
\end{figure}

### 2022-07-26 thesis/prelim.tex
Grammar, improvements in figures, presentation

@@
 Discretizing Equation~\ref{eqn:sir} according the Euler method yields the dynamics:
-%
-\begin{align} \label{eqn:discrete_sir}
+
+\begin{center}
+\label{fig:discrete_sir}
+  \begin{equation*}
       \begin{split}
@@
       \end{split}
-\end{align}
+  \end{equation*}
+  \captionof{figure}{Discretized Dynamics of SIR model}
+\end{center}
  Here, $\Delta$ is the discretization step and the index $k \in \mathbb{N}$ simply represents the current step.
@@
 %
-Note the non-linear terms $s_ki_k$ which precludes the expression of the dynamics as a linear transformation.
+Note the non-linear terms $s_ki_k$ which preclude the expression of the dynamics as a linear transformation.
 \end{example}
@@
 %
-\noindent We will see in a future section an example of the reachable set of the discretized SIR model presented in Equation~\ref{eqn:discrete_sir}.
+\noindent We will see in a future section an example of the reachable set of the discretized SIR model presented in Figure~\ref{fig:discrete_sir}.
 
@@
 
-There is a simple method to convert from the half-space representation of $P$ to its generator representation:
-
-\begin{tcolorbox}
-\hspace{1.5cm} \textbf{Half-Space Representation to Generator Representation Conversion}
-\begin{enumerate}
-
-  \item Obtain vertex $v_1$ by solving the linear equation $\Lambda x = c_{l}$.
-  \item The $j+1$ vertex is obtained by solving the linear equation $\Lambda x = \mu_{j}$ where $\mu_{j}[i] = c_{l}[i]$ when $i \neq j$ and $\mu_{j}[j] = c_{u}[j]$.
-  \item The anchor $v$ of the parallelotope is the vertex $v_1$ and the generators will be $g_{i} = v_{i+1} - v_{1}$.
-\end{enumerate}
-\end{tcolorbox}
-%
+There is a simple method to convert from the half-space representation of $P$ to its equivalent generator representation:
+%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+\input{ptope_rep_convert_alg}
+%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 \noindent There also exists a procedure to perform the reverse direction, namely to convert from the generator representation to the half-space representation.
 However, we will not need this procedure for this thesis.
-Refer to \cite{dang2014parameter} for a more detailed exposition.
+Refer to \cite{dang2014parameter} for a more detailed exposition of these conversions.
 
@@
 The half-space representation in matrix form is given as follows:
-
+%
 \begin{equation}
@@
 \end{equation}
-
+%
 To compute the generator representation of $P$, we need to compute the \emph{anchor} and the \emph{generators}.

### 2022-07-26 thesis/results.tex
Grammar, improvements in figures, presentation

@@
 
-To estimate volume of reachable sets, we employ two techniques for estimating volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize Scipy's convex hull routine.
+To estimate volume of reachable sets, we employ two techniques for estimating the volume of individual parallelotope bundles.
+For systems of dimension fewer than or equal to three, we utilize Scipy's convex hull routine, .
 %
@@
 %
-To be specific, if \texttt{ApproxVol} is the routine used to approximate the volume of a bundle, then by utilizing the notion of line 22 of Algorithm \ref{algo:new}:
+To be specific, if \texttt{ApproxVol} is the routine used to approximate the volume of a bundle, then by the notation introduced in line 22 of Algorithm \ref{algo:new}:
 %
@@
 
+\def \voltablespace {-0.8cm}
+
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-\begin{table}
-%\hspace{-5em}
+\begin{table}[t!]
 \centering
-\begin{tabular}{|p{1.5cm}|c|p{1.7cm}|c|c|p{5cm}|}
+\begin{tabular}{|>{\centering}m{1.5cm}|>{\centering}c|m{1.6cm}|c|c|>{\centering\arraybackslash}m{5cm}|}
 \hline
-Model & Dimension & Parameters & \# steps & $\Delta$ & \hspace{1.5cm}Initial Box \\
+Model & Dimension & Parameters & \# steps & $\Delta$ & Initial Box \\
 \hline
-Vanderpol & 2 & \quad \quad \; - & 70 steps & 0.08 & $x \in [0,0.1], y \in [1.99,2]$ \\
+Vanderpol & 2 & \centering -- & 70  & 0.08 & {\vspace{\voltablespace}\begin{align*} x &\in [0,0.1] \\ y &\in [1.99,2] \end{align*}\vspace{\voltablespace}} \\
 \hline
-Jet Engine& 2 & \quad \quad \; - & 100 steps & 0.2 & $x \in [0.8,1.2], y \in [0,8,1.2]$ \\
+Jet Engine & 2 & \centering -- & 100 & 0.2 &  {\vspace{\voltablespace}\begin{align*} x &\in [0.8,1.2] \\ y &\in [0,8,1.2] \end{align*}\vspace{\voltablespace}} \\
 \hline
-Neuron \cite{fitzhugh1961impulses}& 2 & \quad \quad \; - & 200 steps & 0.2 & $x \in [0.9,1.1], y \in [2.4,2.6]$ \\
+Neuron & 2 & \centering -- & 200 & 0.2 & {\vspace{\voltablespace}\begin{align*} x &\in [0.9,1.1] \\ y &\in [2.4,2.6] \end{align*}\vspace{\voltablespace}} \\
 \hline
-SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 & $s \in [0.79,0.8], i \in [0.19,0.2], r = 0$ \\
+SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150  & 0.1 & {\vspace{\voltablespace}\begin{align*} & s \in [0.79,0.8] \\ & i \in [0.19,0.2] \\ & r = 0 \end{align*}\vspace{\voltablespace}} \\
 \hline
-Coupled \newline Vanderpol & 4 & \quad \quad \; - & 40 steps & 0.08 & $x1 \in [1.25, 2.25], y1 \in [1.25, 2.25]$ \newline $x2 \in [1.25, 2.25], y2 \in [1.25, 2.25]$ \\
+Coupled \newline Vanderpol & 4 & \centering  -- & 40 & 0.08 & {\vspace{\voltablespace}\begin{align*} &x_1 \in [1.25, 2.25] \\ & y_1 \in [1.25, 2.25] \\ &x_2 \in [1.25, 2.25] \\ & y_2 \in [1.25, 2.25] \end{align*}\vspace{\voltablespace}} \\
 \hline
-COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 & \quad \quad \quad \; \; Stated Below\\
+COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 & 0.08 & Stated in Section \ref{sec:covid} \\
 \hline
@@
 \label{sec:compare}
-%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 
@@
 
-From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approxiation directions which perform well on all benchmarks.
+From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approximation directions which perform well on all benchmarks.
 In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the ratio imparts differing effects.
@@
 
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\input{pcalinratio}
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \input{voltable}
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 

### 2022-07-26 thesis/staticalg.tex
Grammar, improvements in figures, presentation

@@
 
-Given that $P$ is a parallelotope, all the states in $P$ can be expressed as a vector summation of anchor and scaled generators.
+Given that $P$ is a parallelotope, all the states in $P$ can be expressed as a vector summation of an anchor and a convex combination of generators.
 %
@@
 Hence, the third parameter $r$ is set to zero.
-
+  %
   Second, since we only have the axis-aligned parallelotope in our initial bundle, the matrix $\T^Q$ will consist of only one row indicing the axis-aligned directions expressed as distinct rows in $\Lambda^Q$.
-
-  \begin{figure}
-  \label{fig:kaa_sir}
-    \hspace*{-2.3cm}
-    \includegraphics[width=1.3\textwidth]{figures/SIRProj.png}
-    \caption{Projection of Reachable Set of SIR propagated 300 steps in time.}
-  \end{figure}
 \end{example}
@@
 \begin{example}
-
 To include an example of a higher-dimensional non-linear system, we introduce the Phosporaley model.
 The Phosphoraley model describes a certain cellular regulatory system.
-It is captured by seven variables governed by the following discretized dynamics:
-\begin{align}
+It is captured by seven variables governed by the discretized dynamics stated in Figure \ref{fig:phos_dynamics}:
+
+\begin{center}
+  \begin{equation*}
     \begin{split}
@@
     \end{split}
-\end{align}
+  \end{equation*}
+  \label{fig:phos_dynamics}
+  \captionof{figure}{Discretized Dynamics of Phosphorlay Model.}
+\end{center}
+%
 Here, we set the two parameters $\alpha,\beta$ as $\alpha=0.5$ and $\beta=5$.
@@
 Additionally, the initial box is set to be $[1.00,1.01]^7 \times [-100, 100]$.
-%
-\begin{figure}[t!]
-\label{fig:kaa_phos}
-  \hspace*{-2.3cm}
-  \includegraphics[width=1.3\textwidth]{figures/PhosporlayProjOn X_1, X_2, X_3.png}
-  \caption{Projection of Reachable Set of the Phosporaley model propagated 300 steps in time.}
-\end{figure}
-%
-Figure \ref{fig:kaa_phos} depicts the projection of the reachable set on the first three variables $x_1, x_2, x_3$. The relevant matrices are defined as:
+Under these parameters, Figure \ref{fig:kaa_phos} depicts the projection of the reachable set on the first three variables $x_1, x_2, x_3$.
+The relevant matrices are defined as:
 
@@
 \end{example}
+
+\def \reachplotwidth {1.3}
+
+\begin{figure}[h!]
+  \hspace*{-2.4cm}
+  \includegraphics[width=\reachplotwidth\textwidth]{figures/SIRProj.png}
+  \caption{Projection of Reachable Set of SIR propagated 300 steps in time.}
+  \label{fig:kaa_sir}
+\end{figure}
+
+\begin{figure}[h!]
+  \hspace*{-2.4cm}
+  \includegraphics[width=\reachplotwidth\textwidth]{figures/PhosporlayProjOn X_1, X_2, X_3.png}
+  \caption{Projection of Reachable Set of the Phosporaley model propagated 300 steps in time.}
+  \label{fig:kaa_phos}
+\end{figure}

### 2022-07-26 thesis/prelim.tex
Fixed up some algorith2e discrepancies

@@
 \label{def:generator_def}
-The \emph{generator representation}of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$.
+The \emph{generator representation} of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$.
 The vector  $v \in \mathbb{R}^n$ is called the \emph{anchor} and the $g_i \in \mathbb{R}^n$, are called the \emph{generators}.
@@
 %
+\begin{remark}
   \noindent This is esstentially a convex reprsentation of the parallelotope, which shares many similarities to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
@@
   In particular, a parallelotope is a special case of a zonotope where the number of generators is exactly the dimension of the system $n$.
+\end{remark}
+%
+There is a simple method to convert from the half-space representation of $P$ to its equivalent generator representation. The algorithm is sketched in Algorithm \ref{alg:rep_conversion}.
+%
+The first line ascertains the lower-most vertex of the input parallelotope.
+This vertex will be our anchor.
+The first loop starting at line two finds all of the vertices which are incident to the lower-most vertex through an edge.
+Finally, the second loop starting at line eight calculates the generator vectors which will span the paralleltope through a convex combination $\bar{\alpha} \in [0,1]^n$.
+Note that the anchor is simply shifting the  generator vectors $g_i$ from the origin to the point required to properly span $P$.
 
-There is a simple method to convert from the half-space representation of $P$ to its equivalent generator representation:
-%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-\input{ptope_rep_convert_alg}
-%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-\noindent There also exists a procedure to perform the reverse direction, namely to convert from the generator representation to the half-space representation.
+There also exists a procedure to perform the reverse direction, namely to convert from the generator representation to the half-space representation.
 However, we will not need this procedure for this thesis.
@@
 We refer to this affine transformation associated to $P$ as $T_P:[0,1]^n \rightarrow P$ when necessary.
-%
+
+%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+\input{ptope_rep_convert_alg}
+%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+
 \begin{example}

### 2022-07-26 thesis/staticalg.tex
Fixed up some algorith2e discrepancies

@@
 %
-To distinguish between the distinct parallelotopes contained in the bundle, we add a new matrix called $\T^Q \in \mathbb{N}^{p \times n}$ such that $\T^Q_i$ is a vector of row indices of $\Lambda^Q$ which specify the template directions defining parallelotope $P_i \in \ptopeset{Q}$.
+To distinguish between the distinct parallelotopes contained in the bundle, we add a new matrix called the \emph{bundle index matrix}, $\T^Q \in \mathbb{N}^{p \times n}$, such that $\T^Q_i$ is a vector of row indices of $\Lambda^Q$ which specify the template directions defining parallelotope $P_i \in \ptopeset{Q}$.
 %
@@
 
-Another input to the algorithm is the initial set, given as a parallelotope $P_0$. When the initial set is a box, $P_0$ will be defined by the axis-aligned template directions.
+Another input to the algorithm is the initial parallelotope bundle, given as $Q$. When the initial set is a box, $P_0$ will be defined by the axis-aligned template directions.
 
@@
 %
-Since computing the maximum value of $f(x)$ along each template direction on the $Q$ in one-shot is computationally difficult, the algorithm instead computes the maximum value over each of the constituent parallelotopes and uses the minimum of all these maximum values.
+Since computing the maximum value of $f(x)$ along each template direction on the $Q$ in one shot is computationally difficult, the algorithm instead computes the maximum value over each of the constituent parallelotopes and uses the minimum of all these maximum values.
 
@@
 
-Given that $P$ is a parallelotope, all the states in $P$ can be expressed as a vector summation of an anchor and a convex combination of generators.
+By , all the states in $P$ can be expressed as a vector summation of an anchor and a convex combination of generators.
 %
@@
 Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of non-linear optimization problems of the form presented in Equation~\ref{eq:maxf}.
+
+% TODO: Expound more on the menaings of constructing ptopes from the bundle index matrix.
 

### 2022-07-27 thesis/appendix.tex
Added appendix and some tikz figures

\appendix

\chapter{Benchmark Model Definitions}
\label{app:benchmark}

\section{Vanderpol Oscillator}
The Vanderpol Oscillator is a classical two-dimensional non-linear dynamical system governed by the following dynamics:
\begin{equation}
  \begin{cases}
  x' = y \\
  y' = \mu \cdot(1 - x^2)\cdot y - x
  \end{cases}
 \end{equation}

 We set $\mu = 1$ in our experiments. The dynamics were gathered from \cite{aachen}.

\section{Jet Engine}
The Jet Engine model is specifically the Moore-Greitzer model. The dynamics are given as follows:
\begin{equation}
  \begin{cases}
  x' = -y - 1.5\cdot x^2 - 0.5\cdot x^3 - 0.5 \\
  y' = 3\cdot x - y
  \end{cases}
 \end{equation}

\noindent These dynamics were studied in \cite{aylward2008stability}, and the dynamics were gathered from \cite{aachen}.

\section{Coupled Vanderpol}
This model couples two Vanderpol Oscillators together, resulting in an non-linear ODE of four variables
\begin{equation}
  \begin{cases}
  x_1' = y_1\\
  y_1' = (1-x_1^2)\cdot y_1 - x_1 + (x_2 - x_1) \\
  x_2 ' = y_2 \\
  y_2' = (1-x_2^2)\cdot y_2 - x_2 + (x_1 - x_2)
  \end{cases}
 \end{equation}
\noindent This model was investigated in \cite{rand1980bifurcation}, and the dynamics were gathered from \cite{aachen}

\section{Neuron Model}
The FitzHugh-Nagumo Model describes the electrical activity of a neuron. The dynamics are given by the two-dimensional non-linear ODE:
\begin{equation}
  \begin{cases}
  x' = x - x^3 - y + 7/8 \\
  y' = 0.08\cdot(x + 0.7 - 0.8\cdot y)
  \end{cases}
 \end{equation}

\noindent The model was first studiend in \cite{fitzhugh1961impulses} and its reachability analysis using Bernstein expansion is studied in \cite{dang2012reachability}.

### 2022-07-27 thesis/formats.tex
Added appendix and some tikz figures

@@
 %
-That is the support points are the set of points $x_i$ such that:
+That is the support points are the set of points $P_{supp}$ such that:
 \begin{equation}
 \label{eq:support}
-x_i = \max_{x \in Q} \Lambda^Q_i\cdot x
+ P_{supp} = \bigcup_{i=1}^n \{ \max_{x \in Q} \Lambda^Q_i\cdot x, \; \min_{x \in Q} \Lambda^Q_i\cdot x \}
 \end{equation}
@@
 %
-where $||\cdot||_2$ is the standard $2$-norm on $\reals^n$.
+where $||\cdot||_2$ is the standard Euclidean norm on $\reals^n$.
 %
@@
 % Give a very light sketch of the procedure here.
+To compute the images of the trajectory points, the algorithm must first compute $P_{supp}$ and propagate them one step forward by the dynamics $f$. These are the points fed into the PCA procedure.
 
@@
 %
- In our benchmarks, we tune each of the options (PCA / linear approximation as well as lifespan option) to demonstrate that specific parameters generate more accurate reachable sets than those generated by the static algorithm.
+ In our benchmarks, we tune each of the options (PCA / linear approximation as well as lifespan option) to demonstrate that specific parameters generate more accurate reachable sets than those generated by the static algorithm presented in Algorithm \ref{alg:old}.
 

### 2022-07-27 thesis/prelim.tex
Added appendix and some tikz figures

@@
 \begin{align} \label{eqn:sir}
-    \begin{split}
+    \begin{cases}
         s' &= \beta \cdot s i \\
@@
         r' &= \gamma \cdot i
-    \end{split}
+    \end{cases}
 \end{align}
@@
   \begin{equation*}
-      \begin{split}
+      \begin{cases}
           s_{k+1} &= s_k - (\beta \cdot s_k i_k)\cdot\Delta \\
@@
           r_{k+1} &= r_k + (\gamma\cdot i_k)\cdot\Delta
-      \end{split}
+      \end{cases}
   \end{equation*}
@@
 \end{equation}
+%
 \end{definition}
@@
 such that
-\begin{equation}
+\begin{align}
 \label{eq:halfspaceconst}
-P =\{x ~|~ c_l \leq \Lambda x \leq c_u\}
-\end{equation}
+P & =\{x ~|~ c_l \leq \Lambda x \leq c_u\}
+\intertext{In a more explicit form:}
+P & = \bigwedge_{i=1}^n \left[c_l(i) \leq \Lambda_i\cdot x \leq c_u(i) \right]
+\end{align}
 \end{definition}
@@
 \begin{equation}
-  c_l = [1,1]^T, \quad c_u = [2,2]^T
+  c_l = [-1,-1]^T, \quad c_u = [1,1]^T
 \end{equation}
@@
 \begin{bmatrix}
-  1 \\ 1
+  -1 \\ -1
 \end{bmatrix} \leq \begin{bmatrix}
@@
   \end{bmatrix} \leq \begin{bmatrix}
-    2 \\ 2
+    1 \\ 1
   \end{bmatrix}
@@
   \begin{split}
-  1 \leq x \leq 2 \\
-  1 \leq y \leq 2
+  -1 \leq x \leq 1 \\
+  -1 \leq y \leq 1
   \end{split}
 \end{align}
-This is exactly the shifted unitbox $[1,2] \times [1,2]$.
+This is exactly the shifted unitbox $[-1,1] \times [-1,1]$.
 In fact, we can easily generalize this a general $n$-dimensional system by considering the template direction matrix $\Lambda = I_n$ where $I_n$ is the $n \times n$ identity matrix and two offset vectors $c_l, c_u$ of length $n$.
@@
 \begin{align}
-  \Lambda = \begin{bmatrix}
+  \Lambda = \begin{bmatrix*}[r]
             1 & 0 \\
@@
             0 & -1 \\
-            \end{bmatrix} , \quad c = [2,2,-1,-1]^T
+            \end{bmatrix*} , \quad c = [2,2,-1,-1]^T
 \end{align}
@@
 %
-For a visual plot of above axis-aligned parallelotope, see Figure \ref{fig:axis_align_ptope}
+For a visual plot of above axis-aligned parallelotope, see Figure \ref{fig:axis_align_ptope}.
 
-\begin{figure}
+\begin{figure}[h!]
   \centering
-  \includegraphics{example-image-a}
-  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}.}
+  \begin{tikzpicture}
+    \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
+    \fill[blue!40!white, opacity=0.4] (-1,-1) rectangle (1,1);
+    \draw[->, thick](0,1) -- (0,2) node[above] (0,2) {$[0,1]$};
+    \draw[->, thick](1,0) -- (2,0) node[right] (2,0) {$[1,0]$};
+    \draw[->, thick](-1,0) -- (-2,0) node[left] (-2,0) {$[-1,0]$};
+    \draw[->, thick](0,-1) -- (0,-2) node[below] (0,-2) {$[0,-1]$};
+  \end{tikzpicture}
+  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}. The template directions are displayed as normal vectors to the pairs of parallel planes defining the parallelotope.}
   \label{fig:axis_align_ptope}
@@
     -1 \\ -1
-  \end{bmatrix} \leq \begin{bmatrix}
+  \end{bmatrix} \leq \begin{bmatrix*}[r]
       1 & 1 \\ -1 & 1
-    \end{bmatrix}\begin{bmatrix}
+    \end{bmatrix*}\begin{bmatrix}
      x \\ y
     \end{bmatrix} \leq \begin{bmatrix}
-      2 \\ 2
+      1 \\ 1
     \end{bmatrix}
@@
 
-  \begin{figure}
+  \begin{figure}[h!]
     \centering
-    \includegraphics{example-image-b}
-    \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.}
+    \begin{tikzpicture}
+      \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
+      \draw[fill=blue!40!white, opacity=0.4] (1,0) -- (0,1) -- (-1,0) -- (0,-1) -- cycle;
+      \draw[->, thick] (0.5,0.5) -- (1,1) node[above] {$[1,1]$};
+      \draw[->, thick] (-0.5,-0.5) -- (-1,-1) node[below] {$[-1,-1]$};
+      \draw[->, thick] (-0.5,0.5) -- (-1,1) node[above] {$[-1,1]$};
+      \draw[->, thick] (0.5,-0.5) -- (1,-1) node[below] {$[1,-1]$};
+
+    \end{tikzpicture}
+    \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.
+Once again, note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes.}
     \label{fig:diagonal_ptope}
@@
   %
-  In particular, a parallelotope is a special case of a zonotope where the number of generators is exactly the dimension of the system $n$.
+  %In particular, a parallelotope is a special case of a zonotope where the number of generators is exactly the dimension of the system $n$.
+  %
+  The most general definition of so-called \emph{template polyhedra} in $\reals^n$ is a tuple $\langle \Lambda, c \rangle$ such that $\Lambda \in \reals^{m \times n}$ for some $m \in \mathbb{N}$ and $c \in \reals^n$.
+The polyhedron is defined by the conjunction of linear
+ inequalities:
+%
+ \begin{equation}
+  \bigwedge_{i=1}^m \Lambda_i \cdot x \leq c_i
+ \end{equation}
+ %
+It follows that parallelotopes are template polyhedra with $m=2n$ and the constraint that $\Lambda_{i+n} = -\Lambda_{i}$ for $1 \leq i \leq n$.
+Several verification tasks of hybrid automata have been investigated in the reachability literature.
+See \cite{dang2011template, gronski2019template, sankaranarayanan2008symbolic} for on related topics.
 \end{remark}
@@
 \begin{equation}
-  \begin{bmatrix} 0 \\ 0 \end{bmatrix} \leq \begin{bmatrix}  1 & -1 \\ 0 &  1 \end{bmatrix}  \begin{bmatrix} x \\ y \end{bmatrix} \leq \begin{bmatrix} 1 \\ 1 \end{bmatrix}. \label{eq:ptopeexample}
+  \begin{bmatrix} 0 \\ 0 \end{bmatrix} \leq \begin{bmatrix*}[r]  1 & -1 \\ 0 &  1 \end{bmatrix*}  \begin{bmatrix} x \\ y \end{bmatrix} \leq \begin{bmatrix} 1 \\ 1 \end{bmatrix}.
+\label{eq:ptopeexample}
 \end{equation}

### 2022-07-27 thesis/staticalg.tex
Added appendix and some tikz figures

@@
 \begin{example}
-
+Set our space to be $\reals^2$. Consider the parallelotope bundle $Q$ containing three individual parallelotopes $P_0, P_1, P_2$ with the following template direction matrices:
+  \begin{equation*}
+    \Lambda^{P_0}  = \begin{bmatrix}
+                  1 & 0 \\
+                  0 & 1
+                  \end{bmatrix}, \quad
+    \Lambda^{P_1}  = \begin{bmatrix}
+                    1 & 0 \\
+                    1 & 1
+                    \end{bmatrix}, \quad
+    \Lambda^{P_2} = \begin{bmatrix*}[r]
+                    0 & 1 \\
+                   -1 & 1
+                    \end{bmatrix*}
+  \end{equation*}
+  The associated template direction matric and bundle index matrix for $Q$ will then be defined as follows:
+  \begin{equation*}
+    \Lambda^{Q} = \begin{bmatrix*}[r]
+                  1 & 0 \\
+                  0 & 1 \\
+                  1 & 1 \\
+                  -1 & 1
+                  \end{bmatrix*}, \quad
+                  \T^Q = \begin{bmatrix}
+                          0 & 1 \\
+                          0 & 2 \\
+                          1 & 3
+                         \end{bmatrix}
+  \end{equation*}
 \end{example}
@@
 %
+\begin{figure}
+\subfloat[A single ptope]{
+  \begin{tikzpicture}
+    \draw (0,0) -- (-1,2) -- (3,2) -- (4,0) -- cycle;
+    \node at (1.5,1) {$P$};
+    \draw[->, very thick] (5,1) -- (7,1) node[above, midway] {$f$};
+    \draw (10,0) .. controls (7,1) ..(9,2) .. controls (10, 1).. (11,2) .. controls (14,1) .. (14,0) .. controls (11,-1) .. cycle;
+    \node at (11,0.75) {$f(P)$};
+  \end{tikzpicture}
+
+}
+
+\caption{Diagrams for the effect of the non-linear optimization procedure.}
+\end{figure}
+%
 Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of non-linear optimization problems of the form presented in Equation~\ref{eq:maxf}.
@@
   \begin{equation*}
-    \begin{split}
+    \begin{cases}
         x^1_{k+1} &= x^1_k + ( -\alpha \cdot x^1_k + \beta \cdot x^3_k x^4_k)\cdot \Delta \\
@@
         x^7_{k+1} &= x^7_k + ( -\alpha\cdot  x^7_k + \beta \cdot  x^5_k x^6_k)\cdot \Delta
-    \end{split}
+    \end{cases}
   \end{equation*}

### 2022-07-28 thesis/bern_prune_deg_output.tex
Draft of RT reachability written

\begin{verbatim}
            -----------------------------------
            Computing Step 0
            ------------------------------------

            BASE VERTEX: [0.005, 2.0]
            GENERATORS: [[-0.004, 0.0], [0.0, -0.004999999999999893]]

            ------------------------------------
            Stats for Template Direction: [1. 0.]
            ------------------------------------

            Chosen Max Monom Deg: (0, 1),  Chosen Min Monom Deg: (0, 0)

            Actual Max Monom Deg: (0, 0), Actual Min Monom Def: (1, 1)

            Actual Bernstein degree: [1, 1]

            Actual Functional Comp Poly: -0.004*x - 0.000499999999999989*y + 0.205
            Returned set of Maximum Monom from Lookup Table:

            [((0, 1), 0.204500000000000),
             ((1, 0), 0.201000000000000),
             ((1, 1), 0.200500000000000)]

            Returned set of Minimum Monom from Lookup Table:
            [((0, 0), 0.205000000000000)]

            Returned set of ALL Actual Monomials and their ALL Computed Acutal Bernstein Coeff:
            [((0, 0), 0.20500000000000002),
             ((0, 1), 0.20450000000000002),
             ((1, 0), 0.201),
             ((1, 1), 0.2005)]

            Difference between Queried Max Coeff and Actual Max Coeff: -0.000500000000000000

            Difference between Queried Min Coeff and Actual Min Coeff: 0.00450000000000000

            ------------------------------------
            Stats for Template Direction: [0. 1.]
            ------------------------------------

            Chosen Max Monom Deg: (3, 3),  Chosen Min Monom Deg: (0, 0)

            Actual Max Monom Deg: (2, 0), Actual Min Monom Def: (0, 1)

            Actual Bernstein degree: [2, 1]

            Actual Functional Comp Poly: 0.0004*x - 0.00499999999999989*y + 0.1*(1 - 2.5e-5*(1 - 0.8*x)**2)*(2.0 - 0.00499999999999989*y) + 1.9995
            Returned set of Maximum Monom from Lookup Table:

            [((3, 3), 2.18359741950000)]

            Returned set of Minimum Monom from Lookup Table:
            [((0, 0), 2.19949500000000)]

            Returned set of ALL Actual Monomials and their ALL Computed Acutal Bernstein Coeff:
            [((0, 0), 2.199495),
             ((0, 1), 2.1939950125000003),
             ((1, 0), 2.1996990000000003),
             ((1, 1), 2.1941990025000004),
             ((2, 0), 2.1998998000000003),
             ((2, 1), 2.1943998005000007)]

            Difference between Queried Max Coeff and Actual Max Coeff: -0.0163023804999995

            Difference between Queried Min Coeff and Actual Min Coeff: 0.00549998749999991
  \end{verbatim}

### 2022-07-28 thesis/bern_prune_wrapping_error.tex
Draft of RT reachability written

\begin{verbatim}
------------------------------------
Computing Step 3
------------------------------------

------------------------------------
Stats for Template Direction: [1. 0.]
------------------------------------

Chosen Max Monom Deg: (1, 1),  Chosen Min Monom Deg: (0, 0)

Actual Max Monom Deg: (1, 1), Actual Min Monom Deg: (0, 0)

Actual Bernstein degree: [1, 1]

Actual Sympy Poly: 0.007306445824668*x + 0.0111954493962119*y + 0.906973678980699
Returned set of Maximum Monom from lookup_table:

[((1, 1), 0.925475574201579)]

Returned set of Minimum Monom from lookup_table:
[((0, 0), 0.906973678980699)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 0.906973678980699),
 ((0, 1), 0.9181691283769109),
 ((1, 0), 0.914280124805367),
 ((1, 1), 0.9254755742015789)]

Difference between Queried Max Coeff and Actual Max Coeff: 0

Difference between Queried Min Coeff and Actual Min Coeff: 0

------------------------------------
Stats for Template Direction: [0. 1.]
------------------------------------

Chosen Max Monom Deg: (0, 3),  Chosen Min Monom Deg: (0, 0)

Actual Max Monom Deg: (0, 1), Actual Min Monom Def: (2, 0)

Actual Bernstein degree: [2, 1]

Actual Sympy Poly: -0.0007306445824668*x + 0.111954493962119*y + 0.1*(1 - 0.435728408759507*(0.0110687411271527*x + 1)**2)*(0.111954493962119*y + 2.46876406785527) + 2.40275434063576

Returned set of Maximum Monom from lookup_table:

[((1, 2), 2.77693923047705),
 ((2, 1), 2.65709768697245),
 ((3, 1), 2.65546013926846),
 ((0, 2), 2.77860321954750),
 ((2, 2), 2.77526086680021),
 ((3, 2), 2.77356812851698),
 ((1, 3), 2.89515700322449),
 ((1, 1), 2.65872145772961),
 ((0, 3), 2.89687498755508),
 ((2, 0), 2.53893450714470),
 ((3, 0), 2.53735215001995),
 ((2, 3), 2.89342404662796),
 ((3, 3), 2.89167611776549)]

Returned set of Minimum Monom from lookup_table:
[((0, 0), 2.54205968353236)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 2.542059683532361),
 ((0, 1), 2.6603314515399323),
 ((1, 0), 2.5405036849821685),
 ((1, 1), 2.658721457729609),
 ((2, 0), 2.5389345071447),
 ((2, 1), 2.657097686972454)]

Difference between Queried Max Coeff and Actual Max Coeff: 0.236543536015143

Difference between Queried Min Coeff and Actual Min Coeff: 0.00312517638766119

------------------------------------
Computing Step 4
------------------------------------

------------------------------------
Stats for Template Direction: [1. 0.]
------------------------------------

Chosen Max Monom Deg: (1, 1),  Chosen Min Monom Deg: (0, 0)\ n
Actual Max Monom Deg: (0, 0), Actual Min Monom Def: (1, 1)

Actual Bernstein degree: [1, 1]

Actual Sympy Poly: -0.0185018952208799*x - 0.0354815304022714*y + 1.21516307295709
Returned set of Maximum Monom from lookup_table:

[((1, 1), 1.16117964733394)]

Returned set of Minimum Monom from lookup_table:
[((0, 0), 1.21516307295709)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 1.2151630729570864),
 ((0, 1), 1.1796815425548148),
 ((1, 0), 1.1966611777362064),
 ((1, 1), 1.161179647333935)]

Difference between Queried Max Coeff and Actual Max Coeff: -0.0539834256231513

Difference between Queried Min Coeff and Actual Min Coeff: 0.0539834256231513

------------------------------------
Stats for Template Direction: [0. 1.]
------------------------------------

Chosen Max Monom Deg: (3, 0),  Chosen Min Monom Deg: (0, 3)\ n
Actual Max Monom Deg: (2, 0), Actual Min Monom Def: (0, 1)

Actual Bernstein degree: [2, 1]

Actual Sympy Poly: 0.00185018952208799*x - 0.354815304022714*y + 0.1*(1 - 0.856505038443742*(1 - 0.0199917704331005*x)**2)*(2.89687498755508 - 0.354815304022714*y) + 2.80432743013492

Returned set of Maximum Monom from lookup_table:

[((0, 1), 2.48598940176842),
 ((1, 2), 2.13075300161303),
 ((2, 1), 2.49645813555339),
 ((3, 1), 2.50156197271691),
 ((0, 2), 2.12608267690468),
 ((2, 2), 2.13534845254406),
 ((1, 0), 2.85178155552809),
 ((3, 2), 2.13986902969777),
 ((1, 3), 1.77023872465550),
 ((1, 1), 2.49126727857056),
 ((0, 3), 1.76617595204093),
 ((2, 0), 2.85756781856273),
 ((3, 0), 2.86325491573606),
 ((2, 3), 1.77423876953473),
 ((3, 3), 1.77817608667862)]

Returned set of Minimum Monom from lookup_table:
[((0, 1), 2.48598940176842),
 ((3, 1), 2.50156197271691),
 ((2, 2), 2.13534845254406),
 ((1, 0), 2.85178155552809),
 ((3, 2), 2.13986902969777),
 ((1, 3), 1.77023872465550),
 ((0, 0), 2.84589612663217),
 ((0, 3), 1.76617595204093),
 ((3, 0), 2.86325491573606),
 ((2, 3), 1.77423876953473),
 ((3, 3), 1.77817608667862)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 2.8458961266321676),
 ((0, 1), 2.485989401768422),
 ((1, 0), 2.8517815555280945),
 ((1, 1), 2.491267278570564),
 ((2, 0), 2.8575678185627256),
 ((2, 1), 2.496458135553395)]

Difference between Queried Max Coeff and Actual Max Coeff: 0.00568709717333471

Difference between Queried Min Coeff and Actual Min Coeff: -0.719813449727491
\end{verbatim}

### 2022-07-28 thesis/realtime.tex
Draft of RT reachability written

\newcommand{\barb}{\overline{b}}

\newcommand{\ComputeBern}{\texttt{ComputeBernsteinCoeffIntervals}}
\newcommand{\FindOverlaps}{\texttt{FindOverlapsWithGrid}}
\newcommand{\ComputeGridRep}{\texttt{ComputeGridGeneratorRep}}

\chapter{Applications to Real-time Reachability}
%
In this chapter, we briefly present some experimental results on an application of the Bernstein expansion on the realm of real-time reachability. First, we outline a few motivating factors as well as our proposed reachability algorithm. Second, we expound on some observations and negative experimental results as well as some early obstacles in the efficacy of our proposed algorithm.
%
\section{Motivation and Algorithm}
The context of these experiments is focused towards the perspectives of real-time reachability. Roughly speaking, the objective of an effective real-time reachability algorithm can be described as covering two criteria:
%
\begin{enumerate}
  \item The computed reachable set must be within some $\epsilon$ distance of the exact reachable set. We can think of this as the constraint dictating that the computed reachable set lies within an $\epsilon$ bloat of the exact reachable set (i.e if $\Theta$ denotes the exact reachable set and $\overline{\Theta}$ its computed over-approximation, then
    $ \overline{\Theta} \subset B_\epsilon(\Theta)$
    where
    $B_\epsilon(\Theta) = \bigcup_{x \in \Theta} B_{\epsilon}(x)$.)
  \item The running time to compute the over-approximation $\overline{\Theta}$ must be strictly upper-bounded by some ``reasonable" time period. Here, by reasonable, we generally speak of a time constraint imposed on real-time systems where predictable running is required for proper functionality.
\end{enumerate}

Recall from Sections \ref{sec:bernstein} and \ref{sec:static} that computing the Bernstein coefficients and taking their maximum and minimum coefficients is tantamount to bounding the solutions to non-linear optimization problems of the form presented in Equation \ref{eq:maxalpha}. As the degrees of the polynomial predicates and dimensions of the system grow, the number of Bernstein coefficients also grows in exponential order. This is evident from the Bernstein coefficient formula defined in Equation \ref{eq:bernstein_coeff}.

To mitigate this explosive growth in coefficients, we can pose the following question: Is there a method for ascertaining a small subset of the basis monomials such that the correct maximum and minimum coefficients can be computed from this subset rather the entire set of Bernstein basis monomials?
%
Towards this end, we propose the following algorithm outlined in Algorithm \ref{alg:bern_prune} and elaborate on individual subroutines. In the context of these experiments, we assume the dynamics $f$ are polynomial. This is to ensure we have a Bernstein expansion of the composed dynamics $f \circ T_p$.

\input{bern_prune_alg}

\begin{description}
\item[Parition Domain into Grid:] We first partition the domain into grid cells of equal dimension. Naturally the entire domain is infinite, so we must restrict the domain into bounded box of sufficiently large dimensions. This box must bound the reachable set computed for a finite number of steps forward. These grid cells will act as constraints to feed into our generator representation.
\item[Initialize Generator Representation:] For each of the cells computed in the previous step, we set the anchor vertex and generator vector components to their corresponding grid cell dimension. In other words, we constrain the parallelotope to be contained within the cell through the generator representation. Formally speaking, suppose we have a grid cell $G$ in $n$-dimensional real vector space $\reals^n$ of the form $G = [l_1,u_1] \times [l_2, u_2] \times \cdots [l_n \times u_n]$. To constrain our parallelotope $P = \langle a, g_1, \cdots, g_n\rangle$ to be contained within $g$, it suffices to set the component variables such that $a, g_1, \cdots, g_n \in G$. For example, to set $a \in G$, the constraints should be initialized as below:
\begin{align}
\label{eq:gen_var_interval}
  a_1 \gets [l_1,u_1] & & a_2 \gets [l_2, u_2] & & \cdots & & a_n \gets [l_n,u_n]
\end{align}
%
%\begin{align*}
%  a_1 & \in [l_1,u_1] & a_1 & \in [l_2, u_2] & \cdots & & a_n \in [l_n,u_n] \\
%  g_{11} & \in [l_1,u_1] & g_{12} & \in [l_2, u_2] & \cdots&   & g_{1n} \in [l_n,u_n] \\
%  & \vdots & & \vdots &
%\end{align*}
%
In Algorithm \ref{alg:bern_prune}, the procedure responsible returning this assignment is $\ComputeGridRep$.

\item[Perform the Functional Composition:] As stated in Equation \ref{eq:maxalpha}, the non-linear optimization predicate should be $\Lambda_i\cdot f$ for fixed template direction $\Lambda_i$. Under the assignments of the variables according to Equation \ref{eq:gen_var_interval} and interval arithmetic, we have the polynomial $\Lambda_i\cdot(f \circ T_p)$ where the coefficients are \emph{intervals} rather than real values.

\item[Compute Bernstein Coefficient Intervals:] Now with our computed polynomial with interval coefficients, we can compute the Bernstein coefficient intervals using the formula displayed in Equation \ref{eq:bernstein_coeff}. By our initialization scheme above, this computation will yield intervals such that the Bernstein coefficients computed from any parallelotope contained in grid cell $P \subseteq G$ should lie within their respective coefficient intervals. Let $\{\barb_{\boldi}\}_{\boldi \in I_{G,\Lambda_i}}$ be the set of computed coefficient intervals where $I_{G,\Lambda_i}$ is an index of degrees which depend on the grid cell $G$ and template direction $\Lambda_i$. For the sake of clarity, we shall just denote this index set as $I$.

\item[Find the Maximum/Minimum Interval:] Using $\{\barb_{\boldi}\}$, we find the maximum and minimum interval by simply finding the interval which has the largest upper-bound and the interval which has the smallest lower-bound. If $\barb_{\boldi} = [l_\boldi, u_\boldi]$, then
%
\begin{align}
  \mathfrak{I}_{G,\Lambda_i, \max} & = \argmax_{\boldi \in I} u_\boldi \\
  \mathfrak{I}_{G,\Lambda_i, \min} & = \argmin_{\boldi \in I} u_\boldi
\end{align}
%
Note that $\mathfrak{I}_{\max}, \mathfrak{I}_{\min}$ are the degrees of the Bernstein basis monomials whose intervals dominate the others. Additionally, we add all the intervals which overlap with the maximum and minimum intervals:
%
\begin{align}
  \begin{split}
      m_{G,\Lambda_i,\max} & = \{\boldi \in I ~|~ \barb_{\boldi} \cap \barb_{\mathfrak{I}_{\max}} \neq \emptyset \}\\
      m_{G,\Lambda_i,\min} & = \{\boldi \in I ~|~ \barb_{\boldi} \cap \barb_{\mathfrak{I}_{\min}} \neq \emptyset \}
  \end{split}
\end{align}

\item[Store into Lookup Table:] The maximum and minimum intervals and their associated degrees are stored into a lookup table indiced by each grid cell and template direction $\Lambda_i$ in $\Lambda$.
\end{description}

The required modifications to the reachability algorithm now turn out to be simple. Instead of computing all of the Bernstein coefficients, the algorithm first determines the grid cells with non-empty overlap with the parallelotope in question ($\FindOverlaps$). It then queries the monomials stored for each overlapping grid cell $G$: $ m_{G,\Lambda_i\max}, m_{G,\Lambda_i \min}$ in respect to the template direction $\Lambda_i$. This retrival is executed on lines 6,7 of Algorithm \ref{alg:bern_prune}. The rest of the reachability algorithm follows exactly as the logic of Algorithm \ref{alg:old}.

\section{Experimental Results}
We move onwards to the experimental results and observations. Our experiments were limited to the Vanderpol Model (Appendix Section \ref{sec:vdp}) and the SIR model (Example \ref{ex:sir}). A few parameters are required to be stated for posterity:

\begin{itemize}
  \item For both models, we attempted to run the modified reachability algorithm for seven steps.
  \item For both models, we partition the domain and fix the initial set for each model according to the below inputs:

  \begin{tabular}{|c|c|c|c|}
  \hline
  Model & Domain & Initial Set & \# of Grid Cells \\
  \hline
  Vanderpol & $[-3, 3] \times [-3, 3]$ & $[0.001, 0.005]\times [1.995, 2]$ & 1600 \\
  \hline
  SIR & $[0, 1]^3$ & $[0.799,0.8] \times [0.199,0.2] \times 0$ & 100 \\
  \hline
  \end{tabular}
\end{itemize}

\subsection{Discrepancy of Degrees}
This issue arises when we compute the actual maximum and minimum intervals during the pre-computation step. We will demonstrate this phenomenon with the Vanderpol System:

Recall that the generator representation for ptope $P$ would yield the linear transformation $T:[0,1] \rightarrow P$ which looks like $$ T(x,y) = q + g_0*x + g_1*y$$ for an anchor (base) vertex $q$ and two generator vectors $g_0,g_1$.
%
The functional composition with the generator representation would be substituting the symbolic expressions as follows:
%
\begin{align} x & \leftarrow q_0 + g_{00}*x + g_{01}*y \\  y & \leftarrow q_1 + g_{01}*x + g_{11}*y\end{align}
%
The composed polynomial $f \circ T$ will look as follows:
%
\begin{align} (f \circ T)_0 & = (q_1 + g_{01}*x + g_{11}*y \\ (f \circ T)_1 & = (1 - (q_0 + g_{00}*x + g_{01}*y)^2) * (q_1 + g_{01}*x + g_{11}*y) - (q_0 + g_{00}*x + g_{01}*y) \end{align}

If you inspect the polynomial $(f \circ T)_1$, the square times another linear factor of $x,y$ gives us a polynomial of degree $(3,3)$. If calculate the non-linear optimization objective function as: $\Lambda \cdot (f \circ T_P)$ where we set $\Lambda = [0,1]^T$ (i.e the y-axis aligned vector), then the polynomial will be precisely be $(f \circ T)_1$, which has degree $(3,3)$.

However, the following occurs during the reachability algorithm: the actual polynomial computed (that is the polynomial computed naturally during the parallelotope reachability algorithm) actually has lower degree than $(3,3)$. Consider the output below computed with just the axis-aligned parallelotope over the VanderPol model with a grid parition of $40 \times 40$:

%-------------------------
%\input{bern_prune_deg_output}
%-------------------------

We could execute an additional pruning method where we simply discard all of the irrelevant monomial degrees that do not show up during the reachability computation. For instance, we can simply remove all the monomial terms if the term has higher degree than $(2,1)$.

However, we run into a subsequent problem of the following sort: if the grid is fine enough to extract a single dominant maximum or minimum coefficient interval and its corresponding monomial, then this may show up as a degree larger than the relevant total degree. In the first above, the maximum degree is a lone monomial:
%
\begin{verbatim}
      Returned set of Maximum Monom from Lookup Table:

      [((3, 3), 2.18359741950000)]
\end{verbatim}
%
This means that the actual monomial with the maximum Bernstein coefficient:
\begin{verbatim}
        Actual Max Monom Deg: (2, 0)
\end{verbatim}
%
was actually pruned out during the pre-computation phase.
%
What seems to happen quite frequently is that the maximum or minimum intervals and their associated overlaps domaintes the actual degree of the ``true" polynomial computed after functional composition during the reachability computation. For instance,

This poses the issue of having to know what the degrees supplied to the algorithm are before it even begins. In other words, we don't exactly know how the functional composition with the generator representation affects the degree of the resulting polynomial. If the degree is actually smaller than that of the monomial with the maximum/minimum coefficient interval (or its overlaps), then it will never show up during our reachability algorithm. This results in either a gross conservative error or an exception indicting that we pruned out too many monomials.
%
This error becomes further compounded if higher degree terms are wiped out by the 0 factors in the computed generator or if the wrapping error becomes worse.

One immediate thought could be to add templates which have non-axis-aligned template directions like $[-1,1]^T$. However, due to the ways higher degree terms can cancel each other out when simplifying after the functional composition, the degree of the polynomial after the composition could still be smaller than that of the maximum and minimum monomial and its overlaps.

To summarize, it appears that pre-computing the Bernstein coefficients and the subsequent analysis on their outputted intervals can overly skew the effect of ``non-relevant coefficients". These coefficients can only be pruned once the generator representation of the specific parallelotope is computed and the exact degree of the composed polynomial is determined.

\subsection{Obstacles to Speedups}
Another issue that seems to occur is that parallelotopes which have great overlap with many grid cells yield the full list of Bernstein coefficients. Hence, there is no speedup gained in these cases. Consider the following output for Steps $3-5$:
%-------------------------
%\input{bern_prune_wrapping_error}
%-------------------------

The main point to take here are the printed values of returned set of monomials from the lookup table:
%
\begin{verbatim}
Returned set of Maximum Monom from lookup_table:

         [((0, 1), 2.48598940176842),
         ((1, 2), 2.13075300161303),
         ((2, 1), 2.49645813555339),
         ((3, 1), 2.50156197271691),
         ((0, 2), 2.12608267690468),
         ((2, 2), 2.13534845254406),
         ((1, 0), 2.85178155552809),
         ((3, 2), 2.13986902969777),
         ((1, 3), 1.77023872465550),
         ((1, 1), 2.49126727857056),
         ((0, 3), 1.76617595204093),
         ((2, 0), 2.85756781856273),
         ((3, 0), 2.86325491573606),
         ((2, 3), 1.77423876953473),
         ((3, 3), 1.77817608667862)]
\end{verbatim}
%
From Step 4 onwards, the Lookup table returns the full list of monomials of degree less than $(3,3)$ i.e all the relevant monomials during the pre-computation stage.
%
We tested this even for reachability computations lasting more than 5 steps. The full set of Bernstein coefficients are always returned after a certain step. This is case it would be around step $4$ or $5$.

As the wrapping error becomes worse, so does the utility of our Lookup table and the monomials it returns. A guess is that as the conservativeness of our reachable becomes greater, it overlaps with more cells. Hence, we need to take into account more and more monomials as the reachability algorithm progresses.

Hence, we speculate that the utility of our pruning method hinges on the careful control of the reachable set error. If the error becomes too great, we end of computing all of the coefficients in the end.

### 2022-07-30 thesis/bern_prune_deg_output.tex
Second draft of thesis finished

@@
+\begin{figure}
 \begin{verbatim}
@@
               Computing Step 0
-            ------------------------------------
+    -----------------------------------
 
     BASE VERTEX: [0.005, 2.0]
-            GENERATORS: [[-0.004, 0.0], [0.0, -0.004999999999999893]]
+    GENERATORS: [[-0.004, 0.0], [0.0, -0.0049999999999999]]
+    ....
+    ....
+    -------------------------------------
+    Stats for Template Direction: [0. 1.]
+    -------------------------------------
 
-            ------------------------------------
-            Stats for Template Direction: [1. 0.]
-            ------------------------------------
+    Queried Max Monom Deg: (3, 3)
+    Queried Min Monom Deg: (0, 0)
 
-            Chosen Max Monom Deg: (0, 1),  Chosen Min Monom Deg: (0, 0)
+    Actual Max Monom Deg: (2, 0)
+    Actual Min Monom Def: (0, 1)
 
-            Actual Max Monom Deg: (0, 0), Actual Min Monom Def: (1, 1)
+    Actual Composed Poly:
+    0.0004*x - 0.00499999999999989*y
+    + 0.1*(1 - 2.5e-5*(1 - 0.8*x)**2)*(2.0 - 0.00499999999999989*y)
+    + 1.9995
 
-            Actual Bernstein degree: [1, 1]
+    Total Poly Degree: [2, 1]
 
-            Actual Functional Comp Poly: -0.004*x - 0.000499999999999989*y + 0.205
     Returned set of Maximum Monom from Lookup Table:
-
-            [((0, 1), 0.204500000000000),
-             ((1, 0), 0.201000000000000),
-             ((1, 1), 0.200500000000000)]
-
-            Returned set of Minimum Monom from Lookup Table:
-            [((0, 0), 0.205000000000000)]
-
-            Returned set of ALL Actual Monomials and their ALL Computed Acutal Bernstein Coeff:
-            [((0, 0), 0.20500000000000002),
-             ((0, 1), 0.20450000000000002),
-             ((1, 0), 0.201),
-             ((1, 1), 0.2005)]
-
-            Difference between Queried Max Coeff and Actual Max Coeff: -0.000500000000000000
-
-            Difference between Queried Min Coeff and Actual Min Coeff: 0.00450000000000000
-
-            ------------------------------------
-            Stats for Template Direction: [0. 1.]
-            ------------------------------------
-
-            Chosen Max Monom Deg: (3, 3),  Chosen Min Monom Deg: (0, 0)
-
-            Actual Max Monom Deg: (2, 0), Actual Min Monom Def: (0, 1)
-
-            Actual Bernstein degree: [2, 1]
-
-            Actual Functional Comp Poly: 0.0004*x - 0.00499999999999989*y + 0.1*(1 - 2.5e-5*(1 - 0.8*x)**2)*(2.0 - 0.00499999999999989*y) + 1.9995
-            Returned set of Maximum Monom from Lookup Table:
-
     [((3, 3), 2.18359741950000)]
@@
 
-            Returned set of ALL Actual Monomials and their ALL Computed Acutal Bernstein Coeff:
+    Set of ALL basis degrees and true Bernstein coeffs:
     [((0, 0), 2.199495),
@@
 
-            Difference between Queried Max Coeff and Actual Max Coeff: -0.0163023804999995
+    Difference between Queried Max Coeff and Actual Max Coeff:
+    -0.0163023804999995
 
-            Difference between Queried Min Coeff and Actual Min Coeff: 0.00549998749999991
+    Difference between Queried Min Coeff and Actual Min Coeff:
+    0.00549998749999991
   \end{verbatim}
+  \caption{Output for First step of Reachability algorithm with the Vanderpol Model.}
+  \label{fig:bern_prune_deg_output}
+\end{figure}

### 2022-07-30 thesis/bern_prune_wrapping_error.tex
Second draft of thesis finished

@@
+\begin{figure}
 \begin{verbatim}
 ------------------------------------
-Computing Step 3
+Computing Step 4
 ------------------------------------
-
-------------------------------------
-Stats for Template Direction: [1. 0.]
-------------------------------------
-
-Chosen Max Monom Deg: (1, 1),  Chosen Min Monom Deg: (0, 0)
-
-Actual Max Monom Deg: (1, 1), Actual Min Monom Deg: (0, 0)
-
-Actual Bernstein degree: [1, 1]
-
-Actual Sympy Poly: 0.007306445824668*x + 0.0111954493962119*y + 0.906973678980699
-Returned set of Maximum Monom from lookup_table:
-
-[((1, 1), 0.925475574201579)]
-
-Returned set of Minimum Monom from lookup_table:
-[((0, 0), 0.906973678980699)]
-
-Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
-[((0, 0), 0.906973678980699),
- ((0, 1), 0.9181691283769109),
- ((1, 0), 0.914280124805367),
- ((1, 1), 0.9254755742015789)]
-
-Difference between Queried Max Coeff and Actual Max Coeff: 0
-
-Difference between Queried Min Coeff and Actual Min Coeff: 0
 
@@
 
-Chosen Max Monom Deg: (0, 3),  Chosen Min Monom Deg: (0, 0)
-
-Actual Max Monom Deg: (0, 1), Actual Min Monom Def: (2, 0)
-
-Actual Bernstein degree: [2, 1]
-
-Actual Sympy Poly: -0.0007306445824668*x + 0.111954493962119*y + 0.1*(1 - 0.435728408759507*(0.0110687411271527*x + 1)**2)*(0.111954493962119*y + 2.46876406785527) + 2.40275434063576
-
+Chosen Max Monom Deg: (3, 0)
+Chosen Min Monom Deg: (0, 3)
+...
 Returned set of Maximum Monom from lookup_table:
-
-[((1, 2), 2.77693923047705),
- ((2, 1), 2.65709768697245),
- ((3, 1), 2.65546013926846),
- ((0, 2), 2.77860321954750),
- ((2, 2), 2.77526086680021),
- ((3, 2), 2.77356812851698),
- ((1, 3), 2.89515700322449),
- ((1, 1), 2.65872145772961),
- ((0, 3), 2.89687498755508),
- ((2, 0), 2.53893450714470),
- ((3, 0), 2.53735215001995),
- ((2, 3), 2.89342404662796),
- ((3, 3), 2.89167611776549)]
-
-Returned set of Minimum Monom from lookup_table:
-[((0, 0), 2.54205968353236)]
-
-Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
-[((0, 0), 2.542059683532361),
- ((0, 1), 2.6603314515399323),
- ((1, 0), 2.5405036849821685),
- ((1, 1), 2.658721457729609),
- ((2, 0), 2.5389345071447),
- ((2, 1), 2.657097686972454)]
-
-Difference between Queried Max Coeff and Actual Max Coeff: 0.236543536015143
-
-Difference between Queried Min Coeff and Actual Min Coeff: 0.00312517638766119
-
-------------------------------------
-Computing Step 4
-------------------------------------
-
-------------------------------------
-Stats for Template Direction: [1. 0.]
-------------------------------------
-
-Chosen Max Monom Deg: (1, 1),  Chosen Min Monom Deg: (0, 0)\ n
-Actual Max Monom Deg: (0, 0), Actual Min Monom Def: (1, 1)
-
-Actual Bernstein degree: [1, 1]
-
-Actual Sympy Poly: -0.0185018952208799*x - 0.0354815304022714*y + 1.21516307295709
-Returned set of Maximum Monom from lookup_table:
-
-[((1, 1), 1.16117964733394)]
-
-Returned set of Minimum Monom from lookup_table:
-[((0, 0), 1.21516307295709)]
-
-Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
-[((0, 0), 1.2151630729570864),
- ((0, 1), 1.1796815425548148),
- ((1, 0), 1.1966611777362064),
- ((1, 1), 1.161179647333935)]
-
-Difference between Queried Max Coeff and Actual Max Coeff: -0.0539834256231513
-
-Difference between Queried Min Coeff and Actual Min Coeff: 0.0539834256231513
-
-------------------------------------
-Stats for Template Direction: [0. 1.]
-------------------------------------
-
-Chosen Max Monom Deg: (3, 0),  Chosen Min Monom Deg: (0, 3)\ n
-Actual Max Monom Deg: (2, 0), Actual Min Monom Def: (0, 1)
-
-Actual Bernstein degree: [2, 1]
-
-Actual Sympy Poly: 0.00185018952208799*x - 0.354815304022714*y + 0.1*(1 - 0.856505038443742*(1 - 0.0199917704331005*x)**2)*(2.89687498755508 - 0.354815304022714*y) + 2.80432743013492
-
-Returned set of Maximum Monom from lookup_table:
-
 [((0, 1), 2.48598940176842),
@@
 
-Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
-[((0, 0), 2.8458961266321676),
- ((0, 1), 2.485989401768422),
- ((1, 0), 2.8517815555280945),
- ((1, 1), 2.491267278570564),
- ((2, 0), 2.8575678185627256),
- ((2, 1), 2.496458135553395)]
+Difference between Queried Max Coeff and Actual Max Coeff:
+0.00568709717333471
 
-Difference between Queried Max Coeff and Actual Max Coeff: 0.00568709717333471
+Difference between Queried Min Coeff and Actual Min Coeff: -
+0.719813449727491
+\end{verbatim}
 
-Difference between Queried Min Coeff and Actual Min Coeff: -0.719813449727491
-\end{verbatim}
+\caption{Example of Wrapping Error Obstacle demonstrated for Step 4 for Vanderpol Model.}
+\label{fig:bern_prune_wrapping_output}
+\end{figure}

### 2022-07-30 thesis/conclusion.tex
Second draft of thesis finished

@@
 \label{chap:conclusion}
-In this paper, we investigated two techniques for generating templates dynamically: first using linear approximation of the dynamics, and second using PCA.
+In this thesis, we covered a specific technique to perform reachability analysis over non-linear dynamical systems by leveraging parallelotope bundles and a specific non-linear optimization problem whose solutions determine certain template directions' offsets required to express the reachable set.
+%
+We presented the original algorithm utilizing static template directions.
+%
+We then investigated two techniques for generating templates \emph{dynamically}: first using linear approximation of the dynamics, and second using PCA.
 %
@@
 %
-In future, we intend to investigate Koopman linearization techniques for computing alternative linear approximation template directions \cite{bak2021reachability}.
+Finally, we experimented with potential applications of the Bernstein expansion technique to real-time reachability.
+We found that several hurdles and barriers arise to effectively utilizing the pruning method developed in the pre-computation phase.
 %
-We also wish to investigate the use of a massively-parallel implementation using HPC hardware such as GPUs for optimizing over an extremely large number of parallelotopes and their template directions.
+Several remarks on ideas towards improvement of the methods we developed are listed below:
+\begin{enumerate}
+  \item Koopman linearization techniques for computing alternative linear approximation template directions with alternative optimization methods could yield interesting modifications to Algorithms \ref{alg:old} and \ref{alg:new} \cite{bak2021reachability}.
+
+  \item The use of a massively-parallel implementation using HPC hardware, such as GPUs, for optimizing over an extremely large number of parallelotopes and their template directions is also an immediate extension.
 This is inspired by the approach behind the recent tool \emph{PIRK} \cite{devonport2020pirk}.
+
+  \item Step five of the pre-computation algorithm, where it simply takes the intervals which dominate all the other intervals, may be too crude to properly yield a proper speed-up.
+Are there any other methods of adding basis monomials to the lookup table which take into account the degree discrepancies outlined in Section \ref{sec:deg_disc}?
+\end{enumerate}
+
+%

### 2022-07-30 thesis/prelim.tex
Draft of RT reachability written
Second draft of thesis finished

@@
 \begin{example}
+\label{ex:sir}
 The SIR Epidemic model is a 3-dimensional dynamical system governed by the following continuous dynamics:
@@
 
-%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+%--------------------------------------------------------------------------------------------------------------------------------------------------------- --------
 \subsection{Bernstein Polynomials}
@@
 where $\boldx^\boldi$ represents the monomial $x_1^{i_1}x_2^{i_2}\cdots x_n^{i_n}$.
-Every such real polynomial $p$ of degree at most $\boldd$ can be represented as linear combination of Bernstein basis polynomials of of degree $\boldd$:
+One of the most important properties of Berstein polynomials is that the Bernstein basis polynomials of degree $\boldd$ span the vector space of real multivariate polynomials of degree at most $\boldd$:
+
+\begin{property}
+\label{prop:bern_basis}
+Every real polynomial $p$ of degree at most $\boldd$ can be represented as linear combination of Bernstein basis polynomials of degree $\boldd$:
 %
 \begin{equation}
+  \label{eq:bernstein_coeff}
   p(\boldx) = \sum_{\boldi \leq \boldd} b_{\boldi} \cdot \mathcal{B}_{(\boldi,\boldd)}(\boldx)
@@
 \end{equation}
-%
+\end{property}
+
 \noindent In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
-%
-%\begin{property}
-%\label{prop:basis_property}
-
-%The Bernstein basis polynomials $\$
-%\end{property}
 
@@
 We state this as a property without proof.
-%
+
 \begin{property} (Enclosure Property)
   \label{prop:bern_enclosure}
-  Let $p: \reals^n \rightarrow \reals$ be a real mutlivarite polynomial of degree $\boldd$, and let $p(\boldx) = \sum_{\boldi \leq \boldd} b_\boldi \cdot \mathcal{B}_{(\boldi,\boldd)}(\boldx)$ be the Bernstein expansion of $p$, then
+  Let $p: \reals^n \rightarrow \reals$ be a real multivariate polynomial of degree $\boldd$, and let $p(\boldx) = \sum_{\boldi \leq \boldd} b_\boldi \cdot \mathcal{B}_{(\boldi,\boldd)}(\boldx)$ be the Bernstein expansion of $p$, then
   $$

### 2022-07-30 thesis/realtime.tex
Second draft of thesis finished

@@
 Recall from Sections \ref{sec:bernstein} and \ref{sec:static} that computing the Bernstein coefficients and taking their maximum and minimum coefficients is tantamount to bounding the solutions to non-linear optimization problems of the form presented in Equation \ref{eq:maxalpha}.
-As the degrees of the polynomial predicates and dimensions of the system grow, the number of Bernstein coefficients also grows in exponential order.
+As the degrees of the polynomial objective functions described by Equation \ref{eq:maxalpha} and dimensions of the system grow, the number of Bernstein coefficients also grows in exponential order.
 This is evident from the Bernstein coefficient formula defined in Equation \ref{eq:bernstein_coeff}.
@@
 
-\begin{description}
-\item[Parition Domain into Grid:] We first partition the domain into grid cells of equal dimension.
+\newcommand\litem[1]{\item{\bfseries #1 \\}}
+
+\begin{enumerate}
+\litem{Parition Domain into Grid Cells:} We first partition the domain into grid cells of equal dimension.
 Naturally the entire domain is infinite, so we must restrict the domain into bounded box of sufficiently large dimensions.
@@
 These grid cells will act as constraints to feed into our generator representation.
-\item[Initialize Generator Representation:] For each of the cells computed in the previous step, we set the anchor vertex and generator vector components to their corresponding grid cell dimension.
+
+\litem{Initialize Generator Representation:} For each of the cells computed in the previous step, we set the anchor vertex and generator vector components to their corresponding grid cell dimension.
 In other words, we constrain the parallelotope to be contained within the cell through the generator representation.
@@
 
-\item[Perform the Functional Composition:] As stated in Equation \ref{eq:maxalpha}, the non-linear optimization predicate should be $\Lambda_i\cdot f$ for fixed template direction $\Lambda_i$.
+\litem{Perform the Functional Composition:} As stated in Equation \ref{eq:maxalpha}, the non-linear optimization predicate should be $\Lambda_i\cdot f$ for fixed template direction $\Lambda_i$.
 Under the assignments of the variables according to Equation \ref{eq:gen_var_interval} and interval arithmetic, we have the polynomial $\Lambda_i\cdot(f \circ T_p)$ where the coefficients are \emph{intervals} rather than real values.
 
-\item[Compute Bernstein Coefficient Intervals:] Now with our computed polynomial with interval coefficients, we can compute the Bernstein coefficient intervals using the formula displayed in Equation \ref{eq:bernstein_coeff}.
+\litem{Compute Bernstein Coefficient Intervals:} Now with our computed polynomial with interval coefficients, we can compute the Bernstein coefficient intervals using the formula displayed in Equation \ref{eq:bernstein_coeff}.
 By our initialization scheme above, this computation will yield intervals such that the Bernstein coefficients computed from any parallelotope contained in grid cell $P \subseteq G$ should lie within their respective coefficient intervals.
@@
 
-\item[Find the Maximum/Minimum Interval:] Using $\{\barb_{\boldi}\}$, we find the maximum and minimum interval by simply finding the interval which has the largest upper-bound and the interval which has the smallest lower-bound.
+\litem{Find the Maximum/Minimum Interval:} Using $\{\barb_{\boldi}\}$, we find the maximum and minimum interval by simply finding the interval which has the largest upper-bound and the interval which has the smallest lower-bound.
 If $\barb_{\boldi} = [l_\boldi, u_\boldi]$, then
@@
 
-\item[Store into Lookup Table:] The maximum and minimum intervals and their associated degrees are stored into a lookup table indiced by each grid cell and template direction $\Lambda_i$ in $\Lambda$.
-\end{description}
+\litem{Store into Lookup Table:} The maximum and minimum intervals and their associated degrees are stored into a lookup table indiced by each grid cell and template direction $\Lambda_i$ in $\Lambda$.
+\end{enumerate}
 
@@
   Vanderpol & $[-3, 3] \times [-3, 3]$ & $[0.001, 0.005]\times [1.995, 2]$ & 1600 \\
-  \hline
-  SIR & $[0, 1]^3$ & $[0.799,0.8] \times [0.199,0.2] \times 0$ & 100 \\
+  %\hline
+  %SIR & $[0, 1]^3$ & $[0.799,0.8] \times [0.199,0.2] \times 0$ & 100 \\
   \hline
@@
 
+During the course of experimentation, several observations hinting towards major obstacles hindering any trivial speed-up have been discovered for the two benchmark models.
+We list those obstacles below.
+
 \subsection{Discrepancy of Degrees}
-This issue arises when we compute the actual maximum and minimum intervals during the pre-computation step. We will demonstrate this phenomenon with the Vanderpol System:
+\label{sec:deg_disc}
+This issue arises when we determine the maximum and minimum intervals and their associated monomials during the pre-computation step. We will demonstrate this phenomenon with the Vanderpol system.
 
-Recall that the generator representation for ptope $P$ would yield the linear transformation $T:[0,1] \rightarrow P$ which looks like $$ T(x,y) = q + g_0*x + g_1*y$$ for an anchor (base) vertex $q$ and two generator vectors $g_0,g_1$.
+Recall that the generator representation for a parallelotope $P$ would yield the linear transformation $T_P:[0,1] \rightarrow P$ which looks like $$ T_P(x,y) = q + g_0\cdot x + g_1\cdot y$$ for an anchor (base) vertex $q$ and two generator vectors $g_0,g_1$.
 %
-The functional composition with the generator representation would be substituting the symbolic expressions as follows:
+The composition with the generator representation would be tantamount to substituting the symbolic expressions as follows:
 %
-\begin{align} x & \leftarrow q_0 + g_{00}*x + g_{01}*y \\  y & \leftarrow q_1 + g_{01}*x + g_{11}*y\end{align}
+\begin{align} x & \leftarrow q_0 + g_{00}\cdot x + g_{01}\cdot y \\  y & \leftarrow q_1 + g_{01}\cdot x + g_{11}\cdot y\end{align}
 %
-The composed polynomial $f \circ T$ will look as follows:
+The components of the composed mapping $f \circ T$ will individually look like:
 %
-\begin{align} (f \circ T)_0 & = (q_1 + g_{01}*x + g_{11}*y \\ (f \circ T)_1 & = (1 - (q_0 + g_{00}*x + g_{01}*y)^2) * (q_1 + g_{01}*x + g_{11}*y) - (q_0 + g_{00}*x + g_{01}*y) \end{align}
+\begin{align} (f \circ T)_0 & = q_1 + g_{01}\cdot x + g_{11}\cdot y \\ (f \circ T)_1 & = (1 - (q_0 + g_{00}\cdot x + g_{01}\cdot y)^2) * (q_1 + g_{01}\cdot x + g_{11}\cdot y) - (q_0 + g_{00}\cdot x + g_{01}\cdot y) \end{align}
 
 If you inspect the polynomial $(f \circ T)_1$, the square times another linear factor of $x,y$ gives us a polynomial of degree $(3,3)$.
-If calculate the non-linear optimization objective function as: $\Lambda \cdot (f \circ T_P)$ where we set $\Lambda = [0,1]^T$ (i.e the y-axis aligned vector), then the polynomial will be precisely be $(f \circ T)_1$, which has degree $(3,3)$.
+%
+Suppose we calculate the non-linear optimization objective function as: $\Lambda_i \cdot (f \circ T_P)$ where we set $\Lambda_i = [0,1]^T$ (i.e the y-axis aligned vector), then the polynomial will be precisely be $(f \circ T)_1$, which has degree $(3,3)$.
 
-However, the following occurs during the reachability algorithm: the actual polynomial computed (that is the polynomial computed naturally during the parallelotope reachability algorithm) actually has lower degree than $(3,3)$.
-Consider the output below computed with just the axis-aligned parallelotope over the VanderPol model with a grid parition of $40 \times 40$:
+However, the following occurs during the reachability algorithm: the actual polynomial computed for template direction $\Lambda_i =[0,1]^T$, that is the polynomial computed naturally during the parallelotope reachability algorithm, actually has lower degree than $(3,3)$.
+%
+Furthermore, the maximum and/or minimum degrees pre-computed actually of \emph{higher degree} than the degree of the actual polynomial computed during the procession of the algorithm.
+This would mean that the pre-computed degrees and their coefficient would never show up in the Bernstein expansion of the actual polynomial.
+To see this, recall Property \ref{prop:bern_basis}.
 
+Consider the output presented in Figure \ref{fig:bern_prune_deg_output} computed with just the axis-aligned parallelotope over the VanderPol model with a grid parition of $40 \times 40$.
+We run the algorithm for one step and display the output during the optimization procedure for template direction $\Lambda_i = [0,1]^T$.
 %-------------------------
-%\input{bern_prune_deg_output}
+\input{bern_prune_deg_output}
 %-------------------------
-
-We could execute an additional pruning method where we simply discard all of the irrelevant monomial degrees that do not show up during the reachability computation.
-For instance, we can simply remove all the monomial terms if the term has higher degree than $(2,1)$.
-
-However, we run into a subsequent problem of the following sort: if the grid is fine enough to extract a single dominant maximum or minimum coefficient interval and its corresponding monomial, then this may show up as a degree larger than the relevant total degree.
-In the first above, the maximum degree is a lone monomial:
+In the first step above, the maximum degree is a lone monomial:
 %
@@
 Returned set of Maximum Monom from Lookup Table:
-
 [((3, 3), 2.18359741950000)]
@@
 %
-What seems to happen quite frequently is that the maximum or minimum intervals and their associated overlaps domaintes the actual degree of the ``true" polynomial computed after functional composition during the reachability computation.
-For instance,
+What seems to happen quite frequently is the maximum or minimum intervals and their associated overlaps dominates the degree of the actual polynomial computed after composition during the reachability computation.
+To see this, note the anchor and generator vectors computed for the generator representation:
+%
+\begin{verbatim}
+BASE VERTEX: [0.005, 2.0]
+GENERATORS: [[-0.004, 0.0], [0.0, -0.0049999999999999]]
+\end{verbatim}
+
+The generators are axis-aligned, meaning that several degrees are annihilated during the composition $\Lambda_i\cdot (f\circ T_P)$.
+Hence, during our pre-computation step two, we consider a more general set of monomial degrees which may not reflect the relevant degree statistics found during the actual computation phase.
 
@@
 
-To summarize, it appears that pre-computing the Bernstein coefficients and the subsequent analysis on their outputted intervals can overly skew the effect of ``non-relevant coefficients".
+To recapitulate, it appears that pre-computing the Bernstein coefficients and the subsequent analysis on their outputted intervals can overly skew the effect of ``non-relevant coefficients".
 These coefficients can only be pruned once the generator representation of the specific parallelotope is computed and the exact degree of the composed polynomial is determined.
 
-\subsection{Obstacles to Speedups}
-Another issue that seems to occur is that parallelotopes which have great overlap with many grid cells yield the full list of Bernstein coefficients.
-Hence, there is no speedup gained in these cases.
-Consider the following output for Steps $3-5$:
+\subsection{Contributions of Many Overlapping Cells}
+Another obstacle stems from the observation that parallelotopes which have great overlap with many grid cells yield the full list of Bernstein coefficients.
+Hence, there is no speed-up gained in these cases.
+Consider the output displayed in Figure \ref{fig:bern_prune_wrapping_output} for step four.
 %-------------------------
-%\input{bern_prune_wrapping_error}
+\input{bern_prune_wrapping_error}
 %-------------------------
-
 The main point to take here are the printed values of returned set of monomials from the lookup table:
@@
 Returned set of Maximum Monom from lookup_table:
-
  [((0, 1), 2.48598940176842),
@@
  ((2, 1), 2.49645813555339),
-         ((3, 1), 2.50156197271691),
-         ((0, 2), 2.12608267690468),
-         ((2, 2), 2.13534845254406),
-         ((1, 0), 2.85178155552809),
-         ((3, 2), 2.13986902969777),
-         ((1, 3), 1.77023872465550),
-         ((1, 1), 2.49126727857056),
-         ((0, 3), 1.76617595204093),
-         ((2, 0), 2.85756781856273),
-         ((3, 0), 2.86325491573606),
-         ((2, 3), 1.77423876953473),
+ ...
  ((3, 3), 1.77817608667862)]
@@
 %
-From Step 4 onwards, the Lookup table returns the full list of monomials of degree less than $(3,3)$ i.e all the relevant monomials during the pre-computation stage.
+From Step 4 onwards, the lookup table returns the full list of monomials of degree less than $(3,3)$ i.e all the relevant monomials during the pre-computation stage.
 %
-We tested this even for reachability computations lasting more than 5 steps.
+We tested this even for reachability computations lasting more than five steps.
 The full set of Bernstein coefficients are always returned after a certain step.
-This is case it would be around step $4$ or $5$.
+For the Vanderpol model initialized with our parameters, this point would be around step $4$ or $5$.
 
-As the wrapping error becomes worse, so does the utility of our Lookup table and the monomials it returns.
-A guess is that as the conservativeness of our reachable becomes greater, it overlaps with more cells.
-Hence, we need to take into account more and more monomials as the reachability algorithm progresses.
+As the wrapping error becomes worse, so does the utility of our lookup table.
+A guess is that as the conservativeness of our reachable becomes greater, it overlaps with more cells of the domain.
+This results in an cascading effect where the reachability algorithm is required to take into account more and more monomials as it progresses.
 
-Hence, we speculate that the utility of our pruning method hinges on the careful control of the reachable set error.
-If the error becomes too great, we end of computing all of the coefficients in the end.
+We speculate that the utility of our pruning method hinges on the careful control of the reachable set error.
+If the error becomes too great, we end of computing all of the coefficients, negating any benefits of a speed-up.

### 2022-07-30 thesis/staticalg.tex
Second draft of thesis finished

@@
 %
-Similarly, we compute the lower-bound of $\Lambda_{i}\cdot f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \Lambda_{i}\cdot f(x)$.
+Similarly, we compute the lower-bound of $\Lambda_{i}\cdot f(x)$ for all $x \in P$ by computing the upper-bound of $-1 \times \Lambda_{i}\cdot f(x)$.
 %
+\begin{remark}
+The original algorithm using Bernstein expansion proposed by \cite{dreossi2016parallelotope} assumed polynomial dynamics i.e $f$ is polynomial in the system variables.
+%
+This is to ensure that a $\Lambda_{i}\cdot f(x)$ is a polynomial after the composition shown in Equation \ref{eq:maxalpha}.
+Proper Bernstein expansion of this polynomial allows us to exploit the enclosure property of Bernstein coefficients.
+However, the use of other non-linear optimization solvers, such as Kodiak \cite{kodiak}, allows us to use more general dynamics involving trigonometric and root functions.
+A Taylor expansion of any analytic function can also be truncated to some suffciently large degree to admit a power series expansion.
+\end{remark}
 
@@
 %
-Therefore, the tightest upper bound on $\Lambda^Q_{i}\cdot f(x)$ over $Q$ is the least of the upper bounds computed from each of the parallelotopes.
+Therefore, the tightest upper bound on $\Lambda^Q_{i}\cdot f(x)$ over $Q$ is the least of the upper-bounds computed from each of the parallelotopes.
 %
@@
 %
-\begin{figure}
-\subfloat[A single ptope]{
-  \begin{tikzpicture}
-    \draw (0,0) -- (-1,2) -- (3,2) -- (4,0) -- cycle;
-    \node at (1.5,1) {$P$};
-    \draw[->, very thick] (5,1) -- (7,1) node[above, midway] {$f$};
-    \draw (10,0) .. controls (7,1) ..(9,2) .. controls (10, 1).. (11,2) .. controls (14,1) .. (14,0) .. controls (11,-1) .. cycle;
-    \node at (11,0.75) {$f(P)$};
-  \end{tikzpicture}
-
-}
-
-\caption{Diagrams for the effect of the non-linear optimization procedure.}
-\end{figure}
+%\begin{figure}
+%\subfloat[A single ptope]{
+%  \begin{tikzpicture}
+%    \draw (0,0) -- (-1,2) -- (3,2) -- (4,0) -- cycle;
+%    \node at (1.5,1) {$P$};
+%    \draw[->, very thick] (5,1) -- (7,1) node[above, midway] {$f$};
+%    \draw (10,0) .. controls (7,1) ..(9,2) .. controls (10, 1).. (11,2) .. controls (14,1) .. (14,0) .. controls (11,-1) .. cycle;
+%    \node at (11,0.75) {$f(P)$};
+%  \end{tikzpicture}
+%}
+%\caption{Diagrams for the effect of the non-linear optimization procedure.}
+%\end{figure}
 %
-Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of non-linear optimization problems of the form presented in Equation~\ref{eq:maxf}.
+Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower-bounds for templates directions are obtained by solving a series of non-linear optimization problems of the form presented in Equation~\ref{eq:maxf}.
 

### 2022-08-02 thesis/conclusion.tex
Major grammatical and formatting changes to third draft of thesis

@@
 \label{chap:conclusion}
-In this thesis, we covered a specific technique to perform reachability analysis over non-linear dynamical systems by leveraging parallelotope bundles and a specific non-linear optimization problem whose solutions determine certain template directions' offsets required to express the reachable set.
+In this thesis, we covered a specific technique to perform reachability analysis over non-linear dynamical systems through leveraging parallelotope bundles.
+A particular non-linear optimization problem determines the offset of template directions required to express the reachable set.
 %
@@
 %
-We then investigated two techniques for generating templates \emph{dynamically}: first using linear approximation of the dynamics, and second using PCA.
+We then investigated two techniques for generating templates \emph{dynamically} and \emph{automatically}: the first using linear approximation of the dynamics, and the second using PCA.
 %
-We demonstrated that these techniques improve the accuracy of reachable set by an order of magnitude when compared to static or random template directions.
-%
-We also observed that both these techniques improve the accuracy of the reachable sets for different benchmarks.
+We demonstrated that these techniques improve the accuracy of the reachable set of several benchmarks by an order of magnitude when compared to static or random template directions.
 %
 Finally, we experimented with potential applications of the Bernstein expansion technique to real-time reachability.
-We found that several hurdles and barriers arise to effectively utilizing the pruning method developed in the pre-computation phase.
+We found that several hurdles and barriers arise to effectively utilizing the pruning method developed in the pre-processing phase.
 %
@@
 \begin{enumerate}
-  \item Koopman linearization techniques for computing alternative linear approximation template directions with alternative optimization methods could yield interesting modifications to Algorithms \ref{alg:old} and \ref{alg:new} \cite{bak2021reachability}.
+  \item Koopman linearization techniques for computing alternative linear approximation template directions with other optimization methods could yield interesting modifications to Algorithms \ref{alg:old} and \ref{alg:new} \cite{bak2021reachability}.
 
@@
 
-  \item Step five of the pre-computation algorithm, where it simply takes the intervals which dominate all the other intervals, may be too crude to properly yield a proper speed-up.
+  \item Simply taking the intervals which dominate all the other intervals may be too crude to properly yield an effective speed-up.
 Are there any other methods of adding basis monomials to the lookup table which take into account the degree discrepancies outlined in Section \ref{sec:deg_disc}?

### 2022-08-02 thesis/covid19.tex
Added appendix and some tikz figures
Major grammatical and formatting changes to third draft of thesis

@@
 \begin{equation*}
-  \begin{split}
+  \begin{cases}
    S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
@@
    D' & = D + (\eta I)\cdot \Delta
-  \end{split}
+  \end{cases}
 \end{equation*}
-\captionof{figure}{Dynamics for the COVID19 Supermodel.}
+\captionof{figure}{Dynamics for the Discretized COVID19 Supermodel.}
 \label{fig:covid19_dynamics}
@@
 
-where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
+The system variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
 We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate.
-The parameters are set based on figures shown in \cite{ansumali2020modelling}.
+The parameters are fixed based on figures shown in \cite{ansumali2020modelling}.
 The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$.
@@
 
-Plots of the reachable set for this model tuned to specific values of parameters $\beta, \gamma, \eta$ were published in an ACM Sigbed Blogpost detailing applications of Formal methods for simulating disease dynamics \cite{bak2021covid}.
-%
-The main theme of the blogpost revolved around the difficulty of extracting accurate parameters from real-world data samples.
-%
-Certainly we can estimate the parameters by analyzing real-world data.
-However, minute changes in the parameters could yield estimates which would vastly overestimate or underestimate the true population of infected or asymptomatic patients.
-%
-This error is further compounded as our time horizon increases due to many of the issues pertaining to wrapping error, as discussed in the Introduction of this thesis (see Section \ref{sec:intro}).
-%
-Reachability analysis provides not only a method of simulating the disease dynamics in order to provide illuminating information for policy decisions but also to demonstrate the effect of slight changes in the parameters on the conservativeness of the outputted reachable set.
-%
-See Figure \ref{fig:covid_plots} for the two plots published in the blogpost.
-The Confirmed population is the sum of the number of Symptomatic ($I$) and Asymptomatic ($A$) populations of the dynamics presented in Equation \ref{eq:covid19_dynamics}.
-The plots were created by dividing the total period into two separate periods.
-%
-We decided to separate the periods as the parameters presented in the original paper \cite{ansumali2020modelling} were estimated separately according to these exact periods.
-Furthermore, we wished to control the conservativeness of the over-approximation of the reachable set.
-%
-The red line represents the real data gathered from India during the prescribed time periods with the light-blue region representing the predicted region based on the parameters given in \cite{ansumali2020modelling}.
-%
-Several lines representing trajectories generated according to specific parameters are also plotted in order to convey the effect of changes in the population under slight perturbations of the underlying parameters.
-%
-\begin{figure}
+\begin{figure}[b!]
   \centering
@@
 \end{figure}
+
+Plots of the reachable set for this model tuned to specific values of parameters $\beta, \gamma, \eta$ were published in an ACM Sigbed Blogpost detailing applications of formal methods for simulating disease dynamics \cite{bak2021covid}.
+%
+The main theme revolved around the difficulty of extracting accurate parameters from real-world data samples.
+%
+Certainly we could attempt to estimate the parameters by analyzing real-world data.
+However, minute changes in the parameters could yield estimates which would vastly overestimate or underestimate the true population of infected or asymptomatic patients.
+%
+This error is further compounded as our time horizon increases, due to many issues pertaining to wrapping error as discussed in the introduction of this thesis (see Section \ref{sec:intro}).
+%
+Reachability analysis provides not only a method of simulating the disease dynamics in order to provide illuminating information for policy decisions but also to demonstrate the effect of slight changes in the parameters on the conservativeness of the outputted reachable set.
+%
+See Figure \ref{fig:covid_plots} for the two plots published in the blogpost.
+The Confirmed population is the sum of the number of Symptomatic ($I$) and Asymptomatic ($A$) populations of the dynamics presented in Equation \ref{fig:covid19_dynamics}.
+The plots were created by dividing the total period into two separate periods.
+%
+We decided to separate the periods as the parameters presented in the original paper \cite{ansumali2020modelling} were estimated separately according to these exact periods.
+Furthermore, we wished to control the conservativeness of the over-approximation of the reachable set.
+%
+The red line represents the real data gathered from India during the prescribed time periods with the light blue region representing the predicted region based on the parameters given in \cite{ansumali2020modelling}.
+%
+Several lines representing trajectories generated according to specific parameter values are also plotted in order to convey the effect on the confirmed population under slight perturbations of the underlying parameters.

### 2022-08-02 thesis/formats.tex
Major grammatical and formatting changes to third draft of thesis

@@
 By automatic, we mean that the template directions require no consideration from the user to proceed with the reachable set computation.
-This is in contrast to the original static algorithm treated in Section \ref{sec:static} where the user must input his or her own template directions to specify the parallelotopes before the computation starts.
+This is in contrast to the original static algorithm treated in Section \ref{sec:static} where the user must input his or her own template directions to specify the parallelotopes before the computation begans.
 %
@@
 %
-The second method is based on Principal Compoenent Analysis (PCA) where the algorithm runs PCA on the end points of the sample trajectories.
+The second method is based on Principal Component Analysis (PCA) where the algorithm runs the PCA procedure on the image points of the sample trajectories.
 %
@@
 %
-Intuitively speaking, if time step is discretized to be sufficiently small, propagating trajectories according to the non-linear dynamics $f$ for one step could lead to good lienar approximations of the dynamics within a small region.
+Intuitively speaking, if time step is discretized to be sufficiently small, propagating trajectories according to the non-linear dynamics $f$ for one time step could lead to good lienar approximations of the dynamics within a small region.
 %
@@
 %
-That is the support points are the set of points $P_{supp}$ such that:
+That is the support points are the set of points $P_{supp}$ defined as:
 \begin{equation}
@@
 We use the support points as a data-driven approach to find the best-fit linear function to use.
-These are all found by a straightforward linear program.
+All points are found by a straightforward linear program.
 %
@@
 %
-If the dynamics of a system is linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \T x \leq c_{u}$, is the set $c_{l} \leq \T \cdot A^{-1} x \leq c_{u}$.
+If the dynamics of a system are linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \Lambda x \leq c_{u}$, is the set $c_{l} \leq \Lambda \cdot A^{-1} x \leq c_{u}$.
 %
 Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the non-linear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
+%
+%This will reflect the approximate template directions for one step forward. To continue to
 %
@@
 %
-PCA is a standard technique in Statistical Machine Learning used for reduction of dimensionality by performing Singular Value Decomposition (SVD) on the covariance matrix generated by a set of data points.
+PCA is a standard technique in Statistical Machine Learning used to reduction of dimensionality by performing Singular Value Decomposition (SVD) on the covariance matrix generated by a set of data points.
 %
-Since the covariance matrix is symmetric (i.e $A = A^T$), the eigenvectors of this matrix will always be a orthonormal basis of the system.
+Since the covariance matrix is symmetric by definition (i.e $A = A^T$), the eigenvectors of this matrix will always be a orthonormal basis of the system.
 %
@@
 % Give a very light sketch of the procedure here.
-To compute the images of the trajectory points, the algorithm must first compute $P_{supp}$ and propagate them one step forward by the dynamics $f$. These are the points fed into the PCA procedure.
+To compute the images of the trajectory points, the algorithm must first compute $P_{supp}$ and propagate them one step forward by the dynamics $f$. These image points fed into the PCA procedure.
 
@@
 %
-To mitigate the drawbacks of this discrepancy, we can improve accuracy of this representation by considering more template directions and more parallelotopes.
+To mitigate the drawbacks of this discrepancy, we can improve the accuracy of this representation by considering more template directions and more parallelotopes.
 %
@@
 %
- In our benchmarks, we tune each of the options (PCA / linear approximation as well as lifespan option) to demonstrate that specific parameters generate more accurate reachable sets than those generated by the static algorithm presented in Algorithm \ref{alg:old}.
+During our benchmarks, we tune each of the options (PCA / linear approximation as well as lifespan option) to demonstrate that specific parameters generate more accurate reachable sets than those generated by the static algorithm presented in Algorithm \ref{alg:old}.
 
@@
 %
-During each step, the algorithm computes a collection of template directions from the two techniques outlined in Sections \ref{sec:lin_app} and \ref{sec:pca}.
-Each technique will be encoded as a subroutine labeled \texttt{ApproxLinearTrans} and \texttt{PCA}.
+During each step, the algorithm computes a collection of template directions from the two techniques outlined in the previous two sections.
+The subroutines will be encoded as a subroutine labeled \texttt{ApproxLinearTrans} and \texttt{PCA} respectively.
 %
-The \texttt{ApproxLinearTrans} function computes the best approximation of a linear transformation given a list of points before and after the one-step transformation $f$.
+The \texttt{ApproxLinearTrans} function computes the best linear approximation of the dynamics as specified in Section \ref{sec:lin_app}.
 %
-The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points.
+The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points as specified in Section \ref{sec:pca}.
 %
@@
 %
-Line~\ref{ln:linearapprox} computes the linear approximation of the non-linear dynamics and this linear approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
+Line~\ref{ln:linearapprox} computes the linear approximation of the non-linear dynamics and this approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
 %
-The PCA directions of the images of support points is computed in line~\ref{ln:pca}.
+The PCA directions of the images of support points is computed in Line~\ref{ln:pca}.
 %
@@
 %
-To improve the accuracy of the reachable set, we compute the overapproximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the template lifespan parameter $L$.
+To improve the accuracy of the reachable set, we compute the over-approximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the template lifespan parameter $L$.
 %

### 2022-08-02 thesis/frontmatter/acknowledgement.tex
Major grammatical and formatting changes to third draft of thesis

@@
 
-I would like to thank the UNC Computer Science community for supporting my work and providing me a warm community for pursuing this research.
+I would like to thank the UNC Computer Science community for supporting my work and providing me a warm community to pursue this research.
 In particular, I would like to thank my advisor, Parasara Sridhar Duggirala, for the opporunity to perform research and participate in his group.
@@
 %
-Bineet Ghosh, Manish Goyal, Abel Karimi, and Meghan Stuart were all great colleagues and I enjoyed my time exchanging ideas, proofreading drafts, and socializing over group lunches.
+Bineet Ghosh, Manish Goyal, Abel Karimi, and Meghan Stuart were all great colleagues and I enjoyed my time exchanging ideas, proof-reading drafts, and socializing over group lunches.
 I would also like to thank Stanley Bak for helping me improve the implementation of Kaa and providing valuable feedback during this research.
-Juan Garcia made a wonderful collegue and an equally wonderful friend.
+Juan Garcia is a wonderful collegue and an equally wonderful friend.
 The nights we raided Peabody and Phillips hall was certainly a memorable time.
-Finally, I would like to thank my parents and my family for their ardent, un-ending support.
+Finally, I would like to thank my parents and my family for their ardent, unending support.
 

### 2022-08-02 thesis/intro.tex
Major grammatical and formatting changes to third draft of thesis

--- 2022-07-02 thesis/intro.tex
@@
 \chapter{Introduction}
+\label{sec:intro}
 
-% Introduction from previous works, sources (ARCH, Formats)_
-% Related works from same papers.
-
-One of the most widely-used techniques for performing safety analysis of non-linear dynamical systems is reachable set computation
+One of the most widely-used techniques for performing safety analysis of non-linear dynamical systems is reachable set computation.
 %
-For example, reachability analysis has found many applications in formally verifying the safety properties of Cyber-physical Systems governed by Neural Network Controllers \cite{tran2019star, fan2020reachnn, bak2021nnenum}.
+For instance, reachability analysis has found a panoply of applications in formally verifying the safety properties of Cyber-physical Systems (CPS), such as autonomous vehicles \cite{althoff2010reachability}, F-16 aircraft \cite{heidlauf2018verification}, and CPS systems governed by Neural Network Controllers \cite{tran2019star, fan2020reachnn, bak2021nnenum}.
 %
-The reachable set is defined to be the set of states visited by at least one of the trajectories of the system starting from an initial set and propagated forward in time by a finite fixed number of steps.
+The reachable set is defined to be the set of states visited by at least one of the trajectories of the system starting from an initial set and propagated forward in time by a fixed number of steps.
 %
@@
 %
-First, unlike linear dynamical systems whose solutions can be expressed as closed form, non-linear dynamical systems generally do not admit such a nice form.
+First, unlike linear dynamical systems whose solutions can be expressed in closed form, non-linear dynamical systems generally do not admit such a nice form.
 %
@@
 %
-That is, the overapproximation acquired at a given step would increase the conservativeness of the overapproximation for all future steps.
+That is, the over-approximation acquired at a given step would increase the conservativeness of the over-approximation for all future steps.
 
-One of the several techniques for computing the overapproximation of reachable sets for discrete non-linear systems is to encode the reachable set through parallelotope bundles.
+One of the several techniques for computing the over-approximation of reachable sets for discrete non-linear systems is encoding the reachable set through parallelotope bundles.
 %
@@
 %
-One of the advantages of this technique is its exploitation of a special form of non-linear optimization problem to overapproximate the reachable set.
+One of the advantages of this technique is its exploitation of a special form of non-linear optimization problem to over-approximate the reachable set.
 %
@@
 %
-An immediate reason stems from the responsbility of the practitioner to define the template directions specifiying the parallelotopes.
+An immediate reason stems from the responsibility of the practitioner to define the template directions specifiying the parallelotopes.
 %
-Often, these template directions are selected to be either the cardinal axis directions or some directions from octahedral domains.
+Often times these template directions are selected to be either the cardinal axis directions or some directions from octahedral domains.
 %
@@
 %
-Finally, we experiment with our dynamic template generation algorithm's effectiveness on approximating the reachable set of high-dimensional COVID19 dynamics proposed by the Indian Supermodel Committee \cite{indiansuper2020supermodel}.
+We experiment with our dynamic template generation algorithm's effectiveness on approximating the reachable set of high-dimensional COVID19 dynamics proposed by the Indian Supermodel Committee \cite{indiansuper2020supermodel}.
 The results were published in an ACM blogpost detailing the utility of reachable set computation in modeling disease dynamics \cite{bak2021covid}.
 
+Finally, we investigate an application of Bernstein expansion-based reachability to the real-time domain.
+%
+We attempt to pre-compute the relevant Bernstein coefficients over the entire domain and prune the coefficients which do not appear as either a maximum or minimum coefficient.
+%
+The idea is to decrease the total number of coefficients the reachability algorithm has to compute in order to gain a speedup.
+%
+However, we show that there are several obstacles which hinder the utility of our pre-processing step.
 % [Introduction para]asdf
@@
 
-Reachable set computation of non-linear systems using template polyhedra and Bernstein polynomials has been first proposed in~\cite{dang2009image}.
+Reachable set computation of non-linear systems using template polyhedra and Bernstein polynomials was first proposed in~\cite{dang2009image}.
 %
@@
 %
-Leveraging the principal component analysis of sample trajectories for computing reachable set has been proposed in~\cite{stursberg2003efficient,chen2011choice,seladji2017finding}.
+Leveraging the Principal Component Analysis of sample trajectories for computing reachable set has been proposed in~\cite{stursberg2003efficient,chen2011choice,seladji2017finding}.
 %

### 2022-08-02 thesis/prelim.tex
Major grammatical and formatting changes to third draft of thesis

@@
 
-We begin with some basic definitions pertaining to reachability and parallelotopes.
+We begin with some preliminaries pertaining to reachability and parallelotopes.
 The definition of Bernstein polynomials and their enclosure properties will be stated.
@@
 As stated in the previous section, this thesis pertains to the reachability analysis of dynamical systems.
-Roughly speaking, a dynamical system describes the behavior of states in space which is governed by a set of differential equations.
+Roughly speaking, a dynamical system describes the behavior of states governed by a set of differential equations.
 The states of the system evolve according to the flow of time and the vector field induced by the governing differential equations.
@@
 %
-In the reachability analysis literature, there are several definitions for dynamical systems which appear depending on the authors' taste of formalizations.
+In the reachability analysis literature, there are several definitions for dynamical systems which appear depending on the authors' taste for formalisms.
 Here is a more rigorous definition appearing in \cite{dang1998reachability}:
@@
 A \emph{continuous dynamical system} is a tuple $\langle X, f \rangle$ where $X = \reals^n$ is finite-dimensional Euclidean space and $f$ is a continuous function on $X$.
-A point (state) of $x_0 \in X$ evolves according to a tracjectory $\xi(t): \reals^+ \rightarrow X$ such that the following hold:
+A point (state) of $x_0 \in X$ evolves according to a trajectory $\xi(t): \reals^+ \rightarrow X$ such that the following hold:
 %
@@
 \begin{equation}
-\xi(t) = x_0 + \int_0^t f(\xi(\tau)) d\tau
+\xi(t) = x_0 + \int_0^t f(\xi(\tau))\; d\tau
 \end{equation}
@@
 %
-Intuitively speaking, the function $f$ generally represents a discretized version of some specified continuous dynamical system.
-Here, a discrete-time dynamical system is considered \emph{linear} if its dynamics can be expressed as
+ The function $f$ generally represents a discretized version of some specified continuous dynamical system.
+Here, a discrete-time dynamical system is considered to be \emph{linear} if its dynamics can be expressed as
 
@@
 %
-Note the non-linear terms $s_ki_k$ which preclude the expression of the dynamics as a linear transformation.
+Note the non-linear terms $s_ki_k$, which preclude the expression of the dynamics as a linear transformation.
 \end{example}
@@
 %
-\noindent We will see in a future section an example of the reachable set of the discretized SIR model presented in Figure~\ref{fig:discrete_sir}.
+\noindent Example \ref{ex:sir_kaa} gives the plot of the reachable set of the discretized SIR model presented in Figure~\ref{fig:discrete_sir}.
 
@@
 \label{sec:parallelotope}
+%
 The heart of our reachability algorithm relies on geometric objects called \emph{parallelotopes}. In this section, we define parallelotopes and two representations for them.
@@
 A \emph{parallelotope} $P \subset \mathbb{R}^{n}$ is captured by the tuple $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ is a matrix and $c \in \reals^{2n}$ is a column vector.
-We impose the condition that $\Lambda_{i+n} = -\Lambda_{i}$ for all $i \in \{1, \ldots, n\}$ such that
+We impose the condition that $\Lambda_{i+n} = -\Lambda_{i}$ for all $1 \leq i \leq n$ such that
 %
@@
 If we unpack Equation~\ref{eq:halfspace}, we can re-express the inequalities as a conjunction of half-space constraints.
-If we define $c_{u} = [c(1), c(2), \cdots, c(n)]^T$ and $c_{l} = [c(n+1), c(n+2), \cdots, c(2n)]^T$, then Equation~\ref{eq:halfspace} tells us that:
+If we define $c_{u} = [c(1), c(2), \cdots, c(n)]^T$ and $c_{l} = [c(n+1), c(n+2), \cdots, c(2n)]^T$, then Equation~\ref{eq:halfspace} tells us that, for $1 \leq i \leq n$:
 %
@@
 \intertext{In a more explicit form:}
-P & = \bigwedge_{i=1}^n \left[c_l(i) \leq \Lambda_i\cdot x \leq c_u(i) \right]
+P & = \bigwedge_{i=1}^n \left[c_l(i) \leq \Lambda_i\cdot x \leq c_u(i) \right]\label{eq:halfspace_ineq}
 \end{align}
@@
 %
-\noindent In particular, as a bounded intersection of pairs of parallel half-spaces, it is convex.
+\noindent In particular, as a bounded intersection of $n$ pairs of parallel half-spaces, it is convex.
 %
@@
 \label{ex:simple_ptope}
-Consider the 2D plane, namely $\reals^2$.
-We can construct a couple of simple examples of parallelotopes.
-First, if we define our parallelotope's template direction matrix to be the rows of the matrix:
+Consider the 2D plane, namely $\reals^2$. We can construct a two simple examples of parallelotopes. First, if we define our parallelotope's template directions to be the rows of the matrix:
 %
@@
 %
-Accordingly, the bounded region in space will be the intersection of the following linear constraints:
+Accordingly, by Equation \ref{eq:halfspace_ineq}, the bounded region in space will be the intersection of the following linear constraints:
 %
@@
 \end{align}
-This is exactly the shifted unitbox $[-1,1] \times [-1,1]$.
-In fact, we can easily generalize this a general $n$-dimensional system by considering the template direction matrix $\Lambda = I_n$ where $I_n$ is the $n \times n$ identity matrix and two offset vectors $c_l, c_u$ of length $n$.
+This is exactly the scaled unitbox $[-1,1] \times [-1,1]$.
+In fact, we can easily generalize this to a general $n$-dimensional system by considering the template direction matrix $\Lambda = I_n$ where $I_n$ is the $n \times n$ identity matrix and two offset vectors $c_l, c_u$ of length $n$.
 This would yield the shifted $n$-dimensional unitbox:
@@
     \draw[->, thick](0,-1) -- (0,-2) node[below] (0,-2) {$[0,-1]$};
+    \draw[fill] (0,0) circle [radius=1pt];
+    \node[above] (0,0) {$\mathcal{O}$};
+    %\foreach \i in {-3,-2,...,3} \node[below] at (\i,0) {\scriptsize \i};
+    %\foreach \i in {-3,-2,...,3} \node[right] at (0,\i) {\scriptsize \i};
   \end{tikzpicture}
-  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}. The template directions are displayed as normal vectors to the pairs of parallel planes defining the parallelotope.}
+
+  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}.
+The template directions are displayed as normal vectors to the pairs of parallel planes defining the parallelotope.
+$\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
   \label{fig:axis_align_ptope}
@@
 This would yield the two diagonal directions $[1,1]^T, [-1,1]^T$.
-Suppose we set the upper and lower offsets to be $c_u = [1,1]^T$ and $c_l = [-1,-1]^T$ respectively, then once again by Definition \ref{def:halfspace_def}:
+Suppose we set the upper and lower offsets to be $c_u = [1,1]^T$ and $c_l = [-1,-1]^T$ respectively.
+Then once again by Definition \ref{def:halfspace_def}:
   %
@@
   In general, we define \emph{diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other.
-As a matter of convenience however, we refer to \emph{diagonal parallelotopes} as those defined by a combination of axis-aligned and diagonal directions.
+As a matter of convenience however, we refer to \emph{diagonal parallelotopes} as those defined by a combination of axis-aligned \emph{and} diagonal directions.
 This definition will be useful when we consider parallelotopes defined by unconventional directions (i.e those template directions which are neither axis-aligned nor diagonal).
@@
       \draw[->, thick] (0.5,-0.5) -- (1,-1) node[below] {$[1,-1]$};
-
+      \draw[fill] (0,0) circle [radius=1pt];
+      \node[above] {$\mathcal{O}$};
     \end{tikzpicture}
     \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.
-Once again, note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes.}
+Once again, note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes.
+Once again, $\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
     \label{fig:diagonal_ptope}
@@
 The \emph{generator representation} of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$.
-The vector  $v \in \mathbb{R}^n$ is called the \emph{anchor} and the $g_i \in \mathbb{R}^n$, are called the \emph{generators}.
+The vector  $v$ is called the \emph{anchor} and the $g_i$ are called the \emph{generators}.
 The parallelotope is defined as the set:
@@
 \begin{remark}
-  \noindent This is esstentially a convex reprsentation of the parallelotope, which shares many similarities to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
+  \noindent This is esstentially a convex representation of the parallelotope, which shares many similarities to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
   %
@@
  %
-It follows that parallelotopes are template polyhedra with $m=2n$ and the constraint that $\Lambda_{i+n} = -\Lambda_{i}$ for $1 \leq i \leq n$.
-Several verification tasks of hybrid automata have been investigated in the reachability literature.
-See \cite{dang2011template, gronski2019template, sankaranarayanan2008symbolic} for on related topics.
+It follows that parallelotopes are template polyhedra with $m=2n$ such that $\Lambda_{i+n} = -\Lambda_{i}$ for $1 \leq i \leq n$.
+Several verification tasks of hybrid automata with template polyhedra have been investigated in the reachability literature.
+See \cite{dang2011template, gronski2019template, sankaranarayanan2008symbolic} for more on related topics.
 \end{remark}
@@
 %
-The first line ascertains the lower-most vertex of the input parallelotope.
+Line \ref{ln:find_anchor} ascertains the lower-most corner vertex of the input parallelotope.
 This vertex will be our anchor.
-The first loop starting at line two finds all of the vertices which are incident to the lower-most vertex through an edge.
+The first loop starting at Line \ref{ln:first_loop} finds all of the vertices which are incident to the lower-most vertex through an boundary edge.
+There will be exactly $n$ of these incident vertices.
+The $i^{th}$ incident vertex can be computed by replacing the value stored in index $i$ of the lower offset vector $c_l$ with the value stored in the corresponding index $i$ of upper offset vector $c_u$.
+%
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
+    \draw[fill=blue!40!white, opacity=0.4] (1,0) -- (0,1) -- (-1,0) -- (0,-1) -- cycle;
+    \draw[fill] (0,0) circle [radius=1pt];
+    \node[above] {$\mathcal{O}$};
+
+    \draw[fill] (0,-1) circle [radius=1.5pt];
+    \node[below] at (0,-1) {$v$};
+
+    \draw[->, very thick, draw=red] (0,-1) -- (-1,0) node[midway, left] {$g_1$};
+    \draw[->, very thick, draw=red] (0,-1) -- (1,0) node[midway, right] {$g_2$};
+  \end{tikzpicture}
+  \caption{Plot of the generator representation for Example \ref{ex:simple_diag_ptope}. The vertex $v$ represents the anchor while the two vectors $g_1,g_2$ are the generator vectors.}
+  \label{fig:axis_align_gen}
+\end{figure}
+%
+Intuively, this process can be visualized as following the lower half-space boundary for each pair of parallel planes until we reach the intersection of the boundary with several other upper half-space boundaries.
+%
+See Figure \ref{fig:axis_align_gen} for a plot of this on the diagonal example considered in Example \ref{ex:simple_diag_ptope}.
+%
 Finally, the second loop starting at line eight calculates the generator vectors which will span the paralleltope through a convex combination $\bar{\alpha} \in [0,1]^n$.
@@
 There also exists a procedure to perform the reverse direction, namely to convert from the generator representation to the half-space representation.
-However, we will not need this procedure for this thesis.
+However, we will not require this procedure for this thesis.
 Refer to \cite{dang2014parameter} for a more detailed exposition of these conversions.
@@
 %
-To obtain the anchor, we end up with the trivial solution $x = 1, y=1$ by adhering to Step 1. Hence, the anchor is set to $v = (1,1)$
+To obtain the anchor, we end up with the trivial solution $x = 1, y=1$ by adhering to Step 1. Hence, the anchor is set to $v = (1,1)$.
 %
@@
 %
-%We denote this affine transformation as $T_{p}$.
+The reachable set will be expressed an intersection of parallelotopes. These parallelotopes will be encoded into a parallelotope bundle.
 %
 \begin{definition}
-A parallelotope bundle $Q$ is a set of parallelotopes $\{P_0, \ldots, P_m\}$ such that
+A \emph{parallelotope bundle} $Q$ is a set of parallelotopes $\{P_0, \ldots, P_m\}$ such that
   $$ Q = \bigcap_{i=1}^{m}P_i $$.
@@
 %
-\noindent This parallelotope bundle will be the geometric data structure which will enclose the region we compute to be the approximation of the exact reachable set.
+\noindent This parallelotope bundle will be the geometric data structure enclosing the region we compute to be the over-approximation of the exact reachable set.
 Observe that $Q$ can be expressed as the conjunction of all the linear constraints defining each parallelotope $P_i \in \ptopeset{Q}$.
@@
 \]
-Finally, we generalize the product of binomial coefficients as:
+Finally, we generalize the product of binomial coefficients over multi-indices as:
 \[
@@
 %
-where for $i,d,x \in \reals$:
+where for $i,d \in \mathbb{N}$ and $x \in \reals$:
 %
@@
 %
-Let $p: \reals^n \rightarrow \reals$ of be a real polynomial of degree at most $\boldd$. We can express $p$ as a linear combination of monomials of degree at most $\boldd$:
+Let $p: \reals^n \rightarrow \reals$ be a real polynomial of degree at most $\boldd$. We can express $p$ as a linear combination of monomials of degree at most $\boldd$:
 \[
@@
 %
-The crux of the reachability algorithm involves exploiting this property of Bernstein polynomials to approximate the solution of certain non-linear optimization problem invovilng polynomial predicates over the unitbox, $[0,1]^n$.
+The crux of the reachability algorithm involves exploiting this property of Bernstein polynomials to approximate the solution of certain non-linear optimization problem involving polynomial predicates over the unitbox, $[0,1]^n$.
 %

### 2022-08-02 thesis/realtime.tex
Major grammatical and formatting changes to third draft of thesis

@@
 In this chapter, we briefly present some experimental results on an application of the Bernstein expansion on the realm of real-time reachability.
-First, we outline a few motivating factors as well as our proposed reachability algorithm.
-Second, we expound on some observations and negative experimental results as well as some early obstacles in the efficacy of our proposed algorithm.
+First, we outline a few motivating factors alongside our proposed reachability algorithm.
+Second, we expound on some observations and negative experimental results as well as some early obstacles working against the efficacy of our proposed algorithm.
 %
@@
   \item The computed reachable set must be within some $\epsilon$ distance of the exact reachable set.
-We can think of this as the constraint dictating that the computed reachable set lies within an $\epsilon$ bloat of the exact reachable set (i.e if $\Theta$ denotes the exact reachable set and $\overline{\Theta}$ its computed over-approximation, then
+We can think of this as the constraint dictating that the computed reachable set lies within an $\epsilon$ bloat of the exact reachable set.
+In more formal terms, if $\Theta$ denotes the exact reachable set and $\overline{\Theta}$ its computed over-approximation, then
     $ \overline{\Theta} \subset B_\epsilon(\Theta)$
     where
-    $B_\epsilon(\Theta) = \bigcup_{x \in \Theta} B_{\epsilon}(x)$.)
+    $B_\epsilon(\Theta) = \bigcup_{x \in \Theta} B_{\epsilon}(x)$. Here, $B_\epsilon(x)$ denotes the $\epsilon$-ball around point $x$.
   \item The running time to compute the over-approximation $\overline{\Theta}$ must be strictly upper-bounded by some ``reasonable" time period.
-Here, by reasonable, we generally speak of a time constraint imposed on real-time systems where predictable running is required for proper functionality.
+Here, by reasonable, we generally speak of a time constraint imposed on real-time systems where predictable running is mandatory for proper functionality.
 \end{enumerate}
 
-Recall from Sections \ref{sec:bernstein} and \ref{sec:static} that computing the Bernstein coefficients and taking their maximum and minimum coefficients is tantamount to bounding the solutions to non-linear optimization problems of the form presented in Equation \ref{eq:maxalpha}.
+\noindent Recall from Sections \ref{sec:bernstein} and \ref{sec:static} that computing the Bernstein coefficients and taking their maximum and minimum coefficients is tantamount to bounding the solutions to non-linear optimization problems of the form presented in Equation \ref{eq:maxalpha}.
 As the degrees of the polynomial objective functions described by Equation \ref{eq:maxalpha} and dimensions of the system grow, the number of Bernstein coefficients also grows in exponential order.
@@
 %
-Towards this end, we propose the following algorithm outlined in Algorithm \ref{alg:bern_prune} and elaborate on individual subroutines.
+Towards this end, we propose the following algorithm outlined in Algorithm \ref{alg:bern_prune} and elaborate on the individual subroutines.
 In the context of these experiments, we assume the dynamics $f$ are polynomial.
@@
 \litem{Parition Domain into Grid Cells:} We first partition the domain into grid cells of equal dimension.
-Naturally the entire domain is infinite, so we must restrict the domain into bounded box of sufficiently large dimensions.
+Naturally the entire domain is infinite, so we must restrict the domain into a bounded box of sufficiently large dimension.
 This box must bound the reachable set computed for a finite number of steps forward.
@@
 In other words, we constrain the parallelotope to be contained within the cell through the generator representation.
-Formally speaking, suppose we have a grid cell $G$ in $n$-dimensional real vector space $\reals^n$ of the form $G = [l_1,u_1] \times [l_2, u_2] \times \cdots [l_n \times u_n]$.
+Formally speaking, suppose we have a grid cell $G$ in $n$-dimensional Euclidean space $\reals^n$ of the form $G = [l_1,u_1] \times [l_2, u_2] \times \cdots [l_n \times u_n]$.
 To constrain our parallelotope $P = \langle a, g_1, \cdots, g_n\rangle$ to be contained within $g$, it suffices to set the component variables such that $a, g_1, \cdots, g_n \in G$.
@@
 \label{eq:gen_var_interval}
-  a_1 \gets [l_1,u_1] & & a_2 \gets [l_2, u_2] & & \cdots & & a_n \gets [l_n,u_n]
+  a(i) \gets [l_1,u_1] & & a(2) \gets [l_2, u_2] & & \cdots & & a(n) \gets [l_n,u_n]
 \end{align}
@@
 %
-In Algorithm \ref{alg:bern_prune}, the procedure responsible returning this assignment is $\ComputeGridRep$.
+In Algorithm \ref{alg:bern_prune}, this is executed in Line \ref{ln:compute_gen}.
 
@@
 
-\litem{Find the Maximum/Minimum Interval:} Using $\{\barb_{\boldi}\}$, we find the maximum and minimum interval by simply finding the interval which has the largest upper-bound and the interval which has the smallest lower-bound.
+\litem{Find the Maximum/Minimum Interval:} Using the coeffcient intervals $\{\barb_{\boldi}\}$, we find the maximum and minimum interval by simply plucking the interval which has the largest upper-bound and the interval which has the smallest lower-bound.
 If $\barb_{\boldi} = [l_\boldi, u_\boldi]$, then
@@
 \begin{align}
-  \mathfrak{I}_{G,\Lambda_i, \max} & = \argmax_{\boldi \in I} u_\boldi \\
-  \mathfrak{I}_{G,\Lambda_i, \min} & = \argmin_{\boldi \in I} u_\boldi
+  \mathfrak{I}_{\max} & := \mathfrak{I}_{G,\Lambda_i, \max}  = \argmax_{\boldi \in I} u_\boldi \\
+  \mathfrak{I}_{\min} & := \mathfrak{I}_{G,\Lambda_i, \min} = \argmin_{\boldi \in I} l_\boldi
 \end{align}
@@
 
-\litem{Store into Lookup Table:} The maximum and minimum intervals and their associated degrees are stored into a lookup table indiced by each grid cell and template direction $\Lambda_i$ in $\Lambda$.
+\litem{Store into Lookup Table:} The maximum and minimum intervals and their associated degrees are stored into a lookup table indiced by each grid cell $G$ and template direction $\Lambda_i$ in $\Lambda$.
 \end{enumerate}
@@
 The required modifications to the reachability algorithm now turn out to be simple.
+%
 Instead of computing all of the Bernstein coefficients, the algorithm first determines the grid cells with non-empty overlap with the parallelotope in question ($\FindOverlaps$).
 It then queries the monomials stored for each overlapping grid cell $G$: $ m_{G,\Lambda_i\max}, m_{G,\Lambda_i \min}$ in respect to the template direction $\Lambda_i$.
-This retrival is executed on lines 6,7 of Algorithm \ref{alg:bern_prune}.
+Lines \ref{ln:max_lookup} and \ref{ln:min_lookup} of Algorithm \ref{alg:modified_reach} reflect this operation.
+The retrival itself is called on Lines \ref{ln:retrieve_max} and \ref{ln:retrieve_min}.
 The rest of the reachability algorithm follows exactly as the logic of Algorithm \ref{alg:old}.
@@
 We move onwards to the experimental results and observations.
-Our experiments were limited to the Vanderpol Model (Appendix Section \ref{sec:vdp}) and the SIR model (Example \ref{ex:sir}).
+Our experiments were limited to the Vanderpol Model (Appendix Section \ref{sec:vdp}).
 A few parameters are required to be stated for posterity:
@@
 \begin{itemize}
-  \item For both models, we attempted to run the modified reachability algorithm for seven steps.
-  \item For both models, we partition the domain and fix the initial set for each model according to the below inputs:
+  \item  We attempted to run the modified reachability algorithm for seven steps.
+  \item We partitioned the domain and fixed the initial set for each model according to the below inputs:
 
@@
 
-During the course of experimentation, several observations hinting towards major obstacles hindering any trivial speed-up have been discovered for the two benchmark models.
-We list those obstacles below.
+During the course of experimentation, several observations hinting towards major obstacles hindering any trivial speed-up have been discovered for the Vanderpol model. We list those obstacles below.
 
@@
 \label{sec:deg_disc}
-This issue arises when we determine the maximum and minimum intervals and their associated monomials during the pre-computation step. We will demonstrate this phenomenon with the Vanderpol system.
+This issue arises when we determine the maximum and minimum intervals and their associated monomials during the pre-computation step. We will demonstrate this phenomenon as such:
 
@@
 %
-\begin{align} x & \leftarrow q_0 + g_{00}\cdot x + g_{01}\cdot y \\  y & \leftarrow q_1 + g_{01}\cdot x + g_{11}\cdot y\end{align}
+\begin{align} x & \leftarrow q_0 + g_{00}\cdot x + g_{10}\cdot y \\  y & \leftarrow q_1 + g_{01}\cdot x + g_{11}\cdot y\end{align}
 %
@@
 %
-\begin{align} (f \circ T)_0 & = q_1 + g_{01}\cdot x + g_{11}\cdot y \\ (f \circ T)_1 & = (1 - (q_0 + g_{00}\cdot x + g_{01}\cdot y)^2) * (q_1 + g_{01}\cdot x + g_{11}\cdot y) - (q_0 + g_{00}\cdot x + g_{01}\cdot y) \end{align}
+\begin{align} (f \circ T)_0 & = q_1 + g_{01}\cdot x + g_{11}\cdot y \\ (f \circ T)_1 & = (1 - (q_0 + g_{00}\cdot x + g_{10}\cdot y)^2) * (q_1 + g_{01}\cdot x + g_{11}\cdot y) - (q_0 + g_{00}\cdot x + g_{10}\cdot y) \end{align}
 
@@
 %
-Suppose we calculate the non-linear optimization objective function as: $\Lambda_i \cdot (f \circ T_P)$ where we set $\Lambda_i = [0,1]^T$ (i.e the y-axis aligned vector), then the polynomial will be precisely be $(f \circ T)_1$, which has degree $(3,3)$.
+Suppose we calculate the non-linear optimization objective function as: $\Lambda_i \cdot (f \circ T_P)$ where we set $\Lambda_i = [0,1]^T$ (i.e the $y$-axis aligned vector), then the polynomial will be precisely be $(f \circ T)_1$, which has degree $(3,3)$.
 
-However, the following occurs during the reachability algorithm: the actual polynomial computed for template direction $\Lambda_i =[0,1]^T$, that is the polynomial computed naturally during the parallelotope reachability algorithm, actually has lower degree than $(3,3)$.
+However, the following occurs during the reachability algorithm: the actual polynomial computed for template direction $\Lambda_i =[0,1]^T$, that is the polynomial computed naturally during the parallelotope reachability algorithm, actually has degree lower than $(3,3)$.
 %
-Furthermore, the maximum and/or minimum degrees pre-computed actually of \emph{higher degree} than the degree of the actual polynomial computed during the procession of the algorithm.
-This would mean that the pre-computed degrees and their coefficient would never show up in the Bernstein expansion of the actual polynomial.
+Furthermore, the maximum and/or minimum degrees pre-computed are actually of \emph{higher degree} than the degree of the polynomial computed during the progression of the algorithm.
+This means that the pre-computed degrees and their coefficients will never show up in the Bernstein expansion of the polynomial arising after the composition shown in Equation \ref{eq:maxalpha}.
 To see this, recall Property \ref{prop:bern_basis}.
@@
 %-------------------------
-In the first step above, the maximum degree is a lone monomial:
+During the first step, the maximum degree is a lone monomial:
 %
@@
 %
-was actually pruned out during the pre-computation phase.
+was pruned out during the pre-computation phase.
 %
-What seems to happen quite frequently is the maximum or minimum intervals and their associated overlaps dominates the degree of the actual polynomial computed after composition during the reachability computation.
+What seems to happen quite frequently is the maximum or minimum intervals and their overlaps dominate the degree of the polynomial computed after composition during the reachability computation.
 To see this, note the anchor and generator vectors computed for the generator representation:
@@
 
-The generators are axis-aligned, meaning that several degrees are annihilated during the composition $\Lambda_i\cdot (f\circ T_P)$.
-Hence, during our pre-computation step two, we consider a more general set of monomial degrees which may not reflect the relevant degree statistics found during the actual computation phase.
+The generators are axis-aligned, meaning that several degrees are annihilated during the composition $\Lambda_i\cdot (f\circ T_P)$ due to the zero components.
+%
+Hence, during our pre-processing stage, we consider a more general set of monomial degrees which may not reflect the relevant degree statistics found during the actual computation phase.
 
-This poses the issue of having to know what the degrees supplied to the algorithm are before it even begins.
-In other words, we don't exactly know how the functional composition with the generator representation affects the degree of the resulting polynomial.
-If the degree is actually smaller than that of the monomial with the maximum/minimum coefficient interval (or its overlaps), then it will never show up during our reachability algorithm.
-This results in either a gross conservative error or an exception indicting that we pruned out too many monomials.
+This poses the issue of being privy to the degrees supplied to the algorithm before it even begins.
 %
-This error becomes further compounded if higher degree terms are wiped out by the 0 factors in the computed generator or if the wrapping error becomes worse.
+In other words, we are uncertain about how the interactions of the functional composition with the generator representation affect the degree of the resulting polynomial.
+If the degree is smaller than that of the monomial with the maximum/minimum coefficient interval or its overlaps, then it will appear during our reachability algorithm.
+This untowardly results in either a gross conservative error or an exception indicting that too many monomials were pruned out.
+%
+This error becomes further compounded if higher degree terms are wiped out by the zero factors in the computed generator or if the wrapping error becomes worse.
 
-One immediate thought could be to add templates which have non-axis-aligned template directions like $[-1,1]^T$.
+One immediate thought could be adding templates which have non-axis-aligned template directions like $[-1,1]^T$.
 However, due to the ways higher degree terms can cancel each other out when simplifying after the functional composition, the degree of the polynomial after the composition could still be smaller than that of the maximum and minimum monomial and its overlaps.
@@
 Hence, there is no speed-up gained in these cases.
-Consider the output displayed in Figure \ref{fig:bern_prune_wrapping_output} for step four.
+Consider the output displayed in Figure \ref{fig:bern_prune_wrapping_output} for step $4$ of the reachable set computation.
 %-------------------------
@@
 %-------------------------
-The main point to take here are the printed values of returned set of monomials from the lookup table:
+The main point to take here lies within the printed values of returned set of monomials from the lookup table:
 %
@@
 A guess is that as the conservativeness of our reachable becomes greater, it overlaps with more cells of the domain.
-This results in an cascading effect where the reachability algorithm is required to take into account more and more monomials as it progresses.
+This results in an cascading effect where the reachability algorithm is required to take into account more monomials as it progresses.
 

### 2022-08-02 thesis/results.tex
Added appendix and some tikz figures
Major grammatical and formatting changes to third draft of thesis

@@
   \item[Numpy:] The \emph{Numpy} module is used to do all matrix computations, such as matrix multiplication and matrix inversions.
-It is also used to efficiently solve systems of linear equations, especially those which arise from converting from half-space representation to generator representation (See Section \ref{sec:parallelotope}), and execute the Least Squares routine required to compute the solution to Problem \ref{eq:least_squares}.
+It is also used to efficiently solve systems of linear equations, especially those which arise from converting from half-space representation to generator representation (See Algorithm \ref{alg:rep_conversion}), and execute the Least Squares routine required to compute the solution to Problem \ref{eq:least_squares}.
   %
@@
   \item[Multiprocessing:] The \emph{multiprocessing} module parallelizes all the non-linear optimization procedures required to compute the new upper and lower offsets of all the template directions of the parallelotope bundle.
-This is expressed by lines 13,14 in Algorithm \ref{alg:old}.
+This is expressed by Lines \ref{ln:bern_opt_max}, \ref{ln:bern_opt_min} in Algorithm \ref{alg:old}.
   %
   \item[Swiglpk:] The \emph{swiglpk} module is a simple Python wrapper over the C library, \emph{GLPK} (GNU Linear Programming Kit).
-Kaa uses swiglpk for all LP problems, such as those which arise from computing the support points over a bundle (see Equation \ref{eq:support}).
+Kaa uses swiglpk for all linear programming problems, such as those which arise from computing the support points over a bundle (see Equation \ref{eq:support}).
 \end{description}
@@
 
-To extend Kaa to handle dynamic parallelotope bundles, we replaced the original optimization procedure leveraging Bernstein polynomials (see Section \ref{sec:bernstein}) to \emph{Kodiak}.
+To extend Kaa to handle dynamic parallelotope bundles, we replaced the original optimization procedure leveraging Bernstein polynomials (see Section \ref{sec:bernstein}) to \emph{Kodiak} \cite{kodiak}.
 %
@@
 \begin{enumerate}
-  \item Kodiak is very fast as a Python wrapper over the original C\texttt{++} implementation. It is much faster than Kaa's original procedure of computing all Bernstein coefficients.
+  \item Kodiak is very fast as a Python wrapper over the original C\texttt{++} implementation. It is much faster than Kaa's original procedure of computing all relevant Bernstein coefficients.
   \item Kodiak can handle a wider variety of dynamics, including those which feature trigonometric terms and square root terms.
@@
 To estimate volume of reachable sets, we employ two techniques for estimating the volume of individual parallelotope bundles.
-For systems of dimension fewer than or equal to three, we utilize Scipy's convex hull routine, .
+For systems of dimension fewer than or equal to three, we utilize Scipy's convex hull routine.
 %
@@
 %
-To be specific, if \texttt{ApproxVol} is the routine used to approximate the volume of a bundle, then by the notation introduced in line 22 of Algorithm \ref{algo:new}:
+To be specific, if \texttt{ApproxVol} is the routine used to approximate the volume of a bundle, then by the notation introduced in Line \ref{ln:theta} of Algorithm \ref{alg:new}:
 %
@@
 %
-Recall through the discussion in Example \ref{ex:simple_diag_ptope} that we refer to parallelotopes defined only by axis-aligned and diagonal directions as \emph{diagonal parallelotopes}.
+Recall through the discussion in Example \ref{ex:simple_diag_ptope} that we refer to parallelotopes defined only by axis-aligned and diagonal directions as diagonal parallelotopes.
 %
 Similarly, \emph{diagonal parallelotope bundles} are parallelotope bundles solely consisting of diagonal parallelotopes.
-Sapo primarily utilizes \emph{static diagonal parallelotope bundles} to perform its reachability computation.
+Sapo primarily utilizes static diagonal parallelotope bundles to perform its reachability computation.
 %
@@
 
-Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation. The last seven-dimensional COVID supermodel is explained in the subsequent section below.
-
-%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-
-\input{covid19}
+Table \ref{tab:modeldyns} summarizes five standard benchmarks used for experimentation.
+The last seven-dimensional COVID supermodel is explained in the subsequent section below.
+The remaining models' dynamics can be found in Appendix \ref{app:benchmark}.
 
@@
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-\begin{table}[t!]
+\begin{table}[h!]
 \centering
@@
 %-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\input{covid19}
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 \section{Comparison of Template Generation Techniques}
@@
 %
-A cursory analysis shows that the number of possible templates with diagonal directions grows with $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is infeasible.
+A cursory analysis shows that the number of possible templates with diagonal directions grows order $O(n^n)$ with the number of dimensions and hence an exhaustive search on optimal template directions is infeasible.
 
@@
 
-From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approximation directions which perform well on all benchmarks.
+From our experiments, we conclude there is no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approximation directions which perform well on every single benchmarks.
 In Figure \ref{fig:PCALinAppRatio}, we demonstrate two cases where varying the ratio imparts differing effects.

### 2022-08-02 thesis/staticalg.tex
Major grammatical and formatting changes to third draft of thesis

@@
 We will end with an outline of the static algorithm first investigated in works \cite{dang2012reachability, dreossi2016parallelotope}.
-As mentioned in the previous section, the building block of the reachability algorithm relies on approximate solutions to a non-linear optimization problem over the unitbox domain.
+As mentioned in the previous section, the building block of the reachability algorithm relies on approximate solutions to a non-linear optimization problem over the unit-box domain.
 Consider a non-linear function $h : \reals^n \rightarrow \reals$.
@@
 \begin{remark}
-If none of the paralleotopes $P_i \in \ptopeset{Q}$ share common template directions, then $\Lambda^Q$ will simply be the template direction matrices $\{\Lambda^P\}_{P \in \ptopeset{Q}}$ concatented along their rows.
+If none of the paralleotopes $P_i \in \ptopeset{Q}$ share common template directions, then $\Lambda^Q$ will simply be the template direction matrices $\{\Lambda^P\}_{P \in \ptopeset{Q}}$ concatenated along their rows.
 This will generally be the matrix generated by the dynamic algorithm we will outline in a future section.
@@
   \end{equation*}
-  The associated template direction matric and bundle index matrix for $Q$ will then be defined as follows:
+  The associated template direction matrix and bundle index matrix for $Q$ will then be defined as follows:
   \begin{equation*}
@@
 
-Another input to the algorithm is the initial parallelotope bundle, given as $Q$. When the initial set is a box, $P_0$ will be defined by the axis-aligned template directions.
+\noindent Another input to the algorithm is the initial parallelotope bundle, given as $Q$. When the initial set is a box, $P_0$ will be defined by the axis-aligned template directions.
 
-The output of the algorithm is, for each step $k$, the set $\overline\Theta_k$, which is an overapproximation of the reachable set at step $k$, $\Theta_k \subseteq \overline\Theta_k$.
-The total overapproximation of the reachable set for a finite number of steps $n$ will be $\Theta = \cup_{k=1}^n \Theta_k$.
+The output of the algorithm is, for each step $k$, the set $\overline\Theta_k$, which is an over-approximation of the reachable set at step $k$, $\Theta_k \subseteq \overline\Theta_k$.
+The total over-approximation of the reachable set for a finite number of steps $n$ will be $\overline\Theta = \cup_{k=1}^n \overline\Theta_k$.
 %
 The high-level pseudo-code is written in Algorithm~\ref{alg:old}.
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+\input{static_alg_fig}
+
+%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
 
@@
 %
-Since computing the maximum value of $f(x)$ along each template direction on the $Q$ in one shot is computationally difficult, the algorithm instead computes the maximum value over each of the constituent parallelotopes and uses the minimum of all these maximum values.
+Since computing the maximum value of $f(x)$ along each template direction over the intersection of the entire bundle $Q$ in one shot is computationally difficult, the algorithm instead computes the maximum value over each of the constituent parallelotopes and uses the minimum of all these maximum values.
 
@@
 % and half-space representation as $\tup{\T, c_{l}, c_{u}$.
-Given a template direction $\Lambda_i^Q$, the maximum value of $\Lambda_{i}^Q f(x)$ for all $x \in Q$ is less than or equal to the maximum value of $\Lambda_{i}^P \cdot f(x)$ for all $x \in P$ such that $\Lambda_i^Q$ is a row in $\Lambda^P$.
+Given a template direction $\Lambda_i^Q$, the maximum value of $\Lambda_{i}^Q \cdot f(x)$ for all $x \in Q$ is less than or equal to the maximum value of $\Lambda_{i}^P \cdot f(x)$ for all $x \in P$ if $\Lambda_i^Q$ is a row in $\Lambda^P$.
 %
@@
 %
-To compute the upper and lower bounds of each template direction $\Lambda_{i} f(x)$, for all $x \in P$, we perform the following optimization.
+\noindent Hence, to compute the upper and lower bounds of each template direction $\Lambda_{i} f(x)$ for all $x \in P$, we must find a solution to the following optimization problem:
 %
@@
 
-By , all the states in $P$ can be expressed as a vector summation of an anchor and a convex combination of generators.
+By Definition \ref{def:generator_def}, all the states in $P$ can be expressed as a vector summation of an anchor and a convex combination of generators.
 %
@@
 
-%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
-
-\input{static_alg_fig}
-
-%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
+\def \reachplotwidth {1.3}
 
 \begin{example}
+\label{ex:sir_kaa}
   We return to the SIR model briefly treated in Section \ref{sec:definitions}.
@@
 
+  \begin{figure}[h!]
+    \hspace*{-2.4cm}
+    \includegraphics[width=\reachplotwidth\textwidth]{figures/SIRProj.png}
+    \caption{Projection of Reachable Set of SIR propagated 300 steps in time.}
+    \label{fig:kaa_sir}
+  \end{figure}
+
   There a few points worth noting here.
@@
 \begin{example}
+\label{ex:phos}
 To include an example of a higher-dimensional non-linear system, we introduce the Phosporaley model.
 The Phosphoraley model describes a certain cellular regulatory system.
-It is captured by seven variables governed by the discretized dynamics stated in Figure \ref{fig:phos_dynamics}:
+It is captured by seven variables governed by the discretized dynamics stated in Figure \ref{fig:phos_dynamics}.
 
@@
 
+\begin{figure}[t!]
+  \hspace*{-2.4cm}
+  \includegraphics[width=\reachplotwidth\textwidth]{figures/PhosporlayProjOn X_1, X_2, X_3.png}
+  \caption{Projection of Reachable Set of the Phosporaley model propagated 300 steps in time.}
+  \label{fig:kaa_phos}
+\end{figure}
+
 \begin{equation}
@@
 \end{example}
-
-\def \reachplotwidth {1.3}
-
-\begin{figure}[h!]
-  \hspace*{-2.4cm}
-  \includegraphics[width=\reachplotwidth\textwidth]{figures/SIRProj.png}
-  \caption{Projection of Reachable Set of SIR propagated 300 steps in time.}
-  \label{fig:kaa_sir}
-\end{figure}
-
-\begin{figure}[h!]
-  \hspace*{-2.4cm}
-  \includegraphics[width=\reachplotwidth\textwidth]{figures/PhosporlayProjOn X_1, X_2, X_3.png}
-  \caption{Projection of Reachable Set of the Phosporaley model propagated 300 steps in time.}
-  \label{fig:kaa_phos}
-\end{figure}

### 2022-08-05 Nonlinear/formats2021/Slides/FORMATS21/formats21.tex
Committing old format slides and more changes to thesis

\documentclass{beamer}
\usepackage{pgfpages}
\usepackage{algorithm2e}
\usepackage{complexity}

\usetheme{Pittsburgh}
\useinnertheme{rectangles}
\usefonttheme{serif}
\newcommand{\T}{\mathcal{T}}

\newcommand{\tup}[1]
           {
             \relax\ifmmode
             \langle #1 \rangle
             \else $\langle$ #1 $\rangle$ \fi
           }

\title{\textbf{Automatic Dynamic Parallelotope Bundles for Reachability of Nonlinear Dynamical Systems}}

\author{\textbf{Edward Kim}\inst{1}}
\institute{\inst{1}University of North Carolina at Chapel Hill}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

%\begin{frame}
%\frametitle{Introduction}
%\begin{enumerate}
%\end{enumerate}
%\end{frame}

%\begin{frame}
%\frametitle{Outline}
%\begin{enumerate}
%  \item Overview of Background
%  \item Our Contributions
%  \item Presentation of Algorithm
%  \item Results on Benchmarks
%  \item Concluding Remarks
%\end{enumerate}
%\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Reachability Computation}
  \begin{columns}
      \column{0.5\textwidth}\footnotesize
        \begin{enumerate}
          \item Reachable set computation is an instrumental tool in performing safety analysis over dynamical systems.
          \item These computations involve taking an initial set of states and propagating them for a set amount of time to understand the possible states the initial states could reach within the alloted time.
          \item Techniques for good over-approximation of reachable sets over classes of \emph{non-linear} systems are topics of active research.
        \end{enumerate}
        \hspace{3em}
        \column{0.6\textwidth}
        \begin{figure}
            \includegraphics[width=1.13\textwidth]{InvertPend}
            \caption{Plot of Inverted Pendulum System under Initial Conditions $x \in [0.25, 0.3], y \in [0.25, 0.3]$}
          \end{figure}
\end{columns}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Reachability Computation}
  \begin{enumerate}
    \item The state of a system, denoted as $x$, lies in a domain $D \subseteq \mathbb{R}^n$.
    A discrete-time nonlinear system is denoted as
    \begin{equation}
      x^{+} = f(x)
    \end{equation}
    where $f: \mathbb{R}^{n} \rightarrow \mathbb{R}^{n}$ is a nonlinear function.
    \item The trajectory of a system that evolves according to $f$, denoted as $\xi(x_0)$ is a sequence $x_0, x_1, \ldots $ where $x_{i+1} = f(x_i)$.
    \item The $k^{th}$ element in this sequence $x_k$ is denoted as $\xi(x_0,k)$.
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Reachability Computation}
  \begin{enumerate}
  \item Given an initial set $\Theta \subseteq \mathbb{R}^n$, the \emph{reachable set} at step $k$, denoted as $\Theta_k$ is defined as
  \begin{equation*}
    \Theta_k = \{ \xi(x,k)\: | \: x \in \Theta\}
  \end{equation*}
  \item Setting number of steps $n$ and $\Theta_0 = \Theta$, the total reachable set for $n$ steps can be defined as
  \begin{equation*}
    \mathcal{R} = \bigcup_{k=0}^n \{ \xi(x,k)\: | \: x \in \Theta\}
  \end{equation*}
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Reachability with Template Polyhedra}
  \begin{enumerate}
  \item One of many techniques in computing the overapproximation of the reachable sets for discrete non-linear systems is to use \textbf{template polyhedra} to bound the reachable set.
  \item Study of template polyhedra find motivations in the static analysis of programs
  \item We are particularly interested in \emph{parallelotopes} as our template polyhedra.
\end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Reachability with Template Polyhedra}
  \begin{definition}
    For $\mathbb{R}^n$, a \emph{template polyhedron} is expressed a tuple $\langle H, \textbf{d} \rangle$ where $H$ is an $m \times n$ real-valued matrix and $\textbf{d} \in \mathbb{R}^n$ is a real-valued vector. The polyhedron is defined by the conjunction of linear inequalities
    \[ \bigwedge_{i=1}^m H_i\cdot \textbf{x} \leq d_i \]
    where $H_i$ is the $i^{th}$ row of template matrix $H$ and $\textbf{x} = (x_1, \cdots, x_m)$
  \end{definition}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Parallelotope Intuition}
  \begin{figure}
    \includegraphics[width=0.7\textwidth]{ptopeiniti}
    \caption{Example Initial Parallelotope with offset distance of 4 for all template directions}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Parallelotope Matrices}
  \begin{figure}
    \includegraphics[width=0.6\textwidth]{initptopemat}
    \caption{Example Initial Parallelotope with template matrix and offset vectors.}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Parallelotope Intuition}
  \begin{figure}
    \includegraphics[width=0.7\textwidth]{diagptope}
    \caption{Example Initial Parallelotope rotated 45 degrees clockwise}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Parallelotope Matrices}
  \begin{figure}
    \includegraphics[width=0.6\textwidth]{diagptopemat}
    \caption{Example Rotated Parallelotope with template matrix and offset vectors.}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Parallelotopes}
  \begin{definition}
    A \emph{parallelotope} in $\mathbb{R}^n$ is represented as a tuple $\langle \mathcal{T}, c_{l}, c_{u} \rangle$ where $\mathcal{T} \in \mathbb{R}^{n \times n}$ are called \emph{template directions} and $c_{l}, c_{u} \in \mathbb{R}^{n}$ such that $\forall_{1 \leq i \leq n} ~  c_{l}[i] \leq c_{u}[i]$ are called \emph{bounds}. The half-space representation defines the set of states
    %
    $$
    P = \{\: x \in \mathbb{R}^n \: | \: c_{l}[i] \leq \mathcal{T}_{i}x \leq c_{u}[i], \; 1 \leq i \leq n \}.
    $$
   \end{definition}

   \begin{itemize}
     \item The rows $\mathcal{T}_i$ are called \emph{template directions} in the parallelotope.
     \item Parallelotopes can be thought of cases of template polyhedra where each template direction $\mathcal{T}_i$ has its negative $-\mathcal{T}_i $ also included.
   \end{itemize}
\end{frame}

\begin{frame}
\frametitle{\textbf{Background:} Parallelotope Bundles}
\begin{definition}
  A \emph{parallelotope bundle} $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$. The set of states represented by a parallelotope bundle is given as the intersection
\[ Q = \bigcap_{i=1}^m P_i \]
\end{definition}

\begin{itemize}
  \item We can think of the bundle's template matrix $\mathcal{T}^Q$ as the conjunction of all the template directions $\mathcal{T}_j^{P_i}$ for the $P_i = \langle \mathcal{T}^{P_i}, c_{l}^{P_i}, c_{u}^{P_i} \rangle$
      \item \textbf{Takeaway:} Any convex initial set can be expressed as a parallelotope bundle.
\end{itemize}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Convex sets can be expressed as Parallelotope Bundles}

  \begin{figure}
    \centering
    \includegraphics[width=0.65\textwidth]{ptopedecomp}
    \caption{Figure 1 from Dreossi et. al: Parallelotope Bundles for Polynomial Reachability (2016)}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Non-linear Optimization}
  \begin{enumerate}
    \item Consider a parallelotopes $P$ in the bundle $Q$ and a non-linear function $f: \mathbb{R}^n \rightarrow \mathbb{R}^n$. We want to be able to bound the image $f(P)$ using the template directions defining the parallelotopes in our bundle $Q$.

    \item This amounts to finding new upper and lower bound vectors $c_{u}', c_{l}' \in \mathbb{R}^n$ respectively such that
    \[c_{l}'[i] \leq \mathcal{T}^Q_{i}\cdot f(x) \leq c_{u}'[i], \; x \in P \]
    for $1 \leq i \leq n$
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Non-linear Optimization}
  \begin{enumerate}
    \item The bounds can be formulated as non-linear optimization problems:
   \begin{align*}
     c_{u}'[i] & = \max_{x \in Q} \mathcal{T}^Q_i\cdot f(x) \\
     c_{l}'[i] & = \min_{x \in Q} \mathcal{T}^Q_i\cdot f(x)
   \end{align*}
     \item The reachable set computation is done by iterating this optimization problem to overapproximate the images of the bundles in each step of the computation.
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Non-linear Optimization}
  \begin{figure}
    \centering
    \includegraphics[width=0.65\textwidth]{transformexam}
    \caption{Toy example of the transformation of a bundle containing two parallelotopes}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Algorithm}
  \begin{algorithm}[H] \scriptsize
    \SetKwInput{KwInput}{Input}
    \SetKwInput{KwOutput}{Output}
    \SetKwFunction{AppendPCA}{AppendNewPCATemplates}
    \SetKwFunction{AppendLin}{AppendNewLinAppTemplates}
    \SetKwFunction{TransformBund}{TransformBundle}
    \SetKwFunction{UpdateTemp}{UpdateTemplates}
    \SetKwFunction{GetSupp}{GetSupportPoints}
    \SetKwFunction{PropPoints}{PropagatePointsOneStep}
    \SetKwFunction{PCA}{PCA}
    \SetKwFunction{ApproxLinearTrans}{ApproxLinearTrans}
    \SetKwFunction{SetLifeSpan}{SetLifeSpan}
    \SetKwFunction{AddTemptoBund}{AddTemplateToBundle}
    \SetKwFunction{RemoveTemp}{RemoveTempFromBund}
    \SetKwProg{Fun}{Proc}{:}{}

  \SetAlgoLined
  \DontPrintSemicolon

  \KwInput{Dynamics $f$, Initial Parallelotope bundle $Q_0$, Step Bound $S$, indexes for parallelotopes $I$}
  \KwOutput{Reachable Set Overapproximation $\overline\Theta_k$ for each step $k$} \;

   \For{$k \in [1, 2, \ldots, S]$}{
      $Q_k$ = \TransformBund($f$, $Q_{k-1}$, $\Lambda^{Q_{k-1}}$) \;
     $\overline\Theta_k = Q_k$ \;
   }
   \Return{$\overline\Theta_1 \ldots \overline\Theta_S$} \;
   \;
   %
   \Fun{\TransformBund{$f$, $Q$, $\Lambda$}}{
     $Q' \gets \{\}$; $c_{u} \gets +\infty$; $c_{l} \gets -\infty$\;
     \For{each parallelotope $P$ in $Q$}{
       % $\tup{\T, c_{l}, c_{u}} \gets \mathsf{half-spaceRepresentation}(P)$\;
       \For{each template direction $\Lambda_i$ in the template directions $\Lambda$ }{
         $c_{u}'[i] \gets \mathsf{min}\{ \mathsf{\Lambda_{i} \cdot f}, c_{u}'[i] \}$ \;
         $c_{l}'[i] \gets \mathsf{max}\{\mathsf{-1\times \Lambda_{i} \cdot f}, c_{l}'[i] \}$\;
       }
     }
     Construct parallelotopes $P_{1}', \ldots, P_{k}'$ from $\Lambda, c_{l}', c_{u}'$ and indexes from $I$\;
     $Q' \gets \{P_{1}', \ldots, P_{k}'\}$\;
       % $P' \gets \tup{\T, c_{l}', c_{u}'}$\;
       % $Q' \gets Q' \cup \{P'\}$\;
     \Return{$Q'$} \;
   }
  \end{algorithm}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Bernstein Expansion}
  \begin{enumerate}
    \item Previous work focused on reachabilty under discrete non-linear \emph{polynomial dynamics} i.e the non-linear function $f: \mathbb{R}^n \rightarrow \mathbb{R}^n$ is polynomial.
    \item Our optimization problems become one of optimizing over polynomial $\mathcal{T}_i\cdot f(x)$ which is generally a costly operation.
    \item The insight comes from leveraging \emph{Bernstein polynomials}. These polynomials yield an efficient manner of bounding the solution to the aforementioned polynomial optimization problems.
    \item An extra linear transformation taking the unit-box domain $[0,1]^n$ to the parallelotope $P$ is required to use Bernstein polynomials.

    %\begin{itemize}
    %  \item Given a polynomial $P$ have the property that their maximum value over the unit box $[0,1]^n$ is upper-bounded by the greatest coefficient
    %\end{enumerate}
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Background:} Drawbacks}
    \begin{columns}
    \column{0.5\textwidth}\scriptsize
    \begin{enumerate}
      \item Hitherto only \emph{static} parallelotopes have been considered. In other words, the template directions specifying the parallelotopes are to be given as user input at the beginning of the reachable set computation.
      \item The template directions chosen are generally the axis-aligned, diagonal directions. However, it's not clear that these directions necessary yield good overapproximations.
      \item Since the template directions are set at the beginning, they cannot adapt to the behavior of the dynamics. This could yield overapproximations which are too conservative for any practical use.
    \end{enumerate}
    \column{0.5\textwidth}
    \begin{figure}
        \includegraphics[width=1.13\textwidth]{cvdp20}
        \caption{The effect of choosing inappropriate template directions for the Coupled VanDerPol system.}
      \end{figure}
  \end{columns}
  %Static, diagonal templates are used.
  %Bad for wrapping error as initial sets increase.
\end{frame}

\begin{frame}
  \frametitle{Contributions}
  \begin{enumerate}
    \item We present a method which is both \emph{dynamic} and \emph{automatic}. Our method utilizes the Principal Component Analysis (PCA) and Local Linear Approximations.
    \item  We extend our tool \emph{Kaa} to leverage NASA's \emph{Kodiak} to perform the optimization procedure.
    \item We parallelize our implementation to scale with an increasing number of parallelotopes in our bundles.
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Contributions:} Automatic, Dynamic Algorithm}
  \begin{enumerate}
    \item We choose template directions at each step by two procdure working in tandem
    \item One procedure generates new template directions via Principal Component Analysis (PCA) and the other generates template directions through Local Linear Approximations.
    \item Each procedure adds one parallelotope to the bundle with a set \emph{template lifespan}. This lifespan dictates the future number of steps the parallelotope and its template directions will exist in the bundle.
  \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{\textbf{Contributions:} PCA Directions}
    \begin{enumerate}
      \item To generate the PCA directions from bundle $Q$, we first calculate the \emph{support points} of all template directions in $\T^Q_i$ over the bundle $Q$.
      \[ p_{i}^u = \max_{x \in Q} T_i^Q \cdot x, \quad  p_{i}^\ell = \min_{x \in Q} T_i^Q \cdot x \]
      \item We then propagate these support points to the next step by the given dynamics $f$. This will gives us set of trajectories.
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{\textbf{Contributions:} PCA Directions}
    \begin{enumerate}
      \item We perform PCA on the endpoints of these trajectories to yield a real-valued $n \times n$ matrix whose rows are the PCA directions. These rows are used as template directions to define a single parallelotope $P'$.
      \item The parallelotope $P'$ with template lifespan $T_{P'}$ is added to the bundle $Q$ to give another bundle $Q' = Q \cup P'$.
    \end{enumerate}
\end{frame}
\begin{frame}
    \frametitle{\textbf{Contributions:} Linear Approx. Directions}
    \begin{enumerate}
      \item Observe that the if the dynamics $f$ were linear i.e $x^+ = Ax$ for some linear transformation $A$, then the image of the parallelotope $c_{l} \leq \T x \leq c_{u}$, will be the set $c_{l} \leq \T \cdot A^{-1} x \leq c_{u}$.
      \item Since the initial parallelotope will be defined by the axis-aligned directions ($\T = I_n$), the image after the first step will be given by
        \[ c_{l} \leq A^{-1} x \leq c_{u}  \]
      Similarly, after the exact image of the initial parallelotope after two steps will problems:
      \[ c_{l} \leq (A^{-1})^2 x \leq c_{u}  \]
    \end{enumerate}
  \end{frame}

\begin{frame}
    \frametitle{\textbf{Contributions:} Linear Approx. Directions}
    \begin{enumerate}
      \item Begin computation with $\T_{lin} = I_n$
      \item To generate the Linear Approximation templates, we once again calculate the support points of all template directions, $\T^Q_i$, over the bundle $Q$.
      \item Propagate the support points using given dynamics $f$ to the next step.   This gives us a set of trajectories $S$.

    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{\textbf{Contributions:} Linear Approx. Directions}
    \begin{enumerate}
      \item Calculate the best-fit linear approximation $\mathcal{A}$ using the start and endpoints of the trajectories. Note that $\mathcal{A}$ will be a real-valued $n \times n$ matrix.
      \item Multiply the template directions $\T_{lin}$ by $\mathcal{A}^{-1}$ to yield $\T_{lin} \cdot \mathcal{A}^{-1}$.
      \item Define a parallelotope $P'$ with template lifespan $T_{P'}$ and add it to bundle $Q$ to get $Q' = Q \cup P'$.
      \item $\T_{lin} \gets \T_{lin} \cdot \mathcal{A}^{-1}$
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{\textbf{Contributions:} Template Lifespan}
    \begin{enumerate}
      \item The template lifespan parameter dictates the number of steps after its creation the parallelotope's template directions exist in the bundle.
      \item The intuition is that certain dynamics may be more amenable to directions generated previously in the reachable set computation.
      \item Having a lifespan allows the algorithm to be adaptive while attenuating the computational cost of optimizing over a large number of parallelotopes.
      \item This is a tunable parameter we vary in our experiments.
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{\textbf{Contributions:} Template Lifespan}
    \begin{figure}
      \includegraphics[width=0.7\textwidth]{templifespan}
      \caption{Instance of the template directions of parallelotope $P$ having a template lifespan of two.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{\textbf{Contributions:} Dynamic Algorithm}
    \begin{algorithm}[H] \scriptsize
      \SetKwInput{KwInput}{Input}
      \SetKwInput{KwOutput}{Output}
      \SetKwFunction{CreatePCA}{CreatePCA}
      \SetKwFunction{CreateLin}{CreateLin}
      \SetKwFunction{TransformBund}{TransformBundle}
      \SetKwFunction{UpdateTemp}{UpdateTemplates}
      \SetKwFunction{GetSupp}{GetSupportPoints}
      \SetKwFunction{PropPoints}{PropagatePointsOneStep}
      \SetKwFunction{PCA}{PCA}
      \SetKwFunction{ApproxLinearTrans}{ApproxLinearTrans}
      \SetKwFunction{SetLifeSpan}{SetLifeSpan}
      \SetKwFunction{ExtractDirections}{ExtractDirections}
      \SetKwFunction{AddTemptoBund}{AddTemplateToBundle}
      \SetKwFunction{Maximize}{Maximize}
      \SetKwFunction{hstack}{hstack}
      \SetKwFunction{RemoveTemp}{RemoveTempFromBund}
      \SetKwProg{Fun}{Proc}{:}{}

    \SetAlgoLined
    \DontPrintSemicolon

    \KwInput{Dynamics $f$, Initial Parallelotope $P_0$, Step Bound $S$}
    \KwOutput{Reachable Set Overapproximation $\overline\Theta_k$ at each step $k$}

    $Q_0 = \{P_0 \}$ \;

    % $\T_0 = \{ \{ P_0.\T_1, \ldots P_0.\T_n \} \}$ \;
    $\T = I$ \tcp{Init Template Directions}

     \For{$k \in [1, 2, \ldots, S]$}{
        $P_{supp}$ = \GetSupp($Q_{k-1}$) (support points of $Q_{k-1}$) \;
        $P_{prop}$ = \PropPoints($P_{supp}$, $f$) (image of support points) \;

        $A$ =  \ApproxLinearTrans($P_{supp}$, $P_{prop}$) \; \label{ln:linearapprox}
        $\T = \T \cdot A^{-1}$ \;

        $\T_k^\text{lin} = \{ \T_{1} , \ldots, \T_{n} \} $ \;

        $\T_k^\text{pca} = \{ \PCA(P_{prop}) \} $\; \label{ln:pca}

        $\T_k = \T_k^\text{lin} \cup \T_k^\text{pca}$ \;
        \;
        \tcc{For lifespan $L > 0$, instead call \TransformBund with $\T_k \cup \T_{k-1} \cup \ldots \cup \T_{k-L}$}
        $Q_k$ = \TransformBund($f$, $Q_{k-1}, \T_k$) \;
       $\overline\Theta_k \gets Q_k$ \;
     }
     \Return{$\overline\Theta_1 \ldots \overline\Theta_S$} \;
     %
     \;

    \end{algorithm}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Contributions:} Kaa}
  \begin{enumerate}
    \item Kaa is written in Python and relies on the \emph{numpy} library for matrix computations, \emph{sympy} library for all symbolic subsitution, and \emph{scipy}, \emph{matplotlib} for plotting the reachable sets and computing the volume for lower-dimensional systems.

    \item The optimization procedure for finding the upper and lower bounds of template directions  is performed through the \emph{Kodiak} library. Finally, parallelization of the offset calculation procedures is implemented through the \emph{multiprocessing} module.

    \item Kodiak allows for more general dynamics beyond the polynomial ones considered in a previous tool, \emph{Sapo}
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Results:} Benchmarks}
  \begin{enumerate}
    \item For benchmarking, we select six non-linear models with polynomial dynamics.
    \item Many of these models are also implemented in \emph{Sapo} which explored reachability with {\bf static} parallelotope bundles. In these cases, we directly compare the performance of our dynamic strategies with the Sapo's static parallelotopes defined by \emph{diagonal} directions.
    \item To provide meaningful comparisions, we set the number of dynamic parallelotopes to be equal to the number of static ones excluding the initial box.
  \end{enumerate}
  %Explain benchmarks and choices. Show Table 1
\end{frame}

\begin{frame}
  \frametitle{\textbf{Results:}  Benchmarks}
  %Explain methodology and why we chose this testing procedure with diagonal temps etc.
  \begin{table}[h!]\scriptsize
  %\hspace{-5em}
    \centering
  \begin{tabular}{|p{1.5cm}|c|p{1.7cm}|c|c|p{5cm}|}
  \hline
  Model & Dimension & Parameters & \# steps & $\Delta$  \\
  \hline
  Vanderpol & 2 & \quad \quad \; - & 70 steps & 0.08 \\
  \hline
  Jet Engine& 2 & \quad \quad \; - & 100 steps & 0.2  \\
  \hline
  Neuron & 2 & \quad \quad \; - & 200 steps & 0.2 \\
  \hline
  SIR& 3 & $\beta=0.05$ \newline $\gamma=0.34$ & 150 steps & 0.1 \\
  \hline
  Coupled \newline Vanderpol & 4 & \quad \quad \; - & 40 steps & 0.08 \\
  \hline
  COVID & 7 & $\beta=0.05$ \newline $\gamma=0.0$ \newline $\eta=0.02$ & 200 steps & 0.08 \\
  \hline
  \end{tabular}
  \label{tab:modeldyns}
  \caption{Benchmark models and truncated information}
  \end{table}
\end{frame}

\begin{frame} \small
  \frametitle{\textbf{Results:}  COVID Supermodel}
A modified SIR model accounting for the possibility of asymptomatic patients. Given by the following discretized seven-dimensional dynamics:

\begin{align*}
  \begin{split}
   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
   R_A' & = R_A + (\gamma A)\cdot \Delta \\
   R_I' & = R_I + (\gamma I)\cdot \Delta \\
   D' & = D + (\eta I)\cdot \Delta
 \end{split}
\end{align*}
where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}.
\end{frame}

\begin{frame}
  \frametitle{\textbf{Results:}  COVID Supermodel}
  \begin{figure}
    \centering
    \includegraphics[width=1.1\textwidth]{IndianRecoveredPlot}
    \caption{Plot of COVID Supermodel reachable set against real-world Indian COVID data.}
  \end{figure}
 \end{frame}

\begin{frame}
  \frametitle{\textbf{Results:} Comparision against Static Parallelotopes}
  \begin{columns}
    \column{0.5\textwidth}\footnotesize
    \begin{enumerate}
      \item For models previously defined in Sapo, we set the static parallelotopes to be exactly those found in Sapo.
      \item If a model is not implemented in Sapo, we simply use the static parallelotopes defined in a model of equal dimension.
      \item To address the unavailability of a four dimensional model implemented in Sapo, we sampled random subsets of five static parallelotopes defined with diagonal directions and chose the flowpipe with smallest volume.
    \end{enumerate}
    \column{0.5\textwidth}
    \includegraphics[width=1.05\textwidth]{voltable}
  \end{columns}

\end{frame}

\begin{frame}
  \frametitle{\textbf{Results:}  Performance under Increasing Initial Sets}
  \begin{columns}
      \column{0.5\textwidth}\footnotesize
      \begin{enumerate}
        \item We vary the initial box dimensions to gradually increase the box's volume, then plot the total flowpipe volumes after running the benchmark on the static strategies and best-performing dynamic strategies.
        \item Our dynamic strategies perform better than static ones in controlling the total flowpipe volume as the initial set becomes larger.
        \item The performance of static parallelotopes tends to degrade rapidly as we increase the volume of the initial box
      \end{enumerate}

      \column{0.5\textwidth}
      \includegraphics[width=1.1\textwidth]{sapoinitplots}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Results:}  Performance against Random Static Bundles}
  \begin{columns}
      \column{0.5\textwidth}\footnotesize
        \begin{enumerate}
          \item We compare our dynamic strategies against static random parallelotope bundles under initial sets with increasing volume.
          \item We sample such parallelotopes in $n$ dimensions by first sampling a set of $n$ directions uniformly on the surface of the unit $(n-1)$-sphere, then defining our parallelotope using these sampled directions
          \item We sample twenty of these parallelotopes for each trial, average the total flowpipe volumes, then plot this against the best-performing dynamic strategy.
        \end{enumerate}

      \column{0.5\textwidth}
      \includegraphics[width=1.1\textwidth]{randomvolplot}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{\textbf{Results:}  No Universal Ratio}
  \begin{columns}
      \column{0.5\textwidth}\footnotesize
        \begin{enumerate}
          \item There seems to be no universal optimal ratio between the number of dynamic parallelotopes defined by PCA and Linear Approximation directions which perform well on all benchmarks.
          \item In the Vanderpol model, using parallelotopes defined by Linear Approximation directions is more effective than those defined by PCA directions.
          \item Neuron model shows the opposite trend where using parallelotopes defined by PCA directions is more effective than using those defined by Linear Approximation directions.
        \end{enumerate}
        \column{0.5\textwidth}
        \includegraphics[width=1.15\textwidth]{universalplot}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Future Directions}
  \begin{enumerate}
    \item Are there other adaptive methods we can utilize to generate template directions dynamically?
    \item What about more sophisticated linear approximation methods through Koopman linearization techniques?
    \item The use of massively parallelized reachability techniques through HPC hardware is applicable here. Specifcally, can we leverage GPUs to parallelize the optimization procedure for larger parallelotope bundles?
    \item Why do PCA directions perform better than Linear Approximation directions on certain systems?
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Concluding Remarks}
  \begin{enumerate}
    \item We investigated two techniques for generating templates dynamically: first using linear approximation of the dynamics, and second using PCA.
    \item We also observed that both these techniques improve the accuracy of the reachable sets for different benchmarks.
    \item We extend Kaa with Kodiak to handle more general dynamics. Systems with more general dynamics are considered in the extended version on arXiv.
  \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{Relevant Links}
    \begin{itemize}
      \item \textbf{Github Repository for Kaa:} \newline
        https://github.com/Tarheel-Formal-Methods/kaa-dynamic
      \item \textbf{Github Repository for Kodiak:} \newline
        https://github.com/nasa/Kodiak
      \item \textbf{COVID Modeling Blog Post:} \newline
      https://sigbed.org/2021/06/21/sidbed-blog-covid-formal-verification/
    \end{itemize}
\end{frame}

\end{document}

### 2022-08-05 thesis/appendix.tex
Draft of RT reachability written
Second draft of thesis finished
Committing old format slides and more changes to thesis

@@
 
-\chapter{Benchmark Model Definitions}
+\chapter{Additional Benchmark Model Definitions}
 \label{app:benchmark}
@@
 \section{Vanderpol Oscillator}
+\label{sec:vdp}
 The Vanderpol Oscillator is a classical two-dimensional non-linear dynamical system governed by the following dynamics:
@@
 \section{Coupled Vanderpol}
+\label{sec:cvdp}
 This model couples two Vanderpol Oscillators together, resulting in an non-linear ODE of four variables

### 2022-08-05 thesis/formats.tex
Committing old format slides and more changes to thesis

@@
 
+\section{Drawbacks to the Static Algorithm}
+Before we embark on introducing the dynamic algorithm, let us first consider a few motivating factors.
+%
+As discussed in Section \ref{sec:static}, the static algorithm requires template directions to be specified by the user in order for the reachable set computation to proceed.
+%
+The parallelotopes tended to be defined by a combination of axis-aligned directions or diagonal directions.
+%
+However, it is unclear whether these template directions work reasonably well over general non-linear dynamics.
+%
+This is a problem which has profound consequences for computing useful over-approximations.
+%
+Choosing the incorrect directions could very well lead to an over-approximation too conservative for any practical use.
+%
+Figure \ref{fig:bloat_cvdp} portrays a simple example of a reachable set whose wrapping error becomes explosive after a few steps.
+%
+In light of these ramifications, it is esstentially incumbent on the user to choose fruitful template directions that control the growth of the over-approximation.
+%
+Due to the inherent difficult nature of choosing these template directions, it may be unreasonable to delegate this responsibility to a practitioner who knows little about parallelotope-based reachability.
+%
+These drawbacks to the static algorithm motivate the rationale behind the \emph{automatic} aspect of our modified algorithm.
+
+Additionally, as the template directions are set at the beginning before the reachable set computation proceeds, these directions cannot adapt to the dynamics as the computation executes.
+%
+There may be template directions which may yield much leaner over-approximations at an intermediate step during the computation.
+%
+Hence, the ability to change the composition of the parallelotope bundle by adding and removing parallelotopes based on a data-driven approach could yield leaner over-approximations.
+%
+These considerations motivate the rationale behind the \emph{dynamic} aspect of our modified algorithm.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width= 0.9\textwidth]{figures/cvdp20.png}
+\caption{Consequences of choosing inappropriate template directions as demonstrated by the Coupled Vanderpol model. See Section \ref{sec:cvdp} for the model definition.}
+\label{fig:bloat_cvdp}
+\end{figure}
+
 \section{Local Linear Approximations}
@@
 for all template directions of the bundle $\Lambda^Q_i$.
-We use the support points as a data-driven approach to find the best-fit linear function to use.
+We use the support points as a data-driven approach to determine the best-fit linear function to use.
 All points are found by a straightforward linear program.
@@
 %
-Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the non-linear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
+To see this, recall the constraint associated to the half-space representation (Definition \ref{def:halfspace_def} and Equation \ref{eq:halfspaceconst}).
+If we set our new template directions to be $\Lambda \cdot A^{-1}$, the image points of $A$ which satisfy the half-space constraint according to fixed offset vectors $c_l, c_u$ should be:
+ \[ c_l \leq (\Lambda \cdot A^{-1})Ax \leq c_u \implies c_{l} \leq \Lambda x \leq c_{u} \]
+%
+This is satisfied by all the points defining parallelotope $P$ as desired.
+Therefore, the new template directions will exactly define the image of $P$ under linear dynamics.
+To exploit this property, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the non-linear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
 %

### 2022-08-05 thesis/prelim.tex
Committing old format slides and more changes to thesis

@@
     \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.
-Once again, note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes.
-Once again, $\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
+Note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes.
+$\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
     \label{fig:diagonal_ptope}
@@
 \label{ex:diag_ptope}
-Once again, consider the space $\reals^2$ and the parallelotope $P$ given in half-plane representation as $0 \leq x-y \leq 1$, $0 \leq y \leq 1$.
+Parallelotope $P$ is given in half-plane representation as $0 \leq x-y \leq 1$, $0 \leq y \leq 1$.
 %
@@
 %
-In the half-space representation, the template directions of the parallelotope $P$ are given by the directions $[1, -1]$ and $[0, 1]$.
+The template directions of the parallelotope $P$ are given by the directions $[1, -1]$ and $[0, 1]$.
 %
@@
 Therefore, all the points in the paralellotope can be written as $(x,y) = (0,0) + \alpha_1 \cdot (1,0) + \alpha_2\cdot (1,1)$, $\alpha_1, \alpha_2 \in [0,1]$.
+%
+Figure \ref{fig:diag_ptope} portrays a visual plot of the derived generator representation.
 \end{example}
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
+    \draw[fill=blue!40!white, opacity=0.4] (0,0) -- (1,0) -- (2,1) -- (1,1) -- cycle;
+    \draw[fill] (0,0) circle [radius=1pt];
+    \node[above] {$\mathcal{O}$};
+
+    \draw[fill] (0,0) circle [radius=1.5pt];
+    \node[below] at (0,0) {$v$};
+
+    \draw[->, very thick, draw=red] (0,0) -- (1,0) node[midway, below] {$g_1$};
+    \draw[->, very thick, draw=red] (0,0) -- (1,1) node[midway, right] {$g_2$};
+  \end{tikzpicture}
+  \caption{Plot of the generator representation for Example \ref{ex:diag_ptope}. The vertex $v$ represents the anchor while the two vectors $g_1,g_2$ are the generator vectors.}
+  \label{fig:diag_ptope}
+\end{figure}
+
 %
@@
 %--------------------------------------------------------------------------------------------------------------------------------------------------------- --------
+\newpage
 \subsection{Bernstein Polynomials}
@@
 One of the most important properties of Berstein polynomials is that the Bernstein basis polynomials of degree $\boldd$ span the vector space of real multivariate polynomials of degree at most $\boldd$:
-
+%
+In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
+\newpage
 \begin{property}
@@
 \end{property}
-
-\noindent In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.
 

### 2022-08-05 thesis/staticalg.tex
Committing old format slides and more changes to thesis

@@
 \begin{remark}
-The original algorithm using Bernstein expansion proposed by \cite{dreossi2016parallelotope} assumed polynomial dynamics i.e $f$ is polynomial in the system variables.
+The original algorithm exploiting Bernstein expansion was proposed by \cite{dreossi2016parallelotope} and assumed polynomial dynamics i.e $f$ is polynomial in the system variables.
 %

### 2022-08-19 thesis/appendix.tex
Formatting, font changes. Rearrangment of figures

@@
-\appendix
+\begin{appendices}
 
-\chapter{Additional Benchmark Model Definitions}
+\chapter{ADDITIONAL BENCHMARK MODEL DEFINITIONS}
 \label{app:benchmark}
@@
 
-\noindent The model was first studiend in \cite{fitzhugh1961impulses} and its reachability analysis using Bernstein expansion is studied in \cite{dang2012reachability}.
+\noindent The model was first studied in \cite{fitzhugh1961impulses} and its reachability analysis using Bernstein expansion is considered in \cite{dang2012reachability}.
+
+\end{appendices}

### 2022-08-19 thesis/bern_prune_wrapping_error.tex
Formatting, font changes. Rearrangment of figures

@@
 
-Difference between Queried Min Coeff and Actual Min Coeff: -
-0.719813449727491
+      Difference between Queried Min Coeff and Actual Min Coeff:
+      -0.719813449727491
 \end{verbatim}
 
-\caption{Example of Wrapping Error Obstacle demonstrated for Step 4 for Vanderpol Model.}
+\caption{Example of Wrapping Error Obstacle Demonstrated for Step 4 for Vanderpol Model.}
 \label{fig:bern_prune_wrapping_output}

### 2022-08-19 thesis/formats.tex
Formatting, font changes. Rearrangment of figures

@@
 %
-The parallelotopes tended to be defined by a combination of axis-aligned directions or diagonal directions.
+In practice, the parallelotopes tend to be defined by a combination of axis-aligned directions or diagonal directions.
 %
@@
 %
-This is a problem which has profound consequences for computing useful over-approximations.
+This is a problem which has profound consequences for the computation of useful over-approximations.
 %
@@
 
-Additionally, as the template directions are set at the beginning before the reachable set computation proceeds, these directions cannot adapt to the dynamics as the computation executes.
+Additionally, as the template directions are set at the beginning, these directions cannot adapt to the dynamics as the computation executes.
 %
@@
 %
-Hence, the ability to change the composition of the parallelotope bundle by adding and removing parallelotopes based on a data-driven approach could yield leaner over-approximations.
+Hence, the ability to change the composition of the parallelotope bundle by adding and removing parallelotopes based on a data-driven approach could yield better over-approximations.
 %

### 2022-08-19 thesis/prelim.tex
Formatting, font changes. Rearrangment of figures

@@
 
+\begin{figure}[t!]
+  \centering
+  \begin{tikzpicture}
+    \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
+    \fill[blue!40!white, opacity=0.4] (-1,-1) rectangle (1,1);
+    \draw[->, thick](0,1) -- (0,2) node[above] (0,2) {$[0,1]$};
+    \draw[->, thick](1,0) -- (2,0) node[right] (2,0) {$[1,0]$};
+    \draw[->, thick](-1,0) -- (-2,0) node[left] (-2,0) {$[-1,0]$};
+    \draw[->, thick](0,-1) -- (0,-2) node[below] (0,-2) {$[0,-1]$};
+    \draw[fill] (0,0) circle [radius=1pt];
+    \node[above] (0,0) {$\mathcal{O}$};
+    %\foreach \i in {-3,-2,...,3} \node[below] at (\i,0) {\scriptsize \i};
+    %\foreach \i in {-3,-2,...,3} \node[right] at (0,\i) {\scriptsize \i};
+  \end{tikzpicture}
+
+  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}.}
+  \label{fig:axis_align_ptope}
+\end{figure}
+
 \begin{equation}
@@
 For a visual plot of above axis-aligned parallelotope, see Figure \ref{fig:axis_align_ptope}.
-
-\begin{figure}[h!]
-  \centering
-  \begin{tikzpicture}
-    \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
-    \fill[blue!40!white, opacity=0.4] (-1,-1) rectangle (1,1);
-    \draw[->, thick](0,1) -- (0,2) node[above] (0,2) {$[0,1]$};
-    \draw[->, thick](1,0) -- (2,0) node[right] (2,0) {$[1,0]$};
-    \draw[->, thick](-1,0) -- (-2,0) node[left] (-2,0) {$[-1,0]$};
-    \draw[->, thick](0,-1) -- (0,-2) node[below] (0,-2) {$[0,-1]$};
-    \draw[fill] (0,0) circle [radius=1pt];
-    \node[above] (0,0) {$\mathcal{O}$};
-    %\foreach \i in {-3,-2,...,3} \node[below] at (\i,0) {\scriptsize \i};
-    %\foreach \i in {-3,-2,...,3} \node[right] at (0,\i) {\scriptsize \i};
-  \end{tikzpicture}
-
-  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}.
 The template directions are displayed as normal vectors to the pairs of parallel planes defining the parallelotope.
-$\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
-  \label{fig:axis_align_ptope}
-\end{figure}
+$\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.
 
@@
 To see a visual plot of the diagonal parallelotope, see Figure \ref{fig:diagonal_ptope}.
+Note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes.
+Again, $\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.
 
-  \begin{figure}[h!]
+  \begin{figure}[t!]
     \centering
@@
     \end{tikzpicture}
-    \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.
-Note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes.
-$\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
+    \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.}
     \label{fig:diagonal_ptope}
@@
 
-As a final remark, notice that for a parallelotope $P$, the generator representation also defines an affine transformation that maps $[0,1]^{n}$ to $P$.
+\begin{remark}
+Notice that for a parallelotope $P$, the generator representation also defines an affine transformation that maps $[0,1]^{n}$ to $P$.
 %
 We refer to this affine transformation associated to $P$ as $T_P:[0,1]^n \rightarrow P$ when necessary.
+\end{remark}
 

### 2022-08-19 thesis/realtime.tex
thesis slide commit
Committing old format slides and more changes to thesis
Formatting, font changes. Rearrangment of figures

@@
 
-To mitigate this explosive growth in coefficients, we can pose the following question: Is there a method for ascertaining a small subset of the basis monomials such that the correct maximum and minimum coefficients can be computed from this subset rather the entire set of Bernstein basis monomials?
+To mitigate this explosive growth in coefficients, we pose the following question:
 %
-Towards this end, we propose the following algorithm outlined in Algorithm \ref{alg:bern_prune} and elaborate on the individual subroutines.
+\begin{tcolorbox}[standard jigsaw, opacityback=0]
+Does there exists a method for ascertaining a small subset of the basis monomials such that the correct maximum and minimum coefficients can be computed from this subset rather the entire set of Bernstein basis monomials?
+\end{tcolorbox}
+%
+To this end, we propose the following algorithm outlined in Algorithm \ref{alg:bern_prune} and elaborate on the individual subroutines.
 In the context of these experiments, we assume the dynamics $f$ are polynomial.
@@
 
-\input{bern_prune_alg}
-
 \newcommand\litem[1]{\item{\bfseries #1 \\}}
 
-\begin{enumerate}
+\begin{enumerate}[leftmargin=*]
 \litem{Parition Domain into Grid Cells:} We first partition the domain into grid cells of equal dimension.
@@
 
+\input{bern_prune_alg}
+
 \litem{Find the Maximum/Minimum Interval:} Using the coeffcient intervals $\{\barb_{\boldi}\}$, we find the maximum and minimum interval by simply plucking the interval which has the largest upper-bound and the interval which has the smallest lower-bound.
@@
 
+\input{mod_reach_alg}
+
 \section{Experimental Results}
@@
   \item  We attempted to run the modified reachability algorithm for seven steps.
-  \item We partitioned the domain and fixed the initial set for each model according to the below inputs:
+  \item We partitioned the domain and fixed the initial set for each model according to the parameters listed in Table~\ref{tab:vdp_rt}.
+\end{itemize}
 
+\begin{table}
+ \centering
  \begin{tabular}{|c|c|c|c|}
@@
  \end{tabular}
-\end{itemize}
+ \caption{Parameters for Vanderpol Model used during Real-time Reachability Experiments}
+ \label{tab:vdp_rt}
+\end{table}
 
@@
 %
-What seems to happen quite frequently is the maximum or minimum intervals and their overlaps dominate the degree of the polynomial computed after composition during the reachability computation.
+Frequently, the maximum or minimum intervals and their overlaps dominate the degree of the polynomial computed after composition during the reachability computation.
 To see this, note the anchor and generator vectors computed for the generator representation:
@@
 Hence, during our pre-processing stage, we consider a more general set of monomial degrees which may not reflect the relevant degree statistics found during the actual computation phase.
-
-This poses the issue of being privy to the degrees supplied to the algorithm before it even begins.
+%
+This poses the issue of being privy to the degrees supplied to algorithm before execution begins.
 %
@@
 A guess is that as the conservativeness of our reachable becomes greater, it overlaps with more cells of the domain.
-This results in an cascading effect where the reachability algorithm is required to take into account more monomials as it progresses.
-
-We speculate that the utility of our pruning method hinges on the careful control of the reachable set error.
+This results in a cascading effect where the reachability algorithm is required to take into account more monomials as it progresses.
+%
+In light of the discussed observations, we speculate that the utility of our pruning method hinges on the careful control of the reachable set error.
 If the error becomes too great, we end of computing all of the coefficients, negating any benefits of a speed-up.

### 2022-08-25 thesis/frontmatter/acknowledgement.tex
Formatting, font changes. Rearrangment of figures
Third formatting revisions for MS thesis

@@
 \vspace*{52pt}
-{\Large \textbf{ACKNOWLEDGEMENTS}}
+{\fontsize{12pt}{14.4pt}\selectfont \textbf{ACKNOWLEDGEMENTS}}
 \end{center}
@@
 I would like to thank the UNC Computer Science community for supporting my work and providing me a warm community to pursue this research.
-In particular, I would like to thank my advisor, Parasara Sridhar Duggirala, for the opporunity to perform research and participate in his group.
+In particular, I would like to thank my advisor, Parasara Sridhar Duggirala, for opporunities to perform research and to participate in his group.
 %
@@
 Juan Garcia is a wonderful collegue and an equally wonderful friend.
-The nights we raided Peabody and Phillips hall was certainly a memorable time.
+The nights we raided Peabody and Phillips hall were certainly memorable times.
 Finally, I would like to thank my parents and my family for their ardent, unending support.

### 2022-08-25 thesis/prelim.tex
Third formatting revisions for MS thesis

@@
 %
-Discretizing Equation~\ref{eqn:sir} according the Euler method yields the dynamics:
+Discretizing Equation~\ref{eqn:sir} according to the Euler method yields the dynamics:
 
@@
 %
-Note the non-linear terms $s_ki_k$, which preclude the expression of the dynamics as a linear transformation.
+Note the non-linear terms $s_ki_k$ which preclude the expression of the dynamics as a linear transformation of the input state variables.
 \end{example}
